### 2017-12-08 ABB_Benthic.Rmd
adding files

---
title: "Antigua-Barbuda Benthic and Rugosity Data"
output: html_document
---

# Benthic and Rugosity data 
*required files: benthic.csv; rugosity.csv*

## Part 1. Setup

### 1.1 Load packages
```{r, echo=FALSE,message=FALSE,warning=FALSE}
# set up
library(lazyeval)
library(rmarkdown)
library(dplyr)
library(plyr)
library(tidyr)
library(ggplot2)
library(grid)
library(stringr)
library(reshape2)
library(data.table)
```
### 1.2 Import and format files
```{r, echo=FALSE,message=FALSE,warning=FALSE}
setwd("~/github/Scarid-Behavior/")
benthic<-read.csv("benthic.csv", stringsAsFactors=FALSE)
rugosity<-read.csv("rugosity.csv", stringsAsFactors=FALSE)
```

## Part 2. Benthic summary data

### 2.1 Percent cover
```{r, echo=FALSE,message=FALSE,warning=FALSE}
# Because not all categories were present within each meter, I need to create template to add 0 values in meters where they were absent.
expand.meter <- benthic %>% group_by(Site,Transect,Meter) %>% dplyr::summarize()
expand.category <- benthic %>% expand(nesting(Code),Site)
temp.benthic <- full_join(expand.meter,expand.category,by="Site")

# Then I can join benthic data with this template, and replace empty rows with 0 values
benthic.meter<- benthic %>% group_by(Site,Transect,Meter,Code) %>% dplyr::summarize(count.m=n()) %>% right_join(temp.benthic, by=c("Site","Transect","Meter","Code")) # join transect-level data with template
benthic.meter[is.na(benthic.meter)] <- 0 # convert NAs to 0 biomass/density in transects where a species is not present

# Calculating percent cover on potential substrate (i.e. not sand), remove meters that were 100% sand cover (no potential susbtrate)
sand.meter <- benthic.meter %>% filter(Code=="sand") %>% dplyr::rename(count.sand=count.m) %>% select(-Code) 
benthic.meter <- left_join(benthic.meter,sand.meter,by=c("Site","Transect","Meter")) %>% mutate(cover.m=100*count.m/(10-count.sand)) %>% filter(!is.nan(cover.m)) %>% filter(Code!="sand")

benthic.trans<-benthic.meter %>% group_by(Site,Transect,Code) %>% dplyr::summarize(cover.t=mean(cover.m))
benthic.site<-benthic.trans %>% group_by(Site,Code) %>% dplyr::summarize(n=n(),cover=mean(cover.t),cover.se=sd(cover.t)/sqrt(n)) %>% select(-n)

# Now need to spread data into wide site-level format
benthic.wide<-dcast(setDT(benthic.site),Site~Code,value.var=c("cover","cover.se"))
```

### 2.2 Algal canopy heights
```{r, echo=FALSE,message=FALSE,warning=FALSE}
benthic$Height<- as.numeric(benthic$Height)
canopy.site <- benthic %>% filter(Height>1) %>% group_by(Site,Transect,Meter,Code) %>% dplyr::summarize(canopy.m=mean(Height)) %>% group_by(Site,Transect,Code) %>% dplyr::summarize(canopy.t=mean(canopy.m)) %>% group_by(Site,Code) %>% dplyr::summarize(n=n(),canopy=mean(canopy.t),canopy.se=sd(canopy.t)/sqrt(n)) %>% select(-n)

# spread into wide site-level format, join with percent cover data in benthic.wide
canopy.wide <- dcast(setDT(canopy.site),Site~Code,value.var=c("canopy","canopy.se"))
site.benthic<- left_join(benthic.wide,canopy.wide,by="Site")
```

### 2.3 Rugosity
```{r, echo=FALSE,message=FALSE,warning=FALSE}
site.rugosity <- rugosity  %>% group_by(Site,Transect) %>% dplyr::summarize(rugosity.t=mean(Rugosity.cm)/100) %>% group_by(Site) %>% dplyr::summarize(n=n(),rugosity=mean(rugosity.t),rugosity.se=sd(rugosity.t)/sqrt(n)) %>% select(-n) %>% rbind(c("Pallaster West",1.7006,NA)) # missing data for Pallaster West, will use mean of Pallaster East and Mid because reef structure across all three was extremely similar
site.rugosity$rugosity<-as.numeric(as.character(site.rugosity$rugosity))
site.rugosity$rugosity.se<-as.numeric(as.character(site.rugosity$rugosity.se))
site.benthic <- full_join(site.benthic, site.rugosity,by="Site")
```

## Part 3. Compile into single df, combine with Bonaire data, and export csv
```{r, echo=FALSE,message=FALSE,warning=FALSE}
bonbenthic <- read.csv("Bonaire_benthic.csv",stringsAsFactors = FALSE)
site.benthic <- bind_rows(site.benthic,bonbenthic)
write.csv(site.benthic,file="site.benthic.csv")
```

### 2017-12-08 ABB_Fish.Rmd
updating files

---
title: "Antigua-Barbuda Fish Community Data: August 2017"
author: "Molly Wilson"
date: "8/8/2017"
output:
  html_document:
    df_print: paged
---
# Fish community data 
*required files: fish.csv*

## Part 1. Setup

### 1.1 Load packages
```{r, echo=FALSE,message=FALSE,warning=FALSE}
# set up
setwd("~/Google Drive/Research + Projects/Dissertation/ABB_Grazing/Analysis")
library(lazyeval)
library(rmarkdown)
library(tidyr)
library(ggplot2)
library(grid)
library(dplyr)
```

### 1.2 Import and format Antigua-Barbuda and Bonaire files
```{r, echo=FALSE,message=FALSE,warning=FALSE}
AB <-read.csv("fish.csv", stringsAsFactors=FALSE)
AB <- AB[rep(row.names(AB), AB$Number), 1:18] # expand to replicate rows if multiple fish were recorded
AB <- AB %>% select(-c(Time,Depth.ft,Surveyor,Number,Date))

bonherb<-read.csv("Bonaire_herbivores.csv", stringsAsFactors=FALSE) %>% filter(Date==2017) %>% select(-Date) 
boncarn<-read.csv("Bonaire_carnivores_2017.csv", stringsAsFactors=FALSE)
boncarn <- boncarn[rep(row.names(boncarn), boncarn$Number), 1:16] %>% select(-c(Number,Diver,Date))
bonaire <- bind_rows(bonherb,boncarn) %>% filter(Site=="Oil Slick" | Site=="Karpata" | Site=="Front Porch" | Site=="18th Palm")

# bind Bonaire and AB data
fish <- bind_rows(AB,bonaire)

fish[fish==""]  <- NA 
fish[fish=="#NUM!"] <- NA
fish$Biomass <- as.numeric(fish$Biomass)
denuder <- fish %>% filter(Family=="Scaridae" | Family=="Acanthuridae")
scarid <- fish %>% filter(Family=="Scaridae")
acan <- fish %>% filter(Family=="Acanthuridae")
carnivore <- fish %>% filter(Family=="Serranidae" | Family=="Lutjanidae" | Family=="Carangidae") # need to expand appropriately, could specify potential predators
```

## Part 2. Summary Data
*2.1 Total fish populations, by site
*2.2 Total fish populations, by species and site
*2.2 Total denuder populations, by site
*2.3 Total scarid populations, by site
*2.4 Scarid populations, by species and site
*2.5 Scarid populations, by phase and site
*2.6 Total carnivore populations, by site

### 2.1 Total fish population, by site
```{r,message=FALSE,warning=FALSE}
totfish.tran<- fish %>% filter(Biomass != 0) %>% group_by(Island,Site,Transect) %>% dplyr::summarize(BM.t=sum(Biomass)/1.2, DEN.t=n()/1.2, L.t=mean(Length))
site.totfish<-totfish.tran %>% group_by(Site) %>% dplyr::summarize(n=n(), tot.BM=mean(BM.t), tot.BM.se=(sd(BM.t))/sqrt(n), tot.DEN=mean(DEN.t), tot.DEN.se=(sd(DEN.t))/sqrt(n), tot.L=mean(L.t), tot.L.se=(sd(L.t))/sqrt(n)) %>% select(-n) # average across transects within each site
boxplot(BM.t~Site, las=2, data = totfish.tran)
```

### 2.2 Total fish populations, by species and site

```{r,message=FALSE,warning=FALSE}
# Because not all species were present at each site/transect, I need to create template to add 0 values for biomass and density in transects where they are absent.
expand.tran <- fish %>% group_by(Site,Transect) %>% dplyr::summarize()
expand.species <- fish %>% expand(nesting(Species,Family),Site)
temp.species <- full_join(expand.species,expand.tran,by="Site")

# Then I can join species/transect data with this species/transect template, and replace empty rows with 0 values

# biomass and density, by site and species:
species.tran<- fish %>% filter(Biomass != 0) %>% group_by(Site,Transect,Species) %>% dplyr::summarize(BM.t=sum(Biomass)/1.2, DEN.t=n()/1.2) %>% right_join(temp.species, by=c("Site","Transect","Species")) # join transect-level data with template
species.tran[is.na(species.tran)] <- 0 # convert NAs to 0 biomass/density in transects where a species is not present
species.site<-species.tran %>% group_by(Site,Species,Family) %>% dplyr::summarize(n=n(), BM=mean(BM.t), BM.se=(sd(BM.t))/sqrt(n), DEN=mean(DEN.t), DEN.se=(sd(DEN.t))/sqrt(n))

# length, by site and species (calculated separately because transects missing length values should not be replaced with 0 values)
species.L.tran <- fish %>% group_by(Site,Transect,Species) %>% dplyr::summarize(L.t=mean(Length))
species.L.site<-species.L.tran %>% group_by(Site,Species) %>% dplyr::summarize(L.n=n(),L=mean(L.t), L.se=(sd(L.t))/sqrt(L.n))
species.site <- species.site %>% left_join(species.L.site) # sample size will be different for length (L.n - only transects where species was observed) vs. biomass/density (n - all transects) calculations
```

### 2.3 Total denuder populations, by site
```{r,message=FALSE,warning=FALSE}
herb.tran<- denuder %>% filter(Biomass != 0) %>% group_by(Site,Transect) %>% dplyr::summarize(BM.t=sum(Biomass)/1.2, DEN.t=n()/1.2, L.t=mean(Length))
site.herb<-herb.tran %>% group_by(Site) %>% dplyr::summarize(n=n(), herb.BM=mean(BM.t), herb.BM.se=(sd(BM.t))/sqrt(n), herb.DEN=mean(DEN.t), herb.DEN.se=(sd(DEN.t))/sqrt(n), herb.L.s=mean(L.t), herb.L.se=(sd(L.t))/sqrt(n)) %>% select(-n)
```

### 2.3 Total acanthurid populations, by site
```{r,message=FALSE,warning=FALSE}
acan.tran<- acan %>% filter(Biomass != 0) %>% group_by(Site,Transect) %>% dplyr::summarize(BM.t=sum(Biomass)/1.2, DEN.t=n()/1.2, L.t=mean(Length))
site.acan<-acan.tran %>% group_by(Site) %>% dplyr::summarize(n=n(), acan.BM=mean(BM.t), acan.BM.se=(sd(BM.t))/sqrt(n), acan.DEN=mean(DEN.t), acan.DEN.se=(sd(DEN.t))/sqrt(n), acan.L=mean(L.t), acan.L.se=(sd(L.t))/sqrt(n)) %>% select(-n)
```

### 2.4 Total scarid populations, by site
```{r,message=FALSE,warning=FALSE}
scarid.tran<- scarid %>% filter(Biomass != 0) %>% group_by(Site,Transect) %>% dplyr::summarize(BM.t=sum(Biomass)/1.2, DEN.t=n()/1.2, L.t=mean(Length))
site.scar<-scarid.tran %>% group_by(Site) %>% dplyr::summarize(n=n(), scar.BM=mean(BM.t), scar.BM.se=(sd(BM.t))/sqrt(n), scar.DEN=mean(DEN.t), scar.DEN.se=(sd(DEN.t))/sqrt(n), scar.L=mean(L.t), scar.L.se=(sd(L.t))/sqrt(n)) %>% select(-n)
```

### 2.5 Scarid populations, by species and site
```{r,message=FALSE,warning=FALSE}
scarid.species <- species.site %>% filter(Family=="Scaridae") %>% select(-Family,-n) %>% ungroup()

# spread to site-level df with columns for each species' variables (ran into errors with unite/spread, so instead will do separately for each species)
viride <- scarid.species %>% filter(Species=="Sparisoma viride") %>% select(-Species)
names(viride)[2:8] <- c("viride.BM","viride.BM.se","viride.DEN","viride.DEN.se","viride.L.n","viride.L","viride.L.se")
vetula <- scarid.species %>% filter(Species=="Scarus vetula") %>% select(-Species)
names(vetula)[2:8] <- c("vetula.BM","vetula.BM.se","vetula.DEN","vetula.DEN.se","vetula.L.n","vetula.L","vetula.L.se")
aurofren <- scarid.species %>% filter(Species=="Sparisoma aurofrenatum") %>% select(-Species)
names(aurofren)[2:8] <- c("aurofren.BM","aurofren.BM.se","aurofren.DEN","aurofren.DEN.se","aurofren.L.n","aurofren.L","aurofren.L.se")
site.scar.sp <- left_join(viride,vetula,by="Site") %>% left_join(aurofren,by="Site")
```

### 2.6 Scarid populations, by phase and site
I will expand again here to make template with all phase/site combinations, then join transect level data
```{r,message=FALSE,warning=FALSE}
expand.tran <- fish %>% group_by(Site,Transect) %>% dplyr::summarize()
expand.phase <- scarid %>% expand(Phase,Site)
temp.phase <- full_join(expand.phase,expand.tran,by="Site")

# scarid biomass and density, by site and phase:
phase.tran <- scarid %>% filter(Biomass != 0) %>% group_by(Site,Transect,Phase) %>% dplyr::summarize(BM.t=sum(Biomass)/1.2, DEN.t=n()/1.2) %>% right_join(temp.phase, by=c("Site","Transect","Phase")) # join transect-level data with template
phase.tran[is.na(phase.tran)] <- 0 # convert NAs to 0 biomass/density in transects where a species is not present
phase.site<-phase.tran %>% group_by(Site,Phase) %>% dplyr::summarize(n=n(), BM=mean(BM.t), BM.se=(sd(BM.t))/sqrt(n), DEN=mean(DEN.t), DEN.se=(sd(DEN.t))/sqrt(n))

# length, by site and species (calculated separately because transects missing length values should not be replaced with 0 values)
phase.L.tran <- scarid %>% group_by(Site,Transect,Phase) %>% dplyr::summarize(L.t=mean(Length))
phase.L.site<-phase.L.tran %>% group_by(Site,Phase) %>% dplyr::summarize(L.n=n(),L=mean(L.t), L.se=(sd(L.t))/sqrt(L.n))
phase.site <- phase.site %>% left_join(phase.L.site) # sample size will be different for length (L.n - only transects where species was observed) vs. biomass/density (n - all transects) calculations

# spread to site-level df with columns for each species' variables (ran into errors with unite/spread, so instead will do separately for each species)
juvenile <- phase.site %>% filter(Phase=="j") %>% select(-Phase,-n)
names(juvenile)[2:8] <- c("juvenile.BM","juvenile.BM.se","juvenile.DEN","juvenile.DEN.se","juvenile.L.n","juvenile.L","juvenile.L.se")
initial <- phase.site %>% filter(Phase=="i") %>% select(-Phase,-n)
names(initial)[2:8] <- c("intial.BM","intial.BM.se","intial.DEN","intial.DEN.se","intial.L.n","intial.L","intial.L.se")
terminal <- phase.site %>% filter(Phase=="t") %>% select(-Phase,-n)
names(terminal)[2:8] <- c("terminal.BM","terminal.BM.se","terminal.DEN","terminal.DEN.se","terminal.L.n","terminal.L","terminal.L.se")
site.scar.ph <- left_join(juvenile,initial,by="Site") %>% left_join(terminal,by="Site")
```

### 2.7 Total carnivore populations, by site
I will need to use expanded template, because some transects had no serranid/lutjanids
*NEED TO ADD LENGTH HERE*
```{r,message=FALSE,warning=FALSE}
carnivore.tran<- carnivore %>% filter(Biomass != 0) %>% group_by(Site,Transect) %>% dplyr::summarize(BM.t=sum(Biomass)/1.2, DEN.t=n()/1.2) %>% right_join(expand.tran,by=c("Site","Transect"))
carnivore.tran[is.na(carnivore.tran)] <- 0 
site.carn<-carnivore.tran %>% group_by(Site) %>% dplyr::summarize(n=n(), carn.BM=mean(BM.t), carn.BM.se=(sd(BM.t))/sqrt(n), carn.DEN=mean(DEN.t), carn.DEN.se=(sd(DEN.t))/sqrt(n)) %>% select(-n)
```

## Part 3. Combine into one wide dataframe by site, export csv for integration with follow data
```{r,message=FALSE,warning=FALSE}
site.fish <- site.totfish %>% left_join(site.herb,by="Site") %>% left_join(site.acan,by="Site") %>% left_join(site.scar,by="Site") %>% left_join(site.scar.sp,by="Site") %>% left_join(site.scar.ph,by="Site") %>% left_join(site.carn,by="Site")

write.csv(site.fish,file="site.fish.csv")
write.csv(species.site,file="species.site.csv")
```

### 2017-12-08 ABB_Integration.Rmd
updating files

---
title: "Data integration"
output: html_document
---
# Integrating fish follow data with site-level fish, benthic, and rugosity data 
*Files: site.fish.csv, site.follows.csv, site.benthic.csv, site.human.csv, species.site.csv, sumID.csv*

## Part 1. Setup

```{r,message=FALSE,warning=FALSE}
# set up
setwd("~/Google Drive/Research + Projects/Dissertation/ABB_Grazing/Analysis")
library(lazyeval)
library(rmarkdown)
library(dplyr)
library(tidyr)
library(ggplot2)
library(grid)
```
### 1.2 Import and format files
```{r, echo=FALSE,message=FALSE,warning=FALSE}
site.follows<-read.csv("site.follows.csv",stringsAsFactors=FALSE)%>% select(-X)
site.fish<-read.csv("site.fish.csv", stringsAsFactors=FALSE)%>% select(-X)
site.benthic<-read.csv("site.benthic.csv", stringsAsFactors=FALSE)%>% select(-X)
site.human<-read.csv("site.human.csv", stringsAsFactors=FALSE)
species.site<-read.csv("species.site.csv",stringsAsFactors = FALSE)%>% filter(Species=="Sparisoma viride" | Species=="Scarus vetula" | Species=="Sparisoma aurofrenatum") %>% select(-X,-Family,-n,-L.n)
names(species.site)[3:8] <- c("conspec.BM","conspec.BM.se","conspec.DEN","conspec.DEN.se","conspec.L","conspec.L.se")
sumID <- read.csv("sumID.csv", stringsAsFactors=FALSE)%>% select(-X)
```
### 1.3 Join datasets, print .csv
```{r, echo=FALSE,message=FALSE,warning=FALSE}
# integrate all site-level variables
site.sum <- site.follows %>% left_join(site.fish,by="Site") %>% left_join(site.benthic,by="Site") %>% left_join(site.human,by="Site") 
# then join site-level data to sumID dataframe
sumID2 <- sumID %>% left_join(site.fish,by="Site") %>% left_join(site.benthic,by="Site") %>% left_join(site.human,by="Site")
# calculate conspecifc scarid variable to add to sumID2 (based on fish site/species)
sumID2 <- sumID2 %>% left_join(species.site,by=c("Site","Species"))

write.csv(site.sum,file="site.sum.csv")
write.csv(sumID2,file="sumID2.csv")
```

### 2017-12-08 Follows.Rmd
updating files

---
title: "Antigua-Barbuda Fish Follows August 2017"
author: "Molly Wilson"
date: "8/8/2017"
output:
  html_document:
    toc: true
    toc_float: true
    theme: united
    number_sections: true
---
```{r,message=FALSE,warning=FALSE}
# set up
library(lazyeval)
library(rmarkdown)
library(dplyr)
library(tidyr)
library(ggplot2)
library(grid)
library(reshape2)
library(data.table)
```

```{r, warning=FALSE, message= FALSE}
# import data
setwd("~/github/Scarid-Behavior/data")
follows <- read.csv("follows.csv", stringsAsFactors=FALSE) 
follows <- follows %>% filter(Date != "" & Date !="3/26/17") %>% select(-Activity.Code) %>% unite("fishID",c(Site,Diver,Date,Time,Follow.Number),sep=".", remove=FALSE)

# variable formatting
follows$Time.sec <- as.numeric(follows$Time.sec)
follows$Time.Fraction <- as.numeric(follows$Time.Fraction)
follows$Length.cm <- as.numeric(follows$Length.cm)
follows$Weight.g <- as.numeric(follows$Weight.g)
follows$Bites <- as.numeric(follows$Bites)
follows[c("Bites")][is.na(follows[c("Bites")])] <- 0
follows[follows == "#N/A"] <- NA

# create df with single row for all fish followed
sumID <- follows %>% group_by(fishID, Date, Island, Site, Species, Species.Code, Phase, Length.cm, Weight.g, Follow.sec) %>% dplyr::summarize()
# summarize sample size by site, species, phase
n.sspph <- group_by(sumID, Site, Species, Phase) %>% dplyr::summarize(n=n())
```

# Grazing behavior

## Time budgets 
Here I create a dataframe of time spent on each activity, per fish, in seconds and time fraction
```{r,warning=FALSE, message= FALSE}
TB <- group_by(follows, fishID, Activity, Species, Phase, Length.cm) %>% dplyr::summarize(time = sum(Time.sec), timefrac = sum(Time.Fraction))

# add time budgets per activity for each fish to 'sumID' (could re-do with dcast....)
TBgraze <- TB %>% filter(Activity == "graze") %>% subset(select=c(1,6,7))
colnames(TBgraze)[colnames(TBgraze)=="time"] <- "g.sec" # keep seconds for calculating grazing rates
colnames(TBgraze)[colnames(TBgraze)=="timefrac"] <- "g.frac"
sumID <- left_join(sumID, TBgraze, by= "fishID")
TBswim <- TB %>% filter(Activity == "swim") %>% subset(select=c(1,7))
colnames(TBswim)[colnames(TBswim)=="timefrac"] <- "swim.frac"
sumID <- left_join(sumID, TBswim, by= "fishID")
TBclean <- TB %>% filter(Activity == "cleaned") %>% subset(select=c(1,7))
colnames(TBclean)[colnames(TBclean)=="timefrac"] <- "clean.frac"
sumID <- left_join(sumID, TBclean, by= "fishID")
sumID[is.na(sumID)] <- 0
sumID$Length.cm[sumID$Length.cm==0]<-NA
```

## Grazing rates
```{r, warning=FALSE, message= FALSE}
# bites/second during total follow time and during grazing forays (including G1 instances)
bites <- follows %>% group_by(fishID) %>% dplyr::summarise(Bites = sum(Bites))
sumID <- left_join(sumID, bites, by="fishID") %>% mutate(BR=Bites/g.sec,FR=Bites/Follow.sec*60*60,TDB=Bites/Follow.sec*60*60*10.5)
sumID$BR[is.na(sumID$BR)] <- 0

# # bites/second during grazing forays, excluding G1 instances
# bites.noG1 <- follows %>% filter(Activity=="graze"&Bites>1&Time.sec>1) %>% group_by(fishID) %>% dplyr::summarize(Bites.noG1 = sum(Bites), g.sec.noG1=sum(Time.sec)) 
# sumID <- left_join(sumID, bites.noG1, by="fishID") %>% mutate(BR.g.noG1=Bites.noG1/g.sec.noG1)
# sumID$BR.g.noG1[is.na(sumID$BR.g.noG1)] <- 0
```

## Foray length and G1 bites
*Should I remove incomplete forays? What about forays >2min??*
```{r,warning=FALSE, message= FALSE}
forays <- follows %>% filter(Activity=="graze")
forays.complete <- follows %>% filter(!(Activity=="graze"&Min==0)&(lead(Activity)=="end")) # only complete forays, but not sure if I should actually exclude incomplete?
sum.forays <- forays %>% group_by(fishID) %>% dplyr::summarize(for.dur=mean(Time.sec),for.bites=mean(Bites))
sumID <- left_join(sumID,sum.forays, by="fishID")

# # frequency of G1 forays
# G1 <- forays %>% filter(Bites==1&Time.sec==1) %>% group_by(fishID) %>% dplyr::summarize(G1.freq=n())
# sumID <- left_join(sumID,G1, by="fishID")
# sumID$G1.freq[is.na(sumID$G1.freq)] <- 0
```

# Competitive interactions

## Summary of interactions per fish
```{r}
int.tot <- follows %>% filter(Initiator>=1) %>% group_by(fishID) %>% dplyr::summarise(int.tot=n())
int.scar <- follows %>% filter(Initiator>=1) %>% group_by(fishID) %>% filter(Competitor.Code=="stop" | Competitor.Code=="qup" | Competitor.Code=="rbp" | Competitor.Code=="prp" | Competitor.Code=="strp" | Competitor.Code=="rtp" | Competitor.Code=="ytp") %>% dplyr::summarise(int.scar=n())# total interaction instances
int.con <- follows %>% filter(Initiator>=1) %>% filter(Species==Competitor.Species) %>% group_by(fishID) %>% dplyr::summarise(int.con=n()) # interactions with conspecifics
int.hetscar <- follows %>% filter(Initiator>=1) %>% filter(Competitor.Code=="stop" | Competitor.Code=="qup" | Competitor.Code=="rbp") %>% filter(Species!=Competitor.Species) %>% group_by(fishID) %>% dplyr::summarise(int.hetscar=n()) # interactions with other scarid species
int.dam <- follows %>% filter(Initiator>=1) %>% filter(Competitor.Code=="dam") %>% group_by(fishID) %>% dplyr::summarise(int.dam=n()) # interactions with damselfish

# join and add to sumID dataframe
sumID <- sumID %>% left_join(int.tot, by="fishID") %>% left_join(int.scar, by="fishID") %>% left_join(int.con, by="fishID") %>% left_join(int.hetscar, by="fishID") %>% left_join(int.dam, by="fishID")
sumID[c("int.tot","int.scar","int.con","int.hetscar","int.dam")][is.na(sumID[c("int.tot","int.scar","int.con","int.hetscar","int.dam")])] <- 0
```

## Matrix of interactions by species
*eventually breakdown by site? region?*
```{r}
stop <- follows %>% filter(Species.Code=="stop") %>% dplyr::mutate(follows=n()) %>% filter(Competitor.Code %in% c("stop","qup","rbp","prp","rtp","strp")) %>% dplyr::group_by(Competitor.Code,Island) %>% dplyr::summarize(stop=n())
qup <- follows %>% filter(Species.Code=="qup") %>% dplyr::mutate(follows=n()) %>% filter(Competitor.Code %in% c("stop","qup","rbp","prp","rtp","strp")) %>% dplyr::group_by(Competitor.Code,Island) %>% dplyr::summarize(qup=n())
rbp <- follows %>% filter(Species.Code=="rbp") %>% dplyr::mutate(follows=n()) %>% filter(Competitor.Code %in% c("stop","qup","rbp","prp","rtp","strp")) %>% dplyr::group_by(Competitor.Code,Island) %>% dplyr::summarize(rbp=n())

intmatrix<-full_join(stop,qup,by=c("Competitor.Code","Island")) %>%full_join(rbp,by=c("Competitor.Code","Island"))
```

# Summarizing data by site, species, and phase
Next I create a summary tables that averages variables within each site/species/phase, then combines into one table
```{r,warning=FALSE, message= FALSE}
# site averages
fol.site <- sumID %>% group_by(Site,Island) %>% dplyr::summarize(n=n(),g.frac.mean=mean(g.frac),g.frac.se=sd(g.frac)/sqrt(n),BR.mean=mean(BR,na.rm=TRUE),BR.se=sd(BR,na.rm=TRUE)/sqrt(n),FR.mean=mean(FR),FR.se=sd(FR)/sqrt(n),for.dur.mean=mean(for.dur,na.rm=TRUE),for.dur.se=sd(for.dur,na.rm=TRUE)/sqrt(n),for.bites.mean=mean(for.bites,na.rm=TRUE),for.bites.se=sd(for.bites,na.rm=TRUE)/sqrt(n),int.tot.mean=mean(int.tot),int.tot.se=sd(int.tot)/sqrt(n), int.scar.mean=mean(int.scar),int.scar.se=sd(int.scar)/sqrt(n), int.con.mean=mean(int.con),int.con.se=sd(int.con)/sqrt(n), int.hetscar.mean=mean(int.hetscar),int.hetscar.se=sd(int.hetscar)/sqrt(n), int.dam.mean=mean(int.dam),int.dam.se=sd(int.dam)/sqrt(n)) %>% select(-n)

# site averages by species
fol.sp.site <- sumID %>% group_by(Site, Island, Species.Code) %>% dplyr::summarize(n=n(),g.frac.mean=mean(g.frac),g.frac.se=sd(g.frac)/sqrt(n),BR.mean=mean(BR,na.rm=TRUE),BR.se=sd(BR,na.rm=TRUE)/sqrt(n),FR.mean=mean(FR),FR.se=sd(FR)/sqrt(n),for.dur.mean=mean(for.dur,na.rm=TRUE),for.dur.se=sd(for.dur,na.rm=TRUE)/sqrt(n),for.bites.mean=mean(for.bites,na.rm=TRUE),for.bites.se=sd(for.bites,na.rm=TRUE)/sqrt(n), int.tot.mean=mean(int.tot),int.tot.se=sd(int.tot)/sqrt(n), int.scar.mean=mean(int.scar),int.scar.se=sd(int.scar)/sqrt(n), int.con.mean=mean(int.con),int.con.se=sd(int.con)/sqrt(n), int.hetscar.mean=mean(int.hetscar),int.hetscar.se=sd(int.hetscar)/sqrt(n), int.dam.mean=mean(int.dam),int.dam.se=sd(int.dam)/sqrt(n)) %>% select(-n) 
site.fol.sp<-dcast(setDT(fol.sp.site),Site~Species.Code,value.var=c("g.frac.mean","g.frac.se","int.scar.mean","int.scar.se"))

# site averages by species and phase
fol.sp.ph.site <- sumID %>% group_by(Site, Island, Species.Code, Phase) %>% dplyr::summarize(n=n(),g.frac.mean=mean(g.frac),g.frac.se=sd(g.frac)/sqrt(n),BR.mean=mean(BR,na.rm=TRUE),BR.se=sd(BR,na.rm=TRUE)/sqrt(n),FR.mean=mean(FR),FR.se=sd(FR)/sqrt(n),for.dur.mean=mean(for.dur,na.rm=TRUE),for.dur.se=sd(for.dur,na.rm=TRUE)/sqrt(n),for.bites.mean=mean(for.bites,na.rm=TRUE),for.bites.se=sd(for.bites,na.rm=TRUE)/sqrt(n), int.tot.mean=mean(int.tot),int.tot.se=sd(int.tot)/sqrt(n), int.scar.mean=mean(int.scar),int.scar.se=sd(int.scar)/sqrt(n), int.con.mean=mean(int.con),int.con.se=sd(int.con)/sqrt(n), int.hetscar.mean=mean(int.hetscar),int.hetscar.se=sd(int.hetscar)/sqrt(n), int.dam.mean=mean(int.dam),int.dam.se=sd(int.dam)/sqrt(n)) %>% select(-n)
site.fol.sp.ph<-dcast(setDT(fol.sp.ph.site),Site~Species.Code+Phase,value.var=c("g.frac.mean","g.frac.se","int.scar.mean","int.scar.se"))

# join into single file
site.follows <- fol.site %>% left_join(site.fol.sp, by="Site") %>% left_join(site.fol.sp.ph, by="Site")
```

## Export csvs for integration with site data
```{r,message=FALSE,warning=FALSE}
# write.csv(sumID,file="sumID.csv")
# write.csv(fol.sp.site,file="fol.sp.site.csv")
# write.csv(fol.sp.ph.site,file="fol.sp.ph.site.csv")
```

### 1.2 Import and format files
```{r, echo=FALSE,message=FALSE,warning=FALSE}
setwd("~/github/Scarid-Behavior/data")
site.fish<-read.csv("site.fish.csv", stringsAsFactors=FALSE)%>% select(-X)
site.benthic<-read.csv("site.benthic.csv", stringsAsFactors=FALSE)%>% select(-X)
site.human<-read.csv("site.human.csv", stringsAsFactors=FALSE)
species.site<-read.csv("species.site.csv",stringsAsFactors = FALSE)%>% filter(Species=="Sparisoma viride" | Species=="Scarus vetula" | Species=="Sparisoma aurofrenatum") %>% select(-X,-Family,-n,-L.n)
names(species.site)[3:8] <- c("conspec.BM","conspec.BM.se","conspec.DEN","conspec.DEN.se","conspec.L","conspec.L.se")
```
### 1.3 Join datasets, print .csv
```{r, echo=FALSE,message=FALSE,warning=FALSE}
# integrate all site-level variables
site.sum <- site.follows %>% left_join(site.fish,by="Site") %>% left_join(site.benthic,by="Site") %>% left_join(site.human,by="Site") 
# then join site-level data to sumID dataframe
sumID2 <- sumID %>% left_join(site.fish,by="Site") %>% left_join(site.benthic,by="Site") %>% left_join(site.human,by="Site")
# calculate conspecifc scarid variable to add to sumID2 (based on fish site/species)
sumID2 <- sumID2 %>% left_join(species.site,by=c("Site","Species"))

write.csv(site.sum,file="site.sum.csv")
write.csv(sumID2,file="sumID2.csv")
```

### 2017-12-08 Sites.Rmd
updating files

---
title: "Fish & Benthic Community Data: Antigua, Barbuda, & Bonaire - Aug. 2017"
author: "Molly Wilson"
date: "8/8/2017"
output:
  html_document:
    toc: true
    toc_float: true
    theme: united
    number_sections: true
---
# Fish community data 
*required files: fish.csv,

```{r, echo=FALSE,message=FALSE,warning=FALSE}
# set up
library(rmarkdown)
library(dplyr)
library(lazyeval)
library(tidyr)
library(ggplot2)
library(grid)
library(reshape2)
library(data.table)
```
```{r, echo=FALSE,message=FALSE,warning=FALSE}
# import files
setwd("~/github/Scarid-Behavior/data")
AB <-read.csv("fish.csv", stringsAsFactors=FALSE)
AB <- AB[rep(row.names(AB), AB$Number), 1:18] # expand to replicate rows if multiple fish were recorded
AB <- AB %>% select(-c(Time,Depth.ft,Surveyor,Number,Date))

bonherb<-read.csv("bon_herb.csv", stringsAsFactors=FALSE) %>% filter(Date==2017) %>% select(-Date) 
boncarn<-read.csv("bon_carn.csv", stringsAsFactors=FALSE)
boncarn <- boncarn[rep(row.names(boncarn), boncarn$Number), 1:16] %>% select(-c(Number,Diver,Date))
bonaire <- bind_rows(bonherb,boncarn) %>% filter(Site=="Oil Slick" | Site=="Karpata" | Site=="Front Porch" | Site=="18th Palm")

# bind Bonaire and AB data
fish <- bind_rows(AB,bonaire)

fish[fish==""]  <- NA 
fish[fish=="#NUM!"] <- NA
fish$Biomass <- as.numeric(fish$Biomass)
denuder <- fish %>% filter(Family=="Scaridae" | Family=="Acanthuridae")
scarid <- fish %>% filter(Family=="Scaridae")
acan <- fish %>% filter(Family=="Acanthuridae")
carnivore <- fish %>% filter(Family=="Serranidae" | Family=="Lutjanidae" | Family=="Carangidae") # need to expand appropriately, could specify potential predators
```

## Total fish population, by site
```{r,message=FALSE,warning=FALSE}
totfish.tran<- fish %>% filter(Biomass != 0) %>% group_by(Island,Site,Transect) %>% dplyr::summarize(BM.t=sum(Biomass)/1.2, DEN.t=n()/1.2, L.t=mean(Length))
site.totfish<-totfish.tran %>% group_by(Site) %>% dplyr::summarize(n=n(), tot.BM=mean(BM.t), tot.BM.se=(sd(BM.t))/sqrt(n), tot.DEN=mean(DEN.t), tot.DEN.se=(sd(DEN.t))/sqrt(n), tot.L=mean(L.t), tot.L.se=(sd(L.t))/sqrt(n)) %>% select(-n) # average across transects within each site
boxplot(BM.t~Site, las=2, data = totfish.tran)
```

## Total fish populations, by species and site
```{r,message=FALSE,warning=FALSE}
# Because not all species were present at each site/transect, I need to create template to add 0 values for biomass and density in transects where they are absent.
expand.tran <- fish %>% group_by(Site,Transect) %>% dplyr::summarize()
expand.species <- fish %>% expand(nesting(Species,Family),Site)
temp.species <- full_join(expand.species,expand.tran,by="Site")

# Then I can join species/transect data with this species/transect template, and replace empty rows with 0 values

# biomass and density, by site and species:
species.tran<- fish %>% filter(Biomass != 0) %>% group_by(Site,Transect,Species) %>% dplyr::summarize(BM.t=sum(Biomass)/1.2, DEN.t=n()/1.2) %>% right_join(temp.species, by=c("Site","Transect","Species")) # join transect-level data with template
species.tran[is.na(species.tran)] <- 0 # convert NAs to 0 biomass/density in transects where a species is not present
species.site<-species.tran %>% group_by(Site,Species,Family) %>% dplyr::summarize(n=n(), BM=mean(BM.t), BM.se=(sd(BM.t))/sqrt(n), DEN=mean(DEN.t), DEN.se=(sd(DEN.t))/sqrt(n))

# length, by site and species (calculated separately because transects missing length values should not be replaced with 0 values)
species.L.tran <- fish %>% group_by(Site,Transect,Species) %>% dplyr::summarize(L.t=mean(Length))
species.L.site<-species.L.tran %>% group_by(Site,Species) %>% dplyr::summarize(L.n=n(),L=mean(L.t), L.se=(sd(L.t))/sqrt(L.n))
species.site <- species.site %>% left_join(species.L.site) # sample size will be different for length (L.n - only transects where species was observed) vs. biomass/density (n - all transects) calculations
```

## Total denuder populations, by site
```{r,message=FALSE,warning=FALSE}
herb.tran<- denuder %>% filter(Biomass != 0) %>% group_by(Site,Transect) %>% dplyr::summarize(BM.t=sum(Biomass)/1.2, DEN.t=n()/1.2, L.t=mean(Length))
site.herb<-herb.tran %>% group_by(Site) %>% dplyr::summarize(n=n(), herb.BM=mean(BM.t), herb.BM.se=(sd(BM.t))/sqrt(n), herb.DEN=mean(DEN.t), herb.DEN.se=(sd(DEN.t))/sqrt(n), herb.L.s=mean(L.t), herb.L.se=(sd(L.t))/sqrt(n)) %>% select(-n)
```

## Total acanthurid populations, by site
```{r,message=FALSE,warning=FALSE}
acan.tran<- acan %>% filter(Biomass != 0) %>% group_by(Site,Transect) %>% dplyr::summarize(BM.t=sum(Biomass)/1.2, DEN.t=n()/1.2, L.t=mean(Length))
site.acan<-acan.tran %>% group_by(Site) %>% dplyr::summarize(n=n(), acan.BM=mean(BM.t), acan.BM.se=(sd(BM.t))/sqrt(n), acan.DEN=mean(DEN.t), acan.DEN.se=(sd(DEN.t))/sqrt(n), acan.L=mean(L.t), acan.L.se=(sd(L.t))/sqrt(n)) %>% select(-n)
```

## Total scarid populations, by site
```{r,message=FALSE,warning=FALSE}
scarid.tran<- scarid %>% filter(Biomass != 0) %>% group_by(Site,Transect) %>% dplyr::summarize(BM.t=sum(Biomass)/1.2, DEN.t=n()/1.2, L.t=mean(Length))
site.scar<-scarid.tran %>% group_by(Site) %>% dplyr::summarize(n=n(), scar.BM=mean(BM.t), scar.BM.se=(sd(BM.t))/sqrt(n), scar.DEN=mean(DEN.t), scar.DEN.se=(sd(DEN.t))/sqrt(n), scar.L=mean(L.t), scar.L.se=(sd(L.t))/sqrt(n)) %>% select(-n)
```

## Scarid populations, by species and site
```{r,message=FALSE,warning=FALSE}
scarid.species <- species.site %>% filter(Family=="Scaridae") %>% select(-Family,-n) %>% ungroup()

# spread to site-level df with columns for each species' variables (ran into errors with unite/spread, so instead will do separately for each species)
viride <- scarid.species %>% filter(Species=="Sparisoma viride") %>% select(-Species)
names(viride)[2:8] <- c("viride.BM","viride.BM.se","viride.DEN","viride.DEN.se","viride.L.n","viride.L","viride.L.se")
vetula <- scarid.species %>% filter(Species=="Scarus vetula") %>% select(-Species)
names(vetula)[2:8] <- c("vetula.BM","vetula.BM.se","vetula.DEN","vetula.DEN.se","vetula.L.n","vetula.L","vetula.L.se")
aurofren <- scarid.species %>% filter(Species=="Sparisoma aurofrenatum") %>% select(-Species)
names(aurofren)[2:8] <- c("aurofren.BM","aurofren.BM.se","aurofren.DEN","aurofren.DEN.se","aurofren.L.n","aurofren.L","aurofren.L.se")
site.scar.sp <- left_join(viride,vetula,by="Site") %>% left_join(aurofren,by="Site")
```

## Scarid populations, by phase and site
I will expand again here to make template with all phase/site combinations, then join transect level data
```{r,message=FALSE,warning=FALSE}
expand.tran <- fish %>% group_by(Site,Transect) %>% dplyr::summarize()
expand.phase <- scarid %>% expand(Phase,Site)
temp.phase <- full_join(expand.phase,expand.tran,by="Site")

# scarid biomass and density, by site and phase:
phase.tran <- scarid %>% filter(Biomass != 0) %>% group_by(Site,Transect,Phase) %>% dplyr::summarize(BM.t=sum(Biomass)/1.2, DEN.t=n()/1.2) %>% right_join(temp.phase, by=c("Site","Transect","Phase")) # join transect-level data with template
phase.tran[is.na(phase.tran)] <- 0 # convert NAs to 0 biomass/density in transects where a species is not present
phase.site<-phase.tran %>% group_by(Site,Phase) %>% dplyr::summarize(n=n(), BM=mean(BM.t), BM.se=(sd(BM.t))/sqrt(n), DEN=mean(DEN.t), DEN.se=(sd(DEN.t))/sqrt(n))

# length, by site and species (calculated separately because transects missing length values should not be replaced with 0 values)
phase.L.tran <- scarid %>% group_by(Site,Transect,Phase) %>% dplyr::summarize(L.t=mean(Length))
phase.L.site<-phase.L.tran %>% group_by(Site,Phase) %>% dplyr::summarize(L.n=n(),L=mean(L.t), L.se=(sd(L.t))/sqrt(L.n))
phase.site <- phase.site %>% left_join(phase.L.site) # sample size will be different for length (L.n - only transects where species was observed) vs. biomass/density (n - all transects) calculations

# spread to site-level df with columns for each species' variables (ran into errors with unite/spread, so instead will do separately for each species)
juvenile <- phase.site %>% filter(Phase=="j") %>% select(-Phase,-n)
names(juvenile)[2:8] <- c("juvenile.BM","juvenile.BM.se","juvenile.DEN","juvenile.DEN.se","juvenile.L.n","juvenile.L","juvenile.L.se")
initial <- phase.site %>% filter(Phase=="i") %>% select(-Phase,-n)
names(initial)[2:8] <- c("intial.BM","intial.BM.se","intial.DEN","intial.DEN.se","intial.L.n","intial.L","intial.L.se")
terminal <- phase.site %>% filter(Phase=="t") %>% select(-Phase,-n)
names(terminal)[2:8] <- c("terminal.BM","terminal.BM.se","terminal.DEN","terminal.DEN.se","terminal.L.n","terminal.L","terminal.L.se")
site.scar.ph <- left_join(juvenile,initial,by="Site") %>% left_join(terminal,by="Site")
```

## Total carnivore populations, by site
I will need to use expanded template, because some transects had no serranid/lutjanids
*Could add gape size restrictions to determine actual predator threat?*
```{r,message=FALSE,warning=FALSE}
carnivore.tran<- carnivore %>% filter(Biomass != 0) %>% group_by(Site,Transect) %>% dplyr::summarize(BM.t=sum(Biomass)/1.2, DEN.t=n()/1.2) %>% right_join(expand.tran,by=c("Site","Transect"))
carnivore.tran[is.na(carnivore.tran)] <- 0 
site.carn<-carnivore.tran %>% group_by(Site) %>% dplyr::summarize(n=n(), carn.BM=mean(BM.t), carn.BM.se=(sd(BM.t))/sqrt(n), carn.DEN=mean(DEN.t), carn.DEN.se=(sd(DEN.t))/sqrt(n)) %>% select(-n)
```

# Benthic & rugosity data
*required files: benthic_anu.csv, benthic_bon.csv, rugosity.csv*

```{r, echo=FALSE,message=FALSE,warning=FALSE}
setwd("~/github/Scarid-Behavior/data")
benthic<-read.csv("benthic_anu.csv", stringsAsFactors=FALSE)
rugosity<-read.csv("rugosity.csv", stringsAsFactors=FALSE)
```

## Percent cover
```{r, echo=FALSE,message=FALSE,warning=FALSE}
# Because not all categories were present within each meter, I need to create template to add 0 values in meters where they were absent.
expand.meter <- benthic %>% group_by(Site,Transect,Meter) %>% dplyr::summarize()
expand.category <- benthic %>% expand(nesting(Code),Site)
temp.benthic <- full_join(expand.meter,expand.category,by="Site")

# Then I can join benthic data with this template, and replace empty rows with 0 values
benthic.meter<- benthic %>% group_by(Site,Transect,Meter,Code) %>% dplyr::summarize(count.m=n()) %>% right_join(temp.benthic, by=c("Site","Transect","Meter","Code")) # join transect-level data with template
benthic.meter[is.na(benthic.meter)] <- 0 # convert NAs to 0 biomass/density in transects where a species is not present

# Calculating percent cover on potential substrate (i.e. not sand), remove meters that were 100% sand cover (no potential susbtrate)
sand.meter <- benthic.meter %>% filter(Code=="sand") %>% dplyr::rename(count.sand=count.m) %>% select(-Code) 
benthic.meter <- left_join(benthic.meter,sand.meter,by=c("Site","Transect","Meter")) %>% mutate(cover.m=100*count.m/(10-count.sand)) %>% filter(!is.nan(cover.m)) %>% filter(Code!="sand")

benthic.trans<-benthic.meter %>% group_by(Site,Transect,Code) %>% dplyr::summarize(cover.t=mean(cover.m))
benthic.site<-benthic.trans %>% group_by(Site,Code) %>% dplyr::summarize(n=n(),cover=mean(cover.t),cover.se=sd(cover.t)/sqrt(n)) %>% select(-n)

# Now need to spread data into wide site-level format
benthic.wide<-dcast(setDT(benthic.site),Site~Code,value.var=c("cover","cover.se"))
```

## Algal canopy heights
```{r, echo=FALSE,message=FALSE,warning=FALSE}
benthic$Height<- as.numeric(benthic$Height)
canopy.site <- benthic %>% filter(Height>1) %>% group_by(Site,Transect,Meter,Code) %>% dplyr::summarize(canopy.m=mean(Height)) %>% group_by(Site,Transect,Code) %>% dplyr::summarize(canopy.t=mean(canopy.m)) %>% group_by(Site,Code) %>% dplyr::summarize(n=n(),canopy=mean(canopy.t),canopy.se=sd(canopy.t)/sqrt(n)) %>% select(-n)

# spread into wide site-level format, join with percent cover data in benthic.wide
canopy.wide <- dcast(setDT(canopy.site),Site~Code,value.var=c("canopy","canopy.se"))
site.benthic<- left_join(benthic.wide,canopy.wide,by="Site")
```

## Rugosity
```{r, echo=FALSE,message=FALSE,warning=FALSE}
site.rugosity <- rugosity  %>% group_by(Site,Transect) %>% dplyr::summarize(rugosity.t=mean(Rugosity.cm)/100) %>% group_by(Site) %>% dplyr::summarize(n=n(),rugosity=mean(rugosity.t),rugosity.se=sd(rugosity.t)/sqrt(n)) %>% select(-n) %>% rbind(c("Pallaster West",1.7006,NA)) # missing data for Pallaster West, will use mean of Pallaster East and Mid because reef structure across all three was extremely similar
site.rugosity$rugosity<-as.numeric(as.character(site.rugosity$rugosity))
site.rugosity$rugosity.se<-as.numeric(as.character(site.rugosity$rugosity.se))
site.benthic <- full_join(site.benthic, site.rugosity,by="Site")
```

# Integrating and exporting.csvs
```{r, echo=FALSE,message=FALSE,warning=FALSE}
setwd("~/github/Scarid-Behavior/data")
bonbenthic <- read.csv("benthic_bon.csv",stringsAsFactors = FALSE)
site.benthic <- bind_rows(site.benthic,bonbenthic)
write.csv(site.benthic,file="site.benthic.csv")

site.fish <- site.totfish %>% left_join(site.herb,by="Site") %>% left_join(site.acan,by="Site") %>% left_join(site.scar,by="Site") %>% left_join(site.scar.sp,by="Site") %>% left_join(site.scar.ph,by="Site") %>% left_join(site.carn,by="Site")
write.csv(site.fish,file="site.fish.csv")
write.csv(species.site,file="species.site.csv")
```

### 2017-12-08 Graphing+Analysis.Rmd
updating files

---
title: "ABB Analysis Tasks"
author: "Molly Wilson"
date: "12/5/2017"
output:
  html_document:
    toc: true
    toc_float: true
    theme: united
    number_sections: true

---

```{r,message=FALSE,warning=FALSE, echo=FALSE}
# set up
library(rmarkdown)
library(dplyr)
library(tidyr)
library(lazyeval)
library(ggplot2)
library(ellipse)
library(RColorBrewer)
library(ggbiplot)
library(corrplot)
library(nlme)
library(pls)
library(factoextra)
```
```{r,message=FALSE,warning=FALSE, echo=FALSE}
setwd("~/github/Scarid-Behavior/data")
site.sum <- read.csv("site.sum.csv", stringsAsFactors=FALSE) %>% select(-X)
data <- read.csv("sumID2.csv", stringsAsFactors=FALSE) %>% select(Island,Site,Spearfishing,Diving,Species,Species.Code,Phase,Length.cm,g.frac,FR,BR,TDB,for.dur,for.bites,tot.BM,tot.DEN,scar.BM,herb.BM,scar.DEN,conspec.BM,conspec.DEN,carn.BM,carn.DEN,cover_CCA,cover_LC,cover_MA,cover_TA,canopy_MA,canopy_TA,rugosity,int.tot,int.scar,int.con,int.hetscar,int.dam)
```

# Data overview:
- Fish follows (2 min) conducted at sites in Antigua (6), Barbuda (3) and Bonaire (4) from March-August 2017
- Follows tracked time spent grazing, bite rates, and competitive interactions (among scarids and with damselfish)
- Follows targeted *Sparisoma viride*, *Scarus vetula*, and *Sparisoma aurofrenatum* of both initial and terminal phases 
  +*Sparisoma aurofrenatum* were not followed in Bonaire, but added in Antigua and Bonaire due to low abundances of other two species. Some sites in Antigua and Barbuda had no/low represenation from terminal phase *viride* and *vetula*
- Site level factors (benthic, fish, and rugosity) assessed at each of the 13 sites
```{r,echo=FALSE}
n.sspph <- data %>% group_by(Island, Site, Species, Phase) %>% dplyr::summarize(n=n())
ggplot(data=n.sspph, aes(x=Site, y=n, fill=Phase)) +
    geom_bar(stat="identity", position=position_dodge()) + facet_grid(.~Species) + theme_minimal() + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + scale_fill_grey()
# counts <- table(data$Species, data$Site)
# barplot(counts, main="Sample size by phase and species",
#   legend = rownames(counts), beside=TRUE, las=2)
```

# Examing variable distributions and relationships

## Site-level predictor variables: fish, benthic, and rugosity

### Distributions
```{r,warning=FALSE,echo=FALSE}
predictors <- site.sum %>% select(tot.BM,tot.DEN,scar.BM,scar.DEN,carn.BM,carn.DEN,cover_CCA,cover_LC,cover_MA,canopy_MA,cover_TA,canopy_TA,rugosity) %>% gather() %>% na.omit()
ggplot(predictors,aes(value))+facet_wrap(~ key, scales = "free")+geom_histogram(aes(y=..density..), colour="black", fill="white")+
 geom_density(alpha=.2, fill="#FF6666") 
```

### Correlations
```{r,echo=FALSE}
site.var <- data %>% select(tot.BM,tot.DEN,scar.BM,scar.DEN,conspec.BM,conspec.DEN,carn.BM,carn.DEN,cover_CCA,cover_LC,cover_MA,canopy_MA,cover_TA,canopy_TA,rugosity)
cor=cor(site.var, use="pairwise.complete.obs")
my_colors <- brewer.pal(5, "Spectral")
my_colors=colorRampPalette(my_colors)(100)
ord <- order(cor[1, ])
cor_ord = cor[ord, ord]
plotcorr(cor_ord , col=my_colors[cor_ord*50+50] , mar=c(1,1,1,1)  )
```

Many correlations here, both within benthic/fish categories and across categories. Rugosity is correlated with both fish and benthic variables. I will examine correlations within fish community and benthic community variables to decide which I can omit and which I will integrate through PCA. Whether I include benthic variables as predictor or response variables will determine if I need to worry about fish~benthic correlations.

```{r,echo=FALSE}
panel.cor <- function(x, y, digits=2, prefix="", cex.cor, ...)
{
    usr <- par("usr"); on.exit(par(usr))
    par(usr = c(0, 1, 0, 1))
    r <- abs(cor(x, y))
    txt <- format(c(r, 0.123456789), digits=digits)[1]
    txt <- paste(prefix, txt, sep="")
    if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
    text(0.5, 0.5, txt, cex = cex.cor * r)
}
pairs(~tot.BM+tot.DEN+scar.BM+scar.DEN+conspec.BM+conspec.DEN+carn.BM+carn.DEN+rugosity, data=data,
      lower.panel=panel.smooth, upper.panel=panel.cor, 
      pch=20, main="Fish Community Variables",na.action = na.omit)
```

Variables carn.DEN and carn.BM are essentially identical - I will remove carn.DEN and use carn.BM as the main preditor metric. Total fish biomass and density (tot.BM and tot.DEN) are highly correlated with both scarid and carnivore variables, so I will remove them. Conspecific biomass is highly correlated with scarid biomass so I will remove it as well (and should possibly also remove conspec.DEN?). The selected variables I will proceed with are as follows:
```{r,echo=FALSE}
panel.cor <- function(x, y, digits=2, prefix="", cex.cor, ...)
{
    usr <- par("usr"); on.exit(par(usr))
    par(usr = c(0, 1, 0, 1))
    r <- abs(cor(x, y))
    txt <- format(c(r, 0.123456789), digits=digits)[1]
    txt <- paste(prefix, txt, sep="")
    if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
    text(0.5, 0.5, txt, cex = cex.cor * r)
}
pairs(~scar.BM+scar.DEN+conspec.DEN+carn.BM+rugosity, data=data,
      lower.panel=panel.smooth, upper.panel=panel.cor, 
      pch=20, main="Selected Fish Community Variables",na.action = na.omit)
```

Scarid biomass, carnivore biomass, and rugosity should be integrated via PCA. Scarid density and conspecific density should also be combined, or one should be omitted.

```{r,echo=FALSE}
panel.cor <- function(x, y, digits=2, prefix="", cex.cor, ...)
{
    usr <- par("usr"); on.exit(par(usr))
    par(usr = c(0, 1, 0, 1))
    r <- abs(cor(x, y))
    txt <- format(c(r, 0.123456789), digits=digits)[1]
    txt <- paste(prefix, txt, sep="")
    if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
    text(0.5, 0.5, txt, cex = cex.cor * r)
}
pairs(~cover_CCA+cover_LC+cover_MA+canopy_MA+cover_TA+canopy_TA+rugosity, data=data,
      lower.panel=panel.smooth, upper.panel=panel.cor, 
      pch=20, main="Benthic Community Variables",na.action = na.omit)
```

All benthic variables are highly correlated. I can omit cover_CCA and cover_MA (accounted for by canopy_MA). The rest should be combined via PCA. First I will assess remaining correlations between selected benthic and fish community variables:

```{r,echo=FALSE}
panel.cor <- function(x, y, digits=2, prefix="", cex.cor, ...)
{
    usr <- par("usr"); on.exit(par(usr))
    par(usr = c(0, 1, 0, 1))
    r <- abs(cor(x, y))
    txt <- format(c(r, 0.123456789), digits=digits)[1]
    txt <- paste(prefix, txt, sep="")
    if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
    text(0.5, 0.5, txt, cex = cex.cor * r)
}
pairs(~scar.BM+scar.DEN+carn.BM+cover_LC+cover_TA+canopy_MA+canopy_TA+rugosity, data=data,
      lower.panel=panel.smooth, upper.panel=panel.cor, 
      pch=20, main="Selected Fish and Benthic Variables",na.action = na.omit)
```

Unsure how to proceed here if I do end up using both benthic and fish community variables as predictors, because some are corellated (esp. scarid.BM.s and all benthic variables....). It looks like I would integrate scar.BM and carn.BM into a PCA with all other benthic/rugosity variables. Variables scar.DEN and conspec.DEN are relatively independent of all other variables (except each other).

Current outline for site-level PCA tasks:
If benthic=predictor:
- cover_LC + cover_TA + canopy_MA + canopy_TA + rugosity + scar.BM + carn.BM
- scar.DEN + conspec.DEN
If benthic=response:
- rugosity + scar.BM + carn.BM
- cover_LC + cover_TA + canopy_MA + canopy_TA
- scar.DEN + conspec.DEN

## Fish-level response variables: grazing behaviors

### Distributions
```{r,echo=FALSE,warning=FALSE}
grazing <- data %>% select(g.frac,FR,BR,for.dur,for.bites) %>% gather() %>% na.omit()
ggplot(grazing,aes(value))+facet_wrap(~ key, scales = "free")+ geom_histogram(aes(y=..density..), colour="black", fill="white")+
 geom_density(alpha=.2, fill="#FF6666") 
```

### Corellations
```{r,echo=FALSE}
grazing <- data %>% select(g.frac,FR,BR,for.dur,for.bites)
cor=cor(grazing, use="pairwise.complete.obs")
ord <- order(cor[1, ])
cor_ord = cor[ord, ord]
plotcorr(cor_ord , col=my_colors[cor_ord*50+50] , mar=c(1,1,1,1)  )
```

```{r,echo=FALSE}
panel.cor <- function(x, y, digits=2, prefix="", cex.cor, ...)
{
    usr <- par("usr"); on.exit(par(usr))
    par(usr = c(0, 1, 0, 1))
    r <- abs(cor(x, y))
    txt <- format(c(r, 0.123456789), digits=digits)[1]
    txt <- paste(prefix, txt, sep="")
    if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
    text(0.5, 0.5, txt, cex = cex.cor * r)
}
pairs(~g.frac+FR+BR+for.dur+for.bites, data=data,
      lower.panel=panel.smooth, upper.panel=panel.cor, 
      pch=20, main="Grazing Response Variables",na.action = na.omit)
```

Variable for.bites is correlated with nearly all other variables, so will be omitted. Variables g.farc, FR. and for.dur are still correlated and should be assessed with a PCA. Variable BR can be a standalone variable.

```{r,echo=FALSE}
panel.cor <- function(x, y, digits=2, prefix="", cex.cor, ...)
{
    usr <- par("usr"); on.exit(par(usr))
    par(usr = c(0, 1, 0, 1))
    r <- abs(cor(x, y))
    txt <- format(c(r, 0.123456789), digits=digits)[1]
    txt <- paste(prefix, txt, sep="")
    if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
    text(0.5, 0.5, txt, cex = cex.cor * r)
}
pairs(~g.frac+FR+BR+for.dur, data=data,
      lower.panel=panel.smooth, upper.panel=panel.cor, 
      pch=20, main="Selected Grazing Response Variables",na.action = na.omit)
```

Grazing PCA:
- g.frac + FR + for.dur
- (BR as standalone)

# Exploratory bivariate plots

## Grazing as a function of fish length
```{r, echo=FALSE,warning=FALSE, eval=FALSE}
ggplot(data,aes(x=Length.cm,y=FR,colour=Island))+facet_grid(.~Species)+geom_point(shape=18)+scale_shape(solid=FALSE)+geom_smooth(method=lm, fullrange=FALSE)+ggtitle("Feeding rate by fish length") + theme_minimal() +labs(y="Feeding Rate (bites/hr)",x="Length (cm)") + theme(plot.title = element_text(hjust = 0.5))  + scale_color_manual(values=c("navy", "darkseagreen", "slategray2"))
```
```{r, echo=FALSE,warning=FALSE, eval=FALSE}
ggplot(data,aes(x=Length.cm,y=BR,colour=Island))+facet_grid(.~Species)+geom_point(shape=18)+scale_shape(solid=FALSE)+geom_smooth(method=lm, fullrange=FALSE)+ggtitle("Bite rate (while grazing) by fish length") + theme_minimal() +labs(y="Bite Rate (bites/sec)",x="Length (cm)") + theme(plot.title = element_text(hjust = 0.5))  + scale_color_manual(values=c("navy", "darkseagreen", "slategray2"))
```

```{r, echo=FALSE,warning=FALSE, eval=FALSE}
ggplot(data,aes(x=Length.cm,y=g.frac,colour=Island))+facet_grid(.~Species)+geom_point(shape=18)+scale_shape(solid=FALSE)+geom_smooth(method=lm, fullrange=FALSE)+ggtitle("Fraction of time spend grazing by fish length") + theme_minimal() +labs(y="Time Fraction",x="Length (cm)") + theme(plot.title = element_text(hjust = 0.5))  + scale_color_manual(values=c("navy", "darkseagreen", "slategray2"))
```

## Grazing as a function of scarid biomass and density

```{r}
ggplot(data,aes(x=scar.BM,y=FR))+facet_grid(.~Species)+geom_smooth(method="lm")+geom_point(shape=18)+ggtitle("Feeding rate by scarid biomass") + theme_minimal() +labs(y="Feeding rate (bites/hr)",x="Scarid Biomass (g/100m2)") + theme(plot.title = element_text(hjust = 0.5))
```

```{r,echo=FALSE,warning=FALSE}
ggplot(data,aes(x=scar.BM,y=FR,colour=Island))+facet_grid(.~Species)+geom_point(shape=18)+geom_smooth(method=lm, fullrange=FALSE)+ggtitle("Feeding rate by phase and scarid biomass") + theme_minimal() +labs(y="Feeding rate (bites/hr)",x="Scarid Biomass (g/100m2)") + theme(plot.title = element_text(hjust = 0.5)) + scale_color_manual(values=c("navy", "darkseagreen", "slategray2"))
```

```{r,echo=FALSE,warning=FALSE}
ggplot(data,aes(x=scar.BM,y=FR,colour=Phase))+facet_grid(.~Species)+geom_point(shape=18)+geom_smooth(method=lm, fullrange=FALSE)+ggtitle("Feeding rate by phase and scarid biomass") + theme_minimal() +labs(y="Feeding rate (bites/hr)",x="Scarid Biomass (g/100m2)") + theme(plot.title = element_text(hjust = 0.5)) + scale_color_manual(values=c("gray40", "aquamarine3"))
```

```{r, echo=FALSE,warning=FALSE, eval=FALSE}
ggplot(data,aes(x=scar.BM,y=BR,colour=Phase))+facet_grid(.~Species)+geom_point(shape=18)+geom_smooth(method=lm, fullrange=FALSE)+ggtitle("Bite rates (while grazing) by scarid biomass") + theme_minimal() +labs(y="Bite rate (bites/sec)",x="Scarid Biomass (g/100m2)") + theme(plot.title = element_text(hjust = 0.5)) + scale_color_manual(values=c("gray40", "aquamarine3"))
```

```{r,echo=FALSE,warning=FALSE}
ggplot(data,aes(x=scar.BM,y=g.frac,colour=Phase))+facet_grid(.~Species)+geom_point(shape=18)+geom_smooth(method=lm, fullrange=FALSE)+ggtitle("Fraction of time spent grazing by phase and scarid biomass") + theme_minimal() +labs(y="Time Fraction",x="Scarid Biomass (g/100m2)") + theme(plot.title = element_text(hjust = 0.5)) + scale_color_manual(values=c("gray40", "aquamarine3"))
```
```{r,echo=FALSE,warning=FALSE}
ggplot(data,aes(x=scar.BM,y=for.bites,colour=Phase))+facet_grid(.~Species)+geom_point(shape=18)+geom_smooth(method=lm, fullrange=FALSE)+ggtitle("Bites per foray by phase and scarid biomass") + theme_minimal() +labs(y="Bites",x="Scarid Biomass (g/100m2)") + theme(plot.title = element_text(hjust = 0.5)) + scale_color_manual(values=c("gray40", "aquamarine3"))
```

## Competitive interactions
```{r,echo=FALSE,warning=FALSE}
ggplot(site.sum,aes(x=scar.BM,y=int.scar.mean, shape=Island))+geom_point(size=3)+ggtitle("Interaction frequency as a function of scarid biomass") + theme_minimal() + theme(plot.title = element_text(hjust = 0.5)) + labs(x="Scarid Biomass (g/100m2)", y="Interactions with other scarids (per 2 min follow)") + geom_errorbar(aes(ymin=int.scar.mean-int.scar.se, ymax=int.scar.mean+int.scar.se))
```
```{r,echo=FALSE,warning=FALSE}
ggplot(site.sum,aes(x=scar.BM,y=int.scar.mean_stop, shape=Island))+geom_point(size=3)+ggtitle("S. viride interaction frequency as a function of scarid biomass") + theme_minimal() + theme(plot.title = element_text(hjust = 0.5)) + labs(x="Scarid Biomass (g/100m2)", y="Interactions with other scarids (per 2 min follow)") + geom_errorbar(aes(ymin=int.scar.mean_stop-int.scar.se_stop, ymax=int.scar.mean_stop+int.scar.se_stop))
```
```{r,echo=FALSE,warning=FALSE}
ggplot(site.sum,aes(x=scar.BM,y=int.scar.mean_qup, shape=Island))+geom_point(size=3)+ggtitle("S. vetula interaction frequency as a function of scarid biomass") + theme_minimal() + theme(plot.title = element_text(hjust = 0.5)) + labs(x="Scarid Biomass (g/100m2)", y="Interactions with other scarids (per 2 min follow)") + geom_errorbar(aes(ymin=int.scar.mean_qup-int.scar.se_qup, ymax=int.scar.mean_qup+int.scar.se_qup))
```
```{r,echo=FALSE,warning=FALSE}
ggplot(data,aes(x=scar.BM,y=int.scar))+facet_grid(.~Species)+geom_point(shape=19)+geom_smooth(method=lm)+ggtitle("Competitive interactions as a function of scarid biomass") + theme_minimal() + theme(plot.title = element_text(hjust = 0.5))
```

# PCAs
*Do I need to normalize my variables pre-PCA? Currently just scaled, except for grazing response variables which are log-transformed*

## Predictors

### All predictors
```{r}
predictors <- site.sum %>% select(cover_LC,cover_TA,canopy_MA,canopy_TA,rugosity,scar.BM,scar.DEN,carn.BM) # should it matter if I use site level summary data vs. full follow dataset (where site-level data has been joined?)
pca <- prcomp(na.omit(predictors),center = TRUE,scale. = TRUE) 
summary(pca)

plot(pca,type="l")
ggbiplot(pca, obs.scale = 1, var.scale = 1, groups=site.sum$Island, ellipse = TRUE, circle = TRUE, varname.size = 2) + scale_color_manual(name="Island", values=c("navy", "darkseagreen", "slategray2")) + theme(legend.direction = 'horizontal', legend.position = 'top') + theme_minimal()

```
### Benthic variables
```{r}
predictors1 <- site.sum %>% select(cover_LC,cover_TA,canopy_MA,canopy_TA,rugosity) # should it matter if I use site level summary data vs. full follow dataset (where site-level data has been joined?)
pca1 <- prcomp(na.omit(predictors1),center = TRUE,scale. = TRUE) 
summary(pca1)

plot(pca1,type="l")
ggbiplot(pca1, obs.scale = 1, var.scale = 1, groups=site.sum$Island, ellipse = TRUE, circle = TRUE, loadings.label.repel=TRUE) + scale_color_manual(name="Island", values=c("navy", "darkseagreen", "slategray2")) + theme(legend.direction = 'horizontal', legend.position = 'top') + theme_minimal()

# add PC1 and PC2 as variables in dataset
pca1.scores <- as.data.frame(pca1$x) %>% select(PC1,PC2)
site.sum <- site.sum  %>% bind_cols(pca1.scores)
pca1.scores <- select(site.sum,Site,PC1,PC2)
data <- data %>% left_join(pca1.scores,by="Site")
```
### Fish community variables
```{r}
predictors2 <- site.sum %>% select(scar.BM,scar.DEN,carn.BM) # should it matter if I use site level summary data vs. full follow dataset (where site-level data has been joined?)
pca2 <- prcomp(na.omit(predictors2),center = TRUE,scale. = TRUE) 
summary(pca2)

plot(pca2,type="l")
ggbiplot(pca2, obs.scale = 1, var.scale = 1, groups=site.sum$Island, ellipse = TRUE, circle = TRUE, loadings.label.repel=TRUE) + scale_color_manual(name="Island", values=c("navy", "darkseagreen", "slategray2")) + theme(legend.direction = 'horizontal', legend.position = 'top') + theme_minimal()

```

## Response variables

### Grazing behaviors
*log transform increases P1 explained variance here...*
*Currently for.dur has several NA values (for individuals that didn't graze)... I am omitting that variable for now because I don't want to omit non-grazing individuals (instead would like to keep them as grazing=0*
```{r}
response <- data %>% select(g.frac,TDB,BR)
log.vars5 <- log(response[,1:3])
pca3 <- prcomp(na.omit(response),center = TRUE,scale. = TRUE) 
summary(pca3)
plot(pca3,type="l")
ggbiplot(pca3, obs.scale = 1, var.scale = 1, groups=data$Island,shape=data$Island,ellipse = TRUE, circle = TRUE, alpha=0.8,varname.size = 5) + theme_minimal() + scale_color_manual(name="Island", values=c("navy", "darkseagreen", "slategray2")) + labs(title="Grazing behaviors by island") + theme(plot.title = element_text(hjust = 0.5))
```
```{r}
response <- data %>% select(g.frac,FR,BR)
log.vars5 <- log(response[,1:3])
pca3 <- prcomp(na.omit(response),center = TRUE,scale. = TRUE) 
summary(pca3)
plot(pca3,type="l")
ggbiplot(pca3, obs.scale = 1, var.scale = 1,ellipse = TRUE, circle = TRUE, alpha=0.4,varname.size = 5) + theme(legend.direction = 'horizontal', legend.position = 'top') + theme_minimal() +geom_point(aes(col=data$scar.BM)) + labs(col="Scarid Biomass (g/100m2)", title="Grazing Behaviors by Scarid Biomass") + theme(plot.title = element_text(hjust = 0.5))
```

```{r,echo=FALSE}
vir <- data %>% filter(Species.Code=="stop")
response <- vir %>% select(g.frac,FR,BR)
log.vars5 <- log(response[,1:3])
pcavir <- prcomp(na.omit(response),center = TRUE,scale. = TRUE) 
ggbiplot(pcavir, obs.scale = 1, var.scale = 1,ellipse = TRUE, circle = TRUE, alpha=1,varname.size = 5) + theme(legend.direction = 'horizontal', legend.position = 'top') + theme_minimal() + labs(col="Scarid Biomass (g/100m2)", title="Grazing Behaviors by Scarid Biomass") + theme(plot.title = element_text(hjust = 0.5)) +geom_point(aes(colour=vir$scar.BM))
```
```{r,echo=FALSE}
vet <- data %>% filter(Species.Code=="qup")
response <- vet %>% select(g.frac,FR,BR)
log.vars5 <- log(response[,1:3])
pcavet <- prcomp(na.omit(response),center = TRUE,scale. = TRUE) 
ggbiplot(pcavet, obs.scale = 1, var.scale = 1,ellipse = TRUE, circle = TRUE, alpha=1,varname.size = 5) + theme(legend.direction = 'horizontal', legend.position = 'top') + theme_minimal() + labs(col="Scarid Biomass (g/100m2)", title="Grazing Behaviors by Scarid Biomass") + theme(plot.title = element_text(hjust = 0.5)) +geom_point(aes(colour=vet$scar.BM))
```

# Model trials: all species compiled

## Basic lm to assess predictors of interest
-site-level predictors: scar.BM,scar.DEN,carn.BM,benthic (PC1,PC2)
-fish-level predictors: species, phase, length
--eventually run separately for different species
--species*scar.BM interaction?
-random effects: island
-response variables: g.frac, BR (?), and FR (run separately)

### Feeding rates
```{r}
initial<-data%>%filter(Phase=="i") # for analysis with initial only
lm_FR <- lm(FR~scar.BM*Species+carn.BM+Length.cm+PC1+PC2, data=data)
summary(lm_FR)
#vif(lm_FR)
```
*Notes: nearly everything is significant? Significant interaction between scar.BM and species for Spar. viride. carn.BM and scar.BM are highly positively correlated (~0.86) - is it ok to include both?*
*R2 goes up slightly with initial only*

### Fraction of time grazing
```{r}
lm_g.frac <- lm(g.frac~Species*scar.BM+scar.DEN+carn.BM+Phase+Length.cm+PC1+PC2, data=data)
summary(lm_g.frac)
```
*Notes: carn.BM and PC1 are strongest predictors of time spent grazing, scar.BM and phase are also significant. Slight significance of species, length, and species-scar.BM interaction*

## Mixed effects models
- random effect: island

### Feeding rates
```{r}
lme_FR <- lme(FR~Phase+Length.cm+Species*scar.BM+scar.DEN+carn.BM+PC1+PC2,data=data,random=~1|Island, na.action = na.omit)
summary(lme_FR)
```
*Notes: nearly everything is significant except for PC2 and scar.DEN?*

### Fraction of time spent grazing
```{r}
lme_g.frac <- lme(g.frac~Phase+Length.cm+Species*scar.BM+scar.DEN+carn.BM+PC1+PC2,data=data,random=~1|Island, na.action = na.omit)
summary(lme_g.frac)
```
*Notes: carn.BM and phase are strongest predictors, followed by scar.BM, PC1, and length. Also significant interaction between *

# Model trials: species run individually
```{r}
stop<-data%>%filter(Species.Code=="stop")
qup<-data%>%filter(Species.Code=="qup")
rbp<-data%>%filter(Species.Code=="rbp")
```

## Feeding rates
```{r}
lme_FR_stop <- lme(FR~Phase+Length.cm+scar.BM+scar.DEN+carn.BM+PC1+PC2,data=stop,random=~1|Island, na.action = na.omit)
summary(lme_FR_stop)
```
*Notes: carn.BM and length are signifigant predictors*

```{r}
lme_FR_qup <- lme(FR~Phase+Length.cm+scar.BM+scar.DEN+carn.BM+PC1+PC2,data=qup,random=~1|Island, na.action = na.omit)
summary(lme_FR_qup)
```
*Notes: PC1 is strongest (magnitude) predictors. Also significant are phase, carn.BM and scar.DEN*

```{r}
lme_FR_rbp <- lme(FR~Phase+Length.cm+scar.BM+scar.DEN+carn.BM+PC1+PC2,data=rbp,random=~1|Island, na.action = na.omit)
summary(lme_FR_rbp)
```
*Notes: much more variable, less predictable. Maybe take out of overall analysis?*
------------------------
To Do as of Nov. 7
* ~~incorporate PCs~~
* boosted regression trees ecosphere 2017 adrians paper
* ~~simpler graphs with just herb density x axis - grazing behaviors y, facet by species~~
* ~~Species*density as fixed effect~~
* Species as random effect
* ~~break out species and run separately~~

### 2017-12-08 Summary.Rmd
updating files

---
title: "Dec. 1 Committee Meeting"
author: "Molly Wilson"
date: "12/1/2017"
output:
  html_document:
    number_sections: yes
    theme: united
    toc: yes
    toc_float: yes
    toc_depth: 3
  pdf_document:
    toc: yes
---

```{r,message=FALSE,warning=FALSE, echo=FALSE}
# set up
setwd("~/Google Drive/Research + Projects/Dissertation/ABB_Grazing/Analysis")
library(rmarkdown)
library(dplyr)
library(tidyr)
library(lazyeval)
library(ggplot2)
library(ellipse)
library(RColorBrewer)
library(ggbiplot)
library(nlme)
library(pls)
```
```{r,message=FALSE,warning=FALSE, echo=FALSE}
site.sum <- read.csv("site.sum.csv", stringsAsFactors=FALSE) %>% select(-X)
data <- read.csv("sumID2.csv", stringsAsFactors=FALSE) %>% select(Island,Site,Spearfishing,Diving,Species,Species.Code,Phase,Length.cm,g.frac,FR,BR,TDB,for.dur,for.bites,tot.BM,tot.DEN,scar.BM,herb.BM,scar.DEN,conspec.BM,conspec.DEN,carn.BM,carn.DEN,cover_CCA,cover_LC,cover_MA,cover_TA,canopy_MA,canopy_TA,rugosity,int.tot,int.con,int.hetscar,int.dam)
vetvir <- data %>% filter(Species.Code!="rbp")
vetvir20 <- vetvir %>% filter(Length.cm>=15 & Length.cm<=25)
init20<- vetvir %>% filter(Length.cm>=15 & Length.cm<=25 & Phase=="i")
```

# Meeting objectives
* Go through where I'm at in terms of chapters: progress and planning 
    + Perspective paper on behavioral effects of human activities
    + Fieldwork investigating variations in Scarid grazing across systems with different levels of human impact
    + Linking behavior to potential ecological effects - likely through herbivory model
    + Additional chapter: social-ecological impacts of parrotfish fishing ban in DR?
* Where I need to get myself in terms of oral exams
* Scheduling next meetings, sign Annual Review form

# Scarid grazing data + analysis

## Intro to the data

**Hypothesis**: herbivorous fish (scarids) have reduced grazing impact in low herbivore biomass (high herb. fishing) systems

* scarids feeding rates increase with scarid biomass  
* foray size (bites/foray) and duration increase with scarid biomass

* 795 follows (2-min)
* 13 sites across 3 islands (Antigua, Barbuda, Bonaire)
* 2 focal species in Bonaire (Scarus vetula, Sparisoma viride - dominant grazers), but added Sparisoma aurofrenatum in Antigua & Barbuda because of low densities of viride and vetula in some sites
* Possibility to integrate fish follow data from FL Keys and St. Croix (Burkepile), and Belize and Bahamas (Mumby)

Site-level predictor variables:

* fish community data: scarid biomass, scarid density, carnivore biomass
* benthic community data: percent coral cover, percent turf algal cover, percent macroalgal cover, turf canopy height, macroalgal canopy height
* rugosity

Indv. fish-level predictor variables:

* species
* phase
* length

Indv. fish-level response variables:

* fraction of time spent grazing (**g.frac**)
* bite rate while grazing (bites/sec) (**BR**)
* feeding rate (overall bite rate, bites/hr) (**FR**)
* total daily bites (**TDB**)
* foray size (bites per foray) (**for.bites**)
* foray duration (seconds) (**for.dur**)

Not done with analyses, but will introduce you to where I'm at....

## Exploring predictors of grazing behavior

### Grazing as a function of scarid biomass
```{r,echo=FALSE,warning=FALSE, echo=FALSE}
ggplot(vetvir,aes(x=scar.BM,y=FR,colour=Island))+facet_grid(.~Species)+geom_point(shape=18)+geom_smooth(method=lm, fullrange=FALSE)+ggtitle("Feeding rate by scarid biomass") + theme_minimal() +labs(y="Feeding rate (bites/hr)",x="Scarid Biomass (g/100m2)") + theme(plot.title = element_text(hjust = 0.5)) + scale_color_manual(values=c("navy", "darkseagreen", "slategray2"))
```
```{r,echo=FALSE,warning=FALSE, echo=FALSE}
ggplot(vetvir,aes(x=scar.BM,y=FR,colour=Phase))+facet_grid(.~Species)+geom_point(shape=18)+geom_smooth(method=lm, fullrange=FALSE)+ggtitle("Feeding rate by phase and scarid biomass") + theme_minimal() +labs(y="Feeding rate (bites/hr)",x="Scarid Biomass (g/100m2)") + theme(plot.title = element_text(hjust = 0.5)) + scale_color_manual(values=c("gray40", "aquamarine3"))
```

```{r, echo=FALSE,warning=FALSE, echo=FALSE}
ggplot(vetvir,aes(x=scar.BM,y=BR,colour=Phase))+facet_grid(.~Species)+geom_point(shape=18)+geom_smooth(method=lm, fullrange=FALSE)+ggtitle("Bite rates (while grazing) by scarid biomass") + theme_minimal() +labs(y="Bite rate (bites/sec)",x="Scarid Biomass (g/100m2)") + theme(plot.title = element_text(hjust = 0.5)) + scale_color_manual(values=c("gray40", "aquamarine3"))
```

```{r,echo=FALSE,warning=FALSE, echo=FALSE}
ggplot(vetvir,aes(x=scar.BM,y=g.frac,colour=Phase))+facet_grid(.~Species)+geom_point(shape=18)+geom_smooth(method=lm, fullrange=FALSE)+ggtitle("Fraction of time spent grazing by phase and scarid biomass") + theme_minimal() +labs(y="Time Fraction",x="Scarid Biomass (g/100m2)") + theme(plot.title = element_text(hjust = 0.5)) + scale_color_manual(values=c("gray40", "aquamarine3"))
```
```{r,echo=FALSE,warning=FALSE, echo=FALSE}
ggplot(vetvir,aes(x=scar.BM,y=for.bites,colour=Phase))+facet_grid(.~Species)+geom_point(shape=18)+geom_smooth(method=lm, fullrange=FALSE)+ggtitle("Bites per foray by phase and scarid biomass") + theme_minimal() +labs(y="Bites",x="Scarid Biomass (g/100m2)") + theme(plot.title = element_text(hjust = 0.5)) + scale_color_manual(values=c("gray40", "aquamarine3"))
```

### Grazing PCAs

```{r, echo=FALSE,eval=FALSE}
response <- vetvir %>% select(g.frac,FR,BR)
log.vars5 <- log(response[,1:3])
pca3 <- prcomp(na.omit(response),center = TRUE,scale. = TRUE) 
summary(pca3)
plot(pca3,type="l")
ggbiplot(pca3, obs.scale = 1, var.scale = 1,ellipse = TRUE, circle = TRUE, alpha=0.4,varname.size = 5) + theme(legend.direction = 'horizontal', legend.position = 'top') + theme_minimal() +geom_point(aes(col=vetvir$scar.BM)) + labs(col="Scarid Biomass (g/100m2)", title="Grazing Behaviors by Scarid Biomass") + theme(plot.title = element_text(hjust = 0.5))
```

```{r,echo=FALSE}
response <- vetvir %>% select(g.frac,FR,BR)
log.vars5 <- log(response[,1:3])
pca3 <- prcomp(na.omit(response),center = TRUE,scale. = TRUE) 
ggbiplot(pca3, obs.scale = 1, var.scale = 1, groups=vetvir$Island,ellipse = TRUE, circle = TRUE, alpha=0.8,varname.size = 5) + theme_minimal() + scale_color_manual(name="Island", values=c("navy", "darkseagreen", "slategray2")) + labs(title="Grazing behaviors by island") + theme(plot.title = element_text(hjust = 0.5))
```

### Grazing as a function of fish length
```{r, echo=FALSE,warning=FALSE}
ggplot(vetvir,aes(x=Length.cm,y=FR,colour=Island))+facet_grid(.~Species)+geom_point(shape=18)+scale_shape(solid=FALSE)+geom_smooth(method=lm, fullrange=FALSE)+ggtitle("Feeding rate by fish length") + theme_minimal() +labs(y="Feeding Rate (bites/hr)",x="Length (cm)") + theme(plot.title = element_text(hjust = 0.5))  + scale_color_manual(values=c("navy", "darkseagreen", "slategray2"))
```

```{r, echo=FALSE,warning=FALSE}
ggplot(vetvir,aes(x=Length.cm,y=BR,colour=Island))+facet_grid(.~Species)+geom_point(shape=18)+scale_shape(solid=FALSE)+geom_smooth(method=lm, fullrange=FALSE)+ggtitle("Bite rate (while grazing) by fish length") + theme_minimal() +labs(y="Bite Rate (bites/sec)",x="Length (cm)") + theme(plot.title = element_text(hjust = 0.5))  + scale_color_manual(values=c("navy", "darkseagreen", "slategray2"))
```

```{r, echo=FALSE,warning=FALSE}
ggplot(vetvir,aes(x=Length.cm,y=g.frac,colour=Island))+facet_grid(.~Species)+geom_point(shape=18)+scale_shape(solid=FALSE)+geom_smooth(method=lm, fullrange=FALSE)+ggtitle("Fraction of time spent grazing by fish length") + theme_minimal() +labs(y="Time Fraction",x="Length (cm)") + theme(plot.title = element_text(hjust = 0.5))  + scale_color_manual(values=c("navy", "darkseagreen", "slategray2"))
```

## Correlations among predictor variables
Strong correlations between scarid biomass and both carnivore biomass and coral cover. I can address benthic corelations via PCA but don't really want to pool the fish predictors I'm interested in?
```{r,echo=FALSE}
panel.cor <- function(x, y, digits=2, prefix="", cex.cor, ...)
{
    usr <- par("usr"); on.exit(par(usr))
    par(usr = c(0, 1, 0, 1))
    r <- abs(cor(x, y))
    txt <- format(c(r, 0.123456789), digits=digits)[1]
    txt <- paste(prefix, txt, sep="")
    if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
    text(0.5, 0.5, txt, cex = cex.cor * r)
}
pairs(~scar.BM+scar.DEN+carn.BM+cover_LC+cover_TA+canopy_MA+canopy_TA+rugosity, data=data,
      lower.panel=panel.smooth, upper.panel=panel.cor, 
      pch=20, main="Selected Fish and Benthic Variables",na.action = na.omit)
```

```{r, echo=FALSE}
predictors1 <- site.sum %>% select(cover_LC,cover_TA,canopy_MA,canopy_TA,rugosity) # should it matter if I use site level summary data vs. full follow dataset (where site-level data has been joined?)
pca1 <- prcomp(na.omit(predictors1),center = TRUE,scale. = TRUE) 
ggbiplot(pca1, obs.scale = 1, var.scale = 1, groups=site.sum$Island, ellipse = TRUE, circle = TRUE, loadings.label.repel=TRUE) + scale_color_manual(name="Island", values=c("navy", "darkseagreen", "slategray2")) + theme(legend.direction = 'horizontal', legend.position = 'top') + theme_minimal() + labs(title="PCA of selected benthic variables")

# add PC1 and PC2 as variables in dataset
pca1.scores <- as.data.frame(pca1$x) %>% select(PC1,PC2)
site.sum <- site.sum  %>% bind_cols(pca1.scores)
pca1.scores <- select(site.sum,Site,PC1,PC2)
data <- data %>% left_join(pca1.scores,by="Site")
```

## Mixed effect models 
Investigating drivers of grazing behavior

### Feeding rates
```{r}
vetvir <- data %>% filter(Species.Code!="rbp")
lme_FR <- lme(FR~Phase+Length.cm+Species*scar.BM+scar.DEN+carn.BM+PC1+PC2,data=vetvir,random=~1|Island, na.action = na.omit)
summary(lme_FR)
```
```{r}
vetvir <- data %>% filter(Species.Code!="rbp")
lme_for.bites <- lme(for.bites~Phase+Length.cm+Species*scar.BM+scar.DEN+carn.BM+PC1+PC2,data=vetvir,random=~1|Island, na.action = na.omit)
summary(lme_for.bites)
```

# Grazing model?

Potential to build off of Mumby & Co. models - primary models I have found that explicitly incorporate scarid feeding rates  
* Mumby et al. (2006): *Fishing, trophic cascades, and the process of grazing on coral reefs*  
* Mumby (2006): *The impact of exploiting grazers (Scaridae) on the dynamics of Caribbean coral reefs*

Essentially use Bruggemann (Bonaire) bite rates from viride and vetula, combined with data on bite size, to calculate total grazing intensity (cm2/hr) and then translates this into percent of reef maintained in grazed state

Planning to work with Holly Moeller/take her ecological modelling class in the spring

# Perspective piece

Introduce outline/framing  
* [human-induced behavioral effect pathways](https://docs.google.com/presentation/d/1l2plflfU6jZJWyrq4kHIunyacIrtJHNpgrbC4t6Lx3A/edit?usp=sharing)  
* [outline](https://docs.google.com/document/d/1tdIV9Wu8Y-o1DvI9YleuZRdH5HVSU1PIX0gQYTnB6lA/edit?usp=sharing)

# Chap. 3 (or 4?): social-ecological impacts of DR parrotfish ban?
Recently implemented (summer 2017) and reportedly (somewhat) enforced 3 year parrotfish ban. We have catch data, ecological data, and fishermen interview data collected over the past 5 years.

**Hypotheses**:  
* shift in catch composition to other species/sizes  
* shift to other gear types  
* shifts to fishing abroad, or other livelihoods altogether  
* economic impacts to fishermen (calculated through changes in catch, potential comparison of fish consumed at home)

# Side projects  
* DR social survey design
* Bonaire long-term monitoring data (recent paper submission)
* KUNI Mariculture (manuscript in prep)
* KUNI TURF project
* potential Fish Forever data analysis with SFG?
* __conferences for 2018?__

### 2017-12-08 ABB_Analysis Tasks.Rmd

--- 2017-12-08 Graphing+Analysis.Rmd
@@
 ---
+Files: sumID2.csv, site.sum.csv
 
@@
 # set up
+setwd("~/Google Drive/Research + Projects/Dissertation/ABB_Grazing/Analysis")
 library(rmarkdown)
@@
 library(factoextra)
-```
-```{r,message=FALSE,warning=FALSE, echo=FALSE}
-setwd("~/github/Scarid-Behavior/data")
 site.sum <- read.csv("site.sum.csv", stringsAsFactors=FALSE) %>% select(-X)

### 2017-12-08 ABB_Benthic.Rmd

@@
 ```{r, echo=FALSE,message=FALSE,warning=FALSE}
-setwd("~/github/Scarid-Behavior/")
-benthic<-read.csv("benthic.csv", stringsAsFactors=FALSE)
+setwd("~/github/Scarid-Behavior/data")
+benthic<-read.csv("benthic_anu.csv", stringsAsFactors=FALSE)
 rugosity<-read.csv("rugosity.csv", stringsAsFactors=FALSE)

### 2017-12-08 ABB_Follows.Rmd

--- 2017-12-08 Follows.Rmd
@@
 ---
+# Antigua + Barbuda Fish Data
+*required files: follows.csv*
+
+## Setup
+
 ```{r,message=FALSE,warning=FALSE}
 # set up
+setwd("~/Google Drive/Research + Projects/Dissertation/ABB_Grazing/Analysis")
 library(lazyeval)
@@
 ```{r, warning=FALSE, message= FALSE}
-# import data
-setwd("~/github/Scarid-Behavior/data")
 follows <- read.csv("follows.csv", stringsAsFactors=FALSE) 
@@
 n.sspph <- group_by(sumID, Site, Species, Phase) %>% dplyr::summarize(n=n())
+
 ```
-
-# Grazing behavior
 
@@
 
-## Grazing rates
+## Bite rates (bites/sec) overall and while grazing
 ```{r, warning=FALSE, message= FALSE}
@@
 
-# Competitive interactions
+## Competitive interactions
 
-## Summary of interactions per fish
+### Summary of interactions per fish
 ```{r}
@@
 
-## Matrix of interactions by species
+### Matrix of interactions by species
 *eventually breakdown by site? region?*
@@
 
-# Summarizing data by site, species, and phase
+## Summarizing data by site, species, and phase
 Next I create a summary tables that averages variables within each site/species/phase, then combines into one table
@@
 
-## Export csvs for integration with site data
+## Export csvs for integration with follow data
 ```{r,message=FALSE,warning=FALSE}
-# write.csv(sumID,file="sumID.csv")
-# write.csv(fol.sp.site,file="fol.sp.site.csv")
-# write.csv(fol.sp.ph.site,file="fol.sp.ph.site.csv")
+write.csv(site.follows,file="site.follows.csv")
+write.csv(sumID,file="sumID.csv")
+write.csv(fol.sp.site,file="fol.sp.site.csv")
+write.csv(fol.sp.ph.site,file="fol.sp.ph.site.csv")
 ```
-
-### 1.2 Import and format files
-```{r, echo=FALSE,message=FALSE,warning=FALSE}
-setwd("~/github/Scarid-Behavior/data")
-site.fish<-read.csv("site.fish.csv", stringsAsFactors=FALSE)%>% select(-X)
-site.benthic<-read.csv("site.benthic.csv", stringsAsFactors=FALSE)%>% select(-X)
-site.human<-read.csv("site.human.csv", stringsAsFactors=FALSE)
-species.site<-read.csv("species.site.csv",stringsAsFactors = FALSE)%>% filter(Species=="Sparisoma viride" | Species=="Scarus vetula" | Species=="Sparisoma aurofrenatum") %>% select(-X,-Family,-n,-L.n)
-names(species.site)[3:8] <- c("conspec.BM","conspec.BM.se","conspec.DEN","conspec.DEN.se","conspec.L","conspec.L.se")
-```
-### 1.3 Join datasets, print .csv
-```{r, echo=FALSE,message=FALSE,warning=FALSE}
-# integrate all site-level variables
-site.sum <- site.follows %>% left_join(site.fish,by="Site") %>% left_join(site.benthic,by="Site") %>% left_join(site.human,by="Site") 
-# then join site-level data to sumID dataframe
-sumID2 <- sumID %>% left_join(site.fish,by="Site") %>% left_join(site.benthic,by="Site") %>% left_join(site.human,by="Site")
-# calculate conspecifc scarid variable to add to sumID2 (based on fish site/species)
-sumID2 <- sumID2 %>% left_join(species.site,by=c("Site","Species"))
-
-write.csv(site.sum,file="site.sum.csv")
-write.csv(sumID2,file="sumID2.csv")
-```

### 2018-01-07 DonorReport.Rmd
1/6

---
title: "Effects of Fishing on Scarid Grazing Behavior in the Caribbean"
author: "Molly Wilson"
date: "1/5/2018"
output:
  pdf_document:
    toc: yes
  html_document:
    number_sections: yes
    theme: united
    toc: yes
    toc_float: yes
    toc_depth: 3
---

```{r,message=FALSE,warning=FALSE, echo=FALSE}
# set up
library(rmarkdown)
library(dplyr)
library(tidyr)
library(lazyeval)
library(ggplot2)
library(ellipse)
library(RColorBrewer)
library(ggbiplot)
library(nlme)
library(pls)
library(mgcv)
library(ggsignif)
```
```{r,message=FALSE,warning=FALSE, echo=FALSE}
setwd("~/github/Scarid-Behavior/data")
site.sum <- read.csv("site.sum.csv", stringsAsFactors=FALSE) %>% select(-X)
data <- read.csv("sumID2.csv", stringsAsFactors=FALSE) %>% select(Island,Site,Spearfishing,Diving,Species,Species.Code,Phase,Length.cm,g.frac,FR,BR,TDB,for.dur,for.bites,tot.BM,tot.DEN,scar.BM,herb.BM,scar.DEN,conspec.BM,conspec.DEN,carn.BM,carn.DEN,cover_CCA,cover_LC,cover_MA,cover_TA,canopy_MA,canopy_TA,rugosity,int.tot,int.con,int.hetscar,int.dam)
vetvir <- data %>% filter(Species.Code!="rbp")
vetvir20 <- vetvir %>% filter(Length.cm>=15 & Length.cm<=25)
init20<- vetvir %>% filter(Length.cm>=15 & Length.cm<=25 & Phase=="i")
vetvirAB <- vetvir %>% filter(Island!="Barbuda")
vetvirAB20 <- vetvir %>% filter(Island!="Barbuda")
vetvirAB20NS <- vetvirAB20 %>% filter(Site!="Snapper Point")
vetvir20NS <- vetvir20 %>% filter(Site!="Snapper Point")
vetvirNS <- vetvir %>% filter(Site!="Snapper Point")
```

# Introduction

Herbivory is a critical ecosystem process on coral reefs. Algae competes with corals for space and light, but herbivores suppress algae and allow corals to grow, reproduce, and recruit new larvae. Up until a mass mortality event in the 1980s, the sea urchin *Diadema antillarum* was a primary herbivore on Caribbean reefs. Since their population crash, most herbivory has been carried out by grazing fish, primarily parrotfish or *Scaridae* species. As these fish are increasingly threatened by overfishing, their grazing impact has been reduced. While numerous studies have documented declines in parrotfish populations and the implications for coral-algal competition, few if any have investigated potential differences in parrotfish grazing behavior between degraded and pristine reef systems. Additionally, most existing data on parrotfish grazing behavior comes from relatively pristine reefs, which might skew our estimates of herbivory when applied to the many degraded systems throughout the Caribbean.

While working on a reef survey project in Bonaire - an island in the Dutch Antilles considered to have some of the most pristine coral reefs in the Caribbean - I began to notice feeding differences between the parrotfish there and those I had been studying in Antigua, a more degraded reef system in the Eastern Caribbean. Numerous differences between degraded and pristine reef systems like Antigua and Bonaire could lead to variations in parrotfish grazing behavior. Shared vigilance theory would predict that parrotfish would have increased feeding in high abundance populations, where they have "safety in numbers" and can share the responsibility of looking out for predators. Conversely, competition between parrotfish might lead individual fish to spend less time feeding in high abundance populations because they have to spend more time defending feeding territories. The availability of algae might also influence feeding behavior, with parrotfish having to spend less time feeding if algae is abundant or of higher nutritional quality.

My research aims to document variations in grazing behavior across different reef conditions, testing for the effects of parrotfish abundance, predator abundance, and benthic condition on the amount of time parrotfish spend feeding, their bite rates, and the duration of their feeding bouts. Bonaire represents a relatively pristine Caribbean system, while Antigua has shifted to a more algal-dominated, low scarid state. I also conducted surveys in Barbuda, where a recent efforts to reduce parrotfish fishing may have interesting implications on grazing behavior.

# Field season summary

In August of 2017, I set out to document parrotfish feeding behaviors in Antigua and Barbuda, two neighboring islands with historical overfishing and relatively degraded reefs systems. While parrotfish fishing continues in Antigua, a parrotfish fishing ban has recently been instated in Barbuda. Both of these islands would serve as comparisons to the pristine system in Bonaire that are more represenative of reef status in the larger Caribbean region.

Thanks to generous funding awarded through the UCSB Research Accellerator Award, I was able to travel to Antigua along with a research assistant also working on her Ph.D. in marine ecoogy and fisheries management. The funds supplied us with SCUBA tanks, survey gear, boat fuel, a car rental to transport gear, and fuel and food to supply a four-day boat trip to Barbuda. During the three weeks we were in the field, we collected benthic and fish survey data at a total of nine sites which, in addition to providing data for this investigation, will also serve as valuable assessment and baseline data for future studies. Across the nine sites we coolected behavioral data from a total of 434 individual parrotfish, making a database of nearly 800 fish follows when integrated with the data I collected in Bonaire. While the data analysis is still ongoing, the preliminary results are extremely interesting. This research would not have been possible without the generous support of my Research Accellerator sponsor.

## Data overview

**Hypotheses**: parrotfish (scarids) have reduced grazing levels in heavily fished (low herbivore biomass) systems  
* Scarid feeding rates increase with scarid biomass
* Foray size (bites/foray) and duration increase with scarid biomass

Research components

* 795 follows (2-min)
* 13 sites across 3 islands (Antigua, Barbuda, Bonaire)
    * Antigua: Rendezvous Bay, Turtle Bay, Windward Bay, Snappers Hole, Pillars of Hercules, Mermaid Gardens
    * Barbuda: Pallaster East, Central, and West
    * Bonaire: 18th Palm, Oil Slick, Karpata, Front Porch
* 2 focal species in Bonaire (Scarus vetula, Sparisoma viride - dominant grazers), but added Sparisoma aurofrenatum in Antigua & Barbuda because of low densities of viride and vetula in some sites
* Possibility to integrate fish follow data from FL Keys and St. Croix (Burkepile), and Belize and Bahamas (Mumby)

Site-level predictor variables:

* Fish community data: scarid biomass, scarid density, carnivore biomass
* Benthic community data: percent coral cover, percent turf algal cover, percent macroalgal cover, turf canopy height, macroalgal canopy height
* Rugosity

Indv. fish-level predictor variables:

* Species
* Phase
* Length

Indv. fish-level response variables:

* Fraction of time spent grazing (**g.frac**)
* Bite rate while grazing (bites/sec) (**BR**)
* Overall feeding rate (overall bite rate, bites/hr) (**FR**)
* Total daily bites (**TDB**)
* Foray size (bites per foray) (**for.bites**)
* Foray duration (seconds) (**for.dur**)

# Initial results

## General grazing behavior

```{r, message=FALSE,warning=FALSE,echo=FALSE}
ggplot(vetvir20, aes(x=Island, y=FR)) + geom_boxplot() + facet_grid(.~Species) +
stat_summary(fun.y = mean, geom="point", shape=5, size=3) + labs(y="Feeding Rate (bites/hr)",x="",title="Feeding Rate by Island and Species") + theme_bw() + theme(plot.title = element_text(hjust = 0.5))
```
```{r, message=FALSE,warning=FALSE,echo=FALSE}
# do I really satisfy ANOVA assumptions?
vet<- vetvir %>% filter(Species.Code=="qup")
vir<- vetvir %>% filter(Species.Code=="stop")
aovFRvet <- aov(FR ~ Island, data=vet) 
#TukeyHSD(aovFRvet)
aovFRvir <- aov(FR ~ Island, data=vir) 
#TukeyHSD(aovFRvir)
```

Boxplots represent median (bar) and mean (diamond) feeding rates for *Scarus vetula* (Queen Parrotfish) and *Sparisoma viride* (Stoplight Parrotfish) in Antigua, Barbuda, and Bonaire. Average feeding rates (bites/hr) of *Scarus vetula*  in Bonaire were nearly three times that of Antigua and Barbuda (p<0.001). For *Sparisoma viride*, feeding rates in Bonaire and Barbuda were significantly higher than that of Bonaire (p<0.001).

```{r, message=FALSE,warning=FALSE,echo=FALSE}
ggplot(vetvir20, aes(x=Island, y=for.bites)) + geom_boxplot()  + labs(y="Bites per foray",x="",title="Grazing intensity by Island and Species") + theme_bw() + theme(plot.title = element_text(hjust = 0.5)) + geom_signif(comparisons = list(c("Antigua", "Barbuda"),c("Antigua","Bonaire"),c("Barbuda","Bonaire")), map_signif_level=TRUE) + facet_grid(.~Species)
```
```{r, message=FALSE,warning=FALSE,echo=FALSE}
# do I really satisfy ANOVA assumptions?
aovforvet <- aov(for.dur ~ Island, data=vet) 
#TukeyHSD(aovforvet)
aovforvir <- aov(for.dur ~ Island, data=vir) 
#TukeyHSD(aovforvir)
```
Grazing intensity, or the number of consecutive bites taken during a feeding foray, has significant implications for the benthic substrate. When parrotfish graze more intensely, they are more likely to crop algae and clear space for recruiting or competing corals. When shorter grazing forays are taken, bites are less concentrated and algae may not be sufficiently reduced in any given area. This plot shows the significant difference in grazing intensity of *Scarus vetula* individuals in Bonaire and of those in Antigua and Barbuda (p<0.001). For *Sparisoma viride*, grazing intensity in both Bonaire and Barbuda were significantly greater than in Antigua (p<0.001 and p<0.05, respectively).

## Grazing impact as a function of fish size

*Due to the lack of larger fish in Barbuda, these length-based relationships focus on comparisons between Antigua and Bonaire*

```{r, message=FALSE,warning=FALSE,echo=FALSE}
ggplot(vetvirAB,aes(x=Length.cm,y=FR,colour=Island))+facet_grid(.~Species)+geom_point(shape=18)+scale_shape(solid=FALSE)+geom_smooth(method=lm, fullrange=FALSE)+ggtitle("Feeding rates by fish length") + theme_minimal() +labs(y="Feeding rates (bites/hr)",x="Length (cm)") + theme(plot.title = element_text(hjust = 0.5))  + scale_color_manual(values=c("navy", "slategray2"))
```
Overall feeding rates decrease with increasing fish length in both Antigua and Bonaire, as has been documented in previous studies. Larger individuals have larger bites, and therefore need to take fewer bites to obtain sufficient food. However, individuals of a given size are taking fewer bites in Antigua than in Bonaire. This is true for both *Scarus vetula* and *Sparisoma viride* individuals.

```{r, message=FALSE,warning=FALSE,echo=FALSE}
ggplot(vetvirAB,aes(x=Length.cm,y=BR,colour=Island))+facet_grid(.~Species)+geom_point(shape=18)+scale_shape(solid=FALSE)+geom_smooth(method=lm, fullrange=FALSE)+ggtitle("Bite rate by fish length") + theme_minimal() +labs(y="Bite Rate (bites/sec)",x="Length (cm)") + theme(plot.title = element_text(hjust = 0.5))  + scale_color_manual(values=c("navy", "slategray2"))
```
As shown here, the differences in instananeous bite rates (bite rates while feeding) do not differ significantly between Antigua and Bonaire. For *Scarus vetula*, bite rates decrease with increasing fish size, while for *Sparisoma viride* they remain relatively constant across fish size.

```{r, message=FALSE,warning=FALSE,echo=FALSE}
ggplot(vetvirAB,aes(x=Length.cm,y=g.frac,colour=Island))+facet_grid(.~Species)+geom_point(shape=18)+scale_shape(solid=FALSE)+geom_smooth(method=lm, fullrange=FALSE)+ggtitle("Fraction of time spent grazing by fish length") + theme_minimal() +labs(y="Fraction of time",x="Length (cm)") + theme(plot.title = element_text(hjust = 0.5))  + scale_color_manual(values=c("navy", "slategray2"))
```
While instantaneous bite rates may be fairly consistent across regions, the fraction of time an individual fish spends grazing is significantly higher in Bonaire than in Antigua. This accounts for the differences in overal feeding rate seen previously.

## Grazing impact as a function of scarid biomass
```{r, message=FALSE,warning=FALSE,echo=FALSE}
ggplot(vetvir20NS,aes(x=scar.BM,y=FR))+facet_grid(.~Species)+geom_point(shape=18)+geom_smooth(method=loess, fullrange=FALSE)+ggtitle("Feeding rate by scarid biomass") + theme_minimal() +labs(y="Feeding rate (bites/hr)",x="Scarid Biomass (g/100m2)") + theme(plot.title = element_text(hjust = 0.5))
```

Overall feeding initially increases with scarid biomass and then eventually decreases as scarid biomass becomes highest. This trend is most pronounced with *Scarus vetula*, but can also be seen in the *Sparisoma viride* data. The initial positive relationship  may include social feeding/shared vigilance theory, in which feeding increases with biomass because individuals can share vigilance for predators and spend more time feeding, and/or an individual's feeding is triggered by the feeding of others. The eventual negative relationship between feeding and scarid biomass is likely a result of competition amongst scarids, which may impede feeding at high biomass levels.

```{r, message=FALSE,warning=FALSE,echo=FALSE}
ggplot(vetvir20NS,aes(x=scar.BM,y=g.frac))+facet_grid(.~Species)+geom_point(shape=18)+geom_smooth(method=loess, fullrange=FALSE)+ggtitle("Fraction of time spent grazing by scarid biomass") + theme_minimal() +labs(y="Time fraction",x="Scarid Biomass (g/100m2)") + theme(plot.title = element_text(hjust = 0.5))
```

The fraction of time spent grazing depicts this trend even more clearly. Individuals initially increase the amount of time they spend grazing with increasing scarid populations, but then decrease at high biomass levels, likely because they have to allocate more time to competitive activities or because it is more challenging for them to find ideal feeding substrates.

```{r, message=FALSE,warning=FALSE,echo=FALSE}
ggplot(vetvir20NS,aes(x=scar.BM,y=BR))+facet_grid(.~Species)+geom_point(shape=18)+geom_smooth(method=loess, fullrange=FALSE)+ggtitle("Bite rates by scarid biomass") + theme_minimal() +labs(y="Bite rate (bites/sec)",x="Scarid Biomass (g/100m2)") + theme(plot.title = element_text(hjust = 0.5))
```

Instantaneous bite rates of *Sparisoma viride* show no relationship with scarid biomass. *Scarus vetula* individuals initially increase bite rates at low scarid biomass levels, but then decrease/remain stable at higher biomasses. Individuals in low scarid biomass settings may feed more slowly if they feel the need to maintain vigilance while feeding.

# Implications and next steps

A key finding of this research is that scarid grazing behaviors can vary greatly between pristine and degraded reef ecosystems, which may have significant implications for their grazing impact and correspondingly on coral reef benthic health. In the case of *Scarus vetula*, individuals in Bonaire are taking an average of twice as many bites than in Antigua. These bites are also more concentrated, which may enhance benefits of grazing on corals. *Sparisoma viride* by nature have lower feeding rates than *Scarus vetula*, but still show elevated feeding rates in Bonaire and Barbuda as opposed to in Antigua.  Because Bonairian, Antiguan, and Barbudan reefs vary across numerous dimensions (e.g. scarid populations, carnivore populations, coral and algal communities, reef structure, fisher and recreational diver presence), it is outside the scope of this initial study to attribute changes in grazing behavior to any one particular factor or mechanism. However, the intial model reports the significant role played by scarid biomass and benthic condition. Further studies are required to determine the exact drivers and mechanisms. Additional research is also required to understand the relationship between feeding behaviors and benthic responses. While this study documents differences in grazing rates and temporal concentration, it does not measure potential differences in bite size, substrate consumed, or spatial distribution of bites. These are all components I hope to explore further during continued fieldwork in Antigua, Barbuda, and Bonaire.

Furthermore, this study provides new grazing data that may help improve coral reef grazing models. While current models base feeding rate estimates on scarids surveyed in Bonaire, these estimates may not be very accurate in low biomass/high fishing systems like Antigua.

### 2018-01-07 Graphing+Analysis.Rmd
1/6

@@
 ```{r}
-ggplot(data,aes(x=scar.BM,y=FR))+facet_grid(.~Species)+geom_smooth(method="lm")+geom_point(shape=18)+ggtitle("Feeding rate by scarid biomass") + theme_minimal() +labs(y="Feeding rate (bites/hr)",x="Scarid Biomass (g/100m2)") + theme(plot.title = element_text(hjust = 0.5))
+anubon <- data %>% filter(Island!="Barbuda")
+ggplot(anubon,aes(x=scar.BM,y=FR))+facet_grid(.~Species)+geom_smooth(method="loess")+geom_point(shape=18)+ggtitle("Feeding rate by scarid biomass") + theme_minimal() +labs(y="Feeding rate (bites/hr)",x="Scarid Biomass (g/100m2)") + theme(plot.title = element_text(hjust = 0.5))
 ```
@@
 ```{r,echo=FALSE,warning=FALSE}
-ggplot(data,aes(x=scar.BM,y=FR,colour=Phase))+facet_grid(.~Species)+geom_point(shape=18)+geom_smooth(method=lm, fullrange=FALSE)+ggtitle("Feeding rate by phase and scarid biomass") + theme_minimal() +labs(y="Feeding rate (bites/hr)",x="Scarid Biomass (g/100m2)") + theme(plot.title = element_text(hjust = 0.5)) + scale_color_manual(values=c("gray40", "aquamarine3"))
+ggplot(data,aes(x=scar.BM,y=FR,colour=Phase))+facet_grid(.~Species)+geom_point(shape=18)+geom_smooth(method=loess, fullrange=FALSE)+ggtitle("Feeding rate by phase and scarid biomass") + theme_minimal() +labs(y="Feeding rate (bites/hr)",x="Scarid Biomass (g/100m2)") + theme(plot.title = element_text(hjust = 0.5)) + scale_color_manual(values=c("gray40", "aquamarine3"))
 ```
@@
 ggbiplot(pca, obs.scale = 1, var.scale = 1, groups=site.sum$Island, ellipse = TRUE, circle = TRUE, varname.size = 2) + scale_color_manual(name="Island", values=c("navy", "darkseagreen", "slategray2")) + theme(legend.direction = 'horizontal', legend.position = 'top') + theme_minimal()
-
 ```
 ### Benthic variables
-```{r}
-predictors1 <- site.sum %>% select(cover_LC,cover_TA,canopy_MA,canopy_TA,rugosity) # should it matter if I use site level summary data vs. full follow dataset (where site-level data has been joined?)
-pca1 <- prcomp(na.omit(predictors1),center = TRUE,scale. = TRUE) 
-summary(pca1)
+```{r, echo=FALSE}
+pred.benthic <- site.sum %>% select(cover_LC,cover_TA,canopy_MA,canopy_TA,rugosity) # should it matter if I use site level summary data vs. full follow dataset (where site-level data has been joined?)
+pca1 <- prcomp(na.omit(pred.benthic),center = TRUE,scale. = TRUE)
 
-plot(pca1,type="l")
-ggbiplot(pca1, obs.scale = 1, var.scale = 1, groups=site.sum$Island, ellipse = TRUE, circle = TRUE, loadings.label.repel=TRUE) + scale_color_manual(name="Island", values=c("navy", "darkseagreen", "slategray2")) + theme(legend.direction = 'horizontal', legend.position = 'top') + theme_minimal()
+ggbiplot(pca1, obs.scale = 1, var.scale = 1, groups=site.sum$Island, ellipse = TRUE, circle = TRUE, loadings.label.repel=TRUE, varname.size=2.5, varname.adjust=1.1) + scale_color_manual(name="Island", values=c("navy", "darkseagreen", "slategray2")) + theme(legend.direction = 'horizontal', legend.position = 'top') + theme_minimal()
 
@@
 pca1.scores <- select(site.sum,Site,PC1,PC2)
-data <- data %>% left_join(pca1.scores,by="Site")
+data2 <- data %>% left_join(pca1.scores,by="Site")
+
+setwd("~/github/Scarid-Behavior/data")
+write.csv(data2,file = "data2.csv")
 ```
@@
 ```{r}
-predictors2 <- site.sum %>% select(scar.BM,scar.DEN,carn.BM) # should it matter if I use site level summary data vs. full follow dataset (where site-level data has been joined?)
-pca2 <- prcomp(na.omit(predictors2),center = TRUE,scale. = TRUE) 
+pred.fish <- site.sum %>% select(scar.BM,scar.DEN,carn.BM) # should it matter if I use site level summary data vs. full follow dataset (where site-level data has been joined?)
+pca2 <- prcomp(na.omit(pred.fish),center = TRUE,scale. = TRUE) 
 summary(pca2)
@@
 initial<-data%>%filter(Phase=="i") # for analysis with initial only
-lm_FR <- lm(FR~scar.BM*Species+carn.BM+Length.cm+PC1+PC2, data=data)
+lm_FR <- lm(FR~scar.BM*Species+Phase+carn.BM+Length.cm+PC1+PC2, data=data)
 summary(lm_FR)
@@
 ```{r}
-lm_g.frac <- lm(g.frac~Species*scar.BM+scar.DEN+carn.BM+Phase+Length.cm+PC1+PC2, data=data)
+lm_g.frac <- lm(g.frac~Species*scar.BM+Phase+scar.DEN+carn.BM+Phase+Length.cm+PC1+PC2, data=data)
 summary(lm_g.frac)
@@
 *Notes: much more variable, less predictable. Maybe take out of overall analysis?*
+
+## GAMs
+
+```{r, echo=FALSE}
+vet<-data %>% filter(Species.Code=="qup")
+gamvet<-gam(FR~Phase+s(Length.cm)+s(scar.BM)+(carn.BM)+(PC1)+(PC2), family=gaussian(link=identity),data=vet, method="REML")
+summary(gamvet)
+plot(gamvet)
+```
+```{r, echo=FALSE}
+vir<-data %>% filter(Species.Code=="stop" & Site!="Snapper Point")
+gammvir<-gamm(FR~Phase+s(Length.cm)+s(scar.BM)+s(carn.BM)+s(PC1)+s(PC2), family=gaussian(link=identity),data=vir, random=list(Island=~1),method="REML")
+summary(gammvir)
+plot(gammvir)
+```
 ------------------------

### 2018-01-07 Models.Rmd
1/6

---
title: "Model Trials"
author: "Molly Wilson"
date: "1/6/2018"
output:
  html_document:
    number_sections: yes
    theme: united
    toc: yes
    toc_float: yes
    toc_depth: 3
  pdf_document:
    toc: yes
---
```{r}
library(rmarkdown)
library(dplyr)
library(nlme)
library(pls)
library(mgcv)
library(ggsignif)
```

```{r}
setwd("~/github/Scarid-Behavior/data")
data <- read.csv("data2.csv", stringsAsFactors=FALSE)
vir<-data%>%filter(Species.Code=="stop")
vet<-data%>%filter(Species.Code=="qup")
auro<-data%>%filter(Species.Code=="rbp")
vetvir<-data%>%filter(Species.Code!="rbp")
```

# Model trials

## Basic lm to assess predictors of interest
-site-level predictors: scar.BM, scar.DEN, benthic (PC1,PC2) (excluding carn.BM because almost fully cor. with scar.BM)
-fish-level predictors: species, phase, length
--eventually run separately for different species
--species*scar.BM interaction?
-response variables: g.frac, BR (?), and FR (run separately) *should I calculate site means at this point a la Nash et al?

### Feeding rates
```{r}
lm_FR <- lm(FR~scar.BM*Species+Phase+Length.cm+PC1+PC2, data=data)
summary(lm_FR)
```
*Notes: nearly everything is significant? Significant interaction between scar.BM and species for Spar. viride.*
*R2 goes up slightly with initial only*

### Fraction of time grazing
```{r}
lm_g.frac <- lm(g.frac~Species*scar.BM+Phase+scar.DEN+carn.BM+Phase+Length.cm+PC1+PC2, data=data)
summary(lm_g.frac)
```
*Notes: carn.BM and PC1 are strongest predictors of time spent grazing, scar.BM and phase are also significant. Slight significance of species, length, and species-scar.BM interaction*

## Mixed effects models
- random effect: island

### Feeding rates
```{r}
lme_FR <- lme(FR~Phase+Length.cm+Species*scar.BM+scar.DEN+carn.BM+PC1+PC2,data=data,random=~1|Island, na.action = na.omit)
summary(lme_FR)
```
*Notes: nearly everything is significant except for PC2 and scar.DEN?*

### Fraction of time spent grazing
```{r}
lme_g.frac <- lme(g.frac~Phase+Length.cm+Species*scar.BM+scar.DEN+carn.BM+PC1+PC2,data=data,random=~1|Island, na.action = na.omit)
summary(lme_g.frac)
```
*Notes: carn.BM and phase are strongest predictors, followed by scar.BM, PC1, and length. Also significant interaction between *

## GAMs
```{r, echo=FALSE}
gam<-gam(FR~Species+Phase+s(Length.cm)+s(scar.BM)+s(PC1)+s(PC2), family=gaussian(link=identity),data=data, method="REML")
summary(gam)
plot(gam,pages=1)
```
```{r, echo=FALSE}
gamvet<-gam(FR~Phase+s(Length.cm)+s(scar.BM)+(carn.BM)+s(PC1)+s(PC2), family=gaussian(link=identity),data=vet, method="REML")
summary(gamvet)
plot(gamvet,pages=1)
```
```{r, echo=FALSE}
vir<-data %>% filter(Species.Code=="stop" & Site!="Snapper Point")
gamvir<-gam(FR~Phase+s(Length.cm)+s(scar.BM)+s(PC1)+s(PC2), family=gaussian(link=identity),data=vir,method="REML")
summary(gamvir)
plot(gamvir,pages=1)
```

## GAMMs
```{r, echo=FALSE}
gamm<-gamm(FR~Species+Phase+s(Length.cm)+s(scar.BM)+s(PC1)+s(PC2), family=gaussian(link=identity),data=data, random=list(Island=~1),method="REML")
summary(gamm$gam)
plot(gamm$gam,pages=1)
```
### Modelling effects of predictors

```{r, message=FALSE,warning=FALSE,echo=FALSE}
gamm<-gamm(FR~Species+Phase+s(Length.cm)+s(scar.BM)+s(PC1)+s(PC2), family=gaussian(link=identity),data=data, random=list(Island=~1),method="REML")
summary(gamm$gam)
plot(gamm$gam,pages=1)
```
```{r, message=FALSE,warning=FALSE,echo=FALSE}
gamm1<-gamm(g.frac~Species+Phase+s(Length.cm)+s(scar.BM)+s(PC1)+s(PC2), family=gaussian(link=identity),data=data, random=list(Island=~1),method="REML")
summary(gamm1$gam)
plot(gamm1$gam,pages=1)
```
------------------------

### 2018-01-07 Summary.Rmd
1/6

@@
 ---
-title: "Dec. 1 Committee Meeting"
+title: "Summary: Dec. 1 Committee Meeting"
 author: "Molly Wilson"
@@
 # set up
-setwd("~/Google Drive/Research + Projects/Dissertation/ABB_Grazing/Analysis")
 library(rmarkdown)
@@
 ```{r,message=FALSE,warning=FALSE, echo=FALSE}
+setwd("~/github/Scarid-Behavior/data")
 site.sum <- read.csv("site.sum.csv", stringsAsFactors=FALSE) %>% select(-X)
@@
 
+### Grazing behaviors by island
+```{r,echo=FALSE}
+response <- vetvir %>% select(g.frac,FR,BR)
+log.vars5 <- log(response[,1:3])
+pca3 <- prcomp(na.omit(response),center = TRUE,scale. = TRUE) 
+ggbiplot(pca3, obs.scale = 1, var.scale = 1, groups=vetvir$Island,ellipse = TRUE, circle = TRUE, alpha=0.8,varname.size = 5) + theme_minimal() + scale_color_manual(name="Island", values=c("navy", "darkseagreen", "slategray2")) + labs(title="Grazing behaviors by island") + theme(plot.title = element_text(hjust = 0.5))
+```
+
+*BOXPLOTS*
+
 ### Grazing as a function of scarid biomass
@@
 
-### Grazing PCAs
-
 ```{r, echo=FALSE,eval=FALSE}
@@
 ggbiplot(pca3, obs.scale = 1, var.scale = 1,ellipse = TRUE, circle = TRUE, alpha=0.4,varname.size = 5) + theme(legend.direction = 'horizontal', legend.position = 'top') + theme_minimal() +geom_point(aes(col=vetvir$scar.BM)) + labs(col="Scarid Biomass (g/100m2)", title="Grazing Behaviors by Scarid Biomass") + theme(plot.title = element_text(hjust = 0.5))
-```
-
-```{r,echo=FALSE}
-response <- vetvir %>% select(g.frac,FR,BR)
-log.vars5 <- log(response[,1:3])
-pca3 <- prcomp(na.omit(response),center = TRUE,scale. = TRUE) 
-ggbiplot(pca3, obs.scale = 1, var.scale = 1, groups=vetvir$Island,ellipse = TRUE, circle = TRUE, alpha=0.8,varname.size = 5) + theme_minimal() + scale_color_manual(name="Island", values=c("navy", "darkseagreen", "slategray2")) + labs(title="Grazing behaviors by island") + theme(plot.title = element_text(hjust = 0.5))
 ```

### 2018-01-07 DonorReport.Rmd
update for antigua fisheries

@@
 My research aims to document variations in grazing behavior across different reef conditions, testing for the effects of parrotfish abundance, predator abundance, and benthic condition on the amount of time parrotfish spend feeding, their bite rates, and the duration of their feeding bouts.
-Bonaire represents a relatively pristine Caribbean system, while Antigua has shifted to a more algal-dominated, low scarid state.
-I also conducted surveys in Barbuda, where a recent efforts to reduce parrotfish fishing may have interesting implications on grazing behavior.
 
@@
 
-In August of 2017, I set out to document parrotfish feeding behaviors in Antigua and Barbuda, two neighboring islands with historical overfishing and relatively degraded reefs systems.
+In January and August of 2017, I set out to document parrotfish feeding behaviors in Antigua and Barbuda, two neighboring islands with historical overfishing and relatively degraded reefs systems.
 While parrotfish fishing continues in Antigua, a parrotfish fishing ban has recently been instated in Barbuda.
 Both of these islands would serve as comparisons to the pristine system in Bonaire that are more represenative of reef status in the larger Caribbean region.
+Thanks to help from Ms.
+Appleton and Ms.
+Lovell at Antigua Fisheries, Dr.
+Cooper at Antigua National Parks, and Mr.
+Morris from Barbuda, I received permission to conduct my surveys as well as useful information about fishing pressure around the islands.
+I would also like to thank Ruleo Camacho for his continued help and feedback.
 
+With the support of generous funding awarded through the UCSB Research Accellerator Award, I was able to travel to Antigua along with a Julia Mason, a research assistant also working on her Ph.D.
+in marine ecoogy and fisheries management at Stanford University.
+The funds supplied us with SCUBA tanks, survey gear, boat fuel, a car rental to transport gear, and fuel and food to supply a four-day boat trip to Barbuda.
+During the three weeks we were in the field, we collected benthic and fish survey data at a total of nine sites which, in addition to providing data for this investigation, will also serve as valuable assessment and baseline data for future studies.
+Across the nine sites we coolected behavioral data from a total of 434 individual parrotfish, making a database of nearly 800 fish follows when integrated with the data I collected in Bonaire.
+While the data analysis is still ongoing, the preliminary results are extremely interesting.
+This research would not have been possible without the generous support of my Research Accellerator sponsor.
+
+```{r, eval=FALSE}
 Thanks to generous funding awarded through the UCSB Research Accellerator Award, I was able to travel to Antigua along with a research assistant also working on her Ph.D.
@@
 This research would not have been possible without the generous support of my Research Accellerator sponsor.
+```
 

### 2018-01-31 DonorReport.Rmd
updates 1/31

@@
 output:
-  pdf_document:
-    toc: yes
   html_document:
@@
     toc_depth: 3
+  pdf_document:
+    toc: yes
 ---
@@
 
-```{r, eval=FALSE}
 Thanks to generous funding awarded through the UCSB Research Accellerator Award, I was able to travel to Antigua along with a research assistant also working on her Ph.D.
@@
 This research would not have been possible without the generous support of my Research Accellerator sponsor.
-```
 

### 2018-01-31 GAMMs.Rmd
updates 1/31

---
title: "GAMM Trials"
author: "Molly Wilson"
date: "1/6/2018"
output:
  html_document:
    number_sections: yes
    theme: united
    toc: yes
    toc_float: yes
    toc_depth: 3
  pdf_document:
    toc: yes
---
```{r,message=FALSE,warning=FALSE}
library(rmarkdown)
library(dplyr)
library(nlme)
library(pls)
library(mgcv)
library(ggsignif)
```

```{r,message=FALSE,warning=FALSE}
setwd("~/github/Scarid-Behavior/data")
data <- read.csv("data2.csv", stringsAsFactors=FALSE)
vir<-data%>%filter(Species.Code=="stop") # Sparisoma viride only
vet<-data%>%filter(Species.Code=="qup") # Scarus vetula only
```

# Feeding rate (bites per hour)

## All species
```{r,message=FALSE,warning=FALSE}
gamm<-gamm(FR~Species+Phase+s(Length.cm)+s(scar.BM)+s(PC1)+s(PC2), family=gaussian(link=identity),data=data, random=list(Island=~1),method="REML")
summary(gamm$gam)
plot(gamm$gam,pages=1)
```

## Scarus vetula
```{r,message=FALSE,warning=FALSE}
gamm<-gamm(FR~Phase+s(Length.cm)+s(scar.BM)+s(PC1)+s(PC2), family=gaussian(link=identity),data=vet, random=list(Island=~1),method="REML")
summary(gamm$gam)
plot(gamm$gam,pages=1)
```

## Sparisoma viride
```{r,message=FALSE,warning=FALSE}
gamm<-gamm(FR~Phase+s(Length.cm)+s(scar.BM)+s(PC1)+s(PC2), family=gaussian(link=identity),data=vir, random=list(Island=~1),method="REML")
summary(gamm$gam)
plot(gamm$gam,pages=1)
```

# Fraction of time spent grazing

## All species
```{r,message=FALSE,warning=FALSE}
gamm<-gamm(g.frac~Species+Phase+s(Length.cm)+s(scar.BM)+s(PC1)+s(PC2), family=gaussian(link=identity),data=data, random=list(Island=~1),method="REML")
summary(gamm$gam)
plot(gamm$gam,pages=1)
```

## Scarus vetula
```{r,message=FALSE,warning=FALSE}
gamm<-gamm(g.frac~Phase+s(Length.cm)+s(scar.BM)+s(PC1)+s(PC2), family=gaussian(link=identity),data=vet, random=list(Island=~1),method="REML")
summary(gamm$gam)
plot(gamm$gam,pages=1)
```

## Sparisoma viride
```{r,message=FALSE,warning=FALSE}
gamm<-gamm(g.frac~Phase+s(Length.cm)+s(scar.BM)+s(PC1)+s(PC2), family=gaussian(link=identity),data=vir, random=list(Island=~1),method="REML")
summary(gamm$gam)
plot(gamm$gam,pages=1)
```

### 2018-01-31 Models.Rmd
update 1/31

@@
 
-## Basic lm to assess predictors of interest
--site-level predictors: scar.BM, scar.DEN, benthic (PC1,PC2) (excluding carn.BM because almost fully cor. with scar.BM)
+## Basic lm
+-site-level predictors: scar.BM, scar.DEN, benthic (PC1,PC2)
 -fish-level predictors: species, phase, length
@@
 --species*scar.BM interaction?
--response variables: g.frac, BR (?), and FR (run separately) *should I calculate site means at this point a la Nash et al?
 
-### Feeding rates
 ```{r}
@@
 
-### Fraction of time grazing
+### Adding spearfishing variable
 ```{r}
-lm_g.frac <- lm(g.frac~Species*scar.BM+Phase+scar.DEN+carn.BM+Phase+Length.cm+PC1+PC2, data=data)
-summary(lm_g.frac)
+lm_FR <- lm(FR~scar.BM*Species+Phase+Length.cm+PC1+PC2+Spearfishing, data=data)
+summary(lm_FR)
 ```
-*Notes: carn.BM and PC1 are strongest predictors of time spent grazing, scar.BM and phase are also significant. Slight significance of species, length, and species-scar.BM interaction*
+*Notes: spearfishing masks benthic (PC1/PC2 effects), because highly correlated...*
 
@@
 
-### Feeding rates
 ```{r}
-lme_FR <- lme(FR~Phase+Length.cm+Species*scar.BM+scar.DEN+carn.BM+PC1+PC2,data=data,random=~1|Island, na.action = na.omit)
+lme_FR <- lme(FR~Phase+Length.cm+Species*scar.BM+PC1+PC2+carn.BM,data=data,random=~1|Island, na.action = na.omit)
 summary(lme_FR)
 ```
-*Notes: nearly everything is significant except for PC2 and scar.DEN?*
+*Notes: including island as a random effect makes all site-level variables non-significant. Adding carn.BM back in makes *
 
@@
 ```{r}
-lme_g.frac <- lme(g.frac~Phase+Length.cm+Species*scar.BM+scar.DEN+carn.BM+PC1+PC2,data=data,random=~1|Island, na.action = na.omit)
+lme_g.frac <- lme(g.frac~Phase+Length.cm+Species*scar.BM+scar.DEN+PC1+PC2,data=data,random=~1|Island, na.action = na.omit)
 summary(lme_g.frac)
 ```
-*Notes: carn.BM and phase are strongest predictors, followed by scar.BM, PC1, and length. Also significant interaction between *
+*Notes: carn.BM and phase are strongest predictors, followed by scar.BM, PC1, and length. Also significant interaction between Spar. viride and scar.BM*
 
@@
 ```
-### Modelling effects of predictors
-
-```{r, message=FALSE,warning=FALSE,echo=FALSE}
-gamm<-gamm(FR~Species+Phase+s(Length.cm)+s(scar.BM)+s(PC1)+s(PC2), family=gaussian(link=identity),data=data, random=list(Island=~1),method="REML")
-summary(gamm$gam)
-plot(gamm$gam,pages=1)
-```
-```{r, message=FALSE,warning=FALSE,echo=FALSE}
-gamm1<-gamm(g.frac~Species+Phase+s(Length.cm)+s(scar.BM)+s(PC1)+s(PC2), family=gaussian(link=identity),data=data, random=list(Island=~1),method="REML")
-summary(gamm1$gam)
-plot(gamm1$gam,pages=1)
-```
-------------------------

### 2018-03-15 DonorReport.Rmd
updates, adding new summary file for orals

@@
 library(ggsignif)
+library(car)
 ```
@@
 vet<- vetvir %>% filter(Species.Code=="qup")
+aovFRvet <- aov(FR ~ Island, data=vet) 
+TukeyHSD(aovFRvet)
+plot(aovFRvet, 2) # check normality
+leveneTest(FR ~ Island, data = vet) # check homogeneity of variance
+
 vir<- vetvir %>% filter(Species.Code=="stop")
-aovFRvet <- aov(FR ~ Island, data=vet) 
-#TukeyHSD(aovFRvet)
 aovFRvir <- aov(FR ~ Island, data=vir) 
-#TukeyHSD(aovFRvir)
+TukeyHSD(aovFRvir)
+plot(aovFRvir, 2) # check normality
+leveneTest(FR ~ Island, data = vir) # check homogeneity of variance
 ```
@@
 aovforvet <- aov(for.dur ~ Island, data=vet) 
-#TukeyHSD(aovforvet)
+TukeyHSD(aovforvet)
 aovforvir <- aov(for.dur ~ Island, data=vir) 
-#TukeyHSD(aovforvir)
+TukeyHSD(aovforvir)
 ```

### 2018-03-15 SummaryOrals.Rmd
updates, adding new summary file for orals

---
title: 'Summary: Dissertation Proposal'
author: "Molly Wilson"
date: "3/14/2018"
output:
  html_document:
    number_sections: yes
    theme: united
    toc: yes
    toc_float: yes
    toc_depth: 3
  pdf_document:
    toc: yes
---

This is a summary of the primary figures and tests I have completed for Chapter 2 as of my oral exams (April 2018), as well as accompanying notes and questions.

```{r,message=FALSE,warning=FALSE, echo=FALSE}
# set up
library(rmarkdown)
library(dplyr)
library(tidyr)
library(lazyeval)
library(ggplot2)
library(ellipse)
library(RColorBrewer)
library(ggbiplot)
library(nlme)
library(pls)
library(mgcv)
library(ggsignif)
library(car)
```
```{r,message=FALSE,warning=FALSE, echo=FALSE}
setwd("~/github/Scarid-Behavior/data")
site.sum <- read.csv("site.sum.csv", stringsAsFactors=FALSE) %>% select(-X)
data <- read.csv("sumID2.csv", stringsAsFactors=FALSE) %>% select(Island,Site,Spearfishing,Diving,Species,Species.Code,Phase,Length.cm,g.frac,FR,BR,TDB,for.dur,for.bites,tot.BM,tot.DEN,scar.BM,herb.BM,scar.DEN,conspec.BM,conspec.DEN,carn.BM,carn.DEN,cover_CCA,cover_LC,cover_MA,cover_TA,canopy_MA,canopy_TA,rugosity,int.tot,int.con,int.hetscar,int.dam)
vetvir <- data %>% filter(Species.Code!="rbp")
vetvir20 <- vetvir %>% filter(Length.cm>=15 & Length.cm<=25)
init20<- vetvir %>% filter(Length.cm>=15 & Length.cm<=25 & Phase=="i")
vetvirAB <- vetvir %>% filter(Island!="Barbuda")
vetvirAB20 <- vetvir %>% filter(Island!="Barbuda")
vetvirAB20NS <- vetvirAB20 %>% filter(Site!="Snapper Point")
vetvir20NS <- vetvir20 %>% filter(Site!="Snapper Point")
vetvirNS <- vetvir %>% filter(Site!="Snapper Point")
```

# Correlations among predictor variables

## Benthic variables
```{r,echo=FALSE}
panel.cor <- function(x, y, digits=2, prefix="", cex.cor, ...)
{
    usr <- par("usr"); on.exit(par(usr))
    par(usr = c(0, 1, 0, 1))
    r <- abs(cor(x, y))
    txt <- format(c(r, 0.123456789), digits=digits)[1]
    txt <- paste(prefix, txt, sep="")
    if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
    text(0.5, 0.5, txt, cex = cex.cor * r)
}
pairs(~cover_CCA+cover_LC+cover_MA+canopy_MA+cover_TA+canopy_TA+rugosity, data=data,
      lower.panel=panel.smooth, upper.panel=panel.cor, 
      pch=20, main="Benthic Community Variables",na.action = na.omit)
```

## Selected fish and benthic variables

```{r,echo=FALSE}
panel.cor <- function(x, y, digits=2, prefix="", cex.cor, ...)
{
    usr <- par("usr"); on.exit(par(usr))
    par(usr = c(0, 1, 0, 1))
    r <- abs(cor(x, y))
    txt <- format(c(r, 0.123456789), digits=digits)[1]
    txt <- paste(prefix, txt, sep="")
    if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
    text(0.5, 0.5, txt, cex = cex.cor * r)
}
pairs(~scar.BM+scar.DEN+carn.BM+cover_LC+cover_TA+canopy_MA+canopy_TA+rugosity, data=data,
      lower.panel=panel.smooth, upper.panel=panel.cor, 
      pch=20, main="Selected Fish and Benthic Variables",na.action = na.omit)
```

# PCAs

```{r, echo=FALSE}
pred.benthic <- site.sum %>% select(cover_LC,cover_TA,canopy_MA,canopy_TA,rugosity) # should it matter if I use site level summary data vs. full follow dataset (where site-level data has been joined?)
pca1 <- prcomp(na.omit(pred.benthic),center = TRUE,scale. = TRUE)

ggbiplot(pca1, obs.scale = 1, var.scale = 1, groups=site.sum$Island, ellipse = TRUE, circle = TRUE, loadings.label.repel=TRUE, varname.size=2.5, varname.adjust=1.1) + scale_color_manual(name="Island", values=c("navy", "darkseagreen", "slategray2")) + theme(legend.direction = 'horizontal', legend.position = 'top') + theme_minimal()

# add PC1 and PC2 as variables in dataset
pca1.scores <- as.data.frame(pca1$x) %>% select(PC1,PC2)
site.sum <- site.sum  %>% bind_cols(pca1.scores)
pca1.scores <- select(site.sum,Site,PC1,PC2)
data2 <- data %>% left_join(pca1.scores,by="Site")

setwd("~/github/Scarid-Behavior/data")
write.csv(data2,file = "data2.csv")
```

# Grazing differences between islands

## Feeding rates

```{r, message=FALSE,warning=FALSE,echo=FALSE}
ggplot(vetvir20, aes(x=Island, y=FR)) + geom_boxplot() + facet_grid(.~Species) +
stat_summary(fun.y = mean, geom="point", shape=5, size=3) + labs(y="Feeding Rate (bites/hr)",x="",title="Feeding Rate by Island and Species") + theme_bw() + theme(plot.title = element_text(hjust = 0.5))
```

```{r, message=FALSE,warning=FALSE,echo=FALSE}
vet<- vetvir %>% filter(Species.Code=="qup")
aovFRvet <- aov(FR ~ Island, data=vet) 
summary.aov(aovFRvet) # significant differences
TukeyHSD(aovFRvet)
plot(aovFRvet, 2) # normally distributed
leveneTest(FR ~ Island, data = vet) # homogeneity of variance

vir<- vetvir %>% filter(Species.Code=="stop")
aovFRvir <- aov(FR ~ Island, data=vir, white.adjust=TRUE) 
summary.aov(aovFRvir) # significant differences
TukeyHSD(aovFRvir)
plot(aovFRvir, 2) # normally distributed
leveneTest(FR ~ Island, data = vir) # heteroskedastic - add white.adjust=TRUE
```

## Bite rates

```{r, message=FALSE,warning=FALSE,echo=FALSE}
ggplot(vetvir20, aes(x=Island, y=BR)) + geom_boxplot() + facet_grid(.~Species) +
stat_summary(fun.y = mean, geom="point", shape=5, size=3) + labs(y="Bite Rate (bites/sec)",x="",title="Bite Rate while Grazing by Island and Species") + theme_bw() + theme(plot.title = element_text(hjust = 0.5))
```

```{r, message=FALSE,warning=FALSE,echo=FALSE}
vet<- vetvir %>% filter(Species.Code=="qup")
aovBRvet <- aov(BR ~ Island, data=vet, white.adjust=TRUE) 
summary.aov(aovBRvet) # no significant differences
suplot(aovBRvet, 2) # fairly normal?
leveneTest(BR ~ Island, data = vet) # heteroskedastic - add white.adjust=TRUE

vir<- vetvir %>% filter(Species.Code=="stop")
aovBRvir <- aov(BR ~ Island, data=vir, white.adjust=TRUE) 
summary.aov(aovBRvir) # no significant differences
plot(aovBRvir, 2) # fairly normal?
leveneTest(BR ~ Island, data = vir) # heteroskedastic - add white.adjust=TRUE
```

## Fraction of time spent grazing

```{r, message=FALSE,warning=FALSE,echo=FALSE}
ggplot(vetvir20, aes(x=Island, y=g.frac)) + geom_boxplot() + facet_grid(.~Species) +
stat_summary(fun.y = mean, geom="point", shape=5, size=3) + labs(y="Time fraction",x="",title="Fraction of Time Spent Grazing by Island and Species") + theme_bw() + theme(plot.title = element_text(hjust = 0.5))
```

```{r, message=FALSE,warning=FALSE,echo=FALSE}
vet<- vetvir %>% filter(Species.Code=="qup")
aovBRvet <- aov(BR ~ Island, data=vet, white.adjust=TRUE) 
summary.aov(aovBRvet) # no significant differences
suplot(aovBRvet, 2) # fairly normal?
leveneTest(BR ~ Island, data = vet) # heteroskedastic - add white.adjust=TRUE

vir<- vetvir %>% filter(Species.Code=="stop")
aovBRvir <- aov(BR ~ Island, data=vir, white.adjust=TRUE) 
summary.aov(aovBRvir) # no significant differences
plot(aovBRvir, 2) # fairly normal?
leveneTest(BR ~ Island, data = vir) # heteroskedastic - add white.adjust=TRUE
```

## Grazing intensity

```{r, message=FALSE,warning=FALSE,echo=FALSE}
ggplot(vetvir20, aes(x=Island, y=for.bites)) + geom_boxplot()  + labs(y="Bites per foray",x="",title="Grazing intensity by Island and Species") + theme_bw() + theme(plot.title = element_text(hjust = 0.5)) + geom_signif(comparisons = list(c("Antigua", "Barbuda"),c("Antigua","Bonaire"),c("Barbuda","Bonaire")), map_signif_level=TRUE) + facet_grid(.~Species)
```
```{r, message=FALSE,warning=FALSE,echo=FALSE}
vet<- vetvir %>% filter(Species.Code=="qup")
aovGIvet <- aov(for.bites ~ Island, data=vet, white.adjust=TRUE) 
summary.aov(aovGIvet) # significant differences
TukeyHSD(aovGIvet)
plot(aovGIvet, 2) # fairly normally distributed?
leveneTest(for.bites ~ Island, data = vet) # heteroskedastic - add white.adjust=TRUE

vir<- vetvir %>% filter(Species.Code=="stop")
aovGIvir <- aov(for.bites ~ Island, data=vir, white.adjust=TRUE) 
summary.aov(aovGIvir) # significant differences
TukeyHSD(aovGIvir)
plot(aovGIvir, 2) # fairly normally distributed?
leveneTest(for.bites ~ Island, data = vir) # heteroskedastic - add white.adjust=TRUE
```

# Bivariate plots

## Grazing as a function of fish size
*Due to the lack of larger fish in Barbuda, these length-based relationships focus on comparisons between Antigua and Bonaire*

```{r, message=FALSE,warning=FALSE,echo=FALSE}
ggplot(vetvirAB,aes(x=Length.cm,y=FR,colour=Island))+facet_grid(.~Species)+geom_point(shape=18)+scale_shape(solid=FALSE)+geom_smooth(method=lm, fullrange=FALSE)+ggtitle("Feeding rates by fish length") + theme_minimal() +labs(y="Feeding rates (bites/hr)",x="Length (cm)") + theme(plot.title = element_text(hjust = 0.5))  + scale_color_manual(values=c("navy", "slategray2"))
```

```{r, message=FALSE,warning=FALSE,echo=FALSE}
ggplot(vetvirAB,aes(x=Length.cm,y=BR,colour=Island))+facet_grid(.~Species)+geom_point(shape=18)+scale_shape(solid=FALSE)+geom_smooth(method=lm, fullrange=FALSE)+ggtitle("Bite rate by fish length") + theme_minimal() +labs(y="Bite Rate (bites/sec)",x="Length (cm)") + theme(plot.title = element_text(hjust = 0.5))  + scale_color_manual(values=c("navy", "slategray2"))
```

```{r, message=FALSE,warning=FALSE,echo=FALSE}
ggplot(vetvirAB,aes(x=Length.cm,y=g.frac,colour=Island))+facet_grid(.~Species)+geom_point(shape=18)+scale_shape(solid=FALSE)+geom_smooth(method=lm, fullrange=FALSE)+ggtitle("Fraction of time spent grazing by fish length") + theme_minimal() +labs(y="Time fraction",x="Length (cm)") + theme(plot.title = element_text(hjust = 0.5))  + scale_color_manual(values=c("navy", "slategray2"))
```

```{r, message=FALSE,warning=FALSE,echo=FALSE}
ggplot(vetvirAB,aes(x=Length.cm,y=for.bites,colour=Island))+facet_grid(.~Species)+geom_point(shape=18)+scale_shape(solid=FALSE)+geom_smooth(method=lm, fullrange=FALSE)+ggtitle("Grazing intensity by fish length") + theme_minimal() +labs(y="Bites/foray",x="Length (cm)") + theme(plot.title = element_text(hjust = 0.5))  + scale_color_manual(values=c("navy", "slategray2"))
```

## Grazing impact as a function of scarid biomass

```{r, message=FALSE,warning=FALSE,echo=FALSE}
ggplot(vetvir20NS,aes(x=scar.BM,y=FR))+facet_grid(.~Species)+geom_point(shape=18)+geom_smooth(method=loess, fullrange=FALSE)+ggtitle("Feeding rate by scarid biomass") + theme_minimal() +labs(y="Feeding rate (bites/hr)",x="Scarid Biomass (g/100m2)") + theme(plot.title = element_text(hjust = 0.5))
```

```{r, message=FALSE,warning=FALSE,echo=FALSE}
ggplot(vetvir20NS,aes(x=scar.BM,y=BR))+facet_grid(.~Species)+geom_point(shape=18)+geom_smooth(method=loess, fullrange=FALSE)+ggtitle("Bite rate while grazing by scarid biomass") + theme_minimal() +labs(y="Bite rate (bites/sec)",x="Scarid Biomass (g/100m2)") + theme(plot.title = element_text(hjust = 0.5))
```

```{r, message=FALSE,warning=FALSE,echo=FALSE}
ggplot(vetvir20NS,aes(x=scar.BM,y=g.frac))+facet_grid(.~Species)+geom_point(shape=18)+geom_smooth(method=loess, fullrange=FALSE)+ggtitle("Fraction of time spent grazing by scarid biomass") + theme_minimal() +labs(y="Time fraction",x="Scarid Biomass (g/100m2)") + theme(plot.title = element_text(hjust = 0.5))
```

```{r, message=FALSE,warning=FALSE,echo=FALSE}
ggplot(vetvir20NS,aes(x=scar.BM,y=for.bites))+facet_grid(.~Species)+geom_point(shape=18)+geom_smooth(method=loess, fullrange=FALSE)+ggtitle("Grazing intensity by scarid biomass") + theme_minimal() +labs(y="Bites/foray",x="Scarid Biomass (g/100m2)") + theme(plot.title = element_text(hjust = 0.5))
```

# GAMMs

## Scarus vetula

### Feeding rates
```{r,message=FALSE,warning=FALSE}
vetinit <- vet %>% filter(Phase=="i")
gamm<-gamm(FR~s(Length.cm)+s(scar.BM)+s(PC1)+s(PC2), family=gaussian(link=identity),data=vetinit, random=list(Island=~1),method="REML")
summary(gamm$gam) # only PC1 and PC2 are significant, scar.BM is almost significant
plot(gamm$gam,pages=1)
```

### Fraction of time spent grazing
```{r,message=FALSE,warning=FALSE}
gamm<-gamm(g.frac~s(Length.cm)+s(scar.BM)+s(PC1)+s(PC2), family=gaussian(link=identity),data=vetinit, random=list(Island=~1),method="REML")
summary(gamm$gam) # PC1 and PC2 are most significant, length is significant, scar.BM is almost significant
plot(gamm$gam,pages=1)
```

### Grazing intensity
```{r,message=FALSE,warning=FALSE}
gamm<-gamm(for.bites~s(Length.cm)+s(scar.BM)+s(PC1)+s(PC2), family=gaussian(link=identity),data=vetinit, random=list(Island=~1),method="REML")
summary(gamm$gam) # very low R2, only PC1 and scar.BM significant
plot(gamm$gam,pages=1)
```

## Sparisoma viride

### Feeding rates
```{r,message=FALSE,warning=FALSE}
virinit<-vir%>% filter(Phase=="i")
gamm<-gamm(FR~Phase+s(Length.cm)+s(scar.BM)+s(PC1)+s(PC2), family=gaussian(link=identity),data=vir, random=list(Island=~1),method="REML")
summary(gamm$gam)
plot(gamm$gam,pages=1)
```

### Fraction of time spent grazing
```{r,message=FALSE,warning=FALSE}
gamm<-gamm(g.frac~s(Length.cm)+s(scar.BM)+s(PC1)+s(PC2), family=gaussian(link=identity),data=virinit, random=list(Island=~1),method="REML")
summary(gamm$gam)
plot(gamm$gam,pages=1)
```

### Grazing intensity
```{r,message=FALSE,warning=FALSE}
gamm<-gamm(for.bites~s(Length.cm)+s(scar.BM)+s(PC1)+s(PC2), family=gaussian(link=identity),data=virinit, random=list(Island=~1),method="REML")
summary(gamm$gam)
plot(gamm$gam,pages=1)
```

### 2018-03-22 Graphing+Analysis.Rmd
updates, adding new summary file for orals
3.21

@@
 vet<-data %>% filter(Species.Code=="qup")
-gamvet<-gam(FR~Phase+s(Length.cm)+s(scar.BM)+(carn.BM)+(PC1)+(PC2), family=gaussian(link=identity),data=vet, method="REML")
+gamvet<-gam(FR~Phase+s(Length.cm)+s(scar.BM)+(PC1)+(PC2), family=gaussian(link=identity),data=vet, method="REML")
 summary(gamvet)

### 2018-03-22 SummaryOrals.Rmd
3.21

@@
 vetvirAB20 <- vetvir %>% filter(Island!="Barbuda")
+vetvirABinit <- vetvirAB %>% filter(Phase=="i")
+
 vetvirAB20NS <- vetvirAB20 %>% filter(Site!="Snapper Point")
@@
 
+## All predictors
+```{r}
+predictors <- site.sum %>% select(cover_LC,cover_TA,canopy_MA,canopy_TA,rugosity,scar.BM,scar.DEN,carn.BM) # should it matter if I use site level summary data vs.
+full follow dataset (where site-level data has been joined?)
+pca <- prcomp(na.omit(predictors),center = TRUE,scale. = TRUE) 
+summary(pca)
+
+plot(pca,type="l")
+ggbiplot(pca, obs.scale = 1, var.scale = 1, groups=site.sum$Island, ellipse = TRUE, circle = TRUE, varname.size = 2) + scale_color_manual(name="Island", values=c("navy", "darkseagreen", "slategray2")) + theme(legend.direction = 'horizontal', legend.position = 'top') + theme_minimal()
+```
+
+## Benthic predictors
+
 ```{r, echo=FALSE}
@@
 ```{r, message=FALSE,warning=FALSE,echo=FALSE}
-ggplot(vetvir20, aes(x=Island, y=FR)) + geom_boxplot() + facet_grid(.~Species) +
-stat_summary(fun.y = mean, geom="point", shape=5, size=3) + labs(y="Feeding Rate (bites/hr)",x="",title="Feeding Rate by Island and Species") + theme_bw() + theme(plot.title = element_text(hjust = 0.5))
+vetvir20init<-vetvir20 %>% filter(Phase=="i")
+ggplot(vetvir20init, aes(x=Island, y=FR)) + geom_boxplot() +
+stat_summary(fun.y = mean, geom="point", shape=5, size=3) + labs(y="Feeding Rate (bites/hr)",x="",title="Feeding Rate by Island and Species (Initial Phase)") + theme_bw() + theme(plot.title = element_text(hjust = 0.5)) + geom_signif(comparisons = list(c("Antigua", "Barbuda"),c("Antigua","Bonaire"),c("Barbuda","Bonaire")), map_signif_level=TRUE) + facet_grid(.~Species)
+ggsave("bp_fr.png")
 ```
@@
 ```{r, message=FALSE,warning=FALSE,echo=FALSE}
-vet<- vetvir %>% filter(Species.Code=="qup")
-aovFRvet <- aov(FR ~ Island, data=vet) 
+vet20init<- vetvir20init %>% filter(Species.Code=="qup")
+aovFRvet <- aov(FR ~ Island, data=vet20init) 
 summary.aov(aovFRvet) # significant differences
@@
 
-vir<- vetvir %>% filter(Species.Code=="stop")
-aovFRvir <- aov(FR ~ Island, data=vir, white.adjust=TRUE) 
+vir20init<- vetvir20init %>% filter(Species.Code=="stop")
+aovFRvir <- aov(FR ~ Island, data=vir20init, white.adjust=TRUE) 
 summary.aov(aovFRvir) # significant differences
@@
 ```{r, message=FALSE,warning=FALSE,echo=FALSE}
-ggplot(vetvir20, aes(x=Island, y=BR)) + geom_boxplot() + facet_grid(.~Species) +
-stat_summary(fun.y = mean, geom="point", shape=5, size=3) + labs(y="Bite Rate (bites/sec)",x="",title="Bite Rate while Grazing by Island and Species") + theme_bw() + theme(plot.title = element_text(hjust = 0.5))
+ggplot(vetvir20init, aes(x=Island, y=BR)) + geom_boxplot() +
+stat_summary(fun.y = mean, geom="point", shape=5, size=3) + labs(y="Bite Rate (bites/sec)",x="",title="Bite Rate while Grazing by Island and Species") + theme_bw() + theme(plot.title = element_text(hjust = 0.5)) + geom_signif(comparisons = list(c("Antigua", "Barbuda"),c("Antigua","Bonaire"),c("Barbuda","Bonaire")), map_signif_level=TRUE) + facet_grid(.~Species)
+ggsave("bp_br.png")
 ```
@@
 ```{r, message=FALSE,warning=FALSE,echo=FALSE}
-vet<- vetvir %>% filter(Species.Code=="qup")
-aovBRvet <- aov(BR ~ Island, data=vet, white.adjust=TRUE) 
+aovBRvet <- aov(BR ~ Island, data=vet20init, white.adjust=TRUE) 
 summary.aov(aovBRvet) # no significant differences
-suplot(aovBRvet, 2) # fairly normal?
+TukeyHSD(aovBRvet)
+plot(aovBRvet, 2) # fairly normal?
 leveneTest(BR ~ Island, data = vet) # heteroskedastic - add white.adjust=TRUE
 
-vir<- vetvir %>% filter(Species.Code=="stop")
-aovBRvir <- aov(BR ~ Island, data=vir, white.adjust=TRUE) 
+aovBRvir <- aov(BR ~ Island, data=vir20init, white.adjust=TRUE) 
 summary.aov(aovBRvir) # no significant differences
+TukeyHSD(aovBRvet)
 plot(aovBRvir, 2) # fairly normal?
@@
 ```{r, message=FALSE,warning=FALSE,echo=FALSE}
-ggplot(vetvir20, aes(x=Island, y=g.frac)) + geom_boxplot() + facet_grid(.~Species) +
-stat_summary(fun.y = mean, geom="point", shape=5, size=3) + labs(y="Time fraction",x="",title="Fraction of Time Spent Grazing by Island and Species") + theme_bw() + theme(plot.title = element_text(hjust = 0.5))
+ggplot(vetvir20init, aes(x=Island, y=g.frac)) + geom_boxplot() +
+stat_summary(fun.y = mean, geom="point", shape=5, size=3) + labs(y="Time fraction",x="",title="Fraction of Time Spent Grazing by Island and Species") + theme_bw() + theme(plot.title = element_text(hjust = 0.5)) + geom_signif(comparisons = list(c("Antigua", "Barbuda"),c("Antigua","Bonaire"),c("Barbuda","Bonaire")), map_signif_level=TRUE) + facet_grid(.~Species)
+ggsave("bp_gfrac.png")
 ```
@@
 ```{r, message=FALSE,warning=FALSE,echo=FALSE}
-vet<- vetvir %>% filter(Species.Code=="qup")
-aovBRvet <- aov(BR ~ Island, data=vet, white.adjust=TRUE) 
+aovBRvet <- aov(BR ~ Island, data=vet20init, white.adjust=TRUE) 
 summary.aov(aovBRvet) # no significant differences
@@
 vir<- vetvir %>% filter(Species.Code=="stop")
-aovBRvir <- aov(BR ~ Island, data=vir, white.adjust=TRUE) 
+aovBRvir <- aov(BR ~ Island, data=vir20init, white.adjust=TRUE) 
 summary.aov(aovBRvir) # no significant differences
@@
 ```{r, message=FALSE,warning=FALSE,echo=FALSE}
-ggplot(vetvir20, aes(x=Island, y=for.bites)) + geom_boxplot()  + labs(y="Bites per foray",x="",title="Grazing intensity by Island and Species") + theme_bw() + theme(plot.title = element_text(hjust = 0.5)) + geom_signif(comparisons = list(c("Antigua", "Barbuda"),c("Antigua","Bonaire"),c("Barbuda","Bonaire")), map_signif_level=TRUE) + facet_grid(.~Species)
+ggplot(vetvir20init, aes(x=Island, y=for.bites)) + geom_boxplot()  + labs(y="Bites per foray",x="",title="Grazing intensity by Island and Species") + theme_bw() + theme(plot.title = element_text(hjust = 0.5)) + geom_signif(comparisons = list(c("Antigua", "Barbuda"),c("Antigua","Bonaire"),c("Barbuda","Bonaire")), map_signif_level=TRUE) + facet_grid(.~Species)
+ggsave("bp_gi.png")
 ```
 ```{r, message=FALSE,warning=FALSE,echo=FALSE}
-vet<- vetvir %>% filter(Species.Code=="qup")
-aovGIvet <- aov(for.bites ~ Island, data=vet, white.adjust=TRUE) 
+aovGIvet <- aov(for.bites ~ Island, data=vet20init, white.adjust=TRUE) 
 summary.aov(aovGIvet) # significant differences
@@
 
-vir<- vetvir %>% filter(Species.Code=="stop")
-aovGIvir <- aov(for.bites ~ Island, data=vir, white.adjust=TRUE) 
+aovGIvir <- aov(for.bites ~ Island, data=vir20init, white.adjust=TRUE) 
 summary.aov(aovGIvir) # significant differences
@@
 ```{r, message=FALSE,warning=FALSE,echo=FALSE}
+
 ggplot(vetvirAB,aes(x=Length.cm,y=FR,colour=Island))+facet_grid(.~Species)+geom_point(shape=18)+scale_shape(solid=FALSE)+geom_smooth(method=lm, fullrange=FALSE)+ggtitle("Feeding rates by fish length") + theme_minimal() +labs(y="Feeding rates (bites/hr)",x="Length (cm)") + theme(plot.title = element_text(hjust = 0.5))  + scale_color_manual(values=c("navy", "slategray2"))
+ggsave("sp_length.png")
 ```
@@
 ```{r, message=FALSE,warning=FALSE,echo=FALSE}
-ggplot(vetvirAB,aes(x=Length.cm,y=BR,colour=Island))+facet_grid(.~Species)+geom_point(shape=18)+scale_shape(solid=FALSE)+geom_smooth(method=lm, fullrange=FALSE)+ggtitle("Bite rate by fish length") + theme_minimal() +labs(y="Bite Rate (bites/sec)",x="Length (cm)") + theme(plot.title = element_text(hjust = 0.5))  + scale_color_manual(values=c("navy", "slategray2"))
+ggplot(vetvirABinit,aes(x=Length.cm,y=BR,colour=Island))+facet_grid(.~Species)+geom_point(shape=18)+scale_shape(solid=FALSE)+geom_smooth(method=lm, fullrange=FALSE)+ggtitle("Bite rate by fish length") + theme_minimal() +labs(y="Bite Rate (bites/sec)",x="Length (cm)") + theme(plot.title = element_text(hjust = 0.5))  + scale_color_manual(values=c("navy", "slategray2"))
 ```
@@
 ```{r, message=FALSE,warning=FALSE,echo=FALSE}
-ggplot(vetvirAB,aes(x=Length.cm,y=g.frac,colour=Island))+facet_grid(.~Species)+geom_point(shape=18)+scale_shape(solid=FALSE)+geom_smooth(method=lm, fullrange=FALSE)+ggtitle("Fraction of time spent grazing by fish length") + theme_minimal() +labs(y="Time fraction",x="Length (cm)") + theme(plot.title = element_text(hjust = 0.5))  + scale_color_manual(values=c("navy", "slategray2"))
+ggplot(vetvirABinit,aes(x=Length.cm,y=g.frac,colour=Island))+facet_grid(.~Species)+geom_point(shape=18)+scale_shape(solid=FALSE)+geom_smooth(method=lm, fullrange=FALSE)+ggtitle("Fraction of time spent grazing by fish length") + theme_minimal() +labs(y="Time fraction",x="Length (cm)") + theme(plot.title = element_text(hjust = 0.5))  + scale_color_manual(values=c("navy", "slategray2"))
 ```
@@
 ```{r, message=FALSE,warning=FALSE,echo=FALSE}
-ggplot(vetvirAB,aes(x=Length.cm,y=for.bites,colour=Island))+facet_grid(.~Species)+geom_point(shape=18)+scale_shape(solid=FALSE)+geom_smooth(method=lm, fullrange=FALSE)+ggtitle("Grazing intensity by fish length") + theme_minimal() +labs(y="Bites/foray",x="Length (cm)") + theme(plot.title = element_text(hjust = 0.5))  + scale_color_manual(values=c("navy", "slategray2"))
+ggplot(vetvirABinit,aes(x=Length.cm,y=for.bites,colour=Island))+facet_grid(.~Species)+geom_point(shape=18)+scale_shape(solid=FALSE)+geom_smooth(method=lm, fullrange=FALSE)+ggtitle("Grazing intensity by fish length") + theme_minimal() +labs(y="Bites/foray",x="Length (cm)") + theme(plot.title = element_text(hjust = 0.5))  + scale_color_manual(values=c("navy", "slategray2"))
 ```
@@
 ```{r, message=FALSE,warning=FALSE,echo=FALSE}
-ggplot(vetvir20NS,aes(x=scar.BM,y=FR))+facet_grid(.~Species)+geom_point(shape=18)+geom_smooth(method=loess, fullrange=FALSE)+ggtitle("Feeding rate by scarid biomass") + theme_minimal() +labs(y="Feeding rate (bites/hr)",x="Scarid Biomass (g/100m2)") + theme(plot.title = element_text(hjust = 0.5))
+ggplot(vetvir20init,aes(x=scar.BM,y=FR))+facet_grid(.~Species)+geom_point(shape=18)+geom_smooth(method=loess, fullrange=FALSE)+ggtitle("Feeding rate by scarid biomass") + theme_minimal() +labs(y="Feeding rate (bites/hr)",x="Scarid Biomass (g/100m2)") + theme(plot.title = element_text(hjust = 0.5))
+ggsave("sp_scarbm.png")
 ```
@@
 ```{r, message=FALSE,warning=FALSE,echo=FALSE}
-ggplot(vetvir20NS,aes(x=scar.BM,y=BR))+facet_grid(.~Species)+geom_point(shape=18)+geom_smooth(method=loess, fullrange=FALSE)+ggtitle("Bite rate while grazing by scarid biomass") + theme_minimal() +labs(y="Bite rate (bites/sec)",x="Scarid Biomass (g/100m2)") + theme(plot.title = element_text(hjust = 0.5))
+ggplot(vetvir20init,aes(x=scar.BM,y=BR))+facet_grid(.~Species)+geom_point(shape=18)+geom_smooth(method=loess, fullrange=FALSE)+ggtitle("Bite rate while grazing by scarid biomass") + theme_minimal() +labs(y="Bite rate (bites/sec)",x="Scarid Biomass (g/100m2)") + theme(plot.title = element_text(hjust = 0.5))
 ```
@@
 ```{r, message=FALSE,warning=FALSE,echo=FALSE}
-ggplot(vetvir20NS,aes(x=scar.BM,y=g.frac))+facet_grid(.~Species)+geom_point(shape=18)+geom_smooth(method=loess, fullrange=FALSE)+ggtitle("Fraction of time spent grazing by scarid biomass") + theme_minimal() +labs(y="Time fraction",x="Scarid Biomass (g/100m2)") + theme(plot.title = element_text(hjust = 0.5))
+ggplot(vetvir20init,aes(x=scar.BM,y=g.frac))+facet_grid(.~Species)+geom_point(shape=18)+geom_smooth(method=loess, fullrange=FALSE)+ggtitle("Fraction of time spent grazing by scarid biomass") + theme_minimal() +labs(y="Time fraction",x="Scarid Biomass (g/100m2)") + theme(plot.title = element_text(hjust = 0.5))
 ```
@@
 ```{r, message=FALSE,warning=FALSE,echo=FALSE}
-ggplot(vetvir20NS,aes(x=scar.BM,y=for.bites))+facet_grid(.~Species)+geom_point(shape=18)+geom_smooth(method=loess, fullrange=FALSE)+ggtitle("Grazing intensity by scarid biomass") + theme_minimal() +labs(y="Bites/foray",x="Scarid Biomass (g/100m2)") + theme(plot.title = element_text(hjust = 0.5))
+ggplot(vetvir20init,aes(x=scar.BM,y=for.bites))+facet_grid(.~Species)+geom_point(shape=18)+geom_smooth(method=loess, fullrange=FALSE)+ggtitle("Grazing intensity by scarid biomass") + theme_minimal() +labs(y="Bites/foray",x="Scarid Biomass (g/100m2)") + theme(plot.title = element_text(hjust = 0.5))
 ```
@@
 vetinit <- vet %>% filter(Phase=="i")
+vetterm <- vet %>% filter(Phase=="t")
 gamm<-gamm(FR~s(Length.cm)+s(scar.BM)+s(PC1)+s(PC2), family=gaussian(link=identity),data=vetinit, random=list(Island=~1),method="REML")
 summary(gamm$gam) # only PC1 and PC2 are significant, scar.BM is almost significant
+AIC(gamm$lme)
 plot(gamm$gam,pages=1)
@@
 ```{r,message=FALSE,warning=FALSE}
-gamm<-gamm(for.bites~s(Length.cm)+s(scar.BM)+s(PC1)+s(PC2), family=gaussian(link=identity),data=virinit, random=list(Island=~1),method="REML")
+gamm<-gamm(for.bites~s(Length.cm)+s(scar.BM)+s(PC1)+s(PC2), family=gaussian(link=identity),data=vir, random=list(Island=~1),method="REML")
 summary(gamm$gam)

### 2018-03-22 SummaryOrals.Rmd
3.21

@@
 gamm<-gamm(FR~s(Length.cm)+s(scar.BM)+s(PC1)+s(PC2), family=gaussian(link=identity),data=vetinit, random=list(Island=~1),method="REML")
-summary(gamm$gam) # only PC1 and PC2 are significant, scar.BM is almost significant
+summary(gamm$gam) # only PC1 and PC2 are significant, length slightly, scar.BM is not
 AIC(gamm$lme)
@@
 ```{r,message=FALSE,warning=FALSE}
-virinit<-vir%>% filter(Phase=="i")
-gamm<-gamm(FR~Phase+s(Length.cm)+s(scar.BM)+s(PC1)+s(PC2), family=gaussian(link=identity),data=vir, random=list(Island=~1),method="REML")
+virinit <- data %>% filter(Species.Code=="stop" & Phase=="i")
+gamm<-gamm(FR~Phase+s(Length.cm)+s(scar.BM)+s(PC1)+s(PC2), family=gaussian(link=identity),data=virinit, random=list(Island=~1),method="REML")
 summary(gamm$gam)

### 2018-03-22 Graphing+Analysis.Rmd
3.21

@@
 ```{r}
-predictors <- site.sum %>% select(cover_LC,cover_TA,canopy_MA,canopy_TA,rugosity,scar.BM,scar.DEN,carn.BM) # should it matter if I use site level summary data vs.
+predictors <- site.sum %>% select(cover_LC,cover_TA,canopy_MA,canopy_TA,rugosity,scar.BM,scar.DEN,tot.BM) # should it matter if I use site level summary data vs.
 full follow dataset (where site-level data has been joined?)
@@
 ggbiplot(pca, obs.scale = 1, var.scale = 1, groups=site.sum$Island, ellipse = TRUE, circle = TRUE, varname.size = 2) + scale_color_manual(name="Island", values=c("navy", "darkseagreen", "slategray2")) + theme(legend.direction = 'horizontal', legend.position = 'top') + theme_minimal()
+ggsave("PCApred.png")
 ```

### 2018-04-04 SummaryOrals.Rmd
4/4

--- 2018-03-22 SummaryOrals.Rmd
@@
 site.sum <- read.csv("site.sum.csv", stringsAsFactors=FALSE) %>% select(-X)
-data <- read.csv("sumID2.csv", stringsAsFactors=FALSE) %>% select(Island,Site,Spearfishing,Diving,Species,Species.Code,Phase,Length.cm,g.frac,FR,BR,TDB,for.dur,for.bites,tot.BM,tot.DEN,scar.BM,herb.BM,scar.DEN,conspec.BM,conspec.DEN,carn.BM,carn.DEN,cover_CCA,cover_LC,cover_MA,cover_TA,canopy_MA,canopy_TA,rugosity,int.tot,int.con,int.hetscar,int.dam)
+data <- read.csv("data2.csv", stringsAsFactors=FALSE)
+
 vetvir <- data %>% filter(Species.Code!="rbp")
@@
 ```{r}
-predictors <- site.sum %>% select(cover_LC,cover_TA,canopy_MA,canopy_TA,rugosity,scar.BM,scar.DEN,carn.BM) # should it matter if I use site level summary data vs.
+predictors <- site.sum %>% select(cover_LC,cover_TA,canopy_MA,canopy_TA,rugosity,scar.BM,scar.DEN) # should it matter if I use site level summary data vs.
 full follow dataset (where site-level data has been joined?)
@@
 ggbiplot(pca, obs.scale = 1, var.scale = 1, groups=site.sum$Island, ellipse = TRUE, circle = TRUE, varname.size = 2) + scale_color_manual(name="Island", values=c("navy", "darkseagreen", "slategray2")) + theme(legend.direction = 'horizontal', legend.position = 'top') + theme_minimal()
+ggsave("pca_pred.png")
 ```
@@
 plot(aovFRvet, 2) # normally distributed
-leveneTest(FR ~ Island, data = vet) # homogeneity of variance
+leveneTest(FR ~ Island, data = vet20init) # homogeneity of variance
 
@@
 plot(aovFRvir, 2) # normally distributed
-leveneTest(FR ~ Island, data = vir) # heteroskedastic - add white.adjust=TRUE
+leveneTest(FR ~ Island, data = vir20init) # heteroskedastic - add white.adjust=TRUE
 ```
@@
 plot(aovBRvet, 2) # fairly normal?
-leveneTest(BR ~ Island, data = vet) # heteroskedastic - add white.adjust=TRUE
+leveneTest(BR ~ Island, data = vet20init) # heteroskedastic - add white.adjust=TRUE
 
@@
 plot(aovBRvir, 2) # fairly normal?
-leveneTest(BR ~ Island, data = vir) # heteroskedastic - add white.adjust=TRUE
+leveneTest(BR ~ Island, data = vir20init) # heteroskedastic - add white.adjust=TRUE
 ```
@@
 summary.aov(aovBRvet) # no significant differences
-suplot(aovBRvet, 2) # fairly normal?
-leveneTest(BR ~ Island, data = vet) # heteroskedastic - add white.adjust=TRUE
+plot(aovBRvet, 2) # fairly normal?
+leveneTest(BR ~ Island, data = vet20init) # heteroskedastic - add white.adjust=TRUE
 
@@
 plot(aovGIvet, 2) # fairly normally distributed?
-leveneTest(for.bites ~ Island, data = vet) # heteroskedastic - add white.adjust=TRUE
+leveneTest(for.bites ~ Island, data = vet20init) # heteroskedastic - add white.adjust=TRUE
 
@@
 plot(aovGIvir, 2) # fairly normally distributed?
-leveneTest(for.bites ~ Island, data = vir) # heteroskedastic - add white.adjust=TRUE
+leveneTest(for.bites ~ Island, data = vir20init) # heteroskedastic - add white.adjust=TRUE
 ```
@@
 
+# Competition
+
+```{r,echo=FALSE,warning=FALSE}
+ggplot(site.sum,aes(x=scar.BM,y=int.scar.mean_stop, shape=Island))+geom_point(size=3)+ggtitle("S.
+viride interaction frequency as a function of scarid biomass") + theme_minimal() + theme(plot.title = element_text(hjust = 0.5)) + labs(x="Scarid Biomass (g/100m2)", y="Interactions with other scarids (per 2 min follow)") + geom_errorbar(aes(ymin=int.scar.mean_stop-int.scar.se_stop, ymax=int.scar.mean_stop+int.scar.se_stop))
+ggsave("comp_vir.png")
+```
+
+```{r,echo=FALSE,warning=FALSE}
+ggplot(site.sum,aes(x=scar.BM,y=int.scar.mean_qup, shape=Island))+geom_point(size=3)+ggtitle("S.
+vetula interaction frequency as a function of scarid biomass") + theme_minimal() + theme(plot.title = element_text(hjust = 0.5)) + labs(x="Scarid Biomass (g/100m2)", y="Interactions with other scarids (per 2 min follow)") + geom_errorbar(aes(ymin=int.scar.mean_qup-int.scar.se_qup, ymax=int.scar.mean_qup+int.scar.se_qup))
+ggsave("comp_vet.png")
+```
+
 # GAMMs
@@
 ```{r,message=FALSE,warning=FALSE}
-vetinit <- vet %>% filter(Phase=="i")
-vetterm <- vet %>% filter(Phase=="t")
+vetinit <- data2 %>% filter(Species.Code=="qup" & Phase=="i")
 gamm<-gamm(FR~s(Length.cm)+s(scar.BM)+s(PC1)+s(PC2), family=gaussian(link=identity),data=vetinit, random=list(Island=~1),method="REML")
-summary(gamm$gam) # only PC1 and PC2 are significant, scar.BM is almost significant
+summary(gamm$gam) # only PC1 and PC2 are significant, length slightly, scar.BM is not
 AIC(gamm$lme)
@@
 summary(gamm$gam) # PC1 and PC2 are most significant, length is significant, scar.BM is almost significant
+AIC(gamm$lme)
 plot(gamm$gam,pages=1)
@@
 summary(gamm$gam) # very low R2, only PC1 and scar.BM significant
+AIC(gamm$lme)
 plot(gamm$gam,pages=1)
@@
 ```{r,message=FALSE,warning=FALSE}
-virinit<-vir%>% filter(Phase=="i")
-gamm<-gamm(FR~Phase+s(Length.cm)+s(scar.BM)+s(PC1)+s(PC2), family=gaussian(link=identity),data=vir, random=list(Island=~1),method="REML")
-summary(gamm$gam)
+virinit <- data2 %>% filter(Species.Code=="stop" & Phase=="i")
+gamm<-gamm(FR~s(Length.cm)+s(scar.BM)+s(PC1)+s(PC2), family=gaussian(link=identity),data=virinit, random=list(Island=~1),method="REML")
+summary(gamm$gam) #PC1, length, and scar.BM are significant predictors
+AIC(gamm$lme)
 plot(gamm$gam,pages=1)
@@
 ```{r,message=FALSE,warning=FALSE}
-gamm<-gamm(g.frac~s(Length.cm)+s(scar.BM)+s(PC1)+s(PC2), family=gaussian(link=identity),data=virinit, random=list(Island=~1),method="REML")
-summary(gamm$gam)
+gamm<-gamm(g.frac~s(Length.cm)+s(scar.BM)+s(PC1), family=gaussian(link=identity),data=virinit, random=list(Island=~1),method="REML")
+summary(gamm$gam) #PC1, length, and scar.BM are significant predictors
+AIC(gamm$lme)
 plot(gamm$gam,pages=1)
@@
 ```{r,message=FALSE,warning=FALSE}
-gamm<-gamm(for.bites~s(Length.cm)+s(scar.BM)+s(PC1)+s(PC2), family=gaussian(link=identity),data=vir, random=list(Island=~1),method="REML")
-summary(gamm$gam)
+gamm<-gamm(for.bites~s(Length.cm)+s(scar.BM)+s(PC1), family=gaussian(link=identity),data=virinit, random=list(Island=~1),method="REML")
+summary(gamm$gam) # no significant predictors (scar.BM almost), low r2
+AIC(gamm$lme)
 plot(gamm$gam,pages=1)

### 2018-04-10 data/SEM.Rmd
4/10

---
title: 'Structural '
author: "Molly Wilson"
date: "3/14/2018"
output:
  html_document:
    number_sections: yes
    theme: united
    toc: yes
    toc_float: yes
    toc_depth: 3
  pdf_document:
    toc: yes
---

```{r,message=FALSE,warning=FALSE, echo=FALSE}
# set up
library(rmarkdown)
library(dplyr)
library(tidyr)
library(lavaan)
```

```{r,message=FALSE,warning=FALSE, echo=FALSE}
setwd("~/github/Scarid-Behavior/data")
site.sum <- read.csv("site.sum.csv", stringsAsFactors=FALSE) %>% select(-X)
data <- read.csv("data2.csv", stringsAsFactors=FALSE)
# add aglal indeces
data <- data %>% mutate(index_MA = cover_MA*canopy_MA) %>% mutate(index_TA = cover_TA*canopy_TA)
# remove Pallaster ?
```

```{r}
# define model pathways
model <- '
    FR ~ Length.cm + Phase + Species + int.scar + RISK + NUTRITION
    int.scar ~ scar.BM + cover_MA + cover_TA
    rugosity ~ cover_LC
    cover_LC ~ cover_MA + cover_TA
    cover_TA ~ scar.BM
    cover_MA ~ scar.BM

    RISK =~ scar.BM + rugosity
    NUTRITION =~ cover_MA + cover_TA
'

fit <- sem(model, data=data)
varTable(fit)
summary(fit)
semPaths(fit)
```
```{r}
# define model pathways
model <- '
    FR ~ Length.cm + Phase + Species + int.scar

'

fit <- sem(model, data=data)
summary(fit)
semPaths(fit)
```

### 2018-04-21 SummaryOrals.Rmd
4/10
4/20

@@
 site.sum <- read.csv("site.sum.csv", stringsAsFactors=FALSE) %>% select(-X)
-data <- read.csv("data2.csv", stringsAsFactors=FALSE)
+data <- read.csv("data2.csv", stringsAsFactors=FALSE) %>% filter(!(Species.Code=="qup" & Site=="Turtle Bay")) %>% filter(!(Species.Code=="qup" & Site=="Pallaster East")) # remove due to low sample size
 
@@
 
-# Grazing differences between islands
+# Grazing differences among islands
 
@@
 ggplot(vetvir20init, aes(x=Island, y=FR)) + geom_boxplot() +
-stat_summary(fun.y = mean, geom="point", shape=5, size=3) + labs(y="Feeding Rate (bites/hr)",x="",title="Feeding Rate by Island and Species (Initial Phase)") + theme_bw() + theme(plot.title = element_text(hjust = 0.5)) + geom_signif(comparisons = list(c("Antigua", "Barbuda"),c("Antigua","Bonaire"),c("Barbuda","Bonaire")), map_signif_level=TRUE) + facet_grid(.~Species)
+stat_summary(fun.y = mean, geom="point", shape=5, size=3) + labs(y="Feeding Rate (bites/hr)",x="",title="Feeding Rate") + theme_bw() + theme(plot.title = element_text(hjust = 0.5)) + geom_signif(comparisons = list(c("Antigua", "Barbuda"),c("Antigua","Bonaire"),c("Barbuda","Bonaire")), map_signif_level=TRUE) + facet_grid(.~Species)
 ggsave("bp_fr.png")
@@
 
+```{r}
+vetAB <- vetvirAB %>% filter(Species.Code=="qup")
+vet_fr <- lm(FR~Length.cm+Island,data=vetAB)
+anova(vet_fr)
+
+virAB <- vetvirAB %>% filter(Species.Code=="stop")
+vir_fr <- lm(FR~Length.cm+Island,data=virAB)
+anova(vir_fr)
+
+```
+
 ```{r, message=FALSE,warning=FALSE,echo=FALSE}
@@
 ggplot(vetvir20init,aes(x=scar.BM,y=for.bites))+facet_grid(.~Species)+geom_point(shape=18)+geom_smooth(method=loess, fullrange=FALSE)+ggtitle("Grazing intensity by scarid biomass") + theme_minimal() +labs(y="Bites/foray",x="Scarid Biomass (g/100m2)") + theme(plot.title = element_text(hjust = 0.5))
+```
+# Comparisons with published data
+```{r}
+# calculate predicted feeding rates from Bruggemann/Bozec:
+FR.vir15=1089-17*15-56
+FR.vir20=1089-17*20-56
+FR.vir25=1089-17*25-56
+
+FR.vet15=3329-33*15
+FR.vet20=3329-33*20
+FR.vet25=3329-33*25
+
+# Antigua feeding rate equation:
+ANUint.vir <- data2 %>% filter(Island=="Antigua"&Phase=="i"&Species.Code=="stop")
+ANUint.vet <- data2 %>% filter(Island=="Antigua"&Phase=="i"&Species.Code=="qup")
+FR.anu.vir <- lm(FR~Length.cm, data=ANUint.vir)
+FR.anu.vet <- lm(FR~Length.cm, data=ANUint.vet)
+summary(FR.anu.vet)
+#FR.vet = 1047-12*L
+summary(FR.anu.vir)
+#FR.vir = 445-8*L
+
+FR.vet15ANU=1047-12*15
+FR.vet20ANU=1047-12*20
+FR.vet25ANU=1047-12*25
+
+FR.vir15ANU=445-8*15
+FR.vir20ANU=445-8*20
+FR.vir25ANU=445-8*25
+
+# calculate predicted bite sizes from Bruggemann/Bozec
+
+BS.vir15=5.257*10^-4*15^2
+BS.vir20=5.257*10^-4*20^2
+BS.vir25=5.257*10^-4*25^2
+
+BS.vet15=4.013*10^-4*15^2
+BS.vet20=4.013*10^-4*20^2
+BS.vet25=4.013*10^-4*25^2
+
+# adding FR and BS for each fish
+feedingrates.vir.ANU <- data2 %>% filter(Species.Code=="stop"&Phase=="i") %>% mutate(FR.ANU = 445-8*Length.cm) %>% mutate(FR.pub=1089-17*Length.cm-56) %>% mutate(BS=5.257*10^-4*Length.cm^2)
+feedingrates.vet.ANU <- data2 %>% filter(Species.Code=="qup"&Phase=="i") %>% mutate(FR.ANU = 1047-12*Length.cm) %>% mutate(FR.pub=3329-33*Length.cm) %>% mutate(BS=4.013*10^-4*Length.cm^2)
+feedingrates <- bind_rows(feedingrates.vir.ANU,feedingrates.vet.ANU) %>% mutate(areagrazed.ANU=FR.ANU*BS) %>% mutate(areagrazed.pub=FR.pub*BS)
+
+areagrazed.site <- feedingrates %>% group_by(Site) %>% summarise(areagrazed.site.ANU=sum(areagrazed.ANU))
 ```

### 2018-04-21 SummaryOrals.Rmd
4/20

@@
 library(car)
+library(reshape2)
 ```
@@
 # Antigua feeding rate equation:
-ANUint.vir <- data2 %>% filter(Island=="Antigua"&Phase=="i"&Species.Code=="stop")
-ANUint.vet <- data2 %>% filter(Island=="Antigua"&Phase=="i"&Species.Code=="qup")
+ANUint.vir <- data %>% filter(Island=="Antigua"&Phase=="i"&Species.Code=="stop")
+ANUint.vet <- data %>% filter(Island=="Antigua"&Phase=="i"&Species.Code=="qup")
 FR.anu.vir <- lm(FR~Length.cm, data=ANUint.vir)
@@
 # adding FR and BS for each fish
-feedingrates.vir.ANU <- data2 %>% filter(Species.Code=="stop"&Phase=="i") %>% mutate(FR.ANU = 445-8*Length.cm) %>% mutate(FR.pub=1089-17*Length.cm-56) %>% mutate(BS=5.257*10^-4*Length.cm^2)
-feedingrates.vet.ANU <- data2 %>% filter(Species.Code=="qup"&Phase=="i") %>% mutate(FR.ANU = 1047-12*Length.cm) %>% mutate(FR.pub=3329-33*Length.cm) %>% mutate(BS=4.013*10^-4*Length.cm^2)
+feedingrates.vir.ANU <- data %>% filter(Species.Code=="stop"&Phase=="i"&Island=="Antigua") %>% mutate(FR.ANU = 445-8*Length.cm) %>% mutate(FR.pub=1089-17*Length.cm-56) %>% mutate(BS=5.257*10^-4*Length.cm^2)
+feedingrates.vet.ANU <- data %>% filter(Species.Code=="qup"&Phase=="i"&Island=="Antigua") %>% mutate(FR.ANU=1047-12*Length.cm) %>% mutate(FR.pub=3329-33*Length.cm) %>% mutate(BS=4.013*10^-4*Length.cm^2)
+
 feedingrates <- bind_rows(feedingrates.vir.ANU,feedingrates.vet.ANU) %>% mutate(areagrazed.ANU=FR.ANU*BS) %>% mutate(areagrazed.pub=FR.pub*BS)
 
-areagrazed.site <- feedingrates %>% group_by(Site) %>% summarise(areagrazed.site.ANU=sum(areagrazed.ANU))
+areagrazed.site <- feedingrates %>% group_by(Site) %>% dplyr::summarize(observed=sum(areagrazed.ANU)/1000,predicted=sum(areagrazed.pub)/1000)
+
+areagrazed.site <- as.data.frame(areagrazed.site)
+AG <- melt(areagrazed.site, id="Site") %>% dplyr::rename("feeding.rates"=variable, "area.grazed"=value)
+```
+```{r}
+ggplot(data=AG, aes(x=Site, y=area.grazed, fill=feeding.rates)) +
+    geom_bar(stat="identity", position=position_dodge()) + theme_bw() + scale_fill_manual(values=c("black","lightslategrey")) + labs(y="Area Grazed (m2)")
 ```

### 2018-04-24 SummaryOrals.Rmd
4/24

@@
 vetvir <- data %>% filter(Species.Code!="rbp")
+vetvirinit <- vetvir %>% filter(Phase=="i")
 vetvir20 <- vetvir %>% filter(Length.cm>=15 & Length.cm<=25)
@@
 ```{r}
-predictors <- site.sum %>% select(cover_LC,cover_TA,canopy_MA,canopy_TA,rugosity,scar.BM,scar.DEN) # should it matter if I use site level summary data vs.
-full follow dataset (where site-level data has been joined?)
+predictors <- site.sum %>% select(cover_LC,cover_TA,canopy_MA,canopy_TA,rugosity) # should it matter if I use site level summary data vs. full follow dataset (where site-level data has been joined?)
 pca <- prcomp(na.omit(predictors),center = TRUE,scale. = TRUE) 
@@
 ggbiplot(pca, obs.scale = 1, var.scale = 1, groups=site.sum$Island, ellipse = TRUE, circle = TRUE, varname.size = 2) + scale_color_manual(name="Island", values=c("navy", "darkseagreen", "slategray2")) + theme(legend.direction = 'horizontal', legend.position = 'top') + theme_minimal()
-ggsave("pca_pred.png")
+ggsave("pca_benth.png")
 ```
@@
 vetvir20init<-vetvir20 %>% filter(Phase=="i")
+
+samples <- vetvir20init %>% dplyr::group_by(Island,Species) %>% dplyr::summarise(n=n())
 ggplot(vetvir20init, aes(x=Island, y=FR)) + geom_boxplot() +
-stat_summary(fun.y = mean, geom="point", shape=5, size=3) + labs(y="Feeding Rate (bites/hr)",x="",title="Feeding Rate") + theme_bw() + theme(plot.title = element_text(hjust = 0.5)) + geom_signif(comparisons = list(c("Antigua", "Barbuda"),c("Antigua","Bonaire"),c("Barbuda","Bonaire")), map_signif_level=TRUE) + facet_grid(.~Species)
+stat_summary(fun.y = mean, geom="point", shape=5, size=3) + labs(y="Feeding Rate (bites/hr)",x="",title="Feeding Rate") + theme_bw() + theme(plot.title = element_text(hjust = 0.5)) + geom_signif(comparisons = list(c("Antigua", "Barbuda"),c("Antigua","Bonaire"),c("Barbuda","Bonaire")), map_signif_level=TRUE) + facet_grid(.~Species) + theme(strip.text = element_text(face = "italic"))
 ggsave("bp_fr.png")
@@
 ggplot(vetvir20init, aes(x=Island, y=BR)) + geom_boxplot() +
-stat_summary(fun.y = mean, geom="point", shape=5, size=3) + labs(y="Bite Rate (bites/sec)",x="",title="Bite Rate while Grazing by Island and Species") + theme_bw() + theme(plot.title = element_text(hjust = 0.5)) + geom_signif(comparisons = list(c("Antigua", "Barbuda"),c("Antigua","Bonaire"),c("Barbuda","Bonaire")), map_signif_level=TRUE) + facet_grid(.~Species)
+stat_summary(fun.y = mean, geom="point", shape=5, size=3) + labs(y="Bite Rate (bites/sec)",x="",title="Bite Rate while Grazing by Island and Species") + theme_bw() + theme(plot.title = element_text(hjust = 0.5)) + geom_signif(comparisons = list(c("Antigua", "Barbuda"),c("Antigua","Bonaire"),c("Barbuda","Bonaire")), map_signif_level=TRUE) + facet_grid(.~Species) + theme(strip.text = element_text(face = "italic"))
 ggsave("bp_br.png")
@@
 ggplot(vetvir20init, aes(x=Island, y=g.frac)) + geom_boxplot() +
-stat_summary(fun.y = mean, geom="point", shape=5, size=3) + labs(y="Time fraction",x="",title="Fraction of Time Spent Grazing by Island and Species") + theme_bw() + theme(plot.title = element_text(hjust = 0.5)) + geom_signif(comparisons = list(c("Antigua", "Barbuda"),c("Antigua","Bonaire"),c("Barbuda","Bonaire")), map_signif_level=TRUE) + facet_grid(.~Species)
+stat_summary(fun.y = mean, geom="point", shape=5, size=3) + labs(y="Time fraction",x="",title="Fraction of Time Spent Grazing by Island and Species") + theme_bw() + theme(plot.title = element_text(hjust = 0.5)) + geom_signif(comparisons = list(c("Antigua", "Barbuda"),c("Antigua","Bonaire"),c("Barbuda","Bonaire")), map_signif_level=TRUE) + facet_grid(.~Species) + theme(strip.text = element_text(face = "italic"))
 ggsave("bp_gfrac.png")
@@
 ```{r, message=FALSE,warning=FALSE,echo=FALSE}
-ggplot(vetvir20init, aes(x=Island, y=for.bites)) + geom_boxplot()  + labs(y="Bites per foray",x="",title="Grazing intensity by Island and Species") + theme_bw() + theme(plot.title = element_text(hjust = 0.5)) + geom_signif(comparisons = list(c("Antigua", "Barbuda"),c("Antigua","Bonaire"),c("Barbuda","Bonaire")), map_signif_level=TRUE) + facet_grid(.~Species)
+ggplot(vetvir20init, aes(x=Island, y=for.bites)) + geom_boxplot()  + labs(y="Bites per foray",x="",title="Grazing intensity by Island and Species") + theme_bw() + theme(plot.title = element_text(hjust = 0.5)) + geom_signif(comparisons = list(c("Antigua", "Barbuda"),c("Antigua","Bonaire"),c("Barbuda","Bonaire")), map_signif_level=TRUE) + facet_grid(.~Species) + theme(strip.text = element_text(face = "italic"))
 ggsave("bp_gi.png")
@@
 
-ggplot(vetvirAB,aes(x=Length.cm,y=FR,colour=Island))+facet_grid(.~Species)+geom_point(shape=18)+scale_shape(solid=FALSE)+geom_smooth(method=lm, fullrange=FALSE)+ggtitle("Feeding rates by fish length") + theme_minimal() +labs(y="Feeding rates (bites/hr)",x="Length (cm)") + theme(plot.title = element_text(hjust = 0.5))  + scale_color_manual(values=c("navy", "slategray2"))
+ggplot(vetvirAB,aes(x=Length.cm,y=FR,colour=Island))+facet_grid(.~Species)+geom_point(shape=18)+scale_shape(solid=FALSE)+geom_smooth(method=lm, fullrange=FALSE)+ggtitle("Feeding rates by fish length") + theme_minimal() +labs(y="Feeding rates (bites/hr)",x="Length (cm)") + theme(plot.title = element_text(hjust = 0.5))  + scale_color_manual(values=c("navy", "slategray2")) + theme(strip.text = element_text(face = "italic"))
 ggsave("sp_length.png")
@@
 ```{r, message=FALSE,warning=FALSE,echo=FALSE}
-ggplot(vetvirABinit,aes(x=Length.cm,y=BR,colour=Island))+facet_grid(.~Species)+geom_point(shape=18)+scale_shape(solid=FALSE)+geom_smooth(method=lm, fullrange=FALSE)+ggtitle("Bite rate by fish length") + theme_minimal() +labs(y="Bite Rate (bites/sec)",x="Length (cm)") + theme(plot.title = element_text(hjust = 0.5))  + scale_color_manual(values=c("navy", "slategray2"))
+ggplot(vetvirABinit,aes(x=Length.cm,y=BR,colour=Island))+facet_grid(.~Species)+geom_point(shape=18)+scale_shape(solid=FALSE)+geom_smooth(method=lm, fullrange=FALSE)+ggtitle("Bite rate by fish length") + theme_minimal() +labs(y="Bite Rate (bites/sec)",x="Length (cm)") + theme(plot.title = element_text(hjust = 0.5))  + scale_color_manual(values=c("navy", "slategray2")) + theme(strip.text = element_text(face = "italic"))
 ```
@@
 ```{r, message=FALSE,warning=FALSE,echo=FALSE}
-ggplot(vetvirABinit,aes(x=Length.cm,y=g.frac,colour=Island))+facet_grid(.~Species)+geom_point(shape=18)+scale_shape(solid=FALSE)+geom_smooth(method=lm, fullrange=FALSE)+ggtitle("Fraction of time spent grazing by fish length") + theme_minimal() +labs(y="Time fraction",x="Length (cm)") + theme(plot.title = element_text(hjust = 0.5))  + scale_color_manual(values=c("navy", "slategray2"))
+ggplot(vetvirABinit,aes(x=Length.cm,y=g.frac,colour=Island))+facet_grid(.~Species)+geom_point(shape=18)+scale_shape(solid=FALSE)+geom_smooth(method=lm, fullrange=FALSE)+ggtitle("Fraction of time spent grazing by fish length") + theme_minimal() +labs(y="Time fraction",x="Length (cm)") + theme(plot.title = element_text(hjust = 0.5))  + scale_color_manual(values=c("navy", "slategray2")) + theme(strip.text = element_text(face = "italic"))
 ```
@@
 ```{r, message=FALSE,warning=FALSE,echo=FALSE}
-ggplot(vetvirABinit,aes(x=Length.cm,y=for.bites,colour=Island))+facet_grid(.~Species)+geom_point(shape=18)+scale_shape(solid=FALSE)+geom_smooth(method=lm, fullrange=FALSE)+ggtitle("Grazing intensity by fish length") + theme_minimal() +labs(y="Bites/foray",x="Length (cm)") + theme(plot.title = element_text(hjust = 0.5))  + scale_color_manual(values=c("navy", "slategray2"))
+ggplot(vetvirABinit,aes(x=Length.cm,y=for.bites,colour=Island))+facet_grid(.~Species)+geom_point(shape=18)+scale_shape(solid=FALSE)+geom_smooth(method=lm, fullrange=FALSE)+ggtitle("Grazing intensity by fish length") + theme_minimal() +labs(y="Bites/foray",x="Length (cm)") + theme(plot.title = element_text(hjust = 0.5))  + scale_color_manual(values=c("navy", "slategray2")) + theme(strip.text = element_text(face = "italic"))
 ```
@@
 ```{r, message=FALSE,warning=FALSE,echo=FALSE}
-ggplot(vetvir20init,aes(x=scar.BM,y=FR))+facet_grid(.~Species)+geom_point(shape=18)+geom_smooth(method=loess, fullrange=FALSE)+ggtitle("Feeding rate by scarid biomass") + theme_minimal() +labs(y="Feeding rate (bites/hr)",x="Scarid Biomass (g/100m2)") + theme(plot.title = element_text(hjust = 0.5))
+ggplot(vetvir20init,aes(x=scar.BM,y=FR))+facet_grid(.~Species)+geom_point(shape=18)+geom_smooth(method=loess, fullrange=FALSE)+ggtitle("Feeding rate by scarid biomass") + theme_minimal() +labs(y="Feeding rate (bites/hr)",x="Scarid Biomass (g/100m2)") + theme(plot.title = element_text(hjust = 0.5)) + theme(strip.text = element_text(face = "italic"))
 ggsave("sp_scarbm.png")
@@
 ```{r, message=FALSE,warning=FALSE,echo=FALSE}
-ggplot(vetvir20init,aes(x=scar.BM,y=BR))+facet_grid(.~Species)+geom_point(shape=18)+geom_smooth(method=loess, fullrange=FALSE)+ggtitle("Bite rate while grazing by scarid biomass") + theme_minimal() +labs(y="Bite rate (bites/sec)",x="Scarid Biomass (g/100m2)") + theme(plot.title = element_text(hjust = 0.5))
+ggplot(vetvir20init,aes(x=scar.BM,y=BR))+facet_grid(.~Species)+geom_point(shape=18)+geom_smooth(method=loess, fullrange=FALSE)+ggtitle("Bite rate while grazing by scarid biomass") + theme_minimal() +labs(y="Bite rate (bites/sec)",x="Scarid Biomass (g/100m2)") + theme(plot.title = element_text(hjust = 0.5)) + theme(strip.text = element_text(face = "italic"))
 ```
@@
 ```{r, message=FALSE,warning=FALSE,echo=FALSE}
-ggplot(vetvir20init,aes(x=scar.BM,y=g.frac))+facet_grid(.~Species)+geom_point(shape=18)+geom_smooth(method=loess, fullrange=FALSE)+ggtitle("Fraction of time spent grazing by scarid biomass") + theme_minimal() +labs(y="Time fraction",x="Scarid Biomass (g/100m2)") + theme(plot.title = element_text(hjust = 0.5))
+ggplot(vetvir20init,aes(x=scar.BM,y=g.frac))+facet_grid(.~Species)+geom_point(shape=18)+geom_smooth(method=loess, fullrange=FALSE)+ggtitle("Fraction of time spent grazing by scarid biomass") + theme_minimal() +labs(y="Time fraction",x="Scarid Biomass (g/100m2)") + theme(plot.title = element_text(hjust = 0.5)) + theme(strip.text = element_text(face = "italic"))
 ```
@@
 ```{r, message=FALSE,warning=FALSE,echo=FALSE}
-ggplot(vetvir20init,aes(x=scar.BM,y=for.bites))+facet_grid(.~Species)+geom_point(shape=18)+geom_smooth(method=loess, fullrange=FALSE)+ggtitle("Grazing intensity by scarid biomass") + theme_minimal() +labs(y="Bites/foray",x="Scarid Biomass (g/100m2)") + theme(plot.title = element_text(hjust = 0.5))
+ggplot(vetvir20init,aes(x=scar.BM,y=for.bites))+facet_grid(.~Species)+geom_point(shape=18)+geom_smooth(method=loess, fullrange=FALSE)+ggtitle("Grazing intensity by scarid biomass") + theme_minimal() +labs(y="Bites/foray",x="Scarid Biomass (g/100m2)") + theme(plot.title = element_text(hjust = 0.5)) + theme(strip.text = element_text(face = "italic"))
 ```
@@
 ```{r}
-# calculate predicted feeding rates from Bruggemann/Bozec:
-FR.vir15=1089-17*15-56
+# calculate predicted feeding rates from Bruggemann/Bozec for boxplot reference lines:
 FR.vir20=1089-17*20-56
-FR.vir25=1089-17*25-56
+FR.vet20=3329-33*20
+```
+```{r}
+# Antigua feeding rate equation:
+ANUinit.vir <- data %>% filter(Island=="Antigua"&Phase=="i"&Species.Code=="stop")
+ANUinit.vet <- data %>% filter(Island=="Antigua"&Phase=="i"&Species.Code=="qup")
+FR.anu.vir <- lm(FR~Length.cm, data=ANUinit.vir)
+FR.anu.vet <- lm(FR~Length.cm, data=ANUinit.vet)
+summary(FR.anu.vet)
+#FR.anu.vet = 1047-12*L
+summary(FR.anu.vir)
+#FR.anu.vir = 445-8*L
 
-FR.vet15=3329-33*15
-FR.vet20=3329-33*20
-FR.vet25=3329-33*25
+# references (20cm, initial phase)
+FR.vet20ANU=1047-12*20
+FR.vir20ANU=445-8*20
 
-# Antigua feeding rate equation:
-ANUint.vir <- data %>% filter(Island=="Antigua"&Phase=="i"&Species.Code=="stop")
-ANUint.vet <- data %>% filter(Island=="Antigua"&Phase=="i"&Species.Code=="qup")
-FR.anu.vir <- lm(FR~Length.cm, data=ANUint.vir)
-FR.anu.vet <- lm(FR~Length.cm, data=ANUint.vet)
-summary(FR.anu.vet)
-#FR.vet = 1047-12*L
-summary(FR.anu.vir)
-#FR.vir = 445-8*L
+```
+```{r}
+# Bonaire feeding rate equation:
+BONinit.vir <- data %>% filter(Island=="Bonaire"&Phase=="i"&Species.Code=="stop")
+BONinit.vet <- data %>% filter(Island=="Bonaire"&Phase=="i"&Species.Code=="qup")
+FR.bon.vir <- lm(FR~Length.cm, data=BONinit.vir)
+FR.bon.vet <- lm(FR~Length.cm, data=BONinit.vet)
 
-FR.vet15ANU=1047-12*15
-FR.vet20ANU=1047-12*20
-FR.vet25ANU=1047-12*25
+summary(FR.bon.vet)
+#FR.bon.vet = 1347+10*L
+summary(FR.bon.vir)
+#FR.bon.vir = 1057-15*L
 
-FR.vir15ANU=445-8*15
-FR.vir20ANU=445-8*20
-FR.vir25ANU=445-8*25
+# references (20cm, initial phase)
+FR.vet20BON=1347+10*20
+FR.vir20BON=1057-15*20
+```
 
+```{r}
 # calculate predicted bite sizes from Bruggemann/Bozec
-
-BS.vir15=5.257*10^-4*15^2
 BS.vir20=5.257*10^-4*20^2
-BS.vir25=5.257*10^-4*25^2
-
-BS.vet15=4.013*10^-4*15^2
 BS.vet20=4.013*10^-4*20^2
-BS.vet25=4.013*10^-4*25^2
 
@@
 
-areagrazed.site <- feedingrates %>% group_by(Site) %>% dplyr::summarize(observed=sum(areagrazed.ANU)/1000,predicted=sum(areagrazed.pub)/1000)
+areagrazed.site <- feedingrates %>% group_by(Site) %>% dplyr::summarize(personal=sum(areagrazed.ANU)/1000,published=sum(areagrazed.pub)/1000)
 
@@
 ggplot(data=AG, aes(x=Site, y=area.grazed, fill=feeding.rates)) +
-    geom_bar(stat="identity", position=position_dodge()) + theme_bw() + scale_fill_manual(values=c("black","lightslategrey")) + labs(y="Area Grazed (m2)")
+    geom_bar(stat="identity", position=position_dodge()) + theme_bw() + scale_fill_manual(values=c("black","lightslategrey"),name="Data Source",breaks=c("personal", "published"), labels=c("Fished", "Unfished")) + labs(y=expression(Area~Grazed~(m^2/hr)), title=("Grazing impact of Antiguan scarids with observed vs.
+predicted feeding rates"))
+ggsave("areagrazed.png")
 ```
@@
 ```{r,echo=FALSE,warning=FALSE}
-ggplot(site.sum,aes(x=scar.BM,y=int.scar.mean_stop, shape=Island))+geom_point(size=3)+ggtitle("S.
-viride interaction frequency as a function of scarid biomass") + theme_minimal() + theme(plot.title = element_text(hjust = 0.5)) + labs(x="Scarid Biomass (g/100m2)", y="Interactions with other scarids (per 2 min follow)") + geom_errorbar(aes(ymin=int.scar.mean_stop-int.scar.se_stop, ymax=int.scar.mean_stop+int.scar.se_stop))
+ggplot(site.sum,aes(x=scar.BM,y=int.scar.mean_stop, shape=Island))+geom_point(size=3)+ggtitle(expression(paste(italic("S.
+viride"), " interaction frequency as a function of scarid biomass"))) + theme_minimal() + theme(plot.title = element_text(hjust = 0.5)) + labs(x="Scarid Biomass (g/100m2)", y="Interactions with other scarids (per 2 min follow)") + geom_errorbar(aes(ymin=int.scar.mean_stop-int.scar.se_stop, ymax=int.scar.mean_stop+int.scar.se_stop))
 ggsave("comp_vir.png")
@@
 ```{r,echo=FALSE,warning=FALSE}
-ggplot(site.sum,aes(x=scar.BM,y=int.scar.mean_qup, shape=Island))+geom_point(size=3)+ggtitle("S.
-vetula interaction frequency as a function of scarid biomass") + theme_minimal() + theme(plot.title = element_text(hjust = 0.5)) + labs(x="Scarid Biomass (g/100m2)", y="Interactions with other scarids (per 2 min follow)") + geom_errorbar(aes(ymin=int.scar.mean_qup-int.scar.se_qup, ymax=int.scar.mean_qup+int.scar.se_qup))
+ggplot(site.sum,aes(x=scar.BM,y=int.scar.mean_qup, shape=Island))+geom_point(size=3)+ggtitle(expression(paste(italic("S.
+vetula"), " interaction frequency as a function of scarid biomass"))) + theme_minimal() + theme(plot.title = element_text(hjust = 0.5)) + labs(x="Scarid Biomass (g/100m2)", y="Interactions with other scarids (per 2 min follow)") + geom_errorbar(aes(ymin=int.scar.mean_qup-int.scar.se_qup, ymax=int.scar.mean_qup+int.scar.se_qup))
 ggsave("comp_vet.png")

### 2018-04-26 SummaryOrals.Rmd
4/26

@@
 library(reshape2)
+library(EnvStats)
 ```
@@
 vetvir20init<-vetvir20 %>% filter(Phase=="i")
+samples <- vetvir20init %>% dplyr::group_by(Island,Species) %>% dplyr::summarise(n=n())
 
-samples <- vetvir20init %>% dplyr::group_by(Island,Species) %>% dplyr::summarise(n=n())
 ggplot(vetvir20init, aes(x=Island, y=FR)) + geom_boxplot() +
-stat_summary(fun.y = mean, geom="point", shape=5, size=3) + labs(y="Feeding Rate (bites/hr)",x="",title="Feeding Rate") + theme_bw() + theme(plot.title = element_text(hjust = 0.5)) + geom_signif(comparisons = list(c("Antigua", "Barbuda"),c("Antigua","Bonaire"),c("Barbuda","Bonaire")), map_signif_level=TRUE) + facet_grid(.~Species) + theme(strip.text = element_text(face = "italic"))
+stat_summary(fun.y = mean, geom="point", shape=5, size=3) + labs(y="Feeding Rate (bites/hr)",x="",title="Feeding rates by island and species") + theme_bw() + theme(plot.title = element_text(hjust = 0.5)) + geom_signif(comparisons = list(c("Antigua", "Barbuda"),c("Antigua","Bonaire"),c("Barbuda","Bonaire")), map_signif_level=TRUE) + facet_grid(.~Species) + theme(strip.text = element_text(face = "italic")) + stat_n_text()
 ggsave("bp_fr.png")
@@
 ggplot(vetvir20init, aes(x=Island, y=BR)) + geom_boxplot() +
-stat_summary(fun.y = mean, geom="point", shape=5, size=3) + labs(y="Bite Rate (bites/sec)",x="",title="Bite Rate while Grazing by Island and Species") + theme_bw() + theme(plot.title = element_text(hjust = 0.5)) + geom_signif(comparisons = list(c("Antigua", "Barbuda"),c("Antigua","Bonaire"),c("Barbuda","Bonaire")), map_signif_level=TRUE) + facet_grid(.~Species) + theme(strip.text = element_text(face = "italic"))
+stat_summary(fun.y = mean, geom="point", shape=5, size=3) + labs(y="Bite Rate (bites/sec)",x="",title="Bite rates by island and species") + theme_bw() + theme(plot.title = element_text(hjust = 0.5)) + geom_signif(comparisons = list(c("Antigua", "Barbuda"),c("Antigua","Bonaire"),c("Barbuda","Bonaire")), map_signif_level=TRUE) + facet_grid(.~Species) + theme(strip.text = element_text(face = "italic")) + stat_n_text()
 ggsave("bp_br.png")
@@
 ggplot(vetvir20init, aes(x=Island, y=g.frac)) + geom_boxplot() +
-stat_summary(fun.y = mean, geom="point", shape=5, size=3) + labs(y="Time fraction",x="",title="Fraction of Time Spent Grazing by Island and Species") + theme_bw() + theme(plot.title = element_text(hjust = 0.5)) + geom_signif(comparisons = list(c("Antigua", "Barbuda"),c("Antigua","Bonaire"),c("Barbuda","Bonaire")), map_signif_level=TRUE) + facet_grid(.~Species) + theme(strip.text = element_text(face = "italic"))
+stat_summary(fun.y = mean, geom="point", shape=5, size=3) + labs(y="Time fraction",x="",title="Fraction of time spent grazing by island and species") + theme_bw() + theme(plot.title = element_text(hjust = 0.5)) + geom_signif(comparisons = list(c("Antigua", "Barbuda"),c("Antigua","Bonaire"),c("Barbuda","Bonaire")), map_signif_level=TRUE) + facet_grid(.~Species) + theme(strip.text = element_text(face = "italic")) + stat_n_text()
 ggsave("bp_gfrac.png")
@@
 ```{r, message=FALSE,warning=FALSE,echo=FALSE}
-ggplot(vetvir20init, aes(x=Island, y=for.bites)) + geom_boxplot()  + labs(y="Bites per foray",x="",title="Grazing intensity by Island and Species") + theme_bw() + theme(plot.title = element_text(hjust = 0.5)) + geom_signif(comparisons = list(c("Antigua", "Barbuda"),c("Antigua","Bonaire"),c("Barbuda","Bonaire")), map_signif_level=TRUE) + facet_grid(.~Species) + theme(strip.text = element_text(face = "italic"))
+ggplot(vetvir20init, aes(x=Island, y=for.bites)) + geom_boxplot()  + labs(y="Bites per foray",x="",title="Grazing intensity by island and species") + theme_bw() + theme(plot.title = element_text(hjust = 0.5)) + geom_signif(comparisons = list(c("Antigua", "Barbuda"),c("Antigua","Bonaire"),c("Barbuda","Bonaire")), map_signif_level=TRUE) + facet_grid(.~Species) + theme(strip.text = element_text(face = "italic")) + stat_n_text()
 ggsave("bp_gi.png")
@@
 ```{r, message=FALSE,warning=FALSE,echo=FALSE}
-ggplot(vetvir20init,aes(x=scar.BM,y=FR))+facet_grid(.~Species)+geom_point(shape=18)+geom_smooth(method=loess, fullrange=FALSE)+ggtitle("Feeding rate by scarid biomass") + theme_minimal() +labs(y="Feeding rate (bites/hr)",x="Scarid Biomass (g/100m2)") + theme(plot.title = element_text(hjust = 0.5)) + theme(strip.text = element_text(face = "italic"))
+ggplot(vetvir20init,aes(x=scar.BM,y=FR))+facet_grid(.~Species)+geom_point(shape=18)+geom_smooth(method=loess, fullrange=FALSE)+ggtitle("Feeding rate by scarid biomass") + theme_minimal() +labs(y="Feeding rate (bites/hr)") + labs(x=expression(Scarid~Biomass~(g/100~m^2))) + theme(plot.title = element_text(hjust = 0.5)) + theme(strip.text = element_text(face = "italic"))
+
 ggsave("sp_scarbm.png")
@@
 # calculate predicted feeding rates from Bruggemann/Bozec for boxplot reference lines:
+FR.vet20=3329-33*20
 FR.vir20=1089-17*20-56
-FR.vet20=3329-33*20
 ```
+
 ```{r}
-# Antigua feeding rate equation:
+# Observed Antigua feeding rate equation:
 ANUinit.vir <- data %>% filter(Island=="Antigua"&Phase=="i"&Species.Code=="stop")
@@
 #FR.anu.vir = 445-8*L
-
-# references (20cm, initial phase)
-FR.vet20ANU=1047-12*20
-FR.vir20ANU=445-8*20
-
 ```
 ```{r}
-# Bonaire feeding rate equation:
+# Observed Bonaire feeding rate equation:
 BONinit.vir <- data %>% filter(Island=="Bonaire"&Phase=="i"&Species.Code=="stop")
@@
 #FR.bon.vir = 1057-15*L
-
-# references (20cm, initial phase)
-FR.vet20BON=1347+10*20
-FR.vir20BON=1057-15*20
 ```
 
+## adding predicted feeding rates and bite sizes from standard equations
+Bite sizes as predicted by Bruggemann/Bozec:
+*S. vetula*: BS.vet=4.013*10^-4*L^2
+*S. viride*: BS.vir=5.257*10^-4*L^2
+
+Feeding rates as predicted by Bruggemann/Bozec
+*S. vetula*: FR.vet=3329-33*L
+*S. viride*: FR.vir=1089-17*L-56 (extra term because phase modifier is not used here (i=1))
+
+Feeding rates as predicted by observed Bonaire data:
+*S. vetula*: FR.bon.vet = 1347+10*L
+*S. viride*: FR.bon.vir = 1057-15*L
+
+Feeding rates as predicted by observed Antigua data:
+*S. vetula*: FR.anu.vet = 1047-12*L
+*S. viride*: FR.anu.vir = 445-8*L
+
+## adding predicted feeding rates from observed relationships
+
 ```{r}
-# calculate predicted bite sizes from Bruggemann/Bozec
-BS.vir20=5.257*10^-4*20^2
-BS.vet20=4.013*10^-4*20^2
+# Antigua
+feedingrates.vet.ANU <- data %>% filter(Species.Code=="qup"&Phase=="i"&Island=="Antigua") %>% mutate(FR.personal=1047-12*Length.cm) %>% mutate(FR.standard=3329-33*Length.cm) %>% mutate(BS=4.013*10^-4*Length.cm^2)
+feedingrates.vir.ANU <- data %>% filter(Species.Code=="stop"&Phase=="i"&Island=="Antigua") %>% mutate(FR.personal = 445-8*Length.cm) %>% mutate(FR.standard=1089-17*Length.cm-56) %>% mutate(BS=5.257*10^-4*Length.cm^2)
 
-# adding FR and BS for each fish
-feedingrates.vir.ANU <- data %>% filter(Species.Code=="stop"&Phase=="i"&Island=="Antigua") %>% mutate(FR.ANU = 445-8*Length.cm) %>% mutate(FR.pub=1089-17*Length.cm-56) %>% mutate(BS=5.257*10^-4*Length.cm^2)
-feedingrates.vet.ANU <- data %>% filter(Species.Code=="qup"&Phase=="i"&Island=="Antigua") %>% mutate(FR.ANU=1047-12*Length.cm) %>% mutate(FR.pub=3329-33*Length.cm) %>% mutate(BS=4.013*10^-4*Length.cm^2)
+# Bonaire
+feedingrates.vet.BON <- data %>% filter(Species.Code=="qup"&Phase=="i"&Island=="Bonaire") %>% mutate(FR.personal=1047-12*Length.cm) %>% mutate(FR.standard=1347+10*Length.cm) %>% mutate(BS=4.013*10^-4*Length.cm^2)
+feedingrates.vir.BON <- data %>% filter(Species.Code=="stop"&Phase=="i"&Island=="Bonaire") %>% mutate(FR.personal = 1057-15*Length.cm) %>% mutate(FR.standard=1089-17*Length.cm-56) %>% mutate(BS=5.257*10^-4*Length.cm^2)
 
-feedingrates <- bind_rows(feedingrates.vir.ANU,feedingrates.vet.ANU) %>% mutate(areagrazed.ANU=FR.ANU*BS) %>% mutate(areagrazed.pub=FR.pub*BS)
+feedingrates <- bind_rows(feedingrates.vir.ANU,feedingrates.vet.ANU,feedingrates.vir.BON,feedingrates.vet.BON) %>% mutate(AG.personal=FR.personal*BS) %>% mutate(AG.standard=FR.standard*BS)
 
-areagrazed.site <- feedingrates %>% group_by(Site) %>% dplyr::summarize(personal=sum(areagrazed.ANU)/1000,published=sum(areagrazed.pub)/1000)
+areagrazed.site <- feedingrates %>% group_by(Site,Island) %>% dplyr::summarize(personal=sum(AG.personal)/1000,standard=sum(AG.standard)/1000)
 
 areagrazed.site <- as.data.frame(areagrazed.site)
-AG <- melt(areagrazed.site, id="Site") %>% dplyr::rename("feeding.rates"=variable, "area.grazed"=value)
+AG <- melt(areagrazed.site, id=c("Site","Island")) %>% dplyr::rename("data.source"=variable, "area.grazed"=value)
 ```
 ```{r}
-ggplot(data=AG, aes(x=Site, y=area.grazed, fill=feeding.rates)) +
-    geom_bar(stat="identity", position=position_dodge()) + theme_bw() + scale_fill_manual(values=c("black","lightslategrey"),name="Data Source",breaks=c("personal", "published"), labels=c("Fished", "Unfished")) + labs(y=expression(Area~Grazed~(m^2/hr)), title=("Grazing impact of Antiguan scarids with observed vs.
-predicted feeding rates"))
+AG.ANU <- AG %>% filter(Island=="Antigua")
+ggplot(data=AG.ANU, aes(x=Site, y=area.grazed, fill=data.source)) +
+    geom_bar(stat="identity", position=position_dodge()) + theme_bw() + scale_fill_manual(values=c("black","lightslategrey"),name=expression(paste("Feeding rates \n obtained from:")),breaks=c("personal", "standard"), labels=c("fished data", "unfished data")) + labs(y=expression(Area~Grazed~(m^2/hr)), title=("Predicted site-level grazing impact of Antiguan scarids")) + theme(plot.title = element_text(hjust = 0.5)) + theme(axis.text.x = element_text(angle = 45, hjust = 1))
 ggsave("areagrazed.png")
@@
 vetinit <- data2 %>% filter(Species.Code=="qup" & Phase=="i")
+vetinit <- vetinit %>% dplyr::rename(PC1='PC1.x',PC2='PC2.x')
 gamm<-gamm(FR~s(Length.cm)+s(scar.BM)+s(PC1)+s(PC2), family=gaussian(link=identity),data=vetinit, random=list(Island=~1),method="REML")
@@
 virinit <- data2 %>% filter(Species.Code=="stop" & Phase=="i")
+virinit <- virinit %>% dplyr::rename(PC1='PC1.x',PC2='PC2.x')
 gamm<-gamm(FR~s(Length.cm)+s(scar.BM)+s(PC1)+s(PC2), family=gaussian(link=identity),data=virinit, random=list(Island=~1),method="REML")

### 2018-11-13 r/Data_prodding.Rmd
11/12

---
title: "Scarid grazing analysis"
author: "Molly Wilson"
date: "Dec. 2018"
output:
  html_document:
    toc: true
    toc_float: true
    theme: united
    number_sections: true

---

Notes 11/11:
* figure out dcast issues
* replacing NA/0s within pipe

```{r,message=FALSE,warning=FALSE, echo=FALSE}
# set up
library(startR)
library(here)
library(tidyverse)
library(RColorBrewer)
library(corrplot)
library(nlme)
library(pls)
library(factoextra)
```
```{r,message=FALSE,warning=FALSE, echo=FALSE}
sum_site <- read.csv(here("data","sum_site.csv"), stringsAsFactors=F) %>%
  select(-X)
sum_id <- read.csv(here("data","sum_id.csv"), stringsAsFactors=F) %>%
  select(-X)

# stop<-sum_id%>%filter(species_code=="stop")
# qup<-sum_id%>%filter(species_code=="qup")
# rbp<-sum_id%>%filter(species_code=="rbp")
```

# Data overview:
- Fish follows (2 min) conducted at sites in Antigua (6), Barbuda (3) and Bonaire (4) from March-August 2017
- Follows tracked time spent grazing, bite rates, and competitive interactions (among scarids and with damselfish)
- Follows targeted *Sparisoma viride*, *Scarus vetula*, and *Sparisoma aurofrenatum* of both initial and terminal phases 
  +*Sparisoma aurofrenatum* were not followed in Bonaire, but added in Antigua and Bonaire due to low abundances of other two species. Some sites in Antigua and Barbuda had no/low represenation from terminal phase *viride* and *vetula*
- site level factors (benthic, fish, and rugosity) assessed at each of the 13 sites

```{r,echo=FALSE}
sample <- group_by(sum_id, site, species, phase) %>% summarize(n=n())
barplot(table(sum_id$species, sum_id$site), main="Sample size by phase and species",
   legend = rownames(counts), beside=TRUE, las=2)
```

# Examing variable distributions and relationships

## Site-level predictor variables: fish, benthic, and rugosity

### Distributions
```{r,warning=FALSE,echo=FALSE}
predictors <- sum_site %>% 
  select(tot_bm,tot_den,scar_bm,scar_den,carn_bm,carn_den,cca_cover,lc_cover,ma_cover,ma_canopy,ta_cover,ta_canopy,rugosity) %>% 
  gather() %>% 
  na.omit()

ggplot(predictors,aes(value)) +
  facet_wrap(~ key, scales = "free") +
  geom_histogram(aes(y=..density..), colour="black", fill="white") +
  geom_density(alpha=.2, fill="#FF6666") 
```
Site-level predictors are not normally distributed....

### Correlations
```{r,echo=FALSE}
cor = cor(sum_id %>% 
  select(tot_bm,tot_den,scar_bm,scar_den,consp_bm,consp_den,
         carn_bm,carn_den,cca_cover,lc_cover,
         ma_cover, ma_canopy,ta_cover,ta_canopy,rugosity), 
  use="pairwise.complete.obs")
corplot <- corrplot(cor,title = "Correlation Plot", method = "square", outline = T, addgrid.col = "darkgray", order="hclust", mar = c(4,0,4,0), addrect = 4, rect.col = "black", rect.lwd = 5,cl.pos = "b", tl.col = "indianred4", tl.cex = .5, cl.cex = .5)

```

Many correlations here, both within benthic/fish categories and across categories. Rugosity is correlated with both fish and benthic variables. I will examine correlations within fish community and benthic community variables to decide which I can omit and which I will integrate through pcA. Whether I include benthic variables as predictor or response variables will determine if I need to worry about fish~benthic correlations.

```{r,echo=FALSE}
panel.cor <- function(x, y, digits=2, prefix="", cex.cor, ...)
{
    usr <- par("usr"); on.exit(par(usr))
    par(usr = c(0, 1, 0, 1))
    r <- abs(cor(x, y))
    txt <- format(c(r, 0.123456789), digits=digits)[1]
    txt <- paste(prefix, txt, sep="")
    if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
    text(0.5, 0.5, txt, cex = cex.cor * r)
}
pairs(~tot_bm+tot_den+scar_bm+scar_den+consp_bm+consp_den+carn_bm+carn_den+rugosity, data=sum_id,
      lower.panel=panel.smooth, upper.panel=panel.cor, 
      pch=20, main="Fish Community Variables", na.action = na.omit)
```

Variables carn_den and carn_bm are essentially identical - I will remove carn_den and use carn_bm as the main preditor metric. Total fish biomass and density (tot_bm and tot_den) are highly correlated with both scarid and carnivore variables, so I will remove them. conspific biomass is highly correlated with scarid biomass so I will remove it as well (and should possibly also remove consp_den?). The selected variables I will proceed with are as follows:
```{r,echo=FALSE}
pairs(~scar_bm+scar_den+consp_den+carn_bm+rugosity, data=sum_id,
      lower.panel = panel.smooth, upper.panel = panel.cor, 
      pch=20, main = "Selected Fish Community Variables", na.action = na.omit)
```

Scarid biomass, carnivore biomass, and rugosity should be integrated via pcA. Scarid density and conspific density should also be combined, or one should be omitted.

```{r,echo=FALSE}
pairs(~cca_cover+lc_cover+ma_cover+ma_canopy+ta_cover+ta_canopy+rugosity, data = sum_id,
      lower.panel = panel.smooth, upper.panel = panel.cor, 
      pch=20, main ="Benthic Community Variables", na.action = na.omit)
```

All benthic variables are highly correlated. I can omit cca_cover and ma_cover (accounted for by ma_canopy). The rest should be combined via pcA. First I will assess remaining correlations between selected benthic and fish community variables:

```{r,echo=FALSE}
pairs(~scar_bm+scar_den+carn_bm+lc_cover+ta_cover+ma_canopy+ta_canopy+rugosity, data=sum_id,
      lower.panel=panel.smooth, upper.panel=panel.cor, 
      pch=20, main="Selected Fish and Benthic Variables",na.action = na.omit)
```

Unsure how to proceed here if I do end up using both benthic and fish community variables as predictors, because some are corellated (esp. scarid_bm.s and all benthic variables....). It looks like I would integrate scar_bm and carn_bm into a pcA with all other benthic/rugosity variables. Variables scar_den and consp_den are relatively independent of all other variables (except each other).

Current outline for site-level pcA tasks:
If benthic=predictor:
- lc_cover + ta_cover + ma_canopy + ta_canopy + rugosity + scar_bm + carn_bm
- scar_den + consp_den
If benthic=response:
- rugosity + scar_bm + carn_bm
- lc_cover + ta_cover + ma_canopy + ta_canopy
- scar_den + consp_den

## Fish-level response variables: grazing behaviors

### Distributions
```{r,echo=FALSE,warning=FALSE}
grazing <- sum_id %>% 
  select(g_frac,fr,br,for_dur,for_bites) %>% 
  gather() %>% 
  na.omit()

ggplot(grazing,aes(value)) + 
  facet_wrap(~ key, scales = "free") + 
  geom_histogram(aes(y=..density..), colour="black", fill="white") +
  geom_density(alpha=.2, fill="#FF6666") 
```

### Corellations
```{r,echo=FALSE}
cor=cor(sum_id %>% 
          select(g_frac,fr,br,for_dur,for_bites), use="pairwise.complete.obs")
corplot <- corrplot(cor,title = "Correlation Plot", method = "square", outline = T, addgrid.col = "darkgray", order="hclust", mar = c(4,0,4,0), addrect = 4, rect.col = "black", rect.lwd = 5,cl.pos = "b", tl.col = "indianred4", tl.cex = .5, cl.cex = .5)
```

```{r,echo=FALSE}
pairs(~g_frac+fr+br+for_dur+for_bites, data=sum_id,
      lower.panel=panel.smooth, upper.panel=panel.cor, 
      pch=20, main="Grazing Response Variables",na.action = na.omit)
```

Variable for_bites is correlated with nearly all other variables, so will be omitted. Variables g.farc, fr. and for_dur are still correlated and should be assessed with a pcA. Variable br can be a standalone variable.

Grazing pcA:
- g_frac + fr + for_dur
- (br as standalone)

# Exploratory bivariate plots

## Grazing as a function of fish length
```{r, echo=FALSE,warning=FALSE, eval=FALSE}
ggplot(sum_id, aes(x = length_cm, y = fr, colour = island)) +
  facet_grid(.~species) +
  geom_point(shape = 18) +
  scale_shape(solid = FALSE) +
  geom_smooth(method = lm, fullrange = FALSE) +
  ggtitle("Feeding rate by fish length") + 
  labs(y = "Feeding Rate (bites/hr)", x = "Length (cm)") + 
  theme_minimal() + 
  theme(plot.title = element_text(hjust = 0.5)) + 
  scale_color_manual(values = c("navy", "darkseagreen", "slategray2"))
```
```{r, echo=FALSE,warning=FALSE, eval=FALSE}
ggplot(sum_id, aes(x = length_cm, y = br, colour = island)) +
  facet_grid(.~species) +
  geom_point(shape = 18) +
  scale_shape(solid = FALSE) +
  geom_smooth(method = lm, fullrange = FALSE) +
  ggtitle("Feeding rate by fish length") + 
  labs(y = "Bite Rate (bites/sec)", x = "Length (cm)") + 
  theme_minimal() + 
  theme(plot.title = element_text(hjust = 0.5)) + 
  scale_color_manual(values = c("navy", "darkseagreen", "slategray2"))
```
```{r, echo=FALSE,warning=FALSE, eval=FALSE}
ggplot(sum_id, aes(x = length_cm, y = g_frac, colour = island)) +
  facet_grid(.~species) +
  geom_point(shape = 18) +
  scale_shape(solid = FALSE) +
  geom_smooth(method = lm, fullrange = FALSE) +
  ggtitle("Feeding rate by fish length") + 
  labs(y = "Time Fraction", x = "Length (cm)") + 
  theme_minimal() + 
  theme(plot.title = element_text(hjust = 0.5)) + 
  scale_color_manual(values = c("navy", "darkseagreen", "slategray2"))
```

## Grazing as a function of scarid biomass

```{r}
ggplot(sum_id %>% filter(island!="Barbuda"), aes(x = scar_bm,y=fr)) +
  facet_grid(.~species) +
  geom_smooth(method="loess") +
  geom_point(shape=18) +
  ggtitle("Feeding rate by scarid biomass") +
  labs(y="Feeding rate (bites/hr)", x="Scarid Biomass (g/100m2)") +
  theme_minimal() + 
  theme(plot.title = element_text(hjust = 0.5))
```
```{r,echo=FALSE,warning=FALSE}
ggplot(sum_id, aes(x = scar_bm, y = fr, colour = island)) +
  facet_grid(. ~ species) +
  geom_point(shape = 18) +
  geom_smooth(method = lm, fullrange=F) +
  ggtitle("Feeding rate by scarid biomass") +
  labs(y = "Feeding rate (bites/hr)", x = "Scarid Biomass (g/100m2)") +
  theme_minimal() + 
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_color_manual(values = c("navy", "darkseagreen", "slategray2"))
```
```{r,echo=FALSE,warning=FALSE}
ggplot(sum_id, aes(x = scar_bm, y = fr, colour = phase)) +
  facet_grid(. ~ species) +
  geom_point(shape = 18) +
  geom_smooth(method = lm, fullrange=F) +
  ggtitle("Feeding rate by phase and scarid biomass") +
  labs(y = "Feeding rate (bites/hr)", x = "Scarid Biomass (g/100m2)") +
  theme_minimal() + 
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_color_manual(values = c("gray40", "aquamarine3"))
```
```{r,echo=FALSE,warning=FALSE}
ggplot(sum_id, aes(x = scar_bm, y = br, colour = phase)) +
  facet_grid(. ~ species) +
  geom_point(shape = 18) +
  geom_smooth(method = lm, fullrange=F) +
  ggtitle("Bite rates (while grazing) by scarid biomass") +
  labs(y = "Bite rate (bites/sec)", x = "Scarid Biomass (g/100m2)") +
  theme_minimal() + 
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_color_manual(values = c("gray40", "aquamarine3"))
```
```{r,echo=FALSE,warning=FALSE}
ggplot(sum_id, aes(x = scar_bm, y = g_frac, colour = phase)) +
  facet_grid(. ~ species) +
  geom_point(shape = 18) +
  geom_smooth(method = lm, fullrange=F) +
  ggtitle("Fraction of time spent grazing by phase and scarid biomass") +
  labs(y = "Time Fraction", x = "Scarid Biomass (g/100m2)") +
  theme_minimal() + 
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_color_manual(values = c("gray40", "aquamarine3"))
```
```{r,echo=FALSE,warning=FALSE}
ggplot(sum_id, aes(x = scar_bm, y = for_bites, colour = phase)) +
  facet_grid(. ~ species) +
  geom_point(shape = 18) +
  geom_smooth(method = lm, fullrange=F) +
  ggtitle("Bites per foray by phase and scarid biomass") +
  labs(y = "Bites", x = "Scarid Biomass (g/100m2)") +
  theme_minimal() + 
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_color_manual(values = c("gray40", "aquamarine3"))
```

## Competitive interactions

```{r,echo=FALSE,warning=FALSE}
ggplot(sum_site,aes(x=scar_bm,y=int_scar.mean, shape=island)) +
  geom_point(size=3) +
  ggtitle("Interaction frequency as a function of scarid biomass") + 
  theme_minimal() + 
  theme(plot.title = element_text(hjust = 0.5)) + 
  labs(x="Scarid Biomass (g/100m2)", y="Interactions with other scarids (per 2 min follow)") +
  geom_errorbar(aes(ymin=int_scar.mean-int_scar.se, ymax=int_scar.mean+int_scar.se))
```
```{r,echo=FALSE,warning=FALSE}
ggplot(sum_site,aes(x=scar_bm,y=int_scar.mean_stop, shape=island)) +
  geom_point(size=3) +
  ggtitle("S. viride interaction frequency as a function of scarid biomass") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(x = "Scarid Biomass (g/100m2)", y = "Interactions with other scarids (per 2 min follow)") +
  geom_errorbar(aes(ymin=int_scar.mean_stop-int_scar.se_stop, ymax=int_scar.mean_stop+int_scar.se_stop))
```
```{r,echo=FALSE,warning=FALSE}
ggplot(sum_site,aes(x=scar_bm,y=int_scar.mean_qup, shape=island)) +
  geom_point(size=3) +
  ggtitle("S. vetula interaction frequency as a function of scarid biomass") + 
  theme_minimal() + 
  theme(plot.title = element_text(hjust = 0.5)) + 
  labs(x="Scarid Biomass (g/100m2)", y="Interactions with other scarids (per 2 min follow)") + 
  geom_errorbar(aes(ymin=int_scar.mean_qup-int_scar.se_qup, ymax=int_scar.mean_qup+int_scar.se_qup))
```
```{r,echo=FALSE,warning=FALSE}
ggplot(sum_id,aes(x=scar_bm,y=int_scar)) +
  facet_grid(.~species)+geom_point(shape=19) +
  geom_smooth(method=lm) +
  ggtitle("Competitive interactions as a function of scarid biomass") + 
  theme_minimal() + 
  theme(plot.title = element_text(hjust = 0.5))
```

# PCAs -->
*Do I need to normalize my variables pre-pcA? Currently just scaled, except for grazing response variables which are log-transformed* -->

## Predictors -->

### All predictors -->
```{r}
pca <- prcomp(na.omit(sum_site %>% 
              remove_rownames %>% 
              column_to_rownames(var="site") %>%
              select(lc_cover,ta_cover,ma_canopy,ta_canopy,rugosity,scar_bm,scar_den,tot_bm)),
  center = TRUE,scale. = TRUE)
summary(pca)

plot(pca,type="l")

fviz_pca_biplot(pca,
                col.var = "black",
                col.ind = as.factor(sum_site$island), # color by groups
                palette = c("#00AFBB", "#E7B800", "#FC4E07"),
                addEllipses = T, ellipse.level=0.95,
                ellipse.type = "confidence",
                label = "var",
                legend.title = "Groups",
                repel = T
                ) +
  scale_color_brewer(palette="Dark")+
  theme_minimal()

fviz_mca_biplot(pca, 
  habillage = sum_id$island, addEllipses = TRUE,
  label = "var", shape.var = 15) +
  scale_color_brewer(palette="Dark3")+
  theme_minimal()
```

<!-- ### Benthic variables -->
<!-- ```{r, echo=FALSE} -->
<!-- pred.benthic <- sum_site %>% select(lc_cover,ta_cover,ma_canopy,ta_canopy,rugosity) # should it matter if I use site level summary data vs. full follow dataset (where site-level data has been joined?) -->
<!-- pca1 <- prcomp(na.omit(pred.benthic),center = TRUE,scale. = TRUE)  -->

<!-- # add pc1 and pc2 as variables in dataset -->
<!-- pca1.scores <- as.data.frame(pca1$x) %>% select(pc1,pc2) -->
<!-- sum_site <- sum_site  %>% bind_cols(pca1.scores) -->
<!-- pca1.scores <- select(sum_site,site,pc1,pc2) -->
<!-- sum_id_pca <- sum_id %>% left_join(pca1.scores,by="site") -->

<!-- write.csv(sum_id_pca,file = here("data/sum_id_pca.csv")) -->
<!-- ``` -->
<!-- ### Fish community variables -->
<!-- ```{r} -->
<!-- pred.fish <- sum_site %>% select(scar_bm,scar_den,carn_bm) # should it matter if I use site level summary data vs. full follow dataset (where site-level data has been joined?) -->
<!-- pca2 <- prcomp(na.omit(pred.fish),center = TRUE,scale. = TRUE)  -->
<!-- summary(pca2) -->

<!-- plot(pca2,type="l") -->
<!-- ggbiplot(pca2, obs.scale = 1, var.scale = 1, groups=sum_site$island, ellipse = TRUE, circle = TRUE, loadings.label.repel=TRUE) + scale_color_manual(name="island", values=c("navy", "darkseagreen", "slategray2")) + theme(legend.direction = 'horizontal', legend.position = 'top') + theme_minimal() -->

<!-- ``` -->

<!-- ## Response variables -->

<!-- ### Grazing behaviors -->
<!-- *log transform increases P1 explained variance here...* -->
<!-- *Currently for_dur has several NA values (for individuals that didn't graze)... I am omitting that variable for now because I don't want to omit non-grazing individuals (instead would like to keep them as grazing=0* -->
<!-- ```{r} -->
<!-- response <- data %>% select(g_frac,tdb,br) -->
<!-- log.vars5 <- log(response[,1:3]) -->
<!-- pca3 <- prcomp(na.omit(response),center = TRUE,scale. = TRUE)  -->
<!-- summary(pca3) -->
<!-- plot(pca3,type="l") -->
<!-- ggbiplot(pca3, obs.scale = 1, var.scale = 1, groups=data$island,shape=data$island,ellipse = TRUE, circle = TRUE, alpha=0.8,varname.size = 5) + theme_minimal() + scale_color_manual(name="island", values=c("navy", "darkseagreen", "slategray2")) + labs(title="Grazing behaviors by island") + theme(plot.title = element_text(hjust = 0.5)) -->
<!-- ``` -->
<!-- ```{r} -->
<!-- response <- data %>% select(g_frac,fr,br) -->
<!-- log.vars5 <- log(response[,1:3]) -->
<!-- pca3 <- prcomp(na.omit(response),center = TRUE,scale. = TRUE)  -->
<!-- summary(pca3) -->
<!-- plot(pca3,type="l") -->
<!-- ggbiplot(pca3, obs.scale = 1, var.scale = 1,ellipse = TRUE, circle = TRUE, alpha=0.4,varname.size = 5) + theme(legend.direction = 'horizontal', legend.position = 'top') + theme_minimal() +geom_point(aes(col=data$scar_bm)) + labs(col="Scarid Biomass (g/100m2)", title="Grazing Behaviors by Scarid Biomass") + theme(plot.title = element_text(hjust = 0.5)) -->
<!-- ``` -->

<!-- ```{r,echo=FALSE} -->
<!-- vir <- data %>% filter(species_code=="stop") -->
<!-- response <- vir %>% select(g_frac,fr,br) -->
<!-- log.vars5 <- log(response[,1:3]) -->
<!-- pcavir <- prcomp(na.omit(response),center = TRUE,scale. = TRUE)  -->
<!-- ggbiplot(pcavir, obs.scale = 1, var.scale = 1,ellipse = TRUE, circle = TRUE, alpha=1,varname.size = 5) + theme(legend.direction = 'horizontal', legend.position = 'top') + theme_minimal() + labs(col="Scarid Biomass (g/100m2)", title="Grazing Behaviors by Scarid Biomass") + theme(plot.title = element_text(hjust = 0.5)) +geom_point(aes(colour=vir$scar_bm)) -->
<!-- ``` -->
<!-- ```{r,echo=FALSE} -->
<!-- vet <- data %>% filter(species_code=="qup") -->
<!-- response <- vet %>% select(g_frac,fr,br) -->
<!-- log.vars5 <- log(response[,1:3]) -->
<!-- pcavet <- prcomp(na.omit(response),center = TRUE,scale. = TRUE)  -->
<!-- ggbiplot(pcavet, obs.scale = 1, var.scale = 1,ellipse = TRUE, circle = TRUE, alpha=1,varname.size = 5) + theme(legend.direction = 'horizontal', legend.position = 'top') + theme_minimal() + labs(col="Scarid Biomass (g/100m2)", title="Grazing Behaviors by Scarid Biomass") + theme(plot.title = element_text(hjust = 0.5)) +geom_point(aes(colour=vet$scar_bm)) -->
<!-- ``` -->

# Model trials

-site-level predictors: scar_bm,scar_den,carn_bm,benthic (pc1,pc2)
-fish-level predictors: species, phase, length
--eventually run separately for different species
--species*scar_bm interaction?
-random effects: island
-response variables: g_frac, br (?), and fr (run separately)

## Basic lm (to assess predictors of interest)
```{r}
initial<-data%>%filter(phase=="i") # for analysis with initial only
lm_fr <- lm(fr~scar_bm*species+phase+carn_bm+length_cm+pc1+pc2, data=data)
summary(lm_fr) # predicting feeding rate of initial phase only, all species
#vif(lm_fr)
```
*Notes: nearly everything is significant? Significant interaction between scar_bm and species for Spar. viride. carn_bm and scar_bm are highly positively correlated (~0.86) - is it ok to include both?*
*R2 goes up slightly with initial only*

```{r}
lm_g_frac <- lm(g_frac~species*scar_bm+phase+scar_den+carn_bm+phase+length_cm+pc1+pc2, data=data)
summary(lm_g_frac)
```
*Notes: carn_bm and pc1 are strongest predictors of time spent grazing, scar_bm and phase are also significant. Slight significance of species, length, and species-scar_bm interaction*

## Mixed effects models
- random effect: island

```{r}
lme_fr <- lme(fr~phase+length_cm+species*scar_bm+scar_den+carn_bm+pc1+pc2,data=data,random=~1|island, na.action = na.omit)
summary(lme_fr)
```
*Notes: nearly everything is significant except for pc2 and scar_den?*
```{r}
lme_g_frac <- lme(g_frac~phase+length_cm+species*scar_bm+scar_den+carn_bm+pc1+pc2,data=data,random=~1|island, na.action = na.omit)
summary(lme_g_frac)
```
```{r}
lme_fr_stop <- lme(fr~phase+length_cm+scar_bm+scar_den+carn_bm+pc1+pc2,data=stop,random=~1|island, na.action = na.omit)
summary(lme_fr_stop)
```
*Notes: carn_bm and length are signifigant predictors*
```{r}
lme_fr_qup <- lme(fr~phase+length_cm+scar_bm+scar_den+carn_bm+pc1+pc2,data=qup,random=~1|island, na.action = na.omit)
summary(lme_fr_qup)
```
*Notes: pc1 is strongest (magnitude) predictors. Also significant are phase, carn_bm and scar_den*
# Model trials: species run individually

## Feeding rates
```{r}
lme_fr_stop <- lme(fr~phase+length_cm+scar_bm+scar_den+carn_bm+pc1+pc2,data=stop,random=~1|island, na.action = na.omit)
summary(lme_fr_stop)
```
*Notes: carn_bm and length are signifigant predictors*

```{r}
lme_fr_qup <- lme(fr~phase+length_cm+scar_bm+scar_den+carn_bm+pc1+pc2,data=qup,random=~1|island, na.action = na.omit)
summary(lme_fr_qup)
```
*Notes: pc1 is strongest (magnitude) predictors. Also significant are phase, carn_bm and scar_den*

## GAMs
```{r, echo=FALSE}
gam<-gam(fr~species+phase+s(length_cm)+s(scar_bm)+s(pc1)+s(pc2), family=gaussian(link=identity),data=data, method="REML")
summary(gam)
plot(gam,pages=1)
```
```{r, echo=FALSE}
gamvet<-gam(fr~phase+s(length_cm)+s(scar_bm)+(carn_bm)+s(pc1)+s(pc2), family=gaussian(link=identity),data=vet, method="REML")
summary(gamvet)
plot(gamvet,pages=1)
```
```{r, echo=FALSE}
vir<-data %>% filter(species_code=="stop" & site!="Snapper Point")
gamvir<-gam(fr~phase+s(length_cm)+s(scar_bm)+s(pc1)+s(pc2), family=gaussian(link=identity),data=vir,method="REML")
summary(gamvir)
plot(gamvir,pages=1)
```

## GAMMs

### Feeding rate (bites per hour)
```{r,message=FALSE,warning=FALSE}
gamm<-gamm(fr~species+phase+s(length_cm)+s(scar_bm)+s(pc1)+s(pc2), family=gaussian(link=identity),data=data, random=list(island=~1),method="REML")
summary(gamm$gam)
plot(gamm$gam,pages=1)
```
```{r,message=FALSE,warning=FALSE}
gamm<-gamm(fr~phase+s(length_cm)+s(scar_bm)+s(pc1)+s(pc2), family=gaussian(link=identity),data=vet, random=list(island=~1),method="REML")
summary(gamm$gam)
plot(gamm$gam,pages=1)
```
```{r,message=FALSE,warning=FALSE}
gamm<-gamm(fr~phase+s(length_cm)+s(scar_bm)+s(pc1)+s(pc2), family=gaussian(link=identity),data=vir, random=list(island=~1),method="REML")
summary(gamm$gam)
plot(gamm$gam,pages=1)
```

Most of variation is explained by species - for individual species models R2 values are very low (~0.3)

### Fraction of time spent grazing
```{r,message=FALSE,warning=FALSE}
gamm<-gamm(g_frac~species+phase+s(length_cm)+s(scar_bm)+s(pc1)+s(pc2), family=gaussian(link=identity),data=data, random=list(island=~1),method="REML")
summary(gamm$gam)
plot(gamm$gam,pages=1)
```
```{r,message=FALSE,warning=FALSE}
gamm<-gamm(g_frac~phase+s(length_cm)+s(scar_bm)+s(pc1)+s(pc2), family=gaussian(link=identity),data=vet, random=list(island=~1),method="REML")
summary(gamm$gam)
plot(gamm$gam,pages=1)
```
```{r,message=FALSE,warning=FALSE}
gamm<-gamm(g_frac~phase+s(length_cm)+s(scar_bm)+s(pc1)+s(pc2), family=gaussian(link=identity),data=vir, random=list(island=~1),method="REML")
summary(gamm$gam)
plot(gamm$gam,pages=1)
```

Low R2 for all trials predicting g_frac (<0.4)(gammvir)
```
------------------------
To Do as of Nov. 7
* ~~incorporate pcs~~
* boosted regression trees ecosphere 2017 adrians paper
* ~~simpler graphs with just herb density x axis - grazing behaviors y, facet by species~~
* ~~species*density as fixed effect~~
* species as random effect
* ~~break out species and run separately~~

### 2018-11-13 r/Follows.Rmd
11/12

---
title: "Antigua-Barbuda Fish Follows August 2017"
author: "Molly Wilson"
date: "8/8/2017"
output:
  html_document:
    toc: true
    toc_float: true
    theme: united
    number_sections: true
---

Notes 11/11:
* figure out dcast issues
* replacing NA/0s within pipe

Output:
* "sum_id.csv": individual fish-level data (behavioral and survey data)
* "sum_site.csv": site-level summary data (behavioral and survey data)

```{r,message=FALSE,warning=FALSE}
# set up
library(startR)
library(tidyverse)
library(here)
```

```{r, warning=FALSE, message= FALSE}
# import data
site <- read.csv(here("data", "site.csv"), strip.white = T, stringsAsFactors = F) %>% select(-X)
sp_site <- read.csv(here("data","sp_site.csv"), stringsAsFactors = F) %>%
  filter(species %in% c("Sparisoma viride","Scarus vetula","Sparisoma aurofrenatum")) %>%
  select(-X,-family) %>%
  rename_at(vars(3:8), ~ paste0("consp_",.))

follows <- read.csv(here("data", "follows.csv"), strip.white = T, stringsAsFactors = F) %>%
  janitor::clean_names(case = "old_janitor") %>%
  filter(date != "" & date != "3/26/17") %>% 
  select(-activity_code) %>% 
  unite("id",c(site,diver,date,time,follow_number),sep="_", remove=FALSE) %>%
  mutate(
    time_sec = as.numeric(time_sec),
    time_frac = as.numeric(time_fraction),
    length_cm = as.numeric(length_cm),
    weight_g=as.numeric(weight_g),
    bites = as.numeric(ifelse(is.na(bites),0,bites))
  ) %>%
  replace(. =="#N/A", NA)

# create df with single row for all fish followed
sum_id <- follows %>% 
  group_by(id, date, island, site, species, species_code, phase, length_cm, weight_g, follow_sec) %>% 
  summarize()
```

# Grazing behavior

## Time budgets 
Here I create a dataframe of time spent on each activity, per fish, in seconds and time fraction
*(could re-do with dcast____)*
```{r,warning=FALSE, message= FALSE}
tb <- follows %>% 
  group_by(id, activity, species, phase, length_cm) %>% 
  summarize(
    time = sum(time_sec), 
    time_frac = sum(time_frac))

# Add time budgets per activity for each fish to 'sum_id'

tbgraze <- tb %>% 
  filter(activity == "graze") %>% 
  subset(select=c(id,time,time_frac)) %>% 
  rename(
    g_sec=time,
    g_frac=time_frac) # keep seconds for calculating grazing rates
tbclean <- tb %>% 
  filter(activity == "cleaned") %>% 
  subset(select=c(id,time_frac)) %>% 
  rename(clean_frac=time_frac)

# join fish-level data to sum_id dataframe
sum_id <- sum_id %>% 
  left_join(tbgraze, by= "id") %>% 
  left_join(tbclean, by = "id") # %>%
  # replace(is.na(.),0) %>%
  # mutate(length_cm = ifelse(0,NA,length_cm)
```

## Grazing rates
```{r, warning=FALSE, message= FALSE}
# bites/second during total follow time and during grazing forays (including G1 instances)
bites <- follows %>% 
  group_by(id) %>% 
  summarise(bites = sum(bites))
sum_id <- sum_id %>% 
  left_join(bites, by="id") %>%
  mutate(
    br=bites/g_sec,
    fr=bites/follow_sec*60*60,
    tdb=bites/follow_sec*60*60*10.5
    ) %>%
  replace(is.na(.),0)

# # Bites/second during grazing forays, excluding G1 instances
# bites_noG1 <- follows %>% filter(activity=="graze"&Bites>1&Time_sec>1) %>% group_by(id) %>% summarize(Bites_noG1 = sum(Bites), g_sec_noG1=sum(Time_sec)) 
# sum_id <- left_join(sum_id, bites_noG1, by="id") %>% mutate(br_g_noG1=Bites_noG1/g_sec_noG1)
# sum_id$br_g_noG1[is.na(sum_id$br_g_noG1)] <- 0
```

## Foray length and G1 bites
*Should I remove incomplete forays? What about forays >2min?*
```{r,warning=FALSE, message= FALSE}
forays <- follows %>% filter(activity == "graze")
forays_c <- follows %>% # currently don't use this, but might want to only use complete forays
  filter(
    !(activity == "graze" & min == 0) | # filters forays starting at time=0
    !(lead(activity) == "end") # filters forays overlapping with end of follow
  )
sum_forays <- forays %>% 
  group_by(id) %>% 
  summarize(for_dur=mean(time_sec),for_bites=mean(bites))

sum_id <- sum_id %>% left_join(sum_forays, by= "id")

# # Frequency of G1 forays
# G1 <- forays %>% filter(Bites==1&Time_sec==1) %>% group_by(id) %>% dplyr::summarize(G1_freq=n())
# sum_id <- left_join(sum_id,G1, by="id")
# sum_id$G1_freq[is.na(sum_id$G1_freq)] <- 0
```

# Competitive interactions

```{r}
# total interaction instances (all species)
int_tot <- follows %>% 
  filter(initiator>=1) %>% 
  group_by(id) %>% summarise(int_tot=n())
# interaction instances with scarids only
int_scar <- follows %>% 
  filter(initiator>=1) %>% 
  filter(competitor_code %in% c("stop","qup","rbp","prp","strp","rtp","ytp")) %>%
  group_by(id) %>% summarise(int_scar=n())
# interaction instances with conspecifics only
int_con <- follows %>% 
  filter(initiator>=1) %>% 
  filter(species==competitor_species) %>% 
  group_by(id) %>% summarise(int_con=n())
# interaction instances with heterospecific scarids
int_hetscar <- follows %>% 
  filter(initiator>=1) %>% 
  filter(competitor_code=="stop" | competitor_code=="qup" | competitor_code=="rbp") %>% 
  filter(species!=competitor_species) %>% 
  group_by(id) %>% summarise(int_hetscar=n())
# interaction instances with damselfish
int_dam <- follows %>% 
  filter(initiator>=1) %>% 
  filter(competitor_code=="dam") %>% 
  group_by(id) %>% summarise(int_dam=n())

# join and add to sum_id dataframe
sum_id <- sum_id %>% 
  left_join(int_tot, by="id") %>% 
  left_join(int_scar, by="id") %>% 
  left_join(int_con, by="id") %>% 
  left_join(int_hetscar, by="id") %>% 
  left_join(int_dam, by="id")
sum_id[c("int_tot","int_scar","int_con","int_hetscar","int_dam")][is.na(sum_id[c("int_tot","int_scar","int_con","int_hetscar","int_dam")])] <- 0

# matrix of competitive interactions by species
stop <- follows %>% 
  filter(species_code=="stop") %>% 
  mutate(follows=n()) %>% 
  filter(competitor_code %in% c("stop","qup","rbp","prp","rtp","strp")) %>% 
  group_by(competitor_code,island) %>% summarize(stop=n())
qup <- follows %>% 
  filter(species_code=="qup") %>% 
  mutate(follows=n()) %>% 
  filter(competitor_code %in% c("stop","qup","rbp","prp","rtp","strp")) %>% 
  group_by(competitor_code,island) %>% summarize(qup=n())
rbp <- follows %>% 
  filter(species_code=="rbp") %>% 
  mutate(follows=n()) %>% 
  filter(competitor_code %in% c("stop","qup","rbp","prp","rtp","strp")) %>% 
  group_by(competitor_code,island) %>% summarize(rbp=n())

int_matrix <- full_join(stop, qup, by = c("competitor_code","island")) %>%
  full_join(rbp, by = c("competitor_code","island"))
```

# Summarizing data by site, species, and phase
Next I create a summary tables that averages variables within each site/species/phase, then combines into one table
```{r,warning=FALSE, message= FALSE}
# site averages
sum_site <- sum_id %>% 
  group_by(site,island) %>% 
  summarize(
    n=n(),
    g_frac_mean=mean(g_frac),
    g_frac_se=sd(g_frac)/sqrt(n),
    br_mean=mean(br,na.rm=TRUE),
    br_se=sd(br,na.rm=TRUE)/sqrt(n),
    fr_mean=mean(fr),
    fr_se=sd(fr)/sqrt(n),
    for_dur_mean=mean(for_dur,na.rm=TRUE),
    for_dur_se=sd(for_dur,na.rm=TRUE)/sqrt(n),
    for_bites_mean=mean(for_bites,na.rm=TRUE),
    for_bites_se=sd(for_bites,na.rm=TRUE)/sqrt(n),
    int_tot_mean=mean(int_tot),
    int_tot_se=sd(int_tot)/sqrt(n),
    int_scar_mean=mean(int_scar),
    int_scar_se=sd(int_scar)/sqrt(n),
    int_con_mean=mean(int_con),
    int_con_se=sd(int_con)/sqrt(n),
    int_hetscar_mean=mean(int_hetscar),
    int_hetscar_se=sd(int_hetscar)/sqrt(n),
    int_dam_mean=mean(int_dam),
    int_dam_se=sd(int_dam)/sqrt(n)
    ) %>% 
  select(-n)

# site averages by species
sum_site_sp <- sum_id %>% 
  group_by(site, island, species_code) %>% 
  summarize(
    n=n(),
    g_frac_mean=mean(g_frac),
    g_frac_se=sd(g_frac)/sqrt(n),
    br_mean=mean(br,na.rm=TRUE),
    br_se=sd(br,na.rm=TRUE)/sqrt(n),
    fr_mean=mean(fr),
    fr_se=sd(fr)/sqrt(n),
    for_dur_mean=mean(for_dur,na.rm=TRUE),
    for_dur_se=sd(for_dur,na.rm=TRUE)/sqrt(n),
    for_bites_mean=mean(for_bites,na.rm=TRUE),
    for_bites_se=sd(for_bites,na.rm=TRUE)/sqrt(n), 
    int_tot_mean=mean(int_tot),
    int_tot_se=sd(int_tot)/sqrt(n), 
    int_scar_mean=mean(int_scar),
    int_scar_se=sd(int_scar)/sqrt(n), 
    int_con_mean=mean(int_con),
    int_con_se=sd(int_con)/sqrt(n), 
    int_hetscar_mean=mean(int_hetscar),
    int_hetscar_se=sd(int_hetscar)/sqrt(n), 
    int_dam_mean=mean(int_dam),
    int_dam_se=sd(int_dam)/sqrt(n)
    ) %>% 
  select(-n)

# site averages by species and phase
sum_site_sp_ph <- sum_id %>% 
  group_by(site, island, species_code, phase) %>% 
  summarize(
    n=n(),
    g_frac_mean=mean(g_frac),
    g_frac_se=sd(g_frac)/sqrt(n),
    br_mean=mean(br,na.rm=TRUE),
    br_se=sd(br,na.rm=TRUE)/sqrt(n),
    fr_mean=mean(fr),
    fr_se=sd(fr)/sqrt(n),
    for_dur_mean=mean(for_dur,na.rm=TRUE),
    for_dur_se=sd(for_dur,na.rm=TRUE)/sqrt(n),
    for_bites_mean=mean(for_bites,na.rm=TRUE),
    for_bites_se=sd(for_bites,na.rm=TRUE)/sqrt(n), 
    int_tot_mean=mean(int_tot),
    int_tot_se=sd(int_tot)/sqrt(n),
    int_scar_mean=mean(int_scar),
    int_scar_se=sd(int_scar)/sqrt(n),
    int_con_mean=mean(int_con),
    int_con_se=sd(int_con)/sqrt(n), 
    int_hetscar_mean=mean(int_hetscar),
    int_hetscar_se=sd(int_hetscar)/sqrt(n),
    int_dam_mean=mean(int_dam),
    int_dam_se=sd(int_dam)/sqrt(n)
    ) %>% 
  select(-n)
```

### 1_3 Join datasets, print .csv
```{r, echo=FALSE,message=FALSE,warning=FALSE}
# integrate all site-level variables
sum_site <- sum_site %>% 
  left_join(site, by="site")
sum_id <- sum_id %>% 
  left_join(site, by="site") %>%
  left_join(sp_site, by=c("species","site")) # add conspecific scarid info (based on site/species)

write.csv(sum_site,file=here("data/sum_site.csv"))
write.csv(sum_id,file=here("data/sum_id.csv"))
```

### 2018-11-13 r/Graphing+Analysis.Rmd
11/12

--- 2018-11-13 r/Data_prodding.Rmd
@@
 library(tidyverse)
+# library(lazyeval)
+# library(ellipse)
 library(RColorBrewer)
+#library(ggbiplot)
 library(corrplot)
@@
 sum_site <- read.csv(here("data","sum_site.csv"), stringsAsFactors=F) %>%
-  select(-X)
+  janitor::clean_names(case="old_janitor")
+
 sum_id <- read.csv(here("data","sum_id.csv"), stringsAsFactors=F) %>%
-  select(-X)
+  janitor::clean_names(case="old_janitor") %>%
+  select(island,site,spearfishing,diving,species,species_code,phase,length_cm,g_frac,fr,br,tdb,for_dur,for_bites,tot_bm,tot_den,scar_bm,herb_bm,scar_den,conspec_bm,conspec_den,carn_bm,carn_den,cover_cca,cover_lc,cover_ma,cover_ta,canopy_ma,canopy_ta,rugosity,int_tot,int_scar,int_con,int_hetscar,int_dam)
 
-# stop<-sum_id%>%filter(species_code=="stop")
-# qup<-sum_id%>%filter(species_code=="qup")
-# rbp<-sum_id%>%filter(species_code=="rbp")
+stop<-sum_id%>%filter(species_code=="stop")
+qup<-sum_id%>%filter(species_code=="qup")
+rbp<-sum_id%>%filter(species_code=="rbp")
 ```
@@
 sample <- group_by(sum_id, site, species, phase) %>% summarize(n=n())
-barplot(table(sum_id$species, sum_id$site), main="Sample size by phase and species",
-   legend = rownames(counts), beside=TRUE, las=2)
+ggplot(data=sample, aes(x=site, y=n, fill=phase)) +
+    geom_bar(stat="identity", position=position_dodge()) + facet_grid(.~species) + theme_minimal() + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + scale_fill_grey()
+# counts <- table(sum_id$species, sum_id$site)
+# barplot(counts, main="Sample size by phase and species",
+#   legend = rownames(counts), beside=TRUE, las=2)
 ```
@@
 predictors <- sum_site %>% 
-  select(tot_bm,tot_den,scar_bm,scar_den,carn_bm,carn_den,cca_cover,lc_cover,ma_cover,ma_canopy,ta_cover,ta_canopy,rugosity) %>% 
+  select(tot_bm,tot_den,scar_bm,scar_den,carn_bm,carn_den,cover_cca,cover_lc,cover_ma,canopy_ma,cover_ta,canopy_ta,rugosity) %>% 
   gather() %>% 
@@
 ```
-Site-level predictors are not normally distributed....
 
@@
 ```{r,echo=FALSE}
-cor = cor(sum_id %>% 
-  select(tot_bm,tot_den,scar_bm,scar_den,consp_bm,consp_den,
-         carn_bm,carn_den,cca_cover,lc_cover,
-         ma_cover, ma_canopy,ta_cover,ta_canopy,rugosity), 
-  use="pairwise.complete.obs")
-corplot <- corrplot(cor,title = "Correlation Plot", method = "square", outline = T, addgrid.col = "darkgray", order="hclust", mar = c(4,0,4,0), addrect = 4, rect.col = "black", rect.lwd = 5,cl.pos = "b", tl.col = "indianred4", tl.cex = .5, cl.cex = .5)
-
+site_var <- sum_id %>% select(tot_bm,tot_den,scar_bm,scar_den,conspec_bm,conspec_den,carn_bm,carn_den,cover_cca,cover_lc,cover_ma,canopy_ma,cover_ta,canopy_ta,rugosity)
+cor = cor(site_var, use="pairwise.complete.obs")
+my_colors <- brewer.pal(5, "Spectral")
+my_colors = colorRampPalette(my_colors)(100)
+ord <- order(cor[1, ])
+cor_ord = cor[ord, ord]
+plotcorr(cor_ord , col=my_colors[cor_ord*50+50] , mar=c(1,1,1,1)  )
 ```
@@
 }
-pairs(~tot_bm+tot_den+scar_bm+scar_den+consp_bm+consp_den+carn_bm+carn_den+rugosity, data=sum_id,
+pairs(~tot_bm+tot_den+scar_bm+scar_den+conspec_bm+conspec_den+carn_bm+carn_den+rugosity, data=sum_id,
       lower.panel=panel.smooth, upper.panel=panel.cor, 
@@
 Total fish biomass and density (tot_bm and tot_den) are highly correlated with both scarid and carnivore variables, so I will remove them.
-conspific biomass is highly correlated with scarid biomass so I will remove it as well (and should possibly also remove consp_den?).
+Conspecific biomass is highly correlated with scarid biomass so I will remove it as well (and should possibly also remove conspec_den?).
 The selected variables I will proceed with are as follows:
 ```{r,echo=FALSE}
-pairs(~scar_bm+scar_den+consp_den+carn_bm+rugosity, data=sum_id,
+panel.cor <- function(x, y, digits=2, prefix="", cex.cor, ...)
+{
+    usr <- par("usr"); on.exit(par(usr))
+    par(usr = c(0, 1, 0, 1))
+    r <- abs(cor(x, y))
+    txt <- format(c(r, 0.123456789), digits=digits)[1]
+    txt <- paste(prefix, txt, sep="")
+    if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
+    text(0.5, 0.5, txt, cex = cex.cor * r)
+}
+pairs(~scar_bm+scar_den+conspec_den+carn_bm+rugosity, data=sum_id,
       lower.panel = panel.smooth, upper.panel = panel.cor, 
@@
 
-Scarid biomass, carnivore biomass, and rugosity should be integrated via pcA. Scarid density and conspific density should also be combined, or one should be omitted.
+Scarid biomass, carnivore biomass, and rugosity should be integrated via pcA. Scarid density and conspecific density should also be combined, or one should be omitted.
 
 ```{r,echo=FALSE}
-pairs(~cca_cover+lc_cover+ma_cover+ma_canopy+ta_cover+ta_canopy+rugosity, data = sum_id,
+panel.cor <- function(x, y, digits=2, prefix="", cex.cor, ...)
+{
+    usr <- par("usr"); on.exit(par(usr))
+    par(usr = c(0, 1, 0, 1))
+    r <- abs(cor(x, y))
+    txt <- format(c(r, 0.123456789), digits=digits)[1]
+    txt <- paste(prefix, txt, sep="")
+    if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
+    text(0.5, 0.5, txt, cex = cex.cor * r)
+}
+pairs(~cover_cca+cover_lc+cover_ma+canopy_ma+cover_ta+canopy_ta+rugosity, data = sum_id,
       lower.panel = panel.smooth, upper.panel = panel.cor, 
@@
 All benthic variables are highly correlated.
-I can omit cca_cover and ma_cover (accounted for by ma_canopy).
+I can omit cover_cca and cover_ma (accounted for by canopy_ma).
 The rest should be combined via pcA.
@@
 ```{r,echo=FALSE}
-pairs(~scar_bm+scar_den+carn_bm+lc_cover+ta_cover+ma_canopy+ta_canopy+rugosity, data=sum_id,
+panel.cor <- function(x, y, digits=2, prefix="", cex.cor, ...)
+{
+    usr <- par("usr"); on.exit(par(usr))
+    par(usr = c(0, 1, 0, 1))
+    r <- abs(cor(x, y))
+    txt <- format(c(r, 0.123456789), digits=digits)[1]
+    txt <- paste(prefix, txt, sep="")
+    if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
+    text(0.5, 0.5, txt, cex = cex.cor * r)
+}
+pairs(~scar_bm+scar_den+carn_bm+cover_lc+cover_ta+canopy_ma+canopy_ta+rugosity, data=sum_id,
       lower.panel=panel.smooth, upper.panel=panel.cor, 
@@
 It looks like I would integrate scar_bm and carn_bm into a pcA with all other benthic/rugosity variables.
-Variables scar_den and consp_den are relatively independent of all other variables (except each other).
+Variables scar_den and conspec_den are relatively independent of all other variables (except each other).
 
@@
 If benthic=predictor:
-- lc_cover + ta_cover + ma_canopy + ta_canopy + rugosity + scar_bm + carn_bm
-- scar_den + consp_den
+- cover_lc + cover_ta + canopy_ma + canopy_ta + rugosity + scar_bm + carn_bm
+- scar_den + conspec_den
 If benthic=response:
 - rugosity + scar_bm + carn_bm
-- lc_cover + ta_cover + ma_canopy + ta_canopy
-- scar_den + consp_den
+- cover_lc + cover_ta + canopy_ma + canopy_ta
+- scar_den + conspec_den
 
@@
 ```{r,echo=FALSE}
-cor=cor(sum_id %>% 
-          select(g_frac,fr,br,for_dur,for_bites), use="pairwise.complete.obs")
-corplot <- corrplot(cor,title = "Correlation Plot", method = "square", outline = T, addgrid.col = "darkgray", order="hclust", mar = c(4,0,4,0), addrect = 4, rect.col = "black", rect.lwd = 5,cl.pos = "b", tl.col = "indianred4", tl.cex = .5, cl.cex = .5)
+grazing <- sum_id %>% select(g_frac,fr,br,for_dur,for_bites)
+cor=cor(grazing, use="pairwise.complete.obs")
+ord <- order(cor[1, ])
+cor_ord = cor[ord, ord]
+plotcorr(cor_ord , col=my_colors[cor_ord*50+50] , mar=c(1,1,1,1)  )
 ```
@@
 ```{r,echo=FALSE}
+panel.cor <- function(x, y, digits=2, prefix="", cex.cor, ...)
+{
+    usr <- par("usr"); on.exit(par(usr))
+    par(usr = c(0, 1, 0, 1))
+    r <- abs(cor(x, y))
+    txt <- format(c(r, 0.123456789), digits=digits)[1]
+    txt <- paste(prefix, txt, sep="")
+    if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
+    text(0.5, 0.5, txt, cex = cex.cor * r)
+}
 pairs(~g_frac+fr+br+for_dur+for_bites, data=sum_id,
@@
 Variable br can be a standalone variable.
+
+```{r,echo=FALSE}
+panel.cor <- function(x, y, digits=2, prefix="", cex.cor, ...)
+{
+    usr <- par("usr"); on.exit(par(usr))
+    par(usr = c(0, 1, 0, 1))
+    r <- abs(cor(x, y))
+    txt <- format(c(r, 0.123456789), digits=digits)[1]
+    txt <- paste(prefix, txt, sep="")
+    if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
+    text(0.5, 0.5, txt, cex = cex.cor * r)
+}
+pairs(~g_frac+fr+br+for_dur, data=sum_id,
+      lower.panel=panel.smooth, upper.panel=panel.cor, 
+      pch=20, main="Selected Grazing Response Variables",na.action = na.omit)
+```
 
@@
 
-# PCAs -->
-*Do I need to normalize my variables pre-pcA? Currently just scaled, except for grazing response variables which are log-transformed* -->
+<!-- # pcAs -->
+<!-- *Do I need to normalize my variables pre-pcA? Currently just scaled, except for grazing response variables which are log-transformed* -->
 
-## Predictors -->
+<!-- ## Predictors -->
 
-### All predictors -->
-```{r}
-pca <- prcomp(na.omit(sum_site %>% 
-              remove_rownames %>% 
-              column_to_rownames(var="site") %>%
-              select(lc_cover,ta_cover,ma_canopy,ta_canopy,rugosity,scar_bm,scar_den,tot_bm)),
-  center = TRUE,scale. = TRUE)
-summary(pca)
+<!-- ### All predictors -->
+<!-- ```{r} -->
+<!-- predictors <- sum_site %>% select(cover_lc,cover_ta,canopy_ma,canopy_ta,rugosity,scar_bm,scar_den,tot_bm) # should it matter if I use site level summary data vs.
+full follow dataset (where site-level data has been joined?) -->
+<!-- pca <- prcomp(na.omit(predictors),center = TRUE,scale. = TRUE)  -->
+<!-- summary(pca) -->
 
-plot(pca,type="l")
-
-fviz_pca_biplot(pca,
-                col.var = "black",
-                col.ind = as.factor(sum_site$island), # color by groups
-                palette = c("#00AFBB", "#E7B800", "#FC4E07"),
-                addEllipses = T, ellipse.level=0.95,
-                ellipse.type = "confidence",
-                label = "var",
-                legend.title = "Groups",
-                repel = T
-                ) +
-  scale_color_brewer(palette="Dark")+
-  theme_minimal()
-
-fviz_mca_biplot(pca, 
-  habillage = sum_id$island, addEllipses = TRUE,
-  label = "var", shape.var = 15) +
-  scale_color_brewer(palette="Dark3")+
-  theme_minimal()
-```
-
+<!-- plot(pca,type="l") -->
+<!-- ggbiplot(pca, obs.scale = 1, var.scale = 1, groups=sum_site$island, ellipse = TRUE, circle = TRUE, varname.size = 2) + scale_color_manual(name="island", values=c("navy", "darkseagreen", "slategray2")) + theme(legend.direction = 'horizontal', legend.position = 'top') + theme_minimal() -->
+<!-- ggsave("pcApred.png") -->
+<!-- ``` -->
 <!-- ### Benthic variables -->
 <!-- ```{r, echo=FALSE} -->
-<!-- pred.benthic <- sum_site %>% select(lc_cover,ta_cover,ma_canopy,ta_canopy,rugosity) # should it matter if I use site level summary data vs.
+<!-- pred.benthic <- sum_site %>% select(cover_lc,cover_ta,canopy_ma,canopy_ta,rugosity) # should it matter if I use site level summary data vs.
 full follow dataset (where site-level data has been joined?) -->
 <!-- pca1 <- prcomp(na.omit(pred.benthic),center = TRUE,scale. = TRUE)  -->
+
+<!-- ggbiplot(pca1, obs.scale = 1, var.scale = 1, groups=sum_site$island, ellipse = TRUE, circle = TRUE, loadings.label.repel=TRUE, varname.size=2.5, varname.adjust=1.1) + scale_color_manual(name="island", values=c("navy", "darkseagreen", "slategray2")) + theme(legend.direction = 'horizontal', legend.position = 'top') + theme_minimal() -->
 

### 2018-11-13 r/SEM.Rmd
11/12

---
title: 'Structural '
author: "Molly Wilson"
date: "3/14/2018"
output:
  html_document:
    number_sections: yes
    theme: united
    toc: yes
    toc_float: yes
    toc_depth: 3
  pdf_document:
    toc: yes
---

```{r,message=FALSE,warning=FALSE, echo=FALSE}
# set up
library(rmarkdown)
library(dplyr)
library(tidyr)
library(lavaan)
library(semPlot)
```

```{r,message=FALSE,warning=FALSE, echo=FALSE}
setwd("~/github/Scarid-Behavior/data")
site.sum <- read.csv("site.sum.csv", stringsAsFactors=FALSE) %>% select(-X)
data <- read.csv("data2.csv", stringsAsFactors=FALSE)
# add aglal indeces
data <- data %>% mutate(index_MA = cover_MA*canopy_MA) %>% mutate(index_TA = cover_TA*canopy_TA)
# recode phase and species with dummy variables
data <- data %>% mutate(spec.dum= 
                          ifelse(Species.Code=="stop", 1, 
                          (ifelse(Species.Code=="qup", 2,
                          (ifelse(Species.Code=="rbp",3,NA)))))) %>%
  mutate(phase.dum=ifelse(Phase=="i",1,2))

```

```{r}
# define model pathways
model <- '
    FR ~ Length.cm + phase.dum + spec.dum+ int.scar + RISK + NUTRITION
    int.scar ~ scar.BM + cover_MA + cover_TA
    rugosity ~ cover_LC
    cover_LC ~ cover_MA + cover_TA
    cover_TA ~ scar.BM
    cover_MA ~ scar.BM

    RISK =~ scar.BM + rugosity
    NUTRITION =~ cover_MA + cover_TA
'

fit <- sem(model, data=data)
varTable(fit)
summary(fit, standardized=T, fit.measures=T,rsq=T)
semPaths(fit)
```

### 2018-11-13 r/Sites.Rmd
11/12

---
title: "Fish & Benthic Community data: Antigua, Barbuda, & Bonaire - Aug. 2017"
author: "Molly Wilson"
date: "8/8/2017"
output:
  html_document:
    toc: true
    toc_float: true
    theme: united
    number_sections: true
---

Notes 11/11:
* figure out dcast issues
* replacing NA/0s within pipe

Final outputs:
* "site.csv": site-level summary data of fish, benthic, rugosity, and human impact variables
* "sp_site.csv": site-level fish data by species (ultimately used to merge conspecific data with fish follows df)

# Fish community data

```{r, echo=F,message=F,warning=F}
# set up
library(startR)
library(here)
library(janitor)
library(tidyverse)
```
```{r, echo=F,message=F,warning=F}
# import files
ab_fish <-read.csv(here("data","ab_fish.csv"), stringsAsFactors=F) %>%
  janitor::clean_names(case="old_janitor") %>%
  uncount(number) %>% # expand to replicate rows if multiple fish were recorded
  select(-c(time,depth_ft,surveyor,date)) 
bon_herb <- read.csv(here("data","bon_herb.csv"), stringsAsFactors=F) %>% # this file is already expanded
  janitor::clean_names(case="old_janitor") %>%
  filter(date==2017) %>% 
  select(-date) %>%
  rename(island = region)
bon_carn <- read.csv(here("data","bon_carn.csv"), stringsAsFactors=F) %>% 
  janitor::clean_names(case="old_janitor") %>%
  rename(island = region) %>%
  uncount(number) %>%
  select(-c(diver,date))
fish <- bind_rows(bon_herb,bon_carn) %>% 
  filter(site %in% c("Oil Slick","Karpata","Front Porch","18th Palm")) %>%
  bind_rows(ab_fish) %>%
  mutate(biomass = as.numeric(biomass)) %>%
  replace(. == "", NA)

ab_benthic <- read.csv(here("data","ab_benthic.csv"), stringsAsFactors=F) %>%
  janitor::clean_names(case="old_janitor") %>%
  mutate(height=as.numeric(height))
ab_rugosity <- read.csv(here("data","ab_rugosity.csv"), stringsAsFactors=F) %>%
  janitor::clean_names(case="old_janitor")
bon_benthic <- read.csv(here("data","bon_benthic.csv"),stringsAsFactors = F) %>%
  janitor::clean_names(case="old_janitor")
human <- read.csv(here("data","site_human.csv"), stringsAsFactors=F) %>%
  janitor::clean_names(case="old_janitor")
```

## Total fish population, by site
```{r,message=F,warning=F}
fish_tran <- fish %>% 
  filter(biomass != 0) %>% 
  group_by(island,site,transect) %>% 
  summarize(
    bm_t=sum(biomass)/1.2, 
    den_t=n()/1.2, 
    l_t=mean(length)
    )

fish_site <-fish_tran %>% 
  group_by(site) %>% 
  summarize(
    n=n(), 
    tot_bm=mean(bm_t), 
    tot_bm_se=(sd(bm_t))/sqrt(n), 
    tot_den=mean(den_t), 
    tot_den_se=(sd(den_t))/sqrt(n), 
    tot_l=mean(l_t), 
    tot_l_se=(sd(l_t))/sqrt(n)
    ) %>% 
  select(-n) # average across transects within each site

boxplot(bm_t~site, las=2, data = fish_tran)
```

## Total fish populations, by species and site
```{r,message=F,warning=F}
# Because not all species were present at each site/transect, I need to create template to add 0 values for biomass and density in transects where they are absent.
exp_tran <- fish %>% 
  group_by(site,transect) %>% 
  summarize() # all possible site/transect combinations
exp_sp <- fish %>% expand(nesting(species,family),site) # all possible species/site combinations
template_sp <- full_join(exp_sp,exp_tran,by="site") # species/site/transect combinations

# Biomass and density, by site and species:
sp_tran <- fish %>% 
  filter(biomass != 0) %>% 
  group_by(site,transect,species) %>% 
  summarize(
    bm_t=sum(biomass)/1.2, 
    den_t=n()/1.2,
    l_t=mean(length)
    ) %>% 
  right_join(template_sp, by=c("site","transect","species")) %>% # join transect-level data with template
  mutate_each(funs(replace(.,is.na(.),0)), -l_t) # missing biomass and density values should be 0

sp_site<-sp_tran %>% 
  group_by(site,species,family) %>% 
  summarize(
    n=n(), # sample size for biomass and density includes 0 values
    l_n=sum(!is.na(l_t)), # only count non-0 values in sample size for length
    bm=mean(bm_t), 
    bm_se=(sd(bm_t))/sqrt(n), 
    den=mean(den_t), 
    den_se=(sd(den_t))/sqrt(n),
    l=mean(l_t, na.rm = T),
    l_se=(sd(l_t, na.rm = T))/sqrt(l_n)
    ) %>%
  select(-c(n,l_n)) %>%
  replace(. == "NaN",NA)

```

## Total fish populations, by functional group and site
```{r,message=F,warning=F}
# Herbivores
herb_tran<- fish %>% 
  filter(family %in% c("Scaridae","Acanthuridae"),
         biomass != 0) %>% 
  group_by(site,transect) %>% 
  summarize(
    bm_t=sum(biomass)/1.2,
    den_t=n()/1.2,
    l_t=mean(length)
    ) %>% 
  right_join(exp_tran, by=c("site","transect")) %>% # join with transect-level template
  mutate_each(funs(replace(.,is.na(.),0)), -l_t) # missing biomass and density values should be 0

herb_site<-herb_tran %>% 
  group_by(site) %>% 
  summarize(
    n=n(),
    l_n=sum(!is.na(l_t)),
    herb_bm=mean(bm_t), 
    herb_bm_se=(sd(bm_t))/sqrt(n), 
    herb_den=mean(den_t), 
    herb_den_se=(sd(den_t))/sqrt(n), 
    herb_l=mean(l_t, na.rm = T),
    herb_l_se=(sd(l_t, na.rm = T))/sqrt(l_n)
    ) %>% 
  select(-c(n,l_n)) %>%
  replace(. == "NaN",NA)

# Acanthurids

acan_tran <- fish %>% 
  filter(family=="Acanthuridae",
         biomass != 0) %>% 
  group_by(site,transect) %>% 
  summarize(
    bm_t=sum(biomass)/1.2,
    den_t=n()/1.2,
    l_t=mean(length)
    ) %>%
  right_join(exp_tran, by=c("site","transect")) %>%
  mutate_each(funs(replace(.,is.na(.),0)), -l_t)

acan_site<-acan_tran %>% group_by(site) %>% 
  summarize(
    n=n(), 
    l_n=sum(!is.na(l_t)),
    acan_bm=mean(bm_t),
    acan_bm_se=(sd(bm_t))/sqrt(n),
    acan_den=mean(den_t),
    acan_den_se=(sd(den_t))/sqrt(n),
    acan_l=mean(l_t, na.rm = T),
    acan_l_se=(sd(l_t, na.rm = T))/sqrt(l_n)
    ) %>% 
  select(-c(n,l_n)) %>%
  replace(. == "NaN",NA)

# Carnivores
carn_tran <- fish %>% 
  filter(family %in% c("Serranidae","lutjanidae","Carangidae"),
         biomass != 0) %>% 
  group_by(site,transect) %>% 
  summarize(
    bm_t=sum(biomass)/1.2,
    den_t=n()/1.2,
    l_t=mean(length)
    ) %>% 
  right_join(exp_tran,by=c("site","transect")) %>%
  mutate_each(funs(replace(.,is.na(.),0)), -l_t)

carn_site<-carn_tran %>% 
  group_by(site) %>% 
  summarize(
    n=n(),
    l_n=sum(!is.na(l_t)),
    carn_bm=mean(bm_t),
    carn_bm_se=(sd(bm_t))/sqrt(n),
    carn_den=mean(den_t),
    carn_den_se=(sd(den_t))/sqrt(n),
    carn_l=mean(l_t, na.rm = T),
    carn_l_se=(sd(l_t, na.rm = T))/sqrt(l_n)
    ) %>% 
  select(-c(n,l_n)) %>%
  replace(. == "NaN",NA)

```

## Scarid populations, by site

```{r,message=F,warning=F}
# Total scarids
scar_tran <- fish %>% 
  filter(family=="Scaridae",
         biomass != 0) %>% 
  group_by(site,transect) %>% 
  summarize(
    bm_t=sum(biomass)/1.2,
    den_t=n()/1.2,
    l_t=mean(length)
    ) %>% 
  right_join(exp_tran,by=c("site","transect")) %>%
  mutate_each(funs(replace(.,is.na(.),0)), -l_t)

scar_site <- scar_tran %>% 
  group_by(site) %>% 
  summarize(
    n=n(),
    l_n=sum(!is.na(l_t)),
    scar_bm=mean(bm_t),
    scar_bm_se=(sd(bm_t))/sqrt(n),
    scar_den=mean(den_t),
    scar_den_se=(sd(den_t))/sqrt(n),
    scar_l=mean(l_t, na.rm = T),
    scar_l_se=(sd(l_t, na.rm = T))/sqrt(l_n)
    ) %>% 
  select(-c(n,l_n)) %>%
  replace(. == "NaN",NA)

# Scarids by species (join wide S. vetula, S. viride, and S. aurofrenatum density, biomass, and length data)

scar_site_sp <- sp_site %>% 
    ungroup() %>%
    filter(species=="Sparisoma viride") %>% 
    select(-c(species,family)) %>%
    rename_at(vars(2:7), ~ paste0("vir_",.)) %>%
  left_join(sp_site %>% 
    ungroup() %>%
    filter(species=="Scarus vetula") %>% 
    select(-c(species,family)) %>%
    rename_at(vars(2:7), ~ paste0("vet_",.)),
    by="site"
  ) %>%
  left_join(sp_site %>% 
    ungroup() %>%
    filter(species=="Sparisoma aurofrenatum") %>% 
    select(-c(species,family)) %>%
    rename_at(vars(2:7), ~ paste0("aur_",.)),
    by="site"
  )

# Scarids by phase (should this be by species, too?)
exp_ph <- fish %>% expand(phase,site)
template_ph <- full_join(exp_ph,exp_tran,by="site") # all site/transect/phase combinations

ph_tran <- fish %>% 
  filter(family=="Scaridae",
         biomass != 0) %>% 
  group_by(site,transect,phase) %>% 
  summarize(
    bm_t=sum(biomass)/1.2,
    den_t=n()/1.2,
    l_t=mean(length)
    ) %>% 
  right_join(template_ph, by=c("site","transect","phase")) %>% # join transect-level data with template
  mutate_each(funs(replace(.,is.na(.),0)), -l_t)

ph_site <- ph_tran %>% 
  group_by(site,phase) %>% 
  summarize(
    n=n(),
    l_n=sum(!is.na(l_t)),
    bm=mean(bm_t),
    bm_se=(sd(bm_t))/sqrt(n),
    den=mean(den_t),
    den_se=(sd(den_t))/sqrt(l_n),
    l=mean(l_t, na.rm = T),
    l_se=(sd(l_t, na.rm = T))/sqrt(l_n)
    ) %>%
  select(-c(n,l_n)) %>%
  replace(. == "NaN",NA)

scar_site_ph <- ph_site %>% 
    ungroup() %>%
    filter(phase=="j") %>% 
    select(-c(phase)) %>%
    rename_at(vars(2:7), ~ paste0("juv_",.)) %>%
  left_join(ph_site %>% 
    ungroup() %>%
    filter(phase=="i") %>% 
    select(-c(phase)) %>%
    rename_at(vars(2:7), ~ paste0("init_",.)),
    by="site"
  ) %>%
  left_join(ph_site %>% 
    ungroup() %>%
    filter(phase=="t") %>% 
    select(-c(phase)) %>%
    rename_at(vars(2:7), ~ paste0("term_",.)),
    by="site"
  )
```

# Benthic & rugosity data

## Percent cover and canopy heights

```{r, echo=F,message=F,warning=F}
# Because not all categories were present within each meter, I need to create template to add 0 values in meters where they were ab_fishsent.
exp_meter <- ab_benthic %>% 
  group_by(site,transect,meter) %>% 
  summarize()
exp_cat <- ab_benthic %>% expand(nesting(code),site)
template_bent <- full_join(exp_meter,exp_cat,by="site")

# Then I can join benthic data with this template, and replace empty rows with 0 values
bent_meter <- ab_benthic %>% 
  group_by(site,transect,meter,code) %>% 
  summarize(count_m=n()) %>% 
  right_join(template_bent, by=c("site","transect","meter","code")) %>% # join transect-level data with template
  replace(is.na(.),0) # convert NAs to 0 biomass/density in transects where a species is not present

# Calculating percent cover on potential substrate (i.e. not sand), and remove meters that were 100% sand cover (no potential susbtrate)

sand_meter <- bent_meter %>% 
  filter(code=="sand") %>% 
  rename(count_sand=count_m) %>% 
  select(-code) 
bent_meter <- left_join(bent_meter,sand_meter,by=c("site","transect","meter")) %>%
  mutate(cover_m=100*count_m/(10-count_sand)) %>% 
  filter(!is.nan(cover_m)) %>% filter(code!="sand") 
bent_tran <- bent_meter %>% 
  group_by(site,transect,code) %>% 
  summarize(cover_t=mean(cover_m))
bent_site <- bent_tran %>% 
  group_by(site,code) %>% 
  summarize(
    n=n(),
    cover=mean(cover_t),
    cover_se=sd(cover_t)/sqrt(n)
    ) %>% 
  select(-n)

# Canopy heights

can_site <- ab_benthic %>% 
  filter(height>=1) %>% 
  group_by(site,transect,meter,code) %>% 
  summarize(canopy_m=mean(height)) %>% 
  group_by(site,transect,code) %>% 
  summarize(canopy_t=mean(canopy_m)) %>% 
  group_by(site,code) %>% 
  summarize(
    n=n(),
    canopy=mean(canopy_t),
    canopy_se=sd(canopy_t)/sqrt(n)
    ) %>% 
  select(-n)

# Spread into wide site-level format, join with percent cover data in benthic.wide

spread_n <- function(df, key, value) {
    keyq <- rlang::enquo(key)
    valueq <- rlang::enquo(value)
    s <- rlang::quos(!!valueq)
    df %>% gather(variable, value, !!!s) %>%
        unite(temp, !!keyq, variable) %>%
        spread(temp, value)
}

bent_site <- bent_site %>% 
      spread_n(code, c(cover, cover_se)) %>%
  left_join(can_site %>% 
            spread_n(code, c(canopy,canopy_se))
  ) %>% 
  janitor::clean_names(case="old_janitor")

```

## Rugosity
```{r, echo=F,message=F,warning=F}
rug_site <- ab_rugosity %>% 
  group_by(site,transect) %>% 
  summarize(rugosity_t=mean(rugosity_cm)/100) %>% 
  group_by(site) %>% 
  summarize(
    n=n(),
    rugosity=mean(rugosity_t),
    rugosity_se=sd(rugosity_t)/sqrt(n)
    ) %>% 
  select(-n) %>% 
  rbind(c("Pallaster West",1.7006,NA)) %>% # missing data for Pallaster West, will use mean of Pallaster East and Mid because reef structure across all three was extremely similar
  mutate(
    rugosity=as.numeric(rugosity),
    rugosity_se=as.numeric(rugosity_se)
  )

bent_site <- bent_site %>% left_join(rug_site,by="site") %>% bind_rows(bon_benthic)
```

# Integrating and exporting .csvs
```{r, echo=F,message=F,warning=F}
site <- fish_site %>% 
  left_join(herb_site,by="site") %>% 
  left_join(acan_site,by="site") %>% 
  left_join(scar_site,by="site") %>% 
  left_join(scar_site_sp,by="site") %>% 
  left_join(scar_site_ph,by="site") %>% 
  left_join(carn_site,by="site") %>%
  left_join(bent_site,by="site") %>%
  left_join(human,by="site")

write.csv(site,file=here("data/site.csv"))
write.csv(sp_site,file=here("data/sp_site.csv"))
```

### 2018-11-13 r/SummaryOrals.Rmd
11/12

--- 2018-04-26 SummaryOrals.Rmd
@@
 library(RColorBrewer)
-library(ggbiplot)
+# library(ggbiplot)
 library(nlme)

### 2018-11-13 r/Analysis.Rmd
11/12 adjustments

---
title: "Figures + Analysis"
author: "Molly Wilson"
date: "Dec. 2018"
output:
  html_document:
    toc: true
    toc_float: true
    theme: united
    number_sections: true
---

Outline of selected figures and analyses:

- Grazing differences between islands: (site or island level?)
-- ANOVAs to test statistical significance of differences here
- Total grazing impact of site-level biomass: measured vs. reported grazing rates (add line or color by feeding rate)
- Something looking at relationship with algal index vs. scarid biomass  
- SEM?

Notes: 
- re: lms/gamms: don't focus on these for now, the data has so many colinearities... if anything, use full PCA in model analysis
- Exclude Barbuda from this analysis...?

### 2018-11-13 r/Data_prodding.Rmd
11/12
11/13

@@
 ---
-title: "Scarid grazing analysis"
+title: "Investigation of scarid data"
 author: "Molly Wilson"
@@
 library(tidyverse)
-library(RColorBrewer)
+#library(RColorBrewer)
 library(corrplot)
+library(ggplot2)
+
 library(nlme)
-library(pls)
-library(factoextra)
+library(gamm4)
 ```
@@
   select(-X)
-
-# stop<-sum_id%>%filter(species_code=="stop")
-# qup<-sum_id%>%filter(species_code=="qup")
-# rbp<-sum_id%>%filter(species_code=="rbp")
 ```
@@
 ```{r,echo=FALSE}
-sample <- group_by(sum_id, site, species, phase) %>% summarize(n=n())
-barplot(table(sum_id$species, sum_id$site), main="Sample size by phase and species",
-   legend = rownames(counts), beside=TRUE, las=2)
+sample <- group_by(sum_id, site, species, phase, island) %>% summarize(n=n())
+
+ggplot(sum_id, aes(site)) +
+        facet_grid(island~species) +
+        geom_bar(aes(fill = phase), position = "dodge") +
+        scale_fill_manual(values = c("gray", "slategray2")) +
+        theme_minimal() +
+        theme(axis.text.x = element_text(angle = 90, hjust = 1))
+
+# playing with plot options...
+ggplot(data = sample, aes(x = site, y = n, fill = phase)) + 
+    geom_bar(stat = "identity", width = 1, position = "dodge") +
+    facet_wrap(~island, strip.position = "bottom", scales = "free_x") +
+    theme(panel.spacing = unit(0, "lines"), 
+         strip.background = element_blank(),
+         strip.placement = "outside") +
+    theme(axis.text.x = element_text(angle = 90, hjust = 1))
 ```
@@
 
-## Predictors -->
+## Predictors
 
-### All predictors -->
 ```{r}
-pca <- prcomp(na.omit(sum_site %>% 
+pca_pred <- prcomp(na.omit(sum_site %>% 
               remove_rownames %>% 
@@
               select(lc_cover,ta_cover,ma_canopy,ta_canopy,rugosity,scar_bm,scar_den,tot_bm)),
-  center = TRUE,scale. = TRUE)
-summary(pca)
+  center = T,scale = T)
 
-plot(pca,type="l")
+summary(pca_pred)
+plot(pca_pred,type="l")
 
-fviz_pca_biplot(pca,
-                col.var = "black",
-                col.ind = as.factor(sum_site$island), # color by groups
-                palette = c("#00AFBB", "#E7B800", "#FC4E07"),
-                addEllipses = T, ellipse.level=0.95,
-                ellipse.type = "confidence",
-                label = "var",
-                legend.title = "Groups",
-                repel = T
-                ) +
-  scale_color_brewer(palette="Dark")+
+autoplot(pca_pred, data = sum_site, colour = 'island',
+         loadings = T, loadings.colour = 'gray',
+         loadings.label = T, loadings.label.size = 3, loadings.label.colour = 'black', 
+            loadings.label.repel = T) +
+    scale_color_manual(name="island", values=c("navy", "darkseagreen", "slategray2")) + 
+    theme(legend.direction = 'horizontal', legend.position = 'top') +
     theme_minimal()
 
-fviz_mca_biplot(pca, 
-  habillage = sum_id$island, addEllipses = TRUE,
-  label = "var", shape.var = 15) +
-  scale_color_brewer(palette="Dark3")+
-  theme_minimal()
+# add PC1 and PC2 as variables in dataset
+sum_site_pca1 <- sum_site %>% 
+  bind_cols(as.data.frame(pca_pred$x) %>% 
+           select(PC1,PC2)) %>%
+  clean_names(case="old_janitor")
+sum_id_pca1 <- sum_id %>% left_join(select(sum_site_pca1,site,pc1,pc2), by = "site")
+write.csv(sum_id_pca1,file=here("data/sum_id_pca1.csv"))
 ```
 
-<!-- ### Benthic variables -->
-<!-- ```{r, echo=FALSE} -->
-<!-- pred.benthic <- sum_site %>% select(lc_cover,ta_cover,ma_canopy,ta_canopy,rugosity) # should it matter if I use site level summary data vs.
-full follow dataset (where site-level data has been joined?) -->
-<!-- pca1 <- prcomp(na.omit(pred.benthic),center = TRUE,scale. = TRUE)  -->
+## Response variables
 
-<!-- # add pc1 and pc2 as variables in dataset -->
-<!-- pca1.scores <- as.data.frame(pca1$x) %>% select(pc1,pc2) -->
-<!-- sum_site <- sum_site  %>% bind_cols(pca1.scores) -->
-<!-- pca1.scores <- select(sum_site,site,pc1,pc2) -->
-<!-- sum_id_pca <- sum_id %>% left_join(pca1.scores,by="site") -->
+```{r}
+pca_resp <- prcomp(na.omit(sum_id %>% 
+              select(g_frac,fr,br)), 
+  center = T,scale = T)
 
-<!-- write.csv(sum_id_pca,file = here("data/sum_id_pca.csv")) -->
-<!-- ``` -->
-<!-- ### Fish community variables -->
-<!-- ```{r} -->
-<!-- pred.fish <- sum_site %>% select(scar_bm,scar_den,carn_bm) # should it matter if I use site level summary data vs. full follow dataset (where site-level data has been joined?) -->
-<!-- pca2 <- prcomp(na.omit(pred.fish),center = TRUE,scale. = TRUE)  -->
-<!-- summary(pca2) -->
+summary(pca_resp)
+plot(pca_resp,type="l")
 
-<!-- plot(pca2,type="l") -->
-<!-- ggbiplot(pca2, obs.scale = 1, var.scale = 1, groups=sum_site$island, ellipse = TRUE, circle = TRUE, loadings.label.repel=TRUE) + scale_color_manual(name="island", values=c("navy", "darkseagreen", "slategray2")) + theme(legend.direction = 'horizontal', legend.position = 'top') + theme_minimal() -->
+autoplot(pca_resp, data = sum_id, colour = "island",
+         loadings = T, loadings.colour = 'gray',
+         loadings.label = T, loadings.label.size = 3, loadings.label.colour = 'black', 
+            loadings.label.repel = T) +
+    scale_color_manual(name="island", values=c("navy", "darkseagreen", "slategray2")) + 
+    theme(legend.direction = 'horizontal', legend.position = 'top') +
+    theme_minimal()
 
-<!-- ``` -->
-
-<!-- ## Response variables -->
-
-<!-- ### Grazing behaviors -->
-<!-- *log transform increases P1 explained variance here...* -->
-<!-- *Currently for_dur has several NA values (for individuals that didn't graze)...
-I am omitting that variable for now because I don't want to omit non-grazing individuals (instead would like to keep them as grazing=0* -->
-<!-- ```{r} -->
-<!-- response <- data %>% select(g_frac,tdb,br) -->
-<!-- log.vars5 <- log(response[,1:3]) -->
-<!-- pca3 <- prcomp(na.omit(response),center = TRUE,scale. = TRUE)  -->
-<!-- summary(pca3) -->
-<!-- plot(pca3,type="l") -->
-<!-- ggbiplot(pca3, obs.scale = 1, var.scale = 1, groups=data$island,shape=data$island,ellipse = TRUE, circle = TRUE, alpha=0.8,varname.size = 5) + theme_minimal() + scale_color_manual(name="island", values=c("navy", "darkseagreen", "slategray2")) + labs(title="Grazing behaviors by island") + theme(plot.title = element_text(hjust = 0.5)) -->
-<!-- ``` -->
-<!-- ```{r} -->
-<!-- response <- data %>% select(g_frac,fr,br) -->
-<!-- log.vars5 <- log(response[,1:3]) -->
-<!-- pca3 <- prcomp(na.omit(response),center = TRUE,scale. = TRUE)  -->
-<!-- summary(pca3) -->
-<!-- plot(pca3,type="l") -->
-<!-- ggbiplot(pca3, obs.scale = 1, var.scale = 1,ellipse = TRUE, circle = TRUE, alpha=0.4,varname.size = 5) + theme(legend.direction = 'horizontal', legend.position = 'top') + theme_minimal() +geom_point(aes(col=data$scar_bm)) + labs(col="Scarid Biomass (g/100m2)", title="Grazing Behaviors by Scarid Biomass") + theme(plot.title = element_text(hjust = 0.5)) -->
-<!-- ``` -->
-
-<!-- ```{r,echo=FALSE} -->
-<!-- vir <- data %>% filter(species_code=="stop") -->
-<!-- response <- vir %>% select(g_frac,fr,br) -->
-<!-- log.vars5 <- log(response[,1:3]) -->
-<!-- pcavir <- prcomp(na.omit(response),center = TRUE,scale. = TRUE)  -->
-<!-- ggbiplot(pcavir, obs.scale = 1, var.scale = 1,ellipse = TRUE, circle = TRUE, alpha=1,varname.size = 5) + theme(legend.direction = 'horizontal', legend.position = 'top') + theme_minimal() + labs(col="Scarid Biomass (g/100m2)", title="Grazing Behaviors by Scarid Biomass") + theme(plot.title = element_text(hjust = 0.5)) +geom_point(aes(colour=vir$scar_bm)) -->
-<!-- ``` -->
-<!-- ```{r,echo=FALSE} -->
-<!-- vet <- data %>% filter(species_code=="qup") -->
-<!-- response <- vet %>% select(g_frac,fr,br) -->
-<!-- log.vars5 <- log(response[,1:3]) -->
-<!-- pcavet <- prcomp(na.omit(response),center = TRUE,scale. = TRUE)  -->
-<!-- ggbiplot(pcavet, obs.scale = 1, var.scale = 1,ellipse = TRUE, circle = TRUE, alpha=1,varname.size = 5) + theme(legend.direction = 'horizontal', legend.position = 'top') + theme_minimal() + labs(col="Scarid Biomass (g/100m2)", title="Grazing Behaviors by Scarid Biomass") + theme(plot.title = element_text(hjust = 0.5)) +geom_point(aes(colour=vet$scar_bm)) -->
-<!-- ``` -->
-
+# autoplot doesn't like when I add it for_dur or for_bites (NA values?)
+```
 # Model trials
@@
 
-## Basic lm (to assess predictors of interest)
+## Basic lm 
 ```{r}
-initial<-data%>%filter(phase=="i") # for analysis with initial only
-lm_fr <- lm(fr~scar_bm*species+phase+carn_bm+length_cm+pc1+pc2, data=data)
-summary(lm_fr) # predicting feeding rate of initial phase only, all species
-#vif(lm_fr)
+lm_fr <- lm(fr~species+phase+length_cm+pc1+pc2, data=filter(sum_id_pca1, species_code!="rbp"))
+summary(lm_fr) 
 ```
-*Notes: nearly everything is significant?
-Significant interaction between scar_bm and species for Spar.
-viride.
-carn_bm and scar_bm are highly positively correlated (~0.86) - is it ok to include both?*
-*R2 goes up slightly with initial only*
-
-```{r}
-lm_g_frac <- lm(g_frac~species*scar_bm+phase+scar_den+carn_bm+phase+length_cm+pc1+pc2, data=data)
-summary(lm_g_frac)
-```
-*Notes: carn_bm and pc1 are strongest predictors of time spent grazing, scar_bm and phase are also significant. Slight significance of species, length, and species-scar_bm interaction*
 
@@
 ```{r}
-lme_fr <- lme(fr~phase+length_cm+species*scar_bm+scar_den+carn_bm+pc1+pc2,data=data,random=~1|island, na.action = na.omit)
+lme_fr <- lme(fr~phase+length_cm+species+pc1+pc2,data=filter(sum_id_pca1, species_code!="rbp"),random=~1|island, na.action = na.omit)
 summary(lme_fr)
 ```
-*Notes: nearly everything is significant except for pc2 and scar_den?*
-```{r}
-lme_g_frac <- lme(g_frac~phase+length_cm+species*scar_bm+scar_den+carn_bm+pc1+pc2,data=data,random=~1|island, na.action = na.omit)
-summary(lme_g_frac)
-```
-```{r}
-lme_fr_stop <- lme(fr~phase+length_cm+scar_bm+scar_den+carn_bm+pc1+pc2,data=stop,random=~1|island, na.action = na.omit)
-summary(lme_fr_stop)
-```
-*Notes: carn_bm and length are signifigant predictors*
-```{r}
-lme_fr_qup <- lme(fr~phase+length_cm+scar_bm+scar_den+carn_bm+pc1+pc2,data=qup,random=~1|island, na.action = na.omit)
-summary(lme_fr_qup)
-```
-*Notes: pc1 is strongest (magnitude) predictors. Also significant are phase, carn_bm and scar_den*
-# Model trials: species run individually
-
-## Feeding rates
-```{r}
-lme_fr_stop <- lme(fr~phase+length_cm+scar_bm+scar_den+carn_bm+pc1+pc2,data=stop,random=~1|island, na.action = na.omit)
-summary(lme_fr_stop)
-```
-*Notes: carn_bm and length are signifigant predictors*
-
-```{r}
-lme_fr_qup <- lme(fr~phase+length_cm+scar_bm+scar_den+carn_bm+pc1+pc2,data=qup,random=~1|island, na.action = na.omit)
-summary(lme_fr_qup)
-```
-*Notes: pc1 is strongest (magnitude) predictors. Also significant are phase, carn_bm and scar_den*
 
@@
 ```{r, echo=FALSE}
-gam<-gam(fr~species+phase+s(length_cm)+s(scar_bm)+s(pc1)+s(pc2), family=gaussian(link=identity),data=data, method="REML")
+gam<-gam(fr~species+phase+s(length_cm)+s(scar_bm)+s(pc1)+s(pc2), family=gaussian(link=identity),sum_id_pca1=data, method="REML")
 summary(gam)
 plot(gam,pages=1)
-```
-```{r, echo=FALSE}
-gamvet<-gam(fr~phase+s(length_cm)+s(scar_bm)+(carn_bm)+s(pc1)+s(pc2), family=gaussian(link=identity),data=vet, method="REML")
-summary(gamvet)
-plot(gamvet,pages=1)
-```
-```{r, echo=FALSE}
-vir<-data %>% filter(species_code=="stop" & site!="Snapper Point")
-gamvir<-gam(fr~phase+s(length_cm)+s(scar_bm)+s(pc1)+s(pc2), family=gaussian(link=identity),data=vir,method="REML")
-summary(gamvir)
-plot(gamvir,pages=1)
 ```
@@
 
-### Feeding rate (bites per hour)
 ```{r,message=FALSE,warning=FALSE}
-gamm<-gamm(fr~species+phase+s(length_cm)+s(scar_bm)+s(pc1)+s(pc2), family=gaussian(link=identity),data=data, random=list(island=~1),method="REML")
+gamm<-gamm(fr~phase+s(length_cm)+s(pc1)+s(pc2), family=gaussian(link=identity),data=filter(sum_id_pca1, species_code=="qup"), random=list(island=~1),method="REML")
 summary(gamm$gam)
@@
 ```{r,message=FALSE,warning=FALSE}
-gamm<-gamm(fr~phase+s(length_cm)+s(scar_bm)+s(pc1)+s(pc2), family=gaussian(link=identity),data=vet, random=list(island=~1),method="REML")
-summary(gamm$gam)
-plot(gamm$gam,pages=1)
-```
-```{r,message=FALSE,warning=FALSE}
-gamm<-gamm(fr~phase+s(length_cm)+s(scar_bm)+s(pc1)+s(pc2), family=gaussian(link=identity),data=vir, random=list(island=~1),method="REML")
+gamm<-gamm(fr~phase+s(length_cm)+s(pc1)+s(pc2), family=gaussian(link=identity),data=filter(sum_id_pca1, species_code=="stop"), random=list(island=~1),method="REML")
 summary(gamm$gam)
@@
 
-Most of variation is explained by species - for individual species models R2 values are very low (~0.3)
-
-### Fraction of time spent grazing
-```{r,message=FALSE,warning=FALSE}
-gamm<-gamm(g_frac~species+phase+s(length_cm)+s(scar_bm)+s(pc1)+s(pc2), family=gaussian(link=identity),data=data, random=list(island=~1),method="REML")
-summary(gamm$gam)
-plot(gamm$gam,pages=1)
-```
-```{r,message=FALSE,warning=FALSE}
-gamm<-gamm(g_frac~phase+s(length_cm)+s(scar_bm)+s(pc1)+s(pc2), family=gaussian(link=identity),data=vet, random=list(island=~1),method="REML")
-summary(gamm$gam)
-plot(gamm$gam,pages=1)
-```
-```{r,message=FALSE,warning=FALSE}
-gamm<-gamm(g_frac~phase+s(length_cm)+s(scar_bm)+s(pc1)+s(pc2), family=gaussian(link=identity),data=vir, random=list(island=~1),method="REML")
-summary(gamm$gam)
-plot(gamm$gam,pages=1)
-```
-
-Low R2 for all trials predicting g_frac (<0.4)(gammvir)
 ```

### 2018-11-14 r/Data_prodding.Rmd
11/13

@@
 # set up
+knitr::opts_chunk$set(echo = FALSE,
+                      message = FALSE,
+                      warning = FALSE)
 library(startR)
@@
 library(tidyverse)
-#library(RColorBrewer)
 library(corrplot)
 library(ggplot2)
-
+library(RColorBrewer)
+library(ggfortify)
+library(ggsignif) # boxplot stats
+library(EnvStats) # boxplot stats
+library(car)
 library(nlme)
@@
 sum_site <- read.csv(here("data","sum_site.csv"), stringsAsFactors=F) %>%
-  select(-X)
+  select(-X) %>%
+  mutate(island = factor(island, levels = c("Bonaire","Antigua","Barbuda"))) %>%
+  arrange(island)
 sum_id <- read.csv(here("data","sum_id.csv"), stringsAsFactors=F) %>%
-  select(-X)
+  select(-X) %>%
+  mutate(island = factor(island, levels = c("Bonaire","Antigua","Barbuda"))) %>%
+  arrange(island) %>%
+  filter(length_cm>=15)
 ```
@@
 - Follows targeted *Sparisoma viride*, *Scarus vetula*, and *Sparisoma aurofrenatum* of both initial and terminal phases 
-  +*Sparisoma aurofrenatum* were not followed in Bonaire, but added in Antigua and Bonaire due to low abundances of other two species.
+  + *Sparisoma aurofrenatum* were not followed in Bonaire, but added in Antigua and Bonaire due to low abundances of other two species.
 Some sites in Antigua and Barbuda had no/low represenation from terminal phase *viride* and *vetula*
+  + Most eventual analyses will likely focus exclusively on initial phase (standardized size window) *S. viride* and *S. vetula*
 - site level factors (benthic, fish, and rugosity) assessed at each of the 13 sites
 
-```{r,echo=FALSE}
-sample <- group_by(sum_id, site, species, phase, island) %>% summarize(n=n())
+```{r}
+sample <- group_by(sum_id, site, species, species_code, phase, island) %>% summarize(n=n())
 
-ggplot(sum_id, aes(site)) +
-        facet_grid(island~species) +
-        geom_bar(aes(fill = phase), position = "dodge") +
-        scale_fill_manual(values = c("gray", "slategray2")) +
-        theme_minimal() +
+ggplot(sample, aes(x=factor(site, levels = sum_site$site), y=n)) +
+        facet_grid(phase~species) +
+        geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
+        scale_fill_manual(values = c("gray", "slategray2", "black")) +
+        labs(title="Total sample") +
+        theme_bw() +
         theme(axis.text.x = element_text(angle = 90, hjust = 1))
+```
+```{r}
+ggplot(filter(sample, species_code!="rbp" & phase == "i"),
+      aes(x=factor(site, levels = sum_site$site), y=n)) +
+        facet_grid(.~species) +
+        geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
+        scale_fill_manual(values = c("gray", "slategray2", "black")) +
+        labs(title="Sample subset") +
+        theme_bw() +
+        theme(axis.text.x = element_text(angle = 90, hjust = 1))
+```
 
-# playing with plot options...
-ggplot(data = sample, aes(x = site, y = n, fill = phase)) + 
-    geom_bar(stat = "identity", width = 1, position = "dodge") +
-    facet_wrap(~island, strip.position = "bottom", scales = "free_x") +
-    theme(panel.spacing = unit(0, "lines"), 
-         strip.background = element_blank(),
-         strip.placement = "outside") +
-    theme(axis.text.x = element_text(angle = 90, hjust = 1))
+Need to determine appropriate size range for comparison.
+Because they are not evenly distributed (i.e.
+larger fish in Bonaire, smaller in Barbuda), I will likely want to compare length-feeding relationships as opposed to pooled averages
+
+```{r}
+ggplot(filter(sum_id, species_code!="rbp" & phase == "i" & length_cm<=30 & length_cm>=15), aes(x = length_cm)) +
+  facet_grid(.~species) +
+  geom_density(aes(fill = island), alpha=0.4) +
+  labs(title="Size distributions within sample subset") +
+  theme_bw() 
 ```
@@
 
-## Site-level predictor variables: fish, benthic, and rugosity
+## Predictor variables
+Potential predictor variables are site-level fish, benthic, and rugosity values.
+These are likely correlated to one another, and I need to determine which ones I ultimately want to use (if modeling behavioral responses via any multivariate regressions).
+I can also move to SEM if I want to keep multiple correlated predictors.
 
-### Distributions
+First, check distribution of predictor variables of interest: not very normally distributed...
 ```{r,warning=FALSE,echo=FALSE}
-predictors <- sum_site %>% 
-  select(tot_bm,tot_den,scar_bm,scar_den,carn_bm,carn_den,cca_cover,lc_cover,ma_cover,ma_canopy,ta_cover,ta_canopy,rugosity) %>% 
+pred <- sum_site %>% 
+  select(tot_bm,tot_den,scar_bm,scar_den,carn_bm,carn_den,
+         cca_cover,lc_cover,ma_cover,ma_canopy,ta_cover,ta_canopy,rugosity) %>% 
   gather() %>% 
@@
 
-ggplot(predictors,aes(value)) +
+ggplot(pred,aes(value)) +
   facet_wrap(~ key, scales = "free") +
@@
 ```
-Site-level predictors are not normally distributed....
 
-### Correlations
 ```{r,echo=FALSE}
-cor = cor(sum_id %>% 
-  select(tot_bm,tot_den,scar_bm,scar_den,consp_bm,consp_den,
+pred <- sum_id %>% select(tot_bm,tot_den,scar_bm,scar_den,consp_bm,consp_den,
                            carn_bm,carn_den,cca_cover,lc_cover,
-         ma_cover, ma_canopy,ta_cover,ta_canopy,rugosity), 
-  use="pairwise.complete.obs")
-corplot <- corrplot(cor,title = "Correlation Plot", method = "square", outline = T, addgrid.col = "darkgray", order="hclust", mar = c(4,0,4,0), addrect = 4, rect.col = "black", rect.lwd = 5,cl.pos = "b", tl.col = "indianred4", tl.cex = .5, cl.cex = .5)
-
+                           ma_cover, ma_canopy,ta_cover,ta_canopy,rugosity)
+c = cor(pred, use="pairwise.complete.obs")
+corplot <- corrplot(c, type = "upper", order = "hclust")
 ```
-
-Many correlations here, both within benthic/fish categories and across categories.
-Rugosity is correlated with both fish and benthic variables.
-I will examine correlations within fish community and benthic community variables to decide which I can omit and which I will integrate through pcA.
-Whether I include benthic variables as predictor or response variables will determine if I need to worry about fish~benthic correlations.
 
@@
 }
-pairs(~tot_bm+tot_den+scar_bm+scar_den+consp_bm+consp_den+carn_bm+carn_den+rugosity, data=sum_id,
+pairs(~consp_l+scar_bm+rugosity+ma_ind+ta_canopy+ta_cover, data=sum_id,
       lower.panel=panel.smooth, upper.panel=panel.cor, 
-      pch=20, main="Fish Community Variables", na.action = na.omit)
+      pch=20, main="Selected Fish & Benthic Community Variables", na.action = na.omit)
 ```
 
-Variables carn_den and carn_bm are essentially identical - I will remove carn_den and use carn_bm as the main preditor metric.
-Total fish biomass and density (tot_bm and tot_den) are highly correlated with both scarid and carnivore variables, so I will remove them.
-conspific biomass is highly correlated with scarid biomass so I will remove it as well (and should possibly also remove consp_den?).
-The selected variables I will proceed with are as follows:
-```{r,echo=FALSE}
-pairs(~scar_bm+scar_den+consp_den+carn_bm+rugosity, data=sum_id,
-      lower.panel = panel.smooth, upper.panel = panel.cor, 
-      pch=20, main = "Selected Fish Community Variables", na.action = na.omit)
+Variable selection notes:
+- excluding both carnivore variables as they are highly correlated with scarid biomass and total biomass, eventually I could make these more nuanced by distinguishing actual predators, but right now I don't think it reflects actual predator populations of >15cm parrotfish
+- rugosity is highly correlated with turf cover, and scarid density
+- scarid density: removing for now, because I think it was a bit skewed from Barbuda juveniles
+- could eventually use consp. scarid length as another indicator of overfishing?
+
+```{r}
+pca_pred <- prcomp(na.omit(sum_site %>% 
+              remove_rownames %>% 
+              column_to_rownames(var="site") %>%
+              select(ma_ind,ta_cover,ta_canopy,rugosity,scar_den,scar_bm)), 
+  center = T,scale = T)
+
+summary(pca_pred)
+plot(pca_pred,type="l")
+
+autoplot(pca_pred, data = sum_site, colour = 'island',
+         # frame = T, frame.type = "norm",
+         loadings = T, loadings.colour = 'gray',
+         loadings.label = T, loadings.label.size = 3, loadings.label.colour = 'black', 
+            loadings.label.repel = T) +
+    scale_color_manual(name="island", values=c("lightblue", "darkseagreen", "lightsalmon")) + 
+    theme(legend.direction = 'horizontal', legend.position = 'top') +
+    theme_minimal()
+
+# add PC1 and PC2 as variables in dataset
+sum_site_pca1 <- sum_site %>% 
+  bind_cols(as.data.frame(pca_pred$x) %>% 
+           select(PC1,PC2)) %>%
+  janitor::clean_names(case="old_janitor")
+sum_id_pca1 <- sum_id %>% left_join(select(sum_site_pca1,site,pc1,pc2), by = "site")
+write.csv(sum_id_pca1,file=here("data/sum_id_pca1.csv"))
 ```
 
-Scarid biomass, carnivore biomass, and rugosity should be integrated via pcA. Scarid density and conspific density should also be combined, or one should be omitted.
-
-```{r,echo=FALSE}
-pairs(~cca_cover+lc_cover+ma_cover+ma_canopy+ta_cover+ta_canopy+rugosity, data = sum_id,
-      lower.panel = panel.smooth, upper.panel = panel.cor, 
-      pch=20, main ="Benthic Community Variables", na.action = na.omit)
-```
-
-All benthic variables are highly correlated.
-I can omit cca_cover and ma_cover (accounted for by ma_canopy).
-The rest should be combined via pcA.
-First I will assess remaining correlations between selected benthic and fish community variables:
-
-```{r,echo=FALSE}
-pairs(~scar_bm+scar_den+carn_bm+lc_cover+ta_cover+ma_canopy+ta_canopy+rugosity, data=sum_id,
-      lower.panel=panel.smooth, upper.panel=panel.cor, 
-      pch=20, main="Selected Fish and Benthic Variables",na.action = na.omit)
-```
-
-Unsure how to proceed here if I do end up using both benthic and fish community variables as predictors, because some are corellated (esp.
-scarid_bm.s and all benthic variables....).
-It looks like I would integrate scar_bm and carn_bm into a pcA with all other benthic/rugosity variables.
-Variables scar_den and consp_den are relatively independent of all other variables (except each other).
-
-Current outline for site-level pcA tasks:
-If benthic=predictor:
-- lc_cover + ta_cover + ma_canopy + ta_canopy + rugosity + scar_bm + carn_bm
-- scar_den + consp_den
-If benthic=response:
-- rugosity + scar_bm + carn_bm
-- lc_cover + ta_cover + ma_canopy + ta_canopy
-- scar_den + consp_den
-
-## Fish-level response variables: grazing behaviors
+## Response variables
+Fish-level grazing behaviors (as well as competitive interaction frequency)
 
@@
 
-### Corellations
-```{r,echo=FALSE}
-cor=cor(sum_id %>% 
-          select(g_frac,fr,br,for_dur,for_bites), use="pairwise.complete.obs")
-corplot <- corrplot(cor,title = "Correlation Plot", method = "square", outline = T, addgrid.col = "darkgray", order="hclust", mar = c(4,0,4,0), addrect = 4, rect.col = "black", rect.lwd = 5,cl.pos = "b", tl.col = "indianred4", tl.cex = .5, cl.cex = .5)
-```
-
 ```{r,echo=FALSE}
@@
 
-Variable for_bites is correlated with nearly all other variables, so will be omitted.
-Variables g.farc, fr.
-and for_dur are still correlated and should be assessed with a pcA.
-Variable br can be a standalone variable.
+Variable selection notes:
+- for_bites is correlated with fr and for_dur, but I will play around with keeping it for now.
 
-Grazing pcA:
-- g_frac + fr + for_dur
-- (br as standalone)
+# Grazing boxplots
+
+```{r, message=FALSE,warning=FALSE,echo=FALSE}
+ggplot(filter(sum_id, 
+              phase=="i" & length_cm>=16 & length_cm<=24 & species_code!="rbp"), 
+       aes(x=island, y=fr)) + 
+  geom_boxplot() +
+  stat_summary(fun.y = mean, geom="point", shape=5, size=3) + 
+  labs(y="Feeding Rate (bites/hr)",x="",title="Hourly feeding rates") +
+  theme_bw() + 
+  theme(plot.title = element_text(hjust = 0.5)) + 
+  geom_signif(comparisons = list(c("Antigua", "Barbuda"),c("Antigua","Bonaire"),c("Barbuda","Bonaire")), map_signif_level=TRUE) +
+  facet_grid(.~species) + 
+  theme(strip.text = element_text(face = "italic")) + 
+  stat_n_text()
+
+ggplot(filter(sum_id, 
+              phase=="i" & length_cm>=16 & length_cm<=24 & species_code!="rbp"), 
+       aes(factor(island), fr)) + 
+  geom_violin(aes(fill=island)) +
+  geom_boxplot(width=0.1) + 
+  labs(y="Feeding Rate (bites/hr)",x="",title="Hourly feeding rates (init. 16-24cm") +
+  geom_signif(comparisons = list(c("Antigua", "Barbuda"),c("Antigua","Bonaire"),c("Barbuda","Bonaire")), map_signif_level=TRUE) +
+  facet_grid(.~species) + 
+  stat_n_text() +
+  scale_fill_brewer(palette="Blues") +
+  theme_bw() +
+  theme(strip.text = element_text(face = "italic")) +
+  theme(plot.title = element_text(hjust = 0.5))   
+```
 
@@
 ## Grazing as a function of fish length
-```{r, echo=FALSE,warning=FALSE, eval=FALSE}
-ggplot(sum_id, aes(x = length_cm, y = fr, colour = island)) +
+
+```{r}
+ggplot(filter(sum_id, species_code != "rbp" & island != "Barbuda"), aes(x = length_cm, y = fr, colour = island)) +
   facet_grid(.~species) +
@@
   geom_smooth(method = lm, fullrange = FALSE) +
-  ggtitle("Feeding rate by fish length") + 
-  labs(y = "Feeding Rate (bites/hr)", x = "Length (cm)") + 
-  theme_minimal() + 
+  labs(y = "Feeding Rate (bites/hr)", x = "Length (cm)", title = "Hourly feeding rate") + 
+  theme_bw() + 
   theme(plot.title = element_text(hjust = 0.5)) + 
@@
 ```
-```{r, echo=FALSE,warning=FALSE, eval=FALSE}
-ggplot(sum_id, aes(x = length_cm, y = br, colour = island)) +
+
+```{r}
+ggplot(filter(sum_id, species_code != "rbp" & island != "Barbuda" & site!="Turtle Bay"), aes(x = length_cm, y = fr, colour = site)) +
   facet_grid(.~species) +
@@
   geom_smooth(method = lm, fullrange = FALSE) +
-  ggtitle("Feeding rate by fish length") + 
-  labs(y = "Bite Rate (bites/sec)", x = "Length (cm)") + 
-  theme_minimal() + 
-  theme(plot.title = element_text(hjust = 0.5)) + 
-  scale_color_manual(values = c("navy", "darkseagreen", "slategray2"))
+  labs(y = "Feeding Rate (bites/hr)", x = "Length (cm)", title = "Hourly feeding rate") + 
+  theme_bw() + 
+  theme(plot.title = element_text(hjust = 0.5))
 ```
-```{r, echo=FALSE,warning=FALSE, eval=FALSE}
-ggplot(sum_id, aes(x = length_cm, y = g_frac, colour = island)) +
+
+```{r}
+ggplot(filter(sum_id, species_code != "rbp" & island != "Barbuda"), aes(x = length_cm, y = br, colour = island)) +
   facet_grid(.~species) +
@@
   geom_smooth(method = lm, fullrange = FALSE) +
-  ggtitle("Feeding rate by fish length") + 
-  labs(y = "Time Fraction", x = "Length (cm)") + 
-  theme_minimal() + 
+  labs(y = "Bite Rate (bites/sec)", x = "Length (cm)", title = "Bite rate while grazing") + 
+  theme_bw() + 
   theme(plot.title = element_text(hjust = 0.5)) + 
@@
 
-## Grazing as a function of scarid biomass
+*Note: remove G1 grazing instances here?*
 
 ```{r}
-ggplot(sum_id %>% filter(island!="Barbuda"), aes(x = scar_bm,y=fr)) +
+ggplot(filter(sum_id, species_code != "rbp" & island != "Barbuda"), aes(x = length_cm, y = g_frac, colour = island)) +
+  facet_grid(.~species) +
+  geom_point(shape = 18) +
+  scale_shape(solid = FALSE) +
+  geom_smooth(method = lm, fullrange = FALSE) +
+  labs(y = "Time Fraction", x = "Length (cm)", title = "Fraction of time spent grazing") + 
+  theme_bw() + 
+  theme(plot.title = element_text(hjust = 0.5)) + 
+  scale_color_manual(values = c("navy", "darkseagreen", "slategray2"))
+```
+
+## Grazing as a function of site traits
+
+```{r}
+ggplot(filter(sum_id, species_code != "rbp" & island != "Barbuda" & length_cm>=16 & length_cm<=24), aes(x = scar_bm,y=fr)) +
   facet_grid(.~species) +
@@
   geom_point(shape=18) +
-  ggtitle("Feeding rate by scarid biomass") +
-  labs(y="Feeding rate (bites/hr)", x="Scarid Biomass (g/100m2)") +
-  theme_minimal() + 
+  labs(y="Feeding rate (bites/hr)", x="Scarid Biomass (g/100m2)", title = "Hourly feeding rate") +
+  theme_bw() + 
   theme(plot.title = element_text(hjust = 0.5))
 ```
+
 ```{r,echo=FALSE,warning=FALSE}
-ggplot(sum_id, aes(x = scar_bm, y = fr, colour = island)) +
+ggplot(filter(sum_id, species_code != "rbp" & island != "Barbuda" & length_cm>=16 & length_cm<=24), aes(x = scar_bm, y = fr, colour = island)) +
   facet_grid(. ~ species) +
@@
   geom_smooth(method = lm, fullrange=F) +
-  ggtitle("Feeding rate by scarid biomass") +
-  labs(y = "Feeding rate (bites/hr)", x = "Scarid Biomass (g/100m2)") +
-  theme_minimal() + 
+  labs(y = "Feeding rate (bites/hr)", x = "Scarid Biomass (g/100m2)", title = "Hourly feeding rate") +
+  theme_bw() + 
   theme(plot.title = element_text(hjust = 0.5)) +
@@
 ```{r,echo=FALSE,warning=FALSE}
-ggplot(sum_id, aes(x = scar_bm, y = fr, colour = phase)) +
+ggplot(filter(sum_id, species_code != "rbp" & island != "Barbuda"), aes(x = scar_bm, y = fr, colour = phase)) +
   facet_grid(. ~ species) +
@@
   geom_smooth(method = lm, fullrange=F) +
-  ggtitle("Feeding rate by phase and scarid biomass") +
-  labs(y = "Feeding rate (bites/hr)", x = "Scarid Biomass (g/100m2)") +
-  theme_minimal() + 
+  labs(y = "Feeding rate (bites/hr)", x = "Scarid Biomass (g/100m2)", title = "Hourly feeding rate") +
+  theme_bw() + 
   theme(plot.title = element_text(hjust = 0.5)) +
@@
 ```
-```{r,echo=FALSE,warning=FALSE}
-ggplot(sum_id, aes(x = scar_bm, y = br, colour = phase)) +
-  facet_grid(. ~ species) +
-  geom_point(shape = 18) +
-  geom_smooth(method = lm, fullrange=F) +
-  ggtitle("Bite rates (while grazing) by scarid biomass") +
-  labs(y = "Bite rate (bites/sec)", x = "Scarid Biomass (g/100m2)") +
-  theme_minimal() + 
-  theme(plot.title = element_text(hjust = 0.5)) +
-  scale_color_manual(values = c("gray40", "aquamarine3"))
-```
-```{r,echo=FALSE,warning=FALSE}
-ggplot(sum_id, aes(x = scar_bm, y = g_frac, colour = phase)) +
-  facet_grid(. ~ species) +
-  geom_point(shape = 18) +
-  geom_smooth(method = lm, fullrange=F) +
-  ggtitle("Fraction of time spent grazing by phase and scarid biomass") +
-  labs(y = "Time Fraction", x = "Scarid Biomass (g/100m2)") +
-  theme_minimal() + 
-  theme(plot.title = element_text(hjust = 0.5)) +
-  scale_color_manual(values = c("gray40", "aquamarine3"))
-```
-```{r,echo=FALSE,warning=FALSE}
-ggplot(sum_id, aes(x = scar_bm, y = for_bites, colour = phase)) +
-  facet_grid(. ~ species) +
-  geom_point(shape = 18) +
-  geom_smooth(method = lm, fullrange=F) +
-  ggtitle("Bites per foray by phase and scarid biomass") +
-  labs(y = "Bites", x = "Scarid Biomass (g/100m2)") +
-  theme_minimal() + 
-  theme(plot.title = element_text(hjust = 0.5)) +
-  scale_color_manual(values = c("gray40", "aquamarine3"))
+
+Notes: 
+- scarid biomass is not the best predictor once I account for differences between my samples in terms of the sizes of fish I was sampling. I think the grazing/length relationships are much stronger.  
+- restraining sample size to length windows lowers sample size and makes trends much less pronounced - esp. for phase differences  
+- reducing sample to individual phase only also blurs trends
+
+```{r}
+ggplot(filter(sum_id, species_code != "rbp" & island != "Barbuda" & length_cm>=16 & length_cm<=24), aes(x = ma_ind,y=fr)) +
+  facet_grid(.~species) +
+  geom_smooth(method="loess") +
+  geom_point(shape=18) +
+  labs(y="Feeding rate (bites/hr)", x="Macroalgal index", title = "Hourly feeding rate") +
+  theme_bw() + 
+  theme(plot.title = element_text(hjust = 0.5))
 ```
@@
 ```{r,echo=FALSE,warning=FALSE}
-ggplot(sum_site,aes(x=scar_bm,y=int_scar.mean, shape=island)) +
+ggplot(sum_site,aes(x=scar_bm,y=int_scar_mean, shape=island)) +
   geom_point(size=3) +
@@
   labs(x="Scarid Biomass (g/100m2)", y="Interactions with other scarids (per 2 min follow)") +
-  geom_errorbar(aes(ymin=int_scar.mean-int_scar.se, ymax=int_scar.mean+int_scar.se))
-```
-```{r,echo=FALSE,warning=FALSE}
-ggplot(sum_site,aes(x=scar_bm,y=int_scar.mean_stop, shape=island)) +
-  geom_point(size=3) +
-  ggtitle("S. viride interaction frequency as a function of scarid biomass") +
-  theme_minimal() +
-  theme(plot.title = element_text(hjust = 0.5)) +
-  labs(x = "Scarid Biomass (g/100m2)", y = "Interactions with other scarids (per 2 min follow)") +
-  geom_errorbar(aes(ymin=int_scar.mean_stop-int_scar.se_stop, ymax=int_scar.mean_stop+int_scar.se_stop))
-```
-```{r,echo=FALSE,warning=FALSE}
-ggplot(sum_site,aes(x=scar_bm,y=int_scar.mean_qup, shape=island)) +
-  geom_point(size=3) +
-  ggtitle("S. vetula interaction frequency as a function of scarid biomass") + 
-  theme_minimal() + 
-  theme(plot.title = element_text(hjust = 0.5)) + 
-  labs(x="Scarid Biomass (g/100m2)", y="Interactions with other scarids (per 2 min follow)") + 
-  geom_errorbar(aes(ymin=int_scar.mean_qup-int_scar.se_qup, ymax=int_scar.mean_qup+int_scar.se_qup))
-```
-```{r,echo=FALSE,warning=FALSE}
-ggplot(sum_id,aes(x=scar_bm,y=int_scar)) +
-  facet_grid(.~species)+geom_point(shape=19) +
-  geom_smooth(method=lm) +
-  ggtitle("Competitive interactions as a function of scarid biomass") + 
-  theme_minimal() + 
-  theme(plot.title = element_text(hjust = 0.5))
+  geom_errorbar(aes(ymin=int_scar_mean-int_scar_se, ymax=int_scar_mean+int_scar_se))
 ```
 
-# PCAs -->
-*Do I need to normalize my variables pre-pcA? Currently just scaled, except for grazing response variables which are log-transformed* -->
-
-## Predictors
-
-```{r}
-pca_pred <- prcomp(na.omit(sum_site %>% 
-              remove_rownames %>% 
-              column_to_rownames(var="site") %>%
-              select(lc_cover,ta_cover,ma_canopy,ta_canopy,rugosity,scar_bm,scar_den,tot_bm)),
-  center = T,scale = T)
-
-summary(pca_pred)
-plot(pca_pred,type="l")
-
-autoplot(pca_pred, data = sum_site, colour = 'island',
-         loadings = T, loadings.colour = 'gray',
-         loadings.label = T, loadings.label.size = 3, loadings.label.colour = 'black', 
-            loadings.label.repel = T) +
-    scale_color_manual(name="island", values=c("navy", "darkseagreen", "slategray2")) + 
-    theme(legend.direction = 'horizontal', legend.position = 'top') +
-    theme_minimal()
-
-# add PC1 and PC2 as variables in dataset
-sum_site_pca1 <- sum_site %>% 
-  bind_cols(as.data.frame(pca_pred$x) %>% 
-           select(PC1,PC2)) %>%
-  clean_names(case="old_janitor")
-sum_id_pca1 <- sum_id %>% left_join(select(sum_site_pca1,site,pc1,pc2), by = "site")
-write.csv(sum_id_pca1,file=here("data/sum_id_pca1.csv"))
-```
-
-## Response variables
-
-```{r}
-pca_resp <- prcomp(na.omit(sum_id %>% 
-              select(g_frac,fr,br)), 
-  center = T,scale = T)
-
-summary(pca_resp)
-plot(pca_resp,type="l")
-
-autoplot(pca_resp, data = sum_id, colour = "island",
-         loadings = T, loadings.colour = 'gray',
-         loadings.label = T, loadings.label.size = 3, loadings.label.colour = 'black', 
-            loadings.label.repel = T) +
-    scale_color_manual(name="island", values=c("navy", "darkseagreen", "slategray2")) + 
-    theme(legend.direction = 'horizontal', legend.position = 'top') +
-    theme_minimal()
-
-# autoplot doesn't like when I add it for_dur or for_bites (NA values?)
-```
 # Model trials
@@
 ```{r}
+# Testing VIF with individual variables of interest
+lm_fr <- lm(fr~species+phase+length_cm+scar_bm+consp_den+ma_canopy+ta_cover+ta_canopy, data=filter(sum_id_pca1, species_code!="rbp"))
+summary(lm_fr) 
+vif(lm_fr)
+```
+
+Only VIF above 5 is rugosity
+```{r}
+# Using PC values as predictors
 lm_fr <- lm(fr~species+phase+length_cm+pc1+pc2, data=filter(sum_id_pca1, species_code!="rbp"))
 summary(lm_fr) 
+vif(lm_fr)
 ```
@@
 ```{r, echo=FALSE}
-gam<-gam(fr~species+phase+s(length_cm)+s(scar_bm)+s(pc1)+s(pc2), family=gaussian(link=identity),sum_id_pca1=data, method="REML")
+gam<-gam(fr~species+phase+s(length_cm)+s(scar_bm)+s(pc1)+s(pc2), family=gaussian(link=identity),data=sum_id_pca1, method="REML")
 summary(gam)
@@
 
-```{r,message=FALSE,warning=FALSE}
+```{r,eval=F}
 gamm<-gamm(fr~phase+s(length_cm)+s(pc1)+s(pc2), family=gaussian(link=identity),data=filter(sum_id_pca1, species_code=="qup"), random=list(island=~1),method="REML")
@@
 ```
-```{r,message=FALSE,warning=FALSE}
+```{r,eval=F}
 gamm<-gamm(fr~phase+s(length_cm)+s(pc1)+s(pc2), family=gaussian(link=identity),data=filter(sum_id_pca1, species_code=="stop"), random=list(island=~1),method="REML")

### 2018-11-14 r/SEM.Rmd
11/13

@@
 # set up
-library(rmarkdown)
-library(dplyr)
 library(tidyr)
@@
 ```{r,message=FALSE,warning=FALSE, echo=FALSE}
-setwd("~/github/Scarid-Behavior/data")
-site.sum <- read.csv("site.sum.csv", stringsAsFactors=FALSE) %>% select(-X)
-data <- read.csv("data2.csv", stringsAsFactors=FALSE)
-# add aglal indeces
-data <- data %>% mutate(index_MA = cover_MA*canopy_MA) %>% mutate(index_TA = cover_TA*canopy_TA)
+sum_id <- read.csv(here("data","sum_id.csv"), stringsAsFactors=FALSE) %>% select(-X)
+sum_site <- read.csv(here("data","sum_site.csv"), stringsAsFactors=FALSE) %>% select(-X)
+
 # recode phase and species with dummy variables
-data <- data %>% mutate(spec.dum= 
-                          ifelse(Species.Code=="stop", 1, 
-                          (ifelse(Species.Code=="qup", 2,
-                          (ifelse(Species.Code=="rbp",3,NA)))))) %>%
-  mutate(phase.dum=ifelse(Phase=="i",1,2))
+# data <- data %>% mutate(spec.dum= 
+#                           ifelse(Species.Code=="stop", 1, 
+#                           (ifelse(Species.Code=="qup", 2,
+#                           (ifelse(Species.Code=="rbp",3,NA)))))) %>%
+#   mutate(phase.dum=ifelse(Phase=="i",1,2))
 
@@
 # define model pathways
-model <- '
-    FR ~ Length.cm + phase.dum + spec.dum+ int.scar + RISK + NUTRITION
-    int.scar ~ scar.BM + cover_MA + cover_TA
-    rugosity ~ cover_LC
-    cover_LC ~ cover_MA + cover_TA
-    cover_TA ~ scar.BM
-    cover_MA ~ scar.BM
+qup_df <- sum_id %>% 
+  select(scar_bm, ma_cover, ta_cover, fr, g_frac, rugosity, int_scar, length_cm, lc_cover) %>% 
+  na.omit()
 
-    RISK =~ scar.BM + rugosity
-    NUTRITION =~ cover_MA + cover_TA
+sem <- '
+  # latent variable indicators  
+      risk =~ scar_bm + rugosity
+      nutrition =~ ma_cover + ma_canopy + ta_cover + ta_canopy
+  # regressions
+      fr ~ length_cm + risk + nutrition
+      rugosity ~ lc_cover
+      lc_cover ~ ma_cover + ta_cover
+      ta_cover ~ scar_bm
+      ma_cover ~ scar_bm
 '
 
-fit <- sem(model, data=data)
+sem1 <- '
+  # latent variable indicators
+      risk =~ scar_bm      
+      nutrition =~ ma_cover + ta_cover
+  # regressions
+      fr ~ length_cm + nutrition + risk
+      lc_cover ~ ma_cover + ta_cover
+'
+
+fit <- sem(sem1, data=qup_df)
 varTable(fit)
@@
 semPaths(fit)
+fitMeasures(fit, c("cfi","rmsea","srmr"))
 ```

### 2018-11-14 r/Sites.Rmd
11/13

@@
 
-Notes 11/11:
-* figure out dcast issues
-* replacing NA/0s within pipe
+Wrangling of fish, benthic, and rugosity data from Antigua, Barbuda, and Bonaire to put in usable site-level format.
 
@@
   summarize() # all possible site/transect combinations
-exp_sp <- fish %>% expand(nesting(species,family),site) # all possible species/site combinations
+exp_sp <- fish %>% tidyr::expand(nesting(species,family),site) # all possible species/site combinations
 template_sp <- full_join(exp_sp,exp_tran,by="site") # species/site/transect combinations
@@
 # Scarids by phase (should this be by species, too?)
-exp_ph <- fish %>% expand(phase,site)
+exp_ph <- fish %>% tidyr::expand(phase,site)
 template_ph <- full_join(exp_ph,exp_tran,by="site") # all site/transect/phase combinations
@@
   summarize()
-exp_cat <- ab_benthic %>% expand(nesting(code),site)
+exp_cat <- ab_benthic %>% tidyr::expand(nesting(code),site)
 template_bent <- full_join(exp_meter,exp_cat,by="site")
@@
   ) %>% 
-  janitor::clean_names(case="old_janitor")
+  janitor::clean_names(case="old_janitor") %>%
+  select(cca_cover, cca_cover_se, lc_cover, lc_cover_se, ma_cover, ma_cover_se, ma_canopy, ma_canopy_se, ta_cover, ta_cover_se, ta_canopy, ta_canopy_se)
 
@@
 
-bent_site <- bent_site %>% left_join(rug_site,by="site") %>% bind_rows(bon_benthic)
+bent_site <- bent_site %>% 
+  left_join(rug_site,by="site") %>% 
+  bind_rows(bon_benthic) %>%
+  mutate(ma_ind = ma_canopy*ma_cover,
+         ta_ind = ta_canopy*ta_cover)
 ```

### 2018-11-15 r/Analysis.Rmd
11/14

@@
 - Exclude Barbuda from this analysis...?
+
+```{r}
+# set up
+knitr::opts_chunk$set(echo = FALSE,
+                      message = FALSE,
+                      warning = FALSE)
+library(startR)
+library(here)
+library(tidyverse)
+library(corrplot)
+library(ggplot2)
+library(RColorBrewer)
+library(ggfortify)
+library(ggsignif) # boxplot stats
+library(EnvStats) # boxplot stats
+library(car)
+library(nlme)
+library(gamm4)
+```
+```{r}
+sum_site <- read.csv(here("data","sum_site.csv"), stringsAsFactors=F) %>%
+  select(-X) %>%
+  mutate(island = factor(island, levels = c("Bonaire","Antigua","Barbuda"))) %>%
+  arrange(island)
+sum_id <- read.csv(here("data","sum_id.csv"), stringsAsFactors=F) %>%
+  select(-X) %>%
+  mutate(island = factor(island, levels = c("Bonaire","Antigua","Barbuda"))) %>%
+  arrange(island) %>%
+  filter(length_cm>=15)
+fish <- read.csv(here("data","fish.csv"), stringsAsFactors=F) %>%
+  select(-X)
+```
+
+# Island - level differences in grazing rates
+
+```{r}
+# Violin plot of inter-island differences
+ggplot(filter(sum_id, 
+              phase=="i" & length_cm>=16 & length_cm<=24 & species_code!="rbp"), 
+       aes(factor(island), fr)) + 
+  geom_violin(aes(fill=island)) +
+  geom_boxplot(width=0.1) + 
+  labs(y = "Feeding Rate (bites/hr)", x = "", title = "Hourly feeding rates (i. phase, 16-24cm)") +
+  geom_signif(comparisons = list(c("Antigua", "Barbuda"),c("Antigua","Bonaire"),c("Barbuda","Bonaire")), map_signif_level=TRUE) +
+  facet_grid(.~species) + 
+  stat_n_text() +
+  scale_fill_brewer(palette="Blues") +
+  theme_bw() +
+  theme(strip.text = element_text(face = "italic")) +
+  theme(plot.title = element_text(hjust = 0.5))
+
+# Length-weight relationships between islands
+ggplot(filter(sum_id, species_code != "rbp" & island != "Barbuda"), 
+       aes(x = length_cm, y = fr, colour = island)) +
+  facet_grid(.~species) +
+  geom_point(shape = 18) +
+  scale_shape(solid = FALSE) +
+  geom_smooth(method = lm, fullrange = FALSE) +
+  labs(y = "Feeding Rate (bites/hr)", x = "Length (cm)", title = "Hourly feeding rate") + 
+  theme_bw() + 
+  theme(plot.title = element_text(hjust = 0.5)) + 
+  scale_color_manual(values = c("navy", "darkseagreen", "slategray2"))
+
+# Calculating length-weight coefficients
+
+fr_anu_vir <- lm(fr~length_cm, data=filter(sum_id, island == "Antigua" & phase == "i" & species_code == "stop"))
+summary(fr_anu_vir) # fr(Antigua, S. viride, initial) = 445-8*L
+
+fr_anu_vet <- lm(fr~length_cm, data=filter(sum_id, island == "Antigua" & phase == "i" & species_code == "qup"))
+summary(fr_anu_vet) # fr(Antigua, S. vetula, initial) = 1030-12*L
+
+fr_bon_vir <- lm(fr~length_cm, data=filter(sum_id, island == "Bonaire" & phase == "i" & species_code == "stop"))
+summary(fr_bon_vir) # fr(Bonaire, S. viride, initial) = 1057-17*L
+
+fr_bon_vet <- lm(fr~length_cm, data=filter(sum_id, island == "Bonaire" & phase == "i" & species_code == "qup"))
+summary(fr_bon_vet) # fr(Bonaire, S. vetula, initial) = 1347+10*L
+```
+
+fr(Antigua, *S. viride*, initial) = 445-8*L
+fr(Antigua, *S. vetula*, initial) = 1030-12*L
+fr(Bonaire, *S. viride*, initial) = 1057-17*L
+fr(Bonaire, *S. vetula*, initial) = 1347+10*L
+
+## Comparison to published equations
+
+## Predicted feeding rates and bite sizes from standard equations
+Bite sizes as predicted by Bruggemann/Bozec:
+bs(*S. vetula*) = 4.013*10^-4*L^2
+bs(*S. viride*) = 5.257*10^-4*L^2
+
+Feeding rates as published by Bruggemann/Bozec
+fr(Published, *S. viride*, initial) = 1089-17*L-56 
+  *(extra term because phase modifier is not used here (i=1))*
+fr(Published, *S. vetula*, initial) = 3329-33*L
+
+## adding predicted feeding rates from observed relationships
+
+```{r}
+fish_fr <- fish %>% 
+  filter(species_code %in% c("stop", "qup") & island %in% c("Antigua","Bonaire")) %>%
+  mutate(
+    fr_pers = ifelse(island == "Antigua" & species_code == "stop", 445-8*length,
+                     ifelse(island == "Antigua" & species_code == "qup", 1030-12*length,
+                            ifelse(island == "Bonaire" & species_code == "stop", 
+                                   1057-17*length, 1347+10*length))),
+    fr_pub = ifelse(species_code == "stop", 1089-17*length-56, 3329-33*length),
+    bs = ifelse(species_code == "stop", 5.257*10^-4*length^2, 4.013*10^-4*length^2),
+    ag_pers = fr_pers*bs,
+    ag_pub = fr_pub*bs
+)
+
+ag_site <- fish_fr %>% group_by(site,island) %>%
+  summarize(personal=sum(ag_pers)/1000,published=sum(ag_pub)/1000)
+ag_melt <- ag_site %>% as.data.frame() %>% 
+  gather(source, area_grazed, -c(site,island))
+
+ggplot(filter(ag_melt, island == "Antigua"),
+      aes(x=reorder(site, -area_grazed), y=area_grazed)) +
+        geom_bar(aes(fill = source), position = "dodge", stat = "identity") +
+        scale_fill_manual(values = c("black", "slategray2")) +
+        labs(title=expression(Predicted~grazing~impact~on~Antiguan~reefs~(m^2/hr))) +
+        theme_bw() +
+        theme(axis.text.x = element_text(angle = 90, hjust = 1))
+```

### 2018-11-15 r/Data_prodding.Rmd
11/14

@@
   stat_summary(fun.y = mean, geom="point", shape=5, size=3) + 
-  labs(y="Feeding Rate (bites/hr)",x="",title="Hourly feeding rates") +
+  labs(y = "Feeding Rate (bites/hr)", x = "", title = "Hourly feeding rates (i. phase, 16-24cm)") +
   theme_bw() + 
@@
   geom_boxplot(width=0.1) + 
-  labs(y="Feeding Rate (bites/hr)",x="",title="Hourly feeding rates (init. 16-24cm") +
+  labs(y = "Feeding Rate (bites/hr)", x = "", title = "Hourly feeding rates (i. phase, 16-24cm)") +
   geom_signif(comparisons = list(c("Antigua", "Barbuda"),c("Antigua","Bonaire"),c("Barbuda","Bonaire")), map_signif_level=TRUE) +
@@
 
+```{r}
+vet <- sum_id %>% filter(phase == "i" & length_cm >= 16 & length_cm <= 24 & species_code=="qup")
+aov_fr_vet <- aov(fr ~ island, data = vet)
+summary.aov(aov_fr_vet) # significant differences
+TukeyHSD(aov_fr_vet)
+plot(aov_fr_vet, 2) # normally distributed
+leveneTest(fr ~ island, data = vet) # homogeneity of variance
+```
+```{r}
+vir <- sum_id %>% filter(phase == "i" & length_cm >= 16 & length_cm <= 24 & species_code=="stop")
+aov_fr_vir <- aov(fr ~ island, data = vir, white.adjust=T) 
+summary.aov(aov_fr_vir) # significant differences
+TukeyHSD(aov_fr_vir)
+plot(aov_fr_vir, 2) # normally distributed
+leveneTest(fr ~ island, data = vir) # heteroskedastic - add white.adjust=TRUE
+```
+
+```{r}
+ggplot(filter(sum_id, 
+              phase=="i" & length_cm>=16 & length_cm<=24 & species_code!="rbp"), 
+       aes(factor(island), g_frac)) + 
+  geom_violin(aes(fill=island)) +
+  geom_boxplot(width=0.1) + 
+  labs(y = "Fraction of time", x = "",title="Grazing time budgets (i. phase, 16-24cm)") +
+  geom_signif(comparisons = list(c("Antigua", "Barbuda"),c("Antigua","Bonaire"),c("Barbuda","Bonaire")), map_signif_level=TRUE) +
+  facet_grid(.~species) + 
+  stat_n_text() +
+  scale_fill_brewer(palette="Blues") +
+  theme_bw() +
+  theme(strip.text = element_text(face = "italic")) +
+  theme(plot.title = element_text(hjust = 0.5)) 
+```
+
+```{r}
+vet <- sum_id %>% filter(phase == "i" & length_cm >= 16 & length_cm <= 24 & species_code=="qup")
+aov_g_vet <- aov(g_frac ~ island, data = vet, white.adjust = T)
+summary.aov(aov_g_vet) # significant differences
+TukeyHSD(aov_g_vet)
+plot(aov_g_vet, 2) # normally distributed
+leveneTest(g_frac ~ island, data = vet) # heteroskedastic - add white.adjust = 2
+```
+```{r}
+vir <- sum_id %>% filter(phase == "i" & length_cm >= 16 & length_cm <= 24 & species_code=="stop")
+aov_g_vir <- aov(g_frac ~ island, data = vir, white.adjust=T) 
+summary.aov(aov_g_vir) # significant differences
+TukeyHSD(aov_g_vir)
+plot(aov_g_vir, 2) # normally distributed
+leveneTest(g_frac ~ island, data = vir) # homogeneity of variance
+```
+
+```{r}
+ggplot(filter(sum_id, 
+              phase=="i" & length_cm>=16 & length_cm<=24 & species_code!="rbp"), 
+       aes(factor(island), br)) + 
+  geom_violin(aes(fill=island)) +
+  geom_boxplot(width=0.1) + 
+  labs(y = "Bite Rate (bites/sec)", x = "", title = "Bite rate while grazing (i. phase, 16-24cm)") +
+  geom_signif(comparisons = list(c("Antigua", "Barbuda"),c("Antigua","Bonaire"),c("Barbuda","Bonaire")), map_signif_level=TRUE) +
+  facet_grid(.~species) + 
+  stat_n_text() +
+  scale_fill_brewer(palette="Blues") +
+  theme_bw() +
+  theme(strip.text = element_text(face = "italic")) +
+  theme(plot.title = element_text(hjust = 0.5)) 
+```
+
+```{r}
+vet <- sum_id %>% filter(phase == "i" & length_cm >= 16 & length_cm <= 24 & species_code=="qup")
+aov_br_vet <- aov(br ~ island, data = vet)
+summary.aov(aov_br_vet) # significant differences
+TukeyHSD(aov_br_vet)
+plot(aov_br_vet, 2) # normally distributed
+leveneTest(br ~ island, data = vet) # homogeneity of variance
+```
+```{r}
+vir <- sum_id %>% filter(phase == "i" & length_cm >= 16 & length_cm <= 24 & species_code=="stop")
+aov_br_vir <- aov(br ~ island, data = vir, white.adjust=T) 
+summary.aov(aov_br_vir) # significant differences
+TukeyHSD(aov_br_vir)
+plot(aov_br_vir, 2) # normally distributed
+leveneTest(br ~ island, data = vir) # heteroskedastic - add white.adjust=TRUE
+```
+
+```{r}
+ggplot(filter(sum_id, 
+              phase=="i" & length_cm>=16 & length_cm<=24 & species_code!="rbp"), 
+       aes(factor(island), for_bites)) + 
+  geom_violin(aes(fill=island)) +
+  geom_boxplot(width=0.1) + 
+  labs(y = "Bite Rate (bites/sec)", x = "", title = "Bite rate while grazing (i. phase, 16-24cm)") +
+  geom_signif(comparisons = list(c("Antigua", "Barbuda"),c("Antigua","Bonaire"),c("Barbuda","Bonaire")), map_signif_level=TRUE) +
+  facet_grid(.~species) + 
+  stat_n_text() +
+  scale_fill_brewer(palette="Blues") +
+  theme_bw() +
+  theme(strip.text = element_text(face = "italic")) +
+  theme(plot.title = element_text(hjust = 0.5))
+
+vet <- sum_id %>% filter(phase == "i" & length_cm >= 16 & length_cm <= 24 & species_code=="qup")
+aov_for_bites_vet <- aov(for_bites ~ island, data = vet)
+summary.aov(aov_for_bites_vet) # significant differences
+TukeyHSD(aov_for_bites_vet)
+plot(aov_for_bites_vet, 2) # normally distributed
+leveneTest(for_bites ~ island, data = vet) # homogeneity of variance
+
+vir <- sum_id %>% filter(phase == "i" & length_cm >= 16 & length_cm <= 24 & species_code=="stop")
+aov_for_bites_vir <- aov(for_bites ~ island, data = vir, white.adjust=T) 
+summary.aov(aov_for_bites_vir) # significant differences
+TukeyHSD(aov_for_bites_vir)
+plot(aov_for_bites_vir, 2) # normally distributed
+leveneTest(for_bites ~ island, data = vir) # heteroskedastic - add white.adjust=TRUE
+```
+
 # Exploratory bivariate plots
@@
 ```{r}
-ggplot(filter(sum_id, species_code != "rbp" & island != "Barbuda"), aes(x = length_cm, y = fr, colour = island)) +
+ggplot(filter(sum_id, species_code != "rbp" & island != "Barbuda"), 
+       aes(x = length_cm, y = fr, colour = island)) +
   facet_grid(.~species) +
@@
 
--site-level predictors: scar_bm,scar_den,carn_bm,benthic (pc1,pc2)  
--fish-level predictors: species, phase, length  
---eventually run separately for different species  
---species*scar_bm interaction?  
--random effects: island  
--response variables: g_frac, br (?), and fr (run separately)
-
-## Basic lm 
-```{r}
-# Testing VIF with individual variables of interest
-lm_fr <- lm(fr~species+phase+length_cm+scar_bm+consp_den+ma_canopy+ta_cover+ta_canopy, data=filter(sum_id_pca1, species_code!="rbp"))
-summary(lm_fr) 
-vif(lm_fr)
-```
-
-Only VIF above 5 is rugosity
-```{r}
-# Using PC values as predictors
-lm_fr <- lm(fr~species+phase+length_cm+pc1+pc2, data=filter(sum_id_pca1, species_code!="rbp"))
-summary(lm_fr) 
-vif(lm_fr)
-```
+- site-level predictors: scar_bm,scar_den,carn_bm,benthic (pc1,pc2)  
+- fish-level predictors: species, phase, length  
+  + eventually run separately for different species  
+  + species*scar_bm interaction?  
+- random effects: island  
+- response variables: g_frac, br (?), and fr (run separately)
 
@@
 summary(lme_fr)
+vif(lme_fr)
 ```
 
-## GAMs
+## GAM + GAMM
 ```{r, echo=FALSE}
@@
 ```
-
-## GAMMs
 
@@
 
-```
 ------------------------
 To Do as of Nov. 7  
-* ~~incorporate pcs~~
 * boosted regression trees ecosphere 2017 adrians paper  
-* ~~simpler graphs with just herb density x axis - grazing behaviors y, facet by species~~
-* ~~species*density as fixed effect~~
 * species as random effect
-* ~~break out species and run separately~~

### 2018-11-15 r/Sites.Rmd
11/14

@@
   bind_rows(ab_fish) %>%
-  mutate(biomass = as.numeric(biomass)) %>%
+  mutate(
+    biomass = as.numeric(biomass),
+    species_code = tolower(species_code)) %>%
   replace(. == "", NA)
@@
 write.csv(sp_site,file=here("data/sp_site.csv"))
+write.csv(fish,file=here("data/fish.csv")) # for predicting grazing impact of observed fish communities at each site (see Analysis.rmd)
 ```

### 2018-11-16 r/Analysis.Rmd
11/15

@@
 Notes: 
-- re: lms/gamms: don't focus on these for now, the data has so many colinearities... if anything, use full PCA in model analysis
+- re: lms/gamms: have been stressed about collinearities in the data, but some aren't that bad...
+return to models once I have a better handle on what I want to use (and make sure I'm looking at mixed effects models only?)
 - Exclude Barbuda from this analysis...?  
+- From Nash et al. 2016:  
+  + used GAMMs with site and reef zone as random effects
+  + aggregated data to the site level
+  + used mean or median values based on skew of different predictor/response variables (within each site)
 

### 2018-11-16 r/Data_prodding.Rmd
11/15

@@
   select(-X) %>%
-  mutate(island = factor(island, levels = c("Bonaire","Antigua","Barbuda"))) %>%
-  arrange(island) %>%
   filter(length_cm>=15)
+sum_site_sp_ph <- read.csv(here("data","sum_site_sp_ph.csv"), stringsAsFactors=F) %>%
+  select(-X) 
 ```
@@
 
+PCA to visualize variable relationships:
+
 ```{r}
@@
               column_to_rownames(var="site") %>%
-              select(ma_ind,ta_cover,ta_canopy,rugosity,scar_den,scar_bm)), 
+              select(ma_ind,ta_cover,ta_canopy,lc_cover,rugosity,scar_den,scar_bm,)), 
   center = T,scale = T)
 
-summary(pca_pred)
-plot(pca_pred,type="l")
+autoplot(pca_pred, data = sum_site, colour = 'island',
+         # frame = T, frame.type = "norm",
+         loadings = T, loadings.colour = 'gray',
+         loadings.label = T, loadings.label.size = 3, loadings.label.colour = 'black',
+            loadings.label.repel = T) +
+    scale_color_manual(name="island", values=c("lightblue", "darkseagreen", "lightsalmon")) + 
+    theme(legend.direction = 'horizontal', legend.position = 'top') +
+    theme_minimal()
+```
 
-autoplot(pca_pred, data = sum_site, colour = 'island',
+PCA for correlated (benthic only?) variables:
+
+```{r}
+pca_bent <- prcomp(na.omit(sum_site %>% 
+              remove_rownames %>% 
+              column_to_rownames(var="site") %>%
+              select(ma_ind,ta_cover,ta_canopy,lc_cover,rugosity,)), 
+  center = T,scale = T)
+
+summary(pca_bent)
+plot(pca_bent,type="l")
+
+autoplot(pca_bent, data = sum_site, colour = 'island',
          # frame = T, frame.type = "norm",
@@
 sum_site_pca1 <- sum_site %>% 
-  bind_cols(as.data.frame(pca_pred$x) %>% 
+  bind_cols(as.data.frame(pca_bent$x) %>% 
            select(PC1,PC2)) %>%
@@
 sum_id_pca1 <- sum_id %>% left_join(select(sum_site_pca1,site,pc1,pc2), by = "site")
+sum_ssp_pca1 <- sum_site_sp_ph %>% left_join(select(sum_site_pca1,site,pc1,pc2), by = "site")
 write.csv(sum_id_pca1,file=here("data/sum_id_pca1.csv"))
+write.csv(sum_ssp_pca1,file=here("data/sum_ssp_pca1.csv"))
 ```
@@
   theme(plot.title = element_text(hjust = 0.5))
+
+ggplot(filter(sum_id, 
+              phase=="i" & length_cm>=16 & length_cm<=24 & species_code!="rbp"), 
+       aes(factor(site), fr)) + 
+  geom_violin(aes(fill=island)) +
+  geom_boxplot(width=0.1) + 
+  labs(y = "Feeding Rate (bites/hr)", x = "", title = "Hourly feeding rates (i. phase, 16-24cm)") +
+  geom_signif(comparisons = list(c("Antigua", "Barbuda"),c("Antigua","Bonaire"),c("Barbuda","Bonaire")), map_signif_level=TRUE) +
+  facet_grid(.~species) + 
+  stat_n_text() +
+  scale_fill_brewer(palette="Blues") +
+  theme_bw() +
+  theme(strip.text = element_text(face = "italic")) +
+  theme(plot.title = element_text(hjust = 0.5), axis.text.x = element_text(angle = 90, hjust = 1)) 
 ```
@@
 ```{r}
-ggplot(filter(sum_id, species_code != "rbp" & island != "Barbuda" & length_cm>=16 & length_cm<=24), aes(x = ma_ind,y=fr)) +
+ggplot(filter(sum_id_pca1, species_code != "rbp" & island != "Barbuda" & length_cm>=16 & length_cm<=24), aes(x = ma_ind,y=fr)) +
   facet_grid(.~species) +
@@
 ## Mixed effects models
-- random effect: island
+- random effect: island or site?
 
 ```{r}
-lme_fr <- lme(fr~phase+length_cm+species+pc1+pc2,data=filter(sum_id_pca1, species_code!="rbp"),random=~1|island, na.action = na.omit)
+lme_fr <- lme(fr~phase+length_cm+species+scar_bm+pc1+pc2,
+              data=filter(sum_id_pca1, species_code!="rbp"),
+              random=~1|site, 
+              na.action = na.omit
+              )
 summary(lme_fr)
-vif(lme_fr)
+vif(lme_fr) # test VIF here, because I'm not sure how to do it in GAMM yet...? All low VIF (<3) when random = island, but when random = site, VIFs jump up (for scarid_den & pc2)
 ```
 
-## GAM + GAMM
-```{r, echo=FALSE}
-gam<-gam(fr~species+phase+s(length_cm)+s(scar_bm)+s(pc1)+s(pc2), family=gaussian(link=identity),data=sum_id_pca1, method="REML")
-summary(gam)
-plot(gam,pages=1)
-```
+## GAMM
+- for GAMMs - seems model results are the same whether random = site or island?
 
 ```{r,eval=F}
-gamm<-gamm(fr~phase+s(length_cm)+s(pc1)+s(pc2), family=gaussian(link=identity),data=filter(sum_id_pca1, species_code=="qup"), random=list(island=~1),method="REML")
+gamm<-gamm(fr~phase+s(length_cm)+s(pc1)+s(pc2)+s(scar_bm)+s(scar_den), 
+           family=gaussian(link=identity),
+           data=filter(sum_id_pca1, species_code=="stop"),
+           random=list(site=~1),
+           method="REML"
+           )
 summary(gamm$gam)
@@
 ```{r,eval=F}
-gamm<-gamm(fr~phase+s(length_cm)+s(pc1)+s(pc2), family=gaussian(link=identity),data=filter(sum_id_pca1, species_code=="stop"), random=list(island=~1),method="REML")
+gamm<-gamm(fr_mean~phase+species+s(pc1)+s(pc2)+s(scar_bm)+s(scar_den), 
+           family=gaussian(link=identity),
+           data=filter(sum_ssp_pca1, species_code!="rbp" & island!="Barbuda"),
+           random=list(site=~1),
+           method="REML"
+           )
 summary(gamm$gam)

### 2018-11-16 r/Follows.Rmd
11/13
11/15

@@
 
-Notes 11/11:
-* figure out dcast issues
-* replacing NA/0s within pipe
+Wrangling of fish follow data from Antigua, Barbuda, and Bonaire, as well as integration of site-level data on fish, benthic, and rugosity (from Sites.rmd).
 
@@
 sum_site_sp <- sum_id %>% 
-  group_by(site, island, species_code) %>% 
+  group_by(site, island, species, species_code) %>% 
   summarize(
@@
 sum_site_sp_ph <- sum_id %>% 
-  group_by(site, island, species_code, phase) %>% 
+  group_by(site, island, species, species_code, phase) %>% 
   summarize(
@@
   left_join(sp_site, by=c("species","site")) # add conspecific scarid info (based on site/species)
+sum_site_sp_ph <- sum_site_sp_ph %>% 
+  left_join(site, by="site") %>%
+  left_join(sp_site, by=c("species","site")) # add conspecific scarid info (based on site/species)
 
@@
 write.csv(sum_id,file=here("data/sum_id.csv"))
+write.csv(sum_site_sp_ph,file=here("data/sum_site_sp_ph.csv"))
 ```

### 2018-11-16 r/SEM.Rmd
11/15

@@
   # latent variable indicators
-      risk =~ scar_bm      
       nutrition =~ ma_cover + ta_cover
   # regressions
-      fr ~ length_cm + nutrition + risk
+      fr ~ length_cm + nutrition
       lc_cover ~ ma_cover + ta_cover

### 2018-11-24 r/Data_prodding.Rmd
11/24

@@
 library(gamm4)
+library(xtable)
 ```
@@
 sum_site_sp_ph <- read.csv(here("data","sum_site_sp_ph.csv"), stringsAsFactors=F) %>%
+  select(-X) 
+fish <- read.csv(here("data","fish.csv"), stringsAsFactors=F) %>%
   select(-X) 
@@
 pred <- sum_site %>% 
-  select(tot_bm,tot_den,scar_bm,scar_den,carn_bm,carn_den,
-         cca_cover,lc_cover,ma_cover,ma_canopy,ta_cover,ta_canopy,rugosity) %>% 
+  select(tot_bm,tot_den,scar_bm,scar_den,carn30_bm,carn30_den,
+         cca_cover,lc_cover,ma_cover,ma_canopy,ta_cover,ta_canopy,rugosity,spear_daily) %>% 
   gather() %>% 
@@
 pred <- sum_id %>% select(tot_bm,tot_den,scar_bm,scar_den,consp_bm,consp_den,
-                           carn_bm,carn_den,cca_cover,lc_cover,
+                           carn30_bm,carn30_den,cca_cover,lc_cover,
                            ma_cover, ma_canopy,ta_cover,ta_canopy,rugosity)
@@
               column_to_rownames(var="site") %>%
-              select(ma_ind,ta_cover,ta_canopy,lc_cover,rugosity,scar_den,scar_bm,)), 
+              select(ma_ind,ta_cover,ta_canopy,lc_cover,rugosity,scar_den,scar_bm,spear_daily)),
   center = T,scale = T)
@@
 ```{r}
-pca_bent <- prcomp(na.omit(sum_site %>% 
+pca_all <- prcomp(na.omit(sum_site %>% 
               remove_rownames %>% 
               column_to_rownames(var="site") %>%
-              select(ma_ind,ta_cover,ta_canopy,lc_cover,rugosity,)), 
+              # filter(island!="Barbuda") %>%
+              select(ma_cover, ma_canopy, ta_cover, ta_canopy, rugosity, scar_bm, scar_den)),
   center = T,scale = T)
 
-summary(pca_bent)
-plot(pca_bent,type="l")
-
-autoplot(pca_bent, data = sum_site, colour = 'island',
+autoplot(pca_all, data = sum_site, colour = 'island',
          # frame = T, frame.type = "norm",
@@
 
+pca_bent <- prcomp(na.omit(sum_site %>% 
+              remove_rownames %>% 
+              column_to_rownames(var="site") %>%
+              select(ma_ind,ta_cover,ta_canopy,lc_cover,rugosity)), 
+  center = T,scale = T)
+summary(pca_bent)
+plot(pca_bent,type="l")
+
 # add PC1 and PC2 as variables in dataset
-sum_site_pca1 <- sum_site %>% 
+sum_site <- sum_site %>% 
+  # filter(island!="Barbuda") %>%
   bind_cols(as.data.frame(pca_bent$x) %>% 
            select(PC1,PC2)) %>%
-  janitor::clean_names(case="old_janitor")
-sum_id_pca1 <- sum_id %>% left_join(select(sum_site_pca1,site,pc1,pc2), by = "site")
-sum_ssp_pca1 <- sum_site_sp_ph %>% left_join(select(sum_site_pca1,site,pc1,pc2), by = "site")
-write.csv(sum_id_pca1,file=here("data/sum_id_pca1.csv"))
-write.csv(sum_ssp_pca1,file=here("data/sum_ssp_pca1.csv"))
+  rename(pc1_bent=PC1,
+         pc2_bent=PC2) %>% 
+  bind_cols(as.data.frame(pca_all$x) %>% 
+           select(PC1,PC2)) %>%
+  rename(pc1_all=PC1,
+         pc2_all=PC2)
+
+sum_id_pca <- sum_id %>% 
+  left_join(select(sum_site,site,pc1_bent,pc2_bent,pc1_all,pc2_all), by = "site")
+
 ```
@@
 ```{r}
-ggplot(filter(sum_id_pca1, species_code != "rbp" & island != "Barbuda" & length_cm>=16 & length_cm<=24), aes(x = ma_ind,y=fr)) +
+ggplot(filter(sum_id_pca, species_code != "rbp" & island != "Barbuda" & length_cm>=16 & length_cm<=24), aes(x = ma_ind,y=fr)) +
   facet_grid(.~species) +
@@
 ```{r}
-lme_fr <- lme(fr~phase+length_cm+species+scar_bm+pc1+pc2,
-              data=filter(sum_id_pca1, species_code!="rbp"),
-              random=~1|site, 
+lme_fr <- lme(fr~phase+length_cm+species+scar_bm+pc1_bent+pc2_bent,
+              data=filter(sum_id_pca, species_code!="rbp"),
+              random=~1|island, # do I need random slope too?
               na.action = na.omit
@@
 ```{r,eval=F}
-gamm<-gamm(fr~phase+s(length_cm)+s(pc1)+s(pc2)+s(scar_bm)+s(scar_den), 
+gamm<-gamm(fr~phase+s(length_cm)+s(pc1_bent)+s(pc2_bent)+s(scar_bm)+s(scar_den), 
            family=gaussian(link=identity),
-           data=filter(sum_id_pca1, species_code=="stop"),
+           data=filter(sum_id_pca, species_code=="qup"),
            random=list(site=~1),
@@
 ```{r,eval=F}
-gamm<-gamm(fr_mean~phase+species+s(pc1)+s(pc2)+s(scar_bm)+s(scar_den), 
+gamm1<-gamm(fr_mean~phase+species+s(pc1_bent)+s(pc2_bent)+s(scar_bm)+s(scar_den), 
            family=gaussian(link=identity),
-           data=filter(sum_ssp_pca1, species_code!="rbp" & island!="Barbuda"),
+           data=filter(sum_ssp_pca, species_code!="rbp" & island!="Barbuda"),
            random=list(site=~1),
@@
            )
-summary(gamm$gam)
-plot(gamm$gam,pages=1)
+summary(gamm1$gam)
+AIC(gamm1$lme)
+plot(gamm1$gam,pages=1)
+
 ```

### 2018-11-24 r/Follows.Rmd
11/24

@@
   ) %>%
-  replace(. =="#N/A", NA)
+  replace(. =="#N/A", NA) %>%
+  filter(length_cm >= 15)
 
@@
 sum_site <- sum_site %>% 
-  left_join(site, by="site")
+  left_join(site, by=c("island","site"))
 sum_id <- sum_id %>% 
-  left_join(site, by="site") %>%
+  left_join(site, by=c("island","site")) %>%
   left_join(sp_site, by=c("species","site")) # add conspecific scarid info (based on site/species)
 sum_site_sp_ph <- sum_site_sp_ph %>% 
-  left_join(site, by="site") %>%
+  left_join(site, by=c("island","site")) %>%
   left_join(sp_site, by=c("species","site")) # add conspecific scarid info (based on site/species)

### 2018-11-24 r/Models.Rmd
11/24

---
title: "Scarid behavior GAM/GAMMS"
author: "Molly Wilson"
date: "Dec. 2018"
output:
  html_document:
    toc: true
    toc_float: true
    theme: united
    number_sections: true
---

```{r}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE)

library(startR)
library(here)
library(tidyverse)
library(mgcv)
library(car)

sum_id <- read.csv(here("data","sum_id.csv"), stringsAsFactors=F) %>%
  select(-X)
sum_site <- read.csv(here("data","sum_site.csv"), stringsAsFactors=F) %>%
  select(-X) %>%
  mutate(island = factor(island, levels = c("Bonaire","Antigua","Barbuda"))) %>%
  arrange(island)
```

# Examining predictor variables
Notes: in models using benthic-only PCA, scarid biomass and scarid density are highly collinear with benthic PCA variables (understandably). I don't think I can justify separating them, even though they are my ultimate variables of interest. I can use PCA loadings to interpret model outcomes. Carnivore biomass and spearfishing presence is also highly correlated with scarid biomass and thus benthic PCA variables - I am excluding it for now.

```{r}
pca_all <- prcomp(na.omit(sum_site %>% 
              remove_rownames %>% 
              column_to_rownames(var="site") %>%
              # filter(island!="Barbuda") %>%
              select(ma_cover, ma_canopy, ta_cover, ta_canopy, rugosity, scar_bm, scar_den)),
  center = T,scale = T)
autoplot(pca_all, data = sum_site, colour = 'island',
         # frame = T, frame.type = "norm",
         loadings = T, loadings.colour = 'gray',
         loadings.label = T, loadings.label.size = 3, loadings.label.colour = 'black',
            loadings.label.repel = T) +
    scale_color_manual(name="island", values=c("lightblue", "darkseagreen", "salmon")) + 
    theme(legend.direction = 'horizontal', legend.position = 'top') +
    theme_minimal()

pca_bent <- prcomp(na.omit(sum_site %>% 
              remove_rownames %>% 
              column_to_rownames(var="site") %>%
              # filter(island!="Barbuda") %>%
              select(ma_cover, ma_canopy, ta_cover, ta_canopy, rugosity)),
  center = T,scale = T)

sum_site <- sum_site %>% 
  # filter(island!="Barbuda") %>%
  bind_cols(as.data.frame(pca_bent$x) %>% 
           select(PC1,PC2)) %>%
  rename(pc1_bent=PC1,
         pc2_bent=PC2) %>% 
  bind_cols(as.data.frame(pca_all$x) %>% 
           select(PC1,PC2)) %>%
  rename(pc1_all=PC1,
         pc2_all=PC2)

sum_id_pca <- sum_id %>% 
  left_join(select(sum_site,site,pc1_bent,pc2_bent,pc1_all,pc2_all), by = "site")

```

```{r,echo=FALSE}
panel.cor <- function(x, y, digits=2, prefix="", cex.cor, ...)
{
    usr <- par("usr"); on.exit(par(usr))
    par(usr = c(0, 1, 0, 1))
    r <- abs(cor(x, y))
    txt <- format(c(r, 0.123456789), digits=digits)[1]
    txt <- paste(prefix, txt, sep="")
    if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
    text(0.5, 0.5, txt, cex = cex.cor * r)
}
pairs(~scar_bm + pc1_bent + pc2_bent, data=sum_site,
      lower.panel=panel.smooth, upper.panel=panel.cor, 
      pch=20, main="Selected Fish & Benthic Community Variables", na.action = na.omit)
```

```{r}
m_data <- sum_id_pca %>% filter(species_code!="rbp")
vif(lm(fr ~ species + phase + length_cm + scar_bm + pc1_bent + pc2_bent, data=m_data))
vif(lme(fr ~ species + phase + length_cm + scar_bm + pc1_bent + pc2_bent, data=m_data,
        random=~1|island, 
        na.action = na.omit
        ))
```

Confirming I can't use scarid biomass and/or density as separate from benthic PCA variables. VIF of scar_bm is just slightly >3. Removing rugosity from benthic PCA reduces scar_bm ~ pc1 correlation slightly, but still >0.7.

BUT: VIF done on LME where random=site or random=island shows values <3 when using benthic PCA values + scarid biomass... *stay tuned...*

# Model as LME first
```{r}
lme1 <- lme(fr ~ species + phase + length_cm + scar_bm + pc1_bent + pc2_bent, data=m_data,
    random = ~1 | island, 
    na.action = na.omit
    )
summary(lme1)
vif(lme1)
```

# Switching to gamm, checking model validity

Notes:  
- GLS: option for model when homogeneity of variances is not satisfied, can add weights for residuals. Keep reading Zuur et al. and decide if I need to go beyond this (e.g. GAMM)

```{r}
m1 <- gam(fr ~ factor(species) + s(length_cm, bs = "cs") + s(pc1_all, bs = "cs") + s(pc2_all, bs = "cs"), data=m_data)
summary(m1)
plot(m1)
gam.check(m1)

gls1 <- gam(fr ~ species + length_cm + pc1_all + pc2_all, data=m_data)
summary(gls1)

e <- resid(m1)
coplot(e ~ length_cm | species, data = m_data) # differences in res. dist. between species
coplot(e ~ pc1_all | species, data = m_data) # differences in res. dist. between species
coplot(e ~ pc2_all | species, data = m_data) # inconsistent res.

m2 <- gam(fr ~ factor(species_code) + s(length_cm) + s(pc1_all) +s (pc2_all) +
          s(length_cm, by = as.numeric(species_code=="qup")), 
          data = m_data)
summary(m2)
gam.check(m2)

m3 <- gam(fr ~ factor(species_code) + s(length_cm) + s(pc1_all) + s(pc2_all) +
          s(pc1_all, by = as.numeric(species_code == "qup")),
          data = m_data)
summary(m3)
gam.check(m3)

m4 <- gam(fr ~ factor(species_code) + s(length_cm) + s(pc1_all) + s(pc2_all) +
          s(pc2_all, by = as.numeric(species_code == "qup")),
          data = m_data)
summary(m4)
gam.check(m4)

m5 <- gam(fr ~ factor(species_code) + s(length_cm) + s(pc1_all) + s(pc2_all) +
          s(pc1_all, by = as.numeric(species_code == "qup"))+
          s(pc2_all, by = as.numeric(species_code == "qup")),
          data = m_data)
summary(m5)
plot(m5)
gam.check(m5)

```

Notes:  
- Violates assumption of homogeneity of variance (increasing variance as predictor values increase)
- Satisfies assumption of normality

```{r}
gamm<-gamm(fr~phase+s(length_cm)+s(pc1)+s(pc2)+s(scar_bm)+s(scar_den), 
           family=gaussian(link=identity),
           data=filter(sum_id_pca1, species_code=="qup"),
           random=list(site=~1),
           method="REML"
           )
summary(gamm$gam)
plot(gamm$gam,pages=1)
```
```{r}
gamm1<-gamm(fr_mean~phase+species+s(pc1)+s(pc2)+s(scar_bm)+s(scar_den), 
           family=gaussian(link=identity),
           data=filter(sum_ssp_pca1, species_code!="rbp" & island!="Barbuda"),
           random=list(site=~1),
           method="REML"
           )
summary(gamm1$gam)
gamtabs(gamm1)
AIC(gamm1$lme)
plot(gamm1$gam,pages=1)

```

### 2018-11-24 r/SEM.Rmd
11/24

@@
 # set up
-library(tidyr)
+library(here)
+library(tidyverse)
 library(lavaan)
@@
 # define model pathways
-qup_df <- sum_id %>% 
-  select(scar_bm, ma_cover, ta_cover, fr, g_frac, rugosity, int_scar, length_cm, lc_cover) %>% 
-  na.omit()
-
-sem <- '
+sem_full <- '
   # latent variable indicators
-      risk =~ scar_bm + rugosity
+      risk =~ scar_bm_sc + rugosity_sc
       nutrition =~ ma_cover + ma_canopy + ta_cover + ta_canopy
   # regressions
-      fr ~ length_cm + risk + nutrition
-      rugosity ~ lc_cover
-      lc_cover ~ ma_cover + ta_cover
-      ta_cover ~ scar_bm
-      ma_cover ~ scar_bm
+      fr_sc ~ length_cm + risk + nutrition
+      rugosity_sc ~ lc_cover
+      lc_cover ~ ma_cover + ma_canopy + ta_cover + ta_canopy
+      ta_cover ~ scar_bm_sc
+      ma_cover ~ scar_bm_sc    
+      ta_canopy ~ scar_bm_sc
+      ma_canopy ~ scar_bm_sc 
 '
+```
+
+```{r}
+# testing partial SEMs
+
+sem_df <- sum_id %>% 
+  filter(island!="Barbuda") %>%
+  select(fr, g_frac, species_code, phase, scar_bm, scar_den, carn30_bm, tot_bm, spear_daily, ma_cover, ma_canopy, ta_cover, ta_canopy, rugosity, int_scar, length_cm, lc_cover) %>% 
+  mutate(phase_dum=ifelse(phase=="i",1,2),
+         fr_sc = fr/100,
+         scar_bm_sc = scar_bm/100,
+         rugosity_sc = rugosity*10) %>%
+  na.omit()
+vet_df <- sem_df %>% filter(species_code=="qup")
+vir_df <- sem_df %>% filter(species_code=="stop")
 
@@
   # latent variable indicators
+      risk =~ scar_bm_sc
       nutrition =~ ma_cover + ta_cover
   # regressions
-      fr ~ length_cm + nutrition
+      fr_sc ~ length_cm + phase_dum + nutrition + risk
+'
+
+sem2 <- '
+  # latent variable indicators
+      risk =~ scar_bm_sc + scar_den + rugosity_sc + spear_daily
+      nutrition =~ ma_cover + ta_cover
+  # regressions
+      fr_sc ~ length_cm + phase_dum + nutrition + risk
+      rugosity_sc ~ lc_cover
       lc_cover ~ ma_cover + ta_cover
@@
 
-fit <- sem(sem1, data=qup_df)
+sem3 <- '
+  # latent variable indicators
+      risk =~ scar_bm_sc + scar_den + spear_daily
+      nutrition =~ ma_cover + ta_cover
+  # regressions
+      fr_sc ~ length_cm + phase_dum + nutrition + risk
+      lc_cover ~ ma_cover + ta_cover
+'
+
+fit <- sem(sem3, data=vet_df, missing="ML", std.lv=T, orthogonal=T)
 varTable(fit)
-summary(fit, standardized=T, fit.measures=T,rsq=T)
+summary(fit, standardized=T, fit.measures=T, rsq=T)
 semPaths(fit)

### 2018-11-24 r/Sites.Rmd
11/24

@@
 fish_site <-fish_tran %>% 
-  group_by(site) %>% 
+  group_by(site, island) %>% 
   summarize(
@@
     ) %>% 
-  select(-n) # average across transects within each site
+  select(-n) %>% # average across transects within each site\
+  mutate(island = factor(island, levels = c("Bonaire","Antigua","Barbuda"))) %>%
+  arrange(island)
 
-boxplot(bm_t~site, las=2, data = fish_tran)
+ggplot(fish_site,
+      aes(x=factor(site, levels = fish_site$site), y=tot_bm)) +
+        geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
+        geom_errorbar(aes(ymin=tot_bm-tot_bm_se, ymax=tot_bm+tot_bm_se),
+                  width=.2,
+                  position=position_dodge(.9)) +
+        scale_fill_manual(values = c("gray", "slategray2", "darkseagreen")) +
+        labs(title="Total fish biomass") +
+        theme_bw() +
+        theme(axis.text.x = element_text(angle = 90, hjust = 1))
 ```
@@
 # Because not all species were present at each site/transect, I need to create template to add 0 values for biomass and density in transects where they are absent.
+
 exp_tran <- fish %>% 
@@
 ```{r,message=F,warning=F}
+island_site <- fish_site %>% select(island,site) # island/site relationships to add to plots
+
 # Herbivores
@@
          biomass != 0) %>% 
-  group_by(site,transect) %>% 
+  group_by(site, transect) %>% 
   summarize(
@@
 
-acan_site<-acan_tran %>% group_by(site) %>% 
+acan_site<-acan_tran %>% 
+  group_by(site) %>% 
   summarize(
@@
 carn_tran <- fish %>% 
-  filter(family %in% c("Serranidae","lutjanidae","Carangidae"),
+  filter(family %in% c("Serranidae", "Lutjanidae", "Carangidae", "Sphyraenidae", "Haemulidae"),
          biomass != 0) %>% 
-  group_by(site,transect) %>% 
+  group_by(site, transect) %>% 
   summarize(
@@
   replace(. == "NaN",NA)
+
+# Carnivores over 30cm
+carn30_tran <- fish %>% 
+  filter(family %in% c("Serranidae","Lutjanidae","Carangidae", "Sphyraenidae", "Haemulidae"),
+         biomass != 0, length>=30) %>% 
+  group_by(site, transect) %>% 
+  summarize(
+    bm_t=sum(biomass)/1.2,
+    den_t=n()/1.2,
+    l_t=mean(length)
+    ) %>% 
+  right_join(exp_tran,by=c("site","transect")) %>%
+  mutate_each(funs(replace(.,is.na(.),0)), -l_t)
+
+carn30_site<-carn30_tran %>% 
+  group_by(site) %>% 
+  summarize(
+    n=n(),
+    l_n=sum(!is.na(l_t)),
+    carn30_bm=mean(bm_t),
+    carn30_bm_se=(sd(bm_t))/sqrt(n),
+    carn30_den=mean(den_t),
+    carn30_den_se=(sd(den_t))/sqrt(n),
+    carn30_l=mean(l_t, na.rm = T),
+    carn30_l_se=(sd(l_t, na.rm = T))/sqrt(l_n)
+    ) %>% 
+  select(-c(n,l_n)) %>%
+  replace(. == "NaN",NA)
+
+ggplot(right_join(carn30_site, island_site, by="site"),
+      aes(x=factor(site, levels = fish_site$site), y=carn30_bm)) +
+        geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
+        geom_errorbar(aes(ymin=carn30_bm-carn30_bm_se, ymax=carn30_bm+carn30_bm_se),
+                      width=.2,
+                      position=position_dodge(.9)) +
+        scale_fill_manual(values = c("gray", "slategray2", "darkseagreen")) +
+        labs(title="Biomass of Piscivores over 25cm") +
+        theme_bw() +
+        theme(axis.text.x = element_text(angle = 90, hjust = 1))
 
@@
   left_join(carn_site,by="site") %>%
+  left_join(carn30_site,by="site") %>%
   left_join(bent_site,by="site") %>%

### 2018-11-26 r/Analysis.Rmd
11/26

@@
 
-fr(Antigua, *S. viride*, initial) = 445-8*L
-fr(Antigua, *S. vetula*, initial) = 1030-12*L
-fr(Bonaire, *S. viride*, initial) = 1057-17*L
-fr(Bonaire, *S. vetula*, initial) = 1347+10*L
+$$fr(Antigua, S. viride, i) = 445-8*L$$
+$$fr(Antigua, S. vetula, i) = 1030-12*L$$
+$$fr(Bonaire, S. viride, i) = 1057-17*L$$
+$$fr(Bonaire, S. vetula, i) = 1347+10*L$$
 
@@
 Bite sizes as predicted by Bruggemann/Bozec:
-bs(*S. vetula*) = 4.013*10^-4*L^2
-bs(*S. viride*) = 5.257*10^-4*L^2
+$$bs(S. vetula) = 4.013*10^{-4}*L^2$$
+$$bs(S. viride) = 5.257*10^{-4}*L^2$$
 
 Feeding rates as published by Bruggemann/Bozec
-fr(Published, *S. viride*, initial) = 1089-17*L-56 
+$$fr(Published, S. viride, i) = 1089-17*L-56 $$
   *(extra term because phase modifier is not used here (i=1))*
-fr(Published, *S. vetula*, initial) = 3329-33*L
+$$fr(Published, S. vetula, initial) = 3329-33*L$$
 
@@
         scale_fill_manual(values = c("black", "slategray2")) +
-        labs(title=expression(Predicted~grazing~impact~on~Antiguan~reefs~(m^2/hr))) +
+        labs(x = "Site", 
+             y = expression(Area~grazed~(m^2)~per~Hour), 
+            title = expression(Predicted~cumulative~grazing~impact~on~Antiguan~reefs~(m^2/hr))) +
         theme_bw() +
         theme(axis.text.x = element_text(angle = 90, hjust = 1))
+ggsave(here("figs/grazing_imp.png"))
 ```

### 2018-11-26 r/Data_prodding.Rmd
11/26

@@
               column_to_rownames(var="site") %>%
-              select(ma_ind,ta_cover,ta_canopy,lc_cover,rugosity,scar_den,scar_bm,spear_daily)),
+              select(ma_cover, ma_canopy, ta_cover, ta_canopy, lc_cover, rugosity, scar_den, 
+                     scar_bm, spear_daily)),
   center = T,scale = T)
@@
 ```{r}
-pca_all <- prcomp(na.omit(sum_site %>% 
+pca_bent <- prcomp(na.omit(sum_site %>% 
               remove_rownames %>% 
               column_to_rownames(var="site") %>%
-              # filter(island!="Barbuda") %>%
-              select(ma_cover, ma_canopy, ta_cover, ta_canopy, rugosity, scar_bm, scar_den)),
+              select(ma_cover, ma_canopy, ta_cover, ta_canopy, lc_cover, rugosity)), 
   center = T,scale = T)
+summary(pca_bent)
+plot(pca_bent,type="l")
 
-autoplot(pca_all, data = sum_site, colour = 'island',
+autoplot(pca_bent, data = sum_site, colour = 'island',
          # frame = T, frame.type = "norm",
@@
 
-pca_bent <- prcomp(na.omit(sum_site %>% 
-              remove_rownames %>% 
-              column_to_rownames(var="site") %>%
-              select(ma_ind,ta_cover,ta_canopy,lc_cover,rugosity)), 
-  center = T,scale = T)
-summary(pca_bent)
-plot(pca_bent,type="l")
+ggsave(filename = here("figs/pca_bent.png"))
 
@@
          pc2_bent=PC2) %>% 
-  bind_cols(as.data.frame(pca_all$x) %>% 
+  bind_cols(as.data.frame(pca_pred$x) %>% 
            select(PC1,PC2)) %>%
@@
   left_join(select(sum_site,site,pc1_bent,pc2_bent,pc1_all,pc2_all), by = "site")
+write.csv(sum_id_pca, file=here("data/sum_id_pca.csv"))
 
+sum_ssp_pca <- sum_site_sp_ph %>%
+  left_join(select(sum_site,site,pc1_bent,pc2_bent,pc1_all,pc2_all), by = "site")
+write.csv(sum_ssp_pca, file=here("data/sum_ssp_pca.csv"))
 ```
@@
   stat_n_text()
+ggsave(filename = here("figs/bp_fr.png"))
 
@@
   theme(plot.title = element_text(hjust = 0.5))
+ggsave(filename = here("figs/vp_fr.png"))
 
@@
        aes(factor(island), g_frac)) + 
+  geom_boxplot() + 
+  labs(y = "Fraction of time", x = "",title="Grazing time budgets (i. phase, 16-24cm)") +
+  geom_signif(comparisons = list(c("Antigua", "Barbuda"),c("Antigua","Bonaire"),c("Barbuda","Bonaire")), map_signif_level=TRUE) +
+  facet_grid(.~species) + 
+  stat_n_text() +
+  scale_fill_brewer(palette="Blues") +
+  theme_bw() +
+  theme(strip.text = element_text(face = "italic")) +
+  theme(plot.title = element_text(hjust = 0.5)) 
+ggsave(here("figs/bp_gfrac.png"))
+
+ggplot(filter(sum_id, 
+              phase=="i" & length_cm>=16 & length_cm<=24 & species_code!="rbp"), 
+       aes(factor(island), g_frac)) + 
   geom_violin(aes(fill=island)) +
@@
   theme(plot.title = element_text(hjust = 0.5))
+ggsave(here("figs/vp_gfrac.png"))
 ```
@@
        aes(factor(island), br)) + 
+  geom_boxplot() + 
+  labs(y = "Bite Rate (bites/sec)", x = "", title = "Bite rate while grazing (i. phase, 16-24cm)") +
+  geom_signif(comparisons = list(c("Antigua", "Barbuda"),c("Antigua","Bonaire"),c("Barbuda","Bonaire")), map_signif_level=TRUE) +
+  facet_grid(.~species) + 
+  stat_n_text() +
+  scale_fill_brewer(palette="Blues") +
+  theme_bw() +
+  theme(strip.text = element_text(face = "italic")) +
+  theme(plot.title = element_text(hjust = 0.5)) 
+ggsave(here("figs/bp_br.png"))
+
+ggplot(filter(sum_id, 
+              phase=="i" & length_cm>=16 & length_cm<=24 & species_code!="rbp"), 
+       aes(factor(island), br)) + 
   geom_violin(aes(fill=island)) +
@@
   theme(plot.title = element_text(hjust = 0.5)) 
+ggsave(here("figs/vp_br.png"))
 ```
@@
        aes(factor(island), for_bites)) + 
-  geom_violin(aes(fill=island)) +
-  geom_boxplot(width=0.1) + 
-  labs(y = "Bite Rate (bites/sec)", x = "", title = "Bite rate while grazing (i. phase, 16-24cm)") +
+  geom_boxplot() + 
+  labs(y = "Bites per foray", x = "", title = "Bites per foray (i. phase, 16-24cm)") +
   geom_signif(comparisons = list(c("Antigua", "Barbuda"),c("Antigua","Bonaire"),c("Barbuda","Bonaire")), map_signif_level=TRUE) +
@@
   theme(plot.title = element_text(hjust = 0.5)) 
+ggsave(here("figs/bp_fordur.png"))
+
+ggplot(filter(sum_id, 
+              phase=="i" & length_cm>=16 & length_cm<=24 & species_code!="rbp"), 
+       aes(factor(island), for_bites)) + 
+  geom_violin(aes(fill=island)) +
+  geom_boxplot(width=0.1) + 
+  labs(y = "Bites per foray", x = "", title = "Bites per foray (i. phase, 16-24cm)") +
+  geom_signif(comparisons = list(c("Antigua", "Barbuda"),c("Antigua","Bonaire"),c("Barbuda","Bonaire")), map_signif_level=TRUE) +
+  facet_grid(.~species) + 
+  stat_n_text() +
+  scale_fill_brewer(palette="Blues") +
+  theme_bw() +
+  theme(strip.text = element_text(face = "italic")) +
+  theme(plot.title = element_text(hjust = 0.5)) 
+ggsave(here("figs/vp_fordur.png"))
 
@@
   theme_bw() + 
+  theme(strip.text = element_text(face = "italic")) +
   theme(plot.title = element_text(hjust = 0.5)) + 
   scale_color_manual(values = c("navy", "darkseagreen", "slategray2"))
+ggsave(here("figs/sp_fr_length.png"))
 ```

### 2018-11-26 r/Follows.Rmd
11/26

@@
     int_dam_mean=mean(int_dam),
-    int_dam_se=sd(int_dam)/sqrt(n)
+    int_dam_se=sd(int_dam)/sqrt(n),
+    length_m = mean(length_cm),
+    length_se = sd(length_cm)/sqrt(n)
     ) %>% 

### 2018-11-26 r/Models.Rmd
11/26

@@
 library(car)
+library(ggfortify)
 
@@
 sum_site <- read.csv(here("data","sum_site.csv"), stringsAsFactors=F) %>%
-  select(-X) %>%
-  mutate(island = factor(island, levels = c("Bonaire","Antigua","Barbuda"))) %>%
-  arrange(island)
+  select(-X)
+sum_ssp <- read.csv(here("data","sum_site_sp_ph.csv"), stringsAsFactors=F) %>%
+  select(-X)
 ```
@@
 Carnivore biomass and spearfishing presence is also highly correlated with scarid biomass and thus benthic PCA variables - I am excluding it for now.
+
+```{r,warning=FALSE,echo=FALSE}
+predictors <- sum_site %>% 
+  select(tot_bm,tot_den,scar_bm,scar_den,carn30_bm,carn30_den,
+         cca_cover,lc_cover,ma_cover,ma_canopy,ta_cover,ta_canopy,rugosity,spear_daily) %>% 
+  gather() %>% 
+  na.omit()
+
+ggplot(predictors,aes(value)) +
+  facet_wrap(~ key, scales = "free") +
+  geom_histogram(aes(y=..density..), colour="black", fill="white") +
+  geom_density(alpha=.2, fill="#FF6666") 
+```
 
@@
   left_join(select(sum_site,site,pc1_bent,pc2_bent,pc1_all,pc2_all), by = "site")
+sum_ssp_pca <- sum_ssp %>% 
+  left_join(select(sum_site,site,pc1_bent,pc2_bent,pc1_all,pc2_all), by = "site")
 
@@
 
-# Model as LME first
+# Examining potential response variables
+Feeding rate (fr) is primary variable of interest
+```{r,echo=FALSE,warning=FALSE}
+response <- sum_id %>% 
+  select(g_frac,fr,br,for_dur,for_bites) %>% 
+  gather() %>% 
+  na.omit()
+
+ggplot(response,aes(value)) + 
+  facet_wrap(~ key, scales = "free") + 
+  geom_histogram(aes(y=..density..), colour="black", fill="white") +
+  geom_density(alpha=.2, fill="#FF6666") 
+```
+
+# Model as LME first, check model validity  
+I won't get too far into dealing with heteroskedasticity here, because it may be improved when switching to GAMM
 ```{r}
+m_data <- sum_id_pca %>% filter(species_code!="rbp" & fr<=3000) # removing outlier
 lme1 <- lme(fr ~ species + phase + length_cm + scar_bm + pc1_bent + pc2_bent, data=m_data,
@@
 vif(lme1)
+plot(lme1, which = c(1), col = 1)
+plot(m_data$length_cm, resid(lme1))
+plot(m_data$scar_bm, resid(lme1)) # highest scar_cm site is big jump from other intervals - scale?
+plot(m_data$pc1_bent, resid(lme1))
+plot(m_data$pc2_bent, resid(lme1)) # potential interval/scale issues?
+plot(factor(m_data$phase), resid(lme1))
+plot(factor(m_data$species), resid(lme1)) # much greater spread in S. vetula data than S. viride
+
 ```
+
+_Dealing with heteroskedasticity_: clear heteroskedasticity in my data, likely with multiple sources...  
+- differences in variance between species, but modelling species independently still does not fully account for heteroskedasticity (below)
+- scar_cm and pc2_bent have unevenly distributed intervals - scale?
+
+_Options/notes from Zuur et al. 2009 (Ch.4)_:  
+- add variance covariate (p.78)
+- log transform DV (p.131)
+
+Dealing with species separately:
+
+```{r}
+# S. vetula only
+qup_data <- sum_id_pca %>% filter(species_code=="qup")
+lme2 <- lme(fr ~ phase + length_cm + scar_bm + pc1_bent + pc2_bent, data=qup_data,
+    random = ~1 | island, 
+    na.action = na.omit
+    )
+summary(lme2)
+vif(lme2)
+plot(lme2, which = c(1), col = 1)
+
+# S. viride only
+vir_data <- sum_id_pca %>% filter(species_code=="stop")
+lme3 <- lme(fr ~ phase + length_cm + scar_bm + pc1_bent + pc2_bent, data=vir_data,
+    random = ~1 | island, 
+    na.action = na.omit
+    )
+summary(lme3)
+vif(lme3)
+plot(lme3, which = c(1), col = 1)
+```
+
+Log transforming feeding rate variable - still heteroskedasticity, largely from observations with fr=0?
+
+```{r}
+m_data_tr <- sum_id_pca %>% 
+  filter(species_code!="rbp" & fr<=3000) # removing outlier
+  # mutate(fr_c = scale(fr, scale = F)) # trying to center fr variable
+lme4 <- lme(log10(fr/100 + 1) ~ species + phase + length_cm + scar_bm + pc1_bent + pc2_bent, data=m_data_tr,
+    random = ~1 | island, 
+    na.action = na.omit
+    )
+summary(lme4)
+vif(lme4)
+plot(lme4, which = c(1), col = 1)
+```
+
+Because of 0 values,
 
@@
 ```{r}
-m1 <- gam(fr ~ factor(species) + s(length_cm, bs = "cs") + s(pc1_all, bs = "cs") + s(pc2_all, bs = "cs"), data=m_data)
+m1 <- gam(fr ~ factor(species) + s(length_cm, bs = "cs") + s(scar_bm, bs = "cs") + 
+            s(pc1_bent, bs = "cs") + s(pc2_bent, bs = "cs"), data = m_data)
 summary(m1)
@@
 gam.check(m1)
-
-gls1 <- gam(fr ~ species + length_cm + pc1_all + pc2_all, data=m_data)
-summary(gls1)
 
@@
 coplot(e ~ pc2_all | species, data = m_data) # inconsistent res.
+coplot(e ~ scar_bm | species, data = m_data)
 
-m2 <- gam(fr ~ factor(species_code) + s(length_cm) + s(pc1_all) +s (pc2_all) +
+m2 <- gam(fr ~ factor(species_code) + s(length_cm) + s(scar_bm) + 
+          s(pc1_bent, bs = "cs") + s(pc2_bent, bs = "cs") +
           s(length_cm, by = as.numeric(species_code=="qup")), 
@@
 
-m3 <- gam(fr ~ factor(species_code) + s(length_cm) + s(pc1_all) + s(pc2_all) +
-          s(pc1_all, by = as.numeric(species_code == "qup")),
+m3 <- gam(fr ~ factor(species_code) + s(length_cm) + s(scar_bm) + s(pc1_bent) + s(pc2_bent) +
+          s(pc1_bent, by = as.numeric(species_code == "qup")),
           data = m_data)
@@
 
-m4 <- gam(fr ~ factor(species_code) + s(length_cm) + s(pc1_all) + s(pc2_all) +
-          s(pc2_all, by = as.numeric(species_code == "qup")),
+m4 <- gam(fr ~ factor(species_code) + s(length_cm) + s(scar_bm) + s(pc1_bent) + s(pc2_bent) +
+          s(pc2_bent, by = as.numeric(species_code == "qup")),
           data = m_data)
@@
 
-m5 <- gam(fr ~ factor(species_code) + s(length_cm) + s(pc1_all) + s(pc2_all) +
-          s(pc1_all, by = as.numeric(species_code == "qup"))+
-          s(pc2_all, by = as.numeric(species_code == "qup")),
+m5 <- gam(fr ~ factor(species_code) + s(length_cm) + s(scar_bm) + s(pc1_bent) + s(pc2_bent) +
+          s(pc1_bent, by = as.numeric(species_code == "qup")) +
+          s(pc2_bent, by = as.numeric(species_code == "qup")) +
+          s(length_cm, by = as.numeric(species_code == "qup")) +
+          s(scar_bm, by = as.numeric(species_code == "qup")),
           data = m_data)
@@
 gam.check(m5)
-
 ```
 
-Notes:  
+_Model validity notes_:  
 - Violates assumption of homogeneity of variance (increasing variance as predictor values increase)
@@
 
+*need to add interaction terms*
 ```{r}
-gamm<-gamm(fr~phase+s(length_cm)+s(pc1)+s(pc2)+s(scar_bm)+s(scar_den), 
-           family=gaussian(link=identity),
-           data=filter(sum_id_pca1, species_code=="qup"),
-           random=list(site=~1),
-           method="REML"
-           )
-summary(gamm$gam)
-plot(gamm$gam,pages=1)
-```
-```{r}
-gamm1<-gamm(fr_mean~phase+species+s(pc1)+s(pc2)+s(scar_bm)+s(scar_den), 
-           family=gaussian(link=identity),
-           data=filter(sum_ssp_pca1, species_code!="rbp" & island!="Barbuda"),
-           random=list(site=~1),
-           method="REML"
+gamm1 <- gamm(fr ~ species_code + phase + s(length_cm) + s(pc1_bent) + s(pc2_bent) + s(scar_bm) +
+                s(length_cm, by = as.numeric(species_code == "qup")) +
+                s(pc1_bent, by = as.numeric(species_code == "qup")) +
+                s(scar_bm, by = as.numeric(species_code == "qup")), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           method = "REML",
+           weights = varIdent(form =~ 1 | species),
+           data = m_data
            )
 summary(gamm1$gam)
-gamtabs(gamm1)
+plot(gamm1$gam,pages=1)
+plot(gamm1$lme) # shows heteroskedasticity
 AIC(gamm1$lme)
-plot(gamm1$gam,pages=1)
+```
 
+*Re: Nash et al. 2016: should I be aggregating data to the site level?*
+```{r}
+m_data2 <- sum_ssp_pca %>% filter(species_code!="rbp") # currently doesn't work when removing rbp data
+gamm2 <- gamm(fr_mean ~ species_code + phase + s(length_m) + s(pc1_bent) + s(pc2_bent) + s(scar_bm), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           method = "REML",
+           data = sum_ssp_pca
+           )
+summary(gamm2$gam)
+plot(gamm2$gam,pages=1)
+plot(gamm2$lme) # not too heteroskedastic?
+AIC(gamm2$lme)
+
+# Confirming low VIF among predictor variables with site-level model
+lme1 <- lme(fr_mean ~ phase + length_m + scar_bm + pc1_bent + pc2_bent, data=sum_ssp_pca,
+    random = ~1 | island, 
+    na.action = na.omit
+    )
+vif(lme1)
 ```

### 2018-11-26 r/SEM.Rmd
11/26

@@
 library(semPlot)
+library(piecewiseSEM)
 ```
@@
 sum_site <- read.csv(here("data","sum_site.csv"), stringsAsFactors=FALSE) %>% select(-X)
+sum_ssp <- read.csv(here("data","sum_site_sp_ph.csv"), stringsAsFactors=FALSE) %>% select(-X)
+```
 
-# recode phase and species with dummy variables
-# data <- data %>% mutate(spec.dum= 
-#                           ifelse(Species.Code=="stop", 1, 
-#                           (ifelse(Species.Code=="qup", 2,
-#                           (ifelse(Species.Code=="rbp",3,NA)))))) %>%
-#   mutate(phase.dum=ifelse(Phase=="i",1,2))
-
-```
+# Full model layout
 
@@
 
+# ID level SEMs
+
 ```{r}
-# testing partial SEMs
+# testing partial SEMs with ID - level dataset
 
 sem_df <- sum_id %>% 
-  filter(island!="Barbuda") %>%
+  # filter(island!="Barbuda") %>%
+  filter(species_code != "rbp") %>%
   select(fr, g_frac, species_code, phase, scar_bm, scar_den, carn30_bm, tot_bm, spear_daily, ma_cover, ma_canopy, ta_cover, ta_canopy, rugosity, int_scar, length_cm, lc_cover) %>% 
-  mutate(phase_dum=ifelse(phase=="i",1,2),
+  mutate(phase_dum=ifelse(phase=="i",0,1),
+         species_dum = ifelse(species_code=="stop",0,1),
          fr_sc = fr/100,
@@
 
-fit <- sem(sem3, data=vet_df, missing="ML", std.lv=T, orthogonal=T)
+fit <- sem(sem1, data=sem_df, missing="ML", std.lv=T, orthogonal=T)
 varTable(fit)
@@
 ```
+
+# Site-level SEM
+
+```{r}
+sem_df <- sum_ssp %>% 
+  select(fr_mean, species_code, phase, scar_bm, scar_den, carn30_bm, tot_bm, spear_daily, ma_cover, ma_canopy, ta_cover, ta_canopy, rugosity, int_scar_mean, length_m, lc_cover) %>% 
+  filter(fr_mean > 0,
+         species_code!="rbp") %>%
+  mutate(phase_dum = ifelse(phase=="i",0,1),
+         species_dum = ifelse(species_code=="stop",0,1),
+         fr_sc = fr_mean/100,
+         scar_bm_sc = scar_bm/100,
+         rugosity_sc = rugosity*10) %>%
+  na.omit()
+vet_df <- sem_df %>% filter(species_code=="qup")
+vir_df <- sem_df %>% filter(species_code=="stop")
+
+sem1 <- '
+  # latent variable indicators
+      risk =~ scar_bm_sc + scar_den + rugosity_sc + spear_daily
+      nutrition =~ ma_cover + ta_cover
+  # regressions
+      fr_sc ~ species_dum + length_m + phase_dum + nutrition + risk
+      rugosity_sc ~ lc_cover
+      lc_cover ~ ma_cover + ta_cover
+'
+
+sem2 <- '
+  # latent variable indicators
+      risk =~ scar_bm_sc + scar_den
+      nutrition =~ ma_cover + ta_cover
+  # regressions
+      fr_sc ~ species_dum + length_m + phase_dum + nutrition + risk
+'
+
+fit <- sem(sem2, data=sem_df, missing="ML", std.lv=T)
+varTable(fit)
+summary(fit, standardized=T, fit.measures=T)
+semPaths(fit, 'std', residuals = F)
+fitMeasures(fit, c("cfi","rmsea","srmr"))
+
+```
+
+# Piecewise SEM
+
+```{r}
+sum_id_nar <- sum_id %>% drop_na()
+psem1 <- psem(
+  lm(fr ~ phase + species_code + scar_bm + ta_cover, sum_id_nar),
+  lm(ta_cover ~ scar_bm, sum_id_nar))
+summary(psem1)
+```

### 2018-11-26 r/Status Nov. 2018.Rmd
11/26

---
title: "Status Nov. 2018"
author: "Molly Wilson"
date: "Dec. 2018"
output:
  html_document:
    toc: true
    toc_float: true
    theme: united
    number_sections: true
---

```{r, echo=F, message=F, warning=F}
library(startR)
library(here)
library(tidyverse)
library(mgcv)
library(car)

sum_id_pca <- read.csv(here("data","sum_id_pca.csv"), stringsAsFactors=F) %>%
  select(-X) 
sum_ssp_pca <- read.csv(here("data","sum_ssp_pca.csv"), stringsAsFactors=F) %>%
  select(-X) 
```

# Intro + motivation

* Social-ecological role of CR herbivores, efforts to manage and maximize grazing impact
* Management targets/thresholds based on direct correlation between biomass and grazing impact
* In the Caribbean, data on parrotfish feeding behaviors are taken largely from Bonaire, one of the regions most pristine reef systems
* But little appreciation of fish behavior, understanding nuances of grazing impact
* Fished systems have been shown to exhibit altered flight behaviors, but implications for ecological function have not been assessed
* In this study I compare parrotfish grazing behavior between Bonaire, where spearfishing has been banned since the 1970s and parrotfish harvest was made illegal in the early 2000s, and Antigua, where spearfishing is prevalent and parrotfish populations have been substantially depleted.

# Methods

## Data collection
* Sites (map), timeline
* Fish and benthic community data
* Behavioral data:
    - Species: *Sparisoma viride* and *Scarus vetula*, initial and terminal phase greater than 15cm
    - 2 minute follows, calibration period, outside of previously established FID
    - Recording time budgets as well as bite rates during grazing forays (define)
    - Discarded data of incomplete follows or where fish were visably disturbed/fled
    - Timing of data collection distributed throughout day

## Analysis
* One-way ANOVAs between islands, run separately for both S. vetula and S. viride on subset of individuals (initial phase, 16-24cm). In the cases where variables showed heteroskedasticity (Levene’s test, p<0.05), I did not assume homogeneity of variance (white.adjust=T within aov() function (car package) in R). If ANOVA results showed significant differences among islands, I used Tukey’s HSD to quantify these differences.
* Predicting community grazing impact: calculated length - feeding rate relationships using linear regression for each phase/island/species, applied these to observed site-level fish communities. Integrated bite size data from Bruggemann et al., then compared with predictions made with published species length - feeding rate relationships
* GAMM: mgcv, island as random effect, tested for collinearity using VIF (cutoff of 3), tested model validity. Used data aggregated to the site level
* SEM: lavaan OR piecewiseSEM?

# Results

## Differences in feeding behaviors among islands

```{r, echo=F, message=F, warning=F}
knitr::include_graphics(here("figs/vp_fr.png"))
knitr::include_graphics(here("figs/vp_br.png"))
knitr::include_graphics(here("figs/vp_gfrac.png"))
```

## Impacts on total grazing pressure

Here I take calculate the predicted reef area grazed per hour of a given scarid (*S. viride* and *S. vetula*) population. I calculate individual predicted grazing impact based on observed and previously published length - feeding rate relationships and multiply these based on published length - bite size relationships for each species.

```{r, echo=F, message=F, warning=F}
knitr::include_graphics(here("figs/sp_fr_length.png"))
```

Length - feeding rate relationships from personal data:  
$$fr(Antigua, S. viride, i) = 445-8*L$$
$$fr(Antigua, S. vetula, i) = 1030-12*L$$
$$fr(Bonaire, S. viride, i) = 1057-17*L$$
$$fr(Bonaire, S. vetula, i) = 1347+10*L$$

Feeding rates as published by Bruggemann/Bozec
$$fr(Published, S. viride, i) = 1089-17*L-56 $$
$$fr(Published, S. vetula, initial) = 3329-33*L$$

Bite sizes as predicted by Bruggemann/Bozec:
$$bs(S. vetula) = 4.013*10^{-4}*L^2$$
$$bs(S. viride) = 5.257*10^{-4}*L^2$$

```{r, echo=F, message=F, warning=F}
knitr::include_graphics(here("figs/areagrazed.png"))
```

## GAMM results

```{r, echo=F, message=F, warning=F}
knitr::include_graphics(here("figs/pca_bent.png"))
```

### Site-level data
```{r, echo=F, message=F, warning=F}
gamm_s <- gamm(fr_mean ~ species_code + phase + s(length_m) + s(pc1_bent) + s(pc2_bent) + s(scar_bm), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = sum_ssp_pca
           )
summary(gamm_s$gam)
plot(gamm_s$gam,pages=1)
plot(gamm_s$lme) # not too heteroskedastic?
AIC(gamm_s$lme)

# Confirming low VIF among predictor variables with site-level model
vif(lme(fr_mean ~ phase + length_m + scar_bm + pc1_bent + pc2_bent, 
        data = sum_ssp_pca,
        random = ~1 | island, 
        na.action = na.omit
        )) # all <3
```

### Individual-level data
```{r}
gamm_id <- gamm(fr ~ species_code + phase + s(length_cm) + s(pc1_bent) + s(pc2_bent) + s(scar_bm) 
              # +
              #   s(length_cm, by = as.numeric(species_code == "qup")) +
              #   s(pc1_bent, by = as.numeric(species_code == "qup")) +
              #   s(scar_bm, by = as.numeric(species_code == "qup"))
              , 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           weights = varIdent(form =~ 1 | species),
           data = sum_id_pca
           )
summary(gamm_id$gam)
plot(gamm_id$gam,pages=1)
plot(gamm_id$lme) # shows some heteroskedasticity
AIC(gamm_id$lme)

# Confirming low VIF among predictor variables with site-level model
vif(lme(fr ~ species + phase + length_cm + scar_bm + pc1_bent + pc2_bent, 
        data = sum_id_pca,
        random = ~1 | island, 
        na.action = na.omit
        )) # prints strange when include RBP data
```

## SEM results

### 2018-11-26 r/Status Nov. 2018.Rmd
11/26 more

@@
 plot(gamm_s$lme) # not too heteroskedastic?
-AIC(gamm_s$lme)
 
-# Confirming low VIF among predictor variables with site-level model
+```
+
+Confirming low VIF among predictor variables with site-level model
+```{r, echo=F, message=F, warning=F}
 vif(lme(fr_mean ~ phase + length_m + scar_bm + pc1_bent + pc2_bent, 
@@
 ### Individual-level data
-```{r}
+```{r, warning=F, message=F, echo=F}
 gamm_id <- gamm(fr ~ species_code + phase + s(length_cm) + s(pc1_bent) + s(pc2_bent) + s(scar_bm) 
@@
 plot(gamm_id$lme) # shows some heteroskedasticity
-AIC(gamm_id$lme)
+```
 
-# Confirming low VIF among predictor variables with site-level model
+Confirming low VIF among predictor variables with id-level model
+```{r, warning=F, message=F, echo=F}
 vif(lme(fr ~ species + phase + length_cm + scar_bm + pc1_bent + pc2_bent, 
@@
 
-## SEM results
+## SEM results?
+
+```{r, echo=F, message=F, warning=F}
+knitr::include_graphics(here("figs/sem.png"))
+```
+
+Notes: Still troubleshooting getting full SEM to run, unsure about how to deal with different data layers/sample sizes.
+If aggregating to the site level, I will be limited in the number of variables I can include (~8), though I could run separate models by species/phase.
+I could also use piecewise SEM (e.g.
+Miller et al.
+2018), though does not allow for latent variables.

### 2018-11-27 r/Data_prodding.Rmd
11/26 updates

@@
 library(nlme)
-library(gamm4)
-library(xtable)
+# library(gamm4)
+# library(xtable)
 ```

### 2018-11-27 r/Models.Rmd
11/27

@@
               # filter(island!="Barbuda") %>%
-              select(ma_cover, ma_canopy, ta_cover, ta_canopy, rugosity, scar_bm, scar_den)),
+              select(ma_cover, ma_canopy, ta_cover, ta_canopy, rugosity, scar_bm, scar_den, lc_cover, carn30_bm)),
   center = T,scale = T)
@@
 ```
+
+Site-level GAMM with pca_all
+```{r}
+m_data <- sum_ssp_pca %>% filter(species_code!="rbp")
+
+# Main effects only
+gamm1 <- gamm(fr_mean ~ species_code + s(length_m) + phase + s(pc1_all) + s(pc2_all), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           method = "REML",
+           data = m_data
+           )
+summary(gamm1$gam)
+plot(gamm1$gam,pages=1)
+plot(gamm1$lme) # not too heteroskedastic?
+AIC(gamm1$lme)
+
+# Interactions
+gamm2 <- gamm(fr_mean ~ species_code + phase + s(length_m, by = factor(species_code)) + s(pc2_all)  + s(pc1_all, by = factor(species_code)), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           method = "REML",
+           data = m_data
+           )
+summary(gamm2$gam)
+plot(gamm2$gam,pages=1)
+plot(gamm2$lme) # not too heteroskedastic?
+AIC(gamm2$lme)
+# Slightly significant interaction between pc1_all and species_code=qup, but this has lower R2 than model with all main effects
+# length/species and pc2/species interactions were not significant
+
+# Reducing non-significant effects
+gamm3 <- gamm(fr_mean ~ species_code + phase + s(pc1_all) + s(pc2_all), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           method = "REML",
+           data = m_data
+           )
+summary(gamm3$gam)
+plot(gamm3$gam,pages=1)
+plot(gamm3$lme) # not too heteroskedastic?
+AIC(gamm3$lme)
+
+# Adding pc1/species interaction back in to model with streamlined main effects
+gamm4 <- gamm(fr_mean ~ species_code + phase + s(pc1_all, by = factor(species_code)), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           method = "REML",
+           data = m_data
+           )
+summary(gamm4$gam)
+plot(gamm4$gam,pages=1)
+plot(gamm4$lme) # not too heteroskedastic?
+AIC(gamm4$lme)
+
+```

### 2018-11-28 r/Models.Rmd
11/27 more

@@
 library(ggfortify)
+library(xtable)
+library(itsadug)
 
@@
 ```{r}
-gamm1 <- gamm(fr ~ species_code + phase + s(length_cm) + s(pc1_bent) + s(pc2_bent) + s(scar_bm) +
-                s(length_cm, by = as.numeric(species_code == "qup")) +
-                s(pc1_bent, by = as.numeric(species_code == "qup")) +
-                s(scar_bm, by = as.numeric(species_code == "qup")), 
-           family = gaussian(link=identity),
-           random = list(island=~1),
-           method = "REML",
-           weights = varIdent(form =~ 1 | species),
-           data = m_data
-           )
-summary(gamm1$gam)
-plot(gamm1$gam,pages=1)
-plot(gamm1$lme) # shows heteroskedasticity
-AIC(gamm1$lme)
+# gamm1 <- gamm(fr ~ species_code + phase + s(length_cm) + s(pc1_bent) + s(pc2_bent) + s(scar_bm) +
+#                 s(length_cm, by = as.numeric(species_code == "qup")) +
+#                 s(pc1_bent, by = as.numeric(species_code == "qup")) +
+#                 s(scar_bm, by = as.numeric(species_code == "qup")), 
+#            family = gaussian(link=identity),
+#            random = list(island=~1),
+#            method = "REML",
+#            weights = varIdent(form =~ 1 | species),
+#            data = m_data
+#            )
+# summary(gamm1$gam)
+# plot(gamm1$gam,pages=1)
+# plot(gamm1$lme) # shows heteroskedasticity
+# AIC(gamm1$lme)
 ```
@@
 
+gamtabs(gamm4, type="HTML")
 ```

### 2018-11-28 r/Manuscript draft.Rmd
11/27 more

--- 2018-11-26 r/Status Nov. 2018.Rmd
@@
 ---
-title: "Status Nov. 2018"
+title: "Manuscript draft"
 author: "Molly Wilson"
@@
 sum_id_pca <- read.csv(here("data","sum_id_pca.csv"), stringsAsFactors=F) %>%
-  select(-X) 
+  select(-X) %>% na.omit()
 sum_ssp_pca <- read.csv(here("data","sum_ssp_pca.csv"), stringsAsFactors=F) %>%
-  select(-X) 
+  select(-X) %>% na.omit()
 ```
@@
 ## Data collection
-* Sites (map), timeline
+* Sites (13, map), timeline
 * Fish and benthic community data
@@
 
-## Differences in feeding behaviors among islands
-
+## Biplots
 ```{r, echo=F, message=F, warning=F}
-knitr::include_graphics(here("figs/vp_fr.png"))
-knitr::include_graphics(here("figs/vp_br.png"))
-knitr::include_graphics(here("figs/vp_gfrac.png"))
-```
-
-## Impacts on total grazing pressure
-
-Here I take calculate the predicted reef area grazed per hour of a given scarid (*S.
-viride* and *S.
-vetula*) population.
-I calculate individual predicted grazing impact based on observed and previously published length - feeding rate relationships and multiply these based on published length - bite size relationships for each species.
-
-```{r, echo=F, message=F, warning=F}
-knitr::include_graphics(here("figs/sp_fr_length.png"))
-```
-
-Length - feeding rate relationships from personal data:  
-$$fr(Antigua, S. viride, i) = 445-8*L$$
-$$fr(Antigua, S. vetula, i) = 1030-12*L$$
-$$fr(Bonaire, S. viride, i) = 1057-17*L$$
-$$fr(Bonaire, S. vetula, i) = 1347+10*L$$
-
-Feeding rates as published by Bruggemann/Bozec
-$$fr(Published, S. viride, i) = 1089-17*L-56 $$
-$$fr(Published, S. vetula, initial) = 3329-33*L$$
-
-Bite sizes as predicted by Bruggemann/Bozec:
-$$bs(S. vetula) = 4.013*10^{-4}*L^2$$
-$$bs(S. viride) = 5.257*10^{-4}*L^2$$
-
-```{r, echo=F, message=F, warning=F}
-knitr::include_graphics(here("figs/areagrazed.png"))
+knitr::include_graphics(here("figs/pca_pred.png"))
 ```
@@
 ```{r, echo=F, message=F, warning=F}
-knitr::include_graphics(here("figs/pca_bent.png"))
+knitr::include_graphics(here("figs/pca_pred.png"))
 ```

### 2019-02-01 r/Analysis.Rmd
feb1 updates

@@
 
-## Comparison to published equations
+# Comparison to published equations
 
@@
 
-## adding predicted feeding rates from observed relationships
+## Adding predicted feeding rates from observed relationships
 
@@
 ```
+
+# Bite content hypothesis
+
+Fish in high algal sites may be taking fewer bites because they can consume more algae per bite. Let's try and add a rough estimate of bite content as a function of algal canopy height.
+
+Notes/assumptions:
+* Using turf algal canopy height (for now, no macroalgal component - though S. viride does eat some macrogalgae)
+* Max bite height will be capped at radius of bite size ($\sqrt{BS}/\pi$)
+
+```{r}
+# grazing impact of fish from fish surveys
+fish_grazed <- fish_fr %>% 
+  left_join(select(sum_site, ta_canopy, site), by = "site") %>%
+  mutate(
+    bh_capped = ifelse(
+      ta_canopy < 2*sqrt(bs)/(22/7), ta_canopy, 2*sqrt(bs)/(22/7)), # if ta_canopy exceeds diameter of bite size, bite height is capped at bite size
+    bv = bs*ta_canopy,
+    bv_capped = bs*bh_capped,
+    vg = fr_pers*bv
+ )
+
+# grazing impact of fish from behavioral surveys
+sum_id_vg <- sum_id %>%
+  filter(species_code != "rbp") %>%
+  mutate(
+    bs = ifelse(species_code == "stop", 5.257*10^-4*length_cm^2, 4.013*10^-4*length_cm^2), # adding bite size equations from Mumby
+    bh_capped = ifelse(ta_canopy < 2*sqrt(bs)/(22/7), ta_canopy, 2*sqrt(bs)/(22/7)), # if ta_canopy exceeds diameter of bite size, bite height is capped at bite size
+    bv = bs*ta_canopy,
+    bv_capped = bs*bh_capped,
+    vg = fr*bv,
+    vg_capped = fr*bv_capped,
+    island = factor(island, levels = c("Bonaire","Antigua","Barbuda"))
+) %>%
+  arrange(island)
+
+sum_ssp_vg <- sum_id_vg %>% 
+  group_by(site, island, species, species_code, phase) %>% 
+  summarize(
+    n=n(),
+    vg_mean=mean(vg),
+    vg_se=sd(vg)/sqrt(n),
+    vg_capped_mean=mean(vg_capped),
+    vg_capped_se=sd(vg_capped)/sqrt(n)
+    ) %>% 
+  select(-n)
+
+ggplot(filter(sum_ssp_grazed, species_code != "rbp" & phase == "i"), aes(x=factor(site, levels = sum_site$site), y=vg_mean)) +
+    geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
+    geom_errorbar(aes(ymin=vg_mean-vg_se, ymax=vg_mean+vg_se), width=.2,
+                 position=position_dodge(.9)) +
+    facet_grid(~species) +
+    scale_fill_brewer(palette = "Blues") +
+    labs(title="Mean grazing impact by volume (initial phase)", 
+         y="Grazing impact (cm3/hr)", 
+         x="Site") +
+    theme_bw() +
+    theme(strip.text = element_text(face = "italic"), 
+          axis.text.x = element_text(angle = 90, hjust = 1), 
+          plot.title = element_text(hjust = 0.5))
+
+ggplot(filter(sum_ssp_grazed, species_code != "rbp" & phase == "i"), aes(x=factor(site, levels = sum_site$site), y=vg_capped_mean)) +
+    geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
+    geom_errorbar(aes(ymin=vg_capped_mean-vg_capped_se, ymax=vg_capped_mean+vg_capped_se), width=.2,
+                 position=position_dodge(.9)) +
+    facet_grid(~species) +
+    scale_fill_brewer(palette = "Blues") +
+    labs(title="Mean grazing impact by volume (capped height, initial phase)", 
+         y="Grazing impact (cm3/hr)", 
+         x="Site") +
+    theme_bw() +
+    theme(strip.text = element_text(face = "italic"), 
+          axis.text.x = element_text(angle = 90, hjust = 1), 
+          plot.title = element_text(hjust = 0.5))
+```

### 2019-02-01 r/Ch2_outline_V2.tex
feb1 updates

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Scarid grazing behavior - manuscript draft},
            pdfauthor={Molly Wilson},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}

  \title{Scarid grazing behavior - manuscript draft}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{Molly Wilson}
    \preauthor{\centering\large\emph}
  \postauthor{\par}
      \predate{\centering\large\emph}
  \postdate{\par}
    \date{Nov. 2018}

\begin{document}
\maketitle

{
\setcounter{tocdepth}{2}
\tableofcontents
}
\section{Intro + motivation}\label{intro-motivation}

\begin{itemize}
\tightlist
\item
  Grazing by herbivorous fish is an established driver of coral reef
  health/benthic dynamics
\item
  As reefs become increasingly degraded/threatened, there is growing
  interest in understanding how much herbivory is needed to sustain
  coral reef health (e.g.~herbivore threshold papers - Karr et al. 2015,
  McClanahan et al. 2011)
\item
  However, not all herbivores are the same in terms of their grazing
  impact\ldots{}

  \begin{itemize}
  \tightlist
  \item
    Recent developments in functional classifications, difference across
    species in terms of feeding selectivity, morphology, etc. (e.g.~Adam
    et al. 2018)
  \end{itemize}
\item
  In addition to these inter-species differences in functional role,
  differences in herbivore behavior also affect grazing impact and may
  drive variation in herbivore function among individuals or across reef
  systems
\item
  Previous work has shown that herbivore behavior is sensitive
  to\ldots{}

  \begin{itemize}
  \tightlist
  \item
    Predator presence (and reef structure) (Catano, Madin)
  \item
    Presence of other herbivores (Gil)
  \item
    Resource availability + distribution (Carlson, Tootell, Mumby, etc.)
  \end{itemize}
\item
  In understanding and predicting herbivore impact on reefs, current
  assumption is that grazing behaviors are constant within species (same
  species-level length\textasciitilde{}feeding rate relationships)
\item
  But there is very limited data on how herbivore feeding behaviors
  might vary across reef systems and whether we can accurately assume
  constant species roles at regional scales
\item
  Here we document variability in multiple components (\emph{if I
  include foraging intensity as well as feeding rate\ldots{}}) of scarid
  feeding behavior across reef sites in three Caribbean islands and
  discuss potential underlying mechanisms, ecological consequences, and
  management implications
\item
  We demonstrate a need for further experimental and modelling work to
  unpack these drivers and understand their ecological consequences
\item
  We urge for the incorporation of behavioral effects in coral reef
  herbivore management and caution against assuming constant species
  function across reef systems
\end{itemize}

\section{Methods}\label{methods}

\subsection{Data collection}\label{data-collection}

\begin{itemize}
\tightlist
\item
  Sites (13 sites, 3 islands {[}map{]}), timeline of data collection
\item
  Fish and benthic community surveys
\item
  Observational behavioral data:

  \begin{itemize}
  \tightlist
  \item
    Species: \emph{Sparisoma viride} and \emph{Scarus vetula}, initial
    and terminal phase greater than 15cm
  \item
    2 minute follows, calibration period, outside of previously
    established FID
  \item
    Recording time budgets as well as bite rates during grazing forays
    (define)
  \item
    Discarded data of incomplete follows or where fish were visably
    disturbed/fled
  \item
    Timing of data collection distributed throughout day, 2 dives per
    site
  \item
    Define behavioral variables ({[}table{]})
  \end{itemize}
\end{itemize}

\subsection{Analysis}\label{analysis}

\begin{itemize}
\tightlist
\item
  Fish biomass using lenth-weight conversions (appendix)
\item
  (stats for inter-island differences?)
\item
  PCA of site-level benthic and fish community variables
\item
  GAMM: mgcv, island as random effect, Gaussian distribution, compared
  using AICc. Used data aggregated to the site level (mean). Average
  length as covariate.
\end{itemize}

\section{Results}\label{results}

\subsection{Variation among sites
(islands)}\label{variation-among-sites-islands}

\includegraphics{Ch2_outline_V2_files/figure-latex/unnamed-chunk-2-1.pdf}

\subsection{Variation in feeding rate across benthic and fish community
variables}\label{variation-in-feeding-rate-across-benthic-and-fish-community-variables}

\begin{itemize}
\tightlist
\item
  Initial phase only
\item
  Don't need to include all of these
\item
  Total scarid density, conspecific density, and predator biomass had
  even more variance - didn't include here for scaling purposes but
  ultimately might want to add back in even variables with no
  discernable trend?
\end{itemize}

\includegraphics{Ch2_outline_V2_files/figure-latex/unnamed-chunk-3-1.pdf}

\subsection{Variation in feeding intensity across benthic and fish
community
variables}\label{variation-in-feeding-intensity-across-benthic-and-fish-community-variables}

\emph{\emph{Note:} Probably not useful to show both feeding rate and
feeding intensity graphs as trends are very similar\ldots{} but does
have potential impacts on grazing concentration which has potential
ecological relevance. But maybe can show that this varies across sites
without showing all the driver biplots}
\includegraphics{Ch2_outline_V2_files/figure-latex/unnamed-chunk-4-1.pdf}

\subsection{PCA}\label{pca}

\includegraphics{Ch2_outline_V2_files/figure-latex/unnamed-chunk-5-1.pdf}
\includegraphics{Ch2_outline_V2_files/figure-latex/unnamed-chunk-6-1.pdf}

\includegraphics{Ch2_outline_V2_files/figure-latex/unnamed-chunk-7-1.pdf}

\subsection{GAMM results}\label{gamm-results}

One of the GAMM iterations - has lowest AICc, eventually would present
table with other trials/results.

\begin{verbatim}
## 
## Family: gaussian 
## Link function: identity 
## 
## Formula:
## fr_mean ~ species_code + phase + s(length_m, by = factor(species_code), 
##     k = -1) + s(pc1_all, by = factor(species_code), k = -1) + 
##     s(pc2_all, k = -1)
## 
## Parametric coefficients:
##                  Estimate Std. Error t value Pr(>|t|)    
## (Intercept)        979.81     111.33   8.801 1.88e-10 ***
## species_codestop  -403.63      62.63  -6.444 1.89e-07 ***
## phaset            -309.86     115.19  -2.690   0.0108 *  
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
## 
## Approximate significance of smooth terms:
##                                        edf Ref.df     F  p-value    
## s(length_m):factor(species_code)qup  3.755  3.755 2.392 0.041677 *  
## s(length_m):factor(species_code)stop 1.000  1.000 2.678 0.110417    
## s(pc1_all):factor(species_code)qup   2.719  2.718 7.031 0.000919 ***
## s(pc1_all):factor(species_code)stop  1.000  1.000 0.000 0.995405    
## s(pc2_all)                           1.000  1.000 0.018 0.894012    
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
## 
## R-sq.(adj) =  0.723   
##   Scale est. = 39781     n = 48
\end{verbatim}

\includegraphics{Ch2_outline_V2_files/figure-latex/unnamed-chunk-9-1.pdf}

\begin{verbatim}
## 
## Family: gaussian 
## Link function: identity 
## 
## Formula:
## fr_mean ~ species_code + phase + s(length_m, by = factor(species_code), 
##     k = 3) + s(pc1_all, by = factor(species_code), k = 3) + s(pc2_all, 
##     k = 3)
## 
## Parametric coefficients:
##                  Estimate Std. Error t value Pr(>|t|)    
## (Intercept)       1001.42     139.66   7.170 1.38e-08 ***
## species_codestop  -387.76      67.45  -5.749 1.22e-06 ***
## phaset            -365.91     119.27  -3.068  0.00394 ** 
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
## 
## Approximate significance of smooth terms:
##                                        edf Ref.df     F p-value  
## s(length_m):factor(species_code)qup  1.886  1.886 3.551  0.0301 *
## s(length_m):factor(species_code)stop 1.000  1.000 3.369  0.0741 .
## s(pc1_all):factor(species_code)qup   1.772  1.772 7.410  0.0141 *
## s(pc1_all):factor(species_code)stop  1.000  1.000 0.566  0.4566  
## s(pc2_all)                           1.000  1.000 0.354  0.5553  
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
## 
## R-sq.(adj) =  0.591   
##   Scale est. = 47706     n = 48
\end{verbatim}

\includegraphics{Ch2_outline_V2_files/figure-latex/unnamed-chunk-9-2.pdf}

\subsection{Bite content hypothesis}\label{bite-content-hypothesis}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# grazing impact of fish from behavioral surveys}
\NormalTok{sum_id_vg <-}\StringTok{ }\NormalTok{sum_id }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{filter}\NormalTok{(species_code }\OperatorTok{!=}\StringTok{ "rbp"}\NormalTok{) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}
    \DataTypeTok{bs =} \KeywordTok{ifelse}\NormalTok{(species_code }\OperatorTok{==}\StringTok{ "stop"}\NormalTok{, }\FloatTok{5.257}\OperatorTok{*}\DecValTok{10}\OperatorTok{^-}\DecValTok{4}\OperatorTok{*}\NormalTok{length_cm}\OperatorTok{^}\DecValTok{2}\NormalTok{, }\FloatTok{4.013}\OperatorTok{*}\DecValTok{10}\OperatorTok{^-}\DecValTok{4}\OperatorTok{*}\NormalTok{length_cm}\OperatorTok{^}\DecValTok{2}\NormalTok{), }\CommentTok{# adding bite size equations from Mumby}
    \DataTypeTok{bh_capped =} \KeywordTok{ifelse}\NormalTok{(ta_canopy }\OperatorTok{<}\StringTok{ }\DecValTok{2}\OperatorTok{*}\KeywordTok{sqrt}\NormalTok{(bs)}\OperatorTok{/}\NormalTok{(}\DecValTok{22}\OperatorTok{/}\DecValTok{7}\NormalTok{), ta_canopy, }\DecValTok{2}\OperatorTok{*}\KeywordTok{sqrt}\NormalTok{(bs)}\OperatorTok{/}\NormalTok{(}\DecValTok{22}\OperatorTok{/}\DecValTok{7}\NormalTok{)), }\CommentTok{# if ta_canopy exceeds diameter of bite size, bite height is capped at bite size}
    \DataTypeTok{bv =}\NormalTok{ bs}\OperatorTok{*}\NormalTok{ta_canopy,}
    \DataTypeTok{bv_capped =}\NormalTok{ bs}\OperatorTok{*}\NormalTok{bh_capped,}
    \DataTypeTok{vg =}\NormalTok{ fr}\OperatorTok{*}\NormalTok{bv,}
    \DataTypeTok{vg_capped =}\NormalTok{ fr}\OperatorTok{*}\NormalTok{bv_capped,}
    \DataTypeTok{island =} \KeywordTok{factor}\NormalTok{(island, }\DataTypeTok{levels =} \KeywordTok{c}\NormalTok{(}\StringTok{"Bonaire"}\NormalTok{,}\StringTok{"Antigua"}\NormalTok{,}\StringTok{"Barbuda"}\NormalTok{))}
\NormalTok{) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{arrange}\NormalTok{(island)}

\NormalTok{sum_ssp_vg <-}\StringTok{ }\NormalTok{sum_id_vg }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{group_by}\NormalTok{(site, island, species, species_code, phase) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{summarize}\NormalTok{(}
    \DataTypeTok{n=}\KeywordTok{n}\NormalTok{(),}
    \DataTypeTok{vg_mean=}\KeywordTok{mean}\NormalTok{(vg),}
    \DataTypeTok{vg_se=}\KeywordTok{sd}\NormalTok{(vg)}\OperatorTok{/}\KeywordTok{sqrt}\NormalTok{(n),}
    \DataTypeTok{vg_capped_mean=}\KeywordTok{mean}\NormalTok{(vg_capped),}
    \DataTypeTok{vg_capped_se=}\KeywordTok{sd}\NormalTok{(vg_capped)}\OperatorTok{/}\KeywordTok{sqrt}\NormalTok{(n)}
\NormalTok{    ) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{select}\NormalTok{(}\OperatorTok{-}\NormalTok{n)}

\KeywordTok{ggplot}\NormalTok{(}\KeywordTok{filter}\NormalTok{(sum_ssp_vg, species_code }\OperatorTok{!=}\StringTok{ "rbp"} \OperatorTok{&}\StringTok{ }\NormalTok{phase }\OperatorTok{==}\StringTok{ "i"}\NormalTok{), }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x=}\KeywordTok{factor}\NormalTok{(site, }\DataTypeTok{levels =}\NormalTok{ sum_site}\OperatorTok{$}\NormalTok{site), }\DataTypeTok{y=}\NormalTok{vg_mean)) }\OperatorTok{+}
\StringTok{    }\KeywordTok{geom_bar}\NormalTok{(}\DataTypeTok{stat =} \StringTok{"identity"}\NormalTok{, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{fill =}\NormalTok{ island), }\DataTypeTok{position =} \StringTok{"dodge"}\NormalTok{) }\OperatorTok{+}
\StringTok{    }\KeywordTok{geom_errorbar}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{ymin=}\NormalTok{vg_mean}\OperatorTok{-}\NormalTok{vg_se, }\DataTypeTok{ymax=}\NormalTok{vg_mean}\OperatorTok{+}\NormalTok{vg_se), }\DataTypeTok{width=}\NormalTok{.}\DecValTok{2}\NormalTok{,}
                 \DataTypeTok{position=}\KeywordTok{position_dodge}\NormalTok{(.}\DecValTok{9}\NormalTok{)) }\OperatorTok{+}
\StringTok{    }\KeywordTok{facet_grid}\NormalTok{(}\OperatorTok{~}\NormalTok{species) }\OperatorTok{+}
\StringTok{    }\KeywordTok{scale_fill_brewer}\NormalTok{(}\DataTypeTok{palette =} \StringTok{"Blues"}\NormalTok{) }\OperatorTok{+}
\StringTok{    }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{title=}\StringTok{"Mean grazing impact by volume (initial phase)"}\NormalTok{, }
         \DataTypeTok{y=}\StringTok{"Grazing impact (cm3/hr)"}\NormalTok{, }
         \DataTypeTok{x=}\StringTok{"Site"}\NormalTok{) }\OperatorTok{+}
\StringTok{    }\KeywordTok{theme_bw}\NormalTok{() }\OperatorTok{+}
\StringTok{    }\KeywordTok{theme}\NormalTok{(}\DataTypeTok{strip.text =} \KeywordTok{element_text}\NormalTok{(}\DataTypeTok{face =} \StringTok{"italic"}\NormalTok{), }
          \DataTypeTok{axis.text.x =} \KeywordTok{element_text}\NormalTok{(}\DataTypeTok{angle =} \DecValTok{90}\NormalTok{, }\DataTypeTok{hjust =} \DecValTok{1}\NormalTok{), }
          \DataTypeTok{plot.title =} \KeywordTok{element_text}\NormalTok{(}\DataTypeTok{hjust =} \FloatTok{0.5}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\includegraphics{Ch2_outline_V2_files/figure-latex/unnamed-chunk-10-1.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{ggplot}\NormalTok{(}\KeywordTok{filter}\NormalTok{(sum_ssp_vg, species_code }\OperatorTok{!=}\StringTok{ "rbp"} \OperatorTok{&}\StringTok{ }\NormalTok{phase }\OperatorTok{==}\StringTok{ "i"}\NormalTok{), }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x=}\KeywordTok{factor}\NormalTok{(site, }\DataTypeTok{levels =}\NormalTok{ sum_site}\OperatorTok{$}\NormalTok{site), }\DataTypeTok{y=}\NormalTok{vg_capped_mean)) }\OperatorTok{+}
\StringTok{    }\KeywordTok{geom_bar}\NormalTok{(}\DataTypeTok{stat =} \StringTok{"identity"}\NormalTok{, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{fill =}\NormalTok{ island), }\DataTypeTok{position =} \StringTok{"dodge"}\NormalTok{) }\OperatorTok{+}
\StringTok{    }\KeywordTok{geom_errorbar}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{ymin=}\NormalTok{vg_capped_mean}\OperatorTok{-}\NormalTok{vg_capped_se, }\DataTypeTok{ymax=}\NormalTok{vg_capped_mean}\OperatorTok{+}\NormalTok{vg_capped_se), }\DataTypeTok{width=}\NormalTok{.}\DecValTok{2}\NormalTok{,}
                 \DataTypeTok{position=}\KeywordTok{position_dodge}\NormalTok{(.}\DecValTok{9}\NormalTok{)) }\OperatorTok{+}
\StringTok{    }\KeywordTok{facet_grid}\NormalTok{(}\OperatorTok{~}\NormalTok{species) }\OperatorTok{+}
\StringTok{    }\KeywordTok{scale_fill_brewer}\NormalTok{(}\DataTypeTok{palette =} \StringTok{"Blues"}\NormalTok{) }\OperatorTok{+}
\StringTok{    }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{title=}\StringTok{"Mean grazing impact by volume (capped height, initial phase)"}\NormalTok{, }
         \DataTypeTok{y=}\StringTok{"Grazing impact (cm3/hr)"}\NormalTok{, }
         \DataTypeTok{x=}\StringTok{"Site"}\NormalTok{) }\OperatorTok{+}
\StringTok{    }\KeywordTok{theme_bw}\NormalTok{() }\OperatorTok{+}
\StringTok{    }\KeywordTok{theme}\NormalTok{(}\DataTypeTok{strip.text =} \KeywordTok{element_text}\NormalTok{(}\DataTypeTok{face =} \StringTok{"italic"}\NormalTok{), }
          \DataTypeTok{axis.text.x =} \KeywordTok{element_text}\NormalTok{(}\DataTypeTok{angle =} \DecValTok{90}\NormalTok{, }\DataTypeTok{hjust =} \DecValTok{1}\NormalTok{), }
          \DataTypeTok{plot.title =} \KeywordTok{element_text}\NormalTok{(}\DataTypeTok{hjust =} \FloatTok{0.5}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\includegraphics{Ch2_outline_V2_files/figure-latex/unnamed-chunk-10-2.pdf}

\section{Discussion}\label{discussion}

\begin{itemize}
\tightlist
\item
  Significant differences in feeding behaviors within species across
  islands (be conservative about conclusions here, just presenting
  observed trends despite relatively small sample sizes)
\item
  GAMM results: explain meaning/significance of PC1 (essentially higher
  feeding rates in reefs with higher scarid biomass and rugosity
  vs.~more algal dominated and lower scarid biomass reefs); \emph{S.
  vetula} seem more impacted by PC1 than \emph{S. viride}.
\item
  Potential mechanisms/hypotheses for these observed trends:

  \begin{itemize}
  \tightlist
  \item
    Social feeding hypothesis: more time must be spent vigilant in
    smaller groups, potential compounding effect with overfishing/loss
    of herbivory
  \item
    Nutrient limitation hypothesis: scarids must graze more (and more
    intensely) in sites with lower algal canopy heights and percent
    cover
  \item
    Predation risk hypothesis: presence of large piscivores does not
    explain feeding trends, but risk from spearfishers may play a bigger
    role (able to target larger fish, i.e.~those sampled here).
    (Incorporate rugosity here - Rogers, Madin, Catano)
  \end{itemize}
\item
  Ecological implications: potential reinforcing feedback loops in
  degraded reefs if high algal/low scarid reefs also have less grazing
  impact per scarid

  \begin{itemize}
  \tightlist
  \item
    Feeding intensity: less intense feeding/more distributed bites in
    degraded reefs may also reduce actual substrate made available to
    recruiting corals (\textasciitilde{} grazing concentration lit.
    here)
  \end{itemize}
\item
  Caveats + limitations: all these potential drivers/mechanisms are
  understandably correlated in observational work

  \begin{itemize}
  \tightlist
  \item
    Room for experimental work to unpack/distinguish drivers
  \end{itemize}
\item
  Need for incorporation of behavior in coral reef grazing models,
  especially those used to set targets for minimum herbivore biomass
  levels, because these may look very different in different reef
  conditions. Current paradigm of constant species-level grazing
  behaviors may be inaccurate. In Caribbean case, most models use
  Bonaire data which may not be representative of dynamics on more
  degraded Caribbean reefs.
\item
  Need for framework to help incorporate behavioral effects into
  ecosystem management
\item
  Conservation + mgmt implications: multiple human impacts on reefs may
  affect herbivore behavior in different ways, spec. fishing + habitat
  degradation - which may have reinforcing/compounding negative effects
  on reef health
\end{itemize}

\end{document}

### 2019-02-01 r/Data_prodding.Rmd
feb1 updates

@@
     theme_minimal()
+ggsave(here("figs/pca_pred.png"))
 ```
@@
 ```{r}
-ggplot(filter(sum_id, species_code != "rbp" & island != "Barbuda" & length_cm>=16 & length_cm<=24), aes(x = scar_bm,y=fr)) +
+ggplot(filter(sum_id, species_code != "rbp" & length_cm>=15 & length_cm<=30), aes(x = scar_bm,y=fr)) +
   facet_grid(.~species) +
@@
 ```{r,echo=FALSE,warning=FALSE}
-ggplot(filter(sum_id, species_code != "rbp" & island != "Barbuda" & length_cm>=16 & length_cm<=24), aes(x = scar_bm, y = fr, colour = island)) +
+ggplot(filter(sum_id, species_code != "rbp" & length_cm>=16 & length_cm<=24), aes(x = scar_bm, y = fr, colour = island)) +
   facet_grid(. ~ species) +
@@
 ```
+
 ```{r,echo=FALSE,warning=FALSE}
@@
 ```
-
-## GAMM
-- for GAMMs - seems model results are the same whether random = site or island?
-
-```{r,eval=F}
-gamm<-gamm(fr~phase+s(length_cm)+s(pc1_bent)+s(pc2_bent)+s(scar_bm)+s(scar_den), 
-           family=gaussian(link=identity),
-           data=filter(sum_id_pca, species_code=="qup"),
-           random=list(site=~1),
-           method="REML"
-           )
-summary(gamm$gam)
-plot(gamm$gam,pages=1)
-```
-```{r,eval=F}
-gamm1<-gamm(fr_mean~phase+species+s(pc1_bent)+s(pc2_bent)+s(scar_bm)+s(scar_den), 
-           family=gaussian(link=identity),
-           data=filter(sum_ssp_pca, species_code!="rbp" & island!="Barbuda"),
-           random=list(site=~1),
-           method="REML"
-           )
-summary(gamm1$gam)
-AIC(gamm1$lme)
-plot(gamm1$gam,pages=1)
-
-```
-
-------------------------
-To Do as of Nov. 7  
-* boosted regression trees ecosphere 2017 adrians paper  
-* species as random effect

### 2019-02-01 r/Follows.Rmd
feb1 updates

@@
   left_join(sp_site, by=c("species","site")) # add conspecific scarid info (based on site/species)
+sum_site_sp <- sum_site_sp %>% 
+  left_join(site, by=c("island","site")) %>%
+  left_join(sp_site, by=c("species","site")) # add conspecific scarid info (based on site/species)
 sum_site_sp_ph <- sum_site_sp_ph %>% 
@@
 write.csv(sum_id,file=here("data/sum_id.csv"))
+write.csv(sum_site_sp,file=here("data/sum_site_sp.csv"))
 write.csv(sum_site_sp_ph,file=here("data/sum_site_sp_ph.csv"))

### 2019-02-05 r/Ch2 outline_V3.Rmd
updated outline

---
title: "Ch2 outline_V3"
author: "Molly Wilson"
date: "2/5/2019"
output: html_document
---

```{r, echo=F, message=F, warning=F}
library(startR)
library(here)
library(tidyverse)

sum_site <- read.csv(here("data","sum_site.csv"), stringsAsFactors=F) %>%
  select(-X) %>% 
  mutate(island = factor(island, levels = c("Bonaire","Antigua","Barbuda"))) %>%
  arrange(island)
sum_id <- read.csv(here("data","sum_id.csv"), stringsAsFactors=F) %>%
  select(-X) %>% 
  mutate(island = factor(island, levels = c("Bonaire","Antigua","Barbuda"))) %>%
  arrange(island) %>%
  filter(fr <= 3000)
sum_ssp <- read.csv(here("data","sum_site_sp_ph.csv"), stringsAsFactors=F) %>%
  select(-X) %>% 
  mutate(island = factor(island, levels = c("Bonaire","Antigua","Barbuda"))) %>%
  arrange(island)
```

# Intro + motivation

* Grazing by herbivorous fish is an established driver of coral reef health/benthic dynamics
* As reefs become increasingly degraded/threatened, there is growing interest in understanding how much herbivory is needed to sustain coral reef health (e.g. herbivore threshold papers - Karr et al. 2015, McClanahan et al. 2011)
* However, not all herbivores are the same in terms of their grazing impact…
    * Recent developments in functional classifications, difference across species in terms of feeding selectivity, morphology, etc. (e.g. Adam et al. 2018)
* In addition to these inter-species differences in functional role, differences in herbivore behavior may also affect grazing impact and may drive variation in herbivore function among individuals or across reef systems
* In understanding and predicting herbivore impact on reefs, current assumption is that grazing behaviors are constant within species (same species-level length~feeding rate relationships)
* But there is very limited data on how herbivore feeding behaviors might vary across reef systems and whether we can accurately assume constant species roles at regional scales
* Here we document variability in multiple components (*if I include foraging intensity as well as feeding rate...*) of scarid feeding behavior across reef sites in three Caribbean islands and discuss potential underlying mechanisms, ecological consequences, and management implications
* We demonstrate a need for further experimental and modelling work to unpack these drivers and understand their ecological consequences
* We urge for the incorporation of behavioral effects in coral reef herbivore management and caution against assuming constant species function across reef systems

# Methods

## Data collection
* Sites (13 sites, 3 islands [map]), timeline of data collection
* Fish and benthic community surveys
* Observational behavioral data:
    - Species: *Sparisoma viride* and *Scarus vetula*, initial and terminal phase greater than 15cm, focus on only initial phase here
    - 2 minute follows, calibration period, outside of previously established FID
    - Recording time budgets as well as bite rates during grazing forays (define)
    - Discarded data of incomplete follows or where fish were visably disturbed/fled
    - Timing of data collection distributed throughout day, 2 dives per site
    - Define behavioral variables ([table])

## Analysis
* Fish biomass using lenth-weight conversions (appendix)
* Site-level behavioral summary stats taken from subset of individuals based on size to avoid bias
* Biplots w predictors - loess smoothers (switch to something else?)
* Bite content: calculated proxy for bite content by multiplying estimated bite sizes (based on species and length of each individual fish surveyed, equations from Bruggemann/Mumby) by site level averages of turf algal canopy height to account for potential differences in amount of algae consumed in low vs. high canopy sites

# Results and discussion

## Variation among sites in multiple components of feeding behavior
Notes:

* Can remove color coding by island if this is distracting. Unsure whether I should be highlighting differences between islands or if I should just be highlighting variation among sites (mixed feedback on this)
* Eventually I'll have a panel that shows differences in feeding rates, time spent feeding, and grazing intensity (trends are similar, but slightly different implications)  
* (Will remove Rendezvous from S. vetula graph in final - no S. vetulas found at this site)

```{r, echo=F, message=F, warning=F}
ggplot(filter(sum_ssp, species_code != "rbp" & phase == "i"), aes(x=factor(site, levels = sum_site$site), y=fr_mean)) +
    geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
    geom_errorbar(aes(ymin=fr_mean-fr_se, ymax=fr_mean+fr_se), width=.2,
                 position=position_dodge(.9)) +
    facet_grid(~species) +
    scale_fill_brewer(palette = "Blues") +
    labs(title="Mean feeding rates of individuals (initial phase)", 
         y="Feeding rate (bites/hr)", 
         x="Site") +
    theme_bw() +
    theme(strip.text = element_text(face = "italic"), 
          axis.text.x = element_text(angle = 90, hjust = 1), 
          plot.title = element_text(hjust = 0.5))
```

Notes:

* Can remove color coding by island if this is distracting. Unsure whether I should be highlighting differences between islands or if I should just be highlighting variation among sites (mixed feedback on this)
* Eventually I'll have a panel that shows differences in feeding rates, time spent feeding, and grazing intensity (trends are similar, but slightly different implications)  
* (Will remove Rendezvous from S. vetula graph in final - no S. vetulas found at this site)

## Several potential underlying mechanisms

While we do not seek to conclusively establish mechanisms here, we discuss potential underlying mechanisms that could be driving these trends in herbivore feeding behaviors

* Bite content/nutrition hypothesis: scarids in higher canopy sites are consuming more per bite, thus taking fewer bites - implications for spatial grazing impact. Previous work showing that scarids vary grazing with nutrient availability/distribution (Carlson, Tootell, Mumby, etc.)

```{r, echo=F, message=F, warning=F}
# grazing impact (vg ~ volume grazed) of fish from behavioral surveys
sum_id_vg <- sum_id %>%
  filter(species_code != "rbp") %>%
  mutate(
    bs = ifelse(species_code == "stop", 5.257*10^-4*length_cm^2, 4.013*10^-4*length_cm^2), # adding bite size equations from Mumby
    bh_capped = ifelse(ta_canopy < 2*sqrt(bs)/(22/7), ta_canopy, 2*sqrt(bs)/(22/7)), # if ta_canopy exceeds diameter of bite size, bite height is capped at bite size
    bv = bs*ta_canopy,
    bv_capped = bs*bh_capped,
    vg = fr*bv,
    vg_capped = fr*bv_capped,
    island = factor(island, levels = c("Bonaire","Antigua","Barbuda"))
) %>%
  arrange(island)

# summing by site, species, and phase
sum_ssp_vg <- sum_id_vg %>% 
  group_by(site, island, species, species_code, phase) %>% 
  summarize(
    n=n(),
    vg_mean=mean(vg),
    vg_se=sd(vg)/sqrt(n),
    vg_capped_mean=mean(vg_capped),
    vg_capped_se=sd(vg_capped)/sqrt(n)
    ) %>% 
  select(-n)

ggplot(filter(sum_ssp_vg, species_code != "rbp" & phase == "i"), aes(x=factor(site, levels = sum_site$site), y=vg_mean)) +
    geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
    geom_errorbar(aes(ymin=vg_mean-vg_se, ymax=vg_mean+vg_se), width=.2,
                 position=position_dodge(.9)) +
    facet_grid(~species) +
    scale_fill_brewer(palette = "Blues") +
    labs(title="Mean grazing impact by volume (initial phase)", 
         y="Grazing impact (cm3/hr)", 
         x="Site") +
    theme_bw() +
    theme(strip.text = element_text(face = "italic"), 
          axis.text.x = element_text(angle = 90, hjust = 1), 
          plot.title = element_text(hjust = 0.5))
```

* Social feeding hypothesis: more time must be spent vigilant in smaller groups, potential compounding effect with overfishing/loss of herbivory (e.g. Gil)
* Predation risk hypothesis: presence of large piscivores does not explain feeding trends, but risk from spearfishers may play a bigger role (able to target larger fish, i.e. those sampled here). (Incorporate rugosity here - Rogers, Madin, Catano). Did not find strong evidence for this here, but carnivore biomass was strongly correllated with herbivore biomass and rugosity.

## Ecological implications

* Potential reinforcing feedback loops in degraded reefs if high algal/low scarid reefs also have less grazing impact per scarid
    * Feeding intensity: less intense feeding/more distributed bites in degraded reefs may also reduce actual substrate made available to recruiting corals (~ grazing concentration lit. here)

## Management implications + future research needs

* Limitations to this study: all these potential drivers/mechanisms are understandably correlated in observational work
    * Room for experimental work to unpack/distinguish drivers
* Need for incorporation of behavior in coral reef grazing models, especially those used to set targets for minimum herbivore biomass levels, because these may look very different in different reef conditions. Current paradigm of constant species-level grazing behaviors may be inaccurate. In Caribbean case, most models use Bonaire data which may not be representative of dynamics on more degraded Caribbean reefs.
* Need for framework to help incorporate behavioral effects into ecosystem management
* Conservation + mgmt implications: multiple human impacts on reefs may affect herbivore behavior in different ways, spec. fishing + habitat degradation - which may have reinforcing/compounding negative effects on reef health

### 2019-03-01 r/Ch2 outline_V3.Rmd
updates

@@
 ---
-title: "Ch2 outline_V3"
+title: "Variance in scarid grazing behavior across Caribbean reef sites"
 author: "Molly Wilson"
-date: "2/5/2019"
+date: "3/1/2019"
 output: html_document
@@
 
+# Variance in scarid grazing behavior across Caribbean reef sites
+### Formatted as: Coral Reefs note (<2900 words)
+
 ```{r, echo=F, message=F, warning=F}
-library(startR)
 library(here)
 library(tidyverse)
+library(ggplot2)
+library(gridExtra)
+library(grid)
+library(ggfortify)
+library(mgcv)
+library(MuMIn)
+library(knitr)
 
@@
 
-# Intro + motivation
+## Introduction
 
-* Grazing by herbivorous fish is an established driver of coral reef health/benthic dynamics
-* As reefs become increasingly degraded/threatened, there is growing interest in understanding how much herbivory is needed to sustain coral reef health (e.g.
-herbivore threshold papers - Karr et al.
+Herbivorous fish play a critical role in coral reef ecosystems by supressing algae and facilitating coral growth, recruitment, and survival ().
+In recent decades, coral reefs have become increasingly threatened by anthropogenic impacts including the fishing down of herbivorous fish populations ().
+Increasing attention has been given to understanding how much herbivory is needed to sustain coral reef health and to identify thresholds of herbivore biomass that can be used to guide fisheries management targets (Karr et al.
 2015, McClanahan et al.
 2011)
-* However, not all herbivores are the same in terms of their grazing impact…
-    * Recent developments in functional classifications, difference across species in terms of feeding selectivity, morphology, etc. (e.g. Adam et al. 2018)
-* In addition to these inter-species differences in functional role, differences in herbivore behavior may also affect grazing impact and may drive variation in herbivore function among individuals or across reef systems
-* In understanding and predicting herbivore impact on reefs, current assumption is that grazing behaviors are constant within species (same species-level length~feeding rate relationships)
-* But there is very limited data on how herbivore feeding behaviors might vary across reef systems and whether we can accurately assume constant species roles at regional scales
-* Here we document variability in multiple components (*if I include foraging intensity as well as feeding rate...*) of scarid feeding behavior across reef sites in three Caribbean islands and discuss potential underlying mechanisms, ecological consequences, and management implications
-* We demonstrate a need for further experimental and modelling work to unpack these drivers and understand their ecological consequences
-* We urge for the incorporation of behavioral effects in coral reef herbivore management and caution against assuming constant species function across reef systems
+
+However, not all herbivores are the same in terms of their grazing impact on reefs.
+Recent studies have highlighted functional differences across herbivorous fish species in terms of their feeding selectivity, feeding morphology, etc.
+(e.g.
+Adam et al.
+2018) [**add more here**]
+
+In addition to these functional differences found across species, herbivore behavior may also drive variation in herbivore function across different reef environments.
+Empirical studies have demonstrated the sensitivity of herbivore behavior to predator presence (), reef rugosity (), grazing behaviors of proximate herbivores (Gil et al.
+2018), and algal abundance, distribution, and nutritional content ().
+[**add more here**]
+[These behavioral variations have not been linked to larger-scale variations - except Madin's seascape papers...]
+[Maybe even highlight Madin papers - these behavioral differences have the potential to drive 
+Madin et al.'s study showing how predator abundance drove herbivore feeding ranges/grazing halos]
+
+Despite a growing understanding of the sensitivity and ecological importance of herbivore behavior, current efforts to predict herbivore impact on reefs assume species-specific feeding rates that vary only with individual fish size.
+There is very limited data on how herbivore feeding behaviors might vary across reef systems and whether we can accurately assume constant species roles at regional scales.
+Here we document variability in multiple components of scarid feeding behavior across reef sites in three Caribbean islands and discuss potential underlying mechanisms, ecological consequences, and management implications.
+We demonstrate a need for further experimental and modelling work to unpack these drivers and determine their ecological consequences.
+We urge for the incorporation of behavioral effects in coral reef herbivore management and caution against assuming constant species function across reef systems.
 
@@
 
-## Data collection
-* Sites (13 sites, 3 islands [map]), timeline of data collection
-* Fish and benthic community surveys
+*Data collection*
+In March through August of 2017, I collected data from 13 reef sites in Bonaire, Antigua, and Barbuda (Fig.
+1).
+At each site I conducted fish, benthic, and rugosity surveys to quantify reef (condition).
+I used a modified AGRRA protocol with 30m by 4m belt transects and 10m point-count transects for fish and benthic survets, respectively.
+To assess rugosity I measured the length of reef contour under each meter of a taut 10m leadline (sensu ___).
+I carried out a minimum of X fish transects and X benthic and rugosity transects per site.
+
+At each site, I observed grazing behavior of two dominant scarid species, *Sparisoma viride* and *Scarus vetula*.
+I targetted these species because of their relative abundance as well as their contrasting grazing mechanisms.
+*S.
+viride* is classified as an excavator, taking [_] (), while *S.
+vetula* is a grazer, taking shallower bites from [_] ().
+I quantified grazing behavior during two minute follows,
 * Observational behavioral data:
-    - Species: *Sparisoma viride* and *Scarus vetula*, initial and terminal phase greater than 15cm, focus on only initial phase here
+    - Species: , initial and terminal phase greater than 15cm, focus on only initial phase here
     - 2 minute follows, calibration period, outside of previously established FID
@@
     - Timing of data collection distributed throughout day, 2 dives per site
-    - Define behavioral variables ([table])
+    - Define behavioral variables ([table]): feeding rate, bite rate, feeding intensity
 
-## Analysis
-* Fish biomass using lenth-weight conversions (appendix)
+```{r}
+Variable <- c('Time spent grazing','Bite rate','Feeding rate','Feeding intensity')
+Definition <- c('Proportion of time spent grazing during follow','Bite rate while actively feeding','Bite rate over total follow time','Number of consecutive bites taken during single grazing foray')
+Unit <- c('fraction','bites/second','bites/minute','bites/foray')
+var.table <- data.frame(Variable,Definition,Unit)
+kable(var.table)
+```
+
+*Analysis*
+
+* Fish biomass using length-weight conversions (appendix)
 * Site-level behavioral summary stats taken from subset of individuals based on size to avoid bias
@@
 
-# Results and discussion
+## Results and discussion
 
 ## Variation among sites in multiple components of feeding behavior
-Notes:
+Fig.
+X shows intraspecific variation in both feeding rate (bites per hour) and feeding intensity (bites per foray) among sites.
+(Bite rates while grazing were relatively constant, indicating that differences in overall feeding rates were driven by differences in the amount of time spent grazing).
 
-* Can remove color coding by island if this is distracting.
-Unsure whether I should be highlighting differences between islands or if I should just be highlighting variation among sites (mixed feedback on this)
-* Eventually I'll have a panel that shows differences in feeding rates, time spent feeding, and grazing intensity (trends are similar, but slightly different implications)  
-* (Will remove Rendezvous from S. vetula graph in final - no S. vetulas found at this site)
-
-```{r, echo=F, message=F, warning=F}
+```{r, eval = F, echo=F, message=F, warning=F}
 ggplot(filter(sum_ssp, species_code != "rbp" & phase == "i"), aes(x=factor(site, levels = sum_site$site), y=fr_mean)) +
@@
 Unsure whether I should be highlighting differences between islands or if I should just be highlighting variation among sites (mixed feedback on this)
-* Eventually I'll have a panel that shows differences in feeding rates, time spent feeding, and grazing intensity (trends are similar, but slightly different implications)  
-* (Will remove Rendezvous from S. vetula graph in final - no S. vetulas found at this site)
+
+```{r, echo = F, message = F, warning = F}
+# Panel graph of site differences
+
+fr_qup <- 
+  ggplot(filter(sum_ssp, species_code == "qup" & phase == "i"), 
+         aes(x=factor(site, levels = sum_site$site), y=fr_mean)) +
+  geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
+  geom_errorbar(aes(ymin=fr_mean-fr_se, ymax=fr_mean+fr_se), width=.2,
+                 position=position_dodge(.9)) +
+  scale_fill_brewer(palette = "Blues") +
+  labs(y = "Feeding rate \n (bites/min)", 
+       title = "S. vetula") +
+  theme_bw() + 
+  theme(plot.title = element_text(hjust = 0.5, face = "italic"),
+        plot.margin = margin(5, 0, 0, 5, unit = "pt"),
+        strip.text = element_text(face = "italic"), 
+        axis.text.x=element_blank(),
+        axis.title.x = element_blank(),
+        axis.title.y = element_text(size=10),
+        legend.position = "none")
+
+fr_stop <- 
+  ggplot(filter(sum_ssp, species_code == "stop" & phase == "i"), 
+         aes(x=factor(site, levels = sum_site$site), y=fr_mean)) +
+  geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
+  geom_errorbar(aes(ymin=fr_mean-fr_se, ymax=fr_mean+fr_se), width=.2,
+                 position=position_dodge(.9)) +
+  scale_fill_brewer(palette = "Blues") +
+  labs(y = "",
+       # y = "S. vetula \nFeeding rate (bites/hr)", 
+       title = "S. viride") +
+  theme_bw() + 
+  theme(plot.title = element_text(hjust = 0.5, face = "italic"),
+        plot.margin = margin(5, 4, 0, 0, unit = "pt"),
+        strip.text = element_text(face = "italic"), 
+        axis.text.x=element_blank(),
+        axis.title.x = element_blank(),
+        legend.position = "none")
+
+br_qup <- 
+  ggplot(filter(sum_ssp, species_code == "qup" & phase == "i"), 
+         aes(x=factor(site, levels = sum_site$site), y=br_mean)) +
+  geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
+  geom_errorbar(aes(ymin=br_mean-br_se, ymax=br_mean+br_se), width=.2,
+                 position=position_dodge(.9)) +
+  scale_fill_brewer(palette = "Blues") +
+  labs(y = "Bite rate \n(bites/sec)", 
+       x = "", 
+       title = "") +
+  theme_bw() + 
+  theme(plot.title = element_text(hjust = 0.5),
+        plot.margin = margin(0, 0, 0, 5, unit = "pt"),
+        strip.text = element_text(face = "italic"), 
+        axis.text.x=element_blank(),
+        axis.title.x = element_blank(),
+        axis.title.y = element_text(size=10),
+        legend.position = "none")
+
+br_stop <- 
+  ggplot(filter(sum_ssp, species_code == "stop" & phase == "i"), 
+         aes(x=factor(site, levels = sum_site$site), y=br_mean)) +
+  geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
+  geom_errorbar(aes(ymin=br_mean-br_se, ymax=br_mean+br_se), width=.2,
+                 position=position_dodge(.9)) +
+  scale_fill_brewer(palette = "Blues") +
+  labs(y = "",
+       # y = "S. vetula \nFeeding rate (bites/hr)", 
+       x = "", 
+       title = "") +
+  theme_bw() + 
+  theme(plot.title = element_text(hjust = 0.5),
+        plot.margin = margin(0, 4, 0, 0, unit = "pt"),
+        strip.text = element_text(face = "italic"), 
+        axis.text.x=element_blank(),
+        axis.title.x = element_blank(),
+        legend.position = "none")
+
+fb_qup <- 
+  ggplot(filter(sum_ssp, species_code == "qup" & phase == "i"), 
+         aes(x=factor(site, levels = sum_site$site), y=for_bites_mean)) +
+  geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
+  geom_errorbar(aes(ymin=for_bites_mean-for_bites_se, ymax=for_bites_mean+for_bites_se), width=.2,
+                 position=position_dodge(.9)) +
+  scale_fill_brewer(palette = "Blues") +
+  labs(y = "Feeding intensity\n(bites/foray)", 
+       x = "", 
+       title = "") +
+  theme_bw() + 
+  theme(plot.title = element_text(hjust = 0.5),
+        plot.margin = margin(0, 0, 0, 5, unit = "pt"),
+        strip.text = element_text(face = "italic"), 
+        axis.text.x=element_blank(),
+        axis.title.x = element_blank(),
+        axis.title.y = element_text(size=10),
+        legend.position = "none")
+
+fb_stop <- 
+  ggplot(filter(sum_ssp, species_code == "stop" & phase == "i"), 
+         aes(x=factor(site, levels = sum_site$site), y=for_bites_mean)) +
+  geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
+  geom_errorbar(aes(ymin=for_bites_mean-for_bites_se, ymax=for_bites_mean+for_bites_se), width=.2,
+                 position=position_dodge(.9)) +
+  scale_fill_brewer(palette = "Blues") +
+  labs(y = "",
+       x = "", 
+       title = "") +
+  theme_bw() + 
+  theme(plot.title = element_text(hjust = 0.5),
+        plot.margin = margin(0, 4, 0, 0, unit = "pt"),
+        strip.text = element_text(face = "italic"), 
+        axis.text.x=element_blank(),
+        axis.title.x = element_blank(),
+        legend.position = "none")
+
+gfrac_qup <- 
+  ggplot(filter(sum_ssp, species_code == "qup" & phase == "i"), 
+         aes(x=factor(site, levels = sum_site$site), y=g_frac_mean)) +
+  geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
+  geom_errorbar(aes(ymin=g_frac_mean-g_frac_se, ymax=g_frac_mean+g_frac_se), width=.2,
+                 position=position_dodge(.9)) +
+  scale_fill_brewer(palette = "Blues") +
+  labs(y = "Time grazing \n(fraction)", 
+       x = "", 
+       title = "") +
+  theme_bw() + 
+  theme(plot.margin = margin(0, 0, 0, 5, unit = "pt"),
+        axis.title.y = element_text(size=10),
+        axis.text.x = element_text(angle = 45, hjust = 1, size=7),
+        legend.position = "none")
+
+gfrac_stop <- 
+  ggplot(filter(sum_ssp, species_code == "stop" & phase == "i"), 
+         aes(x=factor(site, levels = sum_site$site), y=g_frac_mean)) +
+  geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
+  geom_errorbar(aes(ymin=g_frac_mean-g_frac_se, ymax=g_frac_mean+g_frac_se), width=.2,
+                 position=position_dodge(.9)) +
+  scale_fill_brewer(palette = "Blues") +
+  labs(y = "",
+       x = "", 
+       title = "") +
+  theme_bw() + 
+  theme(plot.margin = margin(0, 4, 0, 0, unit = "pt"),
+        axis.text.x = element_text(angle = 45, hjust = 1, size=7),
+        legend.position = "none")
+
+legend_plot <- 
+  ggplot(filter(sum_ssp, species_code == "stop" & phase == "i"), 
+         aes(x=factor(site, levels = sum_site$site), y=g_frac_mean)) +
+  geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
+  geom_errorbar(aes(ymin=g_frac_mean-g_frac_se, ymax=g_frac_mean+g_frac_se), width=.2,
+                 position=position_dodge(.9)) +
+  scale_fill_brewer(palette = "Blues") +
+  labs(y = "",
+       x = "", 
+       title = "") +
+  theme_bw() +
+  theme(legend.position = "right",
+        legend.title = element_blank())
+
+get_legend<-function(myggplot){
+  tmp <- ggplot_gtable(ggplot_build(myggplot))
+  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
+  legend <- tmp$grobs[[leg]]
+  return(legend)
+  }
+legend <- get_legend(legend_plot)
+
+grid.arrange(
+  arrangeGrob(fr_qup, fr_stop, fb_qup, fb_stop, gfrac_qup, gfrac_stop, ncol=2, heights = c(1,1,1.5)), 
+              arrangeGrob(legend), 
+              ncol=2, widths = c(5,1))
+
+```
 
 ## Several potential underlying mechanisms
+
+```{r, echo=F, message=F, warning=F}
+fr_q1 <- ggplot(filter(sum_ssp, species_code == "qup" & phase == "i"), aes(x = scar_bm/1000, y = fr_mean)) +
+  geom_smooth(method="loess", span=1) +
+  geom_point(shape=18) +
+  ylim(-10, 60) +
+  labs(y = "",
+       title = "") +
+  theme_bw() + 
+  theme(plot.title = element_text(hjust = 0.5),
+        plot.margin = margin(0, 1, 0, 0, unit = "pt"),
+        axis.text.x = element_blank(),
+        axis.title.x = element_blank()
+        )
+
+fr_q2 <- ggplot(filter(sum_ssp, species_code == "qup" & phase == "i"), aes(x = consp_den, y = fr_mean)) +
+  geom_smooth(method="loess", span=1) +
+  geom_point(shape=18) +
+  ylim(-10, 60) +
+  labs(y = "", 
+       title = "") +
+  theme_bw() + 
+  theme(plot.title = element_text(hjust = 0.5),
+        axis.ticks.y = element_blank(),
+        axis.text.y = element_blank(),
+        plot.margin = margin(0, 1, 0, 1, unit = "pt"),
+        axis.text.x = element_blank(),
+        axis.title.x = element_blank()
+        )
+
+fr_q3 <- ggplot(filter(sum_ssp, species_code == "qup" & phase == "i"), aes(x = rugosity, y = fr_mean)) +
+  geom_smooth(method="loess", span=1) +
+  geom_point(shape=18) +
+  ylim(-10, 60) +
+  labs(y = "", 
+       title = "") +
+  theme_bw() + 
+  theme(plot.title = element_text(hjust = 0.5),
+        axis.ticks.y = element_blank(),
+        axis.text.y = element_blank(),
+        plot.margin = margin(0, 1, 0, 1, unit = "pt"),
+        axis.text.x = element_blank(),
+        axis.title.x = element_blank()
+        )
+
+fr_q4 <- ggplot(filter(sum_ssp, species_code == "qup" & phase == "i"), aes(x = ta_canopy, y = fr_mean)) +
+  geom_smooth(method="loess", span=1) +
+  geom_point(shape=18) +
+  ylim(-10, 60) +
+  labs(y = "", 
+       title = "") +
+  theme_bw() + 
+  theme(plot.title = element_text(hjust = 0.5),
+        axis.ticks.y = element_blank(),
+        axis.text.y = element_blank(),
+        plot.margin = margin(0, 1, 0, 1, unit = "pt"),
+        axis.text.x = element_blank(),
+        axis.title.x = element_blank()
+        )
+
+fr_q5 <- ggplot(filter(sum_ssp, species_code == "qup" & phase == "i"), aes(x = ta_cover, y = fr_mean)) +
+  geom_smooth(method="loess", span=1) +
+  geom_point(shape=18) +
+  ylim(-10, 60) +
+  labs(y = "",
+       title = "") +
+  theme_bw() + 
+  theme(plot.title = element_text(hjust = 0.5),
+        axis.ticks.y = element_blank(),
+        axis.text.y = element_blank(),
+        plot.margin = margin(0, 1, 0, 1, unit = "pt"),
+        axis.text.x = element_blank(),
+        axis.title.x = element_blank()
+        )
+
+fr_q7 <- ggplot(filter(sum_ssp, species_code == "qup" & phase == "i"), aes(x = ma_cover, y = fr_mean)) +
+  geom_smooth(method="loess", span=1) +
+  geom_point(shape=18) +
+  ylim(-10, 60) +
+  labs(y = "", 
+       title = "") +
+  theme_bw() + 
+  theme(plot.title = element_text(hjust = 0.5),
+        axis.ticks.y = element_blank(),
+        axis.text.y = element_blank(),
+        plot.margin = margin(0, 1, 0, 1, unit = "pt"),
+        axis.text.x = element_blank(),
+        axis.title.x = element_blank()
+        )
+
+fr_s1 <- ggplot(filter(sum_ssp, species_code == "stop" & phase == "i"), aes(x = scar_bm/1000, y = fr_mean)) +
+  geom_smooth(method="loess", span=1) +
+  geom_point(shape=18) +
+  ylim(-5, 20) +
+  labs(y = "",
+       x = expression(atop(NA, atop(textstyle('Scarid biomass'),
+                               textstyle('(kg/100m'^2*')')))),
+       title = "") +
+  theme_bw() + 
+  theme(plot.title = element_text(hjust = 0.5),
+        plot.margin = margin(0, 1, 0, 0, unit = "pt"),
+        axis.title.x = element_text(size = 9)
+        )
+
+fr_s2 <- ggplot(filter(sum_ssp, species_code == "stop" & phase == "i"), aes(x = consp_den, y = fr_mean)) +
+  geom_smooth(method="loess", span=1) +
+  geom_point(shape=18) +
+  ylim(-5, 20) +
+  labs(y = "", 
+       x = expression(atop(NA, atop(textstyle('Conspec. density'),
+                               textstyle('(indv/100m'^2*')')))),
+       title = "") +
+  theme_bw() + 
+  theme(plot.title = element_text(hjust = 0.5),
+        axis.ticks.y = element_blank(),
+        axis.text.y = element_blank(),
+        axis.title.x = element_text(size = 9),
+        plot.margin = margin(0, 1, 0, 1, unit = "pt")
+        )
+
+fr_s3 <- ggplot(filter(sum_ssp, species_code == "stop" & phase == "i"), aes(x = rugosity, y = fr_mean)) +
+  geom_smooth(method="loess", span=1) +
+  geom_point(shape=18) +
+  ylim(-5, 20) +
+  labs(y = "", 
+       x = expression(atop(NA, atop(textstyle('Rugosity'),
+                               textstyle('  ')))), 
+       title = "") +
+  theme_bw() + 
+  theme(plot.title = element_text(hjust = 0.5),
+        axis.ticks.y = element_blank(),
+        axis.text.y = element_blank(),
+        axis.title.x = element_text(size = 9),
+        plot.margin = margin(0, 1, 0, 1, unit = "pt")
+        )
+
+fr_s4 <- ggplot(filter(sum_ssp, species_code == "stop" & phase == "i"), aes(x = ta_canopy, y = fr_mean)) +
+  geom_smooth(method="loess", span=1) +
+  geom_point(shape=18) +
+  ylim(-5, 20) +
+  labs(y = "", 
+       x = expression(atop(NA, atop(textstyle('EAM canopy'),
+                               textstyle('height (mm)')))), 
+       title = "") +
+  theme_bw() + 
+  theme(plot.title = element_text(hjust = 0.5),
+        axis.ticks.y = element_blank(),
+        axis.text.y = element_blank(),
+        axis.title.x = element_text(size = 9),
+        plot.margin = margin(0, 1, 0, 1, unit = "pt")
+        )
+
+fr_s5 <- ggplot(filter(sum_ssp, species_code == "stop" & phase == "i"), aes(x = ta_cover, y = fr_mean)) +
+  geom_smooth(method="loess", span=1) +
+  geom_point(shape=18) +
+  ylim(-5, 20) +
+  labs(y = "", 
+       x = expression(atop(NA, atop(textstyle('EAM cover'),
+                               textstyle('(%)')))), 
+       title = "") +
+  theme_bw() + 
+  theme(plot.title = element_text(hjust = 0.5),
+        axis.ticks.y = element_blank(),
+        axis.text.y = element_blank(),
+        axis.title.x = element_text(size = 9),
+        plot.margin = margin(0, 1, 0, 1, unit = "pt")
+        )
+
+fr_s7 <- ggplot(filter(sum_ssp, species_code == "stop" & phase == "i"), aes(x = ma_cover, y = fr_mean)) +
+  geom_smooth(method="loess", span=1) +
+  geom_point(shape=18) +
+  ylim(-5, 20) +
+  labs(y = "", 
+       x = expression(atop(NA, atop(textstyle('Macroalgal cover'),
+                               textstyle('(%)')))), 
+       title = "") +
+  theme_bw() + 
+  theme(plot.title = element_text(hjust = 0.5),
+        axis.title.x = element_text(size = 9),
+        axis.ticks.y = element_blank(),
+        axis.text.y = element_blank(),
+        plot.margin = margin(0, 1, 0, 1, unit = "pt")
+        )
+
+label_vet = textGrob("S. vetula", gp=gpar(fontface="italic"), rot=270)
+label_vir = textGrob("S. viride", gp=gpar(fontface="italic"), rot=270)
+
+grid.arrange(left = "Feeding rate (bites/min.)",
+  arrangeGrob(fr_q1, fr_q3, fr_q4, fr_q5, fr_q7, ncol=5, right = label_vet),
+  arrangeGrob(fr_s1, fr_s3, fr_s4, fr_s5, fr_s7, ncol=5, right = label_vir), 
+  nrow=2, heights = c(1,1.2)) 
+```
 
@@
 sum_ssp_vg <- sum_id_vg %>% 
-  group_by(site, island, species, species_code, phase) %>% 
+  group_by(site, ta_canopy, island, species, species_code, phase) %>% 
   summarize(
@@
     vg_capped_mean=mean(vg_capped),
-    vg_capped_se=sd(vg_capped)/sqrt(n)
+    vg_capped_se=sd(vg_capped)/sqrt(n),
+    fr_mean=mean(fr)
     ) %>% 
@@
 
-ggplot(filter(sum_ssp_vg, species_code != "rbp" & phase == "i"), aes(x=factor(site, levels = sum_site$site), y=vg_mean)) +
+vg_facet <- ggplot(filter(sum_ssp_vg, species_code != "rbp" & phase == "i"), aes(x=factor(site, levels = sum_site$site), y=vg_mean)) +
     geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
@@
     theme(strip.text = element_text(face = "italic"), 
-          axis.text.x = element_text(angle = 90, hjust = 1), 
-          plot.title = element_text(hjust = 0.5))
+          axis.text.x = element_text(angle = 45, hjust = 1), 
+          legend.title = element_blank())
+
+vg_qup <- ggplot(filter(sum_ssp_vg, species_code == "qup" & phase == "i"), 
+                 aes(x=factor(site, levels = sum_site$site), y=vg_mean)) +
+    geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
+    geom_errorbar(aes(ymin=vg_mean-vg_se, ymax=vg_mean+vg_se), width=.2,
+                 position=position_dodge(.9)) +
+  scale_fill_brewer(palette = "Blues") +
+  labs(y = expression('Grazing impact (cm'^3*'/hr)'), 
+       x = "", 
+       title = "S. vetula") +
+  theme_bw() + 
+  theme(plot.margin = margin(1, 0, 0, 5, unit = "pt"),
+        axis.title.y = element_text(size=10),
+        axis.text.x = element_text(angle = 45, hjust = 1, size=7),
+        legend.position = "none",
+        plot.title = element_text(hjust = 0.5, face = "italic")
+        )
+
+vg_stop <- ggplot(filter(sum_ssp_vg, species_code == "stop" & phase == "i"), 
+                  aes(x=factor(site, levels = sum_site$site), y=vg_mean)) +
+    geom_bar(stat = "identity", aes(fill = island), position = "dodge") +    
+    geom_errorbar(aes(ymin=vg_mean-vg_se, ymax=vg_mean+vg_se), width=.2,
+                 position=position_dodge(.9)) +
+  scale_fill_brewer(palette = "Blues") +
+  labs(y = "",
+       x = "", 
+       title = "S. viride") +
+  theme_bw() + 
+  theme(plot.margin = margin(1, 1, 0, 0, unit = "pt"),
+        axis.text.x = element_text(angle = 45, hjust = 1, size=7),
+        legend.position = "none",
+        plot.title = element_text(hjust = 0.5, face = "italic")
+        )
+
+get_legend<-function(myggplot){
+  tmp <- ggplot_gtable(ggplot_build(myggplot))
+  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
+  legend <- tmp$grobs[[leg]]
+  return(legend)
+  }
+legend <- get_legend(vg_facet)
+
+grid.arrange(
+  arrangeGrob(vg_qup, vg_stop, ncol=2), 
+              arrangeGrob(legend), 
+              ncol=2, widths = c(5,1), heights = c(1))
+
+ggplot(filter(sum_ssp_vg, species_code != "rbp" & phase == "i"), aes(x = fr_mean, y = vg_mean)) +
+  geom_point(shape=18) +
+  facet_grid(~species) +
+  labs(y = "", 
+       title = "") +
+  theme_bw() + 
+  theme(plot.title = element_text(hjust = 0.5),
+        )
 ```
@@
 fishing + habitat degradation - which may have reinforcing/compounding negative effects on reef health
+
+# Potential additions:
+
+## PCA
+```{r, echo=F, message=F, warning=F}
+pca_pred <- prcomp(na.omit(sum_site %>% 
+              select(rugosity, ma_cover, ma_canopy, ta_cover, ta_canopy, scar_den, 
+                     scar_bm, carn30_bm)),
+  center = T, scale = T)
+
+autoplot(pca_pred, data = sum_site, colour = 'island',
+         loadings = T, loadings.colour = 'gray',
+         loadings.label = T, loadings.label.size = 3, loadings.label.colour = 'black',
+            loadings.label.repel = T) +
+    scale_color_brewer(palette = "Blues") +
+    theme_bw() +
+    theme(legend.title = element_blank())
+ggsave(here("figs/pca_pred.png"))
+
+sum_site <- sum_site %>% 
+  bind_cols(as.data.frame(pca_pred$x) %>% 
+           select(PC1,PC2)) %>%
+  rename(pc1_all=PC1,
+         pc2_all=PC2)
+
+sum_id_pca <- sum_id %>% 
+  left_join(select(sum_site,site,pc1_all,pc2_all), by = "site")
+sum_ssp_pca <- sum_ssp %>%
+  left_join(select(sum_site,site,pc1_all,pc2_all), by = "site")
+```
+
+## GAMM results
+
+### Predicting feeding rate:
+
+```{r, echo=F, message=F, warning=F}
+m_data <- sum_ssp_pca %>% filter(species_code!="rbp")
+
+gamm1 <- gamm(fr_mean ~ species_code + phase + s(length_m, k=4) + s(pc1_all, k=4) + s(pc2_all, k=4), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           method = "REML",
+           data = m_data
+           )
+g1 <- c("g1", "Species + Phase + Size + PC1 + PC2", AICc(gamm1$lme), summary(gamm1$gam)$r.sq)
+
+gamm2 <- gamm(fr_mean ~ species_code + phase + s(length_m, k=4) + s(pc1_all, k=4), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           method = "REML",
+           data = m_data
+           )
+g2 <- c("g2", "Species + Phase + Size + PC1", AICc(gamm2$lme), summary(gamm2$gam)$r.sq)
+
+gamm3 <- gamm(fr_mean ~ species_code + phase + s(length_m, by = factor(species_code), k=4) + s(pc1_all, by = factor(species_code), k=4) + s(pc2_all, by = factor(species_code), k=4), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           method = "REML",
+           data = m_data
+           )
+g3 <- c("g3", "Species + Phase + Size~Species + PC1~Species + PC2~Species", AICc(gamm3$lme), summary(gamm3$gam)$r.sq)
+
+gamm4 <- gamm(fr_mean ~ species_code + phase + s(length_m, by = factor(species_code), k=4) + s(pc1_all, by = factor(species_code), k=4) + s(pc2_all, k=4), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           method = "REML",
+           data = m_data
+           )
+g4 <- c("g4", "Species + Phase + Size~Species + PC1~Species + PC2", AICc(gamm4$lme), summary(gamm4$gam)$r.sq)
+
+gamm5 <- gamm(fr_mean ~ species + phase + s(length_m, by = factor(species_code), k=4) + s(pc1_all, by = factor(species_code), k=4), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           method = "REML",
+           data = m_data
+           )
+g5 <- c("g5", "Species + Phase + Size~Species + PC1~Species", AICc(gamm5$lme), summary(gamm5$gam)$r.sq)
+
+gamm6 <- gamm(fr_mean ~ phase + s(length_m, k=4) + s(pc1_all, by = factor(species_code), k=4), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           method = "REML",
+           data = m_data
+           )
+g6 <- c("g6", "Phase + Size + PC1~Species", AICc(gamm6$lme), summary(gamm6$gam)$r.sq)
+
+gamm7 <- gamm(fr_mean ~ phase + s(length_m, by = factor(species_code), k=4) + s(pc1_all, k=4), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           method = "REML",
+           data = m_data
+           )
+g7 <- c("g7", "Phase + Size~Species + PC1", AICc(gamm7$lme), summary(gamm7$gam)$r.sq)
+
+gamm8 <- gamm(fr_mean ~ species + phase + s(length_m, by = factor(species_code), k=4) + s(pc1_all, k=4), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           method = "REML",
+           data = m_data
+           )
+g8 <- c("g8", "Species + Phase + Size~Species + PC1", AICc(gamm8$lme), summary(gamm8$gam)$r.sq)
+
+gamm9 <- gamm(fr_mean ~ species + phase + s(length_m, k=4) + s(pc1_all, by = factor(species_code), k=4), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           method = "REML",
+           data = m_data
+           )
+g9 <- c("g9", "Species + Phase + Size + PC1~Species", AICc(gamm9$lme), summary(gamm9$gam)$r.sq)
+
+gamm10 <- gamm(fr_mean ~ species + phase + s(pc1_all, by = factor(species_code), k=4), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           method = "REML",
+           data = m_data
+           )
+g10 <- c("g10", "Species + Phase + PC1~Species", AICc(gamm10$lme), summary(gamm10$gam)$r.sq)
+
+gamm11 <- gamm(fr_mean ~ phase + s(length_m, by = factor(species_code), k=4) + s(pc1_all, by = factor(species_code), k=4), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           method = "REML",
+           data = m_data
+           )
+g11 <- c("g11", "Phase + Size~Species + PC1~Species", AICc(gamm11$lme), summary(gamm11$gam)$r.sq)
+
+gtab_fr <- as.data.frame(rbind(g1,g2,g3,g4,g5,g6,g7,g8,g9,g10,g11)) %>% setNames(c("g","Model", "AICc", "Adj. R2")) %>% arrange(AICc)
+```
+
+```{r, echo=F, message=F}
+kable(gtab_fr[1:4, ])
+plot(gamm3$gam,pages=1)
+```
+
+### Predicting feeding intensity:
+
+```{r, echo=F, message=F, warning=F}
+m_data <- sum_ssp_pca %>% filter(species_code!="rbp")
+
+fi_gamm1 <- gamm(for_bites_mean ~ species_code + phase + s(length_m, k=4) + s(pc1_all, k=4) + s(pc2_all, k=4), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           method = "REML",
+           data = m_data
+           )
+# summary(fi_gamm1$gam)
+# plot(fi_gamm1$gam,pages=1)
+g1 <- c("g1", "Species + Phase + Size + PC1 + PC2", AICc(fi_gamm1$lme), summary(fi_gamm1$gam)$r.sq)
+
+fi_gamm2 <- gamm(for_bites_mean ~ species_code + phase + s(length_m, k=4) + s(pc1_all, k=4), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           method = "REML",
+           data = m_data
+           )
+g2 <- c("g2", "Species + Phase + Size + PC1", AICc(fi_gamm2$lme), summary(fi_gamm2$gam)$r.sq)
+
+fi_gamm3 <- gamm(for_bites_mean ~ species_code + phase + s(length_m, by = factor(species_code), k=4) + s(pc1_all, by = factor(species_code), k=4) + s(pc2_all, by = factor(species_code), k=4), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           method = "REML",
+           data = m_data
+           )
+g3 <- c("g3", "Species + Phase + Size~Species + PC1~Species", AICc(fi_gamm3$lme), summary(fi_gamm3$gam)$r.sq)
+
+fi_gamm4 <- gamm(for_bites_mean ~ species_code + phase + s(length_m, by = factor(species_code), k=4) + s(pc1_all, by = factor(species_code), k=4) + s(pc2_all, k=4), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           method = "REML",
+           data = m_data
+           )
+g4 <- c("g4", "Species + Phase + Size~Species + PC1~Species + PC2", AICc(fi_gamm4$lme), summary(fi_gamm4$gam)$r.sq)
+
+fi_gamm5 <- gamm(for_bites_mean ~ phase + s(length_m, by = factor(species_code), k=4) + s(pc1_all, by = factor(species_code), k=4), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           method = "REML",
+           data = m_data
+           )
+g5 <- c("g5", "Phase + Size~Species + PC1~Species", AICc(fi_gamm5$lme), summary(fi_gamm5$gam)$r.sq)
+
+fi_gamm6 <- gamm(for_bites_mean ~ species + phase + s(length_m, k=4) + s(pc1_all, by = factor(species_code), k=4), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           method = "REML",
+           data = m_data
+           )
+g6 <- c("g6", "Species + Phase + Size + PC1~Species", AICc(fi_gamm6$lme), summary(fi_gamm6$gam)$r.sq)
+
+fi_gamm7 <- gamm(for_bites_mean ~ species + phase + s(length_m, by = factor(species_code), k=4) + s(pc1_all, k=4), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           method = "REML",
+           data = m_data
+           )
+g7 <- c("g7", "Species + Phase + Size~Species + PC1", AICc(fi_gamm7$lme), summary(fi_gamm7$gam)$r.sq)
+
+fi_gamm8 <- gamm(for_bites_mean ~ phase + s(length_m, k=4) + s(pc1_all, by = factor(species_code), k=4), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           method = "REML",
+           data = m_data
+           )
+g8 <- c("g8", "Phase + Size + PC1~Species", AICc(fi_gamm8$lme), summary(fi_gamm8$gam)$r.sq)
+
+fi_gamm9 <- gamm(for_bites_mean ~ phase + s(length_m, by = factor(species_code), k=4) + s(pc1_all, k=4), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           method = "REML",
+           data = m_data
+           )
+g9 <- c("g9", "Phase + Size~Species + PC1", AICc(fi_gamm9$lme), summary(fi_gamm9$gam)$r.sq)
+
+gtab_fi <- as.data.frame(rbind(g1,g2,g3,g4,g5,g6,g7,g8,g9)) %>% setNames(c("g", "Model", "AICc", "Adj. R2")) %>% arrange(AICc)
+```
+
+```{r, echo=F, message=F}
+kable(gtab_fi[1:3, ])
+plot(gamm6$gam,pages=1)
+```

### 2019-03-01 r/Follows.Rmd
updates

@@
     br=bites/g_sec,
-    fr=bites/follow_sec*60*60,
+    fr=bites/follow_sec*60,
     tdb=bites/follow_sec*60*60*10.5

### 2019-03-02 r/Ch2 outline_V3.Rmd
organizing/updates to v3

@@
 ```
+
+**Abstract:** Herbivorous fish increase coral growth and survival by grazing down algal competitors.
+As coral reef health continues to decline globally, maximizing herbivory on reefs has become a primary management goal.
+However, existing management targets assume consistent grazing behavior across different reef systems.
+Here I document the grazing behavior of two scarid species at 13 sites in Antigua, Barbuda, and Bonaire.
+My analyses show significant differences in overall feeding rates, time spent grazing, and intensity of grazing across reef sites, which may alter the ecological impact of a given scarid population.
+I propose several potential mechanisms for these grazing differences, including variations in bite content which may make bite rates a misleading measure of overall grazing impact.
+As we set management targets to enhance herbivory on coral reefs, it is critical that we are accurately estimating the grazing impact of scarid populations.
+This paper demonstrates the variability of grazing behavior across different reef sites and calls for further investigation of the drivers and ecological implications of these inconsistencies.
 
@@
 ```{r}
-Variable <- c('Time spent grazing','Bite rate','Feeding rate','Feeding intensity')
+Variable <- c('Time spent grazing','Bite rate','Feeding rate','Grazing concentration')
 Definition <- c('Proportion of time spent grazing during follow','Bite rate while actively feeding','Bite rate over total follow time','Number of consecutive bites taken during single grazing foray')
@@
 
-## Variation among sites in multiple components of feeding behavior
-Fig.
-X shows intraspecific variation in both feeding rate (bites per hour) and feeding intensity (bites per foray) among sites.
-(Bite rates while grazing were relatively constant, indicating that differences in overall feeding rates were driven by differences in the amount of time spent grazing).
+Feeding rates, time spent grazing, and grazing concentration varied significantly across sites for both *S.
+vetula* and *S.
+viride* populations (Fig.
+2).
+[Fig.
+2A also shows substantially lower feeding rates in most Antiguan (*S.
+vetula* and *S.
+viride*) and Barbudan (*S.
+vetula*) sites as compared standard values used in Caribbean herbivory models - *if I include reference lines...*].
+Differences in site-level feeding rates stemmed from differences in the fraction of time that fish speng grazing (Fig.
+2B) as opposed to differences in bite rates while activitely feeding, which remained relatively constant accross sites (appendix).
+
+These behavioral variations may have important ecological consequences and mangement implications.
+If we use grazing data from one area to predict the impact of a given species in other parts of the region, we may be overlooking important behavioral differences and misrepresenting grazing levels of different scarid populations.
+Feeding rates are used in herbivory models to calculate the area grazed by a given fish community.
+In the Caribbean case, most standard feeding rate equations are obtained from data collected in Bonaire.
+My results suggest that applying Bonaire feeding rates to other Caribbean reef systems may overestimate grazing impact.
+
+Grazing concentration also plays a key role in determining the ecological impact of coral reef herbivores.
+More concentrated grazing may be more effective in removing algae and clearing substrate for growing or newly recruiting corals than bites dispersed throughout the reefscape.
+[**add citations here**]
 
@@
 
-Notes:
-
-* Can remove color coding by island if this is distracting.
-Unsure whether I should be highlighting differences between islands or if I should just be highlighting variation among sites (mixed feedback on this)
-
 ```{r, echo = F, message = F, warning = F}
 # Panel graph of site differences
+
+ref_qup15 <- (3329-33*15)/60 # published feeding rate (bites/min)
+ref_qup30 <- (3329-33*30)/60
+ref_stop15 <- (1089-17*15-56)/60
+ref_stop30 <- (1089-17*30-56)/60
 
@@
   scale_fill_brewer(palette = "Blues") +
+  #geom_abline(slope=0, intercept=ref_qup15,  col = "red",lty=2) +
+  #geom_abline(slope=0, intercept=ref_qup25,  col = "red",lty=2) +
+  #ylim(0, 50) +
   labs(y = "Feeding rate \n (bites/min)", 
@@
   scale_fill_brewer(palette = "Blues") +
+  #geom_abline(slope=0, intercept=ref_stop15,  col = "red",lty=2) +
+  #geom_abline(slope=0, intercept=ref_stop25,  col = "red",lty=2) +
   labs(y = "",
@@
 
-fb_qup <- 
+gfrac_qup <- 
   ggplot(filter(sum_ssp, species_code == "qup" & phase == "i"), 
-         aes(x=factor(site, levels = sum_site$site), y=for_bites_mean)) +
+         aes(x=factor(site, levels = sum_site$site), y=g_frac_mean)) +
   geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
-  geom_errorbar(aes(ymin=for_bites_mean-for_bites_se, ymax=for_bites_mean+for_bites_se), width=.2,
+  geom_errorbar(aes(ymin=g_frac_mean-g_frac_se, ymax=g_frac_mean+g_frac_se), width=.2,
                  position=position_dodge(.9)) +
   scale_fill_brewer(palette = "Blues") +
-  labs(y = "Feeding intensity\n(bites/foray)", 
+  labs(y = "Time grazing \n(fraction)", 
        x = "", 
@@
 
-fb_stop <- 
+gfrac_stop <- 
   ggplot(filter(sum_ssp, species_code == "stop" & phase == "i"), 
-         aes(x=factor(site, levels = sum_site$site), y=for_bites_mean)) +
+         aes(x=factor(site, levels = sum_site$site), y=g_frac_mean)) +
   geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
-  geom_errorbar(aes(ymin=for_bites_mean-for_bites_se, ymax=for_bites_mean+for_bites_se), width=.2,
+  geom_errorbar(aes(ymin=g_frac_mean-g_frac_se, ymax=g_frac_mean+g_frac_se), width=.2,
                  position=position_dodge(.9)) +
@@
 
-gfrac_qup <- 
+fi_qup <- 
   ggplot(filter(sum_ssp, species_code == "qup" & phase == "i"), 
-         aes(x=factor(site, levels = sum_site$site), y=g_frac_mean)) +
+         aes(x=factor(site, levels = sum_site$site), y=for_bites_mean)) +
   geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
-  geom_errorbar(aes(ymin=g_frac_mean-g_frac_se, ymax=g_frac_mean+g_frac_se), width=.2,
+  geom_errorbar(aes(ymin=for_bites_mean-for_bites_se, ymax=for_bites_mean+for_bites_se), width=.2,
                  position=position_dodge(.9)) +
   scale_fill_brewer(palette = "Blues") +
-  labs(y = "Time grazing \n(fraction)", 
+  labs(y = "Grazing concentration\n(bites/foray)", 
        x = "", 
@@
 
-gfrac_stop <- 
+fi_stop <- 
   ggplot(filter(sum_ssp, species_code == "stop" & phase == "i"), 
-         aes(x=factor(site, levels = sum_site$site), y=g_frac_mean)) +
+         aes(x=factor(site, levels = sum_site$site), y=for_bites_mean)) +
   geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
-  geom_errorbar(aes(ymin=g_frac_mean-g_frac_se, ymax=g_frac_mean+g_frac_se), width=.2,
+  geom_errorbar(aes(ymin=for_bites_mean-for_bites_se, ymax=for_bites_mean+for_bites_se), width=.2,
                  position=position_dodge(.9)) +
@@
   geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
-  geom_errorbar(aes(ymin=g_frac_mean-g_frac_se, ymax=g_frac_mean+g_frac_se), width=.2,
-                 position=position_dodge(.9)) +
   scale_fill_brewer(palette = "Blues") +
@@
 grid.arrange(
-  arrangeGrob(fr_qup, fr_stop, fb_qup, fb_stop, gfrac_qup, gfrac_stop, ncol=2, heights = c(1,1,1.5)), 
+  arrangeGrob(fr_qup, fr_stop, gfrac_qup, gfrac_stop, fi_qup, fi_stop, ncol=2, heights = c(1,1,1.5)), 
               arrangeGrob(legend), 
@@
 
-## Several potential underlying mechanisms
+The variations in grazing behaviors demonstrated here may be driven by several potential underlying mechanisms.
+Figure 3 shows the relationships between various reef characteristics and scarid feeding rate.
+Because study sites expectedly varied across a number of interrelated fish and benthic dimensions (PCA - appendix?), it is impossible to conclusively establish any explicit mechanisms here.
+Instead I present several hypotheses that may be driving these observed behavioral trends and discuss evidence for and implications of each.
 
@@
 
-While we do not seek to conclusively establish mechanisms here, we discuss potential underlying mechanisms that could be driving these trends in herbivore feeding behaviors
+*Bite content hypothesis.* Differences in scarid grazing preferences can be driven by (local differences in) nutrient availability and distribution (Carlson, Tootell, Mumby, etc.).
+Algal communities on Caribbean reefs can vary greatly in terms of canopy height, percent cover, species composition, and nutrient content ().
 
-* Bite content/nutrition hypothesis: scarids in higher canopy sites are consuming more per bite, thus taking fewer bites - implications for spatial grazing impact.
-Previous work showing that scarids vary grazing with nutrient availability/distribution (Carlson, Tootell, Mumby, etc.)
+(no conclusive trends, but FR tended to decrease with increasing EAM canopy height and to some extent EAM and macroalgal percent cover)
+
+In higher canopy sites, scarids may be obtaining more biomass of food per bite than in areas with heavily cropped algae, requiring them to take fewer bites to obtain the same nutritional intake.
+
+Implications: 
+-maybe better metric than bite/feeding rate
+-reduced grazing impact if not clearing substrate
+-reinforcing loop
 
@@
 
-* Social feeding hypothesis: more time must be spent vigilant in smaller groups, potential compounding effect with overfishing/loss of herbivory (e.g. Gil)
-* Predation risk hypothesis: presence of large piscivores does not explain feeding trends, but risk from spearfishers may play a bigger role (able to target larger fish, i.e.
-those sampled here).
-(Incorporate rugosity here - Rogers, Madin, Catano).
-Did not find strong evidence for this here, but carnivore biomass was strongly correllated with herbivore biomass and rugosity.
+*Social feeding hypothesis.* more time must be spent vigilant in smaller groups, potential compounding effect with overfishing/loss of herbivory (e.g. Gil)
+
+*Predation risk hypothesis.* Previous work has documented the suppressive effect of predator presence on herbivore feeding rates (Madin, Catano, McCauley?). 
+- These studies have focused on natural predators (decoys or piscivore populations)
+- While I found no clear relationship between large (>30cm) piscivore biomass and feeding trends (appendix), risk of predation from spearfishers may play a larger role in determining scarid grazing behavior at the reefs studied here.
+All Antiguan and Barbudan sites experience some degree of spearfishing pressure, while spearfishing has been banned in Bonaire since the 1970s.
+The 15-30cm individuals used in my behavioral observations are approaching size refuge from most natural predators, but would be likely targets for spearfishers.
+Both *S.
+vetula* and *S.
+viride* are commonly harvested species in Antigua and Barbuda.
+
+- like with the social feeding hypothesis, spearfishing may have double impact of reducing fish and reducing grazing impact of remaining fish
+
+- often interaction of rugosity (e.g. Madin, stay near structure)
+(Incorporate rugosity here - Rogers, Madin, Catano). Did not find strong evidence for this here, but carnivore biomass was strongly correllated with herbivore biomass and rugosity.
 
@@
 * Potential reinforcing feedback loops in degraded reefs if high algal/low scarid reefs also have less grazing impact per scarid
-    * Feeding intensity: less intense feeding/more distributed bites in degraded reefs may also reduce actual substrate made available to recruiting corals (~ grazing concentration lit. here)
 

### 2019-03-02 r/old drafts/Ch2 outline_V1.Rmd
organizing/updates to v3

---
title: "Manuscript draft"
author: "Molly Wilson"
date: "Dec. 2018"
output:
  html_document:
    toc: true
    toc_float: true
    theme: united
    number_sections: true
---

```{r, echo=F, message=F, warning=F}
library(startR)
library(here)
library(tidyverse)
library(gridExtra)
library(ggfortify) # autoplot(prcomp())
library(mgcv)
library(sme)

sum_site <- read.csv(here("data","sum_site.csv"), stringsAsFactors=F) %>%
  select(-X)
sum_id <- read.csv(here("data","sum_id.csv"), stringsAsFactors=F) %>%
  select(-X) %>% na.omit()
sum_ss <- read.csv(here("data","sum_site_sp.csv"), stringsAsFactors=F) %>%
  select(-X)
sum_ssp <- read.csv(here("data","sum_site_sp_ph.csv"), stringsAsFactors=F) %>%
  select(-X)
```

# Intro + motivation

* Grazing by herbivorous fish is an established driver of coral reef health/dynamics.
* As reefs become increasingly degraded/threatened, there is growing interest in understanding how much herbivory is needed to sustain coral reef health (e.g. Karr et al. 2015, McClanahan et al. 2011).
* However, not all herbivores are the same in terms of their grazing impact…
    * Recent developments in functional classifications, difference across species in terms of selectivity, etc. (e.g. Adam et al. 2018).
* In addition to these inter-species differences in functional role, differences in herbivore behavior may also affect grazing impact and may drive variation in herbivore function among individuals or across reef systems.
* Previous work has shown that herbivore behavior is sensitive to...
    * Predator presence (and reef structure) (Catano, Madin)
    * Presence of other herbivores (Gil)
    * Resource availability (Carlson, Tootell)
* But there is very limited data on how herbivore feeding behaviors might vary across reef systems, as well as the environmental factors associated with grazing differences.
* Here we document variability in multiple components of scarid feeding behavior across reef sites in three Caribbean islands and explore potential underlying mechanisms, ecological impacts, and management implications.
* We demonstrate a need for further experimental and modelling work to unpack these drivers and understand their ecological consequences.

# Methods

## Data collection
* Sites (13, [map]), timeline
* Fish and benthic community data
* Behavioral data:
    - Species: *Sparisoma viride* and *Scarus vetula*, initial and terminal phase greater than 15cm
    - 2 minute follows, calibration period, outside of previously established FID
    - Recording time budgets as well as bite rates during grazing forays (define)
    - Discarded data of incomplete follows or where fish were visably disturbed/fled
    - Timing of data collection distributed throughout day, 2 dives per site
    - Define behavioral variables

## Analysis
* Fish biomass using lenth-weight conversions (appendix)
* PCA of site-level benthic and fish community variables
* GAMM: mgcv, island as random effect, Gaussian distribution, compared using AICc. Used data aggregated to the site level
* (Predicting community grazing impact: calculated length - feeding rate relationships using linear regression for each phase/island/species, applied these to observed site-level fish communities. Integrated bite size data from Bruggemann et al., then compared with predictions made with published species length - feeding rate relationships)

# Results

## Variation in feeding rate across benthic and fish community variables
```{r, echo=F, message=F, warning=F}
q1 <- ggplot(filter(sum_ssp, species_code == "qup" & phase == "i"), aes(x = scar_bm/1000, y = fr_mean/60)) +
  geom_smooth(method="loess") +
  geom_point(shape=18) +
  ylim(-10, 60) +
  labs(y = "",
       # y = "S. vetula \nFeeding rate (bites/min)", 
       x = "", 
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        plot.margin = margin(0, 1, 0, 1, unit = "pt")
        )

q2 <- ggplot(filter(sum_ssp, species_code == "qup" & phase == "i"), aes(x = consp_den, y = fr_mean/60)) +
  geom_smooth(method="loess") +
  geom_point(shape=18) +
  ylim(-10, 60) +
  labs(y = "", 
       x = "",
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        plot.margin = margin(1, 1, 0, 1, unit = "pt")
        )

q3 <- ggplot(filter(sum_ssp, species_code == "qup" & phase == "i"), aes(x = rugosity, y = fr_mean/60)) +
  geom_smooth(method="loess") +
  geom_point(shape=18) +
  ylim(-10, 60) +
  labs(y = "", 
       x = "", 
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        plot.margin = margin(1, 1, 0, 1, unit = "pt")
        )

q4 <- ggplot(filter(sum_ssp, species_code == "qup" & phase == "i"), aes(x = ta_canopy, y = fr_mean/60)) +
  geom_smooth(method="loess") +
  geom_point(shape=18) +
  ylim(-10, 60) +
  labs(y = "", 
       x = "",
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        plot.margin = margin(1, 1, 0, 1, unit = "pt")
        )

q5 <- ggplot(filter(sum_ssp, species_code == "qup" & phase == "i"), aes(x = ta_cover, y = fr_mean/60)) +
  geom_smooth(method="loess") +
  geom_point(shape=18) +
  ylim(-10, 60) +
  labs(y = "",
       x = "",
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        plot.margin = margin(1, 1, 0, 1, unit = "pt")
        )

q7 <- ggplot(filter(sum_ssp, species_code == "qup" & phase == "i"), aes(x = ma_cover, y = fr_mean/60)) +
  geom_smooth(method="loess") +
  geom_point(shape=18) +
  ylim(-10, 60) +
  labs(y = "", 
       x = "",
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        plot.margin = margin(1, 1, 0, 1, unit = "pt")
        )

s1 <- ggplot(filter(sum_ssp, species_code == "stop" & phase == "i"), aes(x = scar_bm/1000, y = fr_mean/60)) +
  geom_smooth(method="loess") +
  geom_point(shape=18) +
  ylim(-5, 20) +
  labs(y = "",
       x = "",
       # y = "S. viride \nFeeding rate (bites/min)", 
       # x = expression(atop(Scarid~biomass, "(kg/100m"^2*")")),
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        plot.margin = margin(0, 1, 0, 1, unit = "pt")
        )

s2 <- ggplot(filter(sum_ssp, species_code == "stop" & phase == "i"), aes(x = consp_den, y = fr_mean/60)) +
  geom_smooth(method="loess") +
  geom_point(shape=18) +
  ylim(-5, 20) +
  labs(y = "", 
       x = "",
       # x = expression(atop(Conspec.~density, "(indv./100m"^2*")")),
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        plot.margin = margin(1, 1, 0, 1, unit = "pt")
        )

s3 <- ggplot(filter(sum_ssp, species_code == "stop" & phase == "i"), aes(x = rugosity, y = fr_mean/60)) +
  geom_smooth(method="loess") +
  geom_point(shape=18) +
  ylim(-5, 20) +
  labs(y = "", 
       x = "",
       #x = expression("Rugosity"), 
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        plot.margin = margin(1, 1, 0, 1, unit = "pt")
        )

s4 <- ggplot(filter(sum_ssp, species_code == "stop" & phase == "i"), aes(x = ta_canopy, y = fr_mean/60)) +
  geom_smooth(method="loess") +
  geom_point(shape=18) +
  ylim(-5, 20) +
  labs(y = "", 
       x = "",
       # x = expression("EAM canopy \nheight (mm)"), 
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        plot.margin = margin(1, 1, 0, 1, unit = "pt")
        )

s5 <- ggplot(filter(sum_ssp, species_code == "stop" & phase == "i"), aes(x = ta_cover, y = fr_mean/60)) +
  geom_smooth(method="loess") +
  geom_point(shape=18) +
  ylim(-5, 20) +
  labs(y = "", 
       x = "",
       # x = expression("EAM percent \ncover"), 
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        plot.margin = margin(1, 1, 0, 1, unit = "pt")
        )

s7 <- ggplot(filter(sum_ssp, species_code == "stop" & phase == "i"), aes(x = ma_cover, y = fr_mean/60)) +
  geom_smooth(method="loess") +
  geom_point(shape=18) +
  ylim(-5, 20) +
  labs(y = "", 
       x = "",
       # x = expression("Macroalgal \npercent cover"), 
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        plot.margin = margin(1, 1, 0, 1, unit = "pt")
        )

grid.arrange(left = "Feeding rate (bites / minute)",
  arrangeGrob(q1, q3, q4, q5, q7, ncol=5, right = "S. vetula"),
  arrangeGrob(
    arrangeGrob(s1, bottom = "Scarid biomass"), 
    arrangeGrob(s3, bottom = "Rugosity"),
    arrangeGrob(s4, bottom = "Turf height"),
    arrangeGrob(s5, bottom = "Turf % cover"),
    arrangeGrob(s7, bottom = "Macro. % cover"),
    ncol=5, right = "S. viride"), 
  nrow=2) 
```

## Variation in feeding intensity across benthic and fish community variables
```{r, echo=F, message=F, warning=F}
q1 <- ggplot(filter(sum_ssp, species_code == "qup" & phase == "i"), aes(x = scar_bm/1000, y = for_bites_mean)) +
  geom_smooth(method="loess") +
  geom_point(shape=18) +
  ylim(-10, 25) +
  labs(y = "",
       # y = "S. vetula \nFeeding rate (bites/min)", 
       x = "", 
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        plot.margin = margin(1, 1, 0, 1, unit = "pt")
        )

q2 <- ggplot(filter(sum_ssp, species_code == "qup" & phase == "i"), aes(x = consp_den, y = for_bites_mean)) +
  geom_smooth(method="loess") +
  geom_point(shape=18) +
  ylim(-10, 25) +
  labs(y = "", 
       x = "",
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        plot.margin = margin(1, 1, 0, 1, unit = "pt")
        )

q3 <- ggplot(filter(sum_ssp, species_code == "qup" & phase == "i"), aes(x = rugosity, y = for_bites_mean)) +
  geom_smooth(method="loess") +
  geom_point(shape=18) +
  ylim(-10, 25) +
  labs(y = "", 
       x = "", 
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        plot.margin = margin(1, 1, 0, 1, unit = "pt")
        )

q4 <- ggplot(filter(sum_ssp, species_code == "qup" & phase == "i"), aes(x = ta_canopy, y = for_bites_mean)) +
  geom_smooth(method="loess") +
  geom_point(shape=18) +
  ylim(-10, 25) +
  labs(y = "", 
       x = "",
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        plot.margin = margin(1, 1, 0, 1, unit = "pt")
        )

q5 <- ggplot(filter(sum_ssp, species_code == "qup" & phase == "i"), aes(x = ta_cover, y = for_bites_mean)) +
  geom_smooth(method="loess") +
  geom_point(shape=18) +
  ylim(-10, 25) +
  labs(y = "",
       x = "",
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        plot.margin = margin(1, 1, 0, 1, unit = "pt")
        )

q7 <- ggplot(filter(sum_ssp, species_code == "qup" & phase == "i"), aes(x = ma_cover, y = for_bites_mean)) +
  geom_smooth(method="loess") +
  geom_point(shape=18) +
  ylim(-10, 25) +
  labs(y = "", 
       x = "",
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        plot.margin = margin(1, 1, 0, 1, unit = "pt")
        )

s1 <- ggplot(filter(sum_ssp, species_code == "stop" & phase == "i"), aes(x = scar_bm/1000, y = for_bites_mean)) +
  geom_smooth(method="loess") +
  geom_point(shape=18) +
  ylim(-10, 25) +
  labs(y = "",
       x = "",
       # y = "S. viride \nFeeding rate (bites/min)", 
       # x = expression(atop(Scarid~biomass, "(kg/100m"^2*")")),
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        plot.margin = margin(1, 1, 0, 1, unit = "pt")
        )

s2 <- ggplot(filter(sum_ssp, species_code == "stop" & phase == "i"), aes(x = consp_den, y = for_bites_mean)) +
  geom_smooth(method="loess") +
  geom_point(shape=18) +
  ylim(-10, 25) +
  labs(y = "", 
       x = "",
       # x = expression(atop(Conspec.~density, "(indv./100m"^2*")")),
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        plot.margin = margin(1, 1, 0, 1, unit = "pt")
        )

s3 <- ggplot(filter(sum_ssp, species_code == "stop" & phase == "i"), aes(x = rugosity, y = for_bites_mean)) +
  geom_smooth(method="loess") +
  geom_point(shape=18) +
  ylim(-10, 25) +
  labs(y = "", 
       x = "",
       #x = expression("Rugosity"), 
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        plot.margin = margin(1, 1, 0, 1, unit = "pt")
        )

s4 <- ggplot(filter(sum_ssp, species_code == "stop" & phase == "i"), aes(x = ta_canopy, y = for_bites_mean)) +
  geom_smooth(method="loess") +
  geom_point(shape=18) +
  ylim(-10, 25) +
  labs(y = "", 
       x = "",
       # x = expression("EAM canopy \nheight (mm)"), 
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        plot.margin = margin(1, 1, 0, 1, unit = "pt")
        )

s5 <- ggplot(filter(sum_ssp, species_code == "stop" & phase == "i"), aes(x = ta_cover, y = for_bites_mean)) +
  geom_smooth(method="loess") +
  geom_point(shape=18) +
  ylim(-10, 25) +
  labs(y = "", 
       x = "",
       # x = expression("EAM percent \ncover"), 
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        plot.margin = margin(1, 1, 0, 1, unit = "pt")
        )

s7 <- ggplot(filter(sum_ssp, species_code == "stop" & phase == "i"), aes(x = ma_cover, y = for_bites_mean)) +
  geom_smooth(method="loess") +
  geom_point(shape=18) +
  ylim(-10, 25) +
  labs(y = "", 
       x = "",
       # x = expression("Macroalgal \npercent cover"), 
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        plot.margin = margin(1, 1, 0, 1, unit = "pt")
        )

grid.arrange(left = "Feeding intensity (bites / foray)",
  arrangeGrob(q1, q3, q4, q5, q7, ncol=5, right = "S. vetula"),
  arrangeGrob(
    arrangeGrob(s1, bottom = "Scarid biomass"), 
    arrangeGrob(s3, bottom = "Rugosity"),
    arrangeGrob(s4, bottom = "Turf height"),
    arrangeGrob(s5, bottom = "Turf % cover"),
    arrangeGrob(s7, bottom = "Macro. % cover"),
    ncol=5, right = "S. viride"), 
  nrow=2) 
```

## PCA
```{r, echo=F, message=F, warning=F}
pca_pred <- prcomp(na.omit(sum_site %>% 
              select(rugosity, ma_cover, ma_canopy, ta_cover, ta_canopy, scar_den, 
                     scar_bm, carn30_bm, spear_daily)),
  center = T, scale = T)

autoplot(pca_pred, data = sum_site, colour = 'island',
         # frame = T, frame.type = "norm",
         loadings = T, loadings.colour = 'gray',
         loadings.label = T, loadings.label.size = 3, loadings.label.colour = 'black',
            loadings.label.repel = T) +
    scale_color_manual(name="island", values=c("lightblue", "darkseagreen", "lightsalmon")) + 
    theme(legend.direction = 'horizontal', legend.position = 'top') +
    theme_minimal()
ggsave(here("figs/pca_pred.png"))

sum_site <- sum_site %>% 
  bind_cols(as.data.frame(pca_pred$x) %>% 
           select(PC1,PC2)) %>%
  rename(pc1_all=PC1,
         pc2_all=PC2)

sum_id_pca <- sum_id %>% 
  left_join(select(sum_site,site,pc1_all,pc2_all), by = "site")
sum_ssp_pca <- sum_ssp %>%
  left_join(select(sum_site,site,pc1_all,pc2_all), by = "site")
```
```{r, eval = F, echo=F, message=F, warning=F}
ggplot(filter(sum_ssp_pca, species_code != "rbp" & phase == "i"), aes(x = pc1_all, y = fr_mean/60)) +
  facet_wrap(.~species) +
  geom_smooth(method="loess") +
  geom_point(shape=18) +
  labs(y = "Feeding rate (bites/min)", 
       x = "PC1")
  theme_bw()
```

## GAMM results

```{r, eval=F, echo=F, message=F, warning=F}
m_data <- sum_ssp_pca %>% filter(species_code!="rbp")
# Main effects only
gamm1 <- gamm(fr_mean ~ species_code + s(length_m) + phase + s(pc1_all) + s(pc2_all), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
summary(gamm1$gam)
plot(gamm1$gam,pages=1)
plot(gamm1$lme) # not too heteroskedastic?

# Length/species and PC1/species interactions
gamm2 <- gamm(fr_mean ~ species_code + phase + s(length_m, by = factor(species_code)) + s(pc1_all, by = factor(species_code)) + s(pc2_all), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
summary(gamm2$gam)
plot(gamm2$gam,pages=1)
plot(gamm2$lme)

# Length/species and PC1/species interactions, no PC2
gamm3 <- gamm(fr_mean ~ species_code + phase + s(length_m, by = factor(species_code)) + s(pc1_all, by = factor(species_code)), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
summary(gamm3$gam)
plot(gamm3$gam,pages=1)
plot(gamm3$lme)

AIC(gamm1$lme, gamm2$lme, gamm3$lme)
AICc(gamm1$lme)
AICc(gamm2$lme)
AICc(gamm3$lme)
```

```{r, echo=F, message=F, warning=F}
m_data <- sum_ssp_pca %>% filter(species_code!="rbp")
# Length/species and PC1/species interactions
gamm2 <- gamm(fr_mean ~ species_code + phase + s(length_m, by = factor(species_code)) + s(pc1_all, by = factor(species_code)) + s(pc2_all), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
summary(gamm2$gam)
plot(gamm2$gam,pages=1)
```

# Discussion

* GAMM results: explain meaning/significance of PC1. *S. vetula* seem more impacted by PC1 than *S. viride* 
* Potential mechanisms/hypotheses for these observed trends:
    * Social feeding hypothesis: more time must be spent vigilant in smaller groups, potential compounding effect with overfishing/loss of herbivory
    * Nutrient limitation hypothesis: scarids must graze more (and more intensely) in sites with lower algal canopy heights and percent cover
    * Predation risk hypothesis: presence of large piscivores does not explain feeding trends, but risk from spearfishers may play a bigger role (able to target larger fish, i.e. those sampled here). (Incorporate rugosity here?)
* Caveats + limitations: all these different factors are understandably correlated in observational work
    * room for experimental work to unpack/distinguish drivers
    * room (/need) for incorporation of behavior in coral reef grazing models, especially those used to set targets for minimum herbivore biomass levels, because these may look very different in different reef conditions.
* Conservation + mgmt implications: multiple human impacts on reefs may affect herbivore behavior in different ways, spec. fishing + habitat degradation - which may have reinforcing/compounding negative effects on reef health

### 2019-03-02 r/old drafts/Manuscript draft.Rmd
organizing/updates to v3

--- 2019-03-02 r/old drafts/Ch2 outline_V1.Rmd
@@
 
-* Grazing by herbivorous fish is an established driver of coral reef health/dynamics.
-* As reefs become increasingly degraded/threatened, there is growing interest in understanding how much herbivory is needed to sustain coral reef health (e.g.
-Karr et al.
-2015, McClanahan et al.
-2011).
+* Grazing by herbivorous fish is an established driver of coral reef health/dynamics
+* As reefs become increasingly degraded/threatened, there is growing interest in understanding how much herbivory is needed to sustain coral reef health (e.g Karr et al., McClanahan et al.)
 * However, not all herbivores are the same in terms of their grazing impact…
-    * Recent developments in functional classifications, difference across species in terms of selectivity, etc. (e.g. Adam et al. 2018).
-* In addition to these inter-species differences in functional role, differences in herbivore behavior may also affect grazing impact and may drive variation in herbivore function among individuals or across reef systems.
-* Previous work has shown that herbivore behavior is sensitive to...
-    * Predator presence (and reef structure) (Catano, Madin)
+    * Recent developments in functional classifications, difference across species in terms of selectivity, etc. (e.g. Adam et al. 2018)
+* In addition to these inter-species differences in functional role, differences in herbivore behavior also affect grazing impact and may drive variation in herbivore function among individuals or across reef systems
+* Herbivore behavior has been shown to be impacted by…
+    * Predator presence (Catano, Madin)
     * Presence of other herbivores (Gil)
     * Resource availability (Carlson, Tootell)
-* But there is very limited data on how herbivore feeding behaviors might vary across reef systems, as well as the environmental factors associated with grazing differences.
-* Here we document variability in multiple components of scarid feeding behavior across reef sites in three Caribbean islands and explore potential underlying mechanisms, ecological impacts, and management implications.
-* We demonstrate a need for further experimental and modelling work to unpack these drivers and understand their ecological consequences.
+* But there is vary limited data on how herbivore feeding behaviors might vary across reef systems, as well as the environmental factors associated with grazing differences
+* Here we document variability in multiple components of scarid feeding behavior across reef sites in three Caribbean islands and explore/enhance our understanding of potential underlying drivers
 
@@
     arrangeGrob(s3, bottom = "Rugosity"),
-    arrangeGrob(s4, bottom = "Turf height"),
-    arrangeGrob(s5, bottom = "Turf % cover"),
-    arrangeGrob(s7, bottom = "Macro. % cover"),
+    arrangeGrob(s4, bottom = "EAM height"),
+    arrangeGrob(s5, bottom = "EAM % cover"),
+    arrangeGrob(s7, bottom = "MA % cover"),
     ncol=5, right = "S. viride"), 
@@
 pca_pred <- prcomp(na.omit(sum_site %>% 
-              select(rugosity, ma_cover, ma_canopy, ta_cover, ta_canopy, scar_den, 
-                     scar_bm, carn30_bm, spear_daily)),
+              remove_rownames %>% 
+              column_to_rownames(var="site") %>%
+              select(ma_cover, ma_canopy, ta_cover, ta_canopy, lc_cover, rugosity, scar_den, 
+                     scar_bm, carn30_bm)),
   center = T, scale = T)
@@
 
+```{r, echo=F, message=F, warning=F}
+m_data <- sum_ssp_pca %>% filter(species_code!="rbp")
+```
+
 ```{r, eval=F, echo=F, message=F, warning=F}
-m_data <- sum_ssp_pca %>% filter(species_code!="rbp")
 # Main effects only
@@
 ```{r, echo=F, message=F, warning=F}
-m_data <- sum_ssp_pca %>% filter(species_code!="rbp")
 # Length/species and PC1/species interactions
@@
 (Incorporate rugosity here?)
-* Caveats + limitations: all these different factors are understandably correlated in observational work
+* Caveats + limitations: all these different factors are understandablycorrelated in observational work
     * room for experimental work to unpack/distinguish drivers

### 2019-04-16 r/Ch2 outline_V3.Rmd
4.15

@@
 
-```{r}
-Variable <- c('Time spent grazing','Bite rate','Feeding rate','Grazing concentration')
+```{r, echo=F, message=F}
+Variable <- c('Time spent grazing','Bite rate','Feeding rate','Foray size')
 Definition <- c('Proportion of time spent grazing during follow','Bite rate while actively feeding','Bite rate over total follow time','Number of consecutive bites taken during single grazing foray')
@@
 
-Feeding rates, time spent grazing, and grazing concentration varied significantly across sites for both *S.
+Feeding rates, time spent grazing, and grazing concentration varied significantly across sites within both *S.
 vetula* and *S.
@@
 2B) as opposed to differences in bite rates while activitely feeding, which remained relatively constant accross sites (appendix).
+Interspecific differences in feeding rates and grazing intensity reflect the different feeding morphologies/functional roles of *S.
+vetula* and *S.
+viride*.
 
-These behavioral variations may have important ecological consequences and mangement implications.
+These behavioral variations may have important ecological consequences and management implications.
 If we use grazing data from one area to predict the impact of a given species in other parts of the region, we may be overlooking important behavioral differences and misrepresenting grazing levels of different scarid populations.
@@
               ncol=2, widths = c(5,1), heights = c(1))
+```
 
+```{r, eval = F}
 ggplot(filter(sum_ssp_vg, species_code != "rbp" & phase == "i"), aes(x = fr_mean, y = vg_mean)) +
@@
   facet_grid(~species) +
-  labs(y = "", 
+  labs(y = "Grazing impact (cm3/hr)",
+       x = "Average feeding rate (bites/min)"
        title = "") +
@@
 
-## Ecological implications
+## Management implications + future research needs
 
 * Potential reinforcing feedback loops in degraded reefs if high algal/low scarid reefs also have less grazing impact per scarid
-
-## Management implications + future research needs
 
@@
 fishing + habitat degradation - which may have reinforcing/compounding negative effects on reef health
+
+In conclusion...
 

### 2019-04-16 r/FID.Rmd
4.15

---
title: "Scarid Flight Response, Antigua"
author: "Molly Wilson"
date: "April 12, 2017"
output: html_document
---

```{r}
# prep
library(lazyeval)
library(here)
library(tidyverse)
library(reshape2)
```
```{r}
# import, format
fid <- read.csv(here("data","fid.csv"), stringsAsFactors=F) %>% 
  select(-c(FID.ft,FID.in,Bump.distance.ft,Bump.distance.in)) %>%
  janitor::clean_names()
```

# FID by species
```{r}
fid.sum <- fid %>%
  group_by(species, phase) %>%
  summarize(fid_mean = mean(fid_cm/100),
            fid_max = max(fid_cm/100)) %>%
  rename(max = fid_max,
         mean = fid_mean) %>%
  melt() %>%
  rename(fid = value)

ggplot(fid, aes(x = species, y = fid_cm)) +
  geom_boxplot()

ggplot(fid.sum %>% filter(phase == "Initial"), aes(x = species, y = fid, group = variable)) +
  geom_point(aes(color = variable))
```

# FID by spear, site
```{r}
lm.fid <-lm(data=fid, fid_cm ~ spear + site + rugosity + pre_approach_activity)
summary(lm.fid)

ggplot(fid, aes(x = pre_approach_activity,y=fid_cm)) + geom_point()
```

# bump distance by spear, site
```{r}
bump <- FID %>% filter(Bump.distance.cm != 0)
lmBump <-lm(data=bump,Bump.distance.cm~Spear + Site + Species + Rugosity + Pre.Approach.Activity)
summary(lmBump)
aov()

ggplot(bump, aes(x=Spear,y=Bump.distance.cm)) + geom_point(aes(color=Species))
```

### 2019-04-16 r/Follows.Rmd
4.15

@@
 sum_site <- sum_id %>% 
+  filter(length_cm >= 15 & length_cm <= 30) %>%
   group_by(site,island) %>% 
@@
 sum_site_sp <- sum_id %>% 
+  filter(length_cm >= 15 & length_cm <= 30) %>%
   group_by(site, island, species, species_code) %>% 
@@
 sum_site_sp_ph <- sum_id %>% 
+  filter(length_cm >= 15 & length_cm <= 30) %>%
   group_by(site, island, species, species_code, phase) %>% 

### 2019-04-16 r/map.Rmd
4.15

---
title: "Site map"
author: "Molly Wilson"
date: "3/2/2019"
output: html_document
---

```{r}
library(ggplot2)
library(ggmap)

map.world <- map_data("world")

map.carib <- ggplot() +
  geom_polygon(data = map.world, aes(x = long, y = lat, group = group))+
  coord_cartesian(xlim = c(-100, -50),ylim = c(30,5)) +
  labs(x = "", y="",title = "")

ggplot() +
  geom_polygon(data = map.world, aes(x = long, y = lat, group = group))+
  coord_cartesian(xlim = c(-62, -60),ylim = c(18,17)) +
  labs(x = "", y="",title = "")

ggplot() +
  theme_bw() +
  geom_polygon(data = map.world, aes(x = long, y = lat, group = group))+
  coord_equal(xlim = c(-100, -60),ylim = c(27,10)) +
  labs(x = "", y="",title = "")
```

### 2019-04-19 r/Ch2 outline_V3.Rmd
4.18

@@
         axis.title.x = element_blank(),
-        axis.title.y = element_text(size=10),
+        axis.title.y = element_text(size=8),
         legend.position = "none")
@@
   scale_fill_brewer(palette = "Blues") +
-  labs(y = "Grazing concentration\n(bites/foray)", 
+  labs(y = "Foray length\n(bites/foray)", 
        x = "", 
@@
   theme(plot.margin = margin(0, 0, 0, 5, unit = "pt"),
-        axis.title.y = element_text(size=10),
+        axis.title.y = element_text(size=8),
         axis.text.x = element_text(angle = 45, hjust = 1, size=7),
@@
 grid.arrange(
-  arrangeGrob(fr_qup, fr_stop, gfrac_qup, gfrac_stop, fi_qup, fi_stop, ncol=2, heights = c(1,1,1.5)), 
+  arrangeGrob(fr_qup, fr_stop, gfrac_qup, gfrac_stop, br_qup, br_stop, fi_qup, fi_stop, ncol=2, heights = c(1.5,1.5,1.5,3)), 
               arrangeGrob(legend), 
@@
 
+g <- arrangeGrob(
+  arrangeGrob(fr_qup, fr_stop, gfrac_qup, gfrac_stop, br_qup, br_stop, fi_qup, fi_stop, ncol=2, heights = c(1,1,1,1.5)), 
+              arrangeGrob(legend), 
+              ncol=2, widths = c(5,1))
+
+ggsave(here("figs/behavior_panel.png"),g)
 ```

### 2019-04-22 r/Analysis.Rmd
adding size range for sum_ssp_vg

--- 2018-11-26 r/Analysis.Rmd
@@
 ```
+
+# Bite content hypothesis
+
+Fish in high algal sites may be taking fewer bites because they can consume more algae per bite. Let's try and add a rough estimate of bite content as a function of algal canopy height.
+
+Notes/assumptions:
+* Using turf algal canopy height (for now, no macroalgal component - though S. viride does eat some macrogalgae)
+* Max bite height will be capped at radius of bite size ($\sqrt{BS}/\pi$)
+
+```{r}
+# grazing impact of fish from fish surveys
+fish_grazed <- fish_fr %>% 
+  left_join(select(sum_site, ta_canopy, site), by = "site") %>%
+  mutate(
+    bh_capped = ifelse(
+      ta_canopy < 2*sqrt(bs)/(22/7), ta_canopy, 2*sqrt(bs)/(22/7)), # if ta_canopy exceeds diameter of bite size, bite height is capped at bite size
+    bv = bs*ta_canopy,
+    bv_capped = bs*bh_capped,
+    vg = fr_pers*bv
+ )
+
+# grazing impact of fish from behavioral surveys
+sum_id_vg <- sum_id %>%
+  filter(species_code != "rbp") %>%
+  mutate(
+    bs = ifelse(species_code == "stop", 5.257*10^-4*length_cm^2, 4.013*10^-4*length_cm^2), # adding bite size equations from Mumby
+    bh_capped = ifelse(ta_canopy < 2*sqrt(bs)/(22/7), ta_canopy, 2*sqrt(bs)/(22/7)), # if ta_canopy exceeds diameter of bite size, bite height is capped at bite size
+    bv = bs*ta_canopy,
+    bv_capped = bs*bh_capped,
+    vg = fr*bv,
+    vg_capped = fr*bv_capped,
+    island = factor(island, levels = c("Bonaire","Antigua","Barbuda"))
+) %>%
+  arrange(island)
+test <- sum_id_vg %>% filter(site=="Pillars") %>% select(species, phase, length_cm, fr, ta_canopy, vg)
+
+# grazing impact averaged across individuals at each site (selecting for size 15-30cm)
+sum_ssp_vg <- sum_id_vg %>% 
+  filter(length_cm >= 15 & length_cm <= 30) %>%
+  group_by(site, island, species, species_code, phase) %>% 
+  summarize(
+    n=n(),
+    vg_mean=mean(vg),
+    vg_se=sd(vg)/sqrt(n),
+    vg_capped_mean=mean(vg_capped),
+    vg_capped_se=sd(vg_capped)/sqrt(n)
+    ) %>% 
+  select(-n)
+
+ggplot(filter(sum_ssp_vg, species_code != "rbp" & phase == "i"), aes(x=factor(site, levels = sum_site$site), y=vg_mean)) +
+    geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
+    geom_errorbar(aes(ymin=vg_mean-vg_se, ymax=vg_mean+vg_se), width=.2,
+                 position=position_dodge(.9)) +
+    facet_grid(~species) +
+    scale_fill_brewer(palette = "Blues") +
+    labs(title="Mean grazing impact by volume (initial phase)", 
+         y="Grazing impact (cm3/hr)", 
+         x="Site") +
+    theme_bw() +
+    theme(strip.text = element_text(face = "italic"), 
+          axis.text.x = element_text(angle = 90, hjust = 1), 
+          plot.title = element_text(hjust = 0.5))
+
+ggplot(filter(sum_ssp_vg, species_code != "rbp" & phase == "i"), aes(x=factor(site, levels = sum_site$site), y=vg_capped_mean)) +
+    geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
+    geom_errorbar(aes(ymin=vg_capped_mean-vg_capped_se, ymax=vg_capped_mean+vg_capped_se), width=.2,
+                 position=position_dodge(.9)) +
+    facet_grid(~species) +
+    scale_fill_brewer(palette = "Blues") +
+    labs(title="Mean grazing impact by volume (capped height, initial phase)", 
+         y="Grazing impact (cm3/hr)", 
+         x="Site") +
+    theme_bw() +
+    theme(strip.text = element_text(face = "italic"), 
+          axis.text.x = element_text(angle = 90, hjust = 1), 
+          plot.title = element_text(hjust = 0.5))
+```

### 2019-04-22 r/Ch2 outline_V3.Rmd
adding size range for sum_ssp_vg

--- 2019-04-16 r/Ch2 outline_V3.Rmd
@@
 sum_ssp_vg <- sum_id_vg %>% 
+  filter(length_cm >= 15 & length_cm <= 30) %>%
   group_by(site, ta_canopy, island, species, species_code, phase) %>% 

### 2019-04-22 r/Follows.Rmd

@@
     int_dam_se=sd(int_dam)/sqrt(n)
-    ) %>% 
-  select(-n)
+    )
 
@@
     length_se = sd(length_cm)/sqrt(n)
-    ) %>% 
-  select(-n)
+    )
+
+test <- sum_id %>% filter(site=="Pallaster Mid")
 ```

### 2019-04-23 r/Ch2 outline_V3.Rmd
4.22 updates

--- 2019-04-19 r/Ch2 outline_V3.Rmd
@@
   nrow=2, heights = c(1,1.2))
+
+g <- grid.arrange(left = "Feeding rate (bites/min.)",
+  arrangeGrob(fr_q1, fr_q3, fr_q4, fr_q5, fr_q7, ncol=5, right = label_vet),
+  arrangeGrob(fr_s1, fr_s3, fr_s4, fr_s5, fr_s7, ncol=5, right = label_vir), 
+  nrow=2, heights = c(1,1.2))
+
+ggsave(here("figs/biplot_panel.png"),g)
 ```

### 2019-04-23 r/Ch2 outline_V3.Rmd

@@
 
+# keep this to get legend for panel
 vg_facet <- ggplot(filter(sum_ssp_vg, species_code != "rbp" & phase == "i"), aes(x=factor(site, levels = sum_site$site), y=vg_mean)) +
@@
               ncol=2, widths = c(5,1), heights = c(1))
+
+g <- grid.arrange(
+  arrangeGrob(vg_qup, vg_stop, ncol=2), 
+              arrangeGrob(legend), 
+              ncol=2, widths = c(5,1), heights = c(1))
+
+ggsave(here("figs/bc_bar_panel.png"),g)
+```
+
+```{r}
+# geom_point version
+# keep this to get legend for panel
+vg_facet <- ggplot(filter(sum_ssp_vg, species_code != "rbp" & phase == "i"), aes(x=factor(site, levels = sum_site$site), y=vg_mean)) +
+    geom_point(stat = "identity", aes(color = fr_mean, size = ta_canopy), position = "dodge") +
+    geom_errorbar(aes(ymin=vg_mean-vg_se, ymax=vg_mean+vg_se), width=.2,
+                 position=position_dodge(.9)) +
+    facet_grid(~species) +
+    labs(title="Mean grazing impact by volume (initial phase)", 
+         y="Grazing impact (cm3/hr)", 
+         x="Site") +
+    theme_bw() +
+    theme(strip.text = element_text(face = "italic"), 
+          axis.text.x = element_text(angle = 45, hjust = 1)
+          )
+
+vg_qup <- ggplot(filter(sum_ssp_vg, species_code == "qup" & phase == "i"), 
+                 aes(x=factor(site, levels = sum_site$site), y=vg_mean)) +
+    geom_point(stat = "identity", aes(color = fr_mean, size = ta_canopy), position = "dodge") +
+    geom_errorbar(aes(ymin=vg_mean-vg_se, ymax=vg_mean+vg_se), width=.2,
+                 position=position_dodge(.9)) +
+  labs(y = expression('Grazing impact (cm'^3*'/hr)'), 
+       x = "", 
+       title = "S. vetula") +
+  theme_bw() + 
+  theme(plot.margin = margin(1, 0, 0, 5, unit = "pt"),
+        axis.title.y = element_text(size=10),
+        axis.text.x = element_text(angle = 45, hjust = 1, size=7),
+        legend.position = "none",
+        plot.title = element_text(hjust = 0.5, face = "italic")
+        )
+
+vg_stop <- ggplot(filter(sum_ssp_vg, species_code == "stop" & phase == "i"), 
+                  aes(x=factor(site, levels = sum_site$site), y=vg_mean)) +
+    geom_point(stat = "identity", aes(color = fr_mean, size = ta_canopy), position = "dodge") +    
+    geom_errorbar(aes(ymin=vg_mean-vg_se, ymax=vg_mean+vg_se), width=.2,
+                 position=position_dodge(.9)) +
+  labs(y = "",
+       x = "", 
+       title = "S. viride") +
+  theme_bw() + 
+  theme(plot.margin = margin(1, 1, 0, 0, unit = "pt"),
+        axis.text.x = element_text(angle = 45, hjust = 1, size=7),
+        legend.position = "none",
+        plot.title = element_text(hjust = 0.5, face = "italic")
+        )
+
+get_legend<-function(myggplot){
+  tmp <- ggplot_gtable(ggplot_build(myggplot))
+  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
+  legend <- tmp$grobs[[leg]]
+  return(legend)
+  }
+legend <- get_legend(vg_facet)
+
+grid.arrange(
+  arrangeGrob(vg_qup, vg_stop, ncol=2), 
+              arrangeGrob(legend), 
+              ncol=2, widths = c(5,1), heights = c(1))
+
+g <- grid.arrange(
+  arrangeGrob(vg_qup, vg_stop, ncol=2), 
+              arrangeGrob(legend), 
+              ncol=2, widths = c(5,1), heights = c(1))
+
+ggsave(here("figs/bc_point_panel.png"),g)
 ```

### 2019-04-24 r/Analysis.Rmd
4.23

@@
 
+Testing for variance in behavior variables among sites
+```{r}
+
+sum_id_stop <- sum_id %>%
+  filter(species_code=="stop" & length_cm > 15 & length_cm < 30 & phase == "i") %>%
+  mutate(site = as.factor(site))
+sum_id_qup <- sum_id %>%
+  filter(species_code=="qup" & length_cm > 15 & length_cm < 30 & phase == "i" ) %>%
+  mutate(site = as.factor(site))
+
+# feeding rates
+leveneTest(fr ~ site, data = sum_id_stop) # variances are slightly heteroskedastic, p ~ 0.04, use KW instead of ANOVA
+kruskal.test(fr ~ site, data = sum_id_stop) # significant variation across groups
+
+leveneTest(fr ~ site, data = sum_id_qup) # variances are homogenous, p > 0.05, can use ANOVA
+aov_fr_qup <- aov(fr ~ site, data = sum_id_qup) 
+summary.aov(aov_fr_qup) # significant variation across sites
+
+# bite rates
+leveneTest(br ~ site, data = sum_id_stop) # variances are heteroskedastic, p < 0.05
+kruskal.test(br ~ site, data = sum_id_stop) # no significant variation across groups
+
+leveneTest(br ~ site, data = sum_id_qup) # variances are are heteroskedastic, p < 0.05
+kruskal.test(br ~ site, data = sum_id_qup) # no significant variation
+
+# time spent feeding
+leveneTest(g_frac ~ site, data = sum_id_stop) # variances are homogenous, p > 0.05
+aov_g_frac_qup <- aov(g_frac ~ site, data = sum_id_stop) 
+summary.aov(aov_g_frac_qup) # significant variation across sites
+
+leveneTest(g_frac ~ site, data = sum_id_qup) # variances are homogenous, p > 0.05
+aov_g_frac_qup <- aov(g_frac ~ site, data = sum_id_qup) 
+summary.aov(aov_g_frac_qup) # significant variation across sites
+
+# foray length
+leveneTest(for_bites ~ site, data = sum_id_stop) # variances are homogenous, p > 0.05
+aov_for_bites_stop <- aov(for_bites ~ site, data = sum_id_stop) 
+summary.aov(aov_for_bites_stop) # significant variation across sites
+
+leveneTest(for_bites ~ site, data = sum_id_qup) # variances are homogenous, p > 0.05
+aov_for_bites_qup <- aov(for_bites ~ site, data = sum_id_qup) 
+summary.aov(aov_for_bites_qup) # significant variation across sites
+
+```
+
 # Bite content hypothesis
@@
 ```
+
+```{r}
+# geom_point version
+# keep this to get legend for panel
+vg_facet <- ggplot(filter(sum_ssp_vg, species_code != "rbp" & phase == "i"), aes(x=factor(site, levels = sum_site$site), y=vg_mean)) +
+    geom_point(stat = "identity", aes(color = fr_mean, size = ta_canopy), position = "dodge") +
+    geom_errorbar(aes(ymin=vg_mean-vg_se, ymax=vg_mean+vg_se), width=.2,
+                 position=position_dodge(.9)) +
+    facet_grid(~species) +
+    labs(title="Mean grazing impact by volume (initial phase)", 
+         y="Grazing impact (cm3/hr)", 
+         x="Site") +
+    theme_bw() +
+    theme(strip.text = element_text(face = "italic"), 
+          axis.text.x = element_text(angle = 45, hjust = 1)
+          )
+
+vg_qup <- ggplot(filter(sum_ssp_vg, species_code == "qup" & phase == "i"), 
+                 aes(x=factor(site, levels = sum_site$site), y=vg_mean)) +
+    geom_point(stat = "identity", aes(color = fr_mean, size = ta_canopy), position = "dodge") +
+    geom_errorbar(aes(ymin=vg_mean-vg_se, ymax=vg_mean+vg_se), width=.2,
+                 position=position_dodge(.9)) +
+  labs(y = expression('Grazing impact (cm'^3*'/hr)'), 
+       x = "", 
+       title = "S. vetula") +
+  theme_bw() + 
+  theme(plot.margin = margin(1, 0, 0, 5, unit = "pt"),
+        axis.title.y = element_text(size=10),
+        axis.text.x = element_text(angle = 45, hjust = 1, size=7),
+        legend.position = "none",
+        plot.title = element_text(hjust = 0.5, face = "italic")
+        )
+
+vg_stop <- ggplot(filter(sum_ssp_vg, species_code == "stop" & phase == "i"), 
+                  aes(x=factor(site, levels = sum_site$site), y=vg_mean)) +
+    geom_point(stat = "identity", aes(color = fr_mean, size = ta_canopy), position = "dodge") +    
+    geom_errorbar(aes(ymin=vg_mean-vg_se, ymax=vg_mean+vg_se), width=.2,
+                 position=position_dodge(.9)) +
+  labs(y = "",
+       x = "", 
+       title = "S. viride") +
+  theme_bw() + 
+  theme(plot.margin = margin(1, 1, 0, 0, unit = "pt"),
+        axis.text.x = element_text(angle = 45, hjust = 1, size=7),
+        legend.position = "none",
+        plot.title = element_text(hjust = 0.5, face = "italic")
+        )
+
+get_legend<-function(myggplot){
+  tmp <- ggplot_gtable(ggplot_build(myggplot))
+  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
+  legend <- tmp$grobs[[leg]]
+  return(legend)
+  }
+legend <- get_legend(vg_facet)
+
+grid.arrange(
+  arrangeGrob(vg_qup, vg_stop, ncol=2), 
+              arrangeGrob(legend), 
+              ncol=2, widths = c(5,1), heights = c(1))
+
+g <- grid.arrange(
+  arrangeGrob(vg_qup, vg_stop, ncol=2), 
+              arrangeGrob(legend), 
+              ncol=2, widths = c(5,1), heights = c(1))
+
+ggsave(here("figs/bc_point_panel.png"),g)
+```

### 2019-04-24 r/Ch2 outline_V3.Rmd
4.23

@@
 
-ref_qup15 <- (3329-33*15)/60 # published feeding rate (bites/min)
+ # published feeding rate (bites/min) for reference lines if needed
+ref_qup15 <- (3329-33*15)/60
 ref_qup30 <- (3329-33*30)/60
@@
 
-```{r}
-# geom_point version
-# keep this to get legend for panel
-vg_facet <- ggplot(filter(sum_ssp_vg, species_code != "rbp" & phase == "i"), aes(x=factor(site, levels = sum_site$site), y=vg_mean)) +
-    geom_point(stat = "identity", aes(color = fr_mean, size = ta_canopy), position = "dodge") +
-    geom_errorbar(aes(ymin=vg_mean-vg_se, ymax=vg_mean+vg_se), width=.2,
-                 position=position_dodge(.9)) +
-    facet_grid(~species) +
-    labs(title="Mean grazing impact by volume (initial phase)", 
-         y="Grazing impact (cm3/hr)", 
-         x="Site") +
-    theme_bw() +
-    theme(strip.text = element_text(face = "italic"), 
-          axis.text.x = element_text(angle = 45, hjust = 1)
-          )
-
-vg_qup <- ggplot(filter(sum_ssp_vg, species_code == "qup" & phase == "i"), 
-                 aes(x=factor(site, levels = sum_site$site), y=vg_mean)) +
-    geom_point(stat = "identity", aes(color = fr_mean, size = ta_canopy), position = "dodge") +
-    geom_errorbar(aes(ymin=vg_mean-vg_se, ymax=vg_mean+vg_se), width=.2,
-                 position=position_dodge(.9)) +
-  labs(y = expression('Grazing impact (cm'^3*'/hr)'), 
-       x = "", 
-       title = "S. vetula") +
-  theme_bw() + 
-  theme(plot.margin = margin(1, 0, 0, 5, unit = "pt"),
-        axis.title.y = element_text(size=10),
-        axis.text.x = element_text(angle = 45, hjust = 1, size=7),
-        legend.position = "none",
-        plot.title = element_text(hjust = 0.5, face = "italic")
-        )
-
-vg_stop <- ggplot(filter(sum_ssp_vg, species_code == "stop" & phase == "i"), 
-                  aes(x=factor(site, levels = sum_site$site), y=vg_mean)) +
-    geom_point(stat = "identity", aes(color = fr_mean, size = ta_canopy), position = "dodge") +    
-    geom_errorbar(aes(ymin=vg_mean-vg_se, ymax=vg_mean+vg_se), width=.2,
-                 position=position_dodge(.9)) +
-  labs(y = "",
-       x = "", 
-       title = "S. viride") +
-  theme_bw() + 
-  theme(plot.margin = margin(1, 1, 0, 0, unit = "pt"),
-        axis.text.x = element_text(angle = 45, hjust = 1, size=7),
-        legend.position = "none",
-        plot.title = element_text(hjust = 0.5, face = "italic")
-        )
-
-get_legend<-function(myggplot){
-  tmp <- ggplot_gtable(ggplot_build(myggplot))
-  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
-  legend <- tmp$grobs[[leg]]
-  return(legend)
-  }
-legend <- get_legend(vg_facet)
-
-grid.arrange(
-  arrangeGrob(vg_qup, vg_stop, ncol=2), 
-              arrangeGrob(legend), 
-              ncol=2, widths = c(5,1), heights = c(1))
-
-g <- grid.arrange(
-  arrangeGrob(vg_qup, vg_stop, ncol=2), 
-              arrangeGrob(legend), 
-              ncol=2, widths = c(5,1), heights = c(1))
-
-ggsave(here("figs/bc_point_panel.png"),g)
-```
-
 ```{r, eval = F}

### 2019-04-24 r/Analysis.Rmd
4/24

@@
 leveneTest(fr ~ site, data = sum_id_stop) # variances are slightly heteroskedastic, p ~ 0.04, use KW instead of ANOVA
-kruskal.test(fr ~ site, data = sum_id_stop) # significant variation across groups
+kruskal.test(fr ~ site, data = sum_id_stop) # significant variation across sites, p < 0.001
 
@@
 aov_fr_qup <- aov(fr ~ site, data = sum_id_qup) 
-summary.aov(aov_fr_qup) # significant variation across sites
+summary.aov(aov_fr_qup) # significant variation across sites, p < 0.001
 
@@
 leveneTest(br ~ site, data = sum_id_stop) # variances are heteroskedastic, p < 0.05
-kruskal.test(br ~ site, data = sum_id_stop) # no significant variation across groups
+kruskal.test(br ~ site, data = sum_id_stop) # no significant variation, p > 0.5
 
 leveneTest(br ~ site, data = sum_id_qup) # variances are are heteroskedastic, p < 0.05
-kruskal.test(br ~ site, data = sum_id_qup) # no significant variation
+kruskal.test(br ~ site, data = sum_id_qup) # no significant variation, p > 0.5
 
@@
 aov_g_frac_qup <- aov(g_frac ~ site, data = sum_id_stop) 
-summary.aov(aov_g_frac_qup) # significant variation across sites
+summary.aov(aov_g_frac_qup) # significant variation across sites, p < 0.001
 
@@
 aov_g_frac_qup <- aov(g_frac ~ site, data = sum_id_qup) 
-summary.aov(aov_g_frac_qup) # significant variation across sites
+summary.aov(aov_g_frac_qup) # significant variation across sites, p < 0.001
 
@@
 aov_for_bites_stop <- aov(for_bites ~ site, data = sum_id_stop) 
-summary.aov(aov_for_bites_stop) # significant variation across sites
+summary.aov(aov_for_bites_stop) # significant variation across sites, p < 0.001
 
@@
 aov_for_bites_qup <- aov(for_bites ~ site, data = sum_id_qup) 
-summary.aov(aov_for_bites_qup) # significant variation across sites
+summary.aov(aov_for_bites_qup) # significant variation across sites, p < 0.001
 

### 2019-04-24 r/Ch2 outline_V3.Rmd
4/24

@@
 
-grid.arrange(left = "Feeding rate (bites/min.)",
-  arrangeGrob(fr_q1, fr_q3, fr_q4, fr_q5, fr_q7, ncol=5, right = label_vet),
-  arrangeGrob(fr_s1, fr_s3, fr_s4, fr_s5, fr_s7, ncol=5, right = label_vir), 
-  nrow=2, heights = c(1,1.2))
-
 g <- grid.arrange(left = "Feeding rate (bites/min.)",
@@
 
-grid.arrange(
-  arrangeGrob(vg_qup, vg_stop, ncol=2), 
-              arrangeGrob(legend), 
-              ncol=2, widths = c(5,1), heights = c(1))
-
 g <- grid.arrange(
@@
 
-```{r, eval = F}
+```{r, eval = F, echo = F, message = F}
 ggplot(filter(sum_ssp_vg, species_code != "rbp" & phase == "i"), aes(x = fr_mean, y = vg_mean)) +
@@
 
-autoplot(pca_pred, data = sum_site, colour = 'island',
+autoplot(pca_pred, data = sum_site, shape = 'island', colour = 'island',
          loadings = T, loadings.colour = 'gray',
@@
             loadings.label.repel = T) +
-    scale_color_brewer(palette = "Blues") +
+    scale_color_manual(values = c("cadetblue3","cadetblue3","cadetblue3")) +
     theme_bw() +
     theme(legend.title = element_blank())
-ggsave(here("figs/pca_pred.png"))
+ggsave(here("figs/pca_pred.png"), width = 8, height = 5)
 
@@
 kable(gtab_fr[1:4, ])
-plot(gamm3$gam,pages=1)
+plot(gamm10$gam,pages=1)
 ```

### 2019-05-08 r/Ch2 outline_V3.Rmd
5/8

@@
 
-gamm1 <- gamm(fr_mean ~ species_code + phase + s(length_m, k=4) + s(pc1_all, k=4) + s(pc2_all, k=4), 
+gamm1 <- gamm(fr_mean ~ species_code + phase + s(length_m, k=5) + s(pc1_all, k=5) + s(pc2_all, k=5), 
            family = gaussian(link=identity),
@@
 
-gamm2 <- gamm(fr_mean ~ species_code + phase + s(length_m, k=4) + s(pc1_all, k=4), 
+gamm2 <- gamm(fr_mean ~ species_code + phase + s(length_m, k=5) + s(pc1_all, k=5), 
            family = gaussian(link=identity),
@@
 
-gamm3 <- gamm(fr_mean ~ species_code + phase + s(length_m, by = factor(species_code), k=4) + s(pc1_all, by = factor(species_code), k=4) + s(pc2_all, by = factor(species_code), k=4), 
+gamm3 <- gamm(fr_mean ~ species_code + phase + s(length_m, by = factor(species_code), k=5) + s(pc1_all, by = factor(species_code), k=5) + s(pc2_all, by = factor(species_code), k=5), 
            family = gaussian(link=identity),
@@
 
-gamm4 <- gamm(fr_mean ~ species_code + phase + s(length_m, by = factor(species_code), k=4) + s(pc1_all, by = factor(species_code), k=4) + s(pc2_all, k=4), 
+gamm4 <- gamm(fr_mean ~ species_code + phase + s(length_m, by = factor(species_code), k=5) + s(pc1_all, by = factor(species_code), k=5) + s(pc2_all, k=5), 
            family = gaussian(link=identity),
@@
 
-gamm5 <- gamm(fr_mean ~ species + phase + s(length_m, by = factor(species_code), k=4) + s(pc1_all, by = factor(species_code), k=4), 
+gamm5 <- gamm(fr_mean ~ species + phase + s(length_m, by = factor(species_code), k=5) + s(pc1_all, by = factor(species_code), k=5), 
            family = gaussian(link=identity),
@@
 
-gamm6 <- gamm(fr_mean ~ phase + s(length_m, k=4) + s(pc1_all, by = factor(species_code), k=4), 
+gamm6 <- gamm(fr_mean ~ phase + s(length_m, k=5) + s(pc1_all, by = factor(species_code), k=5), 
            family = gaussian(link=identity),
@@
 
-gamm7 <- gamm(fr_mean ~ phase + s(length_m, by = factor(species_code), k=4) + s(pc1_all, k=4), 
+gamm7 <- gamm(fr_mean ~ phase + s(length_m, by = factor(species_code), k=5) + s(pc1_all, k=5), 
            family = gaussian(link=identity),
@@
 
-gamm8 <- gamm(fr_mean ~ species + phase + s(length_m, by = factor(species_code), k=4) + s(pc1_all, k=4), 
+gamm8 <- gamm(fr_mean ~ species + phase + s(length_m, by = factor(species_code), k=5) + s(pc1_all, k=5), 
            family = gaussian(link=identity),
@@
 
-gamm9 <- gamm(fr_mean ~ species + phase + s(length_m, k=4) + s(pc1_all, by = factor(species_code), k=4), 
+gamm9 <- gamm(fr_mean ~ species + phase + s(length_m, k=5) + s(pc1_all, by = factor(species_code), k=5), 
            family = gaussian(link=identity),
@@
 
-gamm10 <- gamm(fr_mean ~ species + phase + s(pc1_all, by = factor(species_code), k=4), 
+gamm10 <- gamm(fr_mean ~ species + phase + s(pc1_all, by = factor(species_code), k=5), 
            family = gaussian(link=identity),
@@
 
-gamm11 <- gamm(fr_mean ~ phase + s(length_m, by = factor(species_code), k=4) + s(pc1_all, by = factor(species_code), k=4), 
+gamm11 <- gamm(fr_mean ~ phase + s(length_m, by = factor(species_code), k=5) + s(pc1_all, by = factor(species_code), k=5), 
            family = gaussian(link=identity),
@@
 plot(gamm10$gam,pages=1)
+```
+
+With initial phase individuals only:
+```{r, echo=F, message=F, warning=F}
+m_data <- sum_ssp_pca %>% filter(species_code!="rbp")
+
+gamm1 <- gamm(fr_mean ~ species_code + s(length_m, k=5) + s(pc1_all, k=5) + s(pc2_all, k=5), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           method = "REML",
+           data = m_data
+           )
+g1 <- c("g1", "Species + Size + PC1 + PC2", AICc(gamm1$lme), summary(gamm1$gam)$r.sq)
+
+gamm2 <- gamm(fr_mean ~ species_code + s(length_m, k=5) + s(pc1_all, k=5), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           method = "REML",
+           data = m_data
+           )
+g2 <- c("g2", "Species + Size + PC1", AICc(gamm2$lme), summary(gamm2$gam)$r.sq)
+
+gamm3 <- gamm(fr_mean ~ species_code + s(length_m, by = factor(species_code), k=5) + s(pc1_all, by = factor(species_code), k=5) + s(pc2_all, by = factor(species_code), k=5),
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           method = "REML",
+           data = m_data
+           )
+g3 <- c("g3", "Species + Size~Species + PC1~Species + PC2~Species", AICc(gamm3$lme), summary(gamm3$gam)$r.sq)
+
+gamm4 <- gamm(fr_mean ~ species_code + s(length_m, by = factor(species_code), k=5) + s(pc1_all, by = factor(species_code), k=5) + s(pc2_all, k=5), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           method = "REML",
+           data = m_data
+           )
+g4 <- c("g4", "Species + Size~Species + PC1~Species + PC2", AICc(gamm4$lme), summary(gamm4$gam)$r.sq)
+
+gamm5 <- gamm(fr_mean ~ species + s(length_m, by = factor(species_code), k=5) + s(pc1_all, by = factor(species_code), k=5), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           method = "REML",
+           data = m_data
+           )
+g5 <- c("g5", "Species + Size~Species + PC1~Species", AICc(gamm5$lme), summary(gamm5$gam)$r.sq)
+
+gamm6 <- gamm(fr_mean ~ s(length_m, k=5) + s(pc1_all, by = factor(species_code), k=5), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           method = "REML",
+           data = m_data
+           )
+g6 <- c("g6", "Size + PC1~Species", AICc(gamm6$lme), summary(gamm6$gam)$r.sq)
+
+gamm7 <- gamm(fr_mean ~ s(length_m, by = factor(species_code), k=5) + s(pc1_all, k=5), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           method = "REML",
+           data = m_data
+           )
+g7 <- c("g7", "Size~Species + PC1", AICc(gamm7$lme), summary(gamm7$gam)$r.sq)
+
+gamm8 <- gamm(fr_mean ~ species + s(length_m, by = factor(species_code), k=5) + s(pc1_all, k=5), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           method = "REML",
+           data = m_data
+           )
+g8 <- c("g8", "Species + Size~Species + PC1", AICc(gamm8$lme), summary(gamm8$gam)$r.sq)
+
+gamm9 <- gamm(fr_mean ~ species + s(length_m, k=5) + s(pc1_all, by = factor(species_code), k=5), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           method = "REML",
+           data = m_data
+           )
+g9 <- c("g9", "Species + Size + PC1~Species", AICc(gamm9$lme), summary(gamm9$gam)$r.sq)
+
+gamm10 <- gamm(fr_mean ~ species + s(pc1_all, by = factor(species_code), k=5), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           method = "REML",
+           data = m_data
+           )
+g10 <- c("g10", "Species + PC1~Species", AICc(gamm10$lme), summary(gamm10$gam)$r.sq)
+
+gamm11 <- gamm(fr_mean ~ s(length_m, by = factor(species_code), k=5) + s(pc1_all, by = factor(species_code), k=5), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           method = "REML",
+           data = m_data
+           )
+g11 <- c("g11", "Size~Species + PC1~Species", AICc(gamm11$lme), summary(gamm11$gam)$r.sq)
+
+gtab_fr <- as.data.frame(rbind(g1,g2,g4,g5,g6,g7,g8,g9,g10,g11)) %>% setNames(c("g","Model", "AICc", "Adj. R2")) %>% arrange(AICc)
 ```

### 2019-05-08 r/Follows.Rmd
5/8

--- 2019-03-01 r/Follows.Rmd
@@
   select(-X,-family) %>%
-  rename_at(vars(3:8), ~ paste0("consp_",.))
+  rename_at(vars(3:8), ~ paste0("consp_",.)) # for adding conspecific biomass/density data
 
@@
   filter(
-    !(activity == "graze" & min == 0) | # filters forays starting at time=0
+    !(activity == "graze" & min == 0) | # filters out forays starting at time=0
     !(lead(activity) == "end") # filters forays overlapping with end of follow
@@
 # Summarizing data by site, species, and phase
-Next I create a summary tables that averages variables within each site/species/phase, then combines into one table
+Next I create a summary tables that average variables within each site/species/phase, then combines into one table
 ```{r,warning=FALSE, message= FALSE}
@@
 # site averages by species
-sum_site_sp <- sum_id %>% 
+sum_ss <- sum_id %>% 
+  filter(length_cm >= 15 & length_cm <= 30) %>%
   group_by(site, island, species, species_code) %>% 
@@
     int_dam_se=sd(int_dam)/sqrt(n)
-    ) %>% 
-  select(-n)
+    )
 
 # site averages by species and phase
-sum_site_sp_ph <- sum_id %>% 
+sum_ssp <- sum_id %>% 
+  filter(length_cm >= 15 & length_cm <= 30) %>%
   group_by(site, island, species, species_code, phase) %>% 
@@
     length_se = sd(length_cm)/sqrt(n)
-    ) %>% 
-  select(-n)
+    )
 ```
@@
   left_join(sp_site, by=c("species","site")) # add conspecific scarid info (based on site/species)
-sum_site_sp <- sum_site_sp %>% 
+sum_ss <- sum_ss %>% 
   left_join(site, by=c("island","site")) %>%
   left_join(sp_site, by=c("species","site")) # add conspecific scarid info (based on site/species)
-sum_site_sp_ph <- sum_site_sp_ph %>% 
+sum_ssp <- sum_ssp %>% 
   left_join(site, by=c("island","site")) %>%
@@
 write.csv(sum_id,file=here("data/sum_id.csv"))
-write.csv(sum_site_sp,file=here("data/sum_site_sp.csv"))
-write.csv(sum_site_sp_ph,file=here("data/sum_site_sp_ph.csv"))
+write.csv(sum_ss,file=here("data/sum_ss.csv"))
+write.csv(sum_ssp,file=here("data/sum_ssp.csv"))
 ```

### 2019-05-08 r/Models.Rmd
feb1 updates
5/8

@@
   select(-X)
-sum_ssp <- read.csv(here("data","sum_site_sp_ph.csv"), stringsAsFactors=F) %>%
+sum_ssp <- read.csv(here("data","sum_ssp.csv"), stringsAsFactors=F) %>%
   select(-X)
@@
               # filter(island!="Barbuda") %>%
-              select(ma_cover, ma_canopy, ta_cover, ta_canopy, rugosity)),
+              select(ma_cover, ma_canopy, ta_cover, ta_canopy, rugosity, lc_cover)),
   center = T,scale = T)
@@
 
-gamtabs(gamm4, type="HTML")
+# gamtabs(gamm4, type="HTML")
 ```

### 2019-05-08 r/Sites.Rmd
5/8

@@
 write.csv(site,file=here("data/site.csv"))
-write.csv(sp_site,file=here("data/sp_site.csv"))
+write.csv(sp_site,file=here("data/sp_site.csv")) # for later calculating conspecifics
 write.csv(fish,file=here("data/fish.csv")) # for predicting grazing impact of observed fish communities at each site (see Analysis.rmd)

### 2019-05-08 r/old drafts/Data_prodding.Rmd
5/8

--- 2019-02-01 r/Data_prodding.Rmd
@@
   filter(length_cm>=15)
-sum_site_sp_ph <- read.csv(here("data","sum_site_sp_ph.csv"), stringsAsFactors=F) %>%
+sum_ssp <- read.csv(here("data","sum_ssp.csv"), stringsAsFactors=F) %>%
   select(-X) 
@@
 Need to determine appropriate size range for comparison.
-Because they are not evenly distributed (i.e.
+Because they are not quite evenly distributed (i.e.
 larger fish in Bonaire, smaller in Barbuda), I will likely want to compare length-feeding relationships as opposed to pooled averages
@@
 pred <- sum_site %>% 
-  select(tot_bm,tot_den,scar_bm,scar_den,carn30_bm,carn30_den,
+  select(tot_bm,tot_den,scar_bm,scar_den,carn_bm,carn_den,vir_bm,vet_bm,carn30_bm,carn30_den,
          cca_cover,lc_cover,ma_cover,ma_canopy,ta_cover,ta_canopy,rugosity,spear_daily) %>% 

### 2019-05-08 r/5.8.19.Rmd
5.8 modelling tests

---
title: "GAMM exploration"
author: "Molly Wilson"
date: "5/8/2019"
output: html_document
---

Testing different GAMM iterations to confirm analyses.
Tasks at hand:
- site level analyses vs. individual fish level?
- include phases, or initial phase only?
- add site as random effect (in addition to island?)

```{r, echo=F, message=F, warning=F}
library(here)
library(tidyverse)
library(mgcv)
library(MuMIn)

sum_site <- read.csv(here("data","sum_site.csv"), stringsAsFactors=F) %>%
  select(-X) 
sum_id <- read.csv(here("data","sum_id.csv"), stringsAsFactors=F) %>%
  select(-X) %>% 
  filter(fr <= 3000)
sum_ssp <- read.csv(here("data","sum_ssp.csv"), stringsAsFactors=F) %>%
  select(-X) 
```

Assemble PC1 and PC2 variables
```{r, echo=F, message=F, warning=F}
pca_pred <- prcomp(na.omit(sum_site %>% 
              select(rugosity, ma_cover, ma_canopy, ta_cover, ta_canopy, lc_cover, scar_den, scar_bm, carn30_bm)),
  center = T, scale = T)

sum_site <- sum_site %>% 
  bind_cols(as.data.frame(pca_pred$x) %>% 
           select(PC1,PC2)) %>%
  rename(pc1_all=PC1,
         pc2_all=PC2)

sum_id_pca <- sum_id %>% 
  left_join(select(sum_site,site,pc1_all,pc2_all), by = "site")
sum_ssp_pca <- sum_ssp %>%
  left_join(select(sum_site,site,pc1_all,pc2_all), by = "site")
```

## Site - vs. ID - level GAMMs

ID - level, initial phase only
```{r, echo=F, message=F, warning=F}
m_data <- sum_id_pca %>% filter(species_code!="rbp" & phase == "i")

gamm1 <- gamm(fr ~ species_code + s(length_cm, k=5) + s(pc1_all, k=5) + s(pc2_all, k=5), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
g1 <- c("g1", "Species + Size + PC1 + PC2", AIC(gamm1$lme), summary(gamm1$gam)$r.sq)

gamm2 <- gamm(fr ~ species_code + s(length_cm, k=5) + s(pc1_all, k=5), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
g2 <- c("g2", "Species + Size + PC1", AIC(gamm2$lme), summary(gamm2$gam)$r.sq)

gamm3 <- gamm(fr ~ species_code + s(length_cm, by = factor(species_code), k=5) + s(pc1_all, by = factor(species_code), k=5) + s(pc2_all, by = factor(species_code), k=5), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
g3 <- c("g3", "Species + Size~Species + PC1~Species + PC2~Species", AIC(gamm3$lme), summary(gamm3$gam)$r.sq)

gamm4 <- gamm(fr ~ species_code + s(length_cm, by = factor(species_code), k=5) + s(pc1_all, by = factor(species_code), k=5) + s(pc2_all, k=5), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
g4 <- c("g4", "Species + Size~Species + PC1~Species + PC2", AIC(gamm4$lme), summary(gamm4$gam)$r.sq)

gamm5 <- gamm(fr ~ species + s(length_cm, by = factor(species_code), k=5) + s(pc1_all, by = factor(species_code), k=5), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
g5 <- c("g5", "Species + Size~Species + PC1~Species", AIC(gamm5$lme), summary(gamm5$gam)$r.sq)

gamm6 <- gamm(fr ~  s(length_cm, k=5) + s(pc1_all, by = factor(species_code), k=5), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
g6 <- c("g6", "Size + PC1~Species", AIC(gamm6$lme), summary(gamm6$gam)$r.sq)

gamm7 <- gamm(fr ~ s(length_cm, by = factor(species_code), k=5) + s(pc1_all, k=5), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
g7 <- c("g7", "Size~Species + PC1", AIC(gamm7$lme), summary(gamm7$gam)$r.sq)

gamm8 <- gamm(fr ~ species + s(length_cm, by = factor(species_code), k=5) + s(pc1_all, k=5), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
g8 <- c("g8", "Species + Size~Species + PC1", AIC(gamm8$lme), summary(gamm8$gam)$r.sq)

gamm9 <- gamm(fr ~ species + s(length_cm, k=5) + s(pc1_all, by = factor(species_code), k=5), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
g9 <- c("g9", "Species + Size + PC1~Species", AIC(gamm9$lme), summary(gamm9$gam)$r.sq)

gamm10 <- gamm(fr ~ species + s(pc1_all, by = factor(species_code), k=5), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
g10 <- c("g10", "Species + PC1~Species", AIC(gamm10$lme), summary(gamm10$gam)$r.sq)

gamm11 <- gamm(fr ~ s(length_cm, by = factor(species_code), k=5) + s(pc1_all, by = factor(species_code), k=5), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
g11 <- c("g11", "Size~Species + PC1~Species", AIC(gamm11$lme), summary(gamm11$gam)$r.sq)

gtab_fr <- as.data.frame(rbind(g1,g2,g3,g4,g5,g6,g7,g8,g9,g10,g11)) %>% setNames(c("g","Model", "AIC", "Adj. R2")) %>% arrange(AIC)
```

Site - level GAMMs, initial phase only
```{r, echo=F, message=F, warning=F}
m_data <- sum_ssp_pca %>% filter(species_code!="rbp" & phase == "i")

gamm1 <- gamm(fr_mean ~ species_code + s(length_m, k=3) + s(pc1_all, k=3) + s(pc2_all, k=3), 
           family = gaussian(link=identity),
           random = list(island=~1),
           data = m_data
           )
g1 <- c("g1", "Species + Size + PC1 + PC2", AICc(gamm1$lme), summary(gamm1$gam)$r.sq)

gamm2 <- gamm(fr_mean ~ species_code + s(length_m, k=3) + s(pc1_all, k=3), 
           family = gaussian(link=identity),
           random = list(island=~1),
           data = m_data
           )
g2 <- c("g2", "Species + Size + PC1", AICc(gamm2$lme), summary(gamm2$gam)$r.sq)

gamm3 <- gamm(fr_mean ~ species_code + s(length_m, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3) + s(pc2_all, by = factor(species_code), k=3),
           family = gaussian(link=identity),
           random = list(island=~1),
           data = m_data
           )
g3 <- c("g3", "Species + Size~Species + PC1~Species + PC2~Species", AICc(gamm3$lme), summary(gamm3$gam)$r.sq)

gamm4 <- gamm(fr_mean ~ species_code + s(length_m, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3) + s(pc2_all, k=3), 
           family = gaussian(link=identity),
           random = list(island=~1),
           data = m_data
           )
g4 <- c("g4", "Species + Size~Species + PC1~Species + PC2", AICc(gamm4$lme), summary(gamm4$gam)$r.sq)

gamm5 <- gamm(fr_mean ~ species + s(length_m, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3), 
           family = gaussian(link=identity),
           random = list(island=~1),
           data = m_data
           )
g5 <- c("g5", "Species + Size~Species + PC1~Species", AICc(gamm5$lme), summary(gamm5$gam)$r.sq)

gamm6 <- gamm(fr_mean ~ s(length_m, k=3) + s(pc1_all, by = factor(species_code), k=3), 
           family = gaussian(link=identity),
           random = list(island=~1),
           data = m_data
           )
g6 <- c("g6", "Size + PC1~Species", AICc(gamm6$lme), summary(gamm6$gam)$r.sq)

gamm7 <- gamm(fr_mean ~ s(length_m, by = factor(species_code), k=3) + s(pc1_all, k=3), 
           family = gaussian(link=identity),
           random = list(island=~1),
           data = m_data
           )
g7 <- c("g7", "Size~Species + PC1", AICc(gamm7$lme), summary(gamm7$gam)$r.sq)

gamm8 <- gamm(fr_mean ~ species + s(length_m, by = factor(species_code), k=3) + s(pc1_all, k=3), 
           family = gaussian(link=identity),
           random = list(island=~1),
           data = m_data
           )
g8 <- c("g8", "Species + Size~Species + PC1", AICc(gamm8$lme), summary(gamm8$gam)$r.sq)

gamm9 <- gamm(fr_mean ~ species + s(length_m, k=3) + s(pc1_all, by = factor(species_code), k=3), 
           family = gaussian(link=identity),
           random = list(island=~1),
           data = m_data
           )
g9 <- c("g9", "Species + Size + PC1~Species", AICc(gamm9$lme), summary(gamm9$gam)$r.sq)

gamm10 <- gamm(fr_mean ~ species + s(pc1_all, by = factor(species_code), k=3), 
           family = gaussian(link=identity),
           random = list(island=~1),
           data = m_data
           )
g10 <- c("g10", "Species + PC1~Species", AICc(gamm10$lme), summary(gamm10$gam)$r.sq)

gamm11 <- gamm(fr_mean ~ s(length_m, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3), 
           family = gaussian(link=identity),
           random = list(island=~1),
           data = m_data
           )
g11 <- c("g11", "Size~Species + PC1~Species", AICc(gamm11$lme), summary(gamm11$gam)$r.sq)

gtab_fr <- as.data.frame(rbind(g1,g2,g4,g5,g6,g7,g8,g9,g10,g11)) %>% setNames(c("g","Model", "AICc", "Adj. R2")) %>% arrange(AICc)
```

### 2019-05-10 r/5.8.19.Rmd
5/10

@@
 - add site as random effect (in addition to island?)
+
+Notes after exploration:
+- Both site-level and id-level are 'possible', but still not sure which is most appropriate.
+Best models for id-level have many species interactions.
+I could split these models by species but obviously would lose a lot of R2.
+- Will stick with initial phase only
+- Site-level models for initial phase only cannot have 'method = "REML"'
+- Site as random effect doesn't really change much in terms of best models or R2....
 
@@
   center = T, scale = T)
+pca_bent <- prcomp(na.omit(sum_site %>% 
+              select(rugosity, ma_cover, ma_canopy, ta_cover, ta_canopy)),
+  center = T, scale = T)
+pca_fish <- prcomp(na.omit(sum_site %>% 
+              select(scar_bm, carn30_bm)),
+  center = T, scale = T)
 
-sum_site <- sum_site %>% 
+autoplot(pca_pred, data = sum_site, shape = 'island', colour = 'island',
+         loadings = T, loadings.colour = 'gray',
+         loadings.label = T, loadings.label.size = 3, loadings.label.colour = 'black',
+            loadings.label.repel = T) +
+    scale_color_manual(values = c("cadetblue3","cadetblue3","cadetblue3")) +
+    theme_bw() +
+    theme(legend.title = element_blank())
+ggsave(here("figs/pca_pred.png"), width = 8, height = 5)
+
+sum_site_pca <- sum_site %>% 
   bind_cols(as.data.frame(pca_pred$x) %>% 
@@
          pc2_all=PC2)
+sum_id_pca <- sum_id %>% 
+  left_join(select(sum_site_pca,site,pc1_all,pc2_all), by = "site")
+sum_ssp_pca <- sum_ssp %>%
+  left_join(select(sum_site_pca,site,pc1_all,pc2_all), by = "site")
 
-sum_id_pca <- sum_id %>% 
-  left_join(select(sum_site,site,pc1_all,pc2_all), by = "site")
-sum_ssp_pca <- sum_ssp %>%
-  left_join(select(sum_site,site,pc1_all,pc2_all), by = "site")
+var <- get_pca_var
 ```
@@
 
-gtab_fr <- as.data.frame(rbind(g1,g2,g3,g4,g5,g6,g7,g8,g9,g10,g11)) %>% setNames(c("g","Model", "AIC", "Adj. R2")) %>% arrange(AIC)
+gamm12 <- gamm(fr ~ species + s(pc1_all, k=5), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           method = "REML",
+           data = m_data
+           )
+g12 <- c("g11", "Species + PC1", AIC(gamm11$lme), summary(gamm11$gam)$r.sq)
+
+gtab_fr_id <- as.data.frame(rbind(g1,g2,g3,g4,g5,g6,g7,g8,g9,g10,g11,g12)) %>% setNames(c("g","Model", "AIC", "Adj. R2")) %>% arrange(AIC)
+
+# plots from best models
+plot(gamm3$gam,pages=1)
+plot(gamm5$gam,pages=1)
 ```
@@
 
-gamm1 <- gamm(fr_mean ~ species_code + s(length_m, k=3) + s(pc1_all, k=3) + s(pc2_all, k=3), 
+gamm1 <- gamm(fr_mean ~ species_code + s(length_mean, k=3) + s(pc1_all, k=3) + s(pc2_all, k=3), 
            family = gaussian(link=identity),
@@
 
-gamm2 <- gamm(fr_mean ~ species_code + s(length_m, k=3) + s(pc1_all, k=3), 
+gamm2 <- gamm(fr_mean ~ species_code + s(length_mean, k=3) + s(pc1_all, k=3), 
            family = gaussian(link=identity),
@@
 
-gamm3 <- gamm(fr_mean ~ species_code + s(length_m, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3) + s(pc2_all, by = factor(species_code), k=3),
+gamm3 <- gamm(fr_mean ~ species_code + s(length_mean, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3) + s(pc2_all, by = factor(species_code), k=3),
            family = gaussian(link=identity),
@@
 
-gamm4 <- gamm(fr_mean ~ species_code + s(length_m, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3) + s(pc2_all, k=3), 
+gamm4 <- gamm(fr_mean ~ species_code + s(length_mean, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3) + s(pc2_all, k=3), 
            family = gaussian(link=identity),
@@
 
-gamm5 <- gamm(fr_mean ~ species + s(length_m, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3), 
+gamm5 <- gamm(fr_mean ~ species + s(length_mean, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3), 
            family = gaussian(link=identity),
@@
 
-gamm6 <- gamm(fr_mean ~ s(length_m, k=3) + s(pc1_all, by = factor(species_code), k=3), 
+gamm6 <- gamm(fr_mean ~ s(length_mean, k=3) + s(pc1_all, by = factor(species_code), k=3), 
            family = gaussian(link=identity),
@@
 
-gamm7 <- gamm(fr_mean ~ s(length_m, by = factor(species_code), k=3) + s(pc1_all, k=3), 
+gamm7 <- gamm(fr_mean ~ s(length_mean, by = factor(species_code), k=3) + s(pc1_all, k=3), 
            family = gaussian(link=identity),
@@
 
-gamm8 <- gamm(fr_mean ~ species + s(length_m, by = factor(species_code), k=3) + s(pc1_all, k=3), 
+gamm8 <- gamm(fr_mean ~ species + s(length_mean, by = factor(species_code), k=3) + s(pc1_all, k=3), 
            family = gaussian(link=identity),
@@
 
-gamm9 <- gamm(fr_mean ~ species + s(length_m, k=3) + s(pc1_all, by = factor(species_code), k=3), 
+gamm9 <- gamm(fr_mean ~ species + s(length_mean, k=3) + s(pc1_all, by = factor(species_code), k=3), 
            family = gaussian(link=identity),
@@
 
-gamm11 <- gamm(fr_mean ~ s(length_m, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3), 
+gamm11 <- gamm(fr_mean ~ s(length_mean, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3), 
            family = gaussian(link=identity),
@@
 
-gtab_fr <- as.data.frame(rbind(g1,g2,g4,g5,g6,g7,g8,g9,g10,g11)) %>% setNames(c("g","Model", "AICc", "Adj. R2")) %>% arrange(AICc)
+gamm12 <- gamm(fr_mean ~ species_code + s(pc1_all, k=3), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           data = m_data
+           )
+g12 <- c("g12", "Species + PC1", AICc(gamm12$lme), summary(gamm12$gam)$r.sq)
+
+gtab_fr_site <- as.data.frame(rbind(g1,g2,g4,g5,g6,g7,g8,g9,g10,g11,g12)) %>% setNames(c("g","Model", "AICc", "Adj. R2")) %>% arrange(AICc)
+
+# summary + plots from best models
+summary(gamm12$gam)
+plot(gamm12$gam,pages=1)
 ```
+
+## Random effects: site vs. site + island
+
+Site - level GAMM adding site as random effect
+```{r, echo=F, message=F, warning=F}
+m_data <- sum_ssp_pca %>% filter(species_code!="rbp" & phase == "i")
+
+gamm1 <- gamm(fr_mean ~ species_code + s(length_mean, k=3) + s(pc1_all, k=3) + s(pc2_all, k=3), 
+           family = gaussian(link=identity),
+           random = list(island=~1, site=~1),
+           data = m_data
+           )
+g1 <- c("g1", "Species + Size + PC1 + PC2", AICc(gamm1$lme), summary(gamm1$gam)$r.sq)
+
+gamm2 <- gamm(fr_mean ~ species_code + s(length_mean, k=3) + s(pc1_all, k=3), 
+           family = gaussian(link=identity),
+           random = list(island=~1, site=~1),
+           data = m_data
+           )
+g2 <- c("g2", "Species + Size + PC1", AICc(gamm2$lme), summary(gamm2$gam)$r.sq)
+
+# gamm3 <- gamm(fr_mean ~ species_code + s(length_mean, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3) + s(pc2_all, by = factor(species_code), k=3),
+#            family = gaussian(link=identity),
+#            random = list(island=~1, site=~1),
+#            data = m_data
+#            )
+# g3 <- c("g3", "Species + Size~Species + PC1~Species + PC2~Species", AICc(gamm3$lme), summary(gamm3$gam)$r.sq)
+
+gamm4 <- gamm(fr_mean ~ species_code + s(length_mean, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3) + s(pc2_all, k=3), 
+           family = gaussian(link=identity),
+           random = list(island=~1, site=~1),
+           data = m_data
+           )
+g4 <- c("g4", "Species + Size~Species + PC1~Species + PC2", AICc(gamm4$lme), summary(gamm4$gam)$r.sq)
+
+gamm5 <- gamm(fr_mean ~ species + s(length_mean, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3), 
+           family = gaussian(link=identity),
+           random = list(island=~1, site=~1),
+           data = m_data
+           )
+g5 <- c("g5", "Species + Size~Species + PC1~Species", AICc(gamm5$lme), summary(gamm5$gam)$r.sq)
+
+# gamm6 <- gamm(fr_mean ~ s(length_mean, k=3) + s(pc1_all, by = factor(species_code), k=3), 
+#            family = gaussian(link=identity),
+#            random = list(island=~1, site=~1),
+#            data = m_data
+#            )
+# g6 <- c("g6", "Size + PC1~Species", AICc(gamm6$lme), summary(gamm6$gam)$r.sq)
+
+gamm7 <- gamm(fr_mean ~ s(length_mean, by = factor(species_code), k=3) + s(pc1_all, k=3), 
+           family = gaussian(link=identity),
+           random = list(island=~1, site=~1),
+           data = m_data
+           )
+g7 <- c("g7", "Size~Species + PC1", AICc(gamm7$lme), summary(gamm7$gam)$r.sq)
+
+# gamm8 <- gamm(fr_mean ~ species + s(length_mean, by = factor(species_code), k=3) + s(pc1_all, k=3), 
+#            family = gaussian(link=identity),
+#            random = list(island=~1, site=~1),
+#            data = m_data
+#            )
+# g8 <- c("g8", "Species + Size~Species + PC1", AICc(gamm8$lme), summary(gamm8$gam)$r.sq)
+
+gamm9 <- gamm(fr_mean ~ species + s(length_mean, k=3) + s(pc1_all, by = factor(species_code), k=3), 
+           family = gaussian(link=identity),
+           random = list(island=~1, site=~1),
+           data = m_data
+           )
+g9 <- c("g9", "Species + Size + PC1~Species", AICc(gamm9$lme), summary(gamm9$gam)$r.sq)
+
+gamm10 <- gamm(fr_mean ~ species + s(pc1_all, by = factor(species_code), k=3), 
+           family = gaussian(link=identity),
+           random = list(island=~1, site=~1),
+           data = m_data
+           )
+g10 <- c("g10", "Species + PC1~Species", AICc(gamm10$lme), summary(gamm10$gam)$r.sq)
+
+gamm11 <- gamm(fr_mean ~ s(length_mean, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3), 
+           family = gaussian(link=identity),
+           random = list(island=~1, site=~1),
+           data = m_data
+           )
+g11 <- c("g11", "Size~Species + PC1~Species", AICc(gamm11$lme), summary(gamm11$gam)$r.sq)
+
+gamm12 <- gamm(fr_mean ~ species + s(pc1_all, k=3), 
+           family = gaussian(link=identity),
+           random = list(island=~1, site=~1),
+           data = m_data
+           )
+g12 <- c("g10", "Species + PC1", AICc(gamm10$lme), summary(gamm10$gam)$r.sq)
+
+gtab_fr_site_rd2 <- as.data.frame(rbind(g1,g2,g4,g5,g7,g8,g9,g10,g11,g12)) %>% setNames(c("g","Model", "AICc", "Adj. R2")) %>% arrange(AICc)
+
+# plots from best models
+plot(gamm10$gam,pages=1)
+plot(gamm2$gam,pages=1)
+```
+
+# Feeding intensity
+Site - level GAMMs, initial phase only
+```{r, echo=F, message=F, warning=F}
+m_data <- sum_ssp_pca %>% filter(species_code!="rbp" & phase == "i")
+
+gamm1 <- gamm(for_bites_mean ~ species_code + s(length_mean, k=3) + s(pc1_all, k=3) + s(pc2_all, k=3), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           data = m_data
+           )
+g1 <- c("g1", "Species + Size + PC1 + PC2", AICc(gamm1$lme), summary(gamm1$gam)$r.sq)
+
+gamm2 <- gamm(for_bites_mean ~ species_code + s(length_mean, k=3) + s(pc1_all, k=3), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           data = m_data
+           )
+g2 <- c("g2", "Species + Size + PC1", AICc(gamm2$lme), summary(gamm2$gam)$r.sq)
+
+# gamm3 <- gamm(for_bites_mean ~ species_code + s(length_mean, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3) + s(pc2_all, by = factor(species_code), k=3),
+#            family = gaussian(link=identity),
+#            random = list(island=~1),
+#            data = m_data
+#            )
+# g3 <- c("g3", "Species + Size~Species + PC1~Species + PC2~Species", AICc(gamm3$lme), summary(gamm3$gam)$r.sq)
+
+gamm4 <- gamm(for_bites_mean ~ species_code + s(length_mean, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3) + s(pc2_all, k=3), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           data = m_data
+           )
+g4 <- c("g4", "Species + Size~Species + PC1~Species + PC2", AICc(gamm4$lme), summary(gamm4$gam)$r.sq)
+
+gamm5 <- gamm(for_bites_mean ~ species + s(length_mean, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           data = m_data
+           )
+g5 <- c("g5", "Species + Size~Species + PC1~Species", AICc(gamm5$lme), summary(gamm5$gam)$r.sq)
+
+gamm6 <- gamm(for_bites_mean ~ s(length_mean, k=3) + s(pc1_all, by = factor(species_code), k=3), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           data = m_data
+           )
+g6 <- c("g6", "Size + PC1~Species", AICc(gamm6$lme), summary(gamm6$gam)$r.sq)
+
+gamm7 <- gamm(for_bites_mean ~ s(length_mean, by = factor(species_code), k=3) + s(pc1_all, k=3), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           data = m_data
+           )
+g7 <- c("g7", "Size~Species + PC1", AICc(gamm7$lme), summary(gamm7$gam)$r.sq)
+
+gamm8 <- gamm(for_bites_mean ~ species + s(length_mean, by = factor(species_code), k=3) + s(pc1_all, k=3), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           data = m_data
+           )
+g8 <- c("g8", "Species + Size~Species + PC1", AICc(gamm8$lme), summary(gamm8$gam)$r.sq)
+
+gamm9 <- gamm(for_bites_mean ~ species + s(length_mean, k=3) + s(pc1_all, by = factor(species_code), k=3), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           data = m_data
+           )
+g9 <- c("g9", "Species + Size + PC1~Species", AICc(gamm9$lme), summary(gamm9$gam)$r.sq)
+
+gamm10 <- gamm(for_bites_mean ~ species + s(pc1_all, by = factor(species_code), k=3), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           data = m_data
+           )
+g10 <- c("g10", "Species + PC1~Species", AICc(gamm10$lme), summary(gamm10$gam)$r.sq)
+
+gamm11 <- gamm(for_bites_mean ~ s(length_mean, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           data = m_data
+           )
+g11 <- c("g11", "Size~Species + PC1~Species", AICc(gamm11$lme), summary(gamm11$gam)$r.sq)
+
+gamm12 <- gamm(for_bites_mean ~ species_code + s(pc1_all, k=3), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           data = m_data
+           )
+g12 <- c("g12", "Species + PC1", AICc(gamm12$lme), summary(gamm12$gam)$r.sq)
+
+gtab_fi_site <- as.data.frame(rbind(g1,g2,g4,g5,g6,g7,g8,g9,g10,g11,g12)) %>% setNames(c("g","Model", "AICc", "Adj. R2")) %>% arrange(AICc)
+
+# plots from best models
+plot(gamm12$gam,pages=1)
+plot(gamm10$gam,pages=1)
+
+summary(gamm12$gam)
+```

### 2019-05-10 r/Sites.Rmd
5/10

@@
   select(-c(time,depth_ft,surveyor,date)) 
-bon_herb <- read.csv(here("data","bon_herb.csv"), stringsAsFactors=F) %>% # this file is already expanded
+bon_fish <- read.csv(here("data","bon_fish.csv"), stringsAsFactors=F) %>% # this file is already expanded
   janitor::clean_names(case="old_janitor") %>%
-  filter(date==2017) %>% 
-  select(-date) %>%
-  rename(island = region)
-bon_carn <- read.csv(here("data","bon_carn.csv"), stringsAsFactors=F) %>% 
-  janitor::clean_names(case="old_janitor") %>%
-  rename(island = region) %>%
+  filter(surveyor=="MWW") %>%
+  filter(site %in% c("Oil Slick","Karpata","Front Porch","18th Palm")) %>%
   uncount(number) %>%
-  select(-c(diver,date))
-fish <- bind_rows(bon_herb,bon_carn) %>% 
-  filter(site %in% c("Oil Slick","Karpata","Front Porch","18th Palm")) %>%
-  bind_rows(ab_fish) %>%
+  select(-c(surveyor,date))
+fish <- bind_rows(ab_fish, bon_fish) %>%
   mutate(
@@
 carn30_tran <- fish %>% 
-  filter(family %in% c("Serranidae","Lutjanidae","Carangidae", "Sphyraenidae", "Haemulidae"),
+  filter(family %in% c("Serranidae", "Lutjanidae", "Carangidae", "Sphyraenidae", "Haemulidae"),
          biomass != 0, length>=30) %>% 
@@
         scale_fill_manual(values = c("gray", "slategray2", "darkseagreen")) +
-        labs(title="Biomass of Piscivores over 25cm") +
+        labs(title="Biomass of Piscivores over 30cm") +
         theme_bw() +
@@
   replace(. == "NaN",NA)
+
+# Total scarids above
 

### 2019-05-10 r/old drafts/Data_prodding.Rmd
5/10

@@
 ```{r}
-ggplot(filter(sum_id, species_code!="rbp" & phase == "i" & length_cm<=30 & length_cm>=15), aes(x = length_cm)) +
+ggplot(filter(sum_id, species_code!="rbp" & phase == "i" & length_cm<=25 & length_cm>=17), aes(x = length_cm)) +
   facet_grid(.~species) +
@@
 }
-pairs(~consp_l+scar_bm+rugosity+ma_ind+ta_canopy+ta_cover, data=sum_id,
+pairs(~consp_l+scar_bm+carn30_bm+rugosity+ma_ind+ta_canopy+ta_cover, data=sum_id,
       lower.panel=panel.smooth, upper.panel=panel.cor, 
@@
               select(ma_cover, ma_canopy, ta_cover, ta_canopy, lc_cover, rugosity, scar_den, 
-                     scar_bm, spear_daily)),
+                     scar_bm, carn30_bm, spear_daily)),
   center = T,scale = T)
@@
 
-sum_ssp_pca <- sum_site_sp_ph %>%
+sum_ssp_pca <- sum_ssp %>%
   left_join(select(sum_site,site,pc1_bent,pc2_bent,pc1_all,pc2_all), by = "site")
@@
 ```
+
+Takeaway: positive skew -> should be using median to reflext site averages
 

### 2019-06-05 r/map_exp.Rmd
playing w maps

---
title: "Map"
author: "Molly Wilson"
date: "6/4/2019"
output: html_document
---

```{r}
library(tidyverse)
library(here)
library(sf)
library(ggmap)

site_data <- read.csv(here("data","site_coord.csv"))

# shape files for islands
atg <- read_sf("bes_adm1")

atg.shp <- system.file(here("data", "map", "atg", "atg.shp", package="sf"))
atg <- st_read(atg.shp)
bes.shp <- system.file("bes_adm1", package="sf")
bes <- st_read("bes.shp")

download.file("http://github.com/mgimond/Spatial/raw/master/Data/Income_schooling.zip", 
              destfile = "Income_schooling.zip" , mode='wb')
unzip("Income_schooling.zip", exdir = ".")
file.remove("Income_schooling.zip")
s.sf <- st_read("Income_schooling.shp")

download.file("http://github.com/mgimond/Spatial/raw/master/Data/Income_schooling.zip", 
              destfile = "Income_schooling.zip")
```

```{r}
atg_map <- read_sf(atg)
atg_map <- st_read("data/map/atg/atg.shp")
```

```{r}
map.world <- map_data("world", region = "Bonaire")
ggplot(map.world, aes(x=long, y=lat)) +
  geom_polygon(colour = "white")

islands <- c("Antigua","Barbuda","Bonaire")
ll.islands <- geocode(islands)
islands.x <- ll.islands$lon
islands.y <- ll.islands$lat

ggplot() +
  theme_bw() +
  geom_polygon(data = map.world, aes(x = long, y = lat))+
  coord_equal(xlim = c(-100, -60),ylim = c(27,10)) +
  geom_point(aes(x=islands.x, y=islands.y), color="#f24db3", size=5) +
  labs(x = "", y="",title = "")
```

```{r}
m_antigua <- get_map(location="Antigua", source="stamen", maptype="toner-lite", bbox = c(left = 110, bottom = -40, right = 160, top = -10), zoom=11)
ggmap(m_antigua)

m_carib <- get_map(location="Caribbean")
ggmap(m_carib)

antigua <- map_data("usa")
```

```{r}
ggplot(data = world) +
    geom_sf() +
    coord_sf(xlim = c(-102.15, -74.12), ylim = c(7.65, 33.97), expand = FALSE)
```

# Build base map, which we'll use for rest of maps
base_map <- ggplot()+ 
  geom_sf(data=world_land %>%
            mutate(Country = ifelse(ISO3 %in% c("BRA","PHL","IDN"),"Target","Other")), 
          color = NA, aes(fill=Country)) +
  scale_fill_manual(values=c("grey85","grey60")) +
  scale_color_manual(values = pal,name = "Site Type") +
  scale_shape_manual(values = shape_pal,name = "Site Type") +
  geom_sf(data = site_coords, aes(shape=site_type),color = "black",size=2.5, show.legend = FALSE) +
  geom_sf(data = site_coords,aes(color=site_type,shape=site_type), show.legend = "point",size=2) +
  guides(fill=FALSE)+
  theme_bw() +
  theme(plot.title = element_text(color="black",hjust=0,vjust=1, size=rel(1)),
        panel.border = element_rect(color = 'black', fill = NA,size=1),
        panel.grid.major = element_line(colour = "white"), 
        legend.text = element_text(color = "black", size = rel(2)),
        legend.title = element_text(color = "black", size = rel(2)),
        legend.title.align = 1,
        legend.background = element_rect(fill="white"),
        legend.position = "right",
        legend.direction="vertical",
        legend.text.align = 0,
        axis.text = element_text(color = "black", size = rel(1)),
        axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank()) + 
  guides(colour = guide_legend(override.aes = list(size=4)))

world_map <- base_map +
  geom_sf(data = philippines_bbox%>%
            st_as_sfc(),fill=NA) +
  geom_sf(data = indonesia_bbox%>%
            st_as_sfc(),fill=NA) +
  geom_sf(data = brazil_bbox%>%
            st_as_sfc(),fill=NA) +
  guides(color = FALSE,shape=FALSE)+
  xlim(c(world_bbox[1], world_bbox[3])) +
  ylim(c(world_bbox[2], world_bbox[4]))

brazil_map <- base_map +
  xlim(c(brazil_bbox[1], brazil_bbox[3])) +
  ylim(c(brazil_bbox[2], brazil_bbox[4])) +
  guides(color = FALSE,shape=FALSE)

indonesia_map <- base_map +
  xlim(c(indonesia_bbox[1], indonesia_bbox[3])) +
  ylim(c(indonesia_bbox[2], indonesia_bbox[4])) +
  guides(color = FALSE,shape=FALSE)

philippines_map <- base_map +
  xlim(c(philippines_bbox[1], philippines_bbox[3])) +
  ylim(c(philippines_bbox[2], philippines_bbox[4])) +
  guides(color = FALSE,shape=FALSE)

# Put all maps together
combined_map <- plot_grid(plot_grid(world_map,labels = "A"),
                          plot_grid(brazil_map, philippines_map, labels = c("B", "C"), rel_widths = c(0.65,0.35)),
                          plot_grid(indonesia_map, get_legend(base_map), labels = c("D", ""), rel_widths = c(0.65,0.35)),
                          ncol=1,
                          rel_heights = c(1,0.8,0.5))

ggsave(paste0("paper_package/figures/figure_1_map.png"),combined_map,width = 7.5,height=7.5,device="png",dpi=300)

### 2019-06-05 r/5.8.19.Rmd

@@
 library(tidyverse)
+library(factoextra)
+library(corrplot)
 library(mgcv)
@@
 ```{r, echo=F, message=F, warning=F}
-pca_pred <- prcomp(na.omit(sum_site %>% 
-              select(rugosity, ma_cover, ma_canopy, ta_cover, ta_canopy, lc_cover, scar_den, scar_bm, carn30_bm)),
-  center = T, scale = T)
-pca_bent <- prcomp(na.omit(sum_site %>% 
-              select(rugosity, ma_cover, ma_canopy, ta_cover, ta_canopy)),
-  center = T, scale = T)
-pca_fish <- prcomp(na.omit(sum_site %>% 
-              select(scar_bm, carn30_bm)),
-  center = T, scale = T)
+pca_data <- na.omit(sum_site %>%
+  select(rugosity, ma_cover, ma_canopy, ta_cover, ta_canopy, lc_cover, scar_bm, scar_den, carn30_bm))
+names(pca_data)[1:9] =
+  c("rugosity", "macroalgal cover", "macroalgal canopy", "turf cover", "turf canopy", "coral cover", "scarid biomass", "scarid density", "predator biomass")
 
-autoplot(pca_pred, data = sum_site, shape = 'island', colour = 'island',
-         loadings = T, loadings.colour = 'gray',
-         loadings.label = T, loadings.label.size = 3, loadings.label.colour = 'black',
-            loadings.label.repel = T) +
-    scale_color_manual(values = c("cadetblue3","cadetblue3","cadetblue3")) +
+pca_pred <- prcomp(pca_data, center = T, scale = T)
+
+# pca_bent <- prcomp(na.omit(sum_site %>% 
+#               select(rugosity, ma_cover, ma_canopy, ta_cover, ta_canopy)), center = T, scale = T)
+# pca_fish <- prcomp(na.omit(sum_site %>% 
+#               select(scar_bm, carn30_bm)), center = T, scale = T)
+
+fviz_pca_biplot(pca_pred,
+                col.ind = sum_site$island,
+                mean.point = F,
+                label = "var",
+                repel = T
+                ) +
+  scale_shape_manual(values=c(15,16,17)) +
+  scale_color_manual(values = c("black","black","black")) +
+  labs(x = "PC1 (55.6%)", y = "PC2 (24.5%)", title = "") +
   theme_bw() +
   theme(legend.title = element_blank())
-ggsave(here("figs/pca_pred.png"), width = 8, height = 5)
+
+ggsave(here("figs/pca_pred.png"), width = 7, height = 5)
+
+-var <- get_pca_var(pca_pred)
+corrplot(var$contrib, is.corr=FALSE)
 

### 2019-06-05 r/Ch2 outline_V3.Rmd

--- 2019-04-24 r/Ch2 outline_V3.Rmd
@@
   filter(fr <= 3000)
-sum_ssp <- read.csv(here("data","sum_site_sp_ph.csv"), stringsAsFactors=F) %>%
+sum_ssp <- read.csv(here("data","sum_ssp.csv"), stringsAsFactors=F) %>%
   select(-X) %>% 
@@
     labs(title="Mean feeding rates of individuals (initial phase)", 
-         y="Feeding rate (bites/hr)", 
+         y="Feeding rate (bites/min)", 
          x="Site") +
@@
               select(rugosity, ma_cover, ma_canopy, ta_cover, ta_canopy, scar_den, 
-                     scar_bm, carn30_bm)),
+                     scar_bm, carn30_bm, lc_cover)),
   center = T, scale = T)
@@
 
-## GAMM results
-
-### Predicting feeding rate:
-
-```{r, echo=F, message=F, warning=F}
-m_data <- sum_ssp_pca %>% filter(species_code!="rbp")
-
-gamm1 <- gamm(fr_mean ~ species_code + phase + s(length_m, k=4) + s(pc1_all, k=4) + s(pc2_all, k=4), 
-           family = gaussian(link=identity),
-           random = list(island=~1),
-           method = "REML",
-           data = m_data
-           )
-g1 <- c("g1", "Species + Phase + Size + PC1 + PC2", AICc(gamm1$lme), summary(gamm1$gam)$r.sq)
-
-gamm2 <- gamm(fr_mean ~ species_code + phase + s(length_m, k=4) + s(pc1_all, k=4), 
-           family = gaussian(link=identity),
-           random = list(island=~1),
-           method = "REML",
-           data = m_data
-           )
-g2 <- c("g2", "Species + Phase + Size + PC1", AICc(gamm2$lme), summary(gamm2$gam)$r.sq)
-
-gamm3 <- gamm(fr_mean ~ species_code + phase + s(length_m, by = factor(species_code), k=4) + s(pc1_all, by = factor(species_code), k=4) + s(pc2_all, by = factor(species_code), k=4), 
-           family = gaussian(link=identity),
-           random = list(island=~1),
-           method = "REML",
-           data = m_data
-           )
-g3 <- c("g3", "Species + Phase + Size~Species + PC1~Species + PC2~Species", AICc(gamm3$lme), summary(gamm3$gam)$r.sq)
-
-gamm4 <- gamm(fr_mean ~ species_code + phase + s(length_m, by = factor(species_code), k=4) + s(pc1_all, by = factor(species_code), k=4) + s(pc2_all, k=4), 
-           family = gaussian(link=identity),
-           random = list(island=~1),
-           method = "REML",
-           data = m_data
-           )
-g4 <- c("g4", "Species + Phase + Size~Species + PC1~Species + PC2", AICc(gamm4$lme), summary(gamm4$gam)$r.sq)
-
-gamm5 <- gamm(fr_mean ~ species + phase + s(length_m, by = factor(species_code), k=4) + s(pc1_all, by = factor(species_code), k=4), 
-           family = gaussian(link=identity),
-           random = list(island=~1),
-           method = "REML",
-           data = m_data
-           )
-g5 <- c("g5", "Species + Phase + Size~Species + PC1~Species", AICc(gamm5$lme), summary(gamm5$gam)$r.sq)
-
-gamm6 <- gamm(fr_mean ~ phase + s(length_m, k=4) + s(pc1_all, by = factor(species_code), k=4), 
-           family = gaussian(link=identity),
-           random = list(island=~1),
-           method = "REML",
-           data = m_data
-           )
-g6 <- c("g6", "Phase + Size + PC1~Species", AICc(gamm6$lme), summary(gamm6$gam)$r.sq)
-
-gamm7 <- gamm(fr_mean ~ phase + s(length_m, by = factor(species_code), k=4) + s(pc1_all, k=4), 
-           family = gaussian(link=identity),
-           random = list(island=~1),
-           method = "REML",
-           data = m_data
-           )
-g7 <- c("g7", "Phase + Size~Species + PC1", AICc(gamm7$lme), summary(gamm7$gam)$r.sq)
-
-gamm8 <- gamm(fr_mean ~ species + phase + s(length_m, by = factor(species_code), k=4) + s(pc1_all, k=4), 
-           family = gaussian(link=identity),
-           random = list(island=~1),
-           method = "REML",
-           data = m_data
-           )
-g8 <- c("g8", "Species + Phase + Size~Species + PC1", AICc(gamm8$lme), summary(gamm8$gam)$r.sq)
-
-gamm9 <- gamm(fr_mean ~ species + phase + s(length_m, k=4) + s(pc1_all, by = factor(species_code), k=4), 
-           family = gaussian(link=identity),
-           random = list(island=~1),
-           method = "REML",
-           data = m_data
-           )
-g9 <- c("g9", "Species + Phase + Size + PC1~Species", AICc(gamm9$lme), summary(gamm9$gam)$r.sq)
-
-gamm10 <- gamm(fr_mean ~ species + phase + s(pc1_all, by = factor(species_code), k=4), 
-           family = gaussian(link=identity),
-           random = list(island=~1),
-           method = "REML",
-           data = m_data
-           )
-g10 <- c("g10", "Species + Phase + PC1~Species", AICc(gamm10$lme), summary(gamm10$gam)$r.sq)
-
-gamm11 <- gamm(fr_mean ~ phase + s(length_m, by = factor(species_code), k=4) + s(pc1_all, by = factor(species_code), k=4), 
-           family = gaussian(link=identity),
-           random = list(island=~1),
-           method = "REML",
-           data = m_data
-           )
-g11 <- c("g11", "Phase + Size~Species + PC1~Species", AICc(gamm11$lme), summary(gamm11$gam)$r.sq)
-
-gtab_fr <- as.data.frame(rbind(g1,g2,g3,g4,g5,g6,g7,g8,g9,g10,g11)) %>% setNames(c("g","Model", "AICc", "Adj. R2")) %>% arrange(AICc)
-```
-
-```{r, echo=F, message=F}
-kable(gtab_fr[1:4, ])
-plot(gamm10$gam,pages=1)
-```
-
-### Predicting feeding intensity:
-
-```{r, echo=F, message=F, warning=F}
-m_data <- sum_ssp_pca %>% filter(species_code!="rbp")
-
-fi_gamm1 <- gamm(for_bites_mean ~ species_code + phase + s(length_m, k=4) + s(pc1_all, k=4) + s(pc2_all, k=4), 
-           family = gaussian(link=identity),
-           random = list(island=~1),
-           method = "REML",
-           data = m_data
-           )
-# summary(fi_gamm1$gam)
-# plot(fi_gamm1$gam,pages=1)
-g1 <- c("g1", "Species + Phase + Size + PC1 + PC2", AICc(fi_gamm1$lme), summary(fi_gamm1$gam)$r.sq)
-
-fi_gamm2 <- gamm(for_bites_mean ~ species_code + phase + s(length_m, k=4) + s(pc1_all, k=4), 
-           family = gaussian(link=identity),
-           random = list(island=~1),
-           method = "REML",
-           data = m_data
-           )
-g2 <- c("g2", "Species + Phase + Size + PC1", AICc(fi_gamm2$lme), summary(fi_gamm2$gam)$r.sq)
-
-fi_gamm3 <- gamm(for_bites_mean ~ species_code + phase + s(length_m, by = factor(species_code), k=4) + s(pc1_all, by = factor(species_code), k=4) + s(pc2_all, by = factor(species_code), k=4), 
-           family = gaussian(link=identity),
-           random = list(island=~1),
-           method = "REML",
-           data = m_data
-           )
-g3 <- c("g3", "Species + Phase + Size~Species + PC1~Species", AICc(fi_gamm3$lme), summary(fi_gamm3$gam)$r.sq)
-
-fi_gamm4 <- gamm(for_bites_mean ~ species_code + phase + s(length_m, by = factor(species_code), k=4) + s(pc1_all, by = factor(species_code), k=4) + s(pc2_all, k=4), 
-           family = gaussian(link=identity),
-           random = list(island=~1),
-           method = "REML",
-           data = m_data
-           )
-g4 <- c("g4", "Species + Phase + Size~Species + PC1~Species + PC2", AICc(fi_gamm4$lme), summary(fi_gamm4$gam)$r.sq)
-
-fi_gamm5 <- gamm(for_bites_mean ~ phase + s(length_m, by = factor(species_code), k=4) + s(pc1_all, by = factor(species_code), k=4), 
-           family = gaussian(link=identity),
-           random = list(island=~1),
-           method = "REML",
-           data = m_data
-           )
-g5 <- c("g5", "Phase + Size~Species + PC1~Species", AICc(fi_gamm5$lme), summary(fi_gamm5$gam)$r.sq)
-
-fi_gamm6 <- gamm(for_bites_mean ~ species + phase + s(length_m, k=4) + s(pc1_all, by = factor(species_code), k=4), 
-           family = gaussian(link=identity),
-           random = list(island=~1),
-           method = "REML",
-           data = m_data
-           )
-g6 <- c("g6", "Species + Phase + Size + PC1~Species", AICc(fi_gamm6$lme), summary(fi_gamm6$gam)$r.sq)
-
-fi_gamm7 <- gamm(for_bites_mean ~ species + phase + s(length_m, by = factor(species_code), k=4) + s(pc1_all, k=4), 
-           family = gaussian(link=identity),
-           random = list(island=~1),
-           method = "REML",
-           data = m_data
-           )
-g7 <- c("g7", "Species + Phase + Size~Species + PC1", AICc(fi_gamm7$lme), summary(fi_gamm7$gam)$r.sq)
-
-fi_gamm8 <- gamm(for_bites_mean ~ phase + s(length_m, k=4) + s(pc1_all, by = factor(species_code), k=4), 
-           family = gaussian(link=identity),
-           random = list(island=~1),
-           method = "REML",
-           data = m_data
-           )
-g8 <- c("g8", "Phase + Size + PC1~Species", AICc(fi_gamm8$lme), summary(fi_gamm8$gam)$r.sq)
-
-fi_gamm9 <- gamm(for_bites_mean ~ phase + s(length_m, by = factor(species_code), k=4) + s(pc1_all, k=4), 
-           family = gaussian(link=identity),
-           random = list(island=~1),
-           method = "REML",
-           data = m_data
-           )
-g9 <- c("g9", "Phase + Size~Species + PC1", AICc(fi_gamm9$lme), summary(fi_gamm9$gam)$r.sq)
-
-gtab_fi <- as.data.frame(rbind(g1,g2,g3,g4,g5,g6,g7,g8,g9)) %>% setNames(c("g", "Model", "AICc", "Adj. R2")) %>% arrange(AICc)
-```
-
-```{r, echo=F, message=F}
-kable(gtab_fi[1:3, ])
-plot(gamm6$gam,pages=1)
-```
+## GAMM results (see 5.8.19)

### 2019-06-05 r/Follows.Rmd

@@
     g_frac_mean=mean(g_frac),
+    g_frac_median=median(g_frac),
     g_frac_se=sd(g_frac)/sqrt(n),
     br_mean=mean(br,na.rm=TRUE),
+    br_median=median(br,na.rm=TRUE),
     br_se=sd(br,na.rm=TRUE)/sqrt(n),
     fr_mean=mean(fr),
+    fr_median=median(fr),
     fr_se=sd(fr)/sqrt(n),
     for_dur_mean=mean(for_dur,na.rm=TRUE),
+    for_dur_median=median(for_dur,na.rm=TRUE),
     for_dur_se=sd(for_dur,na.rm=TRUE)/sqrt(n),
     for_bites_mean=mean(for_bites,na.rm=TRUE),
+    for_bites_median=median(for_bites,na.rm=TRUE),
     for_bites_se=sd(for_bites,na.rm=TRUE)/sqrt(n),
     int_tot_mean=mean(int_tot),
+    int_tot_median=median(int_tot),
     int_tot_se=sd(int_tot)/sqrt(n),
     int_scar_mean=mean(int_scar),
+    int_scar_median=median(int_scar),
     int_scar_se=sd(int_scar)/sqrt(n),
     int_con_mean=mean(int_con),
+    int_con_median=median(int_con),
     int_con_se=sd(int_con)/sqrt(n),
     int_hetscar_mean=mean(int_hetscar),
+    int_hetscar_median=median(int_hetscar),
     int_hetscar_se=sd(int_hetscar)/sqrt(n),
     int_dam_mean=mean(int_dam),
+    int_dam_median=median(int_dam),
     int_dam_se=sd(int_dam)/sqrt(n)
@@
     n=n(),
+    length_mean=mean(length_cm),
+    length_se=sd(length_cm)/sqrt(n),
     g_frac_mean=mean(g_frac),
+    g_frac_median=median(g_frac),
     g_frac_se=sd(g_frac)/sqrt(n),
     br_mean=mean(br,na.rm=TRUE),
+    br_median=median(br,na.rm=TRUE),
     br_se=sd(br,na.rm=TRUE)/sqrt(n),
     fr_mean=mean(fr),
+    fr_median=median(fr),
     fr_se=sd(fr)/sqrt(n),
     for_dur_mean=mean(for_dur,na.rm=TRUE),
+    for_dur_median=median(for_dur,na.rm=TRUE),
     for_dur_se=sd(for_dur,na.rm=TRUE)/sqrt(n),
     for_bites_mean=mean(for_bites,na.rm=TRUE),
+    for_bites_median=median(for_bites,na.rm=TRUE),
     for_bites_se=sd(for_bites,na.rm=TRUE)/sqrt(n),
     int_tot_mean=mean(int_tot),
+    int_tot_median=median(int_tot),
     int_tot_se=sd(int_tot)/sqrt(n),
     int_scar_mean=mean(int_scar),
+    int_scar_median=median(int_scar),
     int_scar_se=sd(int_scar)/sqrt(n),
     int_con_mean=mean(int_con),
+    int_con_median=median(int_con),
     int_con_se=sd(int_con)/sqrt(n),
     int_hetscar_mean=mean(int_hetscar),
+    int_hetscar_median=median(int_hetscar),
     int_hetscar_se=sd(int_hetscar)/sqrt(n),
     int_dam_mean=mean(int_dam),
+    int_dam_median=median(int_dam),
     int_dam_se=sd(int_dam)/sqrt(n)
@@
     n=n(),
+    length_mean=mean(length_cm),
+    length_se=sd(length_cm)/sqrt(n),
     g_frac_mean=mean(g_frac),
+    g_frac_median=median(g_frac),
     g_frac_se=sd(g_frac)/sqrt(n),
     br_mean=mean(br,na.rm=TRUE),
+    br_median=median(br,na.rm=TRUE),
     br_se=sd(br,na.rm=TRUE)/sqrt(n),
     fr_mean=mean(fr),
+    fr_median=median(fr),
     fr_se=sd(fr)/sqrt(n),
     for_dur_mean=mean(for_dur,na.rm=TRUE),
+    for_dur_median=median(for_dur,na.rm=TRUE),
     for_dur_se=sd(for_dur,na.rm=TRUE)/sqrt(n),
     for_bites_mean=mean(for_bites,na.rm=TRUE),
+    for_bites_median=median(for_bites,na.rm=TRUE),
     for_bites_se=sd(for_bites,na.rm=TRUE)/sqrt(n),
     int_tot_mean=mean(int_tot),
+    int_tot_median=median(int_tot),
     int_tot_se=sd(int_tot)/sqrt(n),
     int_scar_mean=mean(int_scar),
+    int_scar_median=median(int_scar),
     int_scar_se=sd(int_scar)/sqrt(n),
     int_con_mean=mean(int_con),
+    int_con_median=median(int_con),
     int_con_se=sd(int_con)/sqrt(n),
     int_hetscar_mean=mean(int_hetscar),
+    int_hetscar_median=median(int_hetscar),
     int_hetscar_se=sd(int_hetscar)/sqrt(n),
     int_dam_mean=mean(int_dam),
-    int_dam_se=sd(int_dam)/sqrt(n),
-    length_m = mean(length_cm),
-    length_se = sd(length_cm)/sqrt(n)
+    int_dam_median=median(int_dam),
+    int_dam_se=sd(int_dam)/sqrt(n)
     )

### 2019-06-05 r/Sites.Rmd

@@
   filter(family=="Scaridae",
-         biomass != 0) %>% 
+         biomass != 0,
+         length > 5) %>% 
   group_by(site,transect) %>% 

### 2019-06-05 r/old drafts/Data_prodding.Rmd

@@
 }
-pairs(~consp_l+scar_bm+carn30_bm+rugosity+ma_ind+ta_canopy+ta_cover, data=sum_id,
+pairs(~consp_l+scar_bm+carn30_bm+carn30_den+rugosity+ma_ind+ta_canopy+ta_cover, data=sum_id,
       lower.panel=panel.smooth, upper.panel=panel.cor, 

### 2019-06-13 r/map_exp.Rmd
updating map/shapefiles

@@
 library(ggmap)
+# library(ggspatial)
+library(rnaturalearth)
+library(rnaturalearthdata)
+library(ggrepel)
 
-site_data <- read.csv(here("data","site_coord.csv"))
+# site coordinates
+site_data <- read.csv(here("data","map","site_coord.csv"))
 
-# shape files for islands
-atg <- read_sf("bes_adm1")
+# shapefiles for islands filtered from BES and ATG boundaries
+bonaire <- st_read(here("data","map","BES_adm1.shp")) %>% 
+  filter(NAME_1 == "Bonaire") %>%
+  st_union() %>%
+  st_sf() %>%
+  mutate(island = "Bonaire")
 
-atg.shp <- system.file(here("data", "map", "atg", "atg.shp", package="sf"))
-atg <- st_read(atg.shp)
-bes.shp <- system.file("bes_adm1", package="sf")
-bes <- st_read("bes.shp")
+atg <- st_read(here("data","map","ATG_adm1_gadm.shp")) 
+antigua <- atg %>%
+  filter(!NAME_1 == c("Barbuda","Redonda")) %>%
+  st_union() %>%
+  st_sf() %>%
+  mutate(island = "Antigua")
+barbuda <- atg %>% 
+  filter(NAME_1 == "Barbuda") %>%
+  st_union() %>%
+  st_sf() %>%
+  mutate(island = "Barbuda")
 
-download.file("http://github.com/mgimond/Spatial/raw/master/Data/Income_schooling.zip", 
-              destfile = "Income_schooling.zip" , mode='wb')
-unzip("Income_schooling.zip", exdir = ".")
-file.remove("Income_schooling.zip")
-s.sf <- st_read("Income_schooling.shp")
+islands <- rbind(antigua, barbuda, bonaire) %>%
+  mutate(lon=map_dbl(geometry, ~st_centroid(.x)[[1]]), 
+         lat=map_dbl(geometry, ~st_centroid(.x)[[2]]))
 
-download.file("http://github.com/mgimond/Spatial/raw/master/Data/Income_schooling.zip", 
-              destfile = "Income_schooling.zip")
+# google API (not sure if I can register this permanently?)
+register_google(key = "AIzaSyDvckfjy2riROUUuL9oDg7HBXRBd39X82A")
 ```
 
+### Regional map
+
 ```{r}
-atg_map <- read_sf(atg)
-atg_map <- st_read("data/map/atg/atg.shp")
+world <- ne_countries(scale = "large", returnclass = "sf")
+ggplot() +
+  geom_sf(data = world, fill = "gray80", col = "gray80", size = .25) +
+  geom_sf(data = islands, fill = "gray20", col = "gray20", size = .25) +
+  #geom_text_repel(data = islands %>% filter(island %in% c("Antigua","Barbuda")), aes(x = lon, y = lat, label = island), nudge_x = 6, segment.size = .2) +
+  #geom_text_repel(data = islands %>% filter(island == "Bonaire"), aes(x = lon, y = lat, label = island), nudge_x = -2, nudge_y = 2, segment.size = .2) +
+  #annotate(geom = "text", x = -86.5, y = 26, label = "Gulf of Mexico", fontface = "italic", color = "grey80", size = 3) +
+  annotate(geom = "text", x = -65, y = 29, label = "Atlantic Ocean", fontface = "italic", color = "grey80", size = 4) +
+  annotate(geom = "text", x = -72.5, y = 15, label = "Caribbean Sea", fontface = "italic", color = "grey80", size = 4) +
+  coord_sf(xlim = c(-86, -58), ylim = c(10, 33), expand = FALSE) +
+  labs(x = "", y = "") +
+  theme_bw()
+ggsave(here("figs/map_region.png"))
 ```
 
+Alternate (geocoding) option
 ```{r}
-map.world <- map_data("world", region = "Bonaire")
-ggplot(map.world, aes(x=long, y=lat)) +
-  geom_polygon(colour = "white")
-
-islands <- c("Antigua","Barbuda","Bonaire")
-ll.islands <- geocode(islands)
+#world <- map_data("world") # df
+ll.islands <- geocode(c("Antigua","Barbuda","Bonaire"))
 islands.x <- ll.islands$lon
@@
   theme_bw() +
-  geom_polygon(data = map.world, aes(x = long, y = lat))+
-  coord_equal(xlim = c(-100, -60),ylim = c(27,10)) +
-  geom_point(aes(x=islands.x, y=islands.y), color="#f24db3", size=5) +
-  labs(x = "", y="",title = "")
+  geom_sf(data = world, fill = "gray80", col = "gray80") +
+  annotation_scale(location = "tr", width_hint = 0.25) +
+  coord_sf(xlim = c(-90, -58), ylim = c(10, 33)) +
+  geom_point(aes(x = islands.x, y = islands.y), color = "#f24db3", size = 5, shape = 1) +
+  labs(x = "", y = "",title = "")
+
+```
+
+### Island-level details
+
+```{r}
+sites_anu <- site_data %>% filter(island == "Antigua")
+ggplot() +
+  geom_sf(data = antigua, fill = "slategray4", color = "slategray4") +
+  geom_point(data = sites_anu, aes(x = lon, y = lat), color = "gray20", size = 2) +
+  geom_text_repel(data = sites_anu,
+                  aes(x = lon, y = lat, label = site),
+                  segment.size = .2,
+                  direction = "y",
+                  box.padding = .75,
+                  nudge_x = .18, nudge_y = .08,
+                  size = 5) +
+  coord_sf(xlim = c(-61.92, -61.52), ylim = c(16.98, 17.18)) +
+  annotation_scale(location = "bl", width_hint = 0.25) +
+  theme_bw() +
+  theme(axis.text = element_blank(), axis.ticks = element_blank(), axis.title = element_blank(), panel.grid.major = element_line(color = "transparent"))
+
+ggsave(here("figs/map_anu.png"))
 ```
@@
 ```{r}
-m_antigua <- get_map(location="Antigua", source="stamen", maptype="toner-lite", bbox = c(left = 110, bottom = -40, right = 160, top = -10), zoom=11)
-ggmap(m_antigua)
+sites_brb <- site_data %>% filter(island == "Barbuda")
+ggplot() +
+  geom_sf(data = barbuda, fill = "slategray4", color = "slategray4") +
+  geom_point(data = sites_brb, aes(x = lon, y = lat), color = "gray20", size = 2) +
+  geom_text_repel(data = sites_brb,
+                  aes(x = lon, y = lat, label = site),
+                  segment.size = .2,
+                  direction = "y",
+                  box.padding = .75,
+                  nudge_x = .18, nudge_y = .12,
+                  size = 5) +
+  coord_sf(xlim = c(-61.9, -61.44), ylim = c(17.51, 17.74)) +
+  annotation_scale(location = "bl", width_hint = 0.25) +
+  theme_bw() +
+  theme(axis.text = element_blank(), axis.ticks = element_blank(), axis.title = element_blank(), panel.grid.major = element_line(color = "transparent"))
 
-m_carib <- get_map(location="Caribbean")
-ggmap(m_carib)
-
-antigua <- map_data("usa")
+ggsave(here("figs/map_brb.png"))
 ```
@@
 ```{r}
-ggplot(data = world) +
-    geom_sf() +
-    coord_sf(xlim = c(-102.15, -74.12), ylim = c(7.65, 33.97), expand = FALSE)
+sites_bon <- site_data %>% filter(island == "Bonaire")
+ggplot() +
+  geom_sf(data = bonaire, fill = "slategray4", color = "slategray4") +
+  geom_point(data = sites_bon, aes(x = lon, y = lat), color = "gray20", size = 2) +
+  geom_text_repel(data = sites_bon,
+                  aes(x = lon, y = lat, label = site),
+                  segment.size = .2,
+                  direction = "y",
+                  box.padding = .75,
+                  nudge_x = .22, nudge_y = .02,
+                  size = 5) +
+  coord_sf(xlim = c(-68.45, -67.85), ylim = c(12.02, 12.32)) +
+  annotation_scale(location = "bl", width_hint = 0.25) +
+  theme_bw() +
+  theme(axis.text = element_blank(), axis.ticks = element_blank(), axis.title = element_blank(), panel.grid.major = element_line(color = "transparent"))
+
+ggsave(here("figs/map_bon.png"))
 ```
-
-# Build base map, which we'll use for rest of maps
-base_map <- ggplot()+ 
-  geom_sf(data=world_land %>%
-            mutate(Country = ifelse(ISO3 %in% c("BRA","PHL","IDN"),"Target","Other")), 
-          color = NA, aes(fill=Country)) +
-  scale_fill_manual(values=c("grey85","grey60")) +
-  scale_color_manual(values = pal,name = "Site Type") +
-  scale_shape_manual(values = shape_pal,name = "Site Type") +
-  geom_sf(data = site_coords, aes(shape=site_type),color = "black",size=2.5, show.legend = FALSE) +
-  geom_sf(data = site_coords,aes(color=site_type,shape=site_type), show.legend = "point",size=2) +
-  guides(fill=FALSE)+
-  theme_bw() +
-  theme(plot.title = element_text(color="black",hjust=0,vjust=1, size=rel(1)),
-        panel.border = element_rect(color = 'black', fill = NA,size=1),
-        panel.grid.major = element_line(colour = "white"), 
-        legend.text = element_text(color = "black", size = rel(2)),
-        legend.title = element_text(color = "black", size = rel(2)),
-        legend.title.align = 1,
-        legend.background = element_rect(fill="white"),
-        legend.position = "right",
-        legend.direction="vertical",
-        legend.text.align = 0,
-        axis.text = element_text(color = "black", size = rel(1)),
-        axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank(),
-        axis.title.y=element_blank(),
-        axis.text.y=element_blank(),
-        axis.ticks.y=element_blank()) + 
-  guides(colour = guide_legend(override.aes = list(size=4)))
-
-world_map <- base_map +
-  geom_sf(data = philippines_bbox%>%
-            st_as_sfc(),fill=NA) +
-  geom_sf(data = indonesia_bbox%>%
-            st_as_sfc(),fill=NA) +
-  geom_sf(data = brazil_bbox%>%
-            st_as_sfc(),fill=NA) +
-  guides(color = FALSE,shape=FALSE)+
-  xlim(c(world_bbox[1], world_bbox[3])) +
-  ylim(c(world_bbox[2], world_bbox[4]))
-
-brazil_map <- base_map +
-  xlim(c(brazil_bbox[1], brazil_bbox[3])) +
-  ylim(c(brazil_bbox[2], brazil_bbox[4])) +
-  guides(color = FALSE,shape=FALSE)
-
-indonesia_map <- base_map +
-  xlim(c(indonesia_bbox[1], indonesia_bbox[3])) +
-  ylim(c(indonesia_bbox[2], indonesia_bbox[4])) +
-  guides(color = FALSE,shape=FALSE)
-
-philippines_map <- base_map +
-  xlim(c(philippines_bbox[1], philippines_bbox[3])) +
-  ylim(c(philippines_bbox[2], philippines_bbox[4])) +
-  guides(color = FALSE,shape=FALSE)
-
-# Put all maps together
-combined_map <- plot_grid(plot_grid(world_map,labels = "A"),
-                          plot_grid(brazil_map, philippines_map, labels = c("B", "C"), rel_widths = c(0.65,0.35)),
-                          plot_grid(indonesia_map, get_legend(base_map), labels = c("D", ""), rel_widths = c(0.65,0.35)),
-                          ncol=1,
-                          rel_heights = c(1,0.8,0.5))
-
-ggsave(paste0("paper_package/figures/figure_1_map.png"),combined_map,width = 7.5,height=7.5,device="png",dpi=300)

### 2019-09-17 r/Ch2 outline_V3.Rmd
updating stats

@@
 library(ggfortify)
+library(ggsignif) # boxplot stats
+library(EnvStats) # boxplot stats
 library(mgcv)
@@
   select(-X) %>% 
+  filter(species_code %in% c("stop","qup")) %>%
   mutate(island = factor(island, levels = c("Bonaire","Antigua","Barbuda"))) %>%
@@
   filter(fr <= 3000)
-sum_ssp <- read.csv(here("data","sum_ssp.csv"), stringsAsFactors=F) %>%
-  select(-X) %>% 
-  mutate(island = factor(island, levels = c("Bonaire","Antigua","Barbuda"))) %>%
-  arrange(island)
+
+### data subset: excluding sites with low sample sizes, including only individuals 15-35 cm
+sum_id_sub <- sum_id %>%
+  filter(phase == "i" & length_cm >= 15 & length_cm <= 35 & !(species_code == "stop" & site == "Pallaster West") & !(species_code == "qup" & site %in% c("Pallaster East","Rendezvous","Turtle Bay")))
+
+sum_ssp_sub <- sum_id_sub %>% 
+  group_by(species, species_code, site, island) %>%
+  summarize(
+    n=n(),
+    length_mean=mean(length_cm),
+    length_se=sd(length_cm)/sqrt(n),
+    g_frac_mean=mean(g_frac),
+    g_frac_median=median(g_frac),
+    g_frac_se=sd(g_frac)/sqrt(n),
+    br_mean=mean(br,na.rm=TRUE),
+    br_median=median(br,na.rm=TRUE),
+    br_se=sd(br,na.rm=TRUE)/sqrt(n),
+    fr_mean=mean(fr),
+    fr_median=median(fr),
+    fr_se=sd(fr)/sqrt(n),
+    for_dur_mean=mean(for_dur,na.rm=TRUE),
+    for_dur_median=median(for_dur,na.rm=TRUE),
+    for_dur_se=sd(for_dur,na.rm=TRUE)/sqrt(n),
+    for_bites_mean=mean(for_bites,na.rm=TRUE),
+    for_bites_median=median(for_bites,na.rm=TRUE),
+    for_bites_se=sd(for_bites,na.rm=TRUE)/sqrt(n)
+    )
+
 ```
 
-**Abstract:** Herbivorous fish increase coral growth and survival by grazing down algal competitors.
-As coral reef health continues to decline globally, maximizing herbivory on reefs has become a primary management goal.
-However, existing management targets assume consistent grazing behavior across different reef systems.
-Here I document the grazing behavior of two scarid species at 13 sites in Antigua, Barbuda, and Bonaire.
-My analyses show significant differences in overall feeding rates, time spent grazing, and intensity of grazing across reef sites, which may alter the ecological impact of a given scarid population.
-I propose several potential mechanisms for these grazing differences, including variations in bite content which may make bite rates a misleading measure of overall grazing impact.
-As we set management targets to enhance herbivory on coral reefs, it is critical that we are accurately estimating the grazing impact of scarid populations.
-This paper demonstrates the variability of grazing behavior across different reef sites and calls for further investigation of the drivers and ecological implications of these inconsistencies.
+## Figuring out sample sizes/site consolidations...
 
-## Introduction
+```{r}
+sample_exp <- sum_id %>% 
+  filter(phase == "i" & length_cm >= 15 & length_cm <= 35) %>%
+  group_by(species, species_code, site, island) %>%
+  summarize(
+    n=n())
 
-Herbivorous fish play a critical role in coral reef ecosystems by supressing algae and facilitating coral growth, recruitment, and survival ().
-In recent decades, coral reefs have become increasingly threatened by anthropogenic impacts including the fishing down of herbivorous fish populations ().
-Increasing attention has been given to understanding how much herbivory is needed to sustain coral reef health and to identify thresholds of herbivore biomass that can be used to guide fisheries management targets (Karr et al.
-2015, McClanahan et al.
-2011)
+ggplot(filter(sum_id, 
+              phase=="i" & length_cm >= 15 & length_cm <= 50 & species_code!="rbp"), 
+       aes(factor(site), fr)) + 
+  geom_violin(aes(fill=island)) +
+  geom_boxplot(width=0.1) + 
+  labs(y = "Feeding Rate (bites/hr)", x = "", title = "Hourly feeding rates (i. phase)") +
+  geom_signif(comparisons = list(c("Antigua", "Barbuda"),c("Antigua","Bonaire"),c("Barbuda","Bonaire")), map_signif_level=TRUE) +
+  facet_grid(.~species) + 
+  stat_n_text(angle = 90) +
+  scale_fill_brewer(palette="Blues") +
+  theme_bw() +
+  theme(strip.text = element_text(face = "italic")) +
+  theme(plot.title = element_text(hjust = 0.5), axis.text.x = element_text(angle = 90, hjust = 1)) 
+```
 
-However, not all herbivores are the same in terms of their grazing impact on reefs.
-Recent studies have highlighted functional differences across herbivorous fish species in terms of their feeding selectivity, feeding morphology, etc.
-(e.g.
-Adam et al.
-2018) [**add more here**]
-
-In addition to these functional differences found across species, herbivore behavior may also drive variation in herbivore function across different reef environments.
-Empirical studies have demonstrated the sensitivity of herbivore behavior to predator presence (), reef rugosity (), grazing behaviors of proximate herbivores (Gil et al.
-2018), and algal abundance, distribution, and nutritional content ().
-[**add more here**]
-[These behavioral variations have not been linked to larger-scale variations - except Madin's seascape papers...]
-[Maybe even highlight Madin papers - these behavioral differences have the potential to drive 
-Madin et al.'s study showing how predator abundance drove herbivore feeding ranges/grazing halos]
-
-Despite a growing understanding of the sensitivity and ecological importance of herbivore behavior, current efforts to predict herbivore impact on reefs assume species-specific feeding rates that vary only with individual fish size.
-There is very limited data on how herbivore feeding behaviors might vary across reef systems and whether we can accurately assume constant species roles at regional scales.
-Here we document variability in multiple components of scarid feeding behavior across reef sites in three Caribbean islands and discuss potential underlying mechanisms, ecological consequences, and management implications.
-We demonstrate a need for further experimental and modelling work to unpack these drivers and determine their ecological consequences.
-We urge for the incorporation of behavioral effects in coral reef herbivore management and caution against assuming constant species function across reef systems.
-
-# Methods
-
-*Data collection*
-In March through August of 2017, I collected data from 13 reef sites in Bonaire, Antigua, and Barbuda (Fig.
-1).
-At each site I conducted fish, benthic, and rugosity surveys to quantify reef (condition).
-I used a modified AGRRA protocol with 30m by 4m belt transects and 10m point-count transects for fish and benthic survets, respectively.
-To assess rugosity I measured the length of reef contour under each meter of a taut 10m leadline (sensu ___).
-I carried out a minimum of X fish transects and X benthic and rugosity transects per site.
-
-At each site, I observed grazing behavior of two dominant scarid species, *Sparisoma viride* and *Scarus vetula*.
-I targetted these species because of their relative abundance as well as their contrasting grazing mechanisms.
-*S.
-viride* is classified as an excavator, taking [_] (), while *S.
-vetula* is a grazer, taking shallower bites from [_] ().
-I quantified grazing behavior during two minute follows,
-* Observational behavioral data:
-    - Species: , initial and terminal phase greater than 15cm, focus on only initial phase here
-    - 2 minute follows, calibration period, outside of previously established FID
-    - Recording time budgets as well as bite rates during grazing forays (define)
-    - Discarded data of incomplete follows or where fish were visably disturbed/fled
-    - Timing of data collection distributed throughout day, 2 dives per site
-    - Define behavioral variables ([table]): feeding rate, bite rate, feeding intensity
-
+## Table 1. Variable definitions
 ```{r, echo=F, message=F}
@@
 
-*Analysis*
-
-* Fish biomass using length-weight conversions (appendix)
-* Site-level behavioral summary stats taken from subset of individuals based on size to avoid bias
-* Biplots w predictors - loess smoothers (switch to something else?)
-* Bite content: calculated proxy for bite content by multiplying estimated bite sizes (based on species and length of each individual fish surveyed, equations from Bruggemann/Mumby) by site level averages of turf algal canopy height to account for potential differences in amount of algae consumed in low vs.
-high canopy sites
-
-## Results and discussion
-
-Feeding rates, time spent grazing, and grazing concentration varied significantly across sites within both *S.
-vetula* and *S.
-viride* populations (Fig.
-2).
-[Fig.
-2A also shows substantially lower feeding rates in most Antiguan (*S.
-vetula* and *S.
-viride*) and Barbudan (*S.
-vetula*) sites as compared standard values used in Caribbean herbivory models - *if I include reference lines...*].
-Differences in site-level feeding rates stemmed from differences in the fraction of time that fish speng grazing (Fig.
-2B) as opposed to differences in bite rates while activitely feeding, which remained relatively constant accross sites (appendix).
-Interspecific differences in feeding rates and grazing intensity reflect the different feeding morphologies/functional roles of *S.
-vetula* and *S.
-viride*.
-
-These behavioral variations may have important ecological consequences and management implications.
-If we use grazing data from one area to predict the impact of a given species in other parts of the region, we may be overlooking important behavioral differences and misrepresenting grazing levels of different scarid populations.
-Feeding rates are used in herbivory models to calculate the area grazed by a given fish community.
-In the Caribbean case, most standard feeding rate equations are obtained from data collected in Bonaire.
-My results suggest that applying Bonaire feeding rates to other Caribbean reef systems may overestimate grazing impact.
-
-Grazing concentration also plays a key role in determining the ecological impact of coral reef herbivores.
-More concentrated grazing may be more effective in removing algae and clearing substrate for growing or newly recruiting corals than bites dispersed throughout the reefscape.
-[**add citations here**]
-
+## 
 ```{r, eval = F, echo=F, message=F, warning=F}
@@
 pca_pred <- prcomp(na.omit(sum_site %>%
-              select(rugosity, ma_cover, ma_canopy, ta_cover, ta_canopy, scar_den, 
-                     scar_bm, carn30_bm, lc_cover)),
+                remove_rownames %>% column_to_rownames(var="site") %>%
+                select(rugosity, ma_cover, ma_canopy, ta_cover, ta_canopy, scar_den, scar_bm, carn30_bm, lc_cover)),
   center = T, scale = T)
@@
 autoplot(pca_pred, data = sum_site, shape = 'island', colour = 'island',
+         label = T,
          loadings = T, loadings.colour = 'gray',
@@
 ## GAMM results (see 5.8.19)
+
+# Supplement
+
+## Sample sizes
+Current data subset: initial phase individuals only, above 15cm (eventual summary subset is 16-24cm...)
+
+```{r}
+sample <- sum_id %>% 
+  filter(phase == "i" & species_code != "rbp" & length_cm <= 30) %>% 
+  group_by(site, species, species_code, island) %>% 
+  summarize(n=n())
+
+sample_vir <- sum_id %>% 
+  filter(phase == "i" & species_code == "stop" & length_cm <= 30) %>% 
+  group_by(site, species, species_code, island) %>% 
+  summarize(n=n())
+sample_vir_sum <- sample_vir %>% 
+  filter(site != "Pallaster West") %>% 
+  group_by(species) %>% 
+  summarize(n_sites = n(),
+            n_total = sum(n),
+            n_mean = mean(n),
+            se = sd(n)/sqrt(n_sites))
+
+sample_vet <- sum_id %>% 
+  filter(phase == "i" & species_code == "qup" & length_cm <= 30) %>% 
+  group_by(site, species, species_code, island) %>% 
+  summarize(n=n())
+sample_vet_sum <- sample_vet %>% 
+  filter(!site %in% c("Pallaster East","Rendezvous","Turtle Bay")) %>%
+  group_by(species) %>% 
+  summarize(n_sites = n(),
+            n_total = sum(n),
+            n_mean = mean(n),
+            se = sd(n)/sqrt(n_sites))
+
+ggplot(sample, aes(x=factor(site, levels = sum_site$site), y=n)) +
+        facet_grid(phase~species) +
+        geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
+        scale_fill_manual(values = c("gray", "slategray2", "black")) +
+        labs(title="Total sample") +
+        theme_bw() +
+        theme(axis.text.x = element_text(angle = 90, hjust = 1))
+```

### 2019-09-18 r/FID.Rmd

@@
 fid.sum <- fid %>%
-  group_by(species, phase) %>%
-  summarize(fid_mean = mean(fid_cm/100),
-            fid_max = max(fid_cm/100)) %>%
-  rename(max = fid_max,
-         mean = fid_mean) %>%
-  melt() %>%
-  rename(fid = value)
+  group_by(species) %>%
+  summarize(n=n(),
+            fid_mean = mean(fid_cm/100),
+            fid_mean_se = sd((fid_cm/100))/sqrt(n),
+            fid_max = max(fid_cm/100))
 
-ggplot(fid, aes(x = species, y = fid_cm)) +
-  geom_boxplot()
-
-ggplot(fid.sum %>% filter(phase == "Initial"), aes(x = species, y = fid, group = variable)) +
-  geom_point(aes(color = variable))
+fid.ch2sum <- fid.sum %>%
+  filter(species %in% c("Queen","Stoplight"))
 ```
@@
 ```{r}
-bump <- FID %>% filter(Bump.distance.cm != 0)
-lmBump <-lm(data=bump,Bump.distance.cm~Spear + Site + Species + Rugosity + Pre.Approach.Activity)
+bump <- fid %>% filter(bump_distance_cm != 0)
+lmBump <-lm(data=bump,bump_distance_cm ~ spear + site + species + rugosity + pre_approach_activity)
 summary(lmBump)
-aov()
+aov(lmBump)
 
-ggplot(bump, aes(x=Spear,y=Bump.distance.cm)) + geom_point(aes(color=Species))
+ggplot(bump, aes(x=spear,y=bump_distance_cm)) + geom_point(aes(color=species))
 ```

### 2019-09-18 r/Workflow.Rmd

---
title: "Workflow"
author: "Molly Wilson"
date: "5/8/2019"
output: html_document
---

Data processing and analaysis workflow for 'Scarid Behavior':

1. **Sites.rmd**: process site-level fish, benthic, and rugosity data.
*Exports*:
  - site.csv
  - sp_site.csv (fish species abundance/biomass data for later adding conspecific metrics)
  - fish.csv (processed fish data for later calculating grazing impacts of a given fish community)

2. **Follows.rmd**: process individual fish-level behavioral data, add site-level descriptors into individual-level and site-level summary datasets.
*Exports*:
  - sum_id.csv (indv. fish-level data, including site characteristics)
  - sum_site.csv (site level summary data)
  - sum_ss.csv (summary data by site and species)
  - sum_ssp.csv (summary data by site, species, and phase)

3. Exploration/analysis files:
  - **Data_prodding.rmd**: exploring relationships among variables, distributions, etc.
  - **Analysis.rmd**: initial boxplots/bargraphs looking at trends among islands/sites, testing ANOVAs/assumptions, predicted grazing impact, bite content calculations
  - **Models.rmd**: PCAs, LMEs, GAMMs (exploratory, no .csv exports)
  - **Ch2_outline_V3**: selected analyses for current manuscript outline

### 2019-09-24 r/Ch2 outline_V3.Rmd
updating stats

@@
 library(tidyverse)
+library(R.utils)
 library(ggplot2)
@@
 
-sum_site <- read.csv(here("data","sum_site.csv"), stringsAsFactors=F) %>%
-  select(-X) %>% 
-  mutate(island = factor(island, levels = c("Bonaire","Antigua","Barbuda"))) %>%
-  arrange(island)
+site <- read.csv(here("data", "site.csv"), strip.white = T, stringsAsFactors = F) %>% select(-X)
 sum_id <- read.csv(here("data","sum_id.csv"), stringsAsFactors=F) %>%
   select(-X) %>% 
-  filter(species_code %in% c("stop","qup")) %>%
+  filter(phase == "i" & 
+           species_code %in% c("stop","qup") & 
+           length_cm <= 30 & length_cm >= 15 & 
+           !(species_code == "stop" & site == "Pallaster West") & 
+           !(species_code == "qup" & site %in% c("Pallaster East","Rendezvous","Turtle Bay"))) %>%
   mutate(island = factor(island, levels = c("Bonaire","Antigua","Barbuda"))) %>%
@@
 
-### data subset: excluding sites with low sample sizes, including only individuals 15-35 cm
-sum_id_sub <- sum_id %>%
-  filter(phase == "i" & length_cm >= 15 & length_cm <= 35 & !(species_code == "stop" & site == "Pallaster West") & !(species_code == "qup" & site %in% c("Pallaster East","Rendezvous","Turtle Bay")))
-
-sum_ssp_sub <- sum_id_sub %>% 
+sum_ssp_sub <- sum_id %>% 
   group_by(species, species_code, site, island) %>%
@@
     n=n(),
-    length_mean=mean(length_cm),
+    length_median=median(length_cm),
     length_se=sd(length_cm)/sqrt(n),
-    g_frac_mean=mean(g_frac),
+    g_frac_median=median(g_frac),
     g_frac_median=median(g_frac),
     g_frac_se=sd(g_frac)/sqrt(n),
-    br_mean=mean(br,na.rm=TRUE),
+    br_median=median(br,na.rm=TRUE),
     br_median=median(br,na.rm=TRUE),
     br_se=sd(br,na.rm=TRUE)/sqrt(n),
-    fr_mean=mean(fr),
+    fr_median=median(fr),
     fr_median=median(fr),
     fr_se=sd(fr)/sqrt(n),
-    for_dur_mean=mean(for_dur,na.rm=TRUE),
+    for_dur_median=median(for_dur,na.rm=TRUE),
     for_dur_median=median(for_dur,na.rm=TRUE),
     for_dur_se=sd(for_dur,na.rm=TRUE)/sqrt(n),
-    for_bites_mean=mean(for_bites,na.rm=TRUE),
+    for_bites_median=median(for_bites,na.rm=TRUE),
     for_bites_median=median(for_bites,na.rm=TRUE),
     for_bites_se=sd(for_bites,na.rm=TRUE)/sqrt(n)
-    )
+    ) %>%
+  left_join(site, by = c("island","site"))
 
@@
 
-## Figuring out sample sizes/site consolidations...
+## Sample sizes (Table S1)
+```{r}
+sample <- sum_id %>%
+  group_by(island, site, species) %>% 
+  summarize(n=n()) %>%
+  arrange(island) %>%
+  rename_all(capitalize)
+kable(sample)
+write.csv(sample, file = here("figs/sample.csv"))
 
-```{r}
-sample_exp <- sum_id %>% 
-  filter(phase == "i" & length_cm >= 15 & length_cm <= 35) %>%
-  group_by(species, species_code, site, island) %>%
-  summarize(
-    n=n())
+sample_vir_sum <- sum_id %>% 
+  filter(species_code == "stop") %>% 
+  group_by(site, species, island) %>% 
+  summarize(n=n()) %>%
+  group_by(species) %>% 
+  summarize(n_sites = n(),
+            n_total = sum(n),
+            n_median = median(n),
+            se = sd(n)/sqrt(n_sites))
 
-ggplot(filter(sum_id, 
-              phase=="i" & length_cm >= 15 & length_cm <= 50 & species_code!="rbp"), 
-       aes(factor(site), fr)) + 
-  geom_violin(aes(fill=island)) +
-  geom_boxplot(width=0.1) + 
-  labs(y = "Feeding Rate (bites/hr)", x = "", title = "Hourly feeding rates (i. phase)") +
-  geom_signif(comparisons = list(c("Antigua", "Barbuda"),c("Antigua","Bonaire"),c("Barbuda","Bonaire")), map_signif_level=TRUE) +
-  facet_grid(.~species) + 
-  stat_n_text(angle = 90) +
-  scale_fill_brewer(palette="Blues") +
-  theme_bw() +
-  theme(strip.text = element_text(face = "italic")) +
-  theme(plot.title = element_text(hjust = 0.5), axis.text.x = element_text(angle = 90, hjust = 1)) 
+sample_vet_sum <- sum_id %>% 
+  filter(species_code == "qup") %>% 
+  group_by(site, species, island) %>% 
+  summarize(n=n()) %>%
+  group_by(species) %>% 
+  summarize(n_sites = n(),
+            n_total = sum(n),
+            n_median = median(n),
+            se = sd(n)/sqrt(n_sites))
+
 ```
@@
 
-## 
-```{r, eval = F, echo=F, message=F, warning=F}
-ggplot(filter(sum_ssp, species_code != "rbp" & phase == "i"), aes(x=factor(site, levels = sum_site$site), y=fr_mean)) +
-    geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
-    geom_errorbar(aes(ymin=fr_mean-fr_se, ymax=fr_mean+fr_se), width=.2,
-                 position=position_dodge(.9)) +
-    facet_grid(~species) +
-    scale_fill_brewer(palette = "Blues") +
-    labs(title="Mean feeding rates of individuals (initial phase)", 
-         y="Feeding rate (bites/min)", 
-         x="Site") +
-    theme_bw() +
-    theme(strip.text = element_text(face = "italic"), 
-          axis.text.x = element_text(angle = 90, hjust = 1), 
-          plot.title = element_text(hjust = 0.5))
+## Testing parametric assumptions
+```{r}
+### Normality
+
+hist(sum_id$fr) # positive skew
+shapiro.test(sum_id$fr) # non-normal
+
+hist(sum_id$br) # positive skew
+shapiro.test(sum_id$br) # non-normal
+
+hist(sum_id$g_frac) # positive skew
+shapiro.test(sum_id$g_frac) # non-normal
+
+hist(sum_id$for_bites) # positive skew
+shapiro.test(sum_id$for_bites) # non-normal
+
+# all positively skewed, non-normally distributed -> should use median, should use Kruskall-Wallis
 ```
 
+## Panel graph of site differences
 ```{r, echo = F, message = F, warning = F}
-# Panel graph of site differences
-
 # published feeding rate (bites/min) for reference lines if needed
@@
 fr_qup <- 
-  ggplot(filter(sum_ssp, species_code == "qup" & phase == "i"), 
-         aes(x=factor(site, levels = sum_site$site), y=fr_mean)) +
+  ggplot(filter(sum_ssp_sub, species_code == "qup"), 
+         aes(x=factor(site, levels = sum_site$site), y=fr_median)) +
   geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
-  geom_errorbar(aes(ymin=fr_mean-fr_se, ymax=fr_mean+fr_se), width=.2,
+  geom_errorbar(aes(ymin=fr_median-fr_se, ymax=fr_median+fr_se), width=.2,
                  position=position_dodge(.9)) +
@@
   #ylim(0, 50) +
-  labs(y = "Feeding rate \n (bites/min)", 
+  labs(y = "Grazing rate \n (bites/min)", 
        title = "S. vetula") +
@@
 fr_stop <- 
-  ggplot(filter(sum_ssp, species_code == "stop" & phase == "i"), 
-         aes(x=factor(site, levels = sum_site$site), y=fr_mean)) +
+  ggplot(filter(sum_ssp_sub, species_code == "stop"), 
+         aes(x=factor(site, levels = sum_site$site), y=fr_median)) +
   geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
-  geom_errorbar(aes(ymin=fr_mean-fr_se, ymax=fr_mean+fr_se), width=.2,
+  geom_errorbar(aes(ymin=fr_median-fr_se, ymax=fr_median+fr_se), width=.2,
                  position=position_dodge(.9)) +
@@
 br_qup <- 
-  ggplot(filter(sum_ssp, species_code == "qup" & phase == "i"), 
-         aes(x=factor(site, levels = sum_site$site), y=br_mean)) +
+  ggplot(filter(sum_ssp_sub, species_code == "qup"), 
+         aes(x=factor(site, levels = sum_site$site), y=br_median)) +
   geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
-  geom_errorbar(aes(ymin=br_mean-br_se, ymax=br_mean+br_se), width=.2,
+  geom_errorbar(aes(ymin=br_median-br_se, ymax=br_median+br_se), width=.2,
                  position=position_dodge(.9)) +
@@
 br_stop <- 
-  ggplot(filter(sum_ssp, species_code == "stop" & phase == "i"), 
-         aes(x=factor(site, levels = sum_site$site), y=br_mean)) +
+  ggplot(filter(sum_ssp_sub, species_code == "stop"), 
+         aes(x=factor(site, levels = sum_site$site), y=br_median)) +
   geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
-  geom_errorbar(aes(ymin=br_mean-br_se, ymax=br_mean+br_se), width=.2,
+  geom_errorbar(aes(ymin=br_median-br_se, ymax=br_median+br_se), width=.2,
                  position=position_dodge(.9)) +
@@
 gfrac_qup <- 
-  ggplot(filter(sum_ssp, species_code == "qup" & phase == "i"), 
-         aes(x=factor(site, levels = sum_site$site), y=g_frac_mean)) +
+  ggplot(filter(sum_ssp_sub, species_code == "qup"), 
+         aes(x=factor(site, levels = sum_site$site), y=g_frac_median)) +
   geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
-  geom_errorbar(aes(ymin=g_frac_mean-g_frac_se, ymax=g_frac_mean+g_frac_se), width=.2,
+  geom_errorbar(aes(ymin=g_frac_median-g_frac_se, ymax=g_frac_median+g_frac_se), width=.2,
                  position=position_dodge(.9)) +
@@
 gfrac_stop <- 
-  ggplot(filter(sum_ssp, species_code == "stop" & phase == "i"), 
-         aes(x=factor(site, levels = sum_site$site), y=g_frac_mean)) +
+  ggplot(filter(sum_ssp_sub, species_code == "stop"), 
+         aes(x=factor(site, levels = sum_site$site), y=g_frac_median)) +
   geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
-  geom_errorbar(aes(ymin=g_frac_mean-g_frac_se, ymax=g_frac_mean+g_frac_se), width=.2,
+  geom_errorbar(aes(ymin=g_frac_median-g_frac_se, ymax=g_frac_median+g_frac_se), width=.2,
                  position=position_dodge(.9)) +
@@
 fi_qup <- 
-  ggplot(filter(sum_ssp, species_code == "qup" & phase == "i"), 
-         aes(x=factor(site, levels = sum_site$site), y=for_bites_mean)) +
+  ggplot(filter(sum_ssp_sub, species_code == "qup"), 
+         aes(x=factor(site, levels = sum_site$site), y=for_bites_median)) +
   geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
-  geom_errorbar(aes(ymin=for_bites_mean-for_bites_se, ymax=for_bites_mean+for_bites_se), width=.2,
+  geom_errorbar(aes(ymin=for_bites_median-for_bites_se, ymax=for_bites_median+for_bites_se), width=.2,
                  position=position_dodge(.9)) +
   scale_fill_brewer(palette = "Blues") +
-  labs(y = "Foray length\n(bites/foray)", 
+  labs(y = "Grazing intensity\n(bites/foray)", 
        x = "", 
@@
 fi_stop <- 
-  ggplot(filter(sum_ssp, species_code == "stop" & phase == "i"), 
-         aes(x=factor(site, levels = sum_site$site), y=for_bites_mean)) +
+  ggplot(filter(sum_ssp_sub, species_code == "stop"), 
+         aes(x=factor(site, levels = sum_site$site), y=for_bites_median)) +
   geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
-  geom_errorbar(aes(ymin=for_bites_mean-for_bites_se, ymax=for_bites_mean+for_bites_se), width=.2,
+  geom_errorbar(aes(ymin=for_bites_median-for_bites_se, ymax=for_bites_median+for_bites_se), width=.2,
                  position=position_dodge(.9)) +
@@
   ggplot(filter(sum_ssp, species_code == "stop" & phase == "i"), 
-         aes(x=factor(site, levels = sum_site$site), y=g_frac_mean)) +
+         aes(x=factor(site, levels = sum_site$site), y=g_frac_median)) +
   geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
@@
 
-grid.arrange(
-  arrangeGrob(fr_qup, fr_stop, gfrac_qup, gfrac_stop, br_qup, br_stop, fi_qup, fi_stop, ncol=2, heights = c(1.5,1.5,1.5,3)), 
-              arrangeGrob(legend), 
-              ncol=2, widths = c(5,1))
-
 g <- arrangeGrob(
@@
 
-The variations in grazing behaviors demonstrated here may be driven by several potential underlying mechanisms.
-Figure 3 shows the relationships between various reef characteristics and scarid feeding rate.
-Because study sites expectedly varied across a number of interrelated fish and benthic dimensions (PCA - appendix?), it is impossible to conclusively establish any explicit mechanisms here.
-Instead I present several hypotheses that may be driving these observed behavioral trends and discuss evidence for and implications of each.
+## Kruskal-Wallis tests
+```{r}
+vir <- sum_id %>% filter(species_code == "stop")
+vet <- sum_id %>% filter(species_code == "qup")
+
+kruskal.test(fr ~ site, data = vet) # p<0.001
+kruskal.test(fr ~ site, data = vir) # p<0.001
+
+kruskal.test(br ~ site, data = vet) # p>0.5 (p=0.63)
+kruskal.test(br ~ site, data = vir) # p>0.5 (p=0.55)
+
+kruskal.test(g_frac ~ site, data = vet) # p<0.001
+kruskal.test(g_frac ~ site, data = vir) # p<0.001
+
+kruskal.test(for_bites ~ site, data = vet) # p<0.001
+kruskal.test(for_bites ~ site, data = vir) # p<0.001
+```
+
+## Loess plots of various drivers
 
 ```{r, echo=F, message=F, warning=F}
-fr_q1 <- ggplot(filter(sum_ssp, species_code == "qup" & phase == "i"), aes(x = scar_bm/1000, y = fr_mean)) +
+fr_q1 <- ggplot(filter(sum_ssp_sub, species_code == "qup"), aes(x = scar_bm/1000, y = fr_median)) +
   geom_smooth(method="loess", span=1) +
@@
 
-fr_q2 <- ggplot(filter(sum_ssp, species_code == "qup" & phase == "i"), aes(x = consp_den, y = fr_mean)) +
+fr_q2 <- ggplot(filter(sum_ssp_sub, species_code == "qup"), aes(x = consp_den, y = fr_median)) +
   geom_smooth(method="loess", span=1) +
@@
 
-fr_q3 <- ggplot(filter(sum_ssp, species_code == "qup" & phase == "i"), aes(x = rugosity, y = fr_mean)) +
+fr_q3 <- ggplot(filter(sum_ssp_sub, species_code == "qup"), aes(x = rugosity, y = fr_median)) +
   geom_smooth(method="loess", span=1) +
@@
 
-fr_q4 <- ggplot(filter(sum_ssp, species_code == "qup" & phase == "i"), aes(x = ta_canopy, y = fr_mean)) +
+fr_q4 <- ggplot(filter(sum_ssp_sub, species_code == "qup"), aes(x = ta_canopy, y = fr_median)) +
   geom_smooth(method="loess", span=1) +
@@
 
-fr_q5 <- ggplot(filter(sum_ssp, species_code == "qup" & phase == "i"), aes(x = ta_cover, y = fr_mean)) +
+fr_q5 <- ggplot(filter(sum_ssp_sub, species_code == "qup"), aes(x = ta_cover, y = fr_median)) +
   geom_smooth(method="loess", span=1) +
@@
 
-fr_q7 <- ggplot(filter(sum_ssp, species_code == "qup" & phase == "i"), aes(x = ma_cover, y = fr_mean)) +
+fr_q7 <- ggplot(filter(sum_ssp_sub, species_code == "qup"), aes(x = ma_cover, y = fr_median)) +
   geom_smooth(method="loess", span=1) +
@@
 
-fr_s1 <- ggplot(filter(sum_ssp, species_code == "stop" & phase == "i"), aes(x = scar_bm/1000, y = fr_mean)) +
+fr_s1 <- ggplot(filter(sum_ssp_sub, species_code == "stop"), aes(x = scar_bm/1000, y = fr_median)) +
   geom_smooth(method="loess", span=1) +
@@
 
-fr_s2 <- ggplot(filter(sum_ssp, species_code == "stop" & phase == "i"), aes(x = consp_den, y = fr_mean)) +
+fr_s2 <- ggplot(filter(sum_ssp_sub, species_code == "stop"), aes(x = consp_den, y = fr_median)) +
   geom_smooth(method="loess", span=1) +
@@
 
-fr_s3 <- ggplot(filter(sum_ssp, species_code == "stop" & phase == "i"), aes(x = rugosity, y = fr_mean)) +
+fr_s3 <- ggplot(filter(sum_ssp_sub, species_code == "stop"), aes(x = rugosity, y = fr_median)) +
   geom_smooth(method="loess", span=1) +
@@
 
-fr_s4 <- ggplot(filter(sum_ssp, species_code == "stop" & phase == "i"), aes(x = ta_canopy, y = fr_mean)) +
+fr_s4 <- ggplot(filter(sum_ssp_sub, species_code == "stop"), aes(x = ta_canopy, y = fr_median)) +
   geom_smooth(method="loess", span=1) +
@@
 
-fr_s5 <- ggplot(filter(sum_ssp, species_code == "stop" & phase == "i"), aes(x = ta_cover, y = fr_mean)) +
+fr_s5 <- ggplot(filter(sum_ssp_sub, species_code == "stop"), aes(x = ta_cover, y = fr_median)) +
   geom_smooth(method="loess", span=1) +
@@
 
-fr_s7 <- ggplot(filter(sum_ssp, species_code == "stop" & phase == "i"), aes(x = ma_cover, y = fr_mean)) +
+fr_s7 <- ggplot(filter(sum_ssp_sub, species_code == "stop"), aes(x = ma_cover, y = fr_median)) +
   geom_smooth(method="loess", span=1) +
@@
 
-*Bite content hypothesis.* Differences in scarid grazing preferences can be driven by (local differences in) nutrient availability and distribution (Carlson, Tootell, Mumby, etc.).
-Algal communities on Caribbean reefs can vary greatly in terms of canopy height, percent cover, species composition, and nutrient content ().
-
-(no conclusive trends, but FR tended to decrease with increasing EAM canopy height and to some extent EAM and macroalgal percent cover)
-
-In higher canopy sites, scarids may be obtaining more biomass of food per bite than in areas with heavily cropped algae, requiring them to take fewer bites to obtain the same nutritional intake.
-
-Implications: 
--maybe better metric than bite/feeding rate
--reduced grazing impact if not clearing substrate
--reinforcing loop
-
-```{r, echo=F, message=F, warning=F}
+```{r, eval = F, echo = F, message=F, warning=F}
 # grazing impact (vg ~ volume grazed) of fish from behavioral surveys
@@
     n=n(),
-    vg_mean=mean(vg),
+    vg_median=median(vg),
     vg_se=sd(vg)/sqrt(n),
-    vg_capped_mean=mean(vg_capped),
+    vg_capped_median=median(vg_capped),
     vg_capped_se=sd(vg_capped)/sqrt(n),
-    fr_mean=mean(fr)
+    fr_median=median(fr)
     ) %>% 
@@
 # keep this to get legend for panel
-vg_facet <- ggplot(filter(sum_ssp_vg, species_code != "rbp" & phase == "i"), aes(x=factor(site, levels = sum_site$site), y=vg_mean)) +
+vg_facet <- ggplot(filter(sum_ssp_vg, species_code != "rbp" & phase == "i"), aes(x=factor(site, levels = sum_site$site), y=vg_median)) +
     geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
-    geom_errorbar(aes(ymin=vg_mean-vg_se, ymax=vg_mean+vg_se), width=.2,
+    geom_errorbar(aes(ymin=vg_median-vg_se, ymax=vg_median+vg_se), width=.2,
                  position=position_dodge(.9)) +
@@
     scale_fill_brewer(palette = "Blues") +
-    labs(title="Mean grazing impact by volume (initial phase)", 
+    labs(title="median grazing impact by volume (initial phase)", 
          y="Grazing impact (cm3/hr)", 
@@
 vg_qup <- ggplot(filter(sum_ssp_vg, species_code == "qup" & phase == "i"), 
-                 aes(x=factor(site, levels = sum_site$site), y=vg_mean)) +
+                 aes(x=factor(site, levels = sum_site$site), y=vg_median)) +
     geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
-    geom_errorbar(aes(ymin=vg_mean-vg_se, ymax=vg_mean+vg_se), width=.2,
+    geom_errorbar(aes(ymin=vg_median-vg_se, ymax=vg_median+vg_se), width=.2,
                  position=position_dodge(.9)) +
@@
 vg_stop <- ggplot(filter(sum_ssp_vg, species_code == "stop" & phase == "i"), 
-                  aes(x=factor(site, levels = sum_site$site), y=vg_mean)) +
+                  aes(x=factor(site, levels = sum_site$site), y=vg_median)) +
     geom_bar(stat = "identity", aes(fill = island), position = "dodge") +    
-    geom_errorbar(aes(ymin=vg_mean-vg_se, ymax=vg_mean+vg_se), width=.2,
+    geom_errorbar(aes(ymin=vg_median-vg_se, ymax=vg_median+vg_se), width=.2,
                  position=position_dodge(.9)) +
@@
 
-```{r, eval = F, echo = F, message = F}
-ggplot(filter(sum_ssp_vg, species_code != "rbp" & phase == "i"), aes(x = fr_mean, y = vg_mean)) +
-  geom_point(shape=18) +
-  facet_grid(~species) +
-  labs(y = "Grazing impact (cm3/hr)",
-       x = "Average feeding rate (bites/min)"
-       title = "") +
-  theme_bw() + 
-  theme(plot.title = element_text(hjust = 0.5),
-        )
-```
-
-*Social feeding hypothesis.* more time must be spent vigilant in smaller groups, potential compounding effect with overfishing/loss of herbivory (e.g. Gil)
-
-*Predation risk hypothesis.* Previous work has documented the suppressive effect of predator presence on herbivore feeding rates (Madin, Catano, McCauley?). 
-- These studies have focused on natural predators (decoys or piscivore populations)
-- While I found no clear relationship between large (>30cm) piscivore biomass and feeding trends (appendix), risk of predation from spearfishers may play a larger role in determining scarid grazing behavior at the reefs studied here.
-All Antiguan and Barbudan sites experience some degree of spearfishing pressure, while spearfishing has been banned in Bonaire since the 1970s.
-The 15-30cm individuals used in my behavioral observations are approaching size refuge from most natural predators, but would be likely targets for spearfishers.
-Both *S.
-vetula* and *S.
-viride* are commonly harvested species in Antigua and Barbuda.
-
-- like with the social feeding hypothesis, spearfishing may have double impact of reducing fish and reducing grazing impact of remaining fish
-
-- often interaction of rugosity (e.g. Madin, stay near structure)
-(Incorporate rugosity here - Rogers, Madin, Catano). Did not find strong evidence for this here, but carnivore biomass was strongly correllated with herbivore biomass and rugosity.
-
-## Management implications + future research needs
-
-* Potential reinforcing feedback loops in degraded reefs if high algal/low scarid reefs also have less grazing impact per scarid
-
-* Limitations to this study: all these potential drivers/mechanisms are understandably correlated in observational work
-    * Room for experimental work to unpack/distinguish drivers
-* Need for incorporation of behavior in coral reef grazing models, especially those used to set targets for minimum herbivore biomass levels, because these may look very different in different reef conditions.
-Current paradigm of constant species-level grazing behaviors may be inaccurate.
-In Caribbean case, most models use Bonaire data which may not be representative of dynamics on more degraded Caribbean reefs.
-* Need for framework to help incorporate behavioral effects into ecosystem management
-* Conservation + mgmt implications: multiple human impacts on reefs may affect herbivore behavior in different ways, spec.
-fishing + habitat degradation - which may have reinforcing/compounding negative effects on reef health
-
-In conclusion...
-
-# Potential additions:
-
 ## PCA
 ```{r, echo=F, message=F, warning=F}
-pca_pred <- prcomp(na.omit(sum_site %>%
+pca_pred <- prcomp(na.omit(site %>%
                 remove_rownames %>% column_to_rownames(var="site") %>%
@@
 
-sum_site <- sum_site %>% 
-  bind_cols(as.data.frame(pca_pred$x) %>% 
-           select(PC1,PC2)) %>%
-  rename(pc1_all=PC1,
-         pc2_all=PC2)
-
 sum_id_pca <- sum_id %>% 
   left_join(select(sum_site,site,pc1_all,pc2_all), by = "site")
-sum_ssp_pca <- sum_ssp %>%
+sum_ssp_pca <- sum_ssp_sub %>%
   left_join(select(sum_site,site,pc1_all,pc2_all), by = "site")
@@
 
+Predicting feeding rate (site-level medians, island as random effect)
+```{r, echo=F, message=F, warning=F}
+m_data <- sum_ssp_pca
+
+gamm1 <- gamm(fr_median ~ species_code + s(length_mean, k=3) + s(pc1_all, k=3) + s(pc2_all, k=3), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           data = m_data
+           )
+g1 <- c("g1", "Species + Size + PC1 + PC2", AICc(gamm1$lme), summary(gamm1$gam)$r.sq)
+
+gamm2 <- gamm(fr_median ~ species_code + s(length_mean, k=3) + s(pc1_all, k=3), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           data = m_data
+           )
+g2 <- c("g2", "Species + Size + PC1", AICc(gamm2$lme), summary(gamm2$gam)$r.sq)
+
+gamm3 <- gamm(fr_median ~ species_code + s(length_mean, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3) + s(pc2_all, by = factor(species_code), k=3),
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           data = m_data
+           )
+g3 <- c("g3", "Species + Size~Species + PC1~Species + PC2~Species", AICc(gamm3$lme), summary(gamm3$gam)$r.sq)
+
+# gamm4 <- gamm(fr_median ~ species_code + s(length_mean, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3) + s(pc2_all, k=3), 
+#            family = gaussian(link=identity),
+#            random = list(island=~1),
+#            data = m_data
+#            )
+# g4 <- c("g4", "Species + Size~Species + PC1~Species + PC2", AICc(gamm4$lme), summary(gamm4$gam)$r.sq)
+
+# gamm5 <- gamm(fr_median ~ species + s(length_mean, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3), 
+#            family = gaussian(link=identity),
+#            random = list(island=~1),
+#            data = m_data
+#            )
+# g5 <- c("g5", "Species + Size~Species + PC1~Species", AICc(gamm5$lme), summary(gamm5$gam)$r.sq)
+
+gamm6 <- gamm(fr_median ~ s(length_mean, k=3) + s(pc1_all, by = factor(species_code), k=3), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           data = m_data
+           )
+g6 <- c("g6", "Size + PC1~Species", AICc(gamm6$lme), summary(gamm6$gam)$r.sq)
+
+gamm7 <- gamm(fr_median ~ s(length_mean, by = factor(species_code), k=3) + s(pc1_all, k=3), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           data = m_data
+           )
+g7 <- c("g7", "Size~Species + PC1", AICc(gamm7$lme), summary(gamm7$gam)$r.sq)
+
+gamm8 <- gamm(fr_median ~ species + s(length_mean, by = factor(species_code), k=3) + s(pc1_all, k=3), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           data = m_data
+           )
+g8 <- c("g8", "Species + Size~Species + PC1", AICc(gamm8$lme), summary(gamm8$gam)$r.sq)
+
+gamm9 <- gamm(fr_median ~ species + s(length_mean, k=3) + s(pc1_all, by = factor(species_code), k=3), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           data = m_data
+           )
+g9 <- c("g9", "Species + Size + PC1~Species", AICc(gamm9$lme), summary(gamm9$gam)$r.sq)
+
+gamm10 <- gamm(fr_median ~ species + s(pc1_all, by = factor(species_code), k=3), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           data = m_data
+           )
+g10 <- c("g10", "Species + PC1~Species", AICc(gamm10$lme), summary(gamm10$gam)$r.sq)
+
+gamm11 <- gamm(fr_median ~ s(length_mean, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           data = m_data
+           )
+g11 <- c("g11", "Size~Species + PC1~Species", AICc(gamm11$lme), summary(gamm11$gam)$r.sq)
+
+gamm12 <- gamm(fr_median ~ species_code + s(pc1_all, k=3), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           data = m_data
+           )
+g12 <- c("g12", "Species + PC1", AICc(gamm12$lme), summary(gamm12$gam)$r.sq)
+
+gtab_fr_site <- as.data.frame(rbind(g1,g2,g4,g5,g6,g7,g8,g9,g10,g11,g12)) %>% setNames(c("g","Model", "AICc", "Adj. R2")) %>% arrange(AICc)
+
+# summary + plots from best models
+summary(gamm12$gam)
+plot(gamm12$gam,pages=1)
+```
+
+Predicting feeding intensity (site-level medians, island as random effect)
+```{r, echo=F, message=F, warning=F}
+m_data <- sum_ssp_pca
+
+gamm1 <- gamm(for_bites_median ~ species_code + s(length_mean, k=3) + s(pc1_all, k=3) + s(pc2_all, k=3), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           data = m_data
+           )
+g1 <- c("g1", "Species + Size + PC1 + PC2", AICc(gamm1$lme), summary(gamm1$gam)$r.sq)
+
+gamm2 <- gamm(for_bites_median ~ species_code + s(length_mean, k=3) + s(pc1_all, k=3), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           data = m_data
+           )
+g2 <- c("g2", "Species + Size + PC1", AICc(gamm2$lme), summary(gamm2$gam)$r.sq)
+
+gamm3 <- gamm(for_bites_median ~ species_code + s(length_mean, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3) + s(pc2_all, by = factor(species_code), k=3),
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           data = m_data
+           )
+g3 <- c("g3", "Species + Size~Species + PC1~Species + PC2~Species", AICc(gamm3$lme), summary(gamm3$gam)$r.sq)
+
+gamm4 <- gamm(for_bites_median ~ species_code + s(length_mean, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3) + s(pc2_all, k=3), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           data = m_data
+           )
+g4 <- c("g4", "Species + Size~Species + PC1~Species + PC2", AICc(gamm4$lme), summary(gamm4$gam)$r.sq)
+
+gamm5 <- gamm(for_bites_median ~ species + s(length_mean, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           data = m_data
+           )
+g5 <- c("g5", "Species + Size~Species + PC1~Species", AICc(gamm5$lme), summary(gamm5$gam)$r.sq)
+
+gamm6 <- gamm(for_bites_median ~ s(length_mean, k=3) + s(pc1_all, by = factor(species_code), k=3), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           data = m_data
+           )
+g6 <- c("g6", "Size + PC1~Species", AICc(gamm6$lme), summary(gamm6$gam)$r.sq)
+
+gamm7 <- gamm(for_bites_median ~ s(length_mean, by = factor(species_code), k=3) + s(pc1_all, k=3), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           data = m_data
+           )
+g7 <- c("g7", "Size~Species + PC1", AICc(gamm7$lme), summary(gamm7$gam)$r.sq)
+
+gamm8 <- gamm(for_bites_median ~ species + s(length_mean, by = factor(species_code), k=3) + s(pc1_all, k=3), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           data = m_data
+           )
+g8 <- c("g8", "Species + Size~Species + PC1", AICc(gamm8$lme), summary(gamm8$gam)$r.sq)
+
+gamm9 <- gamm(for_bites_median ~ species + s(length_mean, k=3) + s(pc1_all, by = factor(species_code), k=3), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           data = m_data
+           )
+g9 <- c("g9", "Species + Size + PC1~Species", AICc(gamm9$lme), summary(gamm9$gam)$r.sq)
+
+gamm10 <- gamm(for_bites_median ~ species + s(pc1_all, by = factor(species_code), k=3), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           data = m_data
+           )
+g10 <- c("g10", "Species + PC1~Species", AICc(gamm10$lme), summary(gamm10$gam)$r.sq)
+
+gamm11 <- gamm(for_bites_median ~ s(length_mean, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           data = m_data
+           )
+g11 <- c("g11", "Size~Species + PC1~Species", AICc(gamm11$lme), summary(gamm11$gam)$r.sq)
+
+gamm12 <- gamm(for_bites_median ~ species_code + s(pc1_all, k=3), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           data = m_data
+           )
+g12 <- c("g12", "Species + PC1", AICc(gamm12$lme), summary(gamm12$gam)$r.sq)
+
+gtab_fi_site <- as.data.frame(rbind(g1,g2,g3,g4,g5,g6,g7,g8,g9,g10,g11,g12)) %>% setNames(c("g","Model", "AICc", "Adj. R2")) %>% arrange(AICc)
+
+# plots from best models
+plot(gamm12$gam,pages=1)
+plot(gamm10$gam,pages=1)
+
+summary(gamm12$gam)
+```
+
 # Supplement
 
-## Sample sizes
-Current data subset: initial phase individuals only, above 15cm (eventual summary subset is 16-24cm...)
-
-```{r}
-sample <- sum_id %>% 
-  filter(phase == "i" & species_code != "rbp" & length_cm <= 30) %>% 
-  group_by(site, species, species_code, island) %>% 
-  summarize(n=n())
-
-sample_vir <- sum_id %>% 
-  filter(phase == "i" & species_code == "stop" & length_cm <= 30) %>% 
-  group_by(site, species, species_code, island) %>% 
-  summarize(n=n())
-sample_vir_sum <- sample_vir %>% 
-  filter(site != "Pallaster West") %>% 
-  group_by(species) %>% 
-  summarize(n_sites = n(),
-            n_total = sum(n),
-            n_mean = mean(n),
-            se = sd(n)/sqrt(n_sites))
-
-sample_vet <- sum_id %>% 
-  filter(phase == "i" & species_code == "qup" & length_cm <= 30) %>% 
-  group_by(site, species, species_code, island) %>% 
-  summarize(n=n())
-sample_vet_sum <- sample_vet %>% 
-  filter(!site %in% c("Pallaster East","Rendezvous","Turtle Bay")) %>%
-  group_by(species) %>% 
-  summarize(n_sites = n(),
-            n_total = sum(n),
-            n_mean = mean(n),
-            se = sd(n)/sqrt(n_sites))
-
-ggplot(sample, aes(x=factor(site, levels = sum_site$site), y=n)) +
-        facet_grid(phase~species) +
-        geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
-        scale_fill_manual(values = c("gray", "slategray2", "black")) +
-        labs(title="Total sample") +
-        theme_bw() +
-        theme(axis.text.x = element_text(angle = 90, hjust = 1))
-```
+### loess plots for additional drivers?

### 2019-10-01 r/Ch2 outline_V3.Rmd
updates

@@
 library(knitr)
+library(factoextra)
 
@@
 Variable <- c('Time spent grazing','Bite rate','Feeding rate','Foray size')
-Definition <- c('Proportion of time spent grazing during follow','Bite rate while actively feeding','Bite rate over total follow time','Number of consecutive bites taken during single grazing foray')
+Definition <- c('Proportion of time spent grazing during follow','Bite rate while actively Grazing','Bite rate over total follow time','Number of consecutive bites taken during single grazing foray')
 Unit <- c('fraction','bites/second','bites/minute','bites/foray')
@@
 ```{r, echo = F, message = F, warning = F}
-# published feeding rate (bites/min) for reference lines if needed
+# published Feeding rate (bites/min) for reference lines if needed
 ref_qup15 <- (3329-33*15)/60
@@
   #ylim(0, 50) +
-  labs(y = "Grazing rate \n (bites/min)", 
+  labs(y = "Feeding rate \n (bites/min)", 
        title = "S. vetula") +
@@
 
-autoplot(pca_pred, data = sum_site, shape = 'island', colour = 'island',
+autoplot(pca_pred, data = site, shape = 'island', colour = 'island',
          label = T,
@@
   left_join(select(sum_site,site,pc1_all,pc2_all), by = "site")
+```
+
+### Better PCA?
+```{r, echo=F, message=F, warning=F}
+pca_data <- na.omit(site %>%
+  select(rugosity, ma_cover, ma_canopy, ta_cover, ta_canopy, lc_cover, scar_bm, scar_den, carn30_bm))
+names(pca_data)[1:9] =
+  c("rugosity", "macroalgal cover", "macroalgal canopy", "turf cover", "turf canopy", "coral cover", "scarid biomass", "scarid density", "predator biomass")
+
+pca_pred <- prcomp(pca_data, center = T, scale = T)
+
+fviz_pca_biplot(pca_pred,
+                col.ind = site$island,
+                mean.point = F,
+                label = "var",
+                repel = T
+                ) +
+  scale_shape_manual(values=c(15,16,17)) +
+  scale_color_manual(values = c("black","black","black")) +
+  labs(x = "PC1 (55.6%)", y = "PC2 (24.5%)", title = "") +
+  theme_bw() +
+  theme(legend.title = element_blank())
+
+ggsave(here("figs/pca_pred.png"), width = 7, height = 5)
+
 ```
@@
 
-Predicting feeding intensity (site-level medians, island as random effect)
+Predicting Grazing intensity (site-level medians, island as random effect)
 ```{r, echo=F, message=F, warning=F}
@@
 
+### Predator gape size
+
+From Mumby et al.:
+Grouper *E. striatus*, gape width (G): G = -0.5998 + (0.167 × TL)
+```{r}
+g30_mum <- -0.5998 + (0.167 * 30) # 4
+g40_mum <- -0.5998 + (0.167 * 40) # 6
+```
+
+From Goatley & Bellwood:
+Horizontal gape: G = -0.002 * SL + 6.421
+```{r}
+g30_goat <- -0.002 * 30 + 6.421
+g40_goat <- -0.002 * 40 + 6.421
+```
+
+Dorsoventral height (Rakitin & Kramer 1996):
+S. viride: 15cm FL > 4.5cm body depth
+S. vetula: 15cm FL > 4.5cm body depth
+
 ### loess plots for additional drivers?

### 2019-10-01 r/pca_code.R

@@ -1,0 +1,40 @@
+# PCA options!
+
+### loading any data file with variables as columns/observations as rows
+site <- read.csv(here("data", "site.csv"), strip.white = T, stringsAsFactors = F)
+
+### set up dataframe with only variables that you want to include in the PCA 
+pca_data <- site %>%
+  # so that row/column names are ready to use as labels, if so desired:
+  remove_rownames %>% column_to_rownames(var = "site") %>%
+  select(rugosity, ma_cover, ma_canopy, ta_cover, ta_canopy, lc_cover, scar_bm, scar_den, carn30_bm) %>%
+  na.omit() %>%
+  # note: fviz_pca_biplot can handle names with spaces (e.g., "macroagal cover"), but I've had issues with spaces and autoplot() 
+  setNames(c("rugosity", "macroalgal_cover", "macroalgal_canopy", "turf_cover", "turf_canopy", "coral_cover", "scarid_biomass", "scarid_density", "predator_biomass"))
+
+### this does actual PCA
+pca_pred <- prcomp(pca_data, center = T, scale = T)
+
+### option A for visualization (this one has variables labeled but not sites)
+library(factoextra)
+fviz_pca_biplot(pca_pred,
+                col.ind = site$island, 
+                mean.point = F,
+                label = "var",
+                repel = T) +
+  scale_shape_manual(values=c(15,16,17)) + # changing shape of points by group
+  scale_color_manual(values = c("black","black","black")) +
+  labs(x = "PC1 (55.6%)", y = "PC2 (24.5%)", title = "") +
+  theme_bw() +
+  theme(legend.title = element_blank())
+
+### option B for visualization (this one has sites labeled)
+library(ggfortify)
+autoplot(pca_pred, data = site, shape = 'island', colour = 'island',
+         label = T,
+         loadings = T, loadings.colour = 'gray',
+         loadings.label = T, loadings.label.size = 3, loadings.label.colour = 'black',
+         loadings.label.repel = T) +
+  scale_color_manual(values = c("cadetblue3","cadetblue3","cadetblue3")) +
+  theme_bw() +
+  theme(legend.title = element_blank())

### 2019-12-06 r/Paper_analyses.Rmd
grazing intensity + pca loadings

--- 2019-10-01 r/Ch2 outline_V3.Rmd
@@
 ---
-title: "Variance in scarid grazing behavior across Caribbean reef sites"
-author: "Molly Wilson"
-date: "3/1/2019"
-output: html_document
+title: 'Variance in scarid grazing behavior across Caribbean reef sites'
+author: "*Compiled on `r date()` by Molly Wilson*"
+output: 
+  html_document:
+    toc: true
+    code_folding: hide
+    number_sections: false
+    theme: cerulean
+    highlight: haddock
+
 ---
 
-# Variance in scarid grazing behavior across Caribbean reef sites
-### Formatted as: Coral Reefs note (<2900 words)
+*Formatted as: Coral Reefs note*
 
-```{r, echo=F, message=F, warning=F}
+```{r, message=F, warning=F}
 library(here)
@@
 library(ggfortify)
-library(ggsignif) # boxplot stats
-library(EnvStats) # boxplot stats
 library(mgcv)
@@
 
-site <- read.csv(here("data", "site.csv"), strip.white = T, stringsAsFactors = F) %>% select(-X)
+sum_site <- read.csv(here("data", "site.csv"), strip.white = T, stringsAsFactors = F) %>% select(-X)
 sum_id <- read.csv(here("data","sum_id.csv"), stringsAsFactors=F) %>%
@@
            species_code %in% c("stop","qup") &
+           ### size subset 15 - 30cm
            length_cm <= 30 & length_cm >= 15 &
+           ### excluding sites with less than 7 observations within size subset range
            !(species_code == "stop" & site == "Pallaster West") & 
@@
     n=n(),
+    length_mean=mean(length_cm),
     length_median=median(length_cm),
@@
     ) %>%
-  left_join(site, by = c("island","site"))
+  left_join(sum_site, by = c("island","site"))
 
@@
 
-## Sample sizes (Table S1)
-```{r}
+## Preliminary investigations
+
+### Sample sizes by site and species
+Currently omitting sites with fewer than 7 observations that fall within 15-30cm size subset
+```{r, message=F}
 sample <- sum_id %>%
@@
 
-## Table 1. Variable definitions
+### Variable definitions
 ```{r, echo=F, message=F}
@@
 
-## Testing parametric assumptions
-```{r}
-### Normality
-
+### Testing parametric assumptions
+Testing for heteroscedasticity first
+```{r, message=F}
 hist(sum_id$fr) # positive skew
@@
 ```
+All behavioral variables are positively skewed, non-normally distributed -> should use median values and Kruskall-Wallis instead of ANOVA for site-level differences
 
-## Panel graph of site differences
-```{r, echo = F, message = F, warning = F}
+## Inter-site differences
+
+Panel graph of behavioral variables across sites:
+
+```{r, message = F, warning = F}
 # published Feeding rate (bites/min) for reference lines if needed
@@
 legend_plot <- 
-  ggplot(filter(sum_ssp, species_code == "stop" & phase == "i"), 
+  ggplot(filter(sum_ssp_sub, species_code == "stop"), 
          aes(x=factor(site, levels = sum_site$site), y=g_frac_median)) +
@@
 g <- arrangeGrob(
-  arrangeGrob(fr_qup, fr_stop, gfrac_qup, gfrac_stop, br_qup, br_stop, fi_qup, fi_stop, ncol=2, heights = c(1,1,1,1.5)), 
+  arrangeGrob(fr_qup, fr_stop, gfrac_qup, gfrac_stop, br_qup, br_stop, fi_qup, fi_stop, ncol=2, heights = c(1,1,1,2)), 
               arrangeGrob(legend), 
               ncol=2, widths = c(5,1))
+plot(g)
 
@@
 
-## Kruskal-Wallis tests
+Kruskal-Wallis tests to look at significance of inter-site variance for different behavioral variables:
+
+Feeding rate:
 ```{r}
@@
 kruskal.test(fr ~ site, data = vir) # p<0.001
+```
 
+Bite rate:
+```{r}
 kruskal.test(br ~ site, data = vet) # p>0.5 (p=0.63)
 kruskal.test(br ~ site, data = vir) # p>0.5 (p=0.55)
+```
 
+Fraction of time grazing:
+```{r}
 kruskal.test(g_frac ~ site, data = vet) # p<0.001
 kruskal.test(g_frac ~ site, data = vir) # p<0.001
+```
 
+Grazing intensity:
+```{r}
 kruskal.test(for_bites ~ site, data = vet) # p<0.001
@@
 
+All significant (p < 0.001) except for fraction of time spent grazing.
+
 ## Loess plots of various drivers
 
-```{r, echo=F, message=F, warning=F}
+```{r, message=F, warning=F}
 fr_q1 <- ggplot(filter(sum_ssp_sub, species_code == "qup"), aes(x = scar_bm/1000, y = fr_median)) +
@@
 
-```{r, eval = F, echo = F, message=F, warning=F}
-# grazing impact (vg ~ volume grazed) of fish from behavioral surveys
-sum_id_vg <- sum_id %>%
-  filter(species_code != "rbp") %>%
-  mutate(
-    bs = ifelse(species_code == "stop", 5.257*10^-4*length_cm^2, 4.013*10^-4*length_cm^2), # adding bite size equations from Mumby
-    bh_capped = ifelse(ta_canopy < 2*sqrt(bs)/(22/7), ta_canopy, 2*sqrt(bs)/(22/7)), # if ta_canopy exceeds diameter of bite size, bite height is capped at bite size
-    bv = bs*ta_canopy,
-    bv_capped = bs*bh_capped,
-    vg = fr*bv,
-    vg_capped = fr*bv_capped,
-    island = factor(island, levels = c("Bonaire","Antigua","Barbuda"))
-) %>%
-  arrange(island)
+## PCA
 
-# summing by site, species, and phase
-sum_ssp_vg <- sum_id_vg %>% 
-  filter(length_cm >= 15 & length_cm <= 30) %>%
-  group_by(site, ta_canopy, island, species, species_code, phase) %>% 
-  summarize(
-    n=n(),
-    vg_median=median(vg),
-    vg_se=sd(vg)/sqrt(n),
-    vg_capped_median=median(vg_capped),
-    vg_capped_se=sd(vg_capped)/sqrt(n),
-    fr_median=median(fr)
-    ) %>% 
-  select(-n)
-
-# keep this to get legend for panel
-vg_facet <- ggplot(filter(sum_ssp_vg, species_code != "rbp" & phase == "i"), aes(x=factor(site, levels = sum_site$site), y=vg_median)) +
-    geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
-    geom_errorbar(aes(ymin=vg_median-vg_se, ymax=vg_median+vg_se), width=.2,
-                 position=position_dodge(.9)) +
-    facet_grid(~species) +
-    scale_fill_brewer(palette = "Blues") +
-    labs(title="median grazing impact by volume (initial phase)", 
-         y="Grazing impact (cm3/hr)", 
-         x="Site") +
-    theme_bw() +
-    theme(strip.text = element_text(face = "italic"), 
-          axis.text.x = element_text(angle = 45, hjust = 1), 
-          legend.title = element_blank())
-
-vg_qup <- ggplot(filter(sum_ssp_vg, species_code == "qup" & phase == "i"), 
-                 aes(x=factor(site, levels = sum_site$site), y=vg_median)) +
-    geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
-    geom_errorbar(aes(ymin=vg_median-vg_se, ymax=vg_median+vg_se), width=.2,
-                 position=position_dodge(.9)) +
-  scale_fill_brewer(palette = "Blues") +
-  labs(y = expression('Grazing impact (cm'^3*'/hr)'), 
-       x = "", 
-       title = "S. vetula") +
-  theme_bw() + 
-  theme(plot.margin = margin(1, 0, 0, 5, unit = "pt"),
-        axis.title.y = element_text(size=10),
-        axis.text.x = element_text(angle = 45, hjust = 1, size=7),
-        legend.position = "none",
-        plot.title = element_text(hjust = 0.5, face = "italic")
-        )
-
-vg_stop <- ggplot(filter(sum_ssp_vg, species_code == "stop" & phase == "i"), 
-                  aes(x=factor(site, levels = sum_site$site), y=vg_median)) +
-    geom_bar(stat = "identity", aes(fill = island), position = "dodge") +    
-    geom_errorbar(aes(ymin=vg_median-vg_se, ymax=vg_median+vg_se), width=.2,
-                 position=position_dodge(.9)) +
-  scale_fill_brewer(palette = "Blues") +
-  labs(y = "",
-       x = "", 
-       title = "S. viride") +
-  theme_bw() + 
-  theme(plot.margin = margin(1, 1, 0, 0, unit = "pt"),
-        axis.text.x = element_text(angle = 45, hjust = 1, size=7),
-        legend.position = "none",
-        plot.title = element_text(hjust = 0.5, face = "italic")
-        )
-
-get_legend<-function(myggplot){
-  tmp <- ggplot_gtable(ggplot_build(myggplot))
-  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
-  legend <- tmp$grobs[[leg]]
-  return(legend)
-  }
-legend <- get_legend(vg_facet)
-
-g <- grid.arrange(
-  arrangeGrob(vg_qup, vg_stop, ncol=2), 
-              arrangeGrob(legend), 
-              ncol=2, widths = c(5,1), heights = c(1))
-
-ggsave(here("figs/bc_bar_panel.png"),g)
-```
-
-## PCA
-```{r, echo=F, message=F, warning=F}
-pca_pred <- prcomp(na.omit(site %>%
-                remove_rownames %>% column_to_rownames(var="site") %>%
-                select(rugosity, ma_cover, ma_canopy, ta_cover, ta_canopy, scar_den, scar_bm, carn30_bm, lc_cover)),
-  center = T, scale = T)
-
-autoplot(pca_pred, data = site, shape = 'island', colour = 'island',
-         label = T,
-         loadings = T, loadings.colour = 'gray',
-         loadings.label = T, loadings.label.size = 3, loadings.label.colour = 'black',
-            loadings.label.repel = T) +
-    scale_color_manual(values = c("cadetblue3","cadetblue3","cadetblue3")) +
-    theme_bw() +
-    theme(legend.title = element_blank())
-ggsave(here("figs/pca_pred.png"), width = 8, height = 5)
-
-sum_id_pca <- sum_id %>% 
-  left_join(select(sum_site,site,pc1_all,pc2_all), by = "site")
-sum_ssp_pca <- sum_ssp_sub %>%
-  left_join(select(sum_site,site,pc1_all,pc2_all), by = "site")
-```
-
-### Better PCA?
-```{r, echo=F, message=F, warning=F}
-pca_data <- na.omit(site %>%
+```{r, message=F, warning=F}
+pca_data <- na.omit(sum_site %>%
   select(rugosity, ma_cover, ma_canopy, ta_cover, ta_canopy, lc_cover, scar_bm, scar_den, carn30_bm))
@@
 pca_pred <- prcomp(pca_data, center = T, scale = T)
+print(pca_pred)
+pca_var <- get_pca_var(pca_pred)
+print(pca_var$contrib)
 
 fviz_pca_biplot(pca_pred,
-                col.ind = site$island,
+                col.ind = sum_site$island,
                 mean.point = F,
@@
 ggsave(here("figs/pca_pred.png"), width = 7, height = 5)
+```
 
+```{r, message=F}
+sum_site_pca <- sum_site %>% 
+  bind_cols(as.data.frame(pca_pred$x) %>% 
+           select(PC1,PC2)) %>%
+  rename(pc1_all=PC1,
+         pc2_all=PC2)
+
+sum_id_pca <- sum_id %>% 
+  left_join(select(sum_site_pca,site,pc1_all,pc2_all), by = "site")
+sum_ssp_pca <- sum_ssp_sub %>%
+  left_join(select(sum_site_pca,site,pc1_all,pc2_all), by = "site")
 ```
@@
 
+Note that these GAMMs are using the size subsetted data, as opposed to the full size ranges
+
 Predicting feeding rate (site-level medians, island as random effect)
-```{r, echo=F, message=F, warning=F}
+```{r, message=F, warning=F}
 m_data <- sum_ssp_pca
@@
 
-# gamm4 <- gamm(fr_median ~ species_code + s(length_mean, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3) + s(pc2_all, k=3), 
-#            family = gaussian(link=identity),
-#            random = list(island=~1),
-#            data = m_data
-#            )
-# g4 <- c("g4", "Species + Size~Species + PC1~Species + PC2", AICc(gamm4$lme), summary(gamm4$gam)$r.sq)
+gamm4 <- gamm(fr_median ~ species_code + s(length_mean, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3) + s(pc2_all, k=3),
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           data = m_data
+           )
+g4 <- c("g4", "Species + Size~Species + PC1~Species + PC2", AICc(gamm4$lme), summary(gamm4$gam)$r.sq)
 
-# gamm5 <- gamm(fr_median ~ species + s(length_mean, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3), 
-#            family = gaussian(link=identity),
-#            random = list(island=~1),
-#            data = m_data
-#            )
-# g5 <- c("g5", "Species + Size~Species + PC1~Species", AICc(gamm5$lme), summary(gamm5$gam)$r.sq)
+gamm5 <- gamm(fr_median ~ species + s(length_mean, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3),
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           data = m_data
+           )
+g5 <- c("g5", "Species + Size~Species + PC1~Species", AICc(gamm5$lme), summary(gamm5$gam)$r.sq)
 
@@
 
-gtab_fr_site <- as.data.frame(rbind(g1,g2,g4,g5,g6,g7,g8,g9,g10,g11,g12)) %>% setNames(c("g","Model", "AICc", "Adj. R2")) %>% arrange(AICc)
+gtab_fr_site <- as.data.frame(rbind(g1,g2,g3,g4,g5,g6,g7,g8,g9,g10,g11,g12)) %>% setNames(c("g","Model", "AICc", "Adj. R2")) %>% arrange(AICc)
+kable(gtab_fr_site)
 
 # summary + plots from best models
-summary(gamm12$gam)
-plot(gamm12$gam,pages=1)
+summary(gamm10$gam)
+plot(gamm10$gam,pages=1)
 ```
 
-Predicting Grazing intensity (site-level medians, island as random effect)
+Predicting grazing intensity (site-level medians, island as random effect)
 ```{r, echo=F, message=F, warning=F}
@@
 
-gamm4 <- gamm(for_bites_median ~ species_code + s(length_mean, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3) + s(pc2_all, k=3), 
-           family = gaussian(link=identity),
-           random = list(island=~1),
-           data = m_data
-           )
-g4 <- c("g4", "Species + Size~Species + PC1~Species + PC2", AICc(gamm4$lme), summary(gamm4$gam)$r.sq)
+# gamm4 <- gamm(for_bites_median ~ species_code + s(length_mean, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3) + s(pc2_all, k=3),
+#            family = gaussian(link=identity),
+#            random = list(island=~1),
+#            data = m_data
+#            )
+# g4 <- c("g4", "Species + Size~Species + PC1~Species + PC2", AICc(gamm4$lme), summary(gamm4$gam)$r.sq)
 
@@
 
-gtab_fi_site <- as.data.frame(rbind(g1,g2,g3,g4,g5,g6,g7,g8,g9,g10,g11,g12)) %>% setNames(c("g","Model", "AICc", "Adj. R2")) %>% arrange(AICc)
+gtab_fi_site <- as.data.frame(rbind(g1,g2,g3,g5,g6,g7,g8,g9,g10,g11,g12)) %>% setNames(c("g","Model", "AICc", "Adj. R2")) %>% arrange(AICc)
+kable(gtab_fr_site)
 
@@
 
-# Supplement
+## Supplement/optional
 
@@
 
-### loess plots for additional drivers?
+### Loess plots for additional drivers?

### 2019-12-06 r/grazing intensity.R

@@ -1,0 +1,219 @@
+for_bites <- sum_id %>%
+  select(id, for_bites, site) %>%
+  mutate(for_bites_binned = cut(for_bites, breaks = c(0,10,30,65), labels = c("low","med","high")),
+         for_bites_rounded = round(for_bites, 0)
+         )
+
+ggplot(for_bites, aes(x = for_bites_rounded, fill = factor(site))) +
+  geom_histogram() +
+  geom_density(alpha=0.6) +
+  theme_bw() +
+  labs(x = "Grazing intensity (bites/foray)")
+
+ggsave(here("figs/intensity_hist.png"), width = 3, height = 3)
+
+ggplot(sum_ssp_sub, aes(x = for_bites_median, y = fr_median)) +
+  geom_point() +
+  labs(x = "Grazing intensity (bites/foray)", y = "Feeding rate (bites/min)")
+ggsave(here("figs/intensity_vs_feedingrate.png"), width = 3, height = 3)
+
+ggplot(sum_id, aes(x = for_bites, y = fr)) +
+  geom_point(aes(colour = factor(site)))
+
+## Loess plots of various drivers for grazing intensity
+
+gi_q1 <- ggplot(filter(sum_ssp_sub, species_code == "qup"), aes(x = scar_bm/1000, y = for_bites_median)) +
+  geom_smooth(method="loess", span=1) +
+  geom_point(shape=18) +
+  ylim(-5, 25) +
+  labs(y = "",
+       title = "") +
+  theme_bw() + 
+  theme(plot.title = element_text(hjust = 0.5),
+        plot.margin = margin(0, 1, 0, 0, unit = "pt"),
+        axis.text.x = element_blank(),
+        axis.title.x = element_blank()
+  )
+
+gi_q2 <- ggplot(filter(sum_ssp_sub, species_code == "qup"), aes(x = consp_den, y = for_bites_median)) +
+  geom_smooth(method="loess", span=1) +
+  geom_point(shape=18) +
+  ylim(-5, 25) +
+  labs(y = "", 
+       title = "") +
+  theme_bw() + 
+  theme(plot.title = element_text(hjust = 0.5),
+        axis.ticks.y = element_blank(),
+        axis.text.y = element_blank(),
+        plot.margin = margin(0, 1, 0, 1, unit = "pt"),
+        axis.text.x = element_blank(),
+        axis.title.x = element_blank()
+  )
+
[…]

### 2020-02-19 r/Paper_analyses_means.Rmd
feb updates

---
title: 'Variance in scarid grazing behavior across Caribbean reef sites'
author: "*Compiled on `r date()` by Molly Wilson*"
output: 
  html_document:
    toc: true
    code_folding: hide
    number_sections: false
    theme: cerulean
    highlight: haddock

---

*Formatted as: Coral Reefs note*

**Mean version**

```{r, message=F, warning=F}
library(here)
library(tidyverse)
library(R.utils)
library(ggplot2)
library(gridExtra)
library(grid)
library(ggfortify)
library(mgcv)
library(MuMIn)
library(knitr)
library(factoextra)
library(ggthemes)

sum_site <- read.csv(here("data", "site.csv"), strip.white = T, stringsAsFactors = F) %>% select(-X)
sum_id <- read.csv(here("data","sum_id.csv"), stringsAsFactors=F) %>%
  select(-X) %>% 
  filter(phase == "i" & 
           species_code %in% c("stop","qup") &
           ### size subset 15 - 30cm
           length_cm <= 30 & length_cm >= 15 &
           ### excluding sites with less than 7 observations within size subset range
           !(species_code == "stop" & site == "Pallaster West") & 
           !(species_code == "qup" & site %in% c("Pallaster East","Rendezvous","Turtle Bay"))) %>%
  mutate(island = factor(island, levels = c("Bonaire","Antigua","Barbuda"))) %>%
  arrange(island) %>%
  filter(fr <= 3000)

sum_ssp_sub <- sum_id %>% 
  group_by(species, species_code, site, island) %>%
  summarize(
    n=n(),
    length_mean=mean(length_cm),
    length_median=median(length_cm),
    length_se=sd(length_cm)/sqrt(n),
    g_frac_mean=mean(g_frac),
    g_frac_median=median(g_frac),
    g_frac_se=sd(g_frac)/sqrt(n),
    br_mean=mean(br,na.rm=TRUE),
    br_median=median(br,na.rm=TRUE),
    br_se=sd(br,na.rm=TRUE)/sqrt(n),
    fr_mean=mean(fr),
    fr_median=median(fr),
    fr_se=sd(fr)/sqrt(n),
    for_dur_mean=mean(for_dur,na.rm=TRUE),
    for_dur_median=median(for_dur,na.rm=TRUE),
    for_dur_se=sd(for_dur,na.rm=TRUE)/sqrt(n),
    for_bites_mean=mean(for_bites,na.rm=TRUE),
    for_bites_median=median(for_bites,na.rm=TRUE),
    for_bites_se=sd(for_bites,na.rm=TRUE)/sqrt(n)
    ) %>%
  left_join(sum_site, by = c("island","site"))

```

## Preliminary investigations

### Sample sizes by site and species
Currently omitting sites with fewer than 7 observations that fall within 15-30cm size subset
```{r, message=F}
sample <- sum_id %>%
  group_by(island, site, species) %>% 
  summarize(n=n()) %>%
  arrange(island) %>%
  rename_all(capitalize)
kable(sample)
write.csv(sample, file = here("figs/sample.csv"))

sample_vir_sum <- sum_id %>% 
  filter(species_code == "stop") %>% 
  group_by(site, species, island) %>% 
  summarize(n=n()) %>%
  group_by(species) %>% 
  summarize(n_sites = n(),
            n_total = sum(n),
            n_mean = mean(n),
            se = sd(n)/sqrt(n_sites))

sample_vet_sum <- sum_id %>% 
  filter(species_code == "qup") %>% 
  group_by(site, species, island) %>% 
  summarize(n=n()) %>%
  group_by(species) %>% 
  summarize(n_sites = n(),
            n_total = sum(n),
            n_mean = mean(n),
            se = sd(n)/sqrt(n_sites))

```

### Variable definitions
```{r, echo=F, message=F}
Variable <- c('Time spent grazing','Bite rate','Feeding rate','Foray size')
Definition <- c('Proportion of time spent grazing during follow','Bite rate while actively Grazing','Bite rate over total follow time','Number of consecutive bites taken during single grazing foray')
Unit <- c('fraction','bites/second','bites/minute','bites/foray')
var.table <- data.frame(Variable,Definition,Unit)
kable(var.table)
```

### Testing parametric assumptions
Testing for heteroscedasticity first
```{r, message=F}
hist(sum_id$fr) # positive skew
shapiro.test(sum_id$fr) # non-normal

hist(sum_id$br) # positive skew
shapiro.test(sum_id$br) # non-normal

hist(sum_id$g_frac) # positive skew
shapiro.test(sum_id$g_frac) # non-normal

hist(sum_id$for_bites) # positive skew
shapiro.test(sum_id$for_bites) # non-normal

# all positively skewed, non-normally distributed -> should use median, should use Kruskall-Wallis
```
All behavioral variables are positively skewed, non-normally distributed -> should use median values (?) and Kruskall-Wallis instead of ANOVA for site-level differences

## Inter-site differences

Panel graph of behavioral variables across sites:

```{r, message = F, warning = F}
# published Feeding rate (bites/min) for reference lines if needed
# ref_qup15 <- (3329-33*15)/60
# ref_qup30 <- (3329-33*30)/60
# ref_stop15 <- (1089-17*15-56)/60
# ref_stop30 <- (1089-17*30-56)/60

fr_qup <- 
  ggplot(filter(sum_ssp_sub, species_code == "qup"), 
         aes(x=factor(site, levels = sum_site$site), y=fr_mean)) +
  geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
  geom_errorbar(aes(ymin=fr_mean-fr_se, ymax=fr_mean+fr_se), width=.2,
                 position=position_dodge(.9)) +
  scale_fill_brewer(palette = "Blues") +
  #geom_abline(slope=0, intercept=ref_qup15,  col = "red",lty=2) +
  #geom_abline(slope=0, intercept=ref_qup25,  col = "red",lty=2) +
  #ylim(0, 50) +
  labs(y = "Feeding rate \n (bites/min)", 
       title = "S. vetula") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5, face = "italic"),
        plot.margin = margin(5, 0, 0, 5, unit = "pt"),
        strip.text = element_text(face = "italic"), 
        axis.text.x=element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size=8),
        legend.position = "none")

fr_stop <- 
  ggplot(filter(sum_ssp_sub, species_code == "stop"), 
         aes(x=factor(site, levels = sum_site$site), y=fr_mean)) +
  geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
  geom_errorbar(aes(ymin=fr_mean-fr_se, ymax=fr_mean+fr_se), width=.2,
                 position=position_dodge(.9)) +
  scale_fill_brewer(palette = "Blues") +
  #geom_abline(slope=0, intercept=ref_stop15,  col = "red",lty=2) +
  #geom_abline(slope=0, intercept=ref_stop25,  col = "red",lty=2) +
  labs(y = "",
       # y = "S. vetula \nFeeding rate (bites/hr)", 
       title = "S. viride") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5, face = "italic"),
        plot.margin = margin(5, 4, 0, 0, unit = "pt"),
        strip.text = element_text(face = "italic"), 
        axis.text.x=element_blank(),
        axis.title.x = element_blank(),
        legend.position = "none")

br_qup <- 
  ggplot(filter(sum_ssp_sub, species_code == "qup"), 
         aes(x=factor(site, levels = sum_site$site), y=br_mean)) +
  geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
  geom_errorbar(aes(ymin=br_mean-br_se, ymax=br_mean+br_se), width=.2,
                 position=position_dodge(.9)) +
  scale_fill_brewer(palette = "Blues") +
  labs(y = "Bite rate \n(bites/sec)", 
       x = "", 
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        plot.margin = margin(0, 0, 0, 5, unit = "pt"),
        strip.text = element_text(face = "italic"), 
        axis.text.x=element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size=8),
        legend.position = "none")

br_stop <- 
  ggplot(filter(sum_ssp_sub, species_code == "stop"), 
         aes(x=factor(site, levels = sum_site$site), y=br_mean)) +
  geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
  geom_errorbar(aes(ymin=br_mean-br_se, ymax=br_mean+br_se), width=.2,
                 position=position_dodge(.9)) +
  scale_fill_brewer(palette = "Blues") +
  labs(y = "",
       # y = "S. vetula \nFeeding rate (bites/hr)", 
       x = "", 
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        plot.margin = margin(0, 4, 0, 0, unit = "pt"),
        strip.text = element_text(face = "italic"), 
        axis.text.x=element_blank(),
        axis.title.x = element_blank(),
        legend.position = "none")

gfrac_qup <- 
  ggplot(filter(sum_ssp_sub, species_code == "qup"), 
         aes(x=factor(site, levels = sum_site$site), y=g_frac_mean)) +
  geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
  geom_errorbar(aes(ymin=g_frac_mean-g_frac_se, ymax=g_frac_mean+g_frac_se), width=.2,
                 position=position_dodge(.9)) +
  scale_fill_brewer(palette = "Blues") +
  labs(y = "Time grazing \n(fraction)", 
       x = "", 
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        plot.margin = margin(0, 0, 0, 5, unit = "pt"),
        strip.text = element_text(face = "italic"), 
        axis.text.x=element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size=8),
        legend.position = "none")

gfrac_stop <- 
  ggplot(filter(sum_ssp_sub, species_code == "stop"), 
         aes(x=factor(site, levels = sum_site$site), y=g_frac_mean)) +
  geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
  geom_errorbar(aes(ymin=g_frac_mean-g_frac_se, ymax=g_frac_mean+g_frac_se), width=.2,
                 position=position_dodge(.9)) +
  scale_fill_brewer(palette = "Blues") +
  labs(y = "",
       x = "", 
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        plot.margin = margin(0, 4, 0, 0, unit = "pt"),
        strip.text = element_text(face = "italic"), 
        axis.text.x=element_blank(),
        axis.title.x = element_blank(),
        legend.position = "none")

fi_qup <- 
  ggplot(filter(sum_ssp_sub, species_code == "qup"), 
         aes(x=factor(site, levels = sum_site$site), y=for_bites_mean)) +
  geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
  geom_errorbar(aes(ymin=for_bites_mean-for_bites_se, ymax=for_bites_mean+for_bites_se), width=.2,
                 position=position_dodge(.9)) +
  scale_fill_brewer(palette = "Blues") +
  labs(y = "Grazing intensity\n(bites/foray)", 
       x = "", 
       title = "") +
  theme_bw() + 
  theme(plot.margin = margin(0, 0, 0, 5, unit = "pt"),
        axis.title.y = element_text(size=8),
        axis.text.x = element_text(angle = 45, hjust = 1, size=7),
        legend.position = "none")

fi_stop <- 
  ggplot(filter(sum_ssp_sub, species_code == "stop"), 
         aes(x=factor(site, levels = sum_site$site), y=for_bites_mean)) +
  geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
  geom_errorbar(aes(ymin=for_bites_mean-for_bites_se, ymax=for_bites_mean+for_bites_se), width=.2,
                 position=position_dodge(.9)) +
  scale_fill_brewer(palette = "Blues") +
  labs(y = "",
       x = "", 
       title = "") +
  theme_bw() + 
  theme(plot.margin = margin(0, 4, 0, 0, unit = "pt"),
        axis.text.x = element_text(angle = 45, hjust = 1, size=7),
        legend.position = "none")

legend_plot <- 
  ggplot(filter(sum_ssp_sub, species_code == "stop"), 
         aes(x=factor(site, levels = sum_site$site), y=g_frac_mean)) +
  geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
  scale_fill_brewer(palette = "Blues") +
  labs(y = "",
       x = "", 
       title = "") +
  theme_bw() +
  theme(legend.position = "right",
        legend.title = element_blank())

get_legend<-function(myggplot){
  tmp <- ggplot_gtable(ggplot_build(myggplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)
  }
legend <- get_legend(legend_plot)

g <- arrangeGrob(
  arrangeGrob(fr_qup, fr_stop, gfrac_qup, gfrac_stop, br_qup, br_stop, fi_qup, fi_stop, ncol=2, heights = c(1,1,1,1.75)), 
              arrangeGrob(legend), 
              ncol=2, widths = c(5,1))
plot(g)

ggsave(here("figs/behavior_panel_means.png"), g, width = 5, height = 5)
```

Kruskal-Wallis tests to look at significance of inter-site variance for different behavioral variables:

Feeding rate:
```{r}
vir <- sum_id %>% filter(species_code == "stop")
vet <- sum_id %>% filter(species_code == "qup")

kruskal.test(fr ~ site, data = vet) # p<0.001
kruskal.test(fr ~ site, data = vir) # p<0.001
```

Bite rate:
```{r}
kruskal.test(br ~ site, data = vet) # p>0.5 (p=0.63)
kruskal.test(br ~ site, data = vir) # p>0.5 (p=0.55)
```

Fraction of time grazing:
```{r}
kruskal.test(g_frac ~ site, data = vet) # p<0.001
kruskal.test(g_frac ~ site, data = vir) # p<0.001
```

Grazing intensity:
```{r}
kruskal.test(for_bites ~ site, data = vet) # p<0.001
kruskal.test(for_bites ~ site, data = vir) # p<0.001
```

All significant (p < 0.001) except for fraction of time spent grazing.

## Loess plots of various drivers

```{r, message=F, warning=F}
fr_q1 <- ggplot(filter(sum_ssp_sub, species_code == "qup"), aes(x = scar_bm/1000, y = fr_mean)) +
  geom_smooth(method="loess", span=1) +
  geom_point(shape=18) +
  ylim(-10, 60) +
  labs(y = "",
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        plot.margin = margin(0, 1, 0, 0, unit = "pt"),
        axis.text.x = element_blank(),
        axis.title.x = element_blank()
        )

fr_q2 <- ggplot(filter(sum_ssp_sub, species_code == "qup"), aes(x = carn30_bm/1000, y = fr_mean)) +
  geom_smooth(method="loess", span=1) +
  geom_point(shape=18) +
  ylim(-10, 60) +
  labs(y = "", 
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        plot.margin = margin(0, 1, 0, 1, unit = "pt"),
        axis.text.x = element_blank(),
        axis.title.x = element_blank()
        )

fr_q3 <- ggplot(filter(sum_ssp_sub, species_code == "qup"), aes(x = rugosity, y = fr_mean)) +
  geom_smooth(method="loess", span=1) +
  geom_point(shape=18) +
  ylim(-10, 60) +
  labs(y = "", 
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        plot.margin = margin(0, 1, 0, 1, unit = "pt"),
        axis.text.x = element_blank(),
        axis.title.x = element_blank()
        )

fr_q4 <- ggplot(filter(sum_ssp_sub, species_code == "qup"), aes(x = ta_canopy, y = fr_mean)) +
  geom_smooth(method="loess", span=1) +
  geom_point(shape=18) +
  ylim(-10, 60) +
  labs(y = "", 
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        plot.margin = margin(0, 1, 0, 1, unit = "pt"),
        axis.text.x = element_blank(),
        axis.title.x = element_blank()
        )

fr_q5 <- ggplot(filter(sum_ssp_sub, species_code == "qup"), aes(x = ta_cover, y = fr_mean)) +
  geom_smooth(method="loess", span=1) +
  geom_point(shape=18) +
  ylim(-10, 60) +
  labs(y = "",
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        plot.margin = margin(0, 1, 0, 1, unit = "pt"),
        axis.text.x = element_blank(),
        axis.title.x = element_blank()
        )

fr_q7 <- ggplot(filter(sum_ssp_sub, species_code == "qup"), aes(x = ma_cover, y = fr_mean)) +
  geom_smooth(method="loess", span=1) +
  geom_point(shape=18) +
  ylim(-10, 60) +
  labs(y = "", 
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        plot.margin = margin(0, 1, 0, 1, unit = "pt"),
        axis.text.x = element_blank(),
        axis.title.x = element_blank()
        )

fr_s1 <- ggplot(filter(sum_ssp_sub, species_code == "stop"), aes(x = scar_bm/1000, y = fr_mean)) +
  geom_smooth(method="loess", span=1) +
  geom_point(shape=18) +
  ylim(-5, 20) +
  labs(y = "",
       x = expression(atop(NA, atop(textstyle('Scarid biomass'),
                               textstyle('(kg/100m'^2*')')))),
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        plot.margin = margin(0, 1, 0, 0, unit = "pt"),
        axis.title.x = element_text(size = 7)
        )

fr_s2 <- ggplot(filter(sum_ssp_sub, species_code == "stop"), aes(x = carn30_bm/1000, y = fr_mean)) +
  geom_smooth(method="loess", span=1) +
  geom_point(shape=18) +
  ylim(-5, 20) +
  labs(y = "", 
       x = expression(atop(NA, atop(textstyle('Predator biomass'),
                               textstyle('(kg/100m'^2*')')))),
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        axis.title.x = element_text(size = 7),
        plot.margin = margin(0, 1, 0, 1, unit = "pt")
        )

fr_s3 <- ggplot(filter(sum_ssp_sub, species_code == "stop"), aes(x = rugosity, y = fr_mean)) +
  geom_smooth(method="loess", span=1) +
  geom_point(shape=18) +
  ylim(-5, 20) +
  labs(y = "", 
       x = expression(atop(NA, atop(textstyle('Rugosity'),
                               textstyle('  ')))), 
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        axis.title.x = element_text(size = 7),
        plot.margin = margin(0, 1, 0, 1, unit = "pt")
        )

fr_s4 <- ggplot(filter(sum_ssp_sub, species_code == "stop"), aes(x = ta_canopy, y = fr_mean)) +
  geom_smooth(method="loess", span=1) +
  geom_point(shape=18) +
  ylim(-5, 20) +
  labs(y = "", 
       x = expression(atop(NA, atop(textstyle('Turf canopy'),
                               textstyle('height (mm)')))), 
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        axis.title.x = element_text(size = 7),
        plot.margin = margin(0, 1, 0, 1, unit = "pt")
        )

fr_s5 <- ggplot(filter(sum_ssp_sub, species_code == "stop"), aes(x = ta_cover, y = fr_mean)) +
  geom_smooth(method="loess", span=1) +
  geom_point(shape=18) +
  ylim(-5, 20) +
  labs(y = "", 
       x = expression(atop(NA, atop(textstyle('Turf cover'),
                               textstyle('(%)')))), 
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        axis.title.x = element_text(size = 7),
        plot.margin = margin(0, 1, 0, 1, unit = "pt")
        )

fr_s7 <- ggplot(filter(sum_ssp_sub, species_code == "stop"), aes(x = ma_cover, y = fr_mean)) +
  geom_smooth(method="loess", span=1) +
  geom_point(shape=18) +
  ylim(-5, 20) +
  labs(y = "", 
       x = expression(atop(NA, atop(textstyle('Macroalgal cover'),
                               textstyle('(%)')))), 
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.title.x = element_text(size = 7),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        plot.margin = margin(0, 1, 0, 1, unit = "pt")
        )

label_vet = textGrob("S. vetula", gp=gpar(fontface="italic", fontsize = 10), rot=270)
label_vir = textGrob("S. viride", gp=gpar(fontface="italic", fontsize = 10), rot=270)

g <- grid.arrange(left = textGrob("Feeding rate (bites/min.)", rot = 90, gp = gpar(fontsize = 10)),
  arrangeGrob(fr_q1, fr_q2, fr_q3, fr_q4, fr_q5, fr_q7, ncol=6, right = label_vet),
  arrangeGrob(fr_s1, fr_s2, fr_s3, fr_s4, fr_s5, fr_s7, ncol=6, right = label_vir), 
  nrow=2, heights = c(1,1.2))

ggsave(here("figs/biplot_panel_means.png"), g, width = 7, height = 4)
```
### Carnivore Loess for supplement
```{r}
ggplot(filter(sum_ssp_sub, species_code == "stop"), aes(x = carn30_bm, y = fr_mean)) +
  geom_smooth(method="loess", span=1) +
  geom_point(shape=18) +
  labs(y = "Feeding rate (bites/min.)",
       x = expression(atop(NA, atop(textstyle('Predator abundance')))), 
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5)
        )

ggplot(filter(sum_ssp_sub, species_code == "qup"), aes(x = carn30_bm, y = fr_mean)) +
  geom_smooth(method="loess", span=1) +
  geom_point(shape=18) +
  labs(y = "Feeding rate (bites/min.)",
       x = expression(atop(NA, atop(textstyle('Predator abundance')))), 
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5)
        )
```

## PCA

```{r, message=F, warning=F}
pca_data <- na.omit(sum_site %>%
  select(rugosity, ma_cover, ma_canopy, ta_cover, ta_canopy, lc_cover, scar_bm, scar_den, carn30_bm))
names(pca_data)[1:9] =
  c("rugosity", "macroalgal cover", "macroalgal canopy", "turf cover", "turf canopy", "coral cover", "scarid biomass", "scarid density", "predator biomass")

pca_pred <- prcomp(pca_data, center = T, scale = T)
print(pca_pred)
pca_var <- get_pca_var(pca_pred)
print(pca_var$contrib)

set.seed(5)
fviz_pca_biplot(pca_pred,
                col.ind = sum_site$island,
                mean.point = F,
                label = "var",
                repel = T,
                pointsize = 3,
                alpha.var= 0.3,
                col.var = "gray22",
                labelsize = 4,
                axes.linetype = "dotted"
                ) +
  scale_shape_manual(values = c(15,16,17)) +
  scale_color_manual(values = c("turquoise4", "midnightblue", "steelblue1")) +
  labs(x = "PC1 (55.6%)", y = "PC2 (24.5%)", title = "") +
  theme_bw() +
  theme(legend.title = element_blank())

ggsave(here("figs/pca_pred.png"), width = 7, height = 5)
```

```{r, message=F}
sum_site_pca <- sum_site %>% 
  bind_cols(as.data.frame(pca_pred$x) %>% 
           select(PC1,PC2)) %>%
  rename(pc1_all=PC1,
         pc2_all=PC2)

sum_id_pca <- sum_id %>% 
  left_join(select(sum_site_pca,site,pc1_all,pc2_all), by = "site")
sum_ssp_pca <- sum_ssp_sub %>%
  left_join(select(sum_site_pca,site,pc1_all,pc2_all), by = "site")
```

## GAMM results (see 5.8.19)

Note that these GAMMs are using the size subsetted data, as opposed to the full size ranges

Predicting feeding rate (site-level means, island as random effect)
```{r, message=F, warning=F}
m_data <- sum_ssp_pca

# gamm1 <- gamm(fr_mean ~ species_code + s(length_mean, k=3) + s(pc1_all, k=3) + s(pc2_all, k=3), 
#            family = gaussian(link=identity),
#            random = list(island=~1),
#            data = m_data
#            )
# g1 <- c("g1", "Species + Size + PC1 + PC2", AICc(gamm1$lme), summary(gamm1$gam)$r.sq)

gamm2 <- gamm(fr_mean ~ species_code + s(length_mean, k=3) + s(pc1_all, k=3), 
           family = gaussian(link=identity),
           random = list(island=~1),
           data = m_data
           )
g2 <- c("g2", "Species + Size + PC1", AICc(gamm2$lme), summary(gamm2$gam)$r.sq)

# gamm3 <- gamm(fr_mean ~ species_code + s(length_mean, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3) + s(pc2_all, by = factor(species_code), k=3),
#            family = gaussian(link=identity),
#            random = list(island=~1),
#            data = m_data
#            )
# g3 <- c("g3", "Species + Size~Species + PC1~Species + PC2~Species", AICc(gamm3$lme), summary(gamm3$gam)$r.sq)

# gamm4 <- gamm(fr_mean ~ species_code + s(length_mean, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3) + s(pc2_all, k=3),
#            family = gaussian(link=identity),
#            random = list(island=~1),
#            data = m_data
#            )
# g4 <- c("g4", "Species + Size~Species + PC1~Species + PC2", AICc(gamm4$lme), summary(gamm4$gam)$r.sq)

gamm5 <- gamm(fr_mean ~ species + s(length_mean, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3),
           family = gaussian(link=identity),
           random = list(island=~1),
           data = m_data
           )
g5 <- c("g5", "Species + Size~Species + PC1~Species", AICc(gamm5$lme), summary(gamm5$gam)$r.sq)

gamm6 <- gamm(fr_mean ~ s(length_mean, k=3) + s(pc1_all, by = factor(species_code), k=3), 
           family = gaussian(link=identity),
           random = list(island=~1),
           data = m_data
           )
g6 <- c("g6", "Size + PC1~Species", AICc(gamm6$lme), summary(gamm6$gam)$r.sq)

gamm7 <- gamm(fr_mean ~ s(length_mean, by = factor(species_code), k=3) + s(pc1_all, k=3), 
           family = gaussian(link=identity),
           random = list(island=~1),
           data = m_data
           )
g7 <- c("g7", "Size~Species + PC1", AICc(gamm7$lme), summary(gamm7$gam)$r.sq)

gamm8 <- gamm(fr_mean ~ species + s(length_mean, by = factor(species_code), k=3) + s(pc1_all, k=3), 
           family = gaussian(link=identity),
           random = list(island=~1),
           data = m_data
           )
g8 <- c("g8", "Species + Size~Species + PC1", AICc(gamm8$lme), summary(gamm8$gam)$r.sq)

gamm9 <- gamm(fr_mean ~ species + s(length_mean, k=3) + s(pc1_all, by = factor(species_code), k=3), 
           family = gaussian(link=identity),
           random = list(island=~1),
           data = m_data
           )
g9 <- c("g9", "Species + Size + PC1~Species", AICc(gamm9$lme), summary(gamm9$gam)$r.sq)

gamm10 <- gamm(fr_mean ~ species + s(pc1_all, by = factor(species_code), k=3), 
           family = gaussian(link=identity),
           random = list(island=~1),
           data = m_data
           )
g10 <- c("g10", "Species + PC1~Species", AICc(gamm10$lme), summary(gamm10$gam)$r.sq)

gamm11 <- gamm(fr_mean ~ s(length_mean, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3), 
           family = gaussian(link=identity),
           random = list(island=~1),
           data = m_data
           )
g11 <- c("g11", "Size~Species + PC1~Species", AICc(gamm11$lme), summary(gamm11$gam)$r.sq)

gamm12 <- gamm(fr_mean ~ species_code + s(pc1_all, k=3), 
           family = gaussian(link=identity),
           random = list(island=~1),
           data = m_data
           )
g12 <- c("g12", "Species + PC1", AICc(gamm12$lme), summary(gamm12$gam)$r.sq)

gtab_fr_site <- as.data.frame(rbind(g2,g5,g6,g7,g8,g9,g10,g11,g12)) %>% setNames(c("g","Model", "AICc", "Adj. R2")) %>% arrange(AICc)
kable(gtab_fr_site)

# summary + plots from best models
summary(gamm10$gam)
plot(gamm10$gam,pages=1)
```

Predicting grazing intensity (site-level means, island as random effect)
```{r, echo=F, message=F, warning=F}
m_data <- sum_ssp_pca

gamm1 <- gamm(for_bites_mean ~ species_code + s(length_mean, k=3) + s(pc1_all, k=3) + s(pc2_all, k=3), 
           family = gaussian(link=identity),
           random = list(island=~1),
           data = m_data
           )
g1 <- c("g1", "Species + Size + PC1 + PC2", AICc(gamm1$lme), summary(gamm1$gam)$r.sq)

gamm2 <- gamm(for_bites_mean ~ species_code + s(length_mean, k=3) + s(pc1_all, k=3), 
           family = gaussian(link=identity),
           random = list(island=~1),
           data = m_data
           )
g2 <- c("g2", "Species + Size + PC1", AICc(gamm2$lme), summary(gamm2$gam)$r.sq)

gamm3 <- gamm(for_bites_mean ~ species_code + s(length_mean, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3) + s(pc2_all, by = factor(species_code), k=3),
           family = gaussian(link=identity),
           random = list(island=~1),
           data = m_data
           )
g3 <- c("g3", "Species + Size~Species + PC1~Species + PC2~Species", AICc(gamm3$lme), summary(gamm3$gam)$r.sq)

# gamm4 <- gamm(for_bites_mean ~ species_code + s(length_mean, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3) + s(pc2_all, k=3),
#            family = gaussian(link=identity),
#            random = list(island=~1),
#            data = m_data
#            )
# g4 <- c("g4", "Species + Size~Species + PC1~Species + PC2", AICc(gamm4$lme), summary(gamm4$gam)$r.sq)

gamm5 <- gamm(for_bites_mean ~ species + s(length_mean, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3), 
           family = gaussian(link=identity),
           random = list(island=~1),
           data = m_data
           )
g5 <- c("g5", "Species + Size~Species + PC1~Species", AICc(gamm5$lme), summary(gamm5$gam)$r.sq)

gamm6 <- gamm(for_bites_mean ~ s(length_mean, k=3) + s(pc1_all, by = factor(species_code), k=3), 
           family = gaussian(link=identity),
           random = list(island=~1),
           data = m_data
           )
g6 <- c("g6", "Size + PC1~Species", AICc(gamm6$lme), summary(gamm6$gam)$r.sq)

gamm7 <- gamm(for_bites_mean ~ s(length_mean, by = factor(species_code), k=3) + s(pc1_all, k=3), 
           family = gaussian(link=identity),
           random = list(island=~1),
           data = m_data
           )
g7 <- c("g7", "Size~Species + PC1", AICc(gamm7$lme), summary(gamm7$gam)$r.sq)

gamm8 <- gamm(for_bites_mean ~ species + s(length_mean, by = factor(species_code), k=3) + s(pc1_all, k=3), 
           family = gaussian(link=identity),
           random = list(island=~1),
           data = m_data
           )
g8 <- c("g8", "Species + Size~Species + PC1", AICc(gamm8$lme), summary(gamm8$gam)$r.sq)

gamm9 <- gamm(for_bites_mean ~ species + s(length_mean, k=3) + s(pc1_all, by = factor(species_code), k=3), 
           family = gaussian(link=identity),
           random = list(island=~1),
           data = m_data
           )
g9 <- c("g9", "Species + Size + PC1~Species", AICc(gamm9$lme), summary(gamm9$gam)$r.sq)

gamm10 <- gamm(for_bites_mean ~ species + s(pc1_all, by = factor(species_code), k=3), 
           family = gaussian(link=identity),
           random = list(island=~1),
           data = m_data
           )
g10 <- c("g10", "Species + PC1~Species", AICc(gamm10$lme), summary(gamm10$gam)$r.sq)

gamm11 <- gamm(for_bites_mean ~ s(length_mean, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3), 
           family = gaussian(link=identity),
           random = list(island=~1),
           data = m_data
           )
g11 <- c("g11", "Size~Species + PC1~Species", AICc(gamm11$lme), summary(gamm11$gam)$r.sq)

gamm12 <- gamm(for_bites_mean ~ species_code + s(pc1_all, k=3), 
           family = gaussian(link=identity),
           random = list(island=~1),
           data = m_data
           )
g12 <- c("g12", "Species + PC1", AICc(gamm12$lme), summary(gamm12$gam)$r.sq)

gtab_fi_site <- as.data.frame(rbind(g1,g2,g3,g5,g6,g7,g8,g9,g10,g11,g12)) %>% setNames(c("g","Model", "AICc", "Adj. R2")) %>% arrange(AICc)
kable(gtab_fr_site)

# plots from best models
plot(gamm12$gam,pages=1)
plot(gamm10$gam,pages=1)

summary(gamm12$gam)
```

## Supplement/optional

### Predator gape size

From Mumby et al.:
Grouper *E. striatus*, gape width (G): G = -0.5998 + (0.167 × TL)
```{r}
g30_mum <- -0.5998 + (0.167 * 30) # 4
g40_mum <- -0.5998 + (0.167 * 40) # 6
```

From Goatley & Bellwood:
Horizontal gape: G = -0.002 * SL + 6.421
```{r}
g30_goat <- -0.002 * 30 + 6.421
g40_goat <- -0.002 * 40 + 6.421
```

Dorsoventral height (Rakitin & Kramer 1996):
S. viride: 15cm FL > 4.5cm body depth
S. vetula: 15cm FL > 4.5cm body depth

### Loess plots for additional drivers?

### 2020-02-19 r/Workflow.Rmd
feb updates

@@
 output: html_document
+
 ---
@@
   - **Data_prodding.rmd**: exploring relationships among variables, distributions, etc.
-  - **Analysis.rmd**: initial boxplots/bargraphs looking at trends among islands/sites, testing ANOVAs/assumptions, predicted grazing impact, bite content calculations
+  - **Analysis_misc.rmd**: initial boxplots/bargraphs looking at trends among islands/sites, testing ANOVAs/assumptions, predicted grazing impact, bite content calculations
   - **Models.rmd**: PCAs, LMEs, GAMMs (exploratory, no .csv exports)
-  - **Ch2_outline_V3**: selected analyses for current manuscript outline
+  - **Paper_analyses**: selected analyses for current manuscript outline, as well as relevant precursors (checking sample sizes, testing for normality, etc.)

### 2020-12-06 r/Analysis.Rmd
extra figs for defense slides

@@
   select(-X)
+```
+
+```{r}
+# testing size distribution...
+sum_id_sub <- sum_id %>%
+  filter(length_cm >= 15 & length_cm <= 30 & species_code %in% c("qup", "stop") & phase == "i")
+
+ggplot(data = sum_id_sub, aes(x = site, y = length_cm, group = site)) +
+  geom_boxplot() +
+  facet_grid(. ~ species)
+
+ggplot(data = sum_id_sub, aes(x = length_cm, group = site, fill = site)) +
+  geom_density(alpha = 0.1)
+
+lm <- lm(data = sum_ssp_pca, pc1_all ~ length_mean)
+summary(lm) # not significantly correlated with PC1
 ```

### 2020-12-06 r/Paper_analyses.Rmd
extra figs for defense slides

@@
 plot(g)
+ggsave(here("figs/behavior_panel.png"),g)
+```
 
-ggsave(here("figs/behavior_panel.png"),g)
+```{r}
+# indv plots for slides
+
+library(ggpubr)
+
+fr_qup <- 
+  ggplot(filter(sum_ssp_sub, species_code == "qup"), 
+         aes(x=factor(site, levels = sum_site$site), y=fr_median)) +
+  geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
+  geom_errorbar(aes(ymin=fr_median-fr_se, ymax=fr_median+fr_se), width=.2,
+                 position=position_dodge(.9)) +
+  scale_fill_brewer(palette = "Blues", name = "Island") +
+  labs(y = expression(Feeding~rate~(bites~min^-1)), 
+       title = "S. vetula") +
+  theme_bw() + 
+  theme(plot.title = element_text(hjust = 0.5, face = "italic"),
+        plot.margin = margin(5, 0, 0, 5, unit = "pt"),
+        strip.text = element_text(face = "italic"), 
+        axis.text.x = element_text(angle = 90, hjust = 1),
+        axis.title.x = element_blank(),
+        axis.title.y = element_text(size=12),
+        )
+
+fr_stop <- 
+  ggplot(filter(sum_ssp_sub, species_code == "stop"), 
+         aes(x=factor(site, levels = sum_site$site), y=fr_median)) +
+  geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
+  geom_errorbar(aes(ymin=fr_median-fr_se, ymax=fr_median+fr_se), width=.2,
+                 position=position_dodge(.9)) +
+  scale_fill_brewer(palette = "Blues", name = "Island") +
+  labs(y = "",
+       title = "S. viride") +
+  theme_bw() + 
+  theme(plot.title = element_text(hjust = 0.5, face = "italic"),
+        plot.margin = margin(5, 4, 0, 0, unit = "pt"),
+        strip.text = element_text(face = "italic"), 
+        axis.text.x = element_text(angle = 90, hjust = 1),
+        axis.title.x = element_blank(),
+        )
+
+ggarrange(fr_qup, fr_stop, 
+          #abels = c("A", "B"),
+          ncol = 2, nrow = 1,
+          common.legend = TRUE, legend = "right")
+
+ggsave(here("figs", "fr.png"), height = 3, width = 6)
+
+fi_qup <- 
+  ggplot(filter(sum_ssp_sub, species_code == "qup"), 
+         aes(x=factor(site, levels = sum_site$site), y=for_bites_median)) +
+  geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
+  geom_errorbar(aes(ymin=for_bites_median-for_bites_se, ymax=for_bites_median+for_bites_se), width=.2,
+                 position=position_dodge(.9)) +
+  scale_fill_brewer(palette = "Blues", name = "Island") +
+  labs(y = expression(Grazing~intensity~(bites~foray^-1)), 
+       title = "S. vetula") +
+  theme_bw() + 
+  theme(plot.title = element_text(hjust = 0.5, face = "italic"),
+        plot.margin = margin(5, 0, 0, 5, unit = "pt"),
+        strip.text = element_text(face = "italic"), 
+        axis.text.x = element_text(angle = 90, hjust = 1),
+        axis.title.x = element_blank(),
+        axis.title.y = element_text(size=12),
+        )
+
+fi_stop <- 
+  ggplot(filter(sum_ssp_sub, species_code == "stop"), 
+         aes(x=factor(site, levels = sum_site$site), y=for_bites_median)) +
+  geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
+  geom_errorbar(aes(ymin=for_bites_median-for_bites_se, ymax=for_bites_median+for_bites_se), width=.2,
+                 position=position_dodge(.9)) +
+  scale_fill_brewer(palette = "Blues", name = "Island") +
+  labs(y = "",
+       title = "S. viride") +
+  theme_bw() + 
+  theme(plot.title = element_text(hjust = 0.5, face = "italic"),
+        plot.margin = margin(5, 4, 0, 0, unit = "pt"),
+        strip.text = element_text(face = "italic"), 
+        axis.text.x = element_text(angle = 90, hjust = 1),
+        axis.title.x = element_blank(),
+        )
+
+ggarrange(fi_qup, fi_stop, 
+          #abels = c("A", "B"),
+          ncol = 2, nrow = 1,
+          common.legend = TRUE, legend = "right")
+
+ggsave(here("figs", "fi.png"), height = 3, width = 6)
 ```

### 2020-12-06 r/Sites.Rmd
updating stats
extra figs for defense slides

@@
   janitor::clean_names(case="old_janitor")
+```
+
+## Sample sizes per site
+```{r}
+n_fish_tran <- fish %>% 
+  filter(biomass != 0) %>% 
+  group_by(site) %>%
+  summarize(n_tran = max(transect)) %>%
+  ungroup()
+
+mean_fish_tran <- n_fish_tran %>%
+  summarize(mean = mean(n_tran),
+            sd = sd(n_tran),
+            se = sd/sqrt(n()))
+
+n_benthic_tran <- bent_tran %>% 
+  group_by(site) %>%
+  summarize(n_tran = max(transect)) %>%
+  ungroup()
+
+n_rugosity_tran <- ab_rugosity %>% 
+  group_by(site) %>%
+  summarize(n_tran = max(transect)) %>%
+  ungroup()
 ```

### 2021-01-09 r/Paper_analyses.Rmd
file organization

--- 2019-12-06 r/Paper_analyses.Rmd
@@
 
+**Mean version**
+
 ```{r, message=F, warning=F}
@@
 library(factoextra)
+library(ggthemes)
 
@@
     length_se=sd(length_cm)/sqrt(n),
-    g_frac_median=median(g_frac),
+    g_frac_mean=mean(g_frac),
     g_frac_median=median(g_frac),
     g_frac_se=sd(g_frac)/sqrt(n),
-    br_median=median(br,na.rm=TRUE),
+    br_mean=mean(br,na.rm=TRUE),
     br_median=median(br,na.rm=TRUE),
     br_se=sd(br,na.rm=TRUE)/sqrt(n),
-    fr_median=median(fr),
+    fr_mean=mean(fr),
     fr_median=median(fr),
     fr_se=sd(fr)/sqrt(n),
-    for_dur_median=median(for_dur,na.rm=TRUE),
+    for_dur_mean=mean(for_dur,na.rm=TRUE),
     for_dur_median=median(for_dur,na.rm=TRUE),
     for_dur_se=sd(for_dur,na.rm=TRUE)/sqrt(n),
-    for_bites_median=median(for_bites,na.rm=TRUE),
+    for_bites_mean=mean(for_bites,na.rm=TRUE),
     for_bites_median=median(for_bites,na.rm=TRUE),
@@
             n_total = sum(n),
-            n_median = median(n),
+            n_mean = mean(n),
             se = sd(n)/sqrt(n_sites))
@@
 ```
-All behavioral variables are positively skewed, non-normally distributed -> should use median values and Kruskall-Wallis instead of ANOVA for site-level differences
+All behavioral variables are positively skewed, non-normally distributed -> should use median values (?) and Kruskall-Wallis instead of ANOVA for site-level differences
 
@@
 # published Feeding rate (bites/min) for reference lines if needed
-ref_qup15 <- (3329-33*15)/60
-ref_qup30 <- (3329-33*30)/60
-ref_stop15 <- (1089-17*15-56)/60
-ref_stop30 <- (1089-17*30-56)/60
+# ref_qup15 <- (3329-33*15)/60
+# ref_qup30 <- (3329-33*30)/60
+# ref_stop15 <- (1089-17*15-56)/60
+# ref_stop30 <- (1089-17*30-56)/60
+
+colors <- c("powderblue", "turquoise4", "steelblue4")
 
@@
   ggplot(filter(sum_ssp_sub, species_code == "qup"), 
-         aes(x=factor(site, levels = sum_site$site), y=fr_median)) +
+         aes(x=factor(site, levels = sum_site$site), y=fr_mean)) +
   geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
-  geom_errorbar(aes(ymin=fr_median-fr_se, ymax=fr_median+fr_se), width=.2,
+  geom_errorbar(aes(ymin=fr_mean-fr_se, ymax=fr_mean+fr_se), width=.2,
                  position=position_dodge(.9)) +
-  scale_fill_brewer(palette = "Blues") +
+  scale_fill_manual(values = colors) +
   #geom_abline(slope=0, intercept=ref_qup15,  col = "red",lty=2) +
@@
   ggplot(filter(sum_ssp_sub, species_code == "stop"), 
-         aes(x=factor(site, levels = sum_site$site), y=fr_median)) +
+         aes(x=factor(site, levels = sum_site$site), y=fr_mean)) +
   geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
-  geom_errorbar(aes(ymin=fr_median-fr_se, ymax=fr_median+fr_se), width=.2,
+  geom_errorbar(aes(ymin=fr_mean-fr_se, ymax=fr_mean+fr_se), width=.2,
                  position=position_dodge(.9)) +
-  scale_fill_brewer(palette = "Blues") +
+  scale_fill_manual(values = colors) +
   #geom_abline(slope=0, intercept=ref_stop15,  col = "red",lty=2) +
@@
   ggplot(filter(sum_ssp_sub, species_code == "qup"), 
-         aes(x=factor(site, levels = sum_site$site), y=br_median)) +
+         aes(x=factor(site, levels = sum_site$site), y=br_mean)) +
   geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
-  geom_errorbar(aes(ymin=br_median-br_se, ymax=br_median+br_se), width=.2,
+  geom_errorbar(aes(ymin=br_mean-br_se, ymax=br_mean+br_se), width=.2,
                  position=position_dodge(.9)) +
-  scale_fill_brewer(palette = "Blues") +
+  scale_fill_manual(values = colors) +
   labs(y = "Bite rate \n(bites/sec)", 
@@
   ggplot(filter(sum_ssp_sub, species_code == "stop"), 
-         aes(x=factor(site, levels = sum_site$site), y=br_median)) +
+         aes(x=factor(site, levels = sum_site$site), y=br_mean)) +
   geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
-  geom_errorbar(aes(ymin=br_median-br_se, ymax=br_median+br_se), width=.2,
+  geom_errorbar(aes(ymin=br_mean-br_se, ymax=br_mean+br_se), width=.2,
                  position=position_dodge(.9)) +
-  scale_fill_brewer(palette = "Blues") +
+  scale_fill_manual(values = colors) +
   labs(y = "",
@@
   ggplot(filter(sum_ssp_sub, species_code == "qup"), 
-         aes(x=factor(site, levels = sum_site$site), y=g_frac_median)) +
+         aes(x=factor(site, levels = sum_site$site), y=g_frac_mean)) +
   geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
-  geom_errorbar(aes(ymin=g_frac_median-g_frac_se, ymax=g_frac_median+g_frac_se), width=.2,
+  geom_errorbar(aes(ymin=g_frac_mean-g_frac_se, ymax=g_frac_mean+g_frac_se), width=.2,
                  position=position_dodge(.9)) +
-  scale_fill_brewer(palette = "Blues") +
+  scale_fill_manual(values = colors) +
   labs(y = "Time grazing \n(fraction)", 
@@
   ggplot(filter(sum_ssp_sub, species_code == "stop"), 
-         aes(x=factor(site, levels = sum_site$site), y=g_frac_median)) +
+         aes(x=factor(site, levels = sum_site$site), y=g_frac_mean)) +
   geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
-  geom_errorbar(aes(ymin=g_frac_median-g_frac_se, ymax=g_frac_median+g_frac_se), width=.2,
+  geom_errorbar(aes(ymin=g_frac_mean-g_frac_se, ymax=g_frac_mean+g_frac_se), width=.2,
                  position=position_dodge(.9)) +
-  scale_fill_brewer(palette = "Blues") +
+  scale_fill_manual(values = colors) +
   labs(y = "",
@@
   ggplot(filter(sum_ssp_sub, species_code == "qup"), 
-         aes(x=factor(site, levels = sum_site$site), y=for_bites_median)) +
+         aes(x=factor(site, levels = sum_site$site), y=for_bites_mean)) +
   geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
-  geom_errorbar(aes(ymin=for_bites_median-for_bites_se, ymax=for_bites_median+for_bites_se), width=.2,
+  geom_errorbar(aes(ymin=for_bites_mean-for_bites_se, ymax=for_bites_mean+for_bites_se), width=.2,
                  position=position_dodge(.9)) +
-  scale_fill_brewer(palette = "Blues") +
+  scale_fill_manual(values = colors) +
   labs(y = "Grazing intensity\n(bites/foray)", 
@@
         axis.title.y = element_text(size=8),
-        axis.text.x = element_text(angle = 45, hjust = 1, size=7),
+        axis.text.x = element_text(angle = 45, hjust = 1, size=8),
         legend.position = "none")
@@
   ggplot(filter(sum_ssp_sub, species_code == "stop"), 
-         aes(x=factor(site, levels = sum_site$site), y=for_bites_median)) +
-  geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
-  geom_errorbar(aes(ymin=for_bites_median-for_bites_se, ymax=for_bites_median+for_bites_se), width=.2,
+         aes(x=factor(site, levels = sum_site$site), y=for_bites_mean)) +
+  geom_bar(stat = "identity", aes(fill = island), alpha = 0.9, position = "dodge") +
+  geom_errorbar(aes(ymin=for_bites_mean-for_bites_se, ymax=for_bites_mean+for_bites_se), width=.2,
                  position=position_dodge(.9)) +
-  scale_fill_brewer(palette = "Blues") +
+  scale_fill_manual(values = colors) +
   labs(y = "",
@@
   theme(plot.margin = margin(0, 4, 0, 0, unit = "pt"),
-        axis.text.x = element_text(angle = 45, hjust = 1, size=7),
+        axis.text.x = element_text(angle = 45, hjust = 1, size=8),
         legend.position = "none")
@@
   ggplot(filter(sum_ssp_sub, species_code == "stop"), 
-         aes(x=factor(site, levels = sum_site$site), y=g_frac_median)) +
-  geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
-  scale_fill_brewer(palette = "Blues") +
+         aes(x=factor(site, levels = sum_site$site), y=g_frac_mean)) +
+  geom_bar(stat = "identity", aes(fill = island), alpha = 0.9, position = "dodge") +
+  scale_fill_manual(values = colors) +
   labs(y = "",
@@
 g <- arrangeGrob(
-  arrangeGrob(fr_qup, fr_stop, gfrac_qup, gfrac_stop, br_qup, br_stop, fi_qup, fi_stop, ncol=2, heights = c(1,1,1,2)), 
+  arrangeGrob(fr_qup, fr_stop, gfrac_qup, gfrac_stop, br_qup, br_stop, fi_qup, fi_stop, ncol=2, heights = c(1,1,1,1.75)), 
               arrangeGrob(legend), 
@@
 
-ggsave(here("figs/behavior_panel.png"),g)
+ggsave(here("figs/behavior_panel_means.png"), g, width = 7, height = 7)
+```
+
+Testing with violin plots
+```{r, message = F, warning = F}
+colors <- c("turquoise4", "powderblue", "steelblue4")
+
+fr_qup <- 
+  ggplot(filter(sum_id, species_code == "qup"), 
+         aes(x=factor(site, levels = sum_site$site), y=fr)) +
+  geom_violin(aes(fill = island), alpha = 0.9) +
+  geom_boxplot(width = 0.1) +
+  scale_fill_manual(values = colors) +
+  labs(y = "Feeding rate \n (bites/min)", 
+       title = "S. vetula") +
+  theme_bw() + 
+  theme(plot.title = element_text(hjust = 0.5, face = "italic"),
+        plot.margin = margin(5, 0, 0, 5, unit = "pt"),
+        strip.text = element_text(face = "italic"), 
+        axis.text.x=element_blank(),
+        axis.title.x = element_blank(),
+        axis.title.y = element_text(size=8),
+        legend.position = "none")
+
+fr_stop <- 
+  ggplot(filter(sum_id, species_code == "stop"), 
+         aes(x=factor(site, levels = sum_site$site), y=fr)) +
+  geom_violin(aes(fill = island), alpha = 0.9) +
+  geom_boxplot(width = 0.1) +
+  scale_fill_manual(values = colors) +
+  labs(y = "",
+       # y = "S. vetula \nFeeding rate (bites/hr)", 
+       title = "S. viride") +
+  theme_bw() + 
+  theme(plot.title = element_text(hjust = 0.5, face = "italic"),
+        plot.margin = margin(5, 4, 0, 0, unit = "pt"),
+        strip.text = element_text(face = "italic"), 
+        axis.text.x=element_blank(),
+        axis.title.x = element_blank(),
+        legend.position = "none")
+
+br_qup <- 
+  ggplot(filter(sum_id, species_code == "qup"), 
+         aes(x=factor(site, levels = sum_site$site), y=br)) +
+  geom_violin(aes(fill = island), alpha = 0.9) +
+  geom_boxplot(width = 0.1) +
+  scale_fill_manual(values = colors) +
+  labs(y = "Bite rate \n(bites/sec)", 
+       x = "", 
+       title = "") +
+  theme_bw() + 
+  theme(plot.title = element_text(hjust = 0.5),
+        plot.margin = margin(0, 0, 0, 5, unit = "pt"),
+        strip.text = element_text(face = "italic"), 
+        axis.text.x=element_blank(),
+        axis.title.x = element_blank(),
+        axis.title.y = element_text(size=8),
+        legend.position = "none")
+
+br_stop <- 
+  ggplot(filter(sum_id, species_code == "stop"), 
+         aes(x=factor(site, levels = sum_site$site), y=br)) +
+  geom_violin(aes(fill = island), alpha = 0.9) +
+  geom_boxplot(width = 0.1) +
+  scale_fill_manual(values = colors) +
+  labs(y = "",
+       # y = "S. vetula \nFeeding rate (bites/hr)", 
+       x = "", 
+       title = "") +
+  theme_bw() + 
+  theme(plot.title = element_text(hjust = 0.5),
+        plot.margin = margin(0, 4, 0, 0, unit = "pt"),
+        strip.text = element_text(face = "italic"), 
+        axis.text.x=element_blank(),
+        axis.title.x = element_blank(),
+        legend.position = "none")
+
+gfrac_qup <- 
+  ggplot(filter(sum_id, species_code == "qup"), 
+         aes(x=factor(site, levels = sum_site$site), y=g_frac)) +
+  geom_violin(aes(fill = island), alpha = 0.9) +
+  geom_boxplot(width = 0.1) +
+  scale_fill_manual(values = colors) +
+  labs(y = "Time grazing \n(fraction)", 
+       x = "", 
+       title = "") +
+  theme_bw() + 
+  theme(plot.title = element_text(hjust = 0.5),
+        plot.margin = margin(0, 0, 0, 5, unit = "pt"),
+        strip.text = element_text(face = "italic"), 
+        axis.text.x=element_blank(),
+        axis.title.x = element_blank(),
+        axis.title.y = element_text(size=8),
+        legend.position = "none")
+
+gfrac_stop <- 
+  ggplot(filter(sum_id, species_code == "stop"), 
+         aes(x=factor(site, levels = sum_site$site), y=g_frac)) +
+  geom_violin(aes(fill = island), alpha = 0.9) +
+  geom_boxplot(width = 0.1) +
+  scale_fill_manual(values = colors) +
+  labs(y = "",
+       x = "", 
+       title = "") +
+  theme_bw() + 
+  theme(plot.title = element_text(hjust = 0.5),
+        plot.margin = margin(0, 4, 0, 0, unit = "pt"),
+        strip.text = element_text(face = "italic"), 
+        axis.text.x=element_blank(),
+        axis.title.x = element_blank(),
+        legend.position = "none")
+
+fi_qup <- 
+  ggplot(filter(sum_id, species_code == "qup"), 
+         aes(x=factor(site, levels = sum_site$site), y=for_bites)) +
+  geom_violin(aes(fill = island), alpha = 0.9) +
+  geom_boxplot(width = 0.1) +
+  scale_fill_manual(values = colors) +
+  labs(y = "Grazing intensity\n(bites/foray)", 
+       x = "", 
+       title = "") +
+  theme_bw() + 
+  theme(plot.margin = margin(0, 0, 0, 5, unit = "pt"),
+        axis.title.y = element_text(size=8),
+        axis.text.x = element_text(angle = 45, hjust = 1, size=7),
+        legend.position = "none")
+
+fi_stop <- 
+  ggplot(filter(sum_id, species_code == "stop"), 
+         aes(x=factor(site, levels = sum_site$site), y=for_bites)) +
+  geom_violin(aes(fill = island), alpha = 0.9) +
+  geom_boxplot(width = 0.1) +
+  scale_fill_manual(values = colors) +
+  labs(y = "",
+       x = "", 
+       title = "") +
+  theme_bw() + 
+  theme(plot.margin = margin(0, 4, 0, 0, unit = "pt"),
+        axis.text.x = element_text(angle = 45, hjust = 1, size=7),
+        legend.position = "none")
+
+legend_plot <- 
+  ggplot(filter(sum_id, species_code == "stop"), 
+         aes(x=factor(site, levels = sum_site$site), y=g_frac)) +
+  geom_violin(aes(fill = island), alpha = 0.9) +
+  geom_boxplot(width = 0.1) +
+  scale_fill_manual(values = colors) +
+  labs(y = "",
+       x = "", 
+       title = "") +
+  theme_bw() +
+  theme(legend.position = "right",
+        legend.title = element_blank())
+
+get_legend<-function(myggplot){
+  tmp <- ggplot_gtable(ggplot_build(myggplot))
+  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
+  legend <- tmp$grobs[[leg]]
+  return(legend)
+  }
+legend <- get_legend(legend_plot)
+
+g <- arrangeGrob(
+  arrangeGrob(fr_qup, fr_stop, gfrac_qup, gfrac_stop, br_qup, br_stop, fi_qup, fi_stop, ncol=2, heights = c(1,1,1,1.75)), 
+              arrangeGrob(legend), 
+              ncol=2, widths = c(5,1))
+plot(g)
+
+ggsave(here("figs/behavior_panel_violin.png"), g, width = 8, height = 8)
 ```
@@
 ```{r, message=F, warning=F}
-fr_q1 <- ggplot(filter(sum_ssp_sub, species_code == "qup"), aes(x = scar_bm/1000, y = fr_median)) +
+fr_q1 <- ggplot(filter(sum_ssp_sub, species_code == "qup"), aes(x = scar_bm/1000, y = fr_mean)) +
   geom_smooth(method="loess", span=1) +
@@
 
-fr_q2 <- ggplot(filter(sum_ssp_sub, species_code == "qup"), aes(x = consp_den, y = fr_median)) +
+fr_q2 <- ggplot(filter(sum_ssp_sub, species_code == "qup"), aes(x = carn30_bm/1000, y = fr_mean)) +
   geom_smooth(method="loess", span=1) +
@@
 
-fr_q3 <- ggplot(filter(sum_ssp_sub, species_code == "qup"), aes(x = rugosity, y = fr_median)) +
+fr_q3 <- ggplot(filter(sum_ssp_sub, species_code == "qup"), aes(x = rugosity, y = fr_mean)) +
   geom_smooth(method="loess", span=1) +
@@
 
-fr_q4 <- ggplot(filter(sum_ssp_sub, species_code == "qup"), aes(x = ta_canopy, y = fr_median)) +
+fr_q4 <- ggplot(filter(sum_ssp_sub, species_code == "qup"), aes(x = ta_canopy, y = fr_mean)) +
   geom_smooth(method="loess", span=1) +
@@
 
-fr_q5 <- ggplot(filter(sum_ssp_sub, species_code == "qup"), aes(x = ta_cover, y = fr_median)) +
+fr_q5 <- ggplot(filter(sum_ssp_sub, species_code == "qup"), aes(x = ta_cover, y = fr_mean)) +
   geom_smooth(method="loess", span=1) +
@@
 
-fr_q7 <- ggplot(filter(sum_ssp_sub, species_code == "qup"), aes(x = ma_cover, y = fr_median)) +
+fr_q7 <- ggplot(filter(sum_ssp_sub, species_code == "qup"), aes(x = ma_cover, y = fr_mean)) +
   geom_smooth(method="loess", span=1) +
@@
 
-fr_s1 <- ggplot(filter(sum_ssp_sub, species_code == "stop"), aes(x = scar_bm/1000, y = fr_median)) +
+fr_s1 <- ggplot(filter(sum_ssp_sub, species_code == "stop"), aes(x = scar_bm/1000, y = fr_mean)) +
   geom_smooth(method="loess", span=1) +
@@
         plot.margin = margin(0, 1, 0, 0, unit = "pt"),
-        axis.title.x = element_text(size = 9)
+        axis.title.x = element_text(size = 7)
         )
 
-fr_s2 <- ggplot(filter(sum_ssp_sub, species_code == "stop"), aes(x = consp_den, y = fr_median)) +
+fr_s2 <- ggplot(filter(sum_ssp_sub, species_code == "stop"), aes(x = carn30_bm/1000, y = fr_mean)) +
   geom_smooth(method="loess", span=1) +
@@
   labs(y = "", 
-       x = expression(atop(NA, atop(textstyle('Conspec. density'),
-                               textstyle('(indv/100m'^2*')')))),
+       x = expression(atop(NA, atop(textstyle('Predator biomass'),
+                               textstyle('(kg/100m'^2*')')))),
        title = "") +
@@
         axis.text.y = element_blank(),
-        axis.title.x = element_text(size = 9),
+        axis.title.x = element_text(size = 7),
         plot.margin = margin(0, 1, 0, 1, unit = "pt")
@@
 
-fr_s3 <- ggplot(filter(sum_ssp_sub, species_code == "stop"), aes(x = rugosity, y = fr_median)) +
+fr_s3 <- ggplot(filter(sum_ssp_sub, species_code == "stop"), aes(x = rugosity, y = fr_mean)) +
   geom_smooth(method="loess", span=1) +
@@
 
-fr_s4 <- ggplot(filter(sum_ssp_sub, species_code == "stop"), aes(x = ta_canopy, y = fr_median)) +
+fr_s4 <- ggplot(filter(sum_ssp_sub, species_code == "stop"), aes(x = ta_canopy, y = fr_mean)) +
   geom_smooth(method="loess", span=1) +
@@
   labs(y = "", 
-       x = expression(atop(NA, atop(textstyle('EAM canopy'),
+       x = expression(atop(NA, atop(textstyle('Turf canopy'),
                                textstyle('height (mm)')))), 
@@
 
-fr_s5 <- ggplot(filter(sum_ssp_sub, species_code == "stop"), aes(x = ta_cover, y = fr_median)) +
+fr_s5 <- ggplot(filter(sum_ssp_sub, species_code == "stop"), aes(x = ta_cover, y = fr_mean)) +
   geom_smooth(method="loess", span=1) +
@@
   labs(y = "", 
-       x = expression(atop(NA, atop(textstyle('EAM cover'),
+       x = expression(atop(NA, atop(textstyle('Turf cover'),
                                textstyle('(%)')))), 
@@
 
-fr_s7 <- ggplot(filter(sum_ssp_sub, species_code == "stop"), aes(x = ma_cover, y = fr_median)) +
+fr_s7 <- ggplot(filter(sum_ssp_sub, species_code == "stop"), aes(x = ma_cover, y = fr_mean)) +
   geom_smooth(method="loess", span=1) +
@@
   theme(plot.title = element_text(hjust = 0.5),
-        axis.title.x = element_text(size = 9),
+        axis.title.x = element_text(size = 7),
         axis.ticks.y = element_blank(),
@@
 
-label_vet = textGrob("S. vetula", gp=gpar(fontface="italic"), rot=270)
-label_vir = textGrob("S. viride", gp=gpar(fontface="italic"), rot=270)
+label_vet = textGrob("S. vetula", gp=gpar(fontface="italic", fontsize = 10), rot=270)
+label_vir = textGrob("S. viride", gp=gpar(fontface="italic", fontsize = 10), rot=270)
 
-g <- grid.arrange(left = "Feeding rate (bites/min.)",
-  arrangeGrob(fr_q1, fr_q3, fr_q4, fr_q5, fr_q7, ncol=5, right = label_vet),
-  arrangeGrob(fr_s1, fr_s3, fr_s4, fr_s5, fr_s7, ncol=5, right = label_vir), 
+g <- grid.arrange(left = textGrob("Feeding rate (bites/min.)", rot = 90, gp = gpar(fontsize = 10)),
+  arrangeGrob(fr_q1, fr_q2, fr_q3, fr_q4, fr_q5, fr_q7, ncol=6, right = label_vet),
+  arrangeGrob(fr_s1, fr_s2, fr_s3, fr_s4, fr_s5, fr_s7, ncol=6, right = label_vir), 
   nrow=2, heights = c(1,1.2))
 
-ggsave(here("figs/biplot_panel.png"),g)
+ggsave(here("figs/biplot_panel_means.png"), g, width = 7, height = 4)
+```
+### Carnivore Loess for supplement
+```{r}
+ggplot(filter(sum_ssp_sub, species_code == "stop"), aes(x = carn30_bm, y = fr_mean)) +
+  geom_smooth(method="loess", span=1) +
+  geom_point(shape=18) +
+  labs(y = "Feeding rate (bites/min.)",
+       x = expression(atop(NA, atop(textstyle('Predator abundance')))), 
+       title = "") +
+  theme_bw() + 
+  theme(plot.title = element_text(hjust = 0.5)
+        )
+
+ggplot(filter(sum_ssp_sub, species_code == "qup"), aes(x = carn30_bm, y = fr_mean)) +
+  geom_smooth(method="loess", span=1) +
+  geom_point(shape=18) +
+  labs(y = "Feeding rate (bites/min.)",
+       x = expression(atop(NA, atop(textstyle('Predator abundance')))), 
+       title = "") +
+  theme_bw() + 
+  theme(plot.title = element_text(hjust = 0.5)
+        )
 ```
@@
 
+set.seed(5)
 fviz_pca_biplot(pca_pred,
@@
                 label = "var",
-                repel = T
+                repel = T,
+                pointsize = 3,
+                alpha.var= 0.3,
+                col.var = "gray22",
+                labelsize = 4,
+                axes.linetype = "dotted"
                 ) +
-  scale_shape_manual(values=c(15,16,17)) +
-  scale_color_manual(values = c("black","black","black")) +
+  scale_shape_manual(values = c(15,16,17)) +
+  scale_color_manual(values = colors) +
   labs(x = "PC1 (55.6%)", y = "PC2 (24.5%)", title = "") +
@@
 
-Predicting feeding rate (site-level medians, island as random effect)
+Predicting feeding rate (site-level means, island as random effect)
 ```{r, message=F, warning=F}
@@
 
-gamm1 <- gamm(fr_median ~ species_code + s(length_mean, k=3) + s(pc1_all, k=3) + s(pc2_all, k=3), 
-           family = gaussian(link=identity),
-           random = list(island=~1),
-           data = m_data
-           )
-g1 <- c("g1", "Species + Size + PC1 + PC2", AICc(gamm1$lme), summary(gamm1$gam)$r.sq)
+# gamm1 <- gamm(fr_mean ~ species_code + s(length_mean, k=3) + s(pc1_all, k=3) + s(pc2_all, k=3), 
+#            family = gaussian(link=identity),
+#            random = list(island=~1),
+#            data = m_data
+#            )
+# g1 <- c("g1", "Species + Size + PC1 + PC2", AICc(gamm1$lme), summary(gamm1$gam)$r.sq)
 
-gamm2 <- gamm(fr_median ~ species_code + s(length_mean, k=3) + s(pc1_all, k=3), 
+gamm2 <- gamm(fr_mean ~ species_code + s(length_mean, k=3) + s(pc1_all, k=3), 
            family = gaussian(link=identity),
@@
 
-gamm3 <- gamm(fr_median ~ species_code + s(length_mean, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3) + s(pc2_all, by = factor(species_code), k=3),
-           family = gaussian(link=identity),
-           random = list(island=~1),
-           data = m_data
-           )
-g3 <- c("g3", "Species + Size~Species + PC1~Species + PC2~Species", AICc(gamm3$lme), summary(gamm3$gam)$r.sq)
+# gamm3 <- gamm(fr_mean ~ species_code + s(length_mean, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3) + s(pc2_all, by = factor(species_code), k=3),
+#            family = gaussian(link=identity),
+#            random = list(island=~1),
+#            data = m_data
+#            )
+# g3 <- c("g3", "Species + Size~Species + PC1~Species + PC2~Species", AICc(gamm3$lme), summary(gamm3$gam)$r.sq)
 
-gamm4 <- gamm(fr_median ~ species_code + s(length_mean, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3) + s(pc2_all, k=3),
-           family = gaussian(link=identity),
-           random = list(island=~1),
-           data = m_data
-           )
-g4 <- c("g4", "Species + Size~Species + PC1~Species + PC2", AICc(gamm4$lme), summary(gamm4$gam)$r.sq)
+# gamm4 <- gamm(fr_mean ~ species_code + s(length_mean, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3) + s(pc2_all, k=3),
+#            family = gaussian(link=identity),
+#            random = list(island=~1),
+#            data = m_data
+#            )
+# g4 <- c("g4", "Species + Size~Species + PC1~Species + PC2", AICc(gamm4$lme), summary(gamm4$gam)$r.sq)
 
-gamm5 <- gamm(fr_median ~ species + s(length_mean, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3),
+gamm5 <- gamm(fr_mean ~ species + s(length_mean, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3),
            family = gaussian(link=identity),
@@
 
-gamm6 <- gamm(fr_median ~ s(length_mean, k=3) + s(pc1_all, by = factor(species_code), k=3), 
+gamm6 <- gamm(fr_mean ~ s(length_mean, k=3) + s(pc1_all, by = factor(species_code), k=3), 
            family = gaussian(link=identity),
@@
 
-gamm7 <- gamm(fr_median ~ s(length_mean, by = factor(species_code), k=3) + s(pc1_all, k=3), 
+gamm7 <- gamm(fr_mean ~ s(length_mean, by = factor(species_code), k=3) + s(pc1_all, k=3), 
            family = gaussian(link=identity),
@@
 
-gamm8 <- gamm(fr_median ~ species + s(length_mean, by = factor(species_code), k=3) + s(pc1_all, k=3), 
+gamm8 <- gamm(fr_mean ~ species + s(length_mean, by = factor(species_code), k=3) + s(pc1_all, k=3), 
            family = gaussian(link=identity),
@@
 
-gamm9 <- gamm(fr_median ~ species + s(length_mean, k=3) + s(pc1_all, by = factor(species_code), k=3), 
+gamm9 <- gamm(fr_mean ~ species + s(length_mean, k=3) + s(pc1_all, by = factor(species_code), k=3), 
            family = gaussian(link=identity),
@@
 
-gamm10 <- gamm(fr_median ~ species + s(pc1_all, by = factor(species_code), k=3), 
+gamm10 <- gamm(fr_mean ~ species + s(pc1_all, by = factor(species_code), k=3), 
            family = gaussian(link=identity),
@@
 
-gamm11 <- gamm(fr_median ~ s(length_mean, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3), 
+gamm11 <- gamm(fr_mean ~ s(length_mean, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3), 
            family = gaussian(link=identity),
@@
 
-gamm12 <- gamm(fr_median ~ species_code + s(pc1_all, k=3), 
+gamm12 <- gamm(fr_mean ~ species_code + s(pc1_all, k=3), 
            family = gaussian(link=identity),
@@
 
-gtab_fr_site <- as.data.frame(rbind(g1,g2,g3,g4,g5,g6,g7,g8,g9,g10,g11,g12)) %>% setNames(c("g","Model", "AICc", "Adj. R2")) %>% arrange(AICc)
+gtab_fr_site <- as.data.frame(rbind(g2,g5,g6,g7,g8,g9,g10,g11,g12)) %>% setNames(c("g","Model", "AICc", "Adj. R2")) %>% arrange(AICc)
 kable(gtab_fr_site)
@@
 
-Predicting grazing intensity (site-level medians, island as random effect)
+Predicting grazing intensity (site-level means, island as random effect)
 ```{r, echo=F, message=F, warning=F}
@@
 
-gamm1 <- gamm(for_bites_median ~ species_code + s(length_mean, k=3) + s(pc1_all, k=3) + s(pc2_all, k=3), 
+gamm1 <- gamm(for_bites_mean ~ species_code + s(length_mean, k=3) + s(pc1_all, k=3) + s(pc2_all, k=3), 
            family = gaussian(link=identity),
@@
 
-gamm2 <- gamm(for_bites_median ~ species_code + s(length_mean, k=3) + s(pc1_all, k=3), 
+gamm2 <- gamm(for_bites_mean ~ species_code + s(length_mean, k=3) + s(pc1_all, k=3), 
            family = gaussian(link=identity),
@@
 
-gamm3 <- gamm(for_bites_median ~ species_code + s(length_mean, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3) + s(pc2_all, by = factor(species_code), k=3),
+gamm3 <- gamm(for_bites_mean ~ species_code + s(length_mean, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3) + s(pc2_all, by = factor(species_code), k=3),
            family = gaussian(link=identity),
@@
 
-# gamm4 <- gamm(for_bites_median ~ species_code + s(length_mean, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3) + s(pc2_all, k=3),
+# gamm4 <- gamm(for_bites_mean ~ species_code + s(length_mean, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3) + s(pc2_all, k=3),
 #            family = gaussian(link=identity),
@@
 
-gamm5 <- gamm(for_bites_median ~ species + s(length_mean, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3), 
+gamm5 <- gamm(for_bites_mean ~ species + s(length_mean, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3), 
            family = gaussian(link=identity),
@@
 
-gamm6 <- gamm(for_bites_median ~ s(length_mean, k=3) + s(pc1_all, by = factor(species_code), k=3), 
+gamm6 <- gamm(for_bites_mean ~ s(length_mean, k=3) + s(pc1_all, by = factor(species_code), k=3), 
            family = gaussian(link=identity),
@@
 
-gamm7 <- gamm(for_bites_median ~ s(length_mean, by = factor(species_code), k=3) + s(pc1_all, k=3), 
+gamm7 <- gamm(for_bites_mean ~ s(length_mean, by = factor(species_code), k=3) + s(pc1_all, k=3), 
            family = gaussian(link=identity),
@@
 
-gamm8 <- gamm(for_bites_median ~ species + s(length_mean, by = factor(species_code), k=3) + s(pc1_all, k=3), 
+gamm8 <- gamm(for_bites_mean ~ species + s(length_mean, by = factor(species_code), k=3) + s(pc1_all, k=3), 
            family = gaussian(link=identity),
@@
 
-gamm9 <- gamm(for_bites_median ~ species + s(length_mean, k=3) + s(pc1_all, by = factor(species_code), k=3), 
+gamm9 <- gamm(for_bites_mean ~ species + s(length_mean, k=3) + s(pc1_all, by = factor(species_code), k=3), 
            family = gaussian(link=identity),
@@
 
-gamm10 <- gamm(for_bites_median ~ species + s(pc1_all, by = factor(species_code), k=3), 
+gamm10 <- gamm(for_bites_mean ~ species + s(pc1_all, by = factor(species_code), k=3), 
            family = gaussian(link=identity),
@@
 
-gamm11 <- gamm(for_bites_median ~ s(length_mean, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3), 
+gamm11 <- gamm(for_bites_mean ~ s(length_mean, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3), 
            family = gaussian(link=identity),
@@
 
-gamm12 <- gamm(for_bites_median ~ species_code + s(pc1_all, k=3), 
+gamm12 <- gamm(for_bites_mean ~ species_code + s(pc1_all, k=3), 
            family = gaussian(link=identity),

### 2021-01-09 r/old drafts/Paper_analyses_medians.Rmd
file organization

--- 2019-10-01 r/Ch2 outline_V3.Rmd
@@
 g <- arrangeGrob(
-  arrangeGrob(fr_qup, fr_stop, gfrac_qup, gfrac_stop, br_qup, br_stop, fi_qup, fi_stop, ncol=2, heights = c(1,1,1,1.5)), 
+  arrangeGrob(fr_qup, fr_stop, gfrac_qup, gfrac_stop, br_qup, br_stop, fi_qup, fi_stop, ncol=2, heights = c(1,1,1,2)), 
               arrangeGrob(legend), 
               ncol=2, widths = c(5,1))
-
+plot(g)
 ggsave(here("figs/behavior_panel.png"),g)
@@
 
-## Kruskal-Wallis tests
+```{r}
+# indv plots for slides
+
+library(ggpubr)
+library(ggtext)
+
+fr_qup <- 
+  ggplot(filter(sum_ssp_sub, species_code == "qup"), 
+         aes(x=factor(site, levels = sum_site$site), y=fr_median)) +
+  geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
+  geom_errorbar(aes(ymin=fr_median-fr_se, ymax=fr_median+fr_se), width=.2,
+                 position=position_dodge(.9)) +
+  scale_fill_brewer(palette = "Blues", name = "Island") +
+  labs(y = "", 
+       title = "*S. vetula* (queen parrotfish)") +
+  theme_bw() + 
+  theme(plot.title = element_markdown(hjust = 0.5),
+        plot.margin = margin(5, 0, 0, 5, unit = "pt"),
+        strip.text = element_text(face = "italic"), 
+        axis.text.x = element_text(angle = 45, hjust = 1),
+        axis.title.x = element_blank()
+        )
+
+fr_stop <- 
+  ggplot(filter(sum_ssp_sub, species_code == "stop"), 
+         aes(x=factor(site, levels = sum_site$site), y=fr_median)) +
+  geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
+  geom_errorbar(aes(ymin=fr_median-fr_se, ymax=fr_median+fr_se), width=.2,
+                 position=position_dodge(.9)) +
+  scale_fill_brewer(palette = "Blues", name = "Island") +
+  labs(y = expression(Feeding~rate~(bites~min^-1)),
+       title = "*S. viride* (stoplight parrotfish)") +
+  theme_bw() + 
+  theme(plot.title = element_markdown(hjust = 0.5),
+        plot.margin = margin(5, 4, 0, 0, unit = "pt"),
+        strip.text = element_text(face = "italic"), 
+        axis.text.x = element_text(angle = 45, hjust = 1),
+        axis.title.x = element_blank(),
+        axis.title.y = element_text(size=10)
+        )
+
+ggarrange(fr_stop, fr_qup, 
+          #abels = c("A", "B"),
+          ncol = 2, nrow = 1,
+          common.legend = TRUE, legend = "right")
+
+ggsave(here("figs", "fr.png"), height = 4, width = 8)
+
+fi_qup <- 
+  ggplot(filter(sum_ssp_sub, species_code == "qup"), 
+         aes(x=factor(site, levels = sum_site$site), y=for_bites_median)) +
+  geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
+  geom_errorbar(aes(ymin=for_bites_median-for_bites_se, ymax=for_bites_median+for_bites_se), width=.2,
+                 position=position_dodge(.9)) +
+  scale_fill_brewer(palette = "Blues", name = "Island") +
+  labs(y = "", 
+       title = "*S. vetula* (queen parrotfish)") +
+  theme_bw() + 
+  theme(plot.title = element_markdown(hjust = 0.5),
+        plot.margin = margin(5, 0, 0, 5, unit = "pt"),
+        strip.text = element_text(face = "italic"), 
+        axis.text.x = element_text(angle = 45, hjust = 1),
+        axis.title.x = element_blank()
+        )
+
+fi_stop <- 
+  ggplot(filter(sum_ssp_sub, species_code == "stop"), 
+         aes(x=factor(site, levels = sum_site$site), y=for_bites_median)) +
+  geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
+  geom_errorbar(aes(ymin=for_bites_median-for_bites_se, ymax=for_bites_median+for_bites_se), width=.2,
+                 position=position_dodge(.9)) +
+  scale_fill_brewer(palette = "Blues", name = "Island") +
+  labs(y = expression(Grazing~intensity~(bites~foray^-1)),
+       title = "*S. viride* (stoplight parrotfish)") +
+  theme_bw() + 
+  theme(plot.title = element_markdown(hjust = 0.5),
+        plot.margin = margin(5, 4, 0, 0, unit = "pt"),
+        strip.text = element_text(face = "italic"), 
+        axis.text.x = element_text(angle = 45, hjust = 1),
+        axis.title.x = element_blank(),
+        axis.title.y = element_text(size=10)
+        )
+
+ggarrange(fi_stop, fi_qup, 
+          #abels = c("A", "B"),
+          ncol = 2, nrow = 1,
+          common.legend = TRUE, legend = "right")
+
+ggsave(here("figs", "fi.png"), height = 4, width = 8)
+```
+
+Kruskal-Wallis tests to look at significance of inter-site variance for different behavioral variables:
+
+Feeding rate:
 ```{r}

### 2021-01-09 r/grazing intensity.R

@@ -1,219 +1,0 @@
-for_bites <- sum_id %>%
-  select(id, for_bites, site) %>%
-  mutate(for_bites_binned = cut(for_bites, breaks = c(0,10,30,65), labels = c("low","med","high")),
-         for_bites_rounded = round(for_bites, 0)
-         )
-
-ggplot(for_bites, aes(x = for_bites_rounded, fill = factor(site))) +
-  geom_histogram() +
-  geom_density(alpha=0.6) +
-  theme_bw() +
-  labs(x = "Grazing intensity (bites/foray)")
-
-ggsave(here("figs/intensity_hist.png"), width = 3, height = 3)
-
-ggplot(sum_ssp_sub, aes(x = for_bites_median, y = fr_median)) +
-  geom_point() +
-  labs(x = "Grazing intensity (bites/foray)", y = "Feeding rate (bites/min)")
-ggsave(here("figs/intensity_vs_feedingrate.png"), width = 3, height = 3)
-
-ggplot(sum_id, aes(x = for_bites, y = fr)) +
-  geom_point(aes(colour = factor(site)))
-
-## Loess plots of various drivers for grazing intensity
-
-gi_q1 <- ggplot(filter(sum_ssp_sub, species_code == "qup"), aes(x = scar_bm/1000, y = for_bites_median)) +
-  geom_smooth(method="loess", span=1) +
-  geom_point(shape=18) +
-  ylim(-5, 25) +
-  labs(y = "",
-       title = "") +
-  theme_bw() + 
-  theme(plot.title = element_text(hjust = 0.5),
-        plot.margin = margin(0, 1, 0, 0, unit = "pt"),
-        axis.text.x = element_blank(),
-        axis.title.x = element_blank()
-  )
-
-gi_q2 <- ggplot(filter(sum_ssp_sub, species_code == "qup"), aes(x = consp_den, y = for_bites_median)) +
-  geom_smooth(method="loess", span=1) +
-  geom_point(shape=18) +
-  ylim(-5, 25) +
-  labs(y = "", 
-       title = "") +
-  theme_bw() + 
-  theme(plot.title = element_text(hjust = 0.5),
-        axis.ticks.y = element_blank(),
-        axis.text.y = element_blank(),
-        plot.margin = margin(0, 1, 0, 1, unit = "pt"),
-        axis.text.x = element_blank(),
-        axis.title.x = element_blank()
-  )
-
[…]

### 2021-01-09 r/pca_code.R

@@ -1,40 +1,0 @@
-# PCA options!
-
-### loading any data file with variables as columns/observations as rows
-site <- read.csv(here("data", "site.csv"), strip.white = T, stringsAsFactors = F)
-
-### set up dataframe with only variables that you want to include in the PCA 
-pca_data <- site %>%
-  # so that row/column names are ready to use as labels, if so desired:
-  remove_rownames %>% column_to_rownames(var = "site") %>%
-  select(rugosity, ma_cover, ma_canopy, ta_cover, ta_canopy, lc_cover, scar_bm, scar_den, carn30_bm) %>%
-  na.omit() %>%
-  # note: fviz_pca_biplot can handle names with spaces (e.g., "macroagal cover"), but I've had issues with spaces and autoplot() 
-  setNames(c("rugosity", "macroalgal_cover", "macroalgal_canopy", "turf_cover", "turf_canopy", "coral_cover", "scarid_biomass", "scarid_density", "predator_biomass"))
-
-### this does actual PCA
-pca_pred <- prcomp(pca_data, center = T, scale = T)
-
-### option A for visualization (this one has variables labeled but not sites)
-library(factoextra)
-fviz_pca_biplot(pca_pred,
-                col.ind = site$island, 
-                mean.point = F,
-                label = "var",
-                repel = T) +
-  scale_shape_manual(values=c(15,16,17)) + # changing shape of points by group
-  scale_color_manual(values = c("black","black","black")) +
-  labs(x = "PC1 (55.6%)", y = "PC2 (24.5%)", title = "") +
-  theme_bw() +
-  theme(legend.title = element_blank())
-
-### option B for visualization (this one has sites labeled)
-library(ggfortify)
-autoplot(pca_pred, data = site, shape = 'island', colour = 'island',
-         label = T,
-         loadings = T, loadings.colour = 'gray',
-         loadings.label = T, loadings.label.size = 3, loadings.label.colour = 'black',
-         loadings.label.repel = T) +
-  scale_color_manual(values = c("cadetblue3","cadetblue3","cadetblue3")) +
-  theme_bw() +
-  theme(legend.title = element_blank())

### 2021-01-16 r/Paper_analyses.Rmd
clearing out old files

@@
 
-## GAMM results (see 5.8.19)
+## GAMM results
 
@@
 # summary + plots from best models
+plot(gamm12$gam,pages=1)
+
+summary(gamm12$gam)
 summary(gamm10$gam)
-plot(gamm10$gam,pages=1)
+summary(gamm2$gam)
 ```
@@
 m_data <- sum_ssp_pca
+remove(g1, g2, g4, g6, g7, g8, g9, g10, g11, g12)
 
@@
 gtab_fi_site <- as.data.frame(rbind(g1,g2,g3,g5,g6,g7,g8,g9,g10,g11,g12)) %>% setNames(c("g","Model", "AICc", "Adj. R2")) %>% arrange(AICc)
-kable(gtab_fr_site)
+kable(gtab_fi_site)
 
@@
 summary(gamm12$gam)
+summary(gamm10$gam)
+summary(gamm2$gam)
 ```

### 2021-01-16 r/old_code/Ch2 outline_V2.Rmd
clearing out old files

---
title: "Scarid grazing behavior - manuscript outline"
author: "Molly Wilson"
date: "Jan. 2019"
output:
  html_document:
    number_sections: yes
    theme: united
    toc: yes
    toc_float: yes
---

```{r, echo=F, message=F, warning=F}
library(startR)
library(here)
library(tidyverse)
library(gridExtra)
library(ggfortify) # autoplot(prcomp())
library(mgcv)
library(sme)
library(MuMIn) # AICc
library(knitr)

sum_site <- read.csv(here("data","sum_site.csv"), stringsAsFactors=F) %>%
  select(-X) %>% 
  mutate(island = factor(island, levels = c("Bonaire","Antigua","Barbuda"))) %>%
  arrange(island)
sum_id <- read.csv(here("data","sum_id.csv"), stringsAsFactors=F) %>%
  select(-X) %>% 
  mutate(island = factor(island, levels = c("Bonaire","Antigua","Barbuda"))) %>%
  arrange(island) %>%
  filter(fr <= 3000)
sum_ssp <- read.csv(here("data","sum_site_sp_ph.csv"), stringsAsFactors=F) %>%
  select(-X) %>% 
  mutate(island = factor(island, levels = c("Bonaire","Antigua","Barbuda"))) %>%
  arrange(island)
```

# Intro + motivation

* Grazing by herbivorous fish is an established driver of coral reef health/benthic dynamics
* As reefs become increasingly degraded/threatened, there is growing interest in understanding how much herbivory is needed to sustain coral reef health (e.g. herbivore threshold papers - Karr et al. 2015, McClanahan et al. 2011)
* However, not all herbivores are the same in terms of their grazing impact…
    * Recent developments in functional classifications, difference across species in terms of feeding selectivity, morphology, etc. (e.g. Adam et al. 2018)
* In addition to these inter-species differences in functional role, differences in herbivore behavior also affect grazing impact and may drive variation in herbivore function among individuals or across reef systems
* Previous work has shown that herbivore behavior is sensitive to...
    * Predator presence (and reef structure) (Catano, Madin)
    * Presence of other herbivores (Gil)
    * Resource availability + distribution (Carlson, Tootell, Mumby, etc.)
* In understanding and predicting herbivore impact on reefs, current assumption is that grazing behaviors are constant within species (same species-level length~feeding rate relationships)
* But there is very limited data on how herbivore feeding behaviors might vary across reef systems and whether we can accurately assume constant species roles at regional scales
* Here we document variability in multiple components (*if I include foraging intensity as well as feeding rate...*) of scarid feeding behavior across reef sites in three Caribbean islands and discuss potential underlying mechanisms, ecological consequences, and management implications
* We demonstrate a need for further experimental and modelling work to unpack these drivers and understand their ecological consequences
* We urge for the incorporation of behavioral effects in coral reef herbivore management and caution against assuming constant species function across reef systems

# Methods

## Data collection
* Sites (13 sites, 3 islands [map]), timeline of data collection
* Fish and benthic community surveys
* Observational behavioral data:
    - Species: *Sparisoma viride* and *Scarus vetula*, initial and terminal phase greater than 15cm, focus on only initial phase here
    - 2 minute follows, calibration period, outside of previously established FID
    - Recording time budgets as well as bite rates during grazing forays (define)
    - Discarded data of incomplete follows or where fish were visably disturbed/fled
    - Timing of data collection distributed throughout day, 2 dives per site
    - Define behavioral variables ([table])

## Analysis
* Fish biomass using lenth-weight conversions (appendix)
* Site-level behavioral summary stats taken from subset of individuals based on size to avoid bias
* PCA of site-level benthic and fish community variables
* GAMM: mgcv, island as random effect, Gaussian distribution, compared models using AICc (AIC corrected for small sample size, Burnham & Anderson 2002). Number of knots set to 4 to allow for nonlinearities but prevent overfitting (Zuur et al. 2009). Used data aggregated to the site level (mean). Average length as covariate. 
* Bite content: calculated proxy for bite content by multiplying estimated bite sizes (based on species and length of each individual fish surveyed, equations from Bruggemann/Mumby) by site level averages of turf algal canopy height to account for potential differences in amount of algae consumed in low vs. high canopy sites

# Results
*(Will integrate with discussion depending on target journal)*

## Variation among sites
Notes:

* Can remove color coding by island if this is distracting. Unsure whether I should be highlighting differences between islands or if I should just be highlighting variation among sites (mixed feedback on this)
* Eventually I'll have a panel that shows differences in feeding rates, time spent feeding, and grazing intensity (trends are similar, but slightly different implications)  
* (Will remove Rendezvous from S. vetula graph in final - no S. vetulas found at this site)

```{r, echo=F, message=F, warning=F}
ggplot(filter(sum_ssp, species_code != "rbp" & phase == "i"), aes(x=factor(site, levels = sum_site$site), y=fr_mean)) +
    geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
    geom_errorbar(aes(ymin=fr_mean-fr_se, ymax=fr_mean+fr_se), width=.2,
                 position=position_dodge(.9)) +
    facet_grid(~species) +
    scale_fill_brewer(palette = "Blues") +
    labs(title="Mean feeding rates of individuals (initial phase)", 
         y="Feeding rate (bites/hr)", 
         x="Site") +
    theme_bw() +
    theme(strip.text = element_text(face = "italic"), 
          axis.text.x = element_text(angle = 90, hjust = 1), 
          plot.title = element_text(hjust = 0.5))
```

Demonstrated variance among sites (/islands), what factors might be contributing to these trends?

## Variation in feeding rate across benthic and fish community variables
*Is it worth including this panel, or more productive to focus on GAMMs...?*

* Initial phase only
* Don't need to include all of these
* Total scarid density, conspecific density, and predator biomass had even more variance - didn't include here for scaling purposes 
* Have similar panel for grazing intensity, though probably redundant

```{r, echo=F, message=F, warning=F}
fr_q1 <- ggplot(filter(sum_ssp, species_code == "qup" & phase == "i"), aes(x = scar_bm/1000, y = fr_mean/60)) +
  geom_smooth(method="loess", span=1) +
  geom_point(shape=18) +
  ylim(-10, 60) +
  labs(y = "",
       # y = "S. vetula \nFeeding rate (bites/min)", 
       x = "", 
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        plot.margin = margin(0, 1, 0, 1, unit = "pt")
        )

fr_q2 <- ggplot(filter(sum_ssp, species_code == "qup" & phase == "i"), aes(x = consp_den, y = fr_mean/60)) +
  geom_smooth(method="loess", span=1) +
  geom_point(shape=18) +
  ylim(-10, 60) +
  labs(y = "", 
       x = "",
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        plot.margin = margin(1, 1, 0, 1, unit = "pt")
        )

fr_q3 <- ggplot(filter(sum_ssp, species_code == "qup" & phase == "i"), aes(x = rugosity, y = fr_mean/60)) +
  geom_smooth(method="loess", span=1) +
  geom_point(shape=18) +
  ylim(-10, 60) +
  labs(y = "", 
       x = "", 
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        plot.margin = margin(1, 1, 0, 1, unit = "pt")
        )

fr_q4 <- ggplot(filter(sum_ssp, species_code == "qup" & phase == "i"), aes(x = ta_canopy, y = fr_mean/60)) +
  geom_smooth(method="loess", span=1) +
  geom_point(shape=18) +
  ylim(-10, 60) +
  labs(y = "", 
       x = "",
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        plot.margin = margin(1, 1, 0, 1, unit = "pt")
        )

fr_q5 <- ggplot(filter(sum_ssp, species_code == "qup" & phase == "i"), aes(x = ta_cover, y = fr_mean/60)) +
  geom_smooth(method="loess", span=1) +
  geom_point(shape=18) +
  ylim(-10, 60) +
  labs(y = "",
       x = "",
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        plot.margin = margin(1, 1, 0, 1, unit = "pt")
        )

fr_q7 <- ggplot(filter(sum_ssp, species_code == "qup" & phase == "i"), aes(x = ma_cover, y = fr_mean/60)) +
  geom_smooth(method="loess", span=1) +
  geom_point(shape=18) +
  ylim(-10, 60) +
  labs(y = "", 
       x = "",
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        plot.margin = margin(1, 1, 0, 1, unit = "pt")
        )

fr_s1 <- ggplot(filter(sum_ssp, species_code == "stop" & phase == "i"), aes(x = scar_bm/1000, y = fr_mean/60)) +
  geom_smooth(method="loess", span=1) +
  geom_point(shape=18) +
  ylim(-5, 20) +
  labs(y = "",
       x = "",
       # y = "S. viride \nFeeding rate (bites/min)", 
       # x = expression(atop(Scarid~biomass, "(kg/100m"^2*")")),
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        plot.margin = margin(0, 1, 0, 1, unit = "pt")
        )

fr_s2 <- ggplot(filter(sum_ssp, species_code == "stop" & phase == "i"), aes(x = consp_den, y = fr_mean/60)) +
  geom_smooth(method="loess", span=1) +
  geom_point(shape=18) +
  ylim(-5, 20) +
  labs(y = "", 
       x = "",
       # x = expression(atop(Conspec.~density, "(indv./100m"^2*")")),
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        plot.margin = margin(1, 1, 0, 1, unit = "pt")
        )

fr_s3 <- ggplot(filter(sum_ssp, species_code == "stop" & phase == "i"), aes(x = rugosity, y = fr_mean/60)) +
  geom_smooth(method="loess", span=1) +
  geom_point(shape=18) +
  ylim(-5, 20) +
  labs(y = "", 
       x = "",
       #x = expression("Rugosity"), 
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        plot.margin = margin(1, 1, 0, 1, unit = "pt")
        )

fr_s4 <- ggplot(filter(sum_ssp, species_code == "stop" & phase == "i"), aes(x = ta_canopy, y = fr_mean/60)) +
  geom_smooth(method="loess", span=1) +
  geom_point(shape=18) +
  ylim(-5, 20) +
  labs(y = "", 
       x = "",
       # x = expression("EAM canopy \nheight (mm)"), 
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        plot.margin = margin(1, 1, 0, 1, unit = "pt")
        )

fr_s5 <- ggplot(filter(sum_ssp, species_code == "stop" & phase == "i"), aes(x = ta_cover, y = fr_mean/60)) +
  geom_smooth(method="loess", span=1) +
  geom_point(shape=18) +
  ylim(-5, 20) +
  labs(y = "", 
       x = "",
       # x = expression("EAM percent \ncover"), 
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        plot.margin = margin(1, 1, 0, 1, unit = "pt")
        )

fr_s7 <- ggplot(filter(sum_ssp, species_code == "stop" & phase == "i"), aes(x = ma_cover, y = fr_mean/60)) +
  geom_smooth(method="loess", span=1) +
  geom_point(shape=18) +
  ylim(-5, 20) +
  labs(y = "", 
       x = "",
       # x = expression("Macroalgal \npercent cover"), 
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        plot.margin = margin(1, 1, 0, 1, unit = "pt")
        )

grid.arrange(left = "Feeding rate (bites / minute)",
  arrangeGrob(fr_q1, fr_q3, fr_q4, fr_q5, fr_q7, ncol=5, right = "S. vetula"),
  arrangeGrob(
    arrangeGrob(fr_s1, bottom = "Scarid biomass"), 
    arrangeGrob(fr_s3, bottom = "Rugosity"),
    arrangeGrob(fr_s4, bottom = "Turf height"),
    arrangeGrob(fr_s5, bottom = "Turf % cover"),
    arrangeGrob(fr_s7, bottom = "Macro. % cover"),
    ncol=5, right = "S. viride"), 
  nrow=2) 
```

Many different factors could be contributing, no clear dominant driver

```{r, eval=F, echo=F, message=F, warning=F}
## Variation in feeding intensity across benthic and fish community variables

fi_q1 <- ggplot(filter(sum_ssp, species_code == "qup" & phase == "i"), aes(x = scar_bm/1000, y = for_bites_mean)) +
  geom_smooth(method="loess", span=1) +
  geom_point(shape=18) +
  ylim(-10, 25) +
  labs(y = "",
       # y = "S. vetula \nFeeding rate (bites/min)", 
       x = "", 
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        plot.margin = margin(1, 1, 0, 1, unit = "pt")
        )

fi_q2 <- ggplot(filter(sum_ssp, species_code == "qup" & phase == "i"), aes(x = consp_den, y = for_bites_mean)) +
  geom_smooth(method="loess", span=1) +
  geom_point(shape=18) +
  ylim(-10, 25) +
  labs(y = "", 
       x = "",
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        plot.margin = margin(1, 1, 0, 1, unit = "pt")
        )

fi_q3 <- ggplot(filter(sum_ssp, species_code == "qup" & phase == "i"), aes(x = rugosity, y = for_bites_mean)) +
  geom_smooth(method="loess", span=1) +
  geom_point(shape=18) +
  ylim(-10, 25) +
  labs(y = "", 
       x = "", 
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        plot.margin = margin(1, 1, 0, 1, unit = "pt")
        )

fi_q4 <- ggplot(filter(sum_ssp, species_code == "qup" & phase == "i"), aes(x = ta_canopy, y = for_bites_mean)) +
  geom_smooth(method="loess", span=1) +
  geom_point(shape=18) +
  ylim(-10, 25) +
  labs(y = "", 
       x = "",
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        plot.margin = margin(1, 1, 0, 1, unit = "pt")
        )

fi_q5 <- ggplot(filter(sum_ssp, species_code == "qup" & phase == "i"), aes(x = ta_cover, y = for_bites_mean)) +
  geom_smooth(method="loess", span=1) +
  geom_point(shape=18) +
  ylim(-10, 25) +
  labs(y = "",
       x = "",
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        plot.margin = margin(1, 1, 0, 1, unit = "pt")
        )

fi_q7 <- ggplot(filter(sum_ssp, species_code == "qup" & phase == "i"), aes(x = ma_cover, y = for_bites_mean)) +
  geom_smooth(method="loess", span=1) +
  geom_point(shape=18) +
  ylim(-10, 25) +
  labs(y = "", 
       x = "",
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        plot.margin = margin(1, 1, 0, 1, unit = "pt")
        )

fi_s1 <- ggplot(filter(sum_ssp, species_code == "stop" & phase == "i"), aes(x = scar_bm/1000, y = for_bites_mean)) +
  geom_smooth(method="loess", span=1) +
  geom_point(shape=18) +
  ylim(-10, 25) +
  labs(y = "",
       x = "",
       # y = "S. viride \nFeeding rate (bites/min)", 
       # x = expression(atop(Scarid~biomass, "(kg/100m"^2*")")),
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        plot.margin = margin(1, 1, 0, 1, unit = "pt")
        )

fi_s2 <- ggplot(filter(sum_ssp, species_code == "stop" & phase == "i"), aes(x = consp_den, y = for_bites_mean)) +
  geom_smooth(method="loess", span=1) +
  geom_point(shape=18) +
  ylim(-10, 25) +
  labs(y = "", 
       x = "",
       # x = expression(atop(Conspec.~density, "(indv./100m"^2*")")),
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        plot.margin = margin(1, 1, 0, 1, unit = "pt")
        )

fi_s3 <- ggplot(filter(sum_ssp, species_code == "stop" & phase == "i"), aes(x = rugosity, y = for_bites_mean)) +
  geom_smooth(method="loess", span=1) +
  geom_point(shape=18) +
  ylim(-10, 25) +
  labs(y = "", 
       x = "",
       #x = expression("Rugosity"), 
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        plot.margin = margin(1, 1, 0, 1, unit = "pt")
        )

fi_s4 <- ggplot(filter(sum_ssp, species_code == "stop" & phase == "i"), aes(x = ta_canopy, y = for_bites_mean)) +
  geom_smooth(method="loess", span=1) +
  geom_point(shape=18) +
  ylim(-10, 25) +
  labs(y = "", 
       x = "",
       # x = expression("EAM canopy \nheight (mm)"), 
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        plot.margin = margin(1, 1, 0, 1, unit = "pt")
        )

fi_s5 <- ggplot(filter(sum_ssp, species_code == "stop" & phase == "i"), aes(x = ta_cover, y = for_bites_mean)) +
  geom_smooth(method="loess", span=1) +
  geom_point(shape=18) +
  ylim(-10, 25) +
  labs(y = "", 
       x = "",
       # x = expression("EAM percent \ncover"), 
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        plot.margin = margin(1, 1, 0, 1, unit = "pt")
        )

fi_s7 <- ggplot(filter(sum_ssp, species_code == "stop" & phase == "i"), aes(x = ma_cover, y = for_bites_mean)) +
  geom_smooth(method="loess", span=1) +
  geom_point(shape=18) +
  ylim(-10, 25) +
  labs(y = "", 
       x = "",
       # x = expression("Macroalgal \npercent cover"), 
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        plot.margin = margin(1, 1, 0, 1, unit = "pt")
        )

grid.arrange(left = "Feeding intensity (bites / foray)",
  arrangeGrob(fi_q1, fi_q3, fi_q4, fi_q5, fi_q7, ncol=5, right = "S. vetula"),
  arrangeGrob(
    arrangeGrob(fi_s1, bottom = "Scarid biomass"), 
    arrangeGrob(fi_s3, bottom = "Rugosity"),
    arrangeGrob(fi_s4, bottom = "Turf height"),
    arrangeGrob(fi_s5, bottom = "Turf % cover"),
    arrangeGrob(fi_s7, bottom = "Macro. % cover"),
    ncol=5, right = "S. viride"), 
  nrow=2) 
```

## PCA
```{r, echo=F, message=F, warning=F}
pca_pred <- prcomp(na.omit(sum_site %>% 
              select(rugosity, ma_cover, ma_canopy, ta_cover, ta_canopy, scar_den, 
                     scar_bm, carn30_bm)),
  center = T, scale = T)

autoplot(pca_pred, data = sum_site, colour = 'island',
         # frame = T, frame.type = "norm",
         loadings = T, loadings.colour = 'gray',
         loadings.label = T, loadings.label.size = 3, loadings.label.colour = 'black',
            loadings.label.repel = T) +
    scale_color_manual(name="island", values=c("lightblue", "deepskyblue3", "deepskyblue4")) +
    theme(legend.direction = 'horizontal', legend.position = 'top') +
    theme_minimal()
ggsave(here("figs/pca_pred.png"))

sum_site <- sum_site %>% 
  bind_cols(as.data.frame(pca_pred$x) %>% 
           select(PC1,PC2)) %>%
  rename(pc1_all=PC1,
         pc2_all=PC2)

sum_id_pca <- sum_id %>% 
  left_join(select(sum_site,site,pc1_all,pc2_all), by = "site")
sum_ssp_pca <- sum_ssp %>%
  left_join(select(sum_site,site,pc1_all,pc2_all), by = "site")
```
```{r, eval=F, echo=F, message=F, warning=F}
ggplot(filter(sum_ssp_pca, species_code != "rbp" & phase == "i"), aes(x = pc1_all, y = fr_mean/60)) +
  facet_wrap(.~species) +
  geom_smooth(method="loess", span=1) +
  geom_point(shape=18) +
  labs(y = "Feeding rate (bites/min)", 
       x = "PC1") +
  theme(strip.text = element_text(face = "italic")) +
  theme(plot.title = element_text(hjust = 0.5)) + 
  theme_bw()
```

```{r, eval=F, echo=F, message=F, warning=F}
ggplot(filter(sum_ssp_pca, species_code != "rbp" & phase == "i"), aes(x = pc1_all, y = for_bites_mean)) +
  facet_wrap(.~species) +
  geom_smooth(method="loess", span=1) +
  geom_point(shape=18) +
  labs(y = "Feeding intensity (bites/foray)", 
       x = "PC1") +
  theme(strip.text = element_text(face = "italic")) +
  theme(plot.title = element_text(hjust = 0.5)) + 
  theme_bw()
```

## GAMM results

### Predicting feeding rate:

```{r, echo=F, message=F, warning=F}
m_data <- sum_ssp_pca %>% filter(species_code!="rbp")

gamm1 <- gamm(fr_mean ~ species_code + phase + s(length_m, k=4) + s(pc1_all, k=4) + s(pc2_all, k=4), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
g1 <- c("g1", "Species + Phase + Size + PC1 + PC2", AICc(gamm1$lme), summary(gamm1$gam)$r.sq)

gamm2 <- gamm(fr_mean ~ species_code + phase + s(length_m, k=4) + s(pc1_all, k=4), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
g2 <- c("g2", "Species + Phase + Size + PC1", AICc(gamm2$lme), summary(gamm2$gam)$r.sq)

gamm3 <- gamm(fr_mean ~ species_code + phase + s(length_m, by = factor(species_code), k=4) + s(pc1_all, by = factor(species_code), k=4) + s(pc2_all, by = factor(species_code), k=4), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
g3 <- c("g3", "Species + Phase + Size~Species + PC1~Species + PC2~Species", AICc(gamm3$lme), summary(gamm3$gam)$r.sq)

gamm4 <- gamm(fr_mean ~ species_code + phase + s(length_m, by = factor(species_code), k=4) + s(pc1_all, by = factor(species_code), k=4) + s(pc2_all, k=4), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
g4 <- c("g4", "Species + Phase + Size~Species + PC1~Species + PC2", AICc(gamm4$lme), summary(gamm4$gam)$r.sq)

gamm5 <- gamm(fr_mean ~ species + phase + s(length_m, by = factor(species_code), k=4) + s(pc1_all, by = factor(species_code), k=4), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
g5 <- c("g5", "Species + Phase + Size~Species + PC1~Species", AICc(gamm5$lme), summary(gamm5$gam)$r.sq)

gamm6 <- gamm(fr_mean ~ phase + s(length_m, k=4) + s(pc1_all, by = factor(species_code), k=4), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
g6 <- c("g6", "Phase + Size + PC1~Species", AICc(gamm6$lme), summary(gamm6$gam)$r.sq)

gamm7 <- gamm(fr_mean ~ phase + s(length_m, by = factor(species_code), k=4) + s(pc1_all, k=4), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
g7 <- c("g7", "Phase + Size~Species + PC1", AICc(gamm7$lme), summary(gamm7$gam)$r.sq)

gamm8 <- gamm(fr_mean ~ species + phase + s(length_m, by = factor(species_code), k=4) + s(pc1_all, k=4), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
g8 <- c("g8", "Species + Phase + Size~Species + PC1", AICc(gamm8$lme), summary(gamm8$gam)$r.sq)

gamm9 <- gamm(fr_mean ~ species + phase + s(length_m, k=4) + s(pc1_all, by = factor(species_code), k=4), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
g9 <- c("g9", "Species + Phase + Size + PC1~Species", AICc(gamm9$lme), summary(gamm9$gam)$r.sq)

gamm10 <- gamm(fr_mean ~ species + phase + s(pc1_all, by = factor(species_code), k=4), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
g10 <- c("g10", "Species + Phase + PC1~Species", AICc(gamm10$lme), summary(gamm10$gam)$r.sq)

gamm11 <- gamm(fr_mean ~ phase + s(length_m, by = factor(species_code), k=4) + s(pc1_all, by = factor(species_code), k=4), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
g11 <- c("g11", "Phase + Size~Species + PC1~Species", AICc(gamm11$lme), summary(gamm11$gam)$r.sq)

gtab_fr <- as.data.frame(rbind(g1,g2,g3,g4,g5,g6,g7,g8,g9,g10,g11)) %>% setNames(c("g","Model", "AICc", "Adj. R2")) %>% arrange(AICc)
```

```{r, echo=F, message=F}
kable(gtab_fr[1:4, ])
plot(gamm3$gam,pages=1)
```

### Predicting feeding intensity:

```{r, echo=F, message=F, warning=F}
m_data <- sum_ssp_pca %>% filter(species_code!="rbp")

fi_gamm1 <- gamm(for_bites_mean ~ species_code + phase + s(length_m, k=4) + s(pc1_all, k=4) + s(pc2_all, k=4), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
# summary(fi_gamm1$gam)
# plot(fi_gamm1$gam,pages=1)
g1 <- c("g1", "Species + Phase + Size + PC1 + PC2", AICc(fi_gamm1$lme), summary(fi_gamm1$gam)$r.sq)

fi_gamm2 <- gamm(for_bites_mean ~ species_code + phase + s(length_m, k=4) + s(pc1_all, k=4), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
g2 <- c("g2", "Species + Phase + Size + PC1", AICc(fi_gamm2$lme), summary(fi_gamm2$gam)$r.sq)

fi_gamm3 <- gamm(for_bites_mean ~ species_code + phase + s(length_m, by = factor(species_code), k=4) + s(pc1_all, by = factor(species_code), k=4) + s(pc2_all, by = factor(species_code), k=4), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
g3 <- c("g3", "Species + Phase + Size~Species + PC1~Species", AICc(fi_gamm3$lme), summary(fi_gamm3$gam)$r.sq)

fi_gamm4 <- gamm(for_bites_mean ~ species_code + phase + s(length_m, by = factor(species_code), k=4) + s(pc1_all, by = factor(species_code), k=4) + s(pc2_all, k=4), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
g4 <- c("g4", "Species + Phase + Size~Species + PC1~Species + PC2", AICc(fi_gamm4$lme), summary(fi_gamm4$gam)$r.sq)

fi_gamm5 <- gamm(for_bites_mean ~ phase + s(length_m, by = factor(species_code), k=4) + s(pc1_all, by = factor(species_code), k=4), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
g5 <- c("g5", "Phase + Size~Species + PC1~Species", AICc(fi_gamm5$lme), summary(fi_gamm5$gam)$r.sq)

fi_gamm6 <- gamm(for_bites_mean ~ species + phase + s(length_m, k=4) + s(pc1_all, by = factor(species_code), k=4), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
g6 <- c("g6", "Species + Phase + Size + PC1~Species", AICc(fi_gamm6$lme), summary(fi_gamm6$gam)$r.sq)

fi_gamm7 <- gamm(for_bites_mean ~ species + phase + s(length_m, by = factor(species_code), k=4) + s(pc1_all, k=4), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
g7 <- c("g7", "Species + Phase + Size~Species + PC1", AICc(fi_gamm7$lme), summary(fi_gamm7$gam)$r.sq)

fi_gamm8 <- gamm(for_bites_mean ~ phase + s(length_m, k=4) + s(pc1_all, by = factor(species_code), k=4), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
g8 <- c("g8", "Phase + Size + PC1~Species", AICc(fi_gamm8$lme), summary(fi_gamm8$gam)$r.sq)

fi_gamm9 <- gamm(for_bites_mean ~ phase + s(length_m, by = factor(species_code), k=4) + s(pc1_all, k=4), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
g9 <- c("g9", "Phase + Size~Species + PC1", AICc(fi_gamm9$lme), summary(fi_gamm9$gam)$r.sq)

gtab_fi <- as.data.frame(rbind(g1,g2,g3,g4,g5,g6,g7,g8,g9)) %>% setNames(c("g", "Model", "AICc", "Adj. R2")) %>% arrange(AICc)

```

```{r, echo=F, message=F}
kable(gtab_fi[1:3, ])
plot(gamm6$gam,pages=1)
```

## Bite content hypothesis

Using algal canopy height to estimate the volume of a given bite taken in different sites. This makes estimated grazing impact a lot more consistent across sites than would be inferred from just looking at feeding rates (first bargraph)

```{r, echo=F, message=F, warning=F}
# grazing impact (vg ~ volume grazed) of fish from behavioral surveys
sum_id_vg <- sum_id %>%
  filter(species_code != "rbp") %>%
  mutate(
    bs = ifelse(species_code == "stop", 5.257*10^-4*length_cm^2, 4.013*10^-4*length_cm^2), # adding bite size equations from Mumby
    bh_capped = ifelse(ta_canopy < 2*sqrt(bs)/(22/7), ta_canopy, 2*sqrt(bs)/(22/7)), # if ta_canopy exceeds diameter of bite size, bite height is capped at bite size
    bv = bs*ta_canopy,
    bv_capped = bs*bh_capped,
    vg = fr*bv,
    vg_capped = fr*bv_capped,
    island = factor(island, levels = c("Bonaire","Antigua","Barbuda"))
) %>%
  arrange(island)

# summing by site, species, and phase
sum_ssp_vg <- sum_id_vg %>% 
  group_by(site, island, species, species_code, phase) %>% 
  summarize(
    n=n(),
    vg_mean=mean(vg),
    vg_se=sd(vg)/sqrt(n),
    vg_capped_mean=mean(vg_capped),
    vg_capped_se=sd(vg_capped)/sqrt(n)
    ) %>% 
  select(-n)

ggplot(filter(sum_ssp_vg, species_code != "rbp" & phase == "i"), aes(x=factor(site, levels = sum_site$site), y=vg_mean)) +
    geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
    geom_errorbar(aes(ymin=vg_mean-vg_se, ymax=vg_mean+vg_se), width=.2,
                 position=position_dodge(.9)) +
    facet_grid(~species) +
    scale_fill_brewer(palette = "Blues") +
    labs(title="Mean grazing impact by volume (initial phase)", 
         y="Grazing impact (cm3/hr)", 
         x="Site") +
    theme_bw() +
    theme(strip.text = element_text(face = "italic"), 
          axis.text.x = element_text(angle = 90, hjust = 1), 
          plot.title = element_text(hjust = 0.5))
```

```{r, eval=F, echo=F, message=F, warning=F}
ggplot(filter(sum_ssp_vg, species_code != "rbp" & phase == "i"), aes(x=factor(site, levels = sum_site$site), y=vg_capped_mean)) +
    geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
    geom_errorbar(aes(ymin=vg_capped_mean-vg_capped_se, ymax=vg_capped_mean+vg_capped_se), width=.2,
                 position=position_dodge(.9)) +
    facet_grid(~species) +
    scale_fill_brewer(palette = "Blues") +
    labs(title="Mean grazing impact by volume (capped height, initial phase)", 
         y="Grazing impact (cm3/hr)", 
         x="Site") +
    theme_bw() +
    theme(strip.text = element_text(face = "italic"), 
          axis.text.x = element_text(angle = 90, hjust = 1), 
          plot.title = element_text(hjust = 0.5))
```

# Discussion

* Significant differences in feeding behaviors within species across sites (be conservative about conclusions here, just presenting observed trends)
* GAMM results: explain meaning/significance of PC1 (essentially higher feeding rates in reefs with higher scarid biomass and rugosity vs. more algal dominated and lower scarid biomass reefs); *S. vetula* seem more impacted by PC1 than *S. viride*. 
* Potential mechanisms/hypotheses for these observed trends:
    * Bite content/nutrition hypothesis: scarids in higher canopy sites are consuming more per bite, thus taking fewer bites - implications for spatial grazing impact
    * Social feeding hypothesis: more time must be spent vigilant in smaller groups, potential compounding effect with overfishing/loss of herbivory
    * Predation risk hypothesis: presence of large piscivores does not explain feeding trends, but risk from spearfishers may play a bigger role (able to target larger fish, i.e. those sampled here). (Incorporate rugosity here - Rogers, Madin, Catano). Did not find strong evidence for this here, but carnivore biomass was strongly correllated with herbivore biomass and rugosity.
* Ecological implications: potential reinforcing feedback loops in degraded reefs if high algal/low scarid reefs also have less grazing impact per scarid
    * Feeding intensity: less intense feeding/more distributed bites in degraded reefs may also reduce actual substrate made available to recruiting corals (~ grazing concentration lit. here)
* Caveats + limitations: all these potential drivers/mechanisms are understandably correlated in observational work
    * Room for experimental work to unpack/distinguish drivers
* Need for incorporation of behavior in coral reef grazing models, especially those used to set targets for minimum herbivore biomass levels, because these may look very different in different reef conditions. Current paradigm of constant species-level grazing behaviors may be inaccurate. In Caribbean case, most models use Bonaire data which may not be representative of dynamics on more degraded Caribbean reefs.
* Need for framework to help incorporate behavioral effects into ecosystem management
* Conservation + mgmt implications: multiple human impacts on reefs may affect herbivore behavior in different ways, spec. fishing + habitat degradation - which may have reinforcing/compounding negative effects on reef health

### 2021-01-16 r/old_code/Ch2 outline_V3.Rmd
clearing out old files

--- 2019-10-01 r/Ch2 outline_V3.Rmd
@@
            length_cm <= 30 & length_cm >= 15 &
+           ### excluding sites with less than 7 observations within size subset range
            !(species_code == "stop" & site == "Pallaster West") & 
@@
 
-sum_ssp_sub <- sum_id %>% 
+sum_ssp_sub <- sum_id %>% # size subset 15-30cm
   group_by(species, species_code, site, island) %>%
@@
 
+## PCA
+
+```{r, eval=F, echo=F, message=F, warning=F}
+### old code for other plot version
+# pca_pred <- prcomp(na.omit(sum_site %>%
+#                 remove_rownames %>% column_to_rownames(var="site") %>%
+#                 select(rugosity, ma_cover, ma_canopy, ta_cover, ta_canopy, scar_den, scar_bm, carn30_bm, lc_cover)),
+#   center = T, scale = T)
+# 
+# autoplot(pca_pred, data = sum_site, shape = 'island', colour = 'island',
+#          label = F,
+#          loadings = T, loadings.colour = 'gray',
+#          loadings.label = T, loadings.label.size = 3, loadings.label.colour = 'black',
+#             loadings.label.repel = T) +
+#     scale_color_manual(values = c("cadetblue3","cadetblue3","cadetblue3")) +
+#     theme_bw() +
+#     theme(legend.title = element_blank())
+# ggsave(here("figs/pca_pred.png"), width = 8, height = 5)
+```
+```{r, echo=F, message=F, warning=F}
+pca_data <- na.omit(sum_site %>%
+  select(rugosity, ma_cover, ma_canopy, ta_cover, ta_canopy, lc_cover, scar_bm, scar_den, carn30_bm))
+names(pca_data)[1:9] =
+  c("rugosity", "macroalgal cover", "macroalgal canopy", "turf cover", "turf canopy", "coral cover", "scarid biomass", "scarid density", "predator biomass")
+
+pca_pred <- prcomp(pca_data, center = T, scale = T)
+
+fviz_pca_biplot(pca_pred,
+                col.ind = sum_site$island,
+                mean.point = F,
+                label = "var",
+                repel = T
+                ) +
+  scale_shape_manual(values=c(15,16,17)) +
+  scale_color_manual(values = c("black","black","black")) +
+  labs(x = "PC1 (55.6%)", y = "PC2 (24.5%)", title = "") +
+  theme_bw() +
+  theme(legend.title = element_blank())
+
+ggsave(here("figs/pca_pred.png"), width = 7, height = 5)
+```
+
+```{r}
+sum_site_pca <- sum_site %>% 
+  bind_cols(as.data.frame(pca_pred$x) %>% 
+           select(PC1,PC2)) %>%
+  rename(pc1_all=PC1,
+         pc2_all=PC2)
+
+sum_id_pca <- sum_id %>% 
+  left_join(select(sum_site_pca,site,pc1_all,pc2_all), by = "site")
+sum_ssp_pca <- sum_ssp_sub %>%
+  left_join(select(sum_site_pca,site,pc1_all,pc2_all), by = "site")
+```
+
+## GAMM results (see 5.8.19)
+
+Note that these GAMMs are using the size subsetted data, as opposed to the full size ranges
+
+Predicting feeding rate (site-level medians, island as random effect)
+```{r, echo=F, message=F, warning=F}
+m_data <- sum_ssp_pca
+
+gamm1 <- gamm(fr_median ~ species_code + s(length_mean, k=3) + s(pc1_all, k=3) + s(pc2_all, k=3), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           data = m_data
+           )
+g1 <- c("g1", "Species + Size + PC1 + PC2", AICc(gamm1$lme), summary(gamm1$gam)$r.sq)
+
+gamm2 <- gamm(fr_median ~ species_code + s(length_mean, k=3) + s(pc1_all, k=3), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           data = m_data
+           )
+g2 <- c("g2", "Species + Size + PC1", AICc(gamm2$lme), summary(gamm2$gam)$r.sq)
+
+gamm3 <- gamm(fr_median ~ species_code + s(length_mean, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3) + s(pc2_all, by = factor(species_code), k=3),
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           data = m_data
+           )
+g3 <- c("g3", "Species + Size~Species + PC1~Species + PC2~Species", AICc(gamm3$lme), summary(gamm3$gam)$r.sq)
+
+gamm4 <- gamm(fr_median ~ species_code + s(length_mean, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3) + s(pc2_all, k=3),
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           data = m_data
+           )
+g4 <- c("g4", "Species + Size~Species + PC1~Species + PC2", AICc(gamm4$lme), summary(gamm4$gam)$r.sq)
+
+gamm5 <- gamm(fr_median ~ species + s(length_mean, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3),
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           data = m_data
+           )
+g5 <- c("g5", "Species + Size~Species + PC1~Species", AICc(gamm5$lme), summary(gamm5$gam)$r.sq)
+
+gamm6 <- gamm(fr_median ~ s(length_mean, k=3) + s(pc1_all, by = factor(species_code), k=3), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           data = m_data
+           )
+g6 <- c("g6", "Size + PC1~Species", AICc(gamm6$lme), summary(gamm6$gam)$r.sq)
+
+gamm7 <- gamm(fr_median ~ s(length_mean, by = factor(species_code), k=3) + s(pc1_all, k=3), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           data = m_data
+           )
+g7 <- c("g7", "Size~Species + PC1", AICc(gamm7$lme), summary(gamm7$gam)$r.sq)
+
+gamm8 <- gamm(fr_median ~ species + s(length_mean, by = factor(species_code), k=3) + s(pc1_all, k=3), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           data = m_data
+           )
+g8 <- c("g8", "Species + Size~Species + PC1", AICc(gamm8$lme), summary(gamm8$gam)$r.sq)
+
+gamm9 <- gamm(fr_median ~ species + s(length_mean, k=3) + s(pc1_all, by = factor(species_code), k=3), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           data = m_data
+           )
+g9 <- c("g9", "Species + Size + PC1~Species", AICc(gamm9$lme), summary(gamm9$gam)$r.sq)
+
+gamm10 <- gamm(fr_median ~ species + s(pc1_all, by = factor(species_code), k=3), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           data = m_data
+           )
+g10 <- c("g10", "Species + PC1~Species", AICc(gamm10$lme), summary(gamm10$gam)$r.sq)
+
+gamm11 <- gamm(fr_median ~ s(length_mean, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           data = m_data
+           )
+g11 <- c("g11", "Size~Species + PC1~Species", AICc(gamm11$lme), summary(gamm11$gam)$r.sq)
+
+gamm12 <- gamm(fr_median ~ species_code + s(pc1_all, k=3), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           data = m_data
+           )
+g12 <- c("g12", "Species + PC1", AICc(gamm12$lme), summary(gamm12$gam)$r.sq)
+
+gtab_fr_site <- as.data.frame(rbind(g1,g2,g3,g4,g5,g6,g7,g8,g9,g10,g11,g12)) %>% setNames(c("g","Model", "AICc", "Adj. R2")) %>% arrange(AICc)
+kable(gtab_fr_site)
+
+# summary + plots from best models
+summary(gamm10$gam)
+plot(gamm10$gam,pages=1)
+```
+
+Predicting Grazing intensity (site-level medians, island as random effect)
+```{r, echo=F, message=F, warning=F}
+m_data <- sum_ssp_pca
+
+gamm1 <- gamm(for_bites_median ~ species_code + s(length_mean, k=3) + s(pc1_all, k=3) + s(pc2_all, k=3), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           data = m_data
+           )
+g1 <- c("g1", "Species + Size + PC1 + PC2", AICc(gamm1$lme), summary(gamm1$gam)$r.sq)
+
+gamm2 <- gamm(for_bites_median ~ species_code + s(length_mean, k=3) + s(pc1_all, k=3), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           data = m_data
+           )
+g2 <- c("g2", "Species + Size + PC1", AICc(gamm2$lme), summary(gamm2$gam)$r.sq)
+
+gamm3 <- gamm(for_bites_median ~ species_code + s(length_mean, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3) + s(pc2_all, by = factor(species_code), k=3),
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           data = m_data
+           )
+g3 <- c("g3", "Species + Size~Species + PC1~Species + PC2~Species", AICc(gamm3$lme), summary(gamm3$gam)$r.sq)
+
+# gamm4 <- gamm(for_bites_median ~ species_code + s(length_mean, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3) + s(pc2_all, k=3),
+#            family = gaussian(link=identity),
+#            random = list(island=~1),
+#            data = m_data
+#            )
+# g4 <- c("g4", "Species + Size~Species + PC1~Species + PC2", AICc(gamm4$lme), summary(gamm4$gam)$r.sq)
+
+gamm5 <- gamm(for_bites_median ~ species + s(length_mean, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           data = m_data
+           )
+g5 <- c("g5", "Species + Size~Species + PC1~Species", AICc(gamm5$lme), summary(gamm5$gam)$r.sq)
+
+gamm6 <- gamm(for_bites_median ~ s(length_mean, k=3) + s(pc1_all, by = factor(species_code), k=3), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           data = m_data
+           )
+g6 <- c("g6", "Size + PC1~Species", AICc(gamm6$lme), summary(gamm6$gam)$r.sq)
+
+gamm7 <- gamm(for_bites_median ~ s(length_mean, by = factor(species_code), k=3) + s(pc1_all, k=3), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           data = m_data
+           )
+g7 <- c("g7", "Size~Species + PC1", AICc(gamm7$lme), summary(gamm7$gam)$r.sq)
+
+gamm8 <- gamm(for_bites_median ~ species + s(length_mean, by = factor(species_code), k=3) + s(pc1_all, k=3), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           data = m_data
+           )
+g8 <- c("g8", "Species + Size~Species + PC1", AICc(gamm8$lme), summary(gamm8$gam)$r.sq)
+
+gamm9 <- gamm(for_bites_median ~ species + s(length_mean, k=3) + s(pc1_all, by = factor(species_code), k=3), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           data = m_data
+           )
+g9 <- c("g9", "Species + Size + PC1~Species", AICc(gamm9$lme), summary(gamm9$gam)$r.sq)
+
+gamm10 <- gamm(for_bites_median ~ species + s(pc1_all, by = factor(species_code), k=3), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           data = m_data
+           )
+g10 <- c("g10", "Species + PC1~Species", AICc(gamm10$lme), summary(gamm10$gam)$r.sq)
+
+gamm11 <- gamm(for_bites_median ~ s(length_mean, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           data = m_data
+           )
+g11 <- c("g11", "Size~Species + PC1~Species", AICc(gamm11$lme), summary(gamm11$gam)$r.sq)
+
+gamm12 <- gamm(for_bites_median ~ species_code + s(pc1_all, k=3), 
+           family = gaussian(link=identity),
+           random = list(island=~1),
+           data = m_data
+           )
+g12 <- c("g12", "Species + PC1", AICc(gamm12$lme), summary(gamm12$gam)$r.sq)
+
+gtab_fi_site <- as.data.frame(rbind(g1,g2,g3,g5,g6,g7,g8,g9,g10,g11,g12)) %>% setNames(c("g","Model", "AICc", "Adj. R2")) %>% arrange(AICc)
+
+# plots from best models
+plot(gamm12$gam,pages=1)
+plot(gamm10$gam,pages=1)
+
+summary(gamm12$gam)
+```
+
+# Supplement/optional
+
+### Predator gape size
+
+From Mumby et al.:
+Grouper *E. striatus*, gape width (G): G = -0.5998 + (0.167 × TL)
+```{r}
+g30_mum <- -0.5998 + (0.167 * 30) # 4
+g40_mum <- -0.5998 + (0.167 * 40) # 6
+```
+
+From Goatley & Bellwood:
+Horizontal gape: G = -0.002 * SL + 6.421
+```{r}
+g30_goat <- -0.002 * 30 + 6.421
+g40_goat <- -0.002 * 40 + 6.421
+```
+
+Dorsoventral height (Rakitin & Kramer 1996):
+S. viride: 15cm FL > 4.5cm body depth
+S. vetula: 15cm FL > 4.5cm body depth
+
+### Loess plots for additional drivers?
+
+### Grazing impact evaluation
 ```{r, eval = F, echo = F, message=F, warning=F}
@@
 ```
-
-## PCA
-```{r, echo=F, message=F, warning=F}
-pca_pred <- prcomp(na.omit(site %>%
-                remove_rownames %>% column_to_rownames(var="site") %>%
-                select(rugosity, ma_cover, ma_canopy, ta_cover, ta_canopy, scar_den, scar_bm, carn30_bm, lc_cover)),
-  center = T, scale = T)
-
-autoplot(pca_pred, data = site, shape = 'island', colour = 'island',
-         label = T,
-         loadings = T, loadings.colour = 'gray',
-         loadings.label = T, loadings.label.size = 3, loadings.label.colour = 'black',
-            loadings.label.repel = T) +
-    scale_color_manual(values = c("cadetblue3","cadetblue3","cadetblue3")) +
-    theme_bw() +
-    theme(legend.title = element_blank())
-ggsave(here("figs/pca_pred.png"), width = 8, height = 5)
-
-sum_id_pca <- sum_id %>% 
-  left_join(select(sum_site,site,pc1_all,pc2_all), by = "site")
-sum_ssp_pca <- sum_ssp_sub %>%
-  left_join(select(sum_site,site,pc1_all,pc2_all), by = "site")
-```
-
-### Better PCA?
-```{r, echo=F, message=F, warning=F}
-pca_data <- na.omit(site %>%
-  select(rugosity, ma_cover, ma_canopy, ta_cover, ta_canopy, lc_cover, scar_bm, scar_den, carn30_bm))
-names(pca_data)[1:9] =
-  c("rugosity", "macroalgal cover", "macroalgal canopy", "turf cover", "turf canopy", "coral cover", "scarid biomass", "scarid density", "predator biomass")
-
-pca_pred <- prcomp(pca_data, center = T, scale = T)
-
-fviz_pca_biplot(pca_pred,
-                col.ind = site$island,
-                mean.point = F,
-                label = "var",
-                repel = T
-                ) +
-  scale_shape_manual(values=c(15,16,17)) +
-  scale_color_manual(values = c("black","black","black")) +
-  labs(x = "PC1 (55.6%)", y = "PC2 (24.5%)", title = "") +
-  theme_bw() +
-  theme(legend.title = element_blank())
-
-ggsave(here("figs/pca_pred.png"), width = 7, height = 5)
-
-```
-
-## GAMM results (see 5.8.19)
-
-Predicting feeding rate (site-level medians, island as random effect)
-```{r, echo=F, message=F, warning=F}
-m_data <- sum_ssp_pca
-
-gamm1 <- gamm(fr_median ~ species_code + s(length_mean, k=3) + s(pc1_all, k=3) + s(pc2_all, k=3), 
-           family = gaussian(link=identity),
-           random = list(island=~1),
-           data = m_data
-           )
-g1 <- c("g1", "Species + Size + PC1 + PC2", AICc(gamm1$lme), summary(gamm1$gam)$r.sq)
-
-gamm2 <- gamm(fr_median ~ species_code + s(length_mean, k=3) + s(pc1_all, k=3), 
-           family = gaussian(link=identity),
-           random = list(island=~1),
-           data = m_data
-           )
-g2 <- c("g2", "Species + Size + PC1", AICc(gamm2$lme), summary(gamm2$gam)$r.sq)
-
-gamm3 <- gamm(fr_median ~ species_code + s(length_mean, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3) + s(pc2_all, by = factor(species_code), k=3),
-           family = gaussian(link=identity),
-           random = list(island=~1),
-           data = m_data
-           )
-g3 <- c("g3", "Species + Size~Species + PC1~Species + PC2~Species", AICc(gamm3$lme), summary(gamm3$gam)$r.sq)
-
-# gamm4 <- gamm(fr_median ~ species_code + s(length_mean, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3) + s(pc2_all, k=3), 
-#            family = gaussian(link=identity),
-#            random = list(island=~1),
-#            data = m_data
-#            )
-# g4 <- c("g4", "Species + Size~Species + PC1~Species + PC2", AICc(gamm4$lme), summary(gamm4$gam)$r.sq)
-
-# gamm5 <- gamm(fr_median ~ species + s(length_mean, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3), 
-#            family = gaussian(link=identity),
-#            random = list(island=~1),
-#            data = m_data
-#            )
-# g5 <- c("g5", "Species + Size~Species + PC1~Species", AICc(gamm5$lme), summary(gamm5$gam)$r.sq)
-
-gamm6 <- gamm(fr_median ~ s(length_mean, k=3) + s(pc1_all, by = factor(species_code), k=3), 
-           family = gaussian(link=identity),
-           random = list(island=~1),
-           data = m_data
-           )
-g6 <- c("g6", "Size + PC1~Species", AICc(gamm6$lme), summary(gamm6$gam)$r.sq)
-
-gamm7 <- gamm(fr_median ~ s(length_mean, by = factor(species_code), k=3) + s(pc1_all, k=3), 
-           family = gaussian(link=identity),
-           random = list(island=~1),
-           data = m_data
-           )
-g7 <- c("g7", "Size~Species + PC1", AICc(gamm7$lme), summary(gamm7$gam)$r.sq)
-
-gamm8 <- gamm(fr_median ~ species + s(length_mean, by = factor(species_code), k=3) + s(pc1_all, k=3), 
-           family = gaussian(link=identity),
-           random = list(island=~1),
-           data = m_data
-           )
-g8 <- c("g8", "Species + Size~Species + PC1", AICc(gamm8$lme), summary(gamm8$gam)$r.sq)
-
-gamm9 <- gamm(fr_median ~ species + s(length_mean, k=3) + s(pc1_all, by = factor(species_code), k=3), 
-           family = gaussian(link=identity),
-           random = list(island=~1),
-           data = m_data
-           )
-g9 <- c("g9", "Species + Size + PC1~Species", AICc(gamm9$lme), summary(gamm9$gam)$r.sq)
-
-gamm10 <- gamm(fr_median ~ species + s(pc1_all, by = factor(species_code), k=3), 
-           family = gaussian(link=identity),
-           random = list(island=~1),
-           data = m_data
-           )
-g10 <- c("g10", "Species + PC1~Species", AICc(gamm10$lme), summary(gamm10$gam)$r.sq)
-
-gamm11 <- gamm(fr_median ~ s(length_mean, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3), 
-           family = gaussian(link=identity),
-           random = list(island=~1),
-           data = m_data
-           )
-g11 <- c("g11", "Size~Species + PC1~Species", AICc(gamm11$lme), summary(gamm11$gam)$r.sq)
-
-gamm12 <- gamm(fr_median ~ species_code + s(pc1_all, k=3), 
-           family = gaussian(link=identity),
-           random = list(island=~1),
-           data = m_data
-           )
-g12 <- c("g12", "Species + PC1", AICc(gamm12$lme), summary(gamm12$gam)$r.sq)
-
-gtab_fr_site <- as.data.frame(rbind(g1,g2,g4,g5,g6,g7,g8,g9,g10,g11,g12)) %>% setNames(c("g","Model", "AICc", "Adj. R2")) %>% arrange(AICc)
-
-# summary + plots from best models
-summary(gamm12$gam)
-plot(gamm12$gam,pages=1)
-```
-
-Predicting Grazing intensity (site-level medians, island as random effect)
-```{r, echo=F, message=F, warning=F}
-m_data <- sum_ssp_pca
-
-gamm1 <- gamm(for_bites_median ~ species_code + s(length_mean, k=3) + s(pc1_all, k=3) + s(pc2_all, k=3), 
-           family = gaussian(link=identity),
-           random = list(island=~1),
-           data = m_data
-           )
-g1 <- c("g1", "Species + Size + PC1 + PC2", AICc(gamm1$lme), summary(gamm1$gam)$r.sq)
-
-gamm2 <- gamm(for_bites_median ~ species_code + s(length_mean, k=3) + s(pc1_all, k=3), 
-           family = gaussian(link=identity),
-           random = list(island=~1),
-           data = m_data
-           )
-g2 <- c("g2", "Species + Size + PC1", AICc(gamm2$lme), summary(gamm2$gam)$r.sq)
-
-gamm3 <- gamm(for_bites_median ~ species_code + s(length_mean, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3) + s(pc2_all, by = factor(species_code), k=3),
-           family = gaussian(link=identity),
-           random = list(island=~1),
-           data = m_data
-           )
-g3 <- c("g3", "Species + Size~Species + PC1~Species + PC2~Species", AICc(gamm3$lme), summary(gamm3$gam)$r.sq)
-
-gamm4 <- gamm(for_bites_median ~ species_code + s(length_mean, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3) + s(pc2_all, k=3), 
-           family = gaussian(link=identity),
-           random = list(island=~1),
-           data = m_data
-           )
-g4 <- c("g4", "Species + Size~Species + PC1~Species + PC2", AICc(gamm4$lme), summary(gamm4$gam)$r.sq)
-
-gamm5 <- gamm(for_bites_median ~ species + s(length_mean, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3), 
-           family = gaussian(link=identity),
-           random = list(island=~1),
-           data = m_data
-           )
-g5 <- c("g5", "Species + Size~Species + PC1~Species", AICc(gamm5$lme), summary(gamm5$gam)$r.sq)
-
-gamm6 <- gamm(for_bites_median ~ s(length_mean, k=3) + s(pc1_all, by = factor(species_code), k=3), 
-           family = gaussian(link=identity),
-           random = list(island=~1),
-           data = m_data
-           )
-g6 <- c("g6", "Size + PC1~Species", AICc(gamm6$lme), summary(gamm6$gam)$r.sq)
-
-gamm7 <- gamm(for_bites_median ~ s(length_mean, by = factor(species_code), k=3) + s(pc1_all, k=3), 
-           family = gaussian(link=identity),
-           random = list(island=~1),
-           data = m_data
-           )
-g7 <- c("g7", "Size~Species + PC1", AICc(gamm7$lme), summary(gamm7$gam)$r.sq)
-
-gamm8 <- gamm(for_bites_median ~ species + s(length_mean, by = factor(species_code), k=3) + s(pc1_all, k=3), 
-           family = gaussian(link=identity),
-           random = list(island=~1),
-           data = m_data
-           )
-g8 <- c("g8", "Species + Size~Species + PC1", AICc(gamm8$lme), summary(gamm8$gam)$r.sq)
-
-gamm9 <- gamm(for_bites_median ~ species + s(length_mean, k=3) + s(pc1_all, by = factor(species_code), k=3), 
-           family = gaussian(link=identity),
-           random = list(island=~1),
-           data = m_data
-           )
-g9 <- c("g9", "Species + Size + PC1~Species", AICc(gamm9$lme), summary(gamm9$gam)$r.sq)
-
-gamm10 <- gamm(for_bites_median ~ species + s(pc1_all, by = factor(species_code), k=3), 
-           family = gaussian(link=identity),
-           random = list(island=~1),
-           data = m_data
-           )
-g10 <- c("g10", "Species + PC1~Species", AICc(gamm10$lme), summary(gamm10$gam)$r.sq)
-
-gamm11 <- gamm(for_bites_median ~ s(length_mean, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3), 
-           family = gaussian(link=identity),
-           random = list(island=~1),
-           data = m_data
-           )
-g11 <- c("g11", "Size~Species + PC1~Species", AICc(gamm11$lme), summary(gamm11$gam)$r.sq)
-
-gamm12 <- gamm(for_bites_median ~ species_code + s(pc1_all, k=3), 
-           family = gaussian(link=identity),
-           random = list(island=~1),
-           data = m_data
-           )
-g12 <- c("g12", "Species + PC1", AICc(gamm12$lme), summary(gamm12$gam)$r.sq)
-
-gtab_fi_site <- as.data.frame(rbind(g1,g2,g3,g4,g5,g6,g7,g8,g9,g10,g11,g12)) %>% setNames(c("g","Model", "AICc", "Adj. R2")) %>% arrange(AICc)
-
-# plots from best models
-plot(gamm12$gam,pages=1)
-plot(gamm10$gam,pages=1)
-
-summary(gamm12$gam)
-```
-
-# Supplement
-
-### Predator gape size
-
-From Mumby et al.:
-Grouper *E. striatus*, gape width (G): G = -0.5998 + (0.167 × TL)
-```{r}
-g30_mum <- -0.5998 + (0.167 * 30) # 4
-g40_mum <- -0.5998 + (0.167 * 40) # 6
-```
-
-From Goatley & Bellwood:
-Horizontal gape: G = -0.002 * SL + 6.421
-```{r}
-g30_goat <- -0.002 * 30 + 6.421
-g40_goat <- -0.002 * 40 + 6.421
-```
-
-Dorsoventral height (Rakitin & Kramer 1996):
-S. viride: 15cm FL > 4.5cm body depth
-S. vetula: 15cm FL > 4.5cm body depth
-
-### loess plots for additional drivers?

### 2021-01-16 r/old_code/Data_prodding.Rmd
clearing out old files

--- 2019-06-05 r/old drafts/Data_prodding.Rmd
@@
 ---
-
-Notes 11/11:
-* figure out dcast issues
-* replacing NA/0s within pipe
 
@@
 ```
+
 ```{r,message=FALSE,warning=FALSE, echo=FALSE}
@@
 ```{r}
-sample <- group_by(sum_id, site, species, species_code, phase, island) %>% summarize(n=n())
+sample_tot <- sum_id %>%
+  filter(species_code != "rbp" & phase == "i") %>%
+  group_by(site, species, species_code, phase, island) %>% 
+  summarize(n=n())
+sample_28 <- sum_id %>%
+  filter(species_code != "rbp" & phase == "i" & length_cm <= 28) %>%
+  group_by(site, species, species_code, phase, island) %>% 
+  summarize(n=n())
+sample_30 <- sum_id %>%
+  filter(species_code != "rbp" & phase == "i" & length_cm <= 30) %>%
+  group_by(site, species, species_code, phase, island) %>% 
+  summarize(n=n())
 
-ggplot(sample, aes(x=factor(site, levels = sum_site$site), y=n)) +
+ggplot(sample_tot, aes(x=factor(site, levels = sum_site$site), y=n)) +
         facet_grid(phase~species) +

### 2021-01-16 r/old_code/Grazing_impact.Rmd
clearing out old files

--- 2019-04-24 r/Analysis.Rmd
@@
         theme(axis.text.x = element_text(angle = 90, hjust = 1))
-ggsave(here("figs/grazing_imp.png"))
+#ggsave(here("figs/grazing_imp.png"))
 ```
@@
 
-ggsave(here("figs/bc_point_panel.png"),g)
+#ggsave(here("figs/bc_point_panel.png"),g)
 ```

### 2021-01-16 r/old_code/Models.Rmd
clearing out old files

--- 2019-05-08 r/Models.Rmd
@@
 m_data2 <- sum_ssp_pca %>% filter(species_code!="rbp") # currently doesn't work when removing rbp data
-gamm2 <- gamm(fr_mean ~ species_code + phase + s(length_m) + s(pc1_bent) + s(pc2_bent) + s(scar_bm), 
+gamm2 <- gamm(fr_mean ~ species_code + phase + s(length_mean) + s(pc1_bent) + s(pc2_bent) + s(scar_bm), 
            family = gaussian(link=identity),
@@
 # Confirming low VIF among predictor variables with site-level model
-lme1 <- lme(fr_mean ~ phase + length_m + scar_bm + pc1_bent + pc2_bent, data=sum_ssp_pca,
+lme1 <- lme(fr_mean ~ phase + length_mean + scar_bm + pc1_bent + pc2_bent, data=sum_ssp_pca,
     random = ~1 | island, 
@@
 # Main effects only
-gamm1 <- gamm(fr_mean ~ species_code + s(length_m) + phase + s(pc1_all) + s(pc2_all), 
+gamm1 <- gamm(fr_mean ~ species_code + s(length_mean) + phase + s(pc1_all) + s(pc2_all), 
            family = gaussian(link=identity),
@@
 # Interactions
-gamm2 <- gamm(fr_mean ~ species_code + phase + s(length_m, by = factor(species_code)) + s(pc2_all)  + s(pc1_all, by = factor(species_code)), 
+gamm2 <- gamm(fr_mean ~ species_code + phase + s(length_mean, by = factor(species_code)) + s(pc2_all)  + s(pc1_all, by = factor(species_code)), 
            family = gaussian(link=identity),

### 2021-01-16 r/old_code/SummaryOrals.Rmd
clearing out old files

--- 2018-04-26 SummaryOrals.Rmd
@@
 # set up
-library(rmarkdown)
-library(dplyr)
+library(here)
+library(janitor)
 library(tidyr)
-library(lazyeval)
 library(ggplot2)
@@
 library(RColorBrewer)
-library(ggbiplot)
 library(nlme)
@@
 ```{r,message=FALSE,warning=FALSE, echo=FALSE}
-setwd("~/github/Scarid-Behavior/data")
-site.sum <- read.csv("site.sum.csv", stringsAsFactors=FALSE) %>% select(-X)
-data <- read.csv("data2.csv", stringsAsFactors=FALSE) %>% filter(!(Species.Code=="qup" & Site=="Turtle Bay")) %>% filter(!(Species.Code=="qup" & Site=="Pallaster East")) # remove due to low sample size
-
-vetvir <- data %>% filter(Species.Code!="rbp")
-vetvirinit <- vetvir %>% filter(Phase=="i")
-vetvir20 <- vetvir %>% filter(Length.cm>=15 & Length.cm<=25)
-init20<- vetvir %>% filter(Length.cm>=15 & Length.cm<=25 & Phase=="i")
-vetvirAB <- vetvir %>% filter(Island!="Barbuda")
-vetvirAB20 <- vetvir %>% filter(Island!="Barbuda")
-vetvirABinit <- vetvirAB %>% filter(Phase=="i")
-
-vetvirAB20NS <- vetvirAB20 %>% filter(Site!="Snapper Point")
-vetvir20NS <- vetvir20 %>% filter(Site!="Snapper Point")
-vetvirNS <- vetvir %>% filter(Site!="Snapper Point")
+site.sum <- read.csv(here("data","sum_site.csv"), stringsAsFactors=F) %>% 
+  select(-X) %>%
+  janitor::clean_names(case="old_janitor")
+data <- read.csv(here("data","sum_id.csv"), stringsAsFactors=F) %>%
+  filter(!(species_code=="qup" & site=="Turtle Bay")) %>% 
+  filter(!(species_code=="qup" & Site=="Pallaster East")) # remove due to low sample size
 ```
@@
 
-areagrazed.site <- as.data.frame(areagrazed.site)
+area <- as.data.frame(areagrazed.site)
 AG <- melt(areagrazed.site, id=c("Site","Island")) %>% dplyr::rename("data.source"=variable, "area.grazed"=value)

### 2021-03-12 r/Paper_analyses.Rmd
updating figure axis text for paper

@@
   #ylim(0, 50) +
-  labs(y = "Feeding rate \n (bites/min)", 
+  labs(y = expression(atop("Feeding rate", paste ("(bites ",min^-1,")"))), 
        title = "S. vetula") +
@@
   scale_fill_manual(values = colors) +
-  labs(y = "Bite rate \n(bites/sec)", 
+  labs(y = expression(atop("Bite rate", paste ("(bites ",sec^-1,")"))), 
        x = "", 
@@
   scale_fill_manual(values = colors) +
-  labs(y = "Grazing intensity\n(bites/foray)", 
+  labs(y = expression(atop("Grazing intensity", paste ("(bites ",foray^-1,")"))), 
        x = "", 
